### 2021-02-02 manuscript/manuscript.Rmd
first commit

---
title: Forecasting for emergency medcine
author:
  - name: Author1
    email: emaail@example.com
    affiliation: University1
    footnote: 1
  - name: Author2
    email: email2@example.com
    affiliation: University2
    footnote: 2
address:
  - code: University1
    address: adress1
  - code: University2
    address: adress2
footnote:
  - code: 1
    text: "Corresponding Author"

abstract: |
   The Objective of this paper
keywords:  ""
journal: "Which journal?"
date: "`r Sys.Date()`"
geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
#linenumbers: true
numbersections: true
#output: rticles::elsevier_article
header-includes:  
 \usepackage{adjustbox, float,lscape} #use the 'float' package
output: 
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    extra_dependencies: ["flafter"]
    keep_tex: true
bibliography: mybibfile.bib
biblio-style: authoryear
#csl: elsevier-without-titles.csl
#biblio-style: abbrevnat
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(patchwork)
library(tsibble)
library(kableExtra)
library(fable)
library(ggthemes)
library(readxl)
```

```{r, include=FALSE, cache=FALSE}

```

# Introduction

# Research background {#lit}

# Experimental design {#design}

# Result and discussion {#result}

# Conclusion {#conclusion}

# References {#references .unnumbered}

### 2021-03-01 manuscript/manuscript.Rmd
start writing literature review and intro

@@
 
+An accurate demand forecasting is crucial in Emergency medicine to depict various courses of action that can result in massive savings in terms of patient lives.
+Inability to match the staff with the demand results in an overcrowding care system which is a serious problem causing challenging situations on patient flow.
+Also, it is related with increasing length of stay, low patient satisfaction, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times.without incurring last-minute expenses, such as overtime or supplemental staffing [Reference].
+
+An accurate forecasting of the daily demand enables managers to match staff to meet anticipated patients, reconfigure units and redeploy staff and vehicles.
+This will have many advantages for patients, staff and the quality of healthcare services provided.
+
+Daily forecasts are required to inform the short-term planning for the current and the upcoming shifts of the day.
+This involves the decision making related to the execution of the delivery process for various health care services such as Ambulatory ,Emergency.
+Daily forecasts are important at various levels:
+
+- **Nature of incidents**
+
+- **Category: Red, Amber, Green**
+
+- **Health board**
+
+- **Country level**
+
+The combination of demand forecast, incidents being attended, resource availability and delays at hospitals, provide information on the state of the unscheduled care system across the emergency medicine services.
+Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients.
+
 # Research background {#lit}
+
+Table \@ref(tab:summarylit) summarise studies on forecasting in emergency and urgent care.
+
+```{r summarylit}
+li <- readxl::read_xlsx("../data/literature.xlsx")
+knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("scale_down","hold_position")) %>% column_spec(5,"15em")
+```
+
+<!--
+Is there any study using grouped time series in urgent and emergency care?
+-->
+
+<!--
+Using models for count time series in urgent and emergency care!
+-->
+
+@gijo2016sarima generated a time series model to forecast the daily and hourly call volume at all centre handling emergency ambulance services.
+Since historical data showed seasonality, SARIMA models were investigated.
+Regarding the daily model, the authors generated a SARIMA model, which, however, resulted in the forecast error (MMSE) that significantly increased when the lead time exceeded 8 days.
+On the other hand, the SARIMA model proposed to forecast the log-calls an hourly basis.
+This model was found to fit well the model both for shorter and longer lead times.
+
+Forecasting at the daily level of granularity is generally use for roaster planing and deciding when to contact staff on call for instance.
+@luo2017hospital use a combination of seasonal ARIMA (SARIMA) and a single exponential smoothing (SES) model to forecast daily outpatient visits 1 week ahead.They indicate that using combinatorial model can be more effective than each model separately.
+@whitt2019forecasting examine several alternative models to forecast the total daily arrival for 1-7 days ahead, including a linear regression based on calendar and weather variables, seasonal autoregressive integrated moving average with exogenous regressors (SARIMAX) and the multilayer perceptron (MLP) model.
+Using a daily ED admission for 3 years, they show that SARIMAX provides the most accurate daily forecast by MSE measure.
+@marcilio2013forecasting compare the forecast accuracy of the Generalised Linear Model (GLM), Generalised Estimating Equations (GEE), and Seasonal Autoregressive Integrated Moving Average (SARIMA) methods using total daily patient visits to an ED using MAPE.
+They conclud that GLM and GEE models provide more accurate forecasts than SARIMA model.
+@rostami2020anticipating propose a model to forecast daily ED attendance with consideration of different types of holidays, weekday effects, auto-regressive effects, long-term trends and date effects.
+They provide probabilistic forecasts to quantify uncertainties in future ED attendance and they show that the proposed model outperforms three time series techniques including 1) Naive, 2) AutoRegressive, AR(p), 3)  exponential smoothing state space model (ETS) and a regression model without considering special events as alternatives.
+@zhou2018time SARIMA, NARNN and the hybrid SARIMA-NARNN to forecast the daily number of new admission inpatients.
+The root mean square error (RMSE), mean absolute error (MAE) and mean absolute percentage error (MAPE) were used to compare the forecasting performance among the three models.
+They show that NARNN model outperforms the others.
+@sun2009forecasting use SARIMA model to forecast daily patient volume for each patient acuity level.
+MAPE is used to choose the best-fit model.
+They fitted separate ARIMA models to the three categories of acuity and overall data.
+They conclude that the ARIMA model is effective for both short term (weekly) and long term (three months) forecast horizons.
+Moreover, they observe that the impact of weather is not significant.
+@zinouri2018modelling use SARIMA to develop a statistical prediction model that provides short-term forecasts of daily surgical demand.
+Our results suggest that the proposed SARIMA model can be useful for estimating surgical case volumes 2–4 weeks prior to the day of surgery, which can support robust and reliable staff schedules.
+@moustris2012seven use Artificial Neural Network (ANN) models to forecast the total weekly(7 days) number of Childhood Asthma Admission. Three different ANN models were developed
+and trained to forecast the admission of different age subgroups 
+and for the whole study population.
+@mccoy2018assessment compare SARIMA, Prophet and Snaive methods for prediction 365 days of hospital discharge using a daily hospital discharge volumes at 2 large, New England academic medical centers.
+They show that Prophet model outperforms SARIMA and Snaive.
+
+@khaldi2019forecasting investigate the combination of the Artificial Neural Networks (ANNs) with a signal decomposition technique named Ensemble Empirical Mode decomposition (EEMD), to make one step ahead weekly forecasting of patients arrivals to ED.
+using seven years of weekly demand.
+The results of the proposed model were compared against ANN without signal decomposition, ANN with Discrete wavelet Transform (DWT).
+The results show that the combined forecasts outperform the benchmarking models.
+
+@steins2019forecasting aimed to develop a forecasting model for predicting the number of ambulance services calls per hour and geographical areas, to support managers in decisions-making and to investigate which ones were the factors that affected the number of calls.
+Data collected consisted of a time and location of historical ambulance call data for three counties in Sweden and a list of explanatory factors of the area, such as socioeconomic and geographic. 
+In order to deal with large number of zeros in the data, authors developed zero-inflated Poisson (ZIP) and zero-inflated negative binomial (ZINB) regression models.
+These were then compared to the currently existing forecasting system, based on moving average with seasonality weights, using ME, MAE and RMSE.
+Firstly, the factors affecting the number of ambulance calls were found to be the following: population in different age groups, median income, length of road, number of nightlife spots (the number of restaurants), day of the week and hour of the day.
+Secondly, it was found that the older population (65-100) generated more ambulance calls and that ZIP model performs better than the current model.
+However, the improvement provided by the more advance model was not much greater than the one provided by the existing model.
+Authors suggested that it could be because the population, which was used as an independent variable in both models, was so dominant compared to the other factors.
+Moreover, both models either underestimated or overestimated the number of calls.
+Authors suggested that the inability to capture a positive trend resulted in underestimation, and the opposite was due to a negative trend.
+Therefore, authors recommended that further research should add certain temporal variables able to capture the trend.
 

### 2021-03-01 manuscript/manuscript.tex
start writing literature review and intro

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Which journal?} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Forecasting for emergency medcine},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry sp
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set line spacing
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#3\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% Pandoc header
\usepackage{adjustbox, float,lscape}
\usepackage{flafter}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Forecasting for emergency medcine}
    \author[University1]{Author1\corref{1}}
   \ead{emaail@example.com} 
    \author[University2]{Author2\corref{2}}
   \ead{email2@example.com} 
      \address[University1]{adress1}
    \address[University2]{adress2}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  The Objective of this paper
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

An accurate demand forecasting is crucial in Emergency medicine to depict various courses of action that can result in massive savings in terms of patient lives. Inability to match the staff with the demand results in an overcrowding care system which is a serious problem causing challenging situations on patient flow. Also, it is related with increasing length of stay, low patient satisfaction, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times.without incurring last-minute expenses, such as overtime or supplemental staffing {[}Reference{]}.

An accurate forecasting of the daily demand enables managers to match staff to meet anticipated patients, reconfigure units and redeploy staff and vehicles. This will have many advantages for patients, staff and the quality of healthcare services provided.

Daily forecasts are required to inform the short-term planning for the current and the upcoming shifts of the day. This involves the decision making related to the execution of the delivery process for various health care services such as Ambulatory ,Emergency. Daily forecasts are important at various levels:

\begin{itemize}
\item
  \textbf{Nature of incidents}
\item
  \textbf{Category: Red, Amber, Green}
\item
  \textbf{Health board}
\item
  \textbf{Country level}
\end{itemize}

The combination of demand forecast, incidents being attended, resource availability and delays at hospitals, provide information on the state of the unscheduled care system across the emergency medicine services. Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients.

\hypertarget{lit}{%
\section{Research background}\label{lit}}

Table \ref{tab:summarylit} summarise studies on forecasting in emergency and urgent care.

\begin{table}[!h]

\caption{\label{tab:summarylit}Summary of studies in hourly emergency care forecasting}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrll>{\raggedright\arraybackslash}p{15em}l}
\toprule
Author & Year & Forecast variable & Forecast Horizon & Method & Measure\\
\midrule
McCarthy 
2008 & 2008 & ED arrivals & 24h & Poisson log-linear regression model, including temporal factors, patient characteristics and climatic factors & 95\% CI\\
Asheim et al. & 2019 & ED arrivals & 3h & Poisson regression with weekly and yearly cyclic 
effects. & MAPE\\
Cote et al. & 2013 & ED arrivals & 24h & Fourier regression & R\textasciicircum{}2,  Standard Error\\
Kim et al. & 2014 & Hospital demand & 4h, 24h, 7days, 
30days & Linear regression; Exponential smoothing; ARIMA; GARCH; VAR & MAPE\\
Schweigler et al. & 2009 & ED bed occupancy & 4h and 12h & Hourly historical average; SARIMA; Sinusoidal model with autocorrelated error & RMSE\\
Channouf et al. & 2007 & Ambulance demand & 12h,14h,17h,23h,24h,1h,3h,6h, 13h & Regression & RMSE\\
Hertzum & 2017 & ED arrivals
ED occupancy & 1,2,4,8,24 hours & linear regression; SARIMA; Naïve & MAE, MAPE, 
MASE\\
Choudhury and Urena & 2020 & ED arrivals & 1h to 24h & ARIMA; Holt-winters; TBATS; ANN & RMSE, ME\\
Steins et al. & 2019 & Ambulance call & 24h & ZIP and ZINB regression;
 moving average with seasonality weights & ME, MAE, RMSE\\
Jones et al. & 2009 & ED census & 24h & VAR;  Holt winters & MAE\\
Morzuch and Allen & 2006 & ED arrivals & 168h & Regression; ARIMA; Exponential smoothing & RMSE\\
Chase et al. & 2012 & ED CUR & 30m
1h, 2h, 4h, 8h, 12h & Binary regression & NA\\
Taylor & 2008 & centre volume & 30 minutes 
to 2 weeks & ARMA, Exponential smoothing; Dynamic harmonic regression; Seasonal random walk; Seasonal mean;
Random walk; Mean; Median; Simple exponential smoothing & MAE
RMSE\\
Gijo and Balakrishna & 2016 & call volume & 168h & SARIMA & MMSE\\
\bottomrule
\end{tabular}}
\end{table}

Gijo and Balakrishna (2016) generated a time series model to forecast the daily and hourly call volume at all centre handling emergency ambulance services. Since historical data showed seasonality, SARIMA models were investigated. Regarding the daily model, the authors generated a SARIMA model, which, however, resulted in the forecast error (MMSE) that significantly increased when the lead time exceeded 8 days. On the other hand, the SARIMA model proposed to forecast the log-calls an hourly basis. This model was found to fit well the model both for shorter and longer lead times.

Forecasting at the daily level of granularity is generally use for roaster planing and deciding when to contact staff on call for instance. Luo et al. (2017) use a combination of seasonal ARIMA (SARIMA) and a single exponential smoothing (SES) model to forecast daily outpatient visits 1 week ahead.They indicate that using combinatorial model can be more effective than each model separately. Whitt and Zhang (2019) examine several alternative models to forecast the total daily arrival for 1-7 days ahead, including a linear regression based on calendar and weather variables, seasonal autoregressive integrated moving average with exogenous regressors (SARIMAX) and the multilayer perceptron (MLP) model. Using a daily ED admission for 3 years, they show that SARIMAX provides the most accurate daily forecast by MSE measure. Marcilio, Hajat, and Gouveia (2013) compare the forecast accuracy of the Generalised Linear Model (GLM), Generalised Estimating Equations (GEE), and Seasonal Autoregressive Integrated Moving Average (SARIMA) methods using total daily patient visits to an ED using MAPE. They conclud that GLM and GEE models provide more accurate forecasts than SARIMA model. Rostami-Tabar and Ziel (2020) propose a model to forecast daily ED attendance with consideration of different types of holidays, weekday effects, auto-regressive effects, long-term trends and date effects. They provide probabilistic forecasts to quantify uncertainties in future ED attendance and they show that the proposed model outperforms three time series techniques including 1) Naive, 2) AutoRegressive, AR(p), 3) exponential smoothing state space model (ETS) and a regression model without considering special events as alternatives. Zhou et al. (2018) SARIMA, NARNN and the hybrid SARIMA-NARNN to forecast the daily number of new admission inpatients. The root mean square error (RMSE), mean absolute error (MAE) and mean absolute percentage error (MAPE) were used to compare the forecasting performance among the three models. They show that NARNN model outperforms the others. Sun et al. (2009) use SARIMA model to forecast daily patient volume for each patient acuity level. MAPE is used to choose the best-fit model. They fitted separate ARIMA models to the three categories of acuity and overall data. They conclude that the ARIMA model is effective for both short term (weekly) and long term (three months) forecast horizons. Moreover, they observe that the impact of weather is not significant. Zinouri, Taaffe, and Neyens (2018) use SARIMA to develop a statistical prediction model that provides short-term forecasts of daily surgical demand. Our results suggest that the proposed SARIMA model can be useful for estimating surgical case volumes 2--4 weeks prior to the day of surgery, which can support robust and reliable staff schedules.
Moustris et al. (2012) use Artificial Neural Network (ANN) models to forecast the total weekly(7 days) number of Childhood Asthma Admission. Three different ANN models were developed
and trained to forecast the admission of different age subgroups
and for the whole study population.
McCoy, Pellegrini, and Perlis (2018) compare SARIMA, Prophet and Snaive methods for prediction 365 days of hospital discharge using a daily hospital discharge volumes at 2 large, New England academic medical centers. They show that Prophet model outperforms SARIMA and Snaive.

Khaldi, El Afia, and Chiheb (2019) investigate the combination of the Artificial Neural Networks (ANNs) with a signal decomposition technique named Ensemble Empirical Mode decomposition (EEMD), to make one step ahead weekly forecasting of patients arrivals to ED. using seven years of weekly demand. The results of the proposed model were compared against ANN without signal decomposition, ANN with Discrete wavelet Transform (DWT). The results show that the combined forecasts outperform the benchmarking models.

Steins, Matinrad, and Granberg (2019) aimed to develop a forecasting model for predicting the number of ambulance services calls per hour and geographical areas, to support managers in decisions-making and to investigate which ones were the factors that affected the number of calls.
Data collected consisted of a time and location of historical ambulance call data for three counties in Sweden and a list of explanatory factors of the area, such as socioeconomic and geographic.
In order to deal with large number of zeros in the data, authors developed zero-inflated Poisson (ZIP) and zero-inflated negative binomial (ZINB) regression models. These were then compared to the currently existing forecasting system, based on moving average with seasonality weights, using ME, MAE and RMSE. Firstly, the factors affecting the number of ambulance calls were found to be the following: population in different age groups, median income, length of road, number of nightlife spots (the number of restaurants), day of the week and hour of the day. Secondly, it was found that the older population (65-100) generated more ambulance calls and that ZIP model performs better than the current model. However, the improvement provided by the more advance model was not much greater than the one provided by the existing model. Authors suggested that it could be because the population, which was used as an independent variable in both models, was so dominant compared to the other factors.
Moreover, both models either underestimated or overestimated the number of calls. Authors suggested that the inability to capture a positive trend resulted in underestimation, and the opposite was due to a negative trend. Therefore, authors recommended that further research should add certain temporal variables able to capture the trend.

\hypertarget{design}{%
\section{Experimental design}\label{design}}

\hypertarget{result}{%
\section{Result and discussion}\label{result}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-gijo2016sarima}{}%
Gijo, EV, and N Balakrishna. 2016. {``SARIMA Models for Forecasting Call Volume in Emergency Services.''} \emph{International Journal of Business Excellence} 10 (4): 545--61.

\leavevmode\hypertarget{ref-khaldi2019forecasting}{}%
Khaldi, Rohaifa, Abdellatif El Afia, and Raddouane Chiheb. 2019. {``Forecasting of Weekly Patient Visits to Emergency Department: Real Case Study.''} \emph{Procedia Computer Science} 148: 532--41.

\leavevmode\hypertarget{ref-luo2017hospital}{}%
Luo, Li, Le Luo, Xinli Zhang, and Xiaoli He. 2017. {``Hospital Daily Outpatient Visits Forecasting Using a Combinatorial Model Based on ARIMA and SES Models.''} \emph{BMC Health Services Research} 17 (1): 469.

\leavevmode\hypertarget{ref-marcilio2013forecasting}{}%
Marcilio, Izabel, Shakoor Hajat, and Nelson Gouveia. 2013. {``Forecasting Daily Emergency Department Visits Using Calendar Variables and Ambient Temperature Readings.''} \emph{Academic Emergency Medicine} 20 (8): 769--77.

\leavevmode\hypertarget{ref-mccoy2018assessment}{}%
McCoy, Thomas H, Amelia M Pellegrini, and Roy H Perlis. 2018. {``Assessment of Time-Series Machine Learning Methods for Forecasting Hospital Discharge Volume.''} \emph{JAMA Network Open} 1 (7): e184087--87.

\leavevmode\hypertarget{ref-moustris2012seven}{}%
Moustris, Kostas P, Konstantinos Douros, Panagiotis T Nastos, Ioanna K Larissi, Michael B Anthracopoulos, Athanasios G Paliatsos, and Kostas N Priftis. 2012. {``Seven-Days-Ahead Forecasting of Childhood Asthma Admissions Using Artificial Neural Networks in Athens, Greece.''} \emph{International Journal of Environmental Health Research} 22 (2): 93--104.

\leavevmode\hypertarget{ref-rostami2020anticipating}{}%
Rostami-Tabar, Bahman, and Florian Ziel. 2020. {``Anticipating Special Events in Emergency Department Forecasting.''} \emph{International Journal of Forecasting}.

\leavevmode\hypertarget{ref-steins2019forecasting}{}%
Steins, Krisjanis, Niki Matinrad, and Tobias Granberg. 2019. {``Forecasting the Demand for Emergency Medical Services.''} In \emph{Proceedings of the 52nd Hawaii International Conference on System Sciences}.

\leavevmode\hypertarget{ref-sun2009forecasting}{}%
Sun, Yan, Bee Hoon Heng, Yian Tay Seow, and Eillyne Seow. 2009. {``Forecasting Daily Attendances at an Emergency Department to Aid Resource Planning.''} \emph{BMC Emergency Medicine} 9 (1): 1.

\leavevmode\hypertarget{ref-whitt2019forecasting}{}%
Whitt, Ward, and Xiaopei Zhang. 2019. {``Forecasting Arrivals and Occupancy Levels in an Emergency Department.''} \emph{Operations Research for Health Care} 21: 1--18.

\leavevmode\hypertarget{ref-zhou2018time}{}%
Zhou, Lingling, Ping Zhao, Dongdong Wu, Cheng Cheng, and Hao Huang. 2018. {``Time Series Model for Forecasting the Number of New Admission Inpatients.''} \emph{BMC Medical Informatics and Decision Making} 18 (1): 39.

\leavevmode\hypertarget{ref-zinouri2018modelling}{}%
Zinouri, Nazanin, Kevin M Taaffe, and David M Neyens. 2018. {``Modelling and Forecasting Daily Surgical Case Volume Using Time Series Analysis.''} \emph{Health Systems} 7 (2): 111--19.

\end{CSLReferences}

\end{document}

### 2021-03-01 fit_model.R

@@ -28,21 +28,23 @@ incidents <- incidents_original %>%   mutate_at(.vars = "nature_of_incident", .f
   as_tsibble(index = date, key = c(lhb_code, category, nature_of_incident)) %>%
   fill_gaps(incidents = 0)
 
+
+
 # prepare data
 breatjing_problem_ts <- incidents %>% index_by(date) %>% 
   group_by(nature_of_incident) %>% 
   summarise(incidents = sum(incidents)) %>% 
-  filter(nature_of_incident == "BREATHING PROBLEMS") %>% 
-  select(-nature_of_incident)
+  filter(nature_of_incident == "BREATHING PROBLEMS") %>%  dplyr::select(-nature_of_incident)
+
+ggplot(breatjing_problem_ts, aes(x = incidents)) + 
+  geom_histogram(binwidth = 1, boundary = 0, color = "white")
+
 breatjing_problem_variables <- breatjing_problem_ts %>% mutate(t =row_number(), 
                                 f1_sin = sin((2*pi*t)/7), f1_cos = cos((2*pi*t)/7),
                                 f2_sin = sin((2*2*pi*t)/7), f2_cos = cos((2*2*pi*t)/7),
                                 f3_sin = sin((3*2*pi*t)/7), f3_cos = cos((3*2*pi*t)/7),
-                                f4_sin = sin((4*2*pi*t)/7), f4_cos = cos((4*2*pi*t)/7),
-                                f5_sin = sin((5*2*pi*t)/7), f5_cos = cos((5*2*pi*t)/7),
-                                f6_sin = sin((6*2*pi*t)/7), f6_cos = cos((6*2*pi*t)/7),
-                                f7_sin = sin((7*pi*t)/7), f7_cos = cos((7*2*pi*t)/7))
-
+                                holiday= if_else(incidents > 150, 1,0))
+View(breatjing_problem_variables)
 breatjing_problem_train <-breatjing_problem_variables %>% filter_index(~ "2019-07-24")
 breatjing_problem_test <-breatjing_problem_variables %>% filter_index("2019-07-25" ~ .)
 breatjing_problem_new_data <- breatjing_problem_test %>% as_tibble() %>% select(-date, -incidents)
@@ -55,9 +57,9 @@ breatjing_problem_train_ts <- ts(breatjing_problem_train$incidents, frequency = 
 breatjing_problem_fit <- tsglm(breatjing_problem_train_ts,
[…]

### 2021-03-11 fit_model.R
first forecast model with TSCOUNT

@@ -14,7 +14,8 @@ library(tsibble)
 library(fable)
 library(lubridate)
 library(MASS)
-
+library(fable.tscount)
+library(patchwork)
 # Read data
 incidents_original <- readxl::read_excel("data/Nature_of_Incidents_Attended.xlsx") %>%
   mutate(date = as_date(Incident_Date)) %>%
@@ -24,107 +25,129 @@ incidents <- incidents_original %>%   mutate_at(.vars = "nature_of_incident", .f
   mutate(nature_of_incident= ifelse(!is.na(nature_of_incident),nature_of_incident_description , "other")) %>% 
   group_by(lhb_code, category, nature_of_incident, date) %>% 
   summarise(incidents = sum(total_incidents)) %>% ungroup() %>% 
-  mutate(category = factor(category, level=c("RED","AMBER","GREEN"))) %>% 
   as_tsibble(index = date, key = c(lhb_code, category, nature_of_incident)) %>%
   fill_gaps(incidents = 0)
 
 
 
-# prepare data
-breatjing_problem_ts <- incidents %>% index_by(date) %>% 
+# prepare data, look at two different series 
+breathing_problem_ts <- incidents %>% index_by(date) %>% 
   group_by(nature_of_incident) %>% 
   summarise(incidents = sum(incidents)) %>% 
-  filter(nature_of_incident == "BREATHING PROBLEMS") %>%  dplyr::select(-nature_of_incident)
+  filter(nature_of_incident == "BREATHING PROBLEMS") %>%  
+  dplyr::select(-nature_of_incident) %>% ungroup()
 
-ggplot(breatjing_problem_ts, aes(x = incidents)) + 
-  geom_histogram(binwidth = 1, boundary = 0, color = "white")
+# this ts is ntermittent
+choking_ts <- incidents %>% index_by(date) %>% 
+  group_by(nature_of_incident) %>% 
+  summarise(incidents = sum(incidents)) %>% 
+  filter(nature_of_incident == "CHOKING") %>%  
+  dplyr::select(-nature_of_incident) %>% ungroup()
 
-breatjing_problem_variables <- breatjing_problem_ts %>% mutate(t =row_number(), 
-                                f1_sin = sin((2*pi*t)/7), f1_cos = cos((2*pi*t)/7),
-                                f2_sin = sin((2*2*pi*t)/7), f2_cos = cos((2*2*pi*t)/7),
[…]

### 2022-10-13 manuscript/manuscript.Rmd
first forecast model with TSCOUNT
reorganise files and prepare hts plus with box_cox transformation for ets and arima

@@
 
+Focus on the ambulance services!
+
 @gijo2016sarima generated a time series model to forecast the daily and hourly call volume at all centre handling emergency ambulance services.
@@
 
-# Experimental design {#design}
+# Data
+
+Operations Department and I suppose it's split into three different categories.
+You've got...
+you've got Response which is the ambulances.
+You've got the Control which is the people who take the telephone calls.
+And then you've got Resilience which are the things that people like have the hazardous response teams so they'll go and save people from water and maybe with fires as well.
+They've got...
+they’re specially trained in different things so there's the contingency sort of there.
+So from a Control perspective, they look at...
+they look at it at least twice a week to see what the...
+what it...
+how busy it is.
+And then they staff...
+they… they’re looking to staff to the worst level, knowing that they will get some people...
+there will be some abstractions.
+So they won't manage to...
+to manage that, because their key area [inaudible].
+Excuse me, they try to answer within the...
+the phone within three seconds 95% of the time.
+I'm not sure exactly.
+It's something like that.
+That’s very high up there.
+That's what they're trying to do.
+So...
+and because I got it broken down at the Control level, so we've got North Wales level for calls, and we've also got it broken down by the North, the Mid, and the South Control.
+It's actually called the North, the South East, and Central and West.
+Those are the three control areas.
+So they can see how those break down, knowing that at the moment we've got particularly high demand in the South East, and it's forecast to be just bonkers
+
+# Experimental set {#design}
 

### 2022-10-13 figure/hierarchical_graph.R

@@ -1,0 +1,36 @@
+library(collapsibleTree) 
+
+# input data must be a nested data frame:
+head(warpbreaks)
+
+# Represent this tree:
+p <- collapsibleTree( warpbreaks, 
+                      c("wool", "tension", "breaks"))
+
+Total <- incident_ready %>% 
+  select(region, lhb,category,nature)
+
+p <- collapsibleTree( Total, 
+                      c("region","lhb","category","nature"))
+p
+
+
+
+widgetToPng <- function(widget, file = "widget.pdf",  ...) {
+  temp <- tempfile(fileext = ".html")
+  file <- R.utils::getAbsolutePath(file)
+  htmlwidgets::saveWidget(widget, temp)
+  webshot(
+    temp, file,
+    selector = "#htmlwidget_container",
+    zoom = 2,
+    delay = 0.5,
+    ...
+  )
+}
+
+widgetToPng(
+  collapsibleTreeSummary(warpbreaks, c("wool", "tension", "breaks"), 
+                         maxPercent = 50, collapsed = FALSE),
+  "collapsibleTree.png"
+)

### 2022-10-13 fit_model_fable.R

@@ -1,132 +1,0 @@
-library(tidyverse)
-library(tsibble)
-library(fable)
-library(fabletools)
-library(feasts)
-library(lubridate)
-library(fable.tscount)
-library(tscount)
-
-f_horizon <- 5*7
-# Read data
-incidents_original <- readxl::read_excel("data/Nature_of_Incidents_Attended.xlsx") %>%
-  mutate(date = as_date(Incident_Date)) %>%
-  janitor::clean_names() %>%
-  force_tz(date, tz = "GB") 
-# nature of incidents with low volume
-nature_of_incident_low <- c("AUTOMATIC CRASH NOTIFICATION",
-                            "INACCESSIBLE INCIDENT/OTHER ENTRAP",
-                            "INTERFACILITY EVALUATION/TRANSFER",
-                            "MAJOR INCIDENT - OVERRIDE PROQA",
-                            "TRANSFER/INTERFACILITY/PALLIATIVE")
-
-# read holiday data
-holiday <- readxl::read_excel("data/holiday_rugby.xlsx") %>% 
-  select(date,public_holiday,school_holiday) %>% 
-  mutate(date=as_date(date)) %>% 
-  mutate(public_holiday_d=if_else(is.na(public_holiday), 0,1),
-         school_holiday_d=if_else(is.na(school_holiday), 0,1)) %>% 
-  mutate(xmas=if_else(public_holiday=="Christmas Day","1","0"),
-         new_years_day=if_else(public_holiday=="New Years Day",1,0)) %>% 
-  mutate_at(vars(xmas,new_years_day), ~replace(.,is.na(.),0)) 
-
-# create dummies for holidays
-holiday_dummy <- holiday %>% select(date,
-                                    public_holiday_d,
-                                    school_holiday_d,
-                                    xmas,
-                                    new_years_day) %>% 
-  mutate_at(vars(public_holiday_d,school_holiday_d,xmas,new_years_day),
-            ~as_factor(.)) %>% 
-  as_tsibble(index = date)
-holiday_school <- holiday_dummy %>% select(date,school_holiday_d) %>%
-  filter(school_holiday_d==1) %>% 
-  mutate(holiday = "school_holiday") %>% select(-school_holiday_d)
-holiday_public <- holiday_dummy %>% select(date,public_holiday_d) %>%
[…]

### 2022-10-13 forecast_hts.R

@@ -1,204 +1,0 @@
-library(tidyverse)
-library(hts)
-library(forecast)
-library(tscount)
-
-fh <- 5*7
-incident_all <- read_rds("data/incident_all.rds")
-holidays_p <- read_rds("data/holiday_prophet.rds")
-holiday_dummy <- read_rds("data/holiday_dummy.rds")
-holiday_dummy <- holiday_dummy %>% 
-  mutate_if(is.factor,as.character)%>% 
-  mutate_if(is.character, as.integer)
-
-holiday_dummy_matrix <- holiday_dummy %>% as_tibble() %>% select(-date) %>% as.matrix()
-holiday_dummy_matrix_train <- holiday_dummy_matrix[1:(1400-fh),]
-holiday_dummy_matrix_test <- holiday_dummy_matrix[((1400-fh+1):1400),]
-
-incident_all_train <- incident_all[1:(1400-fh),]
-incident_all_test <- incident_all[((1400-fh+1):1400),]
-#----
-
-#incident_msts_train <- msts(incident_all_train,seasonal.periods = c(7,365.25))
-#create hierarchical/grouped time series
-incident_ts_train <- ts(incident_all_train, frequency = 7)
-gts_incident_train <- hts::gts(incident_ts_train, 
-                         characters = list(c(1,2),3,9))
-#tets set
-#incident_msts_test <- msts(incident_all_test, seasonal.periods = c(7,365.25))
-incident_ts_test <- ts(incident_all_test, frequency = 7)
-gts_incident_test <- hts::gts(incident_ts_test , 
-                              characters = list(c(1,2),3,9))
-#this is to create grouped time series
-#gts_incident <- hts::gts(incident_msts , characters = c(1,2,3,9))
-
-#Forecast with ets and arima
-comb_ets <- forecast(gts_incident_train, 
-                     h=fh,
-                     method = "comb",
-                     fmethod="ets", 
-                     parallel = TRUE)
-bu_ets <- forecast(gts_incident_train, 
-                   h=fh,method="bu", 
-                   fmethod="ets", 
-                   parallel = TRUE)
-comd_arima <- forecast(gts_incident_train, 
-                       h=fh,
-                       method = "comb",
-                       fmethod="arima", 
-                       parallel = TRUE)
-bu_arima <- forecast(gts_incident_train, 
[…]

### 2022-10-13 hierarchical_graph.R

@@ -1,36 +1,0 @@
-library(collapsibleTree) 
-
-# input data must be a nested data frame:
-head(warpbreaks)
-
-# Represent this tree:
-p <- collapsibleTree( warpbreaks, 
-                      c("wool", "tension", "breaks"))
-
-Total <- incident_ready %>% 
-  select(region, lhb,category,nature)
-
-p <- collapsibleTree( Total, 
-                      c("region","lhb","category","nature"))
-p
-
-
-
-widgetToPng <- function(widget, file = "widget.pdf",  ...) {
-  temp <- tempfile(fileext = ".html")
-  file <- R.utils::getAbsolutePath(file)
-  htmlwidgets::saveWidget(widget, temp)
-  webshot(
-    temp, file,
-    selector = "#htmlwidget_container",
-    zoom = 2,
-    delay = 0.5,
-    ...
-  )
-}
-
-widgetToPng(
-  collapsibleTreeSummary(warpbreaks, c("wool", "tension", "breaks"), 
-                         maxPercent = 50, collapsed = FALSE),
-  "collapsibleTree.png"
-)

### 2022-10-13 minimal_examples/reconcile_negative.R

@@ -1,0 +1,51 @@
+library(hts)
+set.seed(1)
+# Bottom level series
+bts <- msts(matrix(rnorm(1500), ncol=2), seasonal.periods=c(7,365.25))
+bts[bts<0] <- 0
+bts <- round(bts)
+
+# Create hts object
+y <- hts(bts)
+plot(y)
+# Create aggregated series
+ally <- aggts(y)
+# Produce forecasts of all series
+h <- 28
+allf <- matrix(NA, nrow = h, ncol = NCOL(ally))
+for(i in seq(NCOL(ally)))
+  allf[,i] <- stlf(ally[,i], h = h)$mean
+allf[allf<0] <- 0
+# Reconcile results
+y.f <- combinef(allf, get_nodes(y))
+y.f_all <- aggts(y.f)
+
+
+#
+library(hts)
+set.seed(1)
+# Bottom level series
+bts <- msts(matrix(rnorm(1500), ncol=2), seasonal.periods=c(7,365.25))
+bts[bts<0] <- 0
+bts <- round(bts)
+
+# Create hts object
+y <- hts(bts)
+# Produce forecasts
+h <- 28
+comb_ets <- forecast(y,
+                     h=h,
+                     method = "comb",
+                     fmethod="ets",
+                     keep_interval = TRUE,
+                     lambda = .5,
+                     parallel = TRUE)
+
+bu_ets <- forecast(y,
+                   h=h,
+                   method="bu",
+                   fmethod="ets",
+                   keep_interval = TRUE,
+                   lambda = .5,
+                   parallel = TRUE
+)

### 2022-10-13 result/accuracy_hts.R

@@ -1,0 +1,173 @@
+# RMSSE <- function (.error, .train, na.rm = TRUE) 
+# {
+#   .train <- diff(.train, lag = 7)
+#   scale <- mean(.train^2, na.rm = na.rm)
+#   sqrt(mean(.error^2/scale, na.rm = na.rm))
+# }
+
+library(tidyverse)
+library(hts)
+#hts_result_ets <- read_rds("result/hts_result_ets.rds")
+#hts_result_ets <- read_rds("result/hts_result_arima.rds")
+hts_result_ets <- read_rds("result/hts_result_tscount.rds")
+
+
+#-----forecast------
+fct <- hts_result_ets$fct_all
+id <- tibble(.id=rep(1,70))
+a1 <- (fct[[1]]$forecast_all_level) %>% bind_cols(id)
+a1[,] <- NA
+for (i in (1:nrow(hts_result_ets))) {
+  a <- (fct[[i]]$forecast_all_level)
+  id <- tibble(.id=rep(i,70))
+  a <- a %>% bind_cols(id)
+  a1 <- a1 %>% bind_rows(a)
+}
+forecast_hts <- a1 %>% drop_na()
+head(forecast_hts[,c(1,1225:1226)])
+write_rds(forecast_hts,"result/forecast/forecast_tscount.rds")
+
+#----error---------
+
+fct <- hts_result_ets$fct_all
+id <- tibble(.id=rep(1,70))
+a1 <- (fct[[1]]$error_all_level) %>% bind_cols(id)
+a1[,] <- NA
+for (i in (1:nrow(hts_result_ets))) {
+  a <- (fct[[i]]$error_all_level)
+  id <- tibble(.id=rep(i,70))
+  a <- a %>% bind_cols(id)
+  a1 <- a1 %>% bind_rows(a)
+}
+error <- a1 %>% drop_na()
+
+write_rds(error,"result/error/error_tscount.rds")
+
+#---test set-------
+# id <- tibble(.id=rep(1,70))
+# a1 <- (fct[[1]]$testset_all_level) %>% bind_cols(id)
+# a1[,] <- NA
+# for (i in (1:nrow(hts_result_ets))) {
+#   a <- (fct[[i]]$testset_all_level)
+#   id <- tibble(.id=rep(i,70))
+#   a <- a %>% bind_cols(id)
+#   a1 <- a1 %>% bind_rows(a)
+# }
+# testset <- a1 %>% drop_na()
+# write_rds(testset,"result/test/test.rds")
+#----RMSSE----------
[…]

### 2022-10-13 result/error_metric.R

@@ -1,0 +1,12 @@
+
+
+d <- .train %>% features(incident, unitroot_ndiffs)
+D <- .train %>% features(incident, unitroot_nsdiffs)
+
+RMSSE <- function (.error, .train, na.rm = TRUE) 
+{
+    .train <- diff(.train, lag = .period, differences = D)
+  scale <- mean(.train^2, na.rm = na.rm)
+  sqrt(mean(.error^2/scale, na.rm = na.rm))
+}
+

### 2022-10-13 rscript/empirical_distribution.R

@@ -1,0 +1,23 @@
+
+# empirical distribution
+
+library(fpp3)
+library(distributional)
+set.seed(123)
+
+train <- tsibble(t = 1:100, y=rnorm(100), index=t)
+test <- tsibble(t = 101:120, y=rnorm(20), index=t)
+fc_ar1 <- train %>%
+  model(
+    ar1 = ARIMA(y ~ pdq(1,0,0))
+  ) %>%
+  fabletools::forecast(h=20)
+fc_benchmark <- fc_ar1 %>% as_tibble() %>% 
+  mutate(
+    y = distributional::dist_sample(list(train$y)),
+    .mean1=mean(train$y),
+    .model = 'benchmark'
+  )
+bind_rows(fc_ar1, fc_benchmark) %>%
+  accuracy(test, measures = distribution_accuracy_measures)
+

### 2022-10-13 rscript/hts/form JASA paper.R

@@ -1,0 +1,38 @@
+It is easy enough to do using the equations from the JASA paper. Here is an example using the infantgts data from the hts package.
+
+library(hts)
+
+h <- 10
+ally <- aggts(infantgts)
+fmean <- fvar <- matrix(NA, nrow = h, ncol = NCOL(ally))
+for(i in seq(NCOL(ally))) {
+  fc <- forecast(auto.arima(ally[,i]), h = h, level=95)
+  fmean[,i] <- fc$mean
+  fvar[,i] <- ((fc$upper - fc$lower)/(2*qnorm(0.975)))^2
+}
+S <- smatrix(infantgts)
+Lambda <- diag(1/fvar[1,])
+G <- solve(t(S) %*% Lambda %*% S) %*% t(S) %*% Lambda
+SG <- S %*% G
+ytilde <- SG %*% t(fmean)
+vartilde <- ytilde * NA
+for(i in seq(h)) {
+  vartilde <- diag(SG %*% diag(fvar[i,]) %*% t(SG))
+}
+lower <- qnorm(0.025, mean = ytilde, sd = sqrt(vartilde))
+upper <- qnorm(0.975, mean = ytilde, sd = sqrt(vartilde))
+
+I have computed the mean directly (as ytilde). This should give the same results as 
+combinef(fmean, groups = get_groups(infantgts), weights=1/fvar[1,])
+
+But since we needed all the matrices for the variance calculation, I thought it would be more obvious what is happening if I computed ytilde as well.
+
+Because we are using a transformation of the data (sqrt) before fitting the models, you need to compute ytilde, lower and upper on the transformed data. Then square them all to give the results on the original scale.
+
+For the Poisson case, you could simply compute
+
+lower <- qpois(0.025, lambda = ytilde)
+upper <- qpois(0.975, lambda = ytilde)
+
+In this case, you don't need a sqrt transformation of the data, and the resulting ytilde, lower and upper are already on the same scale as the data.
+If there are negative values of ytilde, they will need to be set to zero.

### 2022-10-13 rscript/hts/interval and probability without transformation/hts_ets_interval.R

@@ -1,0 +1,168 @@
+library(tidyverse)
+library(purrr)
+library(hts)
+library(forecast)
+library(tscount)
+library(tsibble)
+fcst_inciddent <- function(train, test) {
+incident_all_train <- train %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
+incident_all_test <- test %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
+holiday_dummy_matrix_train <- train %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
+holiday_dummy_matrix_test <- test %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
+
+incident_ts_train <- ts((incident_all_train), frequency = 7)
+#create hierarchical/grouped time series
+gts_incident_train <- hts::gts(incident_ts_train , 
+                               characters = list(c(1,2),3,9))
+incident_ts_test <- ts(incident_all_test, frequency = 7)
+gts_incident_test <- hts::gts(incident_ts_test , 
+                              characters = list(c(1,2),3,9))
+comb_ets <- forecast(gts_incident_train, 
+                     h=fh,
+                     method = "comb",
+                     fmethod="ets", 
+                     lambda = .5,
+                     parallel = TRUE,
+                     )
+
+bu_ets <- forecast(gts_incident_train, 
+                   h=fh,method="bu", 
+                   fmethod="ets", 
+                   lambda = .5,
+                   parallel = TRUE
+)
+
+
+ally <- aggts(gts_incident_train)
+fmean <- fvar <- matrix(NA, nrow = fh, ncol = NCOL(ally))
+for(i in seq(NCOL(ally))) {
+  fc <- forecast(ets(ally[,i]), h = fh, level=95)
+  fmean[,i] <- fc$mean#base forecast
+  fvar[,i] <- ((fc$upper - fc$lower)/(2*qnorm(0.975)))^2
+}
+S <- smatrix(gts_incident_train)
+Lambda <- diag(1/fvar[1,])
+G <- solve(t(S) %*% Lambda %*% S) %*% t(S) %*% Lambda
+SG <- S %*% G
[…]

### 2022-10-13 rscript/hts/interval_probability with transformation/htsplus_with_boxcox_transformation.R

@@ -1,0 +1,87 @@
+library(hts)
+
+# object = gts object
+# h = forecast horizon
+# model_function = function used to model each time series. e.g., ets or auto.arima
+# lambda = Box-Cox transformation parameter
+# level = prediction interval coverage as a percentage
+# method = reconcilation method. either "wls" or "bu"
+# nsim = number of simulations in computing prediction intervals using nonparametric method
+# htsplus = hts with BoxCox transformation
+htsplus <- function(object, h,
+                    model_function = auto.arima,
+                    lambda=NULL, level=95,
+                    method = c("wls","bu"), nsim=2000) {
+  method <- match.arg(method)
+  # Generate all time series at all levels
+  ally <- aggts(object)
+  nseries <- NCOL(ally)
+  # Do Box-Cox transformation if required
+  if(!is.null(lambda)) {
+    ally <- BoxCox(ally, lambda)
+  }
+  # Set up matrices to store results
+  fmean <- fvar <- flower <- fupper <- matrix(NA, nrow = h, ncol = nseries)
+  models <- list()
+  # Loop over all time series and compute forecast means and variances
+  for(i in seq(NCOL(ally))) {
+    models[[i]] <- model_function(ally[,i])
+    fc <- forecast(models[[i]], h=h, level=level)
+    fmean[,i] <- fc$mean
+    fvar[,i] <- ((fc$upper - fc$lower)/(2*qnorm(0.5+level/200)))^2
+  }
+  # Reverse Box-Cox transformation if required
+  # using Taylor series approximation (delta method) to get mean and variance
+  # on original scale
+  if(!is.null(lambda)) {
+    fmean <- InvBoxCox(fmean, lambda)
+    tmp <- lambda * fmean + 1
+    fp2 <- fmean^(2-2*lambda)
+    fmean <- fmean * (1 + 0.5*fvar*(1-lambda)/(fmean)^(2*lambda))
+    fvar <- fvar * fp2
+  }
+  # S matrix
+  S <- smatrix(object)
+  # Find G matrix
+  if(method == "wls") {
+    Lambda <- diag(1/fvar[1,])
+    G <- solve(t(S) %*% Lambda %*% S) %*% t(S) %*% Lambda
+  } else {
+    # Bottom up
+    nbottom <- NCOL(object$bts)
[…]

### 2022-10-13 rscript/hts/point forecast/hts_prophet.R

@@ -1,0 +1,176 @@
+library(tidyverse)
+library(purrr)
+library(hts)
+library(forecast)
+library(tscount)
+library(tsibble)
+library(prophet)
+
+
+#---- prophet parameter setting -------
+prophetfit_control <- prophet(
+  yearly.seasonality = FALSE, 
+  weekly.seasonality = FALSE, 
+  daily.seasonality = FALSE,
+  holidays = holidays_p,
+  interval.width = 0.95,
+  mcmc.samples = 0, # invokes rstan::optimizing (fast MAP estimation)
+  )
+
+prophetfit_control <- add_seasonality(prophetfit_control,
+                                      name = "weekly",
+                                      period = 7,
+                                      fourier.order = 3
+)
+
+prophetfit_control <- add_seasonality(prophetfit_control,
+                                      name = "yearly",
+                                      period = 365.25,
+                                      fourier.order = 10
+)
+
+#-------end of prophet parameter setting-------
+#--- Begin of function to generate prediction-------
+#x is the time series in the training set, h =forecast horizon, date_train=date corresponding to the training,date_test=date corresponding to the test set
+prophet_f <- function(x,h,date_train,date_test) {
+  df <-  data.frame(date_train, 
+                    y = x)
+  df$y = BoxCox(df$y, lambda=0.5)# use Box-Cox transformation to avoid negarive forecast
+  prophetfit <- fit.prophet(
+    m = prophetfit_control,
+    df = df
+  )
+  
+  prophetfitted <- predict(prophetfit, 
+                           date_test)
+  point_forecast <- InvBoxCox(prophetfitted$yhat, lambda=0.5)
+  #sigma <- (InvBoxCox(prophetfitted$yhat_upper,.5)-InvBoxCox(prophetfitted$yhat_lower,.5))/2/qnorm(0.975), Can we reconcile sd?
+  lower <- InvBoxCox(prophetfitted$yhat_lower,lambda=0.5)
+  upper <- InvBoxCox(prophetfitted$yhat_upper,lambda=0.5)
+  output <- as_tibble(point_forecast) %>% 
+    bind_cols( as_tibble(lower))%>% 
+    bind_cols( as_tibble(upper))
[…]

### 2022-10-13 test_msts_hts.R

@@ -1,29 +1,0 @@
-library(hts)
-set.seed(2021)
-# Bottom level series
-bts <- msts(matrix(rnorm(1500), ncol=2), seasonal.periods=c(7,365.25))
-bts[bts<0] <- 0
-bts <- round(bts)
-
-# Create hts object
-y <- hts(bts)
-#plot(y)
-# Create aggregated series
-ally <- aggts(y)
-# Produce forecasts of all series
-h <- 28
-allf <- matrix(NA, nrow = h, ncol = NCOL(ally))
-for(i in seq(NCOL(ally)))
-  allf[,i] <- stlf(ally[,i], h = h)$mean
-allf[allf<0] <- 0
-# Reconcile results
-y.f <- combinef(allf, 
-                nodes = get_nodes(y), 
-                nonnegative = TRUE)
-y.f_all <- aggts(y.f)
-y.f_all[y.f_all<0]
-
-# Add back multiple seasonality attributes
-y.f$bts <- msts(y.f$bts, start=tsp(y$bts)[2]+1/tsp(y$bts)[3],
-                seasonal.periods = attributes(y$bts)$msts)
-plot(y.f)

### 2022-10-13 forecast_hts_with_transformation.R
Split htsplus function into separate file

@@ -1,96 +1,12 @@
 library(hts)
 library(readr)
-# object = gts object
-# h = forecast horizon
-# model_function = function used to model each time series. e.g., ets or auto.arima
-# lambda = Box-Cox transformation parameter
-# level = prediction interval coverage as a percentage
-# method = reconcilation method. either "wls" or "bu"
-# nsim = number of simulations in computing prediction intervals using nonparametric method
-# htsplus = hts with BoxCox transformation
-htsplus <- function(object, h,
-                    model_function = auto.arima,
-                    lambda=NULL, level=95,
-                    method = c("wls","bu"), nsim=2000) {
-  method <- match.arg(method)
-  # Generate all time series at all levels
-  ally <- aggts(object)
-  nseries <- NCOL(ally)
-  # Do Box-Cox transformation if required
-  if(!is.null(lambda)) {
-    ally <- BoxCox(ally, lambda)
-  }
-  # Set up matrices to store results
-  fmean <- fvar <- flower <- fupper <- matrix(NA, nrow = h, ncol = nseries)
-  models <- list()
-  # Loop over all time series and compute forecast means and variances
-  for(i in seq(NCOL(ally))) {
-    models[[i]] <- model_function(ally[,i])
-    fc <- forecast(models[[i]], h=h, level=level)
-    fmean[,i] <- fc$mean
-    fvar[,i] <- ((fc$upper - fc$lower)/(2*qnorm(0.5+level/200)))^2
-  }
-  # Reverse Box-Cox transformation if required
-  # using Taylor series approximation (delta method) to get mean and variance
-  # on original scale
-  if(!is.null(lambda)) {
-    fmean <- InvBoxCox(fmean, lambda)
-    tmp <- lambda * fmean + 1
-    fp2 <- fmean^(2-2*lambda)
-    fmean <- fmean * (1 + 0.5*fvar*(1-lambda)/(fmean)^(2*lambda))
-    fvar <- fvar * fp2
-  }
-  # S matrix
-  S <- smatrix(object)
-  # Find G matrix
-  if(method == "wls") {
-    Lambda <- diag(1/fvar[1,])
-    G <- solve(t(S) %*% Lambda %*% S) %*% t(S) %*% Lambda
-  } else {
-    # Bottom up
-    nbottom <- NCOL(object$bts)
[…]

### 2022-10-13 htsplus.R

@@ -1,0 +1,86 @@
+# object = gts object
+# h = forecast horizon
+# model_function = function used to model each time series. e.g., ets or auto.arima
+# lambda = Box-Cox transformation parameter
+# level = prediction interval coverage as a percentage
+# method = reconcilation method. either "wls" or "bu"
+# nsim = number of simulations in computing prediction intervals using nonparametric method
+# htsplus = hts with BoxCox transformation
+
+htsplus <- function(object, h,
+                    model_function = auto.arima,
+                    lambda = NULL, level = 95,
+                    method = c("wls", "bu"), nsim = 2000) {
+  method <- match.arg(method)
+  # Generate all time series at all levels
+  ally <- aggts(object)
+  nseries <- NCOL(ally)
+  # Do Box-Cox transformation if required
+  if (!is.null(lambda)) {
+    ally <- BoxCox(ally, lambda)
+  }
+  # Set up matrices to store results
+  fmean <- fvar <- flower <- fupper <- matrix(NA, nrow = h, ncol = nseries)
+  models <- list()
+  # Loop over all time series and compute forecast means and variances
+  for (i in seq(NCOL(ally))) {
+    models[[i]] <- model_function(ally[, i])
+    fc <- forecast(models[[i]], h = h, level = level)
+    fmean[, i] <- fc$mean
+    fvar[, i] <- ((fc$upper - fc$lower) / (2 * qnorm(0.5 + level / 200)))^2
+  }
+  # Reverse Box-Cox transformation if required
+  # using Taylor series approximation (delta method) to get mean and variance
+  # on original scale
+  if (!is.null(lambda)) {
+    fmean <- InvBoxCox(fmean, lambda)
+    tmp <- lambda * fmean + 1
+    fp2 <- fmean^(2 - 2 * lambda)
+    fmean <- fmean * (1 + 0.5 * fvar * (1 - lambda) / (fmean)^(2 * lambda))
+    fvar <- fvar * fp2
+  }
+  # S matrix
+  S <- smatrix(object)
+  # Find G matrix
+  if (method == "wls") {
+    Lambda <- diag(1 / fvar[1, ])
+    G <- solve(t(S) %*% Lambda %*% S) %*% t(S) %*% Lambda
+  } else {
+    # Bottom up
+    nbottom <- NCOL(object$bts)
[…]

### 2022-10-14 data/holidays.R
Data processing code

@@ -1,0 +1,91 @@
+# Script to turn original data file of holidays into an RDS for later use
+# Input: holiday_rugby.xlsx 
+# Output: holiday_dummy.rds
+
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+
+# read holiday data
+holiday <- readxl::read_excel(here::here("data/holiday_rugby.xlsx")) %>%
+  select(date, public_holiday, school_holiday) %>%
+  mutate(date = as_date(date)) %>%
+  mutate(
+    public_holiday_d = if_else(is.na(public_holiday), 0, 1),
+    school_holiday_d = if_else(is.na(school_holiday), 0, 1)
+  ) %>%
+  mutate(
+    xmas = if_else(public_holiday == "Christmas Day", "1", "0"),
+    new_years_day = if_else(public_holiday == "New Years Day", 1, 0)
+  ) %>%
+  mutate_at(vars(xmas, new_years_day), ~ replace(., is.na(.), 0))
+
+# create dummies for holidays
+holiday_dummy <- holiday %>%
+  select(
+    date,
+    public_holiday_d,
+    school_holiday_d,
+    xmas,
+    new_years_day
+  ) %>%
+  mutate_at(
+    vars(public_holiday_d, school_holiday_d, xmas, new_years_day),
+    ~ as_factor(.)
+  ) %>%
+  as_tsibble(index = date)
+
+# Save as RDS for reading in by other scripts
+write_rds(holiday_dummy, here::here("data/holiday_dummy.rds"))
+
+
+# Other holiday tibbles created below, no longer used (?)
+
+holiday_school <- holiday_dummy %>%
+  select(date, school_holiday_d) %>%
+  filter(school_holiday_d == 1) %>%
+  mutate(holiday = "school_holiday") %>%
+  select(-school_holiday_d)
+holiday_public <- holiday_dummy %>%
+  select(date, public_holiday_d) %>%
+  filter(public_holiday_d == 1) %>%
+  mutate(holiday = "public_holiday") %>%
+  select(-public_holiday_d)
+holiday_xmas <- holiday_dummy %>%
+  select(date, xmas) %>%
+  filter(xmas == 1) %>%
+  mutate(holiday = "xmas") %>%
+  select(-xmas)
+holiday_newyear <- holiday_dummy %>%
+  select(date, new_years_day) %>%
+  filter(new_years_day == 1) %>%
[…]

### 2022-10-14 data/incidents.R

@@ -1,0 +1,161 @@
+# Script to turn original data file of incidents into an RDS for later use
+# Two different RDS are created with the same data: a tsibble and a gts object
+# Input: Nature_of_Incidents.xlsx
+# Outputs: incidents_tsbl.rds, incidents_gts.rds
+
+library(tidyverse)
+library(lubridate)
+library(tsibble)
+
+# Read data
+incidents_original <- here::here("data/Nature_of_Incidents_Attended.xlsx") |> 
+  readxl::read_excel() |>
+  mutate(date = as_date(Incident_Date)) |>
+  janitor::clean_names() |>
+  force_tz(date, tz = "GB")
+
+# Nature of incidents with low volume
+nature_of_incident_low <- c(
+  "AUTOMATIC CRASH NOTIFICATION",
+  "INACCESSIBLE INCIDENT/OTHER ENTRAP",
+  "INTERFACILITY EVALUATION/TRANSFER",
+  "MAJOR INCIDENT - OVERRIDE PROQA",
+  "TRANSFER/INTERFACILITY/PALLIATIVE"
+)
+
+# Count number of incidents
+incidents <- incidents_original |>
+  mutate(nature_of_incident = as.numeric(nature_of_incident)) |>
+  mutate(nature_of_incident = ifelse(!is.na(nature_of_incident), nature_of_incident_description, "upgrade")) |>
+  mutate(nature_of_incident = ifelse(nature_of_incident %in% nature_of_incident_low, "other", nature_of_incident)) |>
+  group_by(lhb_code, category, nature_of_incident, date) |>
+  summarise(incidents = sum(total_incidents), .groups = "drop") |>
+  mutate(
+    category = factor(category, level = c("RED", "AMBER", "GREEN")),
+    nature_of_incident = factor(nature_of_incident),
+    lhb_code = factor(lhb_code)
+  )
+# Convert to tsibble
+incidents_tsbl <- incidents |>
+  as_tsibble(index = date, key = c(lhb_code, category, nature_of_incident)) |>
+  fill_gaps(incidents = 0, .full = TRUE)
+# Save as rds
+write_rds(incidents_tsbl, here::here("data/incidents_tsbl.rds"))
+
[…]

### 2022-10-14 data/prepate_data_for_hts.R

@@ -1,233 +1,0 @@
-library(glue)
-library(forecast)
-library(hts)
-
-library(tidyverse)
-library(tsibble)
-library(fable)
-library(fabletools)
-library(feasts)
-library(lubridate)
-library(fable.tscount)
-library(tscount)
-
-# import data
-f_horizon <- 5*7
-# Read data
-incidents_original <- readxl::read_excel("data/Nature_of_Incidents_Attended.xlsx") %>%
-  mutate(date = as_date(Incident_Date)) %>%
-  janitor::clean_names() %>%
-  force_tz(date, tz = "GB") 
-# nature of incidents with low volume
-nature_of_incident_low <- c("AUTOMATIC CRASH NOTIFICATION",
-                            "INACCESSIBLE INCIDENT/OTHER ENTRAP",
-                            "INTERFACILITY EVALUATION/TRANSFER",
-                            "MAJOR INCIDENT - OVERRIDE PROQA",
-                            "TRANSFER/INTERFACILITY/PALLIATIVE")
-
-# read holiday data
-holiday <- readxl::read_excel("data/holiday_rugby.xlsx") %>% 
-  select(date,public_holiday,school_holiday) %>% 
-  mutate(date=as_date(date)) %>% 
-  mutate(public_holiday_d=if_else(is.na(public_holiday), 0,1),
-         school_holiday_d=if_else(is.na(school_holiday), 0,1)) %>% 
-  mutate(xmas=if_else(public_holiday=="Christmas Day","1","0"),
-         new_years_day=if_else(public_holiday=="New Years Day",1,0)) %>% 
-  mutate_at(vars(xmas,new_years_day), ~replace(.,is.na(.),0)) 
-
-# create dummies for holidays
-holiday_dummy <- holiday %>% select(date,
-                                    public_holiday_d,
-                                    school_holiday_d,
-                                    xmas,
-                                    new_years_day) %>% 
-  mutate_at(vars(public_holiday_d,school_holiday_d,xmas,new_years_day),
-            ~as_factor(.)) %>% 
-  as_tsibble(index = date)
[…]

### 2022-10-14 rscript/result/analysing_results.Rmd
Moved R scripts into rscript folder

---
title: "analysing results"
author: "Bahman Rostami-Tabar"
date: "12/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r read-result, options}

hts_result_ets <- read_rds("result/ets_hts_rmsse_long_h.rds")
hts_result_arima <- read_rds("result/arima_hts_rmsse_long_h.rds")
result <- hts_result_ets %>% bind_rows(hts_result_arima)
rm <- result %>% filter(!is.nan(error2_scale))
rmse_for_analysis <- rm %>% filter(!is.infinite(error2_scale))

sum(is.infinite(rm$error2_scale))
```

```{r label, options}

rmsse_approach_level_id <-  rmse_for_analysis %>% group_by(level,approach,.id) %>% 
  summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

rmsse_approach_id <-  rmsse_approach_level_id %>% group_by(approach,level) %>% 
  summarise(rmsse=mean(rmsse), .groups = "drop")
```

# overall

```{r label, options}
r <- rmsse_approach_level_id %>% group_by(approach) %>% 
  summarise(mean_rmsse=mean(rmsse), median_rmsse=median(rmsse), .groups = "drop")
r

```

## Total

```{r label, options}
rmsse_total <- rmse_for_analysis %>% filter(level=="Total") %>% dplyr::select(-level) %>% 
  group_by(approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_total, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  ggthemes::theme_few()
```

## control level

```{r label, options}
rmsse_control <- rmse_for_analysis %>% filter(level %in% c("G1/S","G1/N","G1/C")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_control, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()
```

```{r label, options}
rmsse_control_h <- rmse_for_analysis %>% filter(level %in% c("G1/S","G1/N","G1/C")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_control_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

## Health board level

```{r label, options}
rmsse_hb <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G2/")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
  ggthemes::theme_few()
```

```{r label, options}
rmsse_hb_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G2/")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

## Priority

```{r label, options}
rmsse_priority <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G3/")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_priority, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
  geom_boxplot()+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()
```

```{r label, options}
rmsse_priority_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G3/")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_priority_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

## Nature of incidents

```{r label, options}
rmsse_nature <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G4/")) %>%
  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

rmsse_nature_all <- rmsse_nature %>% group_by(approach,level) %>% summarise(rmsse=mean(rmsse), .groups = "drop") 
rmsse_nature_all %>% pivot_wider(names_from = approach, values_from = rmsse)%>% arrange(arima_bu) %>% View()
```

```{r label, options}
rmsse_nature_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G4/")) %>%
  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")

ggplot(rmsse_hb_h, aes(x=h,y=rmsse, colour=factor(approach)))+
  geom_point()+
  geom_line()+
  scale_x_continuous(breaks = seq(1,35,1))+
  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
  ggthemes::theme_few()+
  ggthemes::scale_color_colorblind()
```

##

```{r label, options}
rr <- rmsse_approach_level_id %>% group_by(approach,level) %>% 
  summarise(mean_rmsse=mean(rmsse),.groups = "drop")
ggplot(rmsse_approach_id)+
  geom_boxplot(aes(y=rmsse, x=approach))+
  coord_cartesian(ylim = c(0,2))
```
##

```{r label, options}
rr <- rmsse_approach_level_id %>% group_by(approach,level) %>% 
  summarise(mean_rmsse=mean(rmsse),.groups = "drop")
ggplot(rmsse_approach_id)+
  geom_boxplot(aes(y=rmsse, x=approach))+
  coord_cartesian(ylim = c(0,2))
```

##

```{r label, options}
rmsse_hd <-  rmse_for_analysis %>% group_by(h,.id,approach) %>% 
  summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
ggplot(rmsse_hd)+
  geom_boxplot(aes(y=rmsse, x=factor(h)))+
  facet_wrap(vars(approach))
```

### 2022-10-14 rscript/data/incidents.R

@@ -8,7 +8,7 @@ library(lubridate)
 library(tsibble)
 
 # Read data
-incidents_original <- here::here("data/Nature_of_Incidents_Attended.xlsx") |> 
+incidents_original <- here::here("data/Nature_of_Incidents_Attended.xlsx") |>
   readxl::read_excel() |>
   mutate(date = as_date(Incident_Date)) |>
   janitor::clean_names() |>
@@ -157,5 +157,4 @@ incident_all <- incident_gts |>
   select(-delet, -date)
 
 # Write out final rds
-
 write_rds(incident_all, here::here("data/incidents_gts.rds"))

### 2022-10-14 rscript/fable/fit_model_fable.R

@@ -1,98 +1,101 @@
 library(tidyverse)
 library(tsibble)
 library(fable)
-library(fabletools)
 library(feasts)
 library(lubridate)
 library(fable.tscount)
-library(tscount)
 
-f_horizon <- 5*7
-# Read data
-incidents_original <- readxl::read_excel("data/Nature_of_Incidents_Attended.xlsx") %>%
-  mutate(date = as_date(Incident_Date)) %>%
-  janitor::clean_names() %>%
-  force_tz(date, tz = "GB") 
-# nature of incidents with low volume
-nature_of_incident_low <- c("AUTOMATIC CRASH NOTIFICATION",
-                            "INACCESSIBLE INCIDENT/OTHER ENTRAP",
-                            "INTERFACILITY EVALUATION/TRANSFER",
-                            "MAJOR INCIDENT - OVERRIDE PROQA",
-                            "TRANSFER/INTERFACILITY/PALLIATIVE")
+# Read data previously processed in data folder and add aggregation levels
+incidents <- readRDS(here::here("data/incidents_tsbl.rds")) |>
+  aggregate_key(nature_of_incident * category * lhb_code,
+    incidents = sum(incidents)
+  )
 
-# read holiday data
-holiday <- readxl::read_excel("data/holiday_rugby.xlsx") %>% 
-  select(date,public_holiday,school_holiday) %>% 
-  mutate(date=as_date(date)) %>% 
-  mutate(public_holiday_d=if_else(is.na(public_holiday), 0,1),
-         school_holiday_d=if_else(is.na(school_holiday), 0,1)) %>% 
-  mutate(xmas=if_else(public_holiday=="Christmas Day","1","0"),
-         new_years_day=if_else(public_holiday=="New Years Day",1,0)) %>% 
-  mutate_at(vars(xmas,new_years_day), ~replace(.,is.na(.),0)) 
+# Add holidays
+incidents <- incidents |>
+  left_join(
+    readRDS(here::here("data/holiday_dummy.rds")),
+    by = "date"
+  )
 
-# create dummies for holidays
-holiday_dummy <- holiday %>% select(date,
-                                    public_holiday_d,
-                                    school_holiday_d,
-                                    xmas,
-                                    new_years_day) %>% 
[…]

### 2022-10-14 rscript/fable/fit_model_fable.R
Updated fable code

@@ -7,9 +7,14 @@ library(fable.tscount)
 
 # Read data previously processed in data folder and add aggregation levels
 incidents <- readRDS(here::here("data/incidents_tsbl.rds")) |>
-  aggregate_key(nature_of_incident * category * lhb_code,
-    incidents = sum(incidents)
-  )
+  # Temporarily only consider a small part of data until everything works
+  filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |> 
+  select(-nature_of_incident, -lhb_code) |> 
+  # Create aggregates
+  aggregate_key(category, incidents = sum(incidents))
+  #aggregate_key(nature_of_incident * category * lhb_code,
+  #  incidents = sum(incidents)
+  #)
 
 # Add holidays
 incidents <- incidents |>
@@ -18,19 +23,17 @@ incidents <- incidents |>
     by = "date"
   )
 
-# Check that some of the upper levels look ok
-incidents |>
-  filter(is_aggregated(nature_of_incident), is_aggregated(category)) |>
-  autoplot(incidents)
-incidents |>
-  filter(is_aggregated(nature_of_incident), is_aggregated(lhb_code)) |>
-  autoplot(incidents)
-
 # First pass using a simple training/test set keeping last 6 weeks for testing
-train <- filter(incidents, date <= max(date) - 42)
-test <- filter(incidents, date > max(date) - 42)
+train <- incidents |> 
+  filter(date <= max(date) - 42)
+test <- incidents |> 
+  filter(date > max(date) - 42)
 
 # Fit some models
+# Parallelization doesn't work for TSCOUNT
+# Issue raised at https://github.com/mitchelloharawild/fable.tscount/issues/2
+# library(future)
+# plan(multisession)
 fit_incident <- train %>%
   model(
     NAIVE = NAIVE(sqrt(incidents)),
@@ -50,18 +53,24 @@ fit_reconcile <- fit_incident %>%
     wls_ETS = min_trace(ETS, method = "wls_struct")
   )
 
-# Produce forecasts
+# Produce forecasts (need to use simulated sample paths for TSCOUNT)
 fcst_incident <- fit_reconcile %>%
-  forecast(new_data = test)
+  forecast(new_data = test, simulate = TRUE)
 
 # Look at MASE over several levels for MinT
 fcst_accuracy <- fcst_incident %>%
[…]

### 2022-10-14 rscript/fable/fit_model_fable.R
Code speed up

@@ -10,9 +10,10 @@ incidents <- readRDS(here::here("data/incidents_tsbl.rds")) |>
   # Temporarily only consider a small part of data until everything works
   filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |> 
   select(-nature_of_incident, -lhb_code) |> 
-  # Create aggregates
   aggregate_key(category, incidents = sum(incidents))
-  #aggregate_key(nature_of_incident * category * lhb_code,
+  # End of temporary section of code
+  # Uncomment next 3 lines when ready to scale up
+  #aggregate_key(nature_of_incident * category * lhb_code, 
   #  incidents = sum(incidents)
   #)
 
@@ -24,17 +25,15 @@ incidents <- incidents |>
   )
 
 # First pass using a simple training/test set keeping last 6 weeks for testing
-train <- incidents |> 
-  filter(date <= max(date) - 42)
-test <- incidents |> 
-  filter(date > max(date) - 42)
+train <- incidents |> filter(date <= max(date) - 42)
+test <- incidents |> filter(date > max(date) - 42)
 
 # Fit some models
 # Parallelization doesn't work for TSCOUNT
 # Issue raised at https://github.com/mitchelloharawild/fable.tscount/issues/2
 # library(future)
 # plan(multisession)
-fit_incident <- train %>%
+fit_incident <- train |>
   model(
     NAIVE = NAIVE(sqrt(incidents)),
     ETS = ETS(sqrt(incidents)),
@@ -42,72 +41,80 @@ fit_incident <- train %>%
         + public_holiday_d + school_holiday_d + xmas + new_years_day, 
       link = "log", model = list(past_obs = 1:13))
   )
-# Add reconciliation constraints and methods
-fit_reconcile <- fit_incident %>%
+# Add reconciliation constraints and produce forecasts
+# Need to handle tscount separately so we can simulate forecast distributions
+ets_forecast <- fit_incident |>
+  select(-TSCOUNT) |> 
+  reconcile(
+    td_ETS = top_down(ETS),
+    bu_ETS = bottom_up(ETS),
+    wls_ETS = min_trace(ETS, method = "wls_struct"),
+    wls_ELS = min_trace(ETS, method = "mint_shrink")
+  ) |> 
[…]

### 2022-10-18 rscript/fable/fit_model_fable.R
Write rds file and scale up to all data

@@ -8,14 +8,14 @@ library(fable.tscount)
 # Read data previously processed in data folder and add aggregation levels
 incidents <- readRDS(here::here("data/incidents_tsbl.rds")) |>
   # Temporarily only consider a small part of data until everything works
-  filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |> 
-  select(-nature_of_incident, -lhb_code) |> 
-  aggregate_key(category, incidents = sum(incidents))
+  #filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |> 
+  #select(-nature_of_incident, -lhb_code) |> 
+  #aggregate_key(category, incidents = sum(incidents))
   # End of temporary section of code
   # Uncomment next 3 lines when ready to scale up
-  #aggregate_key(nature_of_incident * category * lhb_code, 
-  #  incidents = sum(incidents)
-  #)
+  aggregate_key(nature_of_incident * category * lhb_code, 
+    incidents = sum(incidents)
+  )
 
 # Add holidays
 incidents <- incidents |>
@@ -41,6 +41,7 @@ fit_incident <- train |>
         + public_holiday_d + school_holiday_d + xmas + new_years_day, 
       link = "log", model = list(past_obs = 1:3))
   )
+write_rds(fit_incident, here::here("rscript/fable/fit_incident.rds"))
 # Add reconciliation constraints and produce forecasts
 # Need to handle tscount separately so we can simulate forecast distributions
 ets_forecast <- fit_incident |>

### 2022-10-18 manuscript/forecasting_ed.tex
manuscript writing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Annals of Emergency Medicine}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Forecasting},
  pdfauthor={Bahman Rostami-Tabar; Rob J. Hyndman},
  pdfkeywords={Forecasting, Ambulance, Hierarchical structure},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Forecasting}
\author[1]{Bahman Rostami-Tabar%
\corref{cor1}%
\fnref{fn1}}
 \ead{rostami-tabarb@cardiff.ac.uk} 
\author[2]{Rob J. Hyndman%
%
}
 \ead{rob.hymdman@monash.ac.edu}

\affiliation[1]{organization={Cardiff University, Cardiff Business
School},country={United Kingdom},countrysep={,},postcode={CF10
3EU},postcodesep={}}
\affiliation[2]{organization={Monash University, Department of
Econometrics and Business
Statistics},country={Australia},countrysep={,},postcode={VIC
3800},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{This is the first author footnote.}

\begin{abstract}
This is the abstract. An accurate demand forecasting is crucial in
Emergency medicine to depict various courses of action that can result
in massive savings in terms of patient lives. Inability to match the
staff with the demand results in an overcrowding care system which is a
serious problem causing challenging situations on patient flow. Also, it
is related with increasing length of stay, low patient satisfaction,
increasing health care costs, inaccuracy in electronic medical record,
and reported waiting times.
\end{abstract}

\begin{keyword}
    Forecasting \sep Ambulance \sep 
    Hierarchical structure
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, breakable, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

An accurate demand forecasting is crucial in Emergency medicine to
depict various courses of action that can result in massive savings in
terms of patient lives. Inability to match the staff with the demand
results in an overcrowding care system which is a serious problem
causing challenging situations on patient flow. Also, it is related with
increasing length of stay, low patient satisfaction, increasing health
care costs, inaccuracy in electronic medical record, and reported
waiting times.without incurring last-minute expenses, such as overtime
or supplemental staffing {[}Reference{]}.

An accurate forecasting of the daily demand enables managers to match
staff to meet anticipated patients, reconfigure units and redeploy staff
and vehicles. This will have many advantages for patients, staff and the
quality of healthcare services provided.

The ability to accurately forecast the demand of future health services
is critical for planning decisions aimed at improving the quality of
service delivery in the health care. Forecasting has been used in
various fields of healthcare to inform planning decisions at the
strategic, tactical and operational levels. There is at least one or
multiple reasons why we need to produce a forecast. Most often reasons
determine clearly what to forecast and which level of forecast
granularities are required. There are typically some planning process at
strategic(long-term) level such as dimensioning resource capacities and
budgeting; tactical(medium-term) level such as temporary capacity
expansions, staff-shift scheduling, and inventory control; and
operational(short-term) level such as planning rosters for staff and
allocate resources and medical item distribution. Each planning level
will typically have different forecasting needs in terms of granularity
and horizon (how far into the future do we forecast). A natural question
that arise from these examples is how to generate forecasts to support
planning decisions at multiple granularities and horizons. In order to
answer this question, we first need to understand some fundamental
aspects of forecasting in healthcare:

Daily forecasts are required to inform the short-term planning for the
current and the upcoming shifts of the day. This involves the decision
making related to the execution of the delivery process for various
health care services such as Ambulatory ,Emergency. Daily forecasts are
important at various levels:

\begin{itemize}
\item
  \textbf{Nature of incidents}
\item
  \textbf{Category: Red, Amber, Green}
\item
  \textbf{Health board}
\item
  \textbf{Country level}
\end{itemize}

Implementing and sustaining improvements in hospital-wide flow requires
alignment, cooperation, and coordination between hospital units and
departments. Without effective executive oversight and collaboration,
teams operate in isolation from one another and the aggregated impact of
their efforts is limited. In many cases, this isolation leads to
duplicative work, rework, or work that runs counter to overall goals to
improve hospital-wide patient flow

The combination of demand forecast, incidents being attended, resource
availability and delays at hospitals, provide information on the state
of the unscheduled care system across the emergency medicine services.
Having this full picture enables the delivery managers to focus on the
areas that require intervention to enable the most effective delivery of
the service to the patients.

Despite a large number of studies dedicated to forecasting in
healthcare{[}ref{]}, this ques- tion unanswered in healthcare
forecasting. Almost all studies assume that the level of data
granularity matches with the level of forecast requirement, e.g.~daily
series is used to generate daily forecast. Moreover, in the presence of
the high data granularity, the recommendation is to first aggregate the
data into a requested forecast granularity level and then generate the
forecast(Goodwin, 2018). However, there is no empirical evidence to
support this claim

\hypertarget{lit}{%
\section{Research background}\label{lit}}

Table Table~\ref{tbl-lit} summarise studies on forecasting in emergency
and urgent care.

\hypertarget{tbl-lit}{}
\begin{table}[!h]
\caption{\label{tbl-lit}summary of literature review }\tabularnewline

\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{l|r|l|l|>{\raggedright\arraybackslash}p{15em}|l}
\hline
Author & Year & Forecast variable & Forecast Horizon & Method & Measure\\
\hline
McCarthy 
2008 & 2008 & ED arrivals & 1 day & Poisson log-linear regression model, including temporal factors, patient characteristics and climatic factors & 95\% CI\\
\hline
Cote et al. & 2013 & ED arrivals & 1 day & Fourier regression & R\textasciicircum{}2,  Standard Error\\
\hline
Kim et al. & 2014 & Hospital demand & 1h, 7days, 
30days & Linear regression; Exponential smoothing; ARIMA; GARCH; VAR & MAPE\\
\hline
1 day & 2007 & Ambulance demand & 1day & Regression & RMSE\\
\hline
Hertzum & 2017 & ED arrivals
ED occupancy & 1day & linear regression; SARIMA; Naïve & MAE, MAPE, 
MASE\\
\hline
Choudhury and Urena & 2020 & ED arrivals & 1day & ARIMA; Holt-winters; TBATS; ANN & RMSE, ME\\
\hline
Steins et al. & 2019 & Ambulance call & 1day & ZIP and ZINB regression;
 moving average with seasonality weights & ME, MAE, RMSE\\
\hline
Jones et al. & 2009 & ED census & 1day & VAR;  Holt winters & MAE\\
\hline
Morzuch and Allen & 2006 & ED arrivals & 7 days & Regression; ARIMA; Exponential smoothing & RMSE\\
\hline
Taylor & 2008 & centre volume & 14 days & ARMA, Exponential smoothing; Dynamic harmonic regression; Seasonal random walk; Seasonal mean;
Random walk; Mean; Median; Simple exponential smoothing & MAE
RMSE\\
\hline
Gijo and Balakrishna & 2016 & call volume & 7 days & SARIMA & MMSE\\
\hline
\end{tabular}}
\end{table}

Focus on the ambulance services!

\citet{gijo2016sarima} generated a time series model to forecast the
daily and hourly call volume at all centre handling emergency ambulance
services. Since historical data showed seasonality, SARIMA models were
investigated. Regarding the daily model, the authors generated a SARIMA
model, which, however, resulted in the forecast error (MMSE) that
significantly increased when the lead time exceeded 8 days. On the other
hand, the SARIMA model proposed to forecast the log-calls an hourly
basis. This model was found to fit well the model both for shorter and
longer lead times.

Forecasting at the daily level of granularity is generally use for
roaster planing and deciding when to contact staff on call for instance.
\citet{luo2017hospital} use a combination of seasonal ARIMA (SARIMA) and
a single exponential smoothing (SES) model to forecast daily outpatient
visits 1 week ahead.They indicate that using combinatorial model can be
more effective than each model separately. \citet{whitt2019forecasting}
examine several alternative models to forecast the total daily arrival
for 1-7 days ahead, including a linear regression based on calendar and
weather variables, seasonal autoregressive integrated moving average
with exogenous regressors (SARIMAX) and the multilayer perceptron (MLP)
model. Using a daily ED admission for 3 years, they show that SARIMAX
provides the most accurate daily forecast by MSE measure.
\citet{marcilio2013forecasting} compare the forecast accuracy of the
Generalised Linear Model (GLM), Generalised Estimating Equations (GEE),
and Seasonal Autoregressive Integrated Moving Average (SARIMA) methods
using total daily patient visits to an ED using MAPE. They conclud that
GLM and GEE models provide more accurate forecasts than SARIMA model.
\citet{rostami2020anticipating} propose a model to forecast daily ED
attendance with consideration of different types of holidays, weekday
effects, auto-regressive effects, long-term trends and date effects.
They provide probabilistic forecasts to quantify uncertainties in future
ED attendance and they show that the proposed model outperforms three
time series techniques including 1) Naive, 2) AutoRegressive, AR(p), 3)
exponential smoothing state space model (ETS) and a regression model
without considering special events as alternatives. \citet{zhou2018time}
SARIMA, NARNN and the hybrid SARIMA-NARNN to forecast the daily number
of new admission inpatients. The root mean square error (RMSE), mean
absolute error (MAE) and mean absolute percentage error (MAPE) were used
to compare the forecasting performance among the three models. They show
that NARNN model outperforms the others. \citet{sun2009forecasting} use
SARIMA model to forecast daily patient volume for each patient acuity
level. MAPE is used to choose the best-fit model. They fitted separate
ARIMA models to the three categories of acuity and overall data. They
conclude that the ARIMA model is effective for both short term (weekly)
and long term (three months) forecast horizons. Moreover, they observe
that the impact of weather is not significant.
\citet{zinouri2018modelling} use SARIMA to develop a statistical
prediction model that provides short-term forecasts of daily surgical
demand. Our results suggest that the proposed SARIMA model can be useful
for estimating surgical case volumes 2--4 weeks prior to the day of
surgery, which can support robust and reliable staff schedules.
\citet{moustris2012seven} use Artificial Neural Network (ANN) models to
forecast the total weekly(7 days) number of Childhood Asthma Admission.
Three different ANN models were developed and trained to forecast the
admission of different age subgroups and for the whole study population.
\citet{mccoy2018assessment} compare SARIMA, Prophet and Snaive methods
for prediction 365 days of hospital discharge using a daily hospital
discharge volumes at 2 large, New England academic medical centers. They
show that Prophet model outperforms SARIMA and Snaive.

\citet{khaldi2019forecasting} investigate the combination of the
Artificial Neural Networks (ANNs) with a signal decomposition technique
named Ensemble Empirical Mode decomposition (EEMD), to make one step
ahead weekly forecasting of patients arrivals to ED. using seven years
of weekly demand. The results of the proposed model were compared
against ANN without signal decomposition, ANN with Discrete wavelet
Transform (DWT). The results show that the combined forecasts outperform
the benchmarking models.

\citet{steins2019forecasting} aimed to develop a forecasting model for
predicting the number of ambulance services calls per hour and
geographical areas, to support managers in decisions-making and to
investigate which ones were the factors that affected the number of
calls. Data collected consisted of a time and location of historical
ambulance call data for three counties in Sweden and a list of
explanatory factors of the area, such as socioeconomic and geographic.
In order to deal with large number of zeros in the data, authors
developed zero-inflated Poisson (ZIP) and zero-inflated negative
binomial (ZINB) regression models. These were then compared to the
currently existing forecasting system, based on moving average with
seasonality weights, using ME, MAE and RMSE. Firstly, the factors
affecting the number of ambulance calls were found to be the following:
population in different age groups, median income, length of road,
number of nightlife spots (the number of restaurants), day of the week
and hour of the day. Secondly, it was found that the older population
(65-100) generated more ambulance calls and that ZIP model performs
better than the current model. However, the improvement provided by the
more advance model was not much greater than the one provided by the
existing model. Authors suggested that it could be because the
population, which was used as an independent variable in both models,
was so dominant compared to the other factors. Moreover, both models
either underestimated or overestimated the number of calls. Authors
suggested that the inability to capture a positive trend resulted in
underestimation, and the opposite was due to a negative trend.
Therefore, authors recommended that further research should add certain
temporal variables able to capture the trend.

\hypertarget{data}{%
\section{Data}\label{data}}

Operations Department and I suppose it's split into three different
categories. You've got\ldots{} you've got Response which is the
ambulances. You've got the Control which is the people who take the
telephone calls. And then you've got Resilience which are the things
that people like have the hazardous response teams so they'll go and
save people from water and maybe with fires as well. They've got\ldots{}
they're specially trained in different things so there's the contingency
sort of there. So from a Control perspective, they look at\ldots{} they
look at it at least twice a week to see what the\ldots{} what it\ldots{}
how busy it is. And then they staff\ldots{} they\ldots{} they're looking
to staff to the worst level, knowing that they will get some
people\ldots{} there will be some abstractions. So they won't manage
to\ldots{} to manage that, because their key area {[}inaudible{]}.
Excuse me, they try to answer within the\ldots{} the phone within three
seconds 95\% of the time. I'm not sure exactly. It's something like
that. That's very high up there. That's what they're trying to do.
So\ldots{} and because I got it broken down at the Control level, so
we've got North Wales level for calls, and we've also got it broken down
by the North, the Mid, and the South Control. It's actually called the
North, the South East, and Central and West. Those are the three control
areas. So they can see how those break down, knowing that at the moment
we've got particularly high demand in the South East, and it's forecast
to be just bonkers

\hypertarget{design}{%
\section{Experimental set}\label{design}}

\hypertarget{result}{%
\section{Result and discussion}\label{result}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\renewcommand\refname{References}
  \bibliography{mybibfile.bib}

\end{document}

### 2022-10-18 manuscript/forecasting_ed.qmd

--- 2022-10-13 manuscript/manuscript.Rmd
@@
 ---
-title: Forecasting for emergency medcine
+title: Forecasting
+format:
+  elsevier-pdf:
+    keep-tex: true
+    journal:
+      name: Annals of Emergency Medicine
+      formatting: preprint
+      model: 3p
+      cite-style: authoryear
+  elsevier-html: 
+    toc: true
 author:
-  - name: Author1
-    email: emaail@example.com
-    affiliation: University1
-    footnote: 1
-  - name: Author2
-    email: email2@example.com
-    affiliation: University2
-    footnote: 2
-address:
-  - code: University1
-    address: adress1
-  - code: University2
-    address: adress2
-footnote:
-  - code: 1
-    text: "Corresponding Author"
-
+  - name: Bahman Rostami-Tabar
+    email: rostami-tabarb@cardiff.ac.uk
+    affiliations: 
+        - id: carbs
+          name: Cardiff University
+          department: Cardiff Business School
+          country: United Kingdom
+          postal-code: CF10 3EU
+    attributes:
+        corresponding: true
+    note: This is the first author footnote.
+  - name: Rob J. Hyndman
+    email: Rob.Hyndman@monash.edu
+    affiliations:
+        - id: monash
+          name: Monash University
+          department: Department of Econometrics and Business Statistics
+          country: Australia
+          postal-code: VIC 3800
 abstract: |
-   The Objective of this paper
-keywords:  ""
-journal: "Which journal?"
-date: "`r Sys.Date()`"
-geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
-#linenumbers: true
-numbersections: true
-#output: rticles::elsevier_article
-header-includes:  
- \usepackage{adjustbox, float,lscape} #use the 'float' package
-output: 
-  bookdown::pdf_book:
-    base_format: rticles::elsevier_article
-    extra_dependencies: ["flafter"]
-    keep_tex: true
+  This is the abstract.
+An accurate demand forecasting is crucial in Emergency medicine to depict various courses of action that can result in massive savings in terms of patient lives.
+Inability to match the staff with the demand results in an overcrowding care system which is a serious problem causing challenging situations on patient flow.
+Also, it is related with increasing length of stay, low patient satisfaction, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times.
+keywords: [Forecasting, Ambulance, Hierarchical structure]
 bibliography: mybibfile.bib
-biblio-style: authoryear
-#csl: elsevier-without-titles.csl
-#biblio-style: abbrevnat
 ---
@@
 library(readxl)
+library(here)
 ```
@@
 
+ The ability to accurately forecast the demand of future health services is critical for planning decisions aimed at improving the quality of service delivery in the health care.
+Forecasting has been used in various fields of healthcare to inform planning decisions at the strategic, tactical and operational levels.
+There is at least one or multiple reasons why we need to produce a forecast.
+Most often reasons determine clearly what to forecast and which level of forecast granularities are required.
+There are typically some planning process at strategic(long-term) level such as dimensioning resource capacities and budgeting; tactical(medium-term) level such as temporary capacity expansions, staff-shift scheduling, and inventory control; and operational(short-term) level such as planning rosters for staff and allocate resources and medical item distribution.
+Each planning level will typically have different forecasting needs in terms of granularity and horizon (how far into the future do we forecast).
+A natural question that arise from these examples is how to generate forecasts to support planning decisions at multiple granularities and horizons.
+In order to answer this question, we first need to understand some fundamental aspects of forecasting in healthcare:
+
 Daily forecasts are required to inform the short-term planning for the current and the upcoming shifts of the day.
@@
 
+Implementing and sustaining improvements in hospital-wide flow requires alignment, cooperation, and coordination between hospital units and departments.
+Without effective executive oversight and collaboration, teams operate in isolation from one another and the aggregated impact of their efforts is limited.
+In many cases, this isolation leads to duplicative work, rework, or work that runs counter to overall goals to improve hospital-wide patient flow
+
 The combination of demand forecast, incidents being attended, resource availability and delays at hospitals, provide information on the state of the unscheduled care system across the emergency medicine services.
@@
 
+<!-- An accurate demand forecast by hour of the day enables planners to match staff to meet anticipated patients, reconfigure units and redeploy staff.
+This will have many advantages for both patients, staff and the quality of provided services.
+Hourly forecasts are required to inform the short-term operational planning for the current and the upcoming shifts of the day.
+This involves the short-term decision making related to the execution of the delivery process in ED.
+The combination of an hourly forecast demand, current staff being occupied, resource availability and waiting times at ED, provide information on the state of the unscheduled care system across the service.
+Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients.
+Moreover, there are often unplanned events, such as walk-in patients, extended consultation times, during a day or a week.
+Staff capacities may be adjusted based on the predicted demand fluctuations by using part-time, on-call nurses, staff overtime and voluntary absenteeism.
+That may also require to adjust the scheduling.
+-->
+
+Despite a large number of studies dedicated to forecasting in healthcare[ref], this ques-
+tion unanswered in healthcare forecasting. Almost all studies assume that the level of
+data granularity matches with the level of forecast requirement, e.g. daily series is used
+to generate daily forecast. Moreover, in the presence of the high data granularity, the
+recommendation is to first aggregate the data into a requested forecast granularity level
+and then generate the forecast(Goodwin, 2018). However, there is no empirical evidence
+to support this claim
+
 # Research background {#lit}
 
-Table \@ref(tab:summarylit) summarise studies on forecasting in emergency and urgent care.
+Table @tbl-lit summarise studies on forecasting in emergency and urgent care.
 
-```{r summarylit}
-li <- readxl::read_xlsx("../data/literature.xlsx")
-knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("scale_down","hold_position")) %>% column_spec(5,"15em")
+```{r}
+#| label: tbl-lit
+#| tbl-cap: "summary of literature review"
+lit <- readxl::read_xlsx(here("data/literature.xlsx"))
+ lit %>% kable() %>%  kableExtra::kable_styling(latex_options = c("scale_down","hold_position")) %>% column_spec(5,"15em")
 ```

### 2022-10-18 rscript/fable/fit_model_fable.R
Updated readme

@@ -13,7 +13,8 @@ incidents <- readRDS(here::here("data/incidents_tsbl.rds")) |>
   #aggregate_key(category, incidents = sum(incidents))
   # End of temporary section of code
   # Uncomment next 3 lines when ready to scale up
-  aggregate_key(nature_of_incident * category * lhb_code, 
+  aggregate_key(
+    nature_of_incident * category * lhb_code, 
     incidents = sum(incidents)
   )
 

### 2022-10-23 manuscript/02-literature.qmd
reorganise manuscript and first draft of intro and literature sections

# Research background {#lit}
Emergency medical services (EMS), also referred to as ambulance, paramedic or pre-hospital emergency services, are a critical component in the delivery of urgent medical care to communities. An effective service delivery requires accurate resource planning that are generally replying on demand forecasts at operational, tactical and strategic levels.

There is a substantial literature on the application of forecasting in the Emergency Medical Services. Various areas have been the main focus. One of the major research topics has been forecasting call volume arrivals. @ibrahim2016modeling provide an extensive review of the forecasting models in this context. Another important focus of the literature in EMS is related to forecasting ambulance demand. Although sometimes the definition of the demand is not cleanly stated, however this is generally the demand where a physical resource has been deployed to respond to an incident, which is also referred to attended incidents. Another demand related variable is the verified incidents. These are all incidents that require an action: either send a physical vehicle, deal with via the Clinical Support Desk, get an external (private) provider to respond to it, or send it through to other channels such as police, firefighters or general practitioners. Another domain that has been extensively studied in the literature and it is also close to ours is forecasting Emergency Department attendance. We refer interested readers to @mingliterature2022, @gul2020exhaustive and @wargon2009systematic for extensive reviews of the literature in these areas. Although crucial to EMS performance, @aringhieri2017emergency state that demand forecasting has received limited research attention in the EMS context. In this section, we provide a brief review of studies on forecasting ambulance demand in EMS.

There are generally two main streams of research in this area: i) the first stream is focusing on the application of time series methods and regression approaches on forecasting aggregate ambulance demand as the most widely used approaches [@vile2012predicting; @sasaki2010using]; ii) another stream considers forecasting EMS demand in more finer temporal and geographical granularitties by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit]. Papers studying the second stream has not used classical regression and time-series approaches like the first one. Our study belongs to the first category, but we briefly review research in both streams.

@sasaki2010using develop a multivariable regression model to estimate future EMS demands. In addition to the historical demand, the population census for different age groups and counts of the number of companies employing more than five people are included in the regression. The census variables describe groups who may be more likely to need an ambulance. A stepwise ordinary least squares regression analysis with SPSS is used for estimating the parameter and generating forecast. The only performance measure reported in this study is $R^2$. The research design of this study is not rigorous and the study is not reproducible. @vile2012predicting explore using a Singular Spectrum Analysis (SSA) method to generate forecasts of the EMS demand at the national level for 7-day, 14-day, 21-day and 28-day forecast horizons using data provided by the Welsh Ambulance Service Trust (WAST). The performance of this approach is compared with Auto-Regressive Moving Average (ARIMA) and Holt-winter time series methods using Root Mean Squared Error (RMSE). They concluded that point forecasts generated by SSA are more accurate for longer-term, but ARIMA and Holt-winter performance is superior for shorter-term horizons. @vile2016time further develop a decision support system to integrate forecasts generated by SSA. However, the study does not compare and contrast of the performance of forecasting methods based on some utility measures such as cost, resource utilization or response time. The tool contains options that allow generating forecasts at various levels of granularity, however, it ignore the hierarchical and grouped relationships structure, preventing  aligned decision making and coordination.

@9659837 investigate forecasting EMS demand in a high spatio-temporal resolution of 1×1km spatial regions and 1-hr time intervals using total incidents in Oslo, Norway from January 1st, 2015 up to and including February 11th, 2019. They used multi-layer perceptron (MLP) and long short-term memory (LSTM) models to forecast the EMS demand, and compare them to simple aggregation methods and baselines. The point forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean Squared Error (MSE) and the forecast distribution is measures by
Categorical Cross-Entropy. They shows that while Neural Network models perform better in producing point forecast, a distribution baseline method based on spatial distribution of the incidents across all time steps provide more accurate forecast distribution. @zhou2016predictinglit propose three methods based on Gaussian mixture models, kernel density estimation, and kernel warping to predict 4 weeks into future for a 1-km2 spatial region over an hour. Two years of incidents attended from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011 and 2012 with  696,975 events) are used to build the model and examine the  performance on test data using mean negative log likelihood. They show that forecasts generated by the proposed methods are significantly more accurate than the current industry practice,  a simple averaging formula. @grekousis2019will investigate the combination of spatial analysis methods with data mining techniques based on  an improved Hungarian algorithm and MLP neural network to identify the most likely locations of future emergency events. The proposed approach is tested using data of 2851 events attended by the EMS in Athens, Greece over 24 weeks. They show that 23.24% of real emergency events lie within 50 m of the predicted ones and nearly 70% of the real emergency events lie no further away than 150 m, which is rather accurate given the granularity of the problem at the city level.

One of the major limitations of the current forecasting research in healthcare and EMS is the lack of a forecasting methodology that accounts for the inherent hierarchical and grouped structure in data to produce coherent forecasts, enabling aligned planning and decision making and possibly forecast accuracy improvement by exploring the demand relationships between different levels. While this methodology has been developed over the past 10 years [@panagiotelis2022probabilistic], it has never been explored in this area. Additionally, we found that most research only deals with point forecast estimation and also reproducibility is still a major concern in this area. Also, given the count nature of Considering these limitations from the literature, we develop a forecasting experiment to fill these gaps.

### 2022-10-23 manuscript/04-htsgts.qmd
reorganise manuscript and first draft of intro and literature sections

--- 2021-03-01 manuscript/manuscript.Rmd
@@
----
-title: Forecasting for emergency medcine
-author:
-  - name: Author1
-    email: emaail@example.com
-    affiliation: University1
-    footnote: 1
-  - name: Author2
-    email: email2@example.com
-    affiliation: University2
-    footnote: 2
-address:
-  - code: University1
-    address: adress1
-  - code: University2
-    address: adress2
-footnote:
-  - code: 1
-    text: "Corresponding Author"
+# Hierachcail and grouped time series forecasting techniques {#htc}
 
-abstract: |
-   The Objective of this paper
-keywords:  ""
-journal: "Which journal?"
-date: "`r Sys.Date()`"
-geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
-#linenumbers: true
-numbersections: true
-#output: rticles::elsevier_article
-header-includes:  
- \usepackage{adjustbox, float,lscape} #use the 'float' package
-output: 
-  bookdown::pdf_book:
-    base_format: rticles::elsevier_article
-    extra_dependencies: ["flafter"]
-    keep_tex: true
-bibliography: mybibfile.bib
-biblio-style: authoryear
-#csl: elsevier-without-titles.csl
-#biblio-style: abbrevnat
----
-
-```{r setup, include=FALSE}
-library(knitr)
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
-# Make sure you have the latest versions of rmarkdown and rticle
-library(tidyverse)
-library(lubridate)
-library(patchwork)
-library(tsibble)
-library(kableExtra)
-library(fable)
-library(ggthemes)
-library(readxl)
-```
-
-```{r, include=FALSE, cache=FALSE}
-
-```
-
-# Introduction
-
-An accurate demand forecasting is crucial in Emergency medicine to depict various courses of action that can result in massive savings in terms of patient lives.
-Inability to match the staff with the demand results in an overcrowding care system which is a serious problem causing challenging situations on patient flow.
-Also, it is related with increasing length of stay, low patient satisfaction, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times.without incurring last-minute expenses, such as overtime or supplemental staffing [Reference].
-
-An accurate forecasting of the daily demand enables managers to match staff to meet anticipated patients, reconfigure units and redeploy staff and vehicles.
-This will have many advantages for patients, staff and the quality of healthcare services provided.
-
-Daily forecasts are required to inform the short-term planning for the current and the upcoming shifts of the day.
-This involves the decision making related to the execution of the delivery process for various health care services such as Ambulatory ,Emergency.
-Daily forecasts are important at various levels:
-
-- **Nature of incidents**
-
-- **Category: Red, Amber, Green**
-
-- **Health board**
-
-- **Country level**
-
-The combination of demand forecast, incidents being attended, resource availability and delays at hospitals, provide information on the state of the unscheduled care system across the emergency medicine services.
-Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients.
-
-# Research background {#lit}
-
-Table \@ref(tab:summarylit) summarise studies on forecasting in emergency and urgent care.
-
-```{r summarylit}
-li <- readxl::read_xlsx("../data/literature.xlsx")
-knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("scale_down","hold_position")) %>% column_spec(5,"15em")
-```
-
-<!--
-Is there any study using grouped time series in urgent and emergency care?
--->
-
-<!--
-Using models for count time series in urgent and emergency care!
--->
-
-@gijo2016sarima generated a time series model to forecast the daily and hourly call volume at all centre handling emergency ambulance services.
-Since historical data showed seasonality, SARIMA models were investigated.
-Regarding the daily model, the authors generated a SARIMA model, which, however, resulted in the forecast error (MMSE) that significantly increased when the lead time exceeded 8 days.
-On the other hand, the SARIMA model proposed to forecast the log-calls an hourly basis.
-This model was found to fit well the model both for shorter and longer lead times.
-
-Forecasting at the daily level of granularity is generally use for roaster planing and deciding when to contact staff on call for instance.
-@luo2017hospital use a combination of seasonal ARIMA (SARIMA) and a single exponential smoothing (SES) model to forecast daily outpatient visits 1 week ahead.They indicate that using combinatorial model can be more effective than each model separately.
-@whitt2019forecasting examine several alternative models to forecast the total daily arrival for 1-7 days ahead, including a linear regression based on calendar and weather variables, seasonal autoregressive integrated moving average with exogenous regressors (SARIMAX) and the multilayer perceptron (MLP) model.
-Using a daily ED admission for 3 years, they show that SARIMAX provides the most accurate daily forecast by MSE measure.
-@marcilio2013forecasting compare the forecast accuracy of the Generalised Linear Model (GLM), Generalised Estimating Equations (GEE), and Seasonal Autoregressive Integrated Moving Average (SARIMA) methods using total daily patient visits to an ED using MAPE.
-They conclud that GLM and GEE models provide more accurate forecasts than SARIMA model.
-@rostami2020anticipating propose a model to forecast daily ED attendance with consideration of different types of holidays, weekday effects, auto-regressive effects, long-term trends and date effects.
-They provide probabilistic forecasts to quantify uncertainties in future ED attendance and they show that the proposed model outperforms three time series techniques including 1) Naive, 2) AutoRegressive, AR(p), 3)  exponential smoothing state space model (ETS) and a regression model without considering special events as alternatives.
-@zhou2018time SARIMA, NARNN and the hybrid SARIMA-NARNN to forecast the daily number of new admission inpatients.
-The root mean square error (RMSE), mean absolute error (MAE) and mean absolute percentage error (MAPE) were used to compare the forecasting performance among the three models.
-They show that NARNN model outperforms the others.
-@sun2009forecasting use SARIMA model to forecast daily patient volume for each patient acuity level.
-MAPE is used to choose the best-fit model.
-They fitted separate ARIMA models to the three categories of acuity and overall data.
-They conclude that the ARIMA model is effective for both short term (weekly) and long term (three months) forecast horizons.
-Moreover, they observe that the impact of weather is not significant.
-@zinouri2018modelling use SARIMA to develop a statistical prediction model that provides short-term forecasts of daily surgical demand.
-Our results suggest that the proposed SARIMA model can be useful for estimating surgical case volumes 2–4 weeks prior to the day of surgery, which can support robust and reliable staff schedules.
-@moustris2012seven use Artificial Neural Network (ANN) models to forecast the total weekly(7 days) number of Childhood Asthma Admission. Three different ANN models were developed
-and trained to forecast the admission of different age subgroups 
-and for the whole study population.
-@mccoy2018assessment compare SARIMA, Prophet and Snaive methods for prediction 365 days of hospital discharge using a daily hospital discharge volumes at 2 large, New England academic medical centers.
-They show that Prophet model outperforms SARIMA and Snaive.
-
-@khaldi2019forecasting investigate the combination of the Artificial Neural Networks (ANNs) with a signal decomposition technique named Ensemble Empirical Mode decomposition (EEMD), to make one step ahead weekly forecasting of patients arrivals to ED.
-using seven years of weekly demand.
-The results of the proposed model were compared against ANN without signal decomposition, ANN with Discrete wavelet Transform (DWT).
-The results show that the combined forecasts outperform the benchmarking models.
-
-@steins2019forecasting aimed to develop a forecasting model for predicting the number of ambulance services calls per hour and geographical areas, to support managers in decisions-making and to investigate which ones were the factors that affected the number of calls.
-Data collected consisted of a time and location of historical ambulance call data for three counties in Sweden and a list of explanatory factors of the area, such as socioeconomic and geographic. 
-In order to deal with large number of zeros in the data, authors developed zero-inflated Poisson (ZIP) and zero-inflated negative binomial (ZINB) regression models.
-These were then compared to the currently existing forecasting system, based on moving average with seasonality weights, using ME, MAE and RMSE.
-Firstly, the factors affecting the number of ambulance calls were found to be the following: population in different age groups, median income, length of road, number of nightlife spots (the number of restaurants), day of the week and hour of the day.
-Secondly, it was found that the older population (65-100) generated more ambulance calls and that ZIP model performs better than the current model.
-However, the improvement provided by the more advance model was not much greater than the one provided by the existing model.
-Authors suggested that it could be because the population, which was used as an independent variable in both models, was so dominant compared to the other factors.
 Moreover, both models either underestimated or overestimated the number of calls.
@@
 
-# Experimental design {#design}
+## Notation
 
-# Result and discussion {#result}
+## Base forecast
 
-# Conclusion {#conclusion}
+## Bottom-up
 
-# References {#references .unnumbered}
+## Optimal reconciliation

### 2022-10-23 manuscript/01-introduction.qmd

# Introduction
Inability to match the resources with the demand in Emergency Medical Services (EMS) results in an overcrowding care system. This is a serious problem causing challenging situations on patient flow, that ultimately leads to increasing patient's length of stay, higher waiting times, low patient satisfaction, high staff pressure, increasing health care costs and inaccuracy in electronic medical record [@ekstrom2015forecasting; @ROSTAMITABAR20221197]. Demand forecasting in EMS is a critical element that helps to depict various courses of action to avoid the mismatch, which can result in massive savings in terms of patient safety and lives. An accurate forecasting of the daily demand enables planners and decision makers to manage resources to meet anticipated patients, reconfigure units, redeploy staff and vehicles, where necessary. This will have many advantages for patients, staff and the quality of healthcare services provided.

Demand forecasts are typically required at multiple levels of cross-sectional granularities to inform various planning and decision-making processes across the EMS [@hulshof2012taxonomic]. There are some planning process at the national level (more strategic or long-term) such as workforce resource planning and budgeting; sub-national and regional level (tactical or medium-term) such as temporary capacity expansions, resource sharing and staff-shift scheduling; and healtboard, hospital or station level (operational or short-term) such as planning rosters for staff and ambulance deployment. Demand forecasts might be also required at different level for a specific group of interest such as nature of demand (e.g. type of incidents) or demand priority. Moreover, there exists an inherent hierarchical and grouped structure relationships that typically occurs in EMS data. Demand for emergency services at the country level can be disaggregated in a geographical hierarchy into sub-national, regions and healthboards or divided into groups such as nature of incidents or demand priority.

Despite a large number of studies dedicated to forecasting for Emergency Medical Services [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the main focus has been on producing independent forecasts at a single level, ignoring the inherent hierarchical or grouped structure relationships of the time series demand. However, forecasting the EMS demand across all hierarchical and grouped levels is paramount for an effective planning and decision making. The hierarchical forecasting approach can not only achieve more accurate forecasts than the independent (base) forecasts, but also creates consistent forecasts [@hyndman2011optimal]. Obtaining consistent forecasts at different levels is important for reliable decision-making to avoid making conflicting decisions. With hierarchical forecasting, plans at any level are based on identical forecasts and therefore can be aligned. Implementing and sustaining improvements in EMS requires alignments and coordination between different stakeholders, without which teams operate in isolation that may lead to conflicts, duplication work, rework, or work that runs counter to overall goals to improve hospital-wide patient flow. Hierarchical forecasting framework can also be used as a tools to improve coordination between teams across the care services at the national, sub-national, regional and hospital levels. To our knowledge, there is not only no research in the EMS forecasting to account for the hierarchical and grouped structure of the system but also in the entire field of forecasting for healthcare management (also referred to as non-clinical healthcare).

In this paper, we address this gap by investigating the application of forecast reconciliation methods in the Emergency Medical Services using the daily time series of verified incidents in the Great Britain from 2015 to 2020. The data has a hierarchical and grouped structures with hierarchies at the national, control (sub-national), healthboard (corresponding to regions) and groups by priority and nature of incidents. We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal combination approaches. We use Exponential smoothing, ARIMA and Poisson regression to generated the independent based forecast followed by applying bottom-up and optimal reconciliation approaches. Forecast performance are assessed by the Root Mean Squared Scaled Error (RMSSE) for point forecasting and the Continuous Ranked Probability Score (CRPS) for the probabilistic forecast. This paper complies with the principles of reproducibility  [@stodden2013best; @boylan2015reproducibility] and utilising fable package [@fable2022] in R software. Therefore, this study could equally be applied to any healthcare service (e.g. emergency department, primary or social care) subject to the time series demand having a hierarchical and/or grouped structure, which is generally the case in the healthcare sector.

The remainder of this article is structured as follows: In section 2, we provide a brief review of the
literature and discuss its limitation to position our work; in section 3, we present the experiment design describing the dataset, forecasting methods and forecast evaluation metrics. In Section 4, we describe the hierarchical and grouped time series forecasting methods for producing point and probabilistic forecasts. In section 5, we present and discuss our results; in section 7, we summarize
our findings and present ideas for future research.

### 2022-10-23 manuscript/03-experiment.qmd

# Experiment setup {#htc}

## Data

## Forecasting methods

## Performance evaluation

### 2022-10-23 manuscript/06-conclusion.qmd

--- 2021-03-01 manuscript/manuscript.Rmd
@@
----
-title: Forecasting for emergency medcine
-author:
-  - name: Author1
-    email: emaail@example.com
-    affiliation: University1
-    footnote: 1
-  - name: Author2
-    email: email2@example.com
-    affiliation: University2
-    footnote: 2
-address:
-  - code: University1
-    address: adress1
-  - code: University2
-    address: adress2
-footnote:
-  - code: 1
-    text: "Corresponding Author"
+# Conclusion {#conclusion}
 
-abstract: |
-   The Objective of this paper
-keywords:  ""
-journal: "Which journal?"
-date: "`r Sys.Date()`"
-geometry: "top=25mm, left=30mm, right=30mm, bottom=25mm,headsep=10mm, footskip=12mm"
-#linenumbers: true
-numbersections: true
-#output: rticles::elsevier_article
-header-includes:  
- \usepackage{adjustbox, float,lscape} #use the 'float' package
-output: 
-  bookdown::pdf_book:
-    base_format: rticles::elsevier_article
-    extra_dependencies: ["flafter"]
-    keep_tex: true
-bibliography: mybibfile.bib
-biblio-style: authoryear
-#csl: elsevier-without-titles.csl
-#biblio-style: abbrevnat
----
-
-```{r setup, include=FALSE}
-library(knitr)
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
-# Make sure you have the latest versions of rmarkdown and rticle
-library(tidyverse)
-library(lubridate)
-library(patchwork)
-library(tsibble)
-library(kableExtra)
-library(fable)
-library(ggthemes)
-library(readxl)
-```
-
-```{r, include=FALSE, cache=FALSE}
-
-```
-
-# Introduction
-
-An accurate demand forecasting is crucial in Emergency medicine to depict various courses of action that can result in massive savings in terms of patient lives.
-Inability to match the staff with the demand results in an overcrowding care system which is a serious problem causing challenging situations on patient flow.
-Also, it is related with increasing length of stay, low patient satisfaction, increasing health care costs, inaccuracy in electronic medical record, and reported waiting times.without incurring last-minute expenses, such as overtime or supplemental staffing [Reference].
-
-An accurate forecasting of the daily demand enables managers to match staff to meet anticipated patients, reconfigure units and redeploy staff and vehicles.
-This will have many advantages for patients, staff and the quality of healthcare services provided.
-
-Daily forecasts are required to inform the short-term planning for the current and the upcoming shifts of the day.
-This involves the decision making related to the execution of the delivery process for various health care services such as Ambulatory ,Emergency.
-Daily forecasts are important at various levels:
-
-- **Nature of incidents**
-
-- **Category: Red, Amber, Green**
-
-- **Health board**
-
-- **Country level**
-
-The combination of demand forecast, incidents being attended, resource availability and delays at hospitals, provide information on the state of the unscheduled care system across the emergency medicine services.
-Having this full picture enables the delivery managers to focus on the areas that require intervention to enable the most effective delivery of the service to the patients.
-
-# Research background {#lit}
-
-Table \@ref(tab:summarylit) summarise studies on forecasting in emergency and urgent care.
-
-```{r summarylit}
-li <- readxl::read_xlsx("../data/literature.xlsx")
-knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("scale_down","hold_position")) %>% column_spec(5,"15em")
-```
-
-<!--
-Is there any study using grouped time series in urgent and emergency care?
--->
-
-<!--
-Using models for count time series in urgent and emergency care!
--->
-
-@gijo2016sarima generated a time series model to forecast the daily and hourly call volume at all centre handling emergency ambulance services.
-Since historical data showed seasonality, SARIMA models were investigated.
-Regarding the daily model, the authors generated a SARIMA model, which, however, resulted in the forecast error (MMSE) that significantly increased when the lead time exceeded 8 days.
-On the other hand, the SARIMA model proposed to forecast the log-calls an hourly basis.
-This model was found to fit well the model both for shorter and longer lead times.
-
-Forecasting at the daily level of granularity is generally use for roaster planing and deciding when to contact staff on call for instance.
-@luo2017hospital use a combination of seasonal ARIMA (SARIMA) and a single exponential smoothing (SES) model to forecast daily outpatient visits 1 week ahead.They indicate that using combinatorial model can be more effective than each model separately.
-@whitt2019forecasting examine several alternative models to forecast the total daily arrival for 1-7 days ahead, including a linear regression based on calendar and weather variables, seasonal autoregressive integrated moving average with exogenous regressors (SARIMAX) and the multilayer perceptron (MLP) model.
-Using a daily ED admission for 3 years, they show that SARIMAX provides the most accurate daily forecast by MSE measure.
-@marcilio2013forecasting compare the forecast accuracy of the Generalised Linear Model (GLM), Generalised Estimating Equations (GEE), and Seasonal Autoregressive Integrated Moving Average (SARIMA) methods using total daily patient visits to an ED using MAPE.
-They conclud that GLM and GEE models provide more accurate forecasts than SARIMA model.
-@rostami2020anticipating propose a model to forecast daily ED attendance with consideration of different types of holidays, weekday effects, auto-regressive effects, long-term trends and date effects.
-They provide probabilistic forecasts to quantify uncertainties in future ED attendance and they show that the proposed model outperforms three time series techniques including 1) Naive, 2) AutoRegressive, AR(p), 3)  exponential smoothing state space model (ETS) and a regression model without considering special events as alternatives.
-@zhou2018time SARIMA, NARNN and the hybrid SARIMA-NARNN to forecast the daily number of new admission inpatients.
-The root mean square error (RMSE), mean absolute error (MAE) and mean absolute percentage error (MAPE) were used to compare the forecasting performance among the three models.
-They show that NARNN model outperforms the others.
-@sun2009forecasting use SARIMA model to forecast daily patient volume for each patient acuity level.
-MAPE is used to choose the best-fit model.
-They fitted separate ARIMA models to the three categories of acuity and overall data.
-They conclude that the ARIMA model is effective for both short term (weekly) and long term (three months) forecast horizons.
-Moreover, they observe that the impact of weather is not significant.
-@zinouri2018modelling use SARIMA to develop a statistical prediction model that provides short-term forecasts of daily surgical demand.
-Our results suggest that the proposed SARIMA model can be useful for estimating surgical case volumes 2–4 weeks prior to the day of surgery, which can support robust and reliable staff schedules.
-@moustris2012seven use Artificial Neural Network (ANN) models to forecast the total weekly(7 days) number of Childhood Asthma Admission. Three different ANN models were developed
-and trained to forecast the admission of different age subgroups 
-and for the whole study population.
-@mccoy2018assessment compare SARIMA, Prophet and Snaive methods for prediction 365 days of hospital discharge using a daily hospital discharge volumes at 2 large, New England academic medical centers.
-They show that Prophet model outperforms SARIMA and Snaive.
-
-@khaldi2019forecasting investigate the combination of the Artificial Neural Networks (ANNs) with a signal decomposition technique named Ensemble Empirical Mode decomposition (EEMD), to make one step ahead weekly forecasting of patients arrivals to ED.
-using seven years of weekly demand.
-The results of the proposed model were compared against ANN without signal decomposition, ANN with Discrete wavelet Transform (DWT).
-The results show that the combined forecasts outperform the benchmarking models.
-
-@steins2019forecasting aimed to develop a forecasting model for predicting the number of ambulance services calls per hour and geographical areas, to support managers in decisions-making and to investigate which ones were the factors that affected the number of calls.
-Data collected consisted of a time and location of historical ambulance call data for three counties in Sweden and a list of explanatory factors of the area, such as socioeconomic and geographic. 
-In order to deal with large number of zeros in the data, authors developed zero-inflated Poisson (ZIP) and zero-inflated negative binomial (ZINB) regression models.
-These were then compared to the currently existing forecasting system, based on moving average with seasonality weights, using ME, MAE and RMSE.
-Firstly, the factors affecting the number of ambulance calls were found to be the following: population in different age groups, median income, length of road, number of nightlife spots (the number of restaurants), day of the week and hour of the day.
-Secondly, it was found that the older population (65-100) generated more ambulance calls and that ZIP model performs better than the current model.
-However, the improvement provided by the more advance model was not much greater than the one provided by the existing model.
-Authors suggested that it could be because the population, which was used as an independent variable in both models, was so dominant compared to the other factors.
 Moreover, both models either underestimated or overestimated the number of calls.
@@
 Therefore, authors recommended that further research should add certain temporal variables able to capture the trend.
-
-# Experimental design {#design}
-
-# Result and discussion {#result}
-
-# Conclusion {#conclusion}
-
-# References {#references .unnumbered}

### 2022-10-23 manuscript/main.qmd

---
title: "Hierarchical and grouped time series forecasting: an application to Emergency Medical Services"
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Annals of Emergency Medicine
      formatting: preprint
      model: 3p
      cite-style: authoryear
  elsevier-html: 
    toc: true
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations: 
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  The primary goal of an emergency service is to minimize the ambulance response time. Accurate forecasts of ambulance demand are crucial inputs into the decisions of planners and policymakers on the resource (e.g. staff and fleet) planning and dynamic deployment. Informing those decisions require detailed forecasting of demand at multiple levels of hierarchy from the national and regional levels to the nature of incidents and their priorities. These forecasts are often generated independently by different teams. However, independent forecasts may not add up to the other levels resulting in mismatched forecasts that lead to conflicting decisions and a lack of coordination in the care service. To address this issue, we explore the inherently hierarchical and grouped structure of Emergency Medcine Services. We apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent, for the first time in the healthcare literature.
keywords: [Emergency services, healthcare, hierarchical time series forecasting, forecast reconciliation]
bibliography: mybibfile.bib
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
# Make sure you have the latest versions of rmarkdown and rticle
library(tidyverse)
library(lubridate)
library(patchwork)
library(tsibble)
library(kableExtra)
library(fable)
library(ggthemes)
library(readxl)
library(here)
```

```{r intro, child="01-introduction.qmd"}

```

```{r literature, child="02-literature.qmd"}

```

```{r experiment, child="03-experiment.qmd"}

```

```{r htc, child="04-htsgts.qmd"}

```

```{r results, child="05-results.qmd"}

```

```{r conclusion, child="06-conclusion.qmd"}

```

# References {#references .unnumbered}

### 2022-10-24 rscript/fable/fit_model_fable.R
Now using tolerant-inverse branch of fabletools

@@ -1,3 +1,6 @@
+# This script needs the "tolerant-inverse" branch of fabletools
+#remotes::install_github("tidyverts/fabletools", "tolerant-inverse")
+
 library(tidyverse)
 library(tsibble)
 library(fable)

### 2022-10-26 rscript/data/incidents.R
Updated generation of gts rds object

@@ -42,8 +42,22 @@ incidents_tsbl <- incidents |>
 # Save as rds
 write_rds(incidents_tsbl, here::here("data/incidents_tsbl.rds"))
 
-# Prepare data for gts format
+# Test data in gts format
+incidents_tsbl |> 
+  as_tibble() |> 
+  filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |> 
+  select(-nature_of_incident, -lhb_code) |> 
+  mutate(category = recode(category, "GREEN" = "GRE", "AMBER" = "AMB")) |> 
+  pivot_wider(names_from = category, values_from = incidents) |> 
+  filter(date <= "2015-12-31") |> 
+  select(-date) |> 
+  mutate(BLU = 0) |> 
+  as.ts(frequency=7) |> 
+  hts() |> 
+  write_rds(here::here("data/incidents_test_gts.rds"))
 
+
+# Prepare data for gts format
 incident_modify <- incidents_tsbl |>
   as_tibble() |>
   # Replace incident names with strings of equal length
@@ -157,4 +171,7 @@ incident_all <- incident_gts |>
   select(-delet, -date)
 
 # Write out final rds
-write_rds(incident_all, here::here("data/incidents_gts.rds"))
+incident_all |> 
+  ts(frequency = 7) |> 
+  gts(characters = list(c(1, 2), 3, 9)) |> 
+  write_rds(here::here("data/incidents_gts.rds"))

### 2022-10-26 rscript/data/incidents.R
Fixed a bug in the stored test data

@@ -42,7 +42,7 @@ incidents_tsbl <- incidents |>
 # Save as rds
 write_rds(incidents_tsbl, here::here("data/incidents_tsbl.rds"))
 
-# Test data in gts format
+# Small data set in gts format
 incidents_tsbl |> 
   as_tibble() |> 
   filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |> 
@@ -52,12 +52,11 @@ incidents_tsbl |>
   filter(date <= "2015-12-31") |> 
   select(-date) |> 
   mutate(BLU = 0) |> 
-  as.ts(frequency=7) |> 
+  ts(frequency=7) |> 
   hts() |> 
   write_rds(here::here("data/incidents_test_gts.rds"))
 
-
-# Prepare data for gts format
+# Prepare full data in gts format
 incident_modify <- incidents_tsbl |>
   as_tibble() |>
   # Replace incident names with strings of equal length

### 2022-10-26 rscript/hts/forecast_hts_with_transformation.R
Working on hts code

@@ -1,12 +1,25 @@
 library(hts)
-library(readr)
-source("htsplus.R")
+source(here::here("rscript/hts/htsplus.R"))
 
-# Test on incident data
-incident_all <- read_rds("data/incident_all.rds")
-incident_ts <- ts(incident_all, frequency = 7)
+# Read hierarchical/grouped time series
+incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
+#incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
 
-# create hierarchical/grouped time series
-gts_incident <- hts::gts(incident_ts, characters = list(c(1, 2), 3, 9))
-fcst_bu <- htsplus(gts_incident, h = 10, model_function = ets, lambda = 0.5, method = "bu")
-fcst_wls <- htsplus(gts_incident, h = 10, model_function = ets, lambda = 0.5, method = "wls")
+train <- window(incident_gts, end = c(12,1))
+test <- window(incident_gts, start=c(12,2))
+
+# Simulated future sample paths
+fcst_ets_bu <- htsplus(train, h = nrow(test$bts), model_function = ets, lambda = 0.5, method = "bu")
+fcst_ets_wls <- htsplus(train, h = nrow(test$bts), model_function = ets, lambda = 0.5, method = "wls")
+fcst_ets_mint <- htsplus(train, h = nrow(test$bts), model_function = ets, lambda = 0.5, method = "mint")
+
+# Means
+mean_ets_bu <- t(apply(fcst_ets_bu, c(1,2), mean))
+mean_ets_wls <- t(apply(fcst_ets_wls, c(1,2), mean))
+mean_ets_mint <- t(apply(fcst_ets_mint, c(1,2), mean))
+
+rmsse(mean_ets_bu, test, train)
+rmsse(mean_ets_wls, test, train)
+rmsse(mean_ets_mint, test, train)
+
+

### 2022-10-26 rscript/hts/htsplus.R

@@ -10,77 +10,87 @@
 htsplus <- function(object, h,
                     model_function = auto.arima,
                     lambda = NULL, level = 95,
-                    method = c("wls", "bu"), nsim = 2000) {
+                    method = c("wls", "bu", "mint"), nsim = 2000) {
   method <- match.arg(method)
   # Generate all time series at all levels
   ally <- aggts(object)
+  nbottom <- NCOL(object$bts)
   nseries <- NCOL(ally)
+  ntime <- NROW(ally)
   # Do Box-Cox transformation if required
   if (!is.null(lambda)) {
     ally <- BoxCox(ally, lambda)
   }
-  # Set up matrices to store results
-  fmean <- fvar <- flower <- fupper <- matrix(NA, nrow = h, ncol = nseries)
+  # Set up matrices to store residuals
+  innov_res <- response_res <- matrix(0, nrow=nrow(ally), ncol=nseries)
   models <- list()
   # Loop over all time series and compute forecast means and variances
-  for (i in seq(NCOL(ally))) {
-    models[[i]] <- model_function(ally[, i])
-    fc <- forecast(models[[i]], h = h, level = level)
-    fmean[, i] <- fc$mean
-    fvar[, i] <- ((fc$upper - fc$lower) / (2 * qnorm(0.5 + level / 200)))^2
-  }
-  # Reverse Box-Cox transformation if required
-  # using Taylor series approximation (delta method) to get mean and variance
-  # on original scale
-  if (!is.null(lambda)) {
-    fmean <- InvBoxCox(fmean, lambda)
-    tmp <- lambda * fmean + 1
-    fp2 <- fmean^(2 - 2 * lambda)
-    fmean <- fmean * (1 + 0.5 * fvar * (1 - lambda) / (fmean)^(2 * lambda))
-    fvar <- fvar * fp2
+  for (i in seq(nseries)) {
+    # Add some tiny positive noise to avoid zero series
+    models[[i]] <- model_function(ally[, i] + abs(rnorm(n=ntime, sd=1e-2)))
+    # Grab the residuals
+    innov_res[,i] <- residuals(models[[i]], type="innovation")
+    response_res[,i] <- residuals(models[[i]], type="response")
   }
   # S matrix
   S <- smatrix(object)
[…]

### 2022-10-26 rscript/data/incidents.R
Computing RMSSE and CRPS, and added tscount

@@ -174,3 +174,12 @@ incident_all |>
   ts(frequency = 7) |> 
   gts(characters = list(c(1, 2), 3, 9)) |> 
   write_rds(here::here("data/incidents_gts.rds"))
+
+# Covariates for gts format
+
+readRDS(here::here("data/holiday_dummy.rds")) |> 
+  as_tibble() |> 
+  select(-date) |> 
+  ts(frequency = 7) |> 
+  write_rds(here::here("data/holidays_ts.rds"))
+  

### 2022-10-26 rscript/hts/forecast_hts_with_transformation.R

@@ -1,25 +1,38 @@
 library(hts)
 source(here::here("rscript/hts/htsplus.R"))
+source(here::here("rscript/hts/tscount.R"))
 
 # Read hierarchical/grouped time series
 incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
 #incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
+holidays <- read_rds(here::here("data/holidays_ts.rds"))
 
 train <- window(incident_gts, end = c(12,1))
 test <- window(incident_gts, start=c(12,2))
+alltest <- aggts(test)
 
-# Simulated future sample paths
-fcst_ets_bu <- htsplus(train, h = nrow(test$bts), model_function = ets, lambda = 0.5, method = "bu")
-fcst_ets_wls <- htsplus(train, h = nrow(test$bts), model_function = ets, lambda = 0.5, method = "wls")
-fcst_ets_mint <- htsplus(train, h = nrow(test$bts), model_function = ets, lambda = 0.5, method = "mint")
+# Reconciliation methods
+methods <- c("bu", "wls", "mint")
 
-# Means
-mean_ets_bu <- t(apply(fcst_ets_bu, c(1,2), mean))
-mean_ets_wls <- t(apply(fcst_ets_wls, c(1,2), mean))
-mean_ets_mint <- t(apply(fcst_ets_mint, c(1,2), mean))
+# Set up storage 
+fcst_ets <- as.list(methods)
+names(fcst_ets) <- methods
+fcst_tscount <- fcst_ets
+rmsse_ets <- matrix(0, nrow = length(methods), ncol = ncol(alltest))
+dimnames(rmsse_ets) <- list(methods, colnames(alltest))
+rmsse_tscount <- crps_tscount <- crps_ets <- rmsse_ets
 
-rmsse(mean_ets_bu, test, train)
-rmsse(mean_ets_wls, test, train)
-rmsse(mean_ets_mint, test, train)
-
-
+# ETS simulations
+for(i in seq_along(methods)) {
+  fcst_ets[[i]] <- htsplus(train, h = nrow(test$bts), 
+                           model_function = ets, method = methods[i])
+  rmsse_ets[i, ] <- rmsse(fcst_ets[[i]], test, train)
+  crps_ets[i, ] <- crps(fcst_ets[[i]], test)
+}
+# TSCOUNT simulations
+for(i in seq_along(methods)) {
[…]

### 2022-10-26 rscript/hts/htsplus.R

@@ -26,11 +26,17 @@ htsplus <- function(object, h,
   models <- list()
   # Loop over all time series and compute forecast means and variances
   for (i in seq(nseries)) {
-    # Add some tiny positive noise to avoid zero series
-    models[[i]] <- model_function(ally[, i] + abs(rnorm(n=ntime, sd=1e-2)))
+    if(forecast:::is.constant(ally[,i])) {
+      # Series is constant. So just fit a mean zero model with small variance
+      models[[i]] <- Arima(ally[,i] + rnorm(ntime, sd=1e-3), order=c(0,0,0), include.mean = FALSE)
+    } else {
+      models[[i]] <- model_function(ally[, i])
+    }
     # Grab the residuals
-    innov_res[,i] <- residuals(models[[i]], type="innovation")
     response_res[,i] <- residuals(models[[i]], type="response")
+    if(!identical(model_function, tscount)) {
+      innov_res[,i] <- residuals(models[[i]], type="innovation")
+    } 
   }
   # S matrix
   S <- smatrix(object)
@@ -84,13 +90,41 @@ htsplus <- function(object, h,
   return(sim)
 }
 
-# Look at RMSSE over several levels for MinT
-rmsse <- function(fmean, test_gts, train_gts) {
+# Compute RMSSE given forecast mean and training/test gts objects
+rmsse <- function(sim, test_gts, train_gts) {
+  fmean <- t(apply(sim, c(1,2), mean))
   alltest <- aggts(test_gts)
   alltrain <- aggts(train_gts)
   scale_factor <- colMeans(diff(alltrain, 7)^2)
   rmsse <- colMeans(sweep((fmean - alltest)^2, 2L, scale_factor, '/'))
+  # Set RMSSE of zero series to zero
   rmsse[scale_factor < .Machine$double.eps] <- 0
   return(rmsse)
 }
+
+# Compute CRPS given simulated values x and actual y
+crps_sample <- function(x, y) {
+  # Set CRPS of zero series to zero
+  if(var(x) < 1e-5)
+    return(0)
+  x <- sort(x)
+  m <- length(x)
+  crps <- (2/m) * mean((x - y) * (m * (y < x) - seq_len(m) +  0.5))
 }
+
+# Compute CRPS given simulated sample paths and test gts object
+crps <- function(sim, test_gts) {
[…]

### 2022-10-26 rscript/hts/tscount.R

@@ -1,0 +1,36 @@
+library(tscount)
+
+# Specific tsglm model used for this data
+tscount <- function(y) {
+  fourier_year <- forecast::fourier(ts(y, frequency=365.25), K=3)
+  season_week <- forecast::seasonaldummy(y)
+  trend <- seq_along(y)
+  X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
+  tsglm(y, model = list(past_obs = 1:3), xreg=X, link="log")
+}
+
+simulate.tsglm <- function(object, innov,...) {
+  n <- length(object$response)
+  h <- NROW(innov)
+  fourier_year <- forecast::fourier(ts(seq(n), frequency=365.25), K=3, h=h)
+  season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
+  trend <- seq(n+h)[-(1:n)]
+  X <- cbind(trend, season_week, fourier_year, holidays[n+seq(h),])
+  output <- tsglm.sim(h, 
+        param = list(
+          intercept = object$coefficient[1],
+          past_obs = object$coefficients[2:4],
+          past_mean = FALSE,
+          xreg = object$coefficients[-(1:4)]
+        ),
+        model = list(
+          past_obs = 1:3, 
+          past_mean = FALSE, 
+          external=FALSE
+        ),
+        xreg = X,
+        link = "log"
+  )$ts
+  ts(output, frequency = frequency(object$response),
+     start = tsp(object$response)[2] + 1/frequency(object$response))
+}

### 2022-10-26 rscript/hts/forecast_hts.R
Added parallel processing over series and removed Box-Cox

@@ -1,20 +1,25 @@
 library(hts)
+library(readr)
+library(furrr)
 source(here::here("rscript/hts/htsplus.R"))
 source(here::here("rscript/hts/tscount.R"))
 
+# Parallelization
+plan(multisession, workers = 2)
+
 # Read hierarchical/grouped time series
 incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
-#incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
+# incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
 holidays <- read_rds(here::here("data/holidays_ts.rds"))
 
-train <- window(incident_gts, end = c(12,1))
-test <- window(incident_gts, start=c(12,2))
+train <- window(incident_gts, end = c(12, 1))
+test <- window(incident_gts, start = c(12, 2))
 alltest <- aggts(test)
 
 # Reconciliation methods
 methods <- c("bu", "wls", "mint")
 
-# Set up storage 
+# Set up storage
 fcst_ets <- as.list(methods)
 names(fcst_ets) <- methods
 fcst_tscount <- fcst_ets
@@ -23,16 +28,22 @@ dimnames(rmsse_ets) <- list(methods, colnames(alltest))
 rmsse_tscount <- crps_tscount <- crps_ets <- rmsse_ets
 
 # ETS simulations
-for(i in seq_along(methods)) {
-  fcst_ets[[i]] <- htsplus(train, h = nrow(test$bts), 
-                           model_function = ets, method = methods[i])
+for (i in seq_along(methods)) {
+  cat(paste("ETS: Method", methods[[i]], "\n"))
+  fcst_ets[[i]] <- htsplus(train,
+    h = nrow(test$bts),
+    model_function = ets, method = methods[i]
+  )
   rmsse_ets[i, ] <- rmsse(fcst_ets[[i]], test, train)
   crps_ets[i, ] <- crps(fcst_ets[[i]], test)
 }
 # TSCOUNT simulations
-for(i in seq_along(methods)) {
-  fcst_tscount[[i]] <- htsplus(train, h = nrow(test$bts), 
-                           model_function = tscount, method = methods[i])
+for (i in seq_along(methods)) {
+  cat(paste("TSCOUNT: Method", methods[[i]], "\n"))
+  fcst_tscount[[i]] <- htsplus(train,
+    h = nrow(test$bts),
[…]

### 2022-10-26 rscript/hts/htsplus.R

@@ -1,42 +1,42 @@
+# htsplus = Use forecast reconciliation to generate simulated future sample paths
+
 # object = gts object
 # h = forecast horizon
-# model_function = function used to model each time series. e.g., ets or auto.arima
-# lambda = Box-Cox transformation parameter
-# level = prediction interval coverage as a percentage
-# method = reconcilation method. either "wls" or "bu"
+# model_function = function used to model each time series. e.g., ets or auto.arima or tscount
+# method = reconcilation method. either "wls" or "bu" or "mint"
 # nsim = number of simulations in computing prediction intervals using nonparametric method
-# htsplus = hts with BoxCox transformation
 
-htsplus <- function(object, h,
+htsplus <- function(object,
+                    h,
                     model_function = auto.arima,
-                    lambda = NULL, level = 95,
-                    method = c("wls", "bu", "mint"), nsim = 2000) {
+                    method = c("wls", "bu", "mint"),
+                    nsim = 50) {
   method <- match.arg(method)
   # Generate all time series at all levels
   ally <- aggts(object)
-  nbottom <- NCOL(object$bts)
+  # Loop over all time series and store models
+  fit_model <- function(y, model_function) {
+    ntime <- length(y)
+    if (forecast:::is.constant(y)) {
+      # Series is constant. So just fit a mean zero model with small variance
+      return(Arima(y + rnorm(ntime, sd = 1e-3), order = c(0, 0, 0), include.mean = FALSE))
+    } else {
+      return(model_function(y))
+    }
+  }
+  models <- furrr::future_map(as.list(ally), fit_model,
+    model_function = model_function,
+    .options = furrr_options(seed = NULL)
+  )
+  # Grab the residuals from each model
   nseries <- NCOL(ally)
   ntime <- NROW(ally)
-  # Do Box-Cox transformation if required
-  if (!is.null(lambda)) {
-    ally <- BoxCox(ally, lambda)
-  }
-  # Set up matrices to store residuals
-  innov_res <- response_res <- matrix(0, nrow=nrow(ally), ncol=nseries)
[…]

### 2022-10-26 rscript/hts/tscount.R

@@ -6,6 +6,9 @@ tscount <- function(y) {
   season_week <- forecast::seasonaldummy(y)
   trend <- seq_along(y)
   X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
+  # Remove constant covariates
+  constant <- apply(X, 2, forecast:::is.constant)
+  X <- X[, !constant]
   tsglm(y, model = list(past_obs = 1:3), xreg=X, link="log")
 }
 
@@ -16,6 +19,8 @@ simulate.tsglm <- function(object, innov,...) {
   season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
   trend <- seq(n+h)[-(1:n)]
   X <- cbind(trend, season_week, fourier_year, holidays[n+seq(h),])
+  # Remove missing columns
+  X <- X[, colnames(X) %in% names(coefficients(object))]
   output <- tsglm.sim(h, 
         param = list(
           intercept = object$coefficient[1],

### 2022-10-26 rscript/data/incidents.R
Now running on small hierarchy but with all dates

@@ -6,6 +6,7 @@
 library(tidyverse)
 library(lubridate)
 library(tsibble)
+library(hts)
 
 # Read data
 incidents_original <- here::here("data/Nature_of_Incidents_Attended.xlsx") |>
@@ -49,7 +50,7 @@ incidents_tsbl |>
   select(-nature_of_incident, -lhb_code) |> 
   mutate(category = recode(category, "GREEN" = "GRE", "AMBER" = "AMB")) |> 
   pivot_wider(names_from = category, values_from = incidents) |> 
-  filter(date <= "2015-12-31") |> 
+  #filter(date <= "2015-12-31") |> 
   select(-date) |> 
   mutate(BLU = 0) |> 
   ts(frequency=7) |> 

### 2022-10-26 rscript/hts/forecast_hts.R

@@ -12,8 +12,8 @@ incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
 # incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
 holidays <- read_rds(here::here("data/holidays_ts.rds"))
 
-train <- window(incident_gts, end = c(12, 1))
-test <- window(incident_gts, start = c(12, 2))
+train <- window(incident_gts, end = c(190, 7))
+test <- window(incident_gts, start = c(191, 1))
 alltest <- aggts(test)
 
 # Reconciliation methods

### 2022-10-26 rscript/hts/forecast_hts.R
Finally trying all series with 10 week test set

@@ -8,8 +8,8 @@ source(here::here("rscript/hts/tscount.R"))
 plan(multisession, workers = 2)
 
 # Read hierarchical/grouped time series
-incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
-# incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
+#incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
+incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
 holidays <- read_rds(here::here("data/holidays_ts.rds"))
 
 train <- window(incident_gts, end = c(190, 7))

### 2022-10-26 rscript/hts/forecast_hts.R
Speed up through using same models for different reconciliation methods

@@ -16,34 +16,16 @@ train <- window(incident_gts, end = c(190, 7))
 test <- window(incident_gts, start = c(191, 1))
 alltest <- aggts(test)
 
-# Reconciliation methods
-methods <- c("bu", "wls", "mint")
-
-# Set up storage
-fcst_ets <- as.list(methods)
-names(fcst_ets) <- methods
-fcst_tscount <- fcst_ets
-rmsse_ets <- matrix(0, nrow = length(methods), ncol = ncol(alltest))
-dimnames(rmsse_ets) <- list(methods, colnames(alltest))
-rmsse_tscount <- crps_tscount <- crps_ets <- rmsse_ets
-
 # ETS simulations
-for (i in seq_along(methods)) {
-  cat(paste("ETS: Method", methods[[i]], "\n"))
-  fcst_ets[[i]] <- htsplus(train,
-    h = nrow(test$bts),
-    model_function = ets, method = methods[i]
+fcst_ets <- htsplus(train, h = nrow(test$bts),
+    model_function = ets, methods = c("bu", "wls", "mint")
   )
-  rmsse_ets[i, ] <- rmsse(fcst_ets[[i]], test, train)
-  crps_ets[i, ] <- crps(fcst_ets[[i]], test)
-}
 # TSCOUNT simulations
-for (i in seq_along(methods)) {
-  cat(paste("TSCOUNT: Method", methods[[i]], "\n"))
-  fcst_tscount[[i]] <- htsplus(train,
-    h = nrow(test$bts),
-    model_function = tscount, method = methods[i]
+fcst_tscount <- htsplus(train, h = nrow(test$bts),
+    model_function = tscount, method = c("bu", "wls", "mint")
   )
-  rmsse_tscount[i, ] <- rmsse(fcst_tscount[[i]], test, train)
-  crps_tscount[i, ] <- crps(fcst_tscount[[i]], test)
-}
+  
+rmsse_ets <- rmsse(fcst_ets, test, train)
+crps_ets <- crps(fcst_ets, test)
+rmsse_tscount <- rmsse(fcst_tscount, test, train)
+crps_tscount <- crps(fcst_tscount, test)

### 2022-10-26 rscript/hts/htsplus.R

@@ -3,15 +3,14 @@
 # object = gts object
 # h = forecast horizon
 # model_function = function used to model each time series. e.g., ets or auto.arima or tscount
-# method = reconcilation method. either "wls" or "bu" or "mint"
+# methods = vector of reconcilation methods. All will be used
 # nsim = number of simulations in computing prediction intervals using nonparametric method
 
 htsplus <- function(object,
                     h,
                     model_function = auto.arima,
-                    method = c("wls", "bu", "mint"),
+                    methods,
                     nsim = 50) {
-  method <- match.arg(method)
   # Generate all time series at all levels
   ally <- aggts(object)
   # Loop over all time series and store models
@@ -40,37 +39,42 @@ htsplus <- function(object,
   }
   # S matrix
   S <- smatrix(object)
-  # Find W matrix
-  if (method == "ols") {
-    W <- diag(nseries)
-  } else {
-    W <- hts:::lowerD(response_res)
-    if (method == "mint") {
-      W <- hts:::shrink.estim(response_res, W)[[1]]
-    }
-  }
-  # Find G matrix
-  if (method == "bu") {
-    nbottom <- NCOL(object$bts)
-    G <- cbind(matrix(0, nrow = nbottom, ncol = nseries - nbottom), diag(nrow = nbottom))
-  } else {
-    Winv <- solve(W)
-    G <- solve(t(S) %*% Winv %*% S) %*% t(S) %*% Winv
-  }
-  # Computing mapping matrix M
-  M <- S %*% G
+  # Find W matrices
+  W1 <- diag(nseries)
+  W2 <- hts:::lowerD(response_res)
+  W3 <- hts:::shrink.estim(response_res, W2)[[1]]
 
+  # Find G matrix
+  nbottom <- NCOL(object$bts)
+  G1 <- cbind(matrix(0, nrow = nbottom, ncol = nseries - nbottom), diag(nrow = nbottom))
+  Winv <- solve(W2)
+  G2 <- solve(t(S) %*% Winv %*% S) %*% t(S) %*% Winv
+  Winv <- solve(W3)
+  G3 <- solve(t(S) %*% Winv %*% S) %*% t(S) %*% Winv
+  
+  # Computing mapping matrix M
[…]

### 2022-10-26 rscript/data/incidents.R
Updated readme

@@ -7,6 +7,7 @@ library(tidyverse)
 library(lubridate)
 library(tsibble)
 library(hts)
+source(here::here("rscript/data/holidays.R"))
 
 # Read data
 incidents_original <- here::here("data/Nature_of_Incidents_Attended.xlsx") |>

### 2022-10-26 rscript/hts/forecast_hts.R
Save simulated sample paths

@@ -5,7 +5,7 @@ source(here::here("rscript/hts/htsplus.R"))
 source(here::here("rscript/hts/tscount.R"))
 
 # Parallelization
-plan(multisession, workers = 2)
+plan(multisession, workers = 3)
 
 # Read hierarchical/grouped time series
 #incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
@@ -17,14 +17,27 @@ test <- window(incident_gts, start = c(191, 1))
 alltest <- aggts(test)
 
 # ETS simulations
-fcst_ets <- htsplus(train, h = nrow(test$bts),
+if (!fs::file_exists(here::here("rscript/hts/fcst_ets.rds"))) {
+  fcst_ets <- htsplus(train,
+    h = nrow(test$bts), nsim = 2000,
     model_function = ets, methods = c("bu", "wls", "mint")
   )
+  write_rds(fcst_ets, here::here("rscript/hts/fcst_ets.rds"))
+} else {
+  fcst_ets <- read_rds(here::here("rscript/hts/fcst_ets.rds"))
+}
 # TSCOUNT simulations
-fcst_tscount <- htsplus(train, h = nrow(test$bts),
+if (!fs::file_exists(here::here("rscript/hts/fcst_tscount.rds"))) {
+  fcst_tscount <- htsplus(train,
+    h = nrow(test$bts), nsim = 2000,
     model_function = tscount, method = c("bu", "wls", "mint")
   )
-  
+  write_rds(fcst_tscount, here::here("rscript/hts/fcst_tscount.rds"))
+} else {
+  fcst_tscount <- read_rds(here::here("rscript/hts/fcst_tscount.rds"))
+}
+
+# Summary statistics of forecast accuracy
 rmsse_ets <- rmsse(fcst_ets, test, train)
 crps_ets <- crps(fcst_ets, test)
 rmsse_tscount <- rmsse(fcst_tscount, test, train)

### 2022-10-26 rscript/hts/htsplus.R

@@ -10,7 +10,7 @@ htsplus <- function(object,
                     h,
                     model_function = auto.arima,
                     methods,
-                    nsim = 50) {
+                    nsim = 2000) {
   # Generate all time series at all levels
   ally <- aggts(object)
   # Loop over all time series and store models
@@ -51,12 +51,12 @@ htsplus <- function(object,
   G2 <- solve(t(S) %*% Winv %*% S) %*% t(S) %*% Winv
   Winv <- solve(W3)
   G3 <- solve(t(S) %*% Winv %*% S) %*% t(S) %*% Winv
-  
+
   # Computing mapping matrix M
   M1 <- S %*% G1
   M2 <- S %*% G2
   M3 <- S %*% G3
-  
+
   # Compute reconciled future sample paths using cross-sectional bootstrap
   nmethods <- length(methods)
   sim <- array(0, c(nseries, h, nsim, nmethods))
@@ -69,8 +69,9 @@ htsplus <- function(object,
     for (i in seq(nseries)) {
       sim[i, , j, 1] <- simulate(models[[i]], innov = bootres[, i])
       # Use same sample paths for other reconciliation methods
-      for(k in seq(nmethods)[-1])
+      for (k in seq(nmethods)[-1]) {
         sim[i, , j, k] <- sim[i, , j, 1]
+      }
     }
     sim[, , j, 1] <- M1 %*% sim[, , j, 1]
     sim[, , j, 2] <- M2 %*% sim[, , j, 2]
@@ -86,9 +87,9 @@ rmsse <- function(sim, test_gts, train_gts) {
   alltest <- aggts(test_gts)
   alltrain <- aggts(train_gts)
   scale_factor <- colMeans(diff(alltrain, 7)^2)
-  rmsse <- fmean[,1,]
-  for(i in NCOL(rmsse)) {
-    rmsse[,i] <- colMeans(sweep((t(fmean[,,i]) - alltest)^2, 2L, scale_factor, "/"))
+  rmsse <- fmean[, 1, ]
+  for (i in NCOL(rmsse)) {
+    rmsse[, i] <- colMeans(sweep((t(fmean[, , i]) - alltest)^2, 2L, scale_factor, "/"))
   }
   # Set RMSSE of zero series to zero
   rmsse[scale_factor < .Machine$double.eps] <- 0
[…]

### 2022-10-26 manuscript/05-results.qmd
reorganise manuscript and first draft of intro and literature sections
adding first draft of section 3 &4

--- 2022-10-23 manuscript/04-htsgts.qmd
@@
-# Hierachcail and grouped time series forecasting techniques {#htc}
+# Results and discussion {#sec-result}
 
-Moreover, both models either underestimated or overestimated the number of calls.
-Authors suggested that the inability to capture a positive trend resulted in underestimation, and the opposite was due to a negative trend.
-Therefore, authors recommended that further research should add certain temporal variables able to capture the trend.
+## Point forecast
 
-## Notation
-
-## Base forecast
-
-## Bottom-up
-
-## Optimal reconciliation
+## Probabilistic forecast

### 2022-10-26 manuscript/main.qmd
adding first draft of section 3 &4

@@
 title: "Hierarchical and grouped time series forecasting: an application to Emergency Medical Services"
+execute:
+  echo: false
+  message: false
+  warning: false
+  cache: true
 format:
@@
 
-```{r setup, include=FALSE}
+```{r}
+#| label: setup
+#| include: false
 library(knitr)
-knitr::opts_chunk$set(echo = FALSE, cache=TRUE, message = FALSE, warning = FALSE)
-# Make sure you have the latest versions of rmarkdown and rticle
 library(tidyverse)
@@
 library(here)
+library(tsibble)
+library(fable)
 ```

### 2022-10-26 rscript/hts/htsplus.R
Using generalized inverse for covariance matrix

@@ -47,9 +47,9 @@ htsplus <- function(object,
   # Find G matrix
   nbottom <- NCOL(object$bts)
   G1 <- cbind(matrix(0, nrow = nbottom, ncol = nseries - nbottom), diag(nrow = nbottom))
-  Winv <- solve(W2)
+  Winv <- MASS::ginv(W2)
   G2 <- solve(t(S) %*% Winv %*% S) %*% t(S) %*% Winv
-  Winv <- solve(W3)
+  Winv <- MASS::ginv(W3)
   G3 <- solve(t(S) %*% Winv %*% S) %*% t(S) %*% Winv
 
   # Computing mapping matrix M

### 2022-10-26 manuscript/03-experiment.qmd

@@
-# Experiment setup {#htc}
+# Experiment setup {#sec-experiment}
 
-## Data
+We are interested in generating forecasts to inform the planning horizon of $ph=7 to 42$ days, interested by planners in the ambulance services trust.
+The forecast horizon in this study is $fh = 2 \times ph$ days ahead (2*42 days planning horizon).
+This is because the planning is generally freezed for ph days and considering a forecast horizon of ph days might not be helpful for planning.
+While forecasts are generated for $2 \times ph$ days ahead, performance evaluation is only assessed based on the last $ph$ days and not the $2 \times ph$ days.
+The forecasts are produced for the holdout of 365 days using time series cross-validation [@hyndman2021forecasting].
+In the following section, we discuss the dataset, describe the forecasting methods used to generate base forecasts and present the point and probabilistic accuracy measures.
+
+## Data {#sec-data}
+
+The dataset used in this study is from a major ambulance service trust in the United Kingdom.
+It contains information relating to the daily number of attended incidents from 10 October 2015 to 31 July 2019 by nature of incidents, priority, the health board managing the service and the control area (or region).
+Figure @fig-hierarchy depicts both the hierarchical and grouped structure of the data.
+Figure @fig-hierarchy-1 illustrates the nested hierarchical structure based on control area and health board and Figure @fig-hierarchy-2 shows the grouped structure by priority and the nature of incident.
+
+```{r}
+#| label: fig-hierarchy
+#| out.width: "60%"
+#| fig-cap: "The hierarchical and grouped structure of attended inciddents (ambulance demand)"
+#| fig-subcap: 
+#|   - "Hierarchical structure: Attended inciddents in all country is disaggregated into 3 control areas /regions and then into 7 different healrhboard at the bottom"
+#|   - "Grouped structure:  Incidents could be grouped into priority (i.e.
+Red, Amber & Green) and the nature of attenedd incident (i.e.
+there are 35 different nature of incidents including chest pain , breahtign problems, heart attack, stroke, and son on).
+The symbol * refers to the crossed attributes between hierarchical and grouped levels."
+#| layout: [[50],[-2], [38]]
+
+library(ggraph)
+library(igraph)
+library(tidyverse)
+
+data <- data.frame(
+  level1="Total",
+  level2=c("Central & West", "Central & West","Central & West","North", "South & East","South & East","South & East"),
+  level3=c("HD", "SB", "PO", "BC","CV", "CT", "AB")
+)
+# transform it to a edge list!
+edges_level1_2 = data %>% select(level1, level2) %>% 
+  unique %>% rename(from=level1, to=level2)
+edges_level2_3 = data %>% select(level2, level3) %>% 
+  unique %>% rename(from=level2, to=level3)
+edge_list=rbind(edges_level1_2, edges_level2_3)
+
+mygraph <- graph_from_data_frame( edge_list )
+ggraph(mygraph, layout = 'dendrogram', circular = FALSE) + 
+  geom_edge_diagonal() +
+  geom_node_point(colour="#dddddd",size=10) +
+  geom_node_text(
+    aes(label=c("All country", "Central & West", "North","South & East", "HD", "SB", "PO", "BC","CV", "CT", "AB") )
+  )+
+  theme_void()
+
+knitr::include_graphics(here::here("img/group.png"))
+```
+
+Table @tbl-hierarchy also displays the structure of data with the total number of series at each level.
+At the top level, we have the total attended incidents for the country.
+We can split these total attended incidents by control area, by health board, by priority or by nature of incident.
+There are 3 control areas breakdown by 7 local health boards.
+Attended incident data are categorized into 3 priority classes of red, amber and green.
+There are also 35 different nature of incidents such as chest pain, stroke, breathing problem, etc.
+In total, across all levels of disaggregation, there are 1530 time series.
+
+```{r}
+#| label: tbl-hierarchy
+#| tbl-cap: "Number of time series in each level for the hierarchical & grouped structure of attended incidents"
+agg_level <- tibble::tribble(
+  ~Level, ~`Number of series`,
+  "All country", 1,
+  "Control", 3,
+  "Health board", 7,
+  "Priority", 3,
+  "Priority * Control", 9,
+  "Priority * Health board", 21,
+  "Nature of incident", 35,
+  "Nature of incident * Control", 105,
+  "Nature of incident * Health board", 245,
+  "Priority * Nature of incident", 104,
+  "Control * Priority * Nature of incident", 306,
+  "Control * Health board * Priority * Nature of incident (Bottom level)", 691,
+  "Total", 1530
+)
+knitr::kable(agg_level, booktabs = T, position = "left") %>% kable_classic(full_width = F)
+```
+
+Given the total number of time series, direct visual analysis is infeasible.
+Therefore, we first compute features of all 1530 time series and display the strength of trend and weekly seasonality strength in Figure @fig-feature.
+Each point represents one time series with strength of trend in  x-axis and strength of seasonality in y-axis.
+It is clear that there are some series showing strong trends and/or seasonality, corresponding to series at the higher levels of the hierarchy.
+The majority show low trend and seasonality with .
+These are time series belonging to the bottom of the structure, series related to the nature of incidents for each given control, health board and priority.
+
+```{r}
+#| label: fig-feature
+#| cache: true
+#| out.width: "70%"
+#| fig.align: center
+#| fig-cap: "Time series features of attended incidents across all levels (1530 series)"
+library(feasts)
+incident_gthf <- read_rds(here::here("data/incident_gt.rds"))
+incident_gthf %>% 
+  features(incidents,feat_stl) %>%
+  ggplot(aes(x = trend_strength, y = seasonal_strength_week)) +
+  geom_point(pch=1)+
+  ggthemes::theme_few()
+```
+
+In addition to displaying the trend and seasonality features, we also highlight few time series at various levels of the aggregation.
+Figure @fig-dataviz reveals different information such as trend, seasonality and noise.
+For example, some series depict seasonality and trend, whereas some other series report low volume of attended incidents and entropy, making them more volatile and difficult to forecast.
+At the level on nature of incidents combined with categories of other levels, there are many series that contain zeros with low counts.
+As such, the data set represents a diverse set of daily time series patterns.
+
+```{r}
+#| label: fig-dataviz
+#| cache: true
+#| dependson: "fig-feature"
+#| out.width: "70%"
+#| fig-cap: "Time series of attended incidents at various levels"
+#| fig-subcap: 
+#|   - "all country"
+#|   - "Control area"
+#|   - "Health board and control area"
+#|   - "Healthboard and priority"
+#|   - "Nature of incident and health board"
+#| layout: [[44,-2, 44], [44,-2, 44], [100]]
+p_total <- incident_gthf %>%
+    filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
+    autoplot()+
+    labs(x="Date", y="Incidents")+
+    ggthemes::theme_few()
+
+p_total_control <- incident_gthf %>%
+  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
+  as_tibble() %>% select(-nature,-category) %>% 
+  group_by(date,region) %>% summarise(incidents=sum(incidents), .groups="drop") %>% 
+  ggplot(aes(x=date, y= incidents, colour = factor(region))) +
+  geom_line()+
+  labs(x="Date", y="Incidents", colour = "Control")+
+  ggthemes::scale_color_colorblind()+
+  ggthemes::theme_few()+
+  theme(legend.position = "bottom")
+
+p_control <- incident_gthf %>%
+  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
+  as_tibble() %>% select(-nature,-category) %>% 
+  ggplot(aes(x=date, y= incidents, colour = factor(lhb))) +
+  geom_line()+
+  facet_wrap(vars(factor(region)), scales = "free_y")+
+  labs(x="Date", y="Incidents", colour = "Health board")+
+  ggthemes::scale_color_colorblind()+
+  ggthemes::theme_few()+
+  theme(legend.position = "bottom")
+
+p_priority <- incident_gthf %>%
+  filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & is_aggregated(nature)) %>%
+  as_tibble() %>% select(-nature,-region) %>% 
+  ggplot(aes(x=date, y= incidents, colour = factor(lhb))) +
+  geom_line()+
+  facet_wrap(vars(factor(category)), scales = "free_y")+
+  labs(x="Date", y="Incidents", colour = "Health board")+
+  ggthemes::scale_color_colorblind()+
+  ggthemes::theme_few()+
+  theme(legend.position = "bottom")
+
+selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
+
+p_nature_healthboard <- incident_gthf %>%
+  filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & !is_aggregated(nature)) %>% 
+as_tibble() %>% mutate(nature=as.character(nature)) %>% 
+  filter(nature %in% selected_nature) %>% group_by(date,nature,lhb) %>% summarise(incidents=sum(incidents), .groups="drop") %>% 
+  ggplot(aes(x=date, y= incidents, colour = factor(lhb))) +
+  geom_line()+
+  facet_wrap(vars(factor(nature)), scales = "free_y")+
+  labs(x="Date", y="Incidents", colour = "Health board")+
+  ggthemes::scale_color_colorblind()+
+  ggthemes::theme_few()+
+  theme(legend.position = "none")
+
+p_total
+p_total_control
+p_control
+p_priority
+p_nature_healthboard
+```
 
@@
 
+Given the presence of various significant patterns in the past attended incidents, we consider three different forecasting models to generate the base forecasts.
+Once the base forecasts are produced, hierarchical and grouped time series methods are used to reconcile them across the all levels.
+We briefly discussed forecasting models in the following sections, and the hierarchical forecasting methods are discussed in @sec-htc.
+
+**Exponential Smoothing State Space model (ETS):** ETS models [@hyndman2021forecasting] can combine trend, seasonality and error components in a time series through various forms such as additive, multiplicative or mixed.
+The trend component can be none ("N"), Additive ("A"), damped ("Ad") or multiplicative ("M").
+The seasonality can be none ("N"), Additive ("A"), or multiplicative ("M").
+The error term can also be additive ("A") or multiplicative ("M").
+To forecast the attended incidents at each level, we use the `ETS()` function in the `fable` package [@fable2022] in R.
+To identify the best model for a given time series, the ETS function uses the corrected Akaike’s Information Criterion (AICc).
+
+**Auto-Regressive Integrated Moving Average (ARIMA):** ARIMA [@hyndman2021forecasting] is another effective family of forecasting models that capture autocorrelation in time series.
+They can model various range of time series considering the stationarity of series, the Auto-Regressive (AR) and moving average (MA) orders.
+In this study, an automatic ARIMA algorithm implemented using `ARIMA()` function of fable package in R is used to forecast the attended incidents.
+To select the best model for a given time series, the algorithm uses unit root test, Maximum Likelihood Estimation (MLE) and the AICc.
+
+**Poisson Regression: ** Despite the popularity and the relevant of automatic `ETS` and `ARIMA` in this study, but they produce non-integer attended incidents and forecasts might also be negative.
+However, the number of attended incidents is an integer and non-negative.
+When using `ETS` and `ARIMA`, a time series transformation approach such as `sqrt()` could be used to generate strictly positive forecasts, however forecasts are still not integer.
+Another alternative is to use  forecasting models that produce integer, non-negative forecasts.
+One of the models that is frequently used in practice is the Poisson regression [see, for example, @mccarthy2008challenge], which can be summarized as
+$$
+  \begin{aligned}
+    & {y}_t \sim \mathrm{Poisson} \left( \lambda_t \right) \\
+    & \log \lambda_t = \mathbf{x}_t' \boldsymbol{\beta}
+  \end{aligned} \quad
+$$ {#eq-poisson-regression}
+
+The logarithm in @eq-poisson-regression is needed to ensure that the parameter of Poisson distribution is always positive.
+This model can be estimated via maximization of the likelihood function based on Poisson mass function.
+For taking into account serial dependence we include a regression on the previous observation.
+We try to capture the short range serial dependence by a first order autoregressive term and the weekly seasonality by a 7th order autoregressive term.
+We use the `tscount()` function in the fable package in R, which is a wrapper function written based on `tscount` package.
+We fit a model to each time series using the function tsglm, with the identity link function, defined by the argument link.
+
 ## Performance evaluation
+
+Forecasting performance is evaluated using both point and probabilistic error measures.
+
+The point forecast accuracy is measured via Root Mean Squared Scaled Error (RMSSE):
+
+$$
+\text{RMSSE} = \sqrt{\text{mean}(q_{j}^2)},
+$$ {#eq-RMSSE}
+
+where,
+
+$$
+q^2_{j} = \frac{\displaystyle e^2_{j}}
+    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
+$$
+
+$e_{j}$ is the point forecast error $j$ and $m = 1$ for non-seasonal series and $m = 7$ for daily seasonal series, $y_t$ is the observation for period $t$ and $T$ is the sample size (observations used for training the forecasting model).
+Smaller RMSSE values suggest more accurate forecasts.
+Note that the measure is scale-independent, thus allowing us to average the results across series.
+
+To measure the forecast distribution performance, we calculate  the Continuous Rank Probability Score [@gneiting2014probabilistic].
+It rewards sharpness and penalizes miscalibration, so it measures overall performance.
+
+$$
+CRPS=\frac{1}{h} \sum_{j=1}^{h} \int_{-\infty}^{\infty} \left(F^f_j(x) - F^0_j(x)\right)^2dx
+$$ {#eq-CRPS}
+
+where $F^f_j(x)$ is the forecasted Cumulative Density Function (CDF) of period $j$ and 𝐹$F^0_j(x)$ is the true CDF of period $j$.

### 2022-10-26 manuscript/04-htsgts.qmd

@@
-# Hierachcail and grouped time series forecasting techniques {#htc}
+# Hierachcail and grouped time series forecasting techniques {#sec-htc}
 
-Moreover, both models either underestimated or overestimated the number of calls.
-Authors suggested that the inability to capture a positive trend resulted in underestimation, and the opposite was due to a negative trend.
-Therefore, authors recommended that further research should add certain temporal variables able to capture the trend.
+There are many applications in the healthcare and in particular in EMS where a collection of time series is available.
+These series are generally hierarchically organized based on multiple levels such as area/region, health board and/or are aggregated at different levels in groups based on nature of demand, priority of demand or some other attributes.
+While series could be strictly hierarchical or only grouped bases on some attributes, in many situation a more complex structures arise when attributes of interest are both nested and crossed, having hierarchical and grouped structure.
+This is also the case as discussed in @sec-data
 
-## Notation
+## Independent (base forecast)
 
-## Base forecast
+A common practice in healthcare (and EMS) to predict hierarchical and grouped series relies on producing independent forecast, also refereed to as base forecast, typically by different teams as the need for such forecasts arise.
+This is also known as base forecast.
+We observe $n$ time series at time $t$, across the entire hierarchical and grouped structure, written as $y_t$.
+The base forecasts of $y_{T+h}$ given data $y_1,\dots,y_T$ are denoted by $\hat{y}_h$ for h steps-ahead for all $n$ series ($n=1530$ in this study).
+Forecasts generated in this way are nor coherent.
 
-## Bottom-up
+## Reconciliation methos
 
-## Optimal reconciliation
+Traditionally, alternative approaches to produce coherent forecasts for hierarchical and grouped time series involves using bottom-up and top-down methods by generating forecasts at a single level and then aggregate or disaggregate.
+Top-down requires having a unique structure to disaggregated forecasts generated at top level by proportions.
+However, given that we have multiple grouped attributes combined with the hierarchical structure, there is no unique way to disaggregate top forecast.
+Hence the top-down can not be used in this case, so either we can do some kind of reconciliation or must define our own top-down method for each hierarchy.
+The recommended approach is to use reconciliation.
+In the following sections, we first discuss some notations and then present bottom-up and forecast reconciliation approach used in this study to generate coherent forecasts.
+
+### Notations
+
+Let $b_t$ be a vector of $n_b$ "bottom-level" time series at time $t$, and let $a_t$ be a corresponding vector of $n_a = n-n_b$ aggregated time series, where
+
+$$
+ a_t = Ab_t
+$$
+
+and $A$ is the $n_a\times n_b$ "aggregation" matrix specifying how the bottom-level series $b_t$ are to be aggregated to form $a_t$. The full vector of time series is given by
+$$
+ y_t = \begin{bmatrix}a_t \\b_t\end{bmatrix}.
+$$
+This leads to the $n\times n_b$ "summing" or "structural" matrix given by
+$$
+  S = \begin{bmatrix}A \\ I_{n_b}\end{bmatrix}
+$$
+such that $y_t = Sb_t$.
+
+### Bottom-up (BU) and linear reconciliation methods
+
+Bottom-Up is a simple approach to generate coherent forecasts.
+It first involves creating the base forecasts for the bottom level series.
+These forecasts are then aggregated to the upper levels which results in generating coherent forecasts.
+BU approach can capture the dynamics of the series at the bottom level, but they may be noisy and difficult to forecast as well.
+BU approach is limited on using only forecasts at the bottom level and does not utilize all the information available across the hierarchical and grouped structure.
+Forecast reconciliation approaches fill this gap and combine and reconcile all the base forecasts in order to produce coherent forecasts.
+
+Given the summing matrix and base forecasts, bottom-up and linear reconciliation methods can be written as $\tilde{y}_h = SG\hat{y}_h$ for different matrices $G$.
+
+Optimal reconciled forecasts are obtained with $G=(S'W^{-1}S)^{-1}W^{-1}$, or $\tilde{y}_h = M\hat{y}_h,$ where the $n\times n$ "mapping" matrix is given by $M = S(S'W^{-1}S)^{-1}W^{-1},$ where $\hat{y}_h$ are the $h$-step forecasts of $y_{T+h}$ given data to time $T$, and $W$ is an $n \times n$ positive definite matrix.
+Different choices for $W$ lead to different solutions such as Ordinary Least Square (OLS) , Weighted Least Square (WLS) and Minimum Trace (MinT) [@WicEtAl2019].
+We use the implementation of these methods in the fable package in R in the experiment.

### 2022-10-27 rscript/hts/htsplus.R
Refactored simulation code to store everything on disk

@@ -1,44 +1,104 @@
-# htsplus = Use forecast reconciliation to generate simulated future sample paths
+# fit_models: Function to fit models to incidents data and save results as a list
 
 # object = gts object
-# h = forecast horizon
-# model_function = function used to model each time series. e.g., ets or auto.arima or tscount
-# methods = vector of reconcilation methods. All will be used
-# nsim = number of simulations in computing prediction intervals using nonparametric method
+# model_function = function used to model each time series.
+# e.g., ets or auto.arima or tscount
 
-htsplus <- function(object,
-                    h,
-                    model_function = auto.arima,
-                    methods,
-                    nsim = 2000) {
+fit_models <- function(object, model_function = "ets") {
+  # Number of observations in training set
+  ntime <- NROW(object$bts)
+
+  # Form file name for saving results
+  filename <- here::here(paste0("rscript/hts/", model_function, "_", ntime, ".rds"))
+  # Check if this has already been run
+  if (fs::file_exists(filename)) {
+    # Just use the previous results
+    return(invisible(read_rds(filename)))
+  }
+
   # Generate all time series at all levels
   ally <- aggts(object)
   # Loop over all time series and store models
   fit_model <- function(y, model_function) {
     ntime <- length(y)
+    modelfn <- get(model_function)
     if (forecast:::is.constant(y)) {
       # Series is constant. So just fit a mean zero model with small variance
       return(Arima(y + rnorm(ntime, sd = 1e-3), order = c(0, 0, 0), include.mean = FALSE))
     } else {
-      return(model_function(y))
+      return(modelfn(y))
     }
   }
   models <- furrr::future_map(as.list(ally), fit_model,
     model_function = model_function,
     .options = furrr_options(seed = NULL)
   )
-  # Grab the residuals from each model
-  nseries <- NCOL(ally)
-  ntime <- NROW(ally)
-  innov_res <- response_res <- matrix(0, nrow = ntime, ncol = nseries)
-  for (i in seq(nseries)) {
-    response_res[, i] <- residuals(models[[i]], type = "response")
-    if (!identical(model_function, tscount)) {
-      innov_res[, i] <- residuals(models[[i]], type = "innovation")
-    }
[…]

### 2022-10-27 rscript/hts/tscount.R
Spline trend rather than linear trend

@@ -4,7 +4,7 @@ library(tscount)
 tscount <- function(y) {
   fourier_year <- forecast::fourier(ts(y, frequency=365.25), K=3)
   season_week <- forecast::seasonaldummy(y)
-  trend <- seq_along(y)
+  trend <- splines::ns(seq(length(y)), df=10)
   X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
   # Remove constant covariates
   constant <- apply(X, 2, forecast:::is.constant)
@@ -17,11 +17,11 @@ simulate.tsglm <- function(object, innov,...) {
   h <- NROW(innov)
   fourier_year <- forecast::fourier(ts(seq(n), frequency=365.25), K=3, h=h)
   season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
-  trend <- seq(n+h)[-(1:n)]
+  trend <- predict(splines::ns(seq(n), df=10), newx=seq(n+h)[-(seq(n))])
   X <- cbind(trend, season_week, fourier_year, holidays[n+seq(h),])
   # Remove missing columns
   X <- X[, colnames(X) %in% names(coefficients(object))]
-  output <- tsglm.sim(h, 
+  output <- tsglm.sim(h,
         param = list(
           intercept = object$coefficient[1],
           past_obs = object$coefficients[2:4],
@@ -29,8 +29,8 @@ simulate.tsglm <- function(object, innov,...) {
           xreg = object$coefficients[-(1:4)]
         ),
         model = list(
-          past_obs = 1:3, 
-          past_mean = FALSE, 
+          past_obs = 1:3,
+          past_mean = FALSE,
           external=FALSE
         ),
         xreg = X,

### 2022-10-27 manuscript/03-experiment.qmd
adding the missing img (img/group) and the r code to create incident_gt in the incidents.R script in data

@@
 
-We are interested in generating forecasts to inform the planning horizon of $ph=7 to 42$ days, interested by planners in the ambulance services trust.
+We are interested in generating forecasts to inform the planning horizon of $ph= 42$ days, interested by planners in the ambulance services trust.
 The forecast horizon in this study is $fh = 2 \times ph$ days ahead (2*42 days planning horizon).
-This is because the planning is generally freezed for ph days and considering a forecast horizon of ph days might not be helpful for planning.
+This is because the planning is generally freezed for $ph$ days and considering a forecast horizon of $ph$ days might not be helpful for planning.
 While forecasts are generated for $2 \times ph$ days ahead, performance evaluation is only assessed based on the last $ph$ days and not the $2 \times ph$ days.
@@
 library(feasts)
+
 incident_gthf <- read_rds(here::here("data/incident_gt.rds"))

### 2022-10-27 rscript/hts/htsplus.R
Simulations constrained to be non-negative

@@ -159,6 +159,8 @@ future_sample_paths <- function(object, model_function = "ets", h = 84, nsim = 1
       sim[i, , j] <- simulate(models[[i]], innov = bootres[, i])
     }
   }
+  # Set negative to zero
+  sim[sim < 0] <- 0
   # Save results to file and then return them
   write_rds(sim, filename)
   return(invisible(sim))
@@ -192,6 +194,8 @@ reconcile_sample_paths <- function(object, model_function = "ets", methods = c("
     for(j in seq(nsim)) {
       newsim[,,j] <- M[[k]] %*% sim[,,j]
     }
+    # Set negative to zero
+    newsim[newsim < 0] <- 0
     write_rds(newsim, paste0(filestem,methods[k],".rds"))
   }
   return(invisible(newsim))
@@ -213,6 +217,8 @@ rmsse <- function(train_gts, test_gts, model_function, method) {
   }
   
   # Find the simulation files that exist
+  stem <- filename <- here::here(paste0("rscript/hts/", model_function, "_*_sim_", ".rds"))
+  
   sim <- reconcile_sample_paths(train_gts, model_function, methods = method)
 
   # Set up rmsse object

### 2022-10-27 manuscript/main.qmd

@@
 title: "Hierarchical and grouped time series forecasting: an application to Emergency Medical Services"
-execute:
-  echo: false
-  message: false
-  warning: false
-  cache: true
-format:
-  elsevier-pdf:
-    keep-tex: true
-    journal:
-      name: Annals of Emergency Medicine
-      formatting: preprint
-      model: 3p
-      cite-style: authoryear
-  elsevier-html: 
-    toc: true
 author:
@@
 We apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent, for the first time in the healthcare literature.
-keywords: [Emergency services, healthcare, hierarchical time series forecasting, forecast reconciliation]
-bibliography: mybibfile.bib
+keywords: 
+  - Healthcare
+  - Emergency services
+  - forecast reconciliation
+  - hierarchical time series forecasting
+bibliography: bibliography.bib
+execute:
+  echo: false
+  message: false
+  warning: false
+  cache: true
+format:
+  elsevier-pdf:
+    keep-tex: true
+    journal:
+      name: Annals of Emergency Medicine
+      formatting: preprint
+      model: 3p
+      cite-style: authoryear
+  elsevier-html: 
+    toc: true
 ---

### 2022-10-27 rscript/data/incidents.R
incidents.R now running in a fresh session

@@ -6,6 +6,7 @@
 library(tidyverse)
 library(lubridate)
 library(tsibble)
+library(fabletools)
 library(hts)
 source(here::here("rscript/data/holidays.R"))
 
@@ -131,6 +132,8 @@ incident_ready <- incident_modify |>
 incident_gt <- incident_ready |>
   as_tsibble(index = date, key = c(region, lhb, category, nature)) |>
   aggregate_key((region / lhb) * category * nature, incident = sum(incident))
+write_rds(incident_gt, here::here("data/incidents_gt.rds"))
+
 
 a1 <- incident_ready |>
   filter(lhb == "CV", category == "RED", nature == "FALLSTHRO") |>

### 2022-10-27 rscript/hts/forecast_hts.R
Added expanding window to forecast_hts.R

@@ -12,13 +12,18 @@ source(here::here("rscript/hts/tscount.R"))
 incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
 holidays <- read_rds(here::here("data/holidays_ts.rds"))
 
-# Keep last 12 weeks for evaluation
-train <- window(incident_gts, end = c(188, 7))
-test <- window(incident_gts, start = c(188, 8))
-
-# Create reconciled sample paths
-reconcile_sample_paths(train, model_function = "ets")
-reconcile_sample_paths(train, model_function = "tscount")
+nobs <- NROW(incident_gts$bts)
+horizons <- 42 * seq(10)
+for(i in seq(horizons)) {
+  # Set up training and test sets
+  n <- nobs - horizons[i]
+  weeks <- trunc(n/7)
+  train <- window(incident_gts, end = c(weeks, n - weeks*7 + 7))
+  test <- window(incident_gts, start = c(weeks, n - weeks*7 + 8))
+  # Create reconciled sample paths
+  reconcile_sample_paths(train, model_function = "ets")
+  #reconcile_sample_paths(train, model_function = "tscount")
+}
 
 # Summary statistics of forecast accuracy
 #rmsse_ets <- rmsse(train, test, model_function = "ets", method = "wls")

### 2022-10-27 rscript/data/holidays.R
Storing rds files on Google Drive

@@ -36,7 +36,7 @@ holiday_dummy <- holiday %>%
   as_tsibble(index = date)
 
 # Save as RDS for reading in by other scripts
-write_rds(holiday_dummy, here::here("data/holiday_dummy.rds"))
+write_rds(holiday_dummy, paste0(storage_folder, "holiday_dummy.rds"))
 
 
 # Other holiday tibbles created below, no longer used (?)

### 2022-10-27 rscript/data/incidents.R

@@ -43,7 +43,7 @@ incidents_tsbl <- incidents |>
   as_tsibble(index = date, key = c(lhb_code, category, nature_of_incident)) |>
   fill_gaps(incidents = 0, .full = TRUE)
 # Save as rds
-write_rds(incidents_tsbl, here::here("data/incidents_tsbl.rds"))
+write_rds(incidents_tsbl, paste0(storage_folder, "incidents_tsbl.rds"))
 
 # Small data set in gts format
 incidents_tsbl |> 
@@ -57,7 +57,7 @@ incidents_tsbl |>
   mutate(BLU = 0) |> 
   ts(frequency=7) |> 
   hts() |> 
-  write_rds(here::here("data/incidents_test_gts.rds"))
+  write_rds(paste0(storage_folder, "incidents_test_gts.rds"))
 
 # Prepare full data in gts format
 incident_modify <- incidents_tsbl |>
@@ -132,7 +132,7 @@ incident_ready <- incident_modify |>
 incident_gt <- incident_ready |>
   as_tsibble(index = date, key = c(region, lhb, category, nature)) |>
   aggregate_key((region / lhb) * category * nature, incident = sum(incident))
-write_rds(incident_gt, here::here("data/incidents_gt.rds"))
+write_rds(incident_gt, paste0(storage_folder, "incidents_gt.rds"))
 
 
 a1 <- incident_ready |>
@@ -183,13 +183,13 @@ incident_all <- incident_gts |>
 incident_all |> 
   ts(frequency = 7) |> 
   gts(characters = list(c(1, 2), 3, 9)) |> 
-  write_rds(here::here("data/incidents_gts.rds"))
+  write_rds(paste0(storage_folder, "incidents_gts.rds"))
 
 # Covariates for gts format
 
-readRDS(here::here("data/holiday_dummy.rds")) |> 
+readRDS(paste0(storage_folder, "holiday_dummy.rds")) |> 
   as_tibble() |> 
   select(-date) |> 
   ts(frequency = 7) |> 
-  write_rds(here::here("data/holidays_ts.rds"))
+  write_rds(paste0(storage_folder, "holidays_ts.rds"))
   

### 2022-10-27 rscript/hts/forecast_hts.R
Running ets code over 10 test sets of size 84 with 42 day overlap

@@ -8,22 +8,22 @@ source(here::here("rscript/hts/tscount.R"))
 #plan(multisession, workers = 3)
 
 # Read hierarchical/grouped time series
-#incident_gts <- read_rds(here::here("data/incidents_test_gts.rds"))
-incident_gts <- read_rds(here::here("data/incidents_gts.rds"))
-holidays <- read_rds(here::here("data/holidays_ts.rds"))
+#incident_gts <- read_rds(paste0(storage_folder, "incidents_test_gts.rds"))
+incident_gts <- read_rds(paste0(storage_folder, "incidents_gts.rds"))
+holidays <- read_rds(paste0(storage_folder, "holidays_ts.rds"))
 
-nobs <- NROW(incident_gts$bts)
-horizons <- 42 * seq(10)
-for(i in seq(horizons)) {
-  # Set up training and test sets
-  n <- nobs - horizons[i]
-  weeks <- trunc(n/7)
-  train <- window(incident_gts, end = c(weeks, n - weeks*7 + 7))
-  test <- window(incident_gts, start = c(weeks, n - weeks*7 + 8))
+# Test sets of size 84, 
+origins <- 42 * seq(10) + 42
+for(i in seq(origins)) {
+  # Set up training set
+  train <- incident_gts
+  train$bts <- subset(train$bts, end = nrow(incident_gts$bts) - origins[i])
   # Create reconciled sample paths
   reconcile_sample_paths(train, model_function = "ets")
   #reconcile_sample_paths(train, model_function = "tscount")
 }
+
+
 
 # Summary statistics of forecast accuracy
 #rmsse_ets <- rmsse(train, test, model_function = "ets", method = "wls")

### 2022-10-27 rscript/hts/htsplus.R

@@ -9,7 +9,7 @@ fit_models <- function(object, model_function = "ets") {
   ntime <- NROW(object$bts)
 
   # Form file name for saving results
-  filename <- here::here(paste0("rscript/hts/", model_function, "_", ntime, ".rds"))
+  filename <- paste0(storage_folder, model_function, "_", ntime, ".rds")
   # Check if this has already been run
   if (fs::file_exists(filename)) {
     # Just use the previous results
@@ -52,7 +52,7 @@ calculate_residuals <- function(object, model_function, type=c("innovation","res
   ntime <- NROW(object$bts)
   
   # Form file name for saving results
-  filename <- here::here(paste0("rscript/hts/", model_function, "_", ntime, "_res_",type,".rds"))
+  filename <- paste0(storage_folder, model_function, "_", ntime, "_res_",type,".rds")
   # Check if this has already been run
   if (fs::file_exists(filename)) {
     return(read_rds(filename))
@@ -85,7 +85,7 @@ make_mapping_matrices <- function(object, model_function) {
   ntime <- NROW(object$bts)
   
   # Form file name for saving results
-  filename <- here::here(paste0("rscript/hts/", model_function, "_", ntime, "_mapping.rds"))
+  filename <- paste0(storage_folder, model_function, "_", ntime, "_mapping.rds")
   # Check if this has already been run
   if (fs::file_exists(filename)) {
     # Just use the previous results
@@ -136,7 +136,7 @@ future_sample_paths <- function(object, model_function = "ets", h = 84, nsim = 1
   ntime <- NROW(object$bts)
 
   # Form file name for saving results
-  filename <- here::here(paste0("rscript/hts/", model_function, "_", ntime, "_sim.rds"))
+  filename <- paste0(storage_folder, model_function, "_", ntime, "_sim_base.rds")
   # Check if this has already been run
   if (fs::file_exists(filename)) {
     # Just use the previous results
@@ -176,11 +176,10 @@ reconcile_sample_paths <- function(object, model_function = "ets", methods = c("
   ntime <- NROW(object$bts)
   
   # Has this already been run?
-  filename <- here::here(paste0("rscript/hts/", model_function, "_", ntime, 
-                                  "_sim_",methods[1],".rds"))
+  filename <- paste0(storage_folder, model_function, "_", ntime, "_sim_",methods[1],".rds")
   # Check if this has already been run
[…]

### 2022-10-28 rscript/hts/forecast_hts.R
Added glm model

@@ -18,12 +18,11 @@ for(i in seq(origins)) {
   # Set up training set
   train <- incident_gts
   train$bts <- subset(train$bts, end = nrow(incident_gts$bts) - origins[i])
-  # Create reconciled sample paths
+  # Create reconciled sample paths for different models
   reconcile_sample_paths(train, model_function = "ets")
   #reconcile_sample_paths(train, model_function = "tscount")
+  reconcile_sample_paths(train, model_function = "iglm")
 }
-
-
 
 # Summary statistics of forecast accuracy
 #rmsse_ets <- rmsse(train, test, model_function = "ets", method = "wls")

### 2022-10-28 rscript/hts/glm.R

@@ -1,0 +1,39 @@
+# Specific tsglm model used for this data
+iglm <- function(y) {
+  n <- length(y)
+  fourier_year <- forecast::fourier(ts(y, frequency=365.25), K=3)
+  season_week <- forecast::seasonaldummy(y)
+  trend <- splines::ns(seq(n), df=round(n/100))
+  X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
+  # Remove constant covariates
+  constant <- apply(X, 2, forecast:::is.constant)
+  X <- X[, !constant]
+  object <- suppressWarnings(try(glm(y ~ X, family=poisson()), silent=TRUE))
+  if(inherits(object, "try-error"))
+    object <- glm(y ~ 1, family = poisson())
+  # Slim down return object and add ts attributes
+  object <- list(coefficients=object$coefficients, 
+                 residuals=residuals(object, type="response"),
+                 y=y)
+  return(structure(object, class='iglm'))
+}
+
+simulate.iglm <- function(object, innov,...) {
+  n <- length(object$y)
+  h <- NROW(innov)
+  fourier_year <- forecast::fourier(ts(seq(n), frequency=365.25), K=3, h=h)
+  season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
+  # Use last value of trend for future
+  lasttrend <- tail(splines::ns(seq(n), df=round(n/100)),1)
+  trend <- matrix(rep(lasttrend,h), nrow=h, byrow=TRUE)
+  colnames(trend) <- colnames(lasttrend)
+  X <- cbind(rep(1, h), trend, season_week, fourier_year, holidays[n+seq(h),])
+  colnames(X) <- c("(Intercept)", paste0("X",colnames(X[,-1])))
+  # Remove missing columns
+  X <- X[, colnames(X) %in% names(coefficients(object)), drop=FALSE]
+  # Find mean of future periods
+  mean_future <- c(exp(X %*% matrix(coefficients(object), ncol=1)))
+  # Return Poisson series with this mean
+  ts(rpois(h, lambda=mean_future), frequency = frequency(object$y),
+     start = tsp(object$y)[2] + 1/frequency(object$y))
+}

### 2022-10-28 rscript/hts/tscount.R
Better trend for tscount

@@ -4,7 +4,7 @@ library(tscount)
 tscount <- function(y) {
   fourier_year <- forecast::fourier(ts(y, frequency=365.25), K=3)
   season_week <- forecast::seasonaldummy(y)
-  trend <- splines::ns(seq(length(y)), df=10)
+  trend <- splines::ns(seq(n), df=round(n/100))
   X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
   # Remove constant covariates
   constant <- apply(X, 2, forecast:::is.constant)
@@ -17,7 +17,10 @@ simulate.tsglm <- function(object, innov,...) {
   h <- NROW(innov)
   fourier_year <- forecast::fourier(ts(seq(n), frequency=365.25), K=3, h=h)
   season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
-  trend <- predict(splines::ns(seq(n), df=10), newx=seq(n+h)[-(seq(n))])
+  # Use last value of trend for future
+  lasttrend <- tail(splines::ns(seq(n), df=round(n/100)),1)
+  trend <- matrix(rep(lasttrend,h), nrow=h, byrow=TRUE)
+  colnames(trend) <- colnames(lasttrend)
   X <- cbind(trend, season_week, fourier_year, holidays[n+seq(h),])
   # Remove missing columns
   X <- X[, colnames(X) %in% names(coefficients(object))]

### 2022-10-28 rscript/hts/forecast_hts.R
Use a simple model if fewer than 5% positive values

@@ -5,27 +5,27 @@ source(here::here("rscript/hts/htsplus.R"))
 source(here::here("rscript/hts/tscount.R"))
 
 # Parallelization
-#plan(multisession, workers = 3)
+# plan(multisession, workers = 3)
 
 # Read hierarchical/grouped time series
-#incident_gts <- read_rds(paste0(storage_folder, "incidents_test_gts.rds"))
+# incident_gts <- read_rds(paste0(storage_folder, "incidents_test_gts.rds"))
 incident_gts <- read_rds(paste0(storage_folder, "incidents_gts.rds"))
 holidays <- read_rds(paste0(storage_folder, "holidays_ts.rds"))
 
-# Test sets of size 84, 
+# Test sets of size 84,
 origins <- 42 * seq(10) + 42
-for(i in seq(origins)) {
+for (i in seq(origins)) {
   # Set up training set
   train <- incident_gts
   train$bts <- subset(train$bts, end = nrow(incident_gts$bts) - origins[i])
   # Create reconciled sample paths for different models
   reconcile_sample_paths(train, model_function = "ets")
-  #reconcile_sample_paths(train, model_function = "tscount")
+  # reconcile_sample_paths(train, model_function = "tscount")
   reconcile_sample_paths(train, model_function = "iglm")
 }
 
 # Summary statistics of forecast accuracy
-#rmsse_ets <- rmsse(train, test, model_function = "ets", method = "wls")
-#crps_ets <- crps(fcst_ets, test)
-#rmsse_tscount <- rmsse(fcst_tscount, test, train)
-#crps_tscount <- crps(fcst_tscount, test)
+# rmsse_ets <- rmsse(train, test, model_function = "ets", method = "wls")
+# crps_ets <- crps(fcst_ets, test)
+# rmsse_tscount <- rmsse(fcst_tscount, test, train)
+# crps_tscount <- crps(fcst_tscount, test)

### 2022-10-28 rscript/hts/glm.R

@@ -1,39 +1,50 @@
 # Specific tsglm model used for this data
 iglm <- function(y) {
   n <- length(y)
-  fourier_year <- forecast::fourier(ts(y, frequency=365.25), K=3)
-  season_week <- forecast::seasonaldummy(y)
-  trend <- splines::ns(seq(n), df=round(n/100))
-  X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
-  # Remove constant covariates
-  constant <- apply(X, 2, forecast:::is.constant)
-  X <- X[, !constant]
-  object <- suppressWarnings(try(glm(y ~ X, family=poisson()), silent=TRUE))
-  if(inherits(object, "try-error"))
+  simple_fit <- mean(y > 0) < 0.05
+  if(!simple_fit) {
+    fourier_year <- forecast::fourier(ts(y, frequency = 365.25), K = 3)
+    season_week <- forecast::seasonaldummy(y)
+    trend <- splines::ns(seq(n), df = round(n / 100))
+    X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y), ])
+    # Remove constant covariates
+    constant <- apply(X, 2, forecast:::is.constant)
+    X <- X[, !constant]
+    object <- suppressWarnings(try(glm(y ~ X, family = poisson()), silent = TRUE))
+    if (inherits(object, "try-error")) {
+      simple_fit <- TRUE
+    }
+  }
+  if(simple_fit) {
     object <- glm(y ~ 1, family = poisson())
+  }
   # Slim down return object and add ts attributes
-  object <- list(coefficients=object$coefficients, 
-                 residuals=residuals(object, type="response"),
-                 y=y)
-  return(structure(object, class='iglm'))
+  object <- list(
+    coefficients = object$coefficients,
+    residuals = residuals(object, type = "response"),
+    y = y
+  )
+  return(structure(object, class = "iglm"))
 }
 
-simulate.iglm <- function(object, innov,...) {
+simulate.iglm <- function(object, innov, ...) {
   n <- length(object$y)
   h <- NROW(innov)
-  fourier_year <- forecast::fourier(ts(seq(n), frequency=365.25), K=3, h=h)
-  season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
[…]

### 2022-10-28 rscript/hts/htsplus.R

@@ -142,9 +142,9 @@ future_sample_paths <- function(object, model_function = "ets", h = 84, nsim = 1
     # Just use the previous results
     return(invisible(read_rds(filename)))
   }
-  
+
   models <- fit_models(object, model_function)
-  restype <- dplyr::if_else(identical(model_function,"tscount"), "response", "innovation")
+  restype <- dplyr::if_else(identical(model_function, "ets"), "innovation", "response")
   innov_res <- calculate_residuals(object, model_function, type = restype)
   nseries <- length(models)
 

### 2022-10-28 rscript/hts/glm.R
Stylistic changes only

@@ -1,39 +1,44 @@
 # Specific tsglm model used for this data
 iglm <- function(y) {
   n <- length(y)
-  fourier_year <- forecast::fourier(ts(y, frequency=365.25), K=3)
+  fourier_year <- forecast::fourier(ts(y, frequency = 365.25), K = 3)
   season_week <- forecast::seasonaldummy(y)
-  trend <- splines::ns(seq(n), df=round(n/100))
-  X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
+  trend <- splines::ns(seq(n), df = round(n / 100))
+  X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y), ])
   # Remove constant covariates
   constant <- apply(X, 2, forecast:::is.constant)
   X <- X[, !constant]
-  object <- suppressWarnings(try(glm(y ~ X, family=poisson()), silent=TRUE))
-  if(inherits(object, "try-error"))
+  object <- suppressWarnings(try(glm(y ~ X, family = poisson()), silent = TRUE))
+  if (inherits(object, "try-error")) {
     object <- glm(y ~ 1, family = poisson())
+  }
   # Slim down return object and add ts attributes
-  object <- list(coefficients=object$coefficients, 
-                 residuals=residuals(object, type="response"),
-                 y=y)
-  return(structure(object, class='iglm'))
+  object <- list(
+    coefficients = object$coefficients,
+    residuals = residuals(object, type = "response"),
+    y = y
+  )
+  return(structure(object, class = "iglm"))
 }
 
-simulate.iglm <- function(object, innov,...) {
+simulate.iglm <- function(object, innov, ...) {
   n <- length(object$y)
   h <- NROW(innov)
-  fourier_year <- forecast::fourier(ts(seq(n), frequency=365.25), K=3, h=h)
-  season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
+  fourier_year <- forecast::fourier(ts(seq(n), frequency = 365.25), K = 3, h = h)
+  season_week <- forecast::seasonaldummy(ts(seq(n), frequency = 7), h = h)
   # Use last value of trend for future
-  lasttrend <- tail(splines::ns(seq(n), df=round(n/100)),1)
-  trend <- matrix(rep(lasttrend,h), nrow=h, byrow=TRUE)
[…]

### 2022-10-28 rscript/hts/tscount.R

@@ -2,43 +2,45 @@ library(tscount)
 
 # Specific tsglm model used for this data
 tscount <- function(y) {
-  fourier_year <- forecast::fourier(ts(y, frequency=365.25), K=3)
+  fourier_year <- forecast::fourier(ts(y, frequency = 365.25), K = 3)
   season_week <- forecast::seasonaldummy(y)
-  trend <- splines::ns(seq(n), df=round(n/100))
-  X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y),])
+  trend <- splines::ns(seq(n), df = round(n / 100))
+  X <- cbind(trend, season_week, fourier_year, holidays[seq_along(y), ])
   # Remove constant covariates
   constant <- apply(X, 2, forecast:::is.constant)
   X <- X[, !constant]
-  tsglm(y, model = list(past_obs = 1:3), xreg=X, link="log")
+  tsglm(y, model = list(past_obs = 1:3), xreg = X, link = "log")
 }
 
-simulate.tsglm <- function(object, innov,...) {
+simulate.tsglm <- function(object, innov, ...) {
   n <- length(object$response)
   h <- NROW(innov)
-  fourier_year <- forecast::fourier(ts(seq(n), frequency=365.25), K=3, h=h)
-  season_week <- forecast::seasonaldummy(ts(seq(n), frequency=7), h=h)
+  fourier_year <- forecast::fourier(ts(seq(n), frequency = 365.25), K = 3, h = h)
+  season_week <- forecast::seasonaldummy(ts(seq(n), frequency = 7), h = h)
   # Use last value of trend for future
-  lasttrend <- tail(splines::ns(seq(n), df=round(n/100)),1)
-  trend <- matrix(rep(lasttrend,h), nrow=h, byrow=TRUE)
+  lasttrend <- tail(splines::ns(seq(n), df = round(n / 100)), 1)
+  trend <- matrix(rep(lasttrend, h), nrow = h, byrow = TRUE)
   colnames(trend) <- colnames(lasttrend)
-  X <- cbind(trend, season_week, fourier_year, holidays[n+seq(h),])
+  X <- cbind(trend, season_week, fourier_year, holidays[n + seq(h), ])
   # Remove missing columns
   X <- X[, colnames(X) %in% names(coefficients(object))]
   output <- tsglm.sim(h,
-        param = list(
-          intercept = object$coefficient[1],
-          past_obs = object$coefficients[2:4],
[…]

### 2022-11-13 rscript/hts/ensemble.R
Fixed bugs

@@ -45,9 +45,9 @@ create_specific_ensemble <- function(models, method, origin) {
   # Find simulation files
   files <- fs::dir_ls(storage_folder, glob = paste0("*_", origin, "_sim_", method, ".rds"))
   # For each training set, combine available simulation files
-  new_sim <- array(0, c(1224, 84, 1000 * length(files_i)))
-  for (j in seq_along(files_i)) {
-    new_sim[, , 1000 * (j - 1) + seq(1000)] <- read_rds(files_i[j])
+  new_sim <- array(0, c(1224, 84, 1000 * length(files)))
+  for (j in seq_along(files)) {
+    new_sim[, , 1000 * (j - 1) + seq(1000)] <- read_rds(files[j])
   }
   write_rds(new_sim, filename, compress = "bz2")
 }

### 2022-11-13 rscript/hts/tscount.R

@@ -30,7 +30,7 @@ tscount <- function(y) {
 }
 
 simulate.tsglm <- function(object, innov, ...) {
-  n <- length(object$response)
+  n <- length(object$y)
   h <- NROW(innov)
   fourier_year <- forecast::fourier(ts(seq(n), frequency = 365.25), K = 3, h = h)
   season_week <- forecast::seasonaldummy(ts(seq(n), frequency = 7), h = h)
@@ -62,7 +62,7 @@ simulate.tsglm <- function(object, innov, ...) {
     )$ts
   }
   ts(output,
-    frequency = frequency(object$response),
-    start = tsp(object$response)[2] + 1 / frequency(object$response)
+    frequency = frequency(object$y),
+    start = tsp(object$y)[2] + 1 / frequency(object$y)
   )
 }

### 2022-11-13 rscript/hts/forecast_hts.R
Added quantile combinations

@@ -4,10 +4,14 @@ library(readr)
 library(furrr)
 source(here::here("rscript/hts/htsplus.R"))
 source(here::here("rscript/hts/glm.R"))
+source(here::here("rscript/hts/tscount.R"))
 source(here::here("rscript/hts/naive.R"))
+source(here::here("rscript/hts/ensemble.R"))
+source(here::here("rscript/hts/qcombinations.R"))
+source(here::here("rscript/hts/accuracy.R"))
 
 # Parallelization
-# plan(multisession, workers = 3)
+plan(multisession, workers = 3)
 
 # Read hierarchical/grouped time series
 # incident_gts <- read_rds(paste0(storage_folder, "incidents_test_gts.rds"))
@@ -28,6 +32,7 @@ for (i in seq(origins)) {
 }
 
 create_ensembles()
+create_qcomb()
 
 # Accuracy
 

### 2022-11-13 rscript/hts/qcombinations.R

@@ -1,0 +1,69 @@
+library(stringr)
+library(readr)
+
+create_qcomb <- function(models_to_use = "all") {
+  # Find simulation files
+  files <- fs::dir_ls(storage_folder, glob = paste0("*_sim_*.rds"))
+  # Find models
+  models <- str_remove(files, storage_folder) |>
+    str_extract("[a-zA-Z]*_") |>
+    str_remove("_") |>
+    unique()
+  if (!identical(models_to_use, "all")) {
+    models <- models[models %in% models_to_use]
+    if (!identical(models, models_to_use)) {
+      stop("Some models not available")
+    }
+  }
+  # Find methods
+  methods <- str_remove(files, storage_folder) |>
+    str_extract("[a-zA-Z]*.rds") |>
+    str_remove(".rds") |>
+    unique()
+  methods <- methods[methods != ""]
+  # Find separate training sets
+  origins <- readr::parse_number(files) |>
+    unique() |>
+    sort() |>
+    as.character()
+  for (j in seq_along(methods)) {
+    for (k in seq_along(origins)) {
+      create_specific_qcomb(models_to_use, methods[j], origins[k])
+    }
+  }
+}
+
+create_specific_qcomb <- function(models, method, origin) {
+  # Form file name for saving results
+  filename <- paste0(storage_folder, "qcomb_", origin, "_sim_", method, ".rds")
+  # Check if this has already been run
+  if (fs::file_exists(filename)) {
+    # Just use the previous results
+    return(invisible(read_rds(filename)))
+  }
+
+  # Find simulation files
+  files <- fs::dir_ls(storage_folder, glob = paste0("*_", origin, "_sim_", method, ".rds"))
+  files <- files[!grepl("ensemble", files)]
+  # For each training set, load available simulation files and compute quantiles
+  # with same average quantiles
+  prob <- seq(0.001, 0.999, by=0.001)
+  q <- array(0, c(1224, 84, length(files), length(prob)))
+  for(k in seq_along(files)) {
+    tmp <- readr::read_rds(files[k])
+    for(i in seq(1224)) {
+      for(j in seq(84)) {
+        q[i, j, k, ] <- quantile(tmp[i,j,], prob=prob)
+      }
+    }
+  }
+  aveq <- apply(q, c(1,2,4), mean) 
+  # Generate simulated data with same distribution
+  sim <- array(0, c(1224, 84, 1000))
[…]

### 2022-11-14 manuscript/02-literature.qmd
update section 1 and 2

@@
 # Research background {#lit}
-Emergency medical services (EMS), also referred to as ambulance, paramedic or pre-hospital emergency services, are a critical component in the delivery of urgent medical care to communities.
+Emergency medical services (EMS) are a critical component in the delivery of urgent medical care to communities.
 An effective service delivery requires accurate resource planning that are generally replying on demand forecasts at operational, tactical and strategic levels.
 
-There is a substantial literature on the application of forecasting in the Emergency Medical Services.
-Various areas have been the main focus.
-One of the major research topics has been forecasting call volume arrivals.
+There is a substantial number of studies on the application of time series forecasting in the Emergency Medical Services.
+Various areas have been the focus of the literature.
+Forecasting call volume arrivals is one of the major research topics.
 @ibrahim2016modeling provide an extensive review of the forecasting models in this context.
-Another important focus of the literature in EMS is related to forecasting ambulance demand.
-Although sometimes the definition of the demand is not cleanly stated, however this is generally the demand where a physical resource has been deployed to respond to an incident, which is also referred to attended incidents.
-Another demand related variable is the verified incidents.
-These are all incidents that require an action: either send a physical vehicle, deal with via the Clinical Support Desk, get an external (private) provider to respond to it, or send it through to other channels such as police, firefighters or general practitioners.
-Another domain that has been extensively studied in the literature and it is also close to ours is forecasting Emergency Department attendance.
-We refer interested readers to @mingliterature2022, @gul2020exhaustive and @wargon2009systematic for extensive reviews of the literature in these areas.
+Another important area is related to forecasting ambulance demand.
+Although the definition of demand might not be always clearly stated, however, this is typically referring to a situation where a physical resource has been deployed to respond to an incident.
+This might be also called _attended incidents_.
+Another demand related variable is verified incidents.
+These are all incidents that require an action: either send a physical vehicle, deal with via the Clinical Support Desk (e.g.
+calls), get an external (private) provider to respond to it, or send it through to other channels such as police, firefighters or general practitioners.
+Our study is aligned with this stream of the literature.
+Another similar area that is largely studied in the literature, is Emergency Department forecasting.
+We refer interested readers to @mingliterature2022, @gul2020exhaustive and @wargon2009systematic for extensive reviews of the literature on Emergency Department forecasting.
 Although crucial to EMS performance, @aringhieri2017emergency state that demand forecasting has received limited research attention in the EMS context.
@@
 
-There are generally two main streams of research in this area: i) the first stream is focusing on the application of time series methods and regression approaches on forecasting aggregate ambulance demand as the most widely used approaches [@vile2012predicting; @sasaki2010using]; ii) another stream considers forecasting EMS demand in more finer temporal and geographical granularitties by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit].
-Papers studying the second stream has not used classical regression and time-series approaches like the first one.
-Our study belongs to the first category, but we briefly review research in both streams.
+There are generally two main streams of research related to forecasting ambulance demand in EMS: i) the first stream focuses on the application of time series methods and regression approaches on forecasting aggregate ambulance demand [@vile2012predicting; @sasaki2010using]; ii) the second stream considers forecasting EMS demand in a more finer temporal and geographical granularitties by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit].
+The focus of our study is related to the first stream of research.
 
@@
 @vile2016time further develop a decision support system to integrate forecasts generated by SSA.
-However, the study does not compare and contrast of the performance of forecasting methods based on some utility measures such as cost, resource utilization or response time.
+However, the study does not compare and contrast the performance of forecasting methods based on utility measures such as cost, resource utilization or response time.
 The tool contains options that allow generating forecasts at various levels of granularity, however, it ignore the hierarchical and grouped relationships structure, preventing  aligned decision making and coordination.
@@
 They used multi-layer perceptron (MLP) and long short-term memory (LSTM) models to forecast the EMS demand, and compare them to simple aggregation methods and baselines.
-The point forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean Squared Error (MSE) and the forecast distribution is measures by
-Categorical Cross-Entropy.
+The point forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean Squared Error (MSE) and the forecast distribution is measures by Categorical Cross-Entropy.
 They shows that while Neural Network models perform better in producing point forecast, a distribution baseline method based on spatial distribution of the incidents across all time steps provide more accurate forecast distribution.
@@
 The proposed approach is tested using data of 2851 events attended by the EMS in Athens, Greece over 24 weeks.
-They show that 23.24% of real emergency events lie within 50 m of the predicted ones and nearly 70% of the real emergency events lie no further away than 150 m, which is rather accurate given the granularity of the problem at the city level.
+They show that $23.24\%$ of real emergency events lie within 50 meter of the predicted ones and nearly $70\%$ of the real emergency events lie no further away than 150 meter, which is rather accurate given the granularity of the problem at the city level.
 
-One of the major limitations of the current forecasting research in healthcare and EMS is the lack of a forecasting methodology that accounts for the inherent hierarchical and grouped structure in data to produce coherent forecasts, enabling aligned planning and decision making and possibly forecast accuracy improvement by exploring the demand relationships between different levels.
-While this methodology has been developed over the past 10 years [@panagiotelis2022probabilistic], it has never been explored in this area.
-Additionally, we found that most research only deals with point forecast estimation and also reproducibility is still a major concern in this area.
-Also, given the count nature of Considering these limitations from the literature, we develop a forecasting experiment to fill these gaps.
+We observed a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research. These limitations are summarized as following:
+
+1.
+Current studies ignore the inherent hierarchical and/or grouped structure of the time series data and the relationship between series at different levels of hierarchy.
+This may result in incoherent forecasts leading to misaligned planning and decision making.
+While the hierarchical forecasting methodology has been developed and applied in various domains over the past 10 years [@panagiotelis2022probabilistic], it has never been explored in this area.
+
+2.
+Current research is mainly concerned with generating point forecast at a single level of hierarchy.
+There is a lack of studies presenting the entire forecast distribution of daily ambulance demand for the entire hierarchy to inform the whole decision-making process and to better represent the uncertainty of future demand, providing a risk management tool for planners.
+
+3. Reproducibility is still a major challenge in EMS forecasting, as it is unlikely to reproduce the results without the help of the authors of those papers.
+
+4.
+Another limitation is related to the generated forecasts that are not integer counts.
+Since actual ambulance counts cannot be negative or fractions, ambulance demand forecasts should be the same.
+While this might not be an issue when producing forecasts at a single level, producing non-negative count forecasts in a hierarchical/grouped structure is challenging and requires further investigation in the future.
+
+This paper concerns the problem of hierarchical forecasting in EMS and generates and evaluates both point and probabilistic forecast across different levels of the hierarchy, hence addressing some important gaps identified in the literature.

### 2022-11-14 manuscript/main.qmd
update section 1 and 2

--- 2022-10-26 manuscript/main.qmd
@@
 ---
-title: "Hierarchical and grouped time series forecasting: an application to Emergency Medical Services"
-execute:
-  echo: false
-  message: false
-  warning: false
-  cache: true
-format:
-  elsevier-pdf:
-    keep-tex: true
-    journal:
-      name: Annals of Emergency Medicine
-      formatting: preprint
-      model: 3p
-      cite-style: authoryear
-  elsevier-html: 
-    toc: true
+title: "Hierarchical time series forecasting in Emergency Medical Services"
 author:
@@
 We apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent, for the first time in the healthcare literature.
-keywords: [Emergency services, healthcare, hierarchical time series forecasting, forecast reconciliation]
-bibliography: mybibfile.bib
+keywords: 
+  - forecasting
+  - healthcare
+  - emergency services
+  - forecast reconciliation
+  - hierarchical time series
+  - ambulance demand
+  - atteneded incidents
+
+bibliography: bibliography.bib
+execute:
+  echo: false
+  message: false
+  warning: false
+  cache: true
+format:
+  elsevier-pdf:
+    keep-tex: true
+    journal:
+      name: Annals of Emergency Medicine
+      formatting: preprint
+      model: 3p
+      cite-style: authoryear
+  elsevier-html: 
+    toc: true
 ---

### 2022-11-14 rscript/hts/accuracy.R
Remove stray stop call

@@ -83,7 +83,6 @@ compute_accuracy_specific <- function(train, model_function = "ets", method = "w
         accuracy[j, k] <- crps_sample(e[j, k, ], 0)
       }
     }
-    stop("not yet implemented")
   }
 
   # Overall accuracy

### 2022-11-14 manuscript/01-introduction.qmd
adding first draft of section 3 &4

@@
-# Introduction
+# Introduction {#sec-intro}
+
 Inability to match the resources with the demand in Emergency Medical Services (EMS) results in an overcrowding care system.
-This is a serious problem causing challenging situations on patient flow, that ultimately leads to increasing patient's length of stay, higher waiting times, low patient satisfaction, high staff pressure, increasing health care costs and inaccuracy in electronic medical record [@ekstrom2015forecasting; @ROSTAMITABAR20221197].
-Demand forecasting in EMS is a critical element that helps to depict various courses of action to avoid the mismatch, which can result in massive savings in terms of patient safety and lives.
-An accurate forecasting of the daily demand enables planners and decision makers to manage resources to meet anticipated patients, reconfigure units, redeploy staff and vehicles, where necessary.
-This will have many advantages for patients, staff and the quality of healthcare services provided.
+This is a serious problem causing challenging situations on patient flow with serious consequences on patients, staff and the entire care system [@ekstrom2015forecasting; @ROSTAMITABAR20221197].
+Demand forecasting in EMS is a vital element that helps to depict various courses of action to avoid the mismatch, which can result in massive savings in terms of patient safety and lives.
+An accurate daily demand forecasting enables planners and decision makers to manage resources to meet anticipated patients, reconfigure units, redeploy staff and vehicles, where necessary.
 
-Demand forecasts are typically required at multiple levels of cross-sectional granularities to inform various planning and decision-making processes across the EMS [@hulshof2012taxonomic].
-There are some planning process at the national level (more strategic or long-term) such as workforce resource planning and budgeting; sub-national and regional level (tactical or medium-term) such as temporary capacity expansions, resource sharing and staff-shift scheduling; and healtboard, hospital or station level (operational or short-term) such as planning rosters for staff and ambulance deployment.
-Demand forecasts might be also required at different level for a specific group of interest such as nature of demand (e.g.
-type of incidents) or demand priority.
-Moreover, there exists an inherent hierarchical and grouped structure relationships that typically occurs in EMS data.
-Demand for emergency services at the country level can be disaggregated in a geographical hierarchy into sub-national, regions and healthboards or divided into groups such as nature of incidents or demand priority.
+Demand forecasts at EMS are typically required at multiple levels of cross-sectional granularities to inform various planning and decision-making processes [@hulshof2012taxonomic].
+There are some planning process at the national level (more strategic or long-term) such as workforce resource planning and budgeting; sub-national, regional or healthcare level (tactical or medium-term) such as temporary capacity expansions, resource sharing and staff-shift scheduling; and hospital or station level (operational or short-term) such as planning rosters for staff and ambulance deployment.
+Demand forecasts might also be required at different level for a specific group of interest such as nature of demand or priority.
+Moreover, the time series data in EMS has an inherent hierarchical and grouped structure to support such forecasting requirements.
+Demand for emergency medical services at the country level can be disaggregated in a geographical hierarchy into sub-national, regions, health boards, stations/hospitals or divided into groups such as nature of incidents or demand priority.
+Therefore, using forecasting methodologies that account for hierarchical and/or grouped structures of time series in EMS seems to be a natural fit.
 
-Despite a large number of studies dedicated to forecasting for Emergency Medical Services [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the main focus has been on producing independent forecasts at a single level, ignoring the inherent hierarchical or grouped structure relationships of the time series demand.
-However, forecasting the EMS demand across all hierarchical and grouped levels is paramount for an effective planning and decision making.
-The hierarchical forecasting approach can not only achieve more accurate forecasts than the independent (base) forecasts, but also creates consistent forecasts [@hyndman2011optimal].
-Obtaining consistent forecasts at different levels is important for reliable decision-making to avoid making conflicting decisions.
-With hierarchical forecasting, plans at any level are based on identical forecasts and therefore can be aligned.
-Implementing and sustaining improvements in EMS requires alignments and coordination between different stakeholders, without which teams operate in isolation that may lead to conflicts, duplication work, rework, or work that runs counter to overall goals to improve hospital-wide patient flow.
-Hierarchical forecasting framework can also be used as a tools to improve coordination between teams across the care services at the national, sub-national, regional and hospital levels.
-To our knowledge, there is not only no research in the EMS forecasting to account for the hierarchical and grouped structure of the system but also in the entire field of forecasting for healthcare management (also referred to as non-clinical healthcare).
+However, despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], this area is neglected and the main focus has been on producing forecasts at a single level, independently.
+Generating independent forecasts not only ignore the inherent hierarchical and/or grouped structure relationships of the time series demand but also results in a lack of consistency and coordination.
+Consistent forecasts of the EMS demand across all hierarchical and grouped levels are paramount for an effective planning and decision making.
+The hierarchical forecasting approaches can not only create consistent forecasts but also achieve more accurate forecasts than the independent (base) forecasts [@hyndman2011optimal].
+Obtaining consistent forecasts at different levels is important as it helps to avoid making conflicting decisions.
+With hierarchical forecasting, plans at any level are based on coherent forecasts and therefore can be aligned.
+Implementing and sustaining improvements in EMS require alignments and coordination between different stakeholders, without which teams operate in isolation leading to conflicts, duplication work, rework, or work that runs counter to the overall goal to improve the quality of delivery service.
+Hierarchical forecasting framework can be used as a tool to improve coordination between teams across the care services at the national, sub-national, regional and local levels.
+To our knowledge, there is not only no research in the EMS forecasting to account for the hierarchical and grouped structure of the system but also in the entire field of forecasting for healthcare management.
 
-In this paper, we address this gap by investigating the application of forecast reconciliation methods in the Emergency Medical Services using the daily time series of verified incidents in the Great Britain from 2015 to 2020.
-The data has a hierarchical and grouped structures with hierarchies at the national, control (sub-national), healthboard (corresponding to regions) and groups by priority and nature of incidents.
-We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal combination approaches.
-We use Exponential smoothing, ARIMA and Poisson regression to generated the independent based forecast followed by applying bottom-up and optimal reconciliation approaches.
-Forecast performance are assessed by the Root Mean Squared Scaled Error (RMSSE) for point forecasting and the Continuous Ranked Probability Score (CRPS) for the probabilistic forecast.
-This paper complies with the principles of reproducibility  [@stodden2013best; @boylan2015reproducibility] and utilising fable package [@fable2022] in R software.
-Therefore, this study could equally be applied to any healthcare service (e.g.
-emergency department, primary or social care) subject to the time series demand having a hierarchical and/or grouped structure, which is generally the case in the healthcare sector.
+In this paper, we address this gap by investigating the application of  hierarchical forecasting approaches in the EMS using daily time series of verified incidents from 2015 to 2020 in a major ambulance service in Great Britain.
+The data has hierarchical and grouped structures with hierarchies at the national, control (i.e.
+sub-national), health board (i.e.
+regional) and groups by priority and nature of incidents.
+We produce not only the point forecast but also the forecast distribution across all levels, which is critical for an effective planning and associated risk management.
+We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal reconciliation approaches.
+We first generate independent/based forecasts using Exponential Smoothing State Space (ETS),  Generalised Linear Model (GLM), Poisson regression, a simple empirical distribution and an ensemble method followed by applying bottom-up and optimal reconciliation approaches.
+Forecast performance is assessed by the Root Mean Squared Scaled Error (RMSSE) for point forecasts and the Continuous Ranked Probability Score (CRPS) for the probabilistic forecasts.
+This paper complies with the principles of the reproducibility  [@stodden2013best; @boylan2015reproducibility].
+Therefore, the study could equally be applied to any healthcare service (e.g.
+emergency department, primary or social care) subject to the time series having a hierarchical and/or grouped structure, which is generally the case in the healthcare sector.
 
-The remainder of this article is structured as follows: In section 2, we provide a brief review of the
-literature and discuss its limitation to position our work; in section 3, we present the experiment design describing the dataset, forecasting methods and forecast evaluation metrics.
-In Section 4, we describe the hierarchical and grouped time series forecasting methods for producing point and probabilistic forecasts.
-In section 5, we present and discuss our results; in section 7, we summarize
-our findings and present ideas for future research.
+The remainder of this article is structured as follows: In section 2, we provide a brief review of the literature and discuss its limitation to position our work; in section 3, we present the experiment design describing the data set, forecasting methods and forecast evaluation metrics.
+In Section 4, we discuss the hierarchical time series forecasting approaches to generate both point and probabilistic forecasts.
+In section 5, we present and discuss our results; in section 7, we summarize our findings and present ideas for future research.

### 2022-11-14 manuscript/03-experiment.qmd

@@
 library(feasts)
-
+incident_gthf <- read_rds(paste0(storage_folder, "incidents_gt.rds"))
 incident_gthf <- read_rds(here::here("data/incident_gt.rds"))
 incident_gthf %>% 
-  features(incidents,feat_stl) %>%
+  features(incident,feat_stl) %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_week)) +
@@
 
+**Naive:** We start with one of the simplest forecasting approaches used in practice - assuming that in the next few days, everything will be the same as in the similar days in the past.
+This is called "Naïve".
+In our case, given that we need a distribution of values, we will use a modified approach, where the empirical distribution of the daily attended incidents time series is used to forecast the future attended incidents distribution.
+We consider the empirical distribution of the most recent year of historic data on a rolling basis to capture potential changes in behavior over time.
+
 **Exponential Smoothing State Space model (ETS):** ETS models [@hyndman2021forecasting] can combine trend, seasonality and error components in a time series through various forms such as additive, multiplicative or mixed.
@@
 
-**Auto-Regressive Integrated Moving Average (ARIMA):** ARIMA [@hyndman2021forecasting] is another effective family of forecasting models that capture autocorrelation in time series.
+<!-- **Auto-Regressive Integrated Moving Average (ARIMA):** ARIMA [@hyndman2021forecasting] is another effective family of forecasting models that capture autocorrelation in time series.
 They can model various range of time series considering the stationarity of series, the Auto-Regressive (AR) and moving average (MA) orders.
@@
 To select the best model for a given time series, the algorithm uses unit root test, Maximum Likelihood Estimation (MLE) and the AICc.
+-->
 

### 2022-11-14 rscript/hts/forecast_hts.R
Updated accuracy summaries

@@ -52,12 +52,6 @@ mase |>
   summarise(mase = mean(mase)) |>
   arrange(series, mase) |>
   print(n = 200)
-mase |>
-  filter(series == "Overall") |>
-  group_by(method, model, series) |>
-  summarise(mase = mean(mase)) |>
-  arrange(mase) |>
-  print(n = 200)
 
 rmsse |>
   group_by(method, model, series) |>
@@ -65,9 +59,8 @@ rmsse |>
   arrange(series, rmsse) |>
   print(n = 200)
 
-rmsse |>
-  filter(series == "Overall") |>
+crps |>
   group_by(method, model, series) |>
-  summarise(rmsse = sqrt(mean(rmsse^2))) |>
-  arrange(rmsse) |>
+  summarise(crps = mean(crps)) |>
+  arrange(series, crps) |>
   print(n = 200)

### 2022-11-16 manuscript/graphs.R
Added graph of accuracy vs forecast horizon

@@ -1,0 +1,44 @@
+library(tidyverse)
+
+options(
+  ggplot2.continuous.colour = "viridis",
+  ggplot2.continuous.fill = "viridis",
+  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9", "#F0E442"),
+  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9", "#F0E442")
+)
+
+rmsse <- readr::read_rds(here::here("results/rmsse.rds"))
+mase <- readr::read_rds(here::here("results/mase.rds"))
+crps <- readr::read_rds(here::here("results/crps.rds")) |>
+  pivot_wider(names_from = model, values_from = crps) |>
+  group_by(method, h, series) |>
+  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
+  ungroup() |>
+  pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")
+
+accuracy <- bind_rows(
+  rmsse |> mutate(measure = "msse", accuracy = rmsse^2),
+  mase |> mutate(measure = "mase", accuracy = mase),
+  crps |> mutate(measure = "crps", accuracy = crps),
+) |>
+  select(-rmsse, -mase, -crps)
+
+# Plot of average accuracy vs week for each method for Total
+acc_summary <- accuracy |>
+  filter(
+    series %in% c("Total", "Control areas", "Health boards"),
+    method == "mint", model != "qcomb"
+  ) |>
+  mutate(
+    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
+    model = factor(model, levels = c("naiveecdf", "ets", "tscount", "iglm", "ensemble")),
+    week = factor(trunc((h - 1) / 7) + 1)
+  ) |>
+  group_by(week, model, measure, series) |>
+  summarise(accuracy = mean(accuracy), .groups = "drop")
+
+acc_summary |>
+  ggplot(aes(x = week, y = accuracy, group = model)) +
+  geom_line(aes(col = model), size = 1) +
+  facet_grid(measure ~ series, scales = "free_y") +
+  labs(y = "Average accuracy", x = "Week ahead")

### 2022-11-16 manuscript/01-introduction.qmd
Spelling

@@
 We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal reconciliation approaches.
-We first generate independent/based forecasts using Exponential Smoothing State Space (ETS),  Generalised Linear Model (GLM), Poisson regression, a simple empirical distribution and an ensemble method followed by applying bottom-up and optimal reconciliation approaches.
+We first generate independent/based forecasts using Exponential Smoothing State Space (ETS),  Generalized Linear Model (GLM), Poisson regression, a simple empirical distribution and an ensemble method followed by applying bottom-up and optimal reconciliation approaches.
 Forecast performance is assessed by the Root Mean Squared Scaled Error (RMSSE) for point forecasts and the Continuous Ranked Probability Score (CRPS) for the probabilistic forecasts.

### 2022-11-16 manuscript/02-literature.qmd
Spelling

@@
 
-There are generally two main streams of research related to forecasting ambulance demand in EMS: i) the first stream focuses on the application of time series methods and regression approaches on forecasting aggregate ambulance demand [@vile2012predicting; @sasaki2010using]; ii) the second stream considers forecasting EMS demand in a more finer temporal and geographical granularitties by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit].
+There are generally two main streams of research related to forecasting ambulance demand in EMS: i) the first stream focuses on the application of time series methods and regression approaches on forecasting aggregate ambulance demand [@vile2012predicting; @sasaki2010using]; ii) the second stream considers forecasting EMS demand in a more finer temporal and geographical granularities by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit].
 The focus of our study is related to the first stream of research.

### 2022-11-16 manuscript/03-experiment.qmd
Spelling

@@
 The forecast horizon in this study is $fh = 2 \times ph$ days ahead (2*42 days planning horizon).
-This is because the planning is generally freezed for $ph$ days and considering a forecast horizon of $ph$ days might not be helpful for planning.
+This is because the planning is generally frozen for $ph$ days and considering a forecast horizon of $ph$ days might not be helpful for planning.
 While forecasts are generated for $2 \times ph$ days ahead, performance evaluation is only assessed based on the last $ph$ days and not the $2 \times ph$ days.
@@
 #| out.width: "60%"
-#| fig-cap: "The hierarchical and grouped structure of attended inciddents (ambulance demand)"
+#| fig-cap: "The hierarchical and grouped structure of attended incidents (ambulance demand)"
 #| fig-subcap:
-#|   - "Hierarchical structure: Attended inciddents in all country is disaggregated into 3 control areas /regions and then into 7 different healrhboard at the bottom"
+#|   - "Hierarchical structure: Attended incidents in all country is disaggregated into 3 control areas /regions and then into 7 different healthboard at the bottom"
 #|   - "Grouped structure:  Incidents could be grouped into priority (i.e.
-Red, Amber & Green) and the nature of attenedd incident (i.e.
-there are 35 different nature of incidents including chest pain , breahtign problems, heart attack, stroke, and son on).
+Red, Amber & Green) and the nature of attended incident (i.e.
+there are 35 different nature of incidents including chest pain , breathing problems, heart attack, stroke, and so on).
 The symbol * refers to the crossed attributes between hierarchical and grouped levels."

### 2022-11-16 manuscript/main.qmd
Updated abstract

@@
 abstract: |
-  The primary goal of an emergency service is to minimize the ambulance response time.
-Accurate forecasts of ambulance demand are crucial inputs into the decisions of planners and policymakers on the resource (e.g.
-staff and fleet) planning and dynamic deployment.
-Informing those decisions require detailed forecasting of demand at multiple levels of hierarchy from the national and regional levels to the nature of incidents and their priorities.
-These forecasts are often generated independently by different teams.
-However, independent forecasts may not add up to the other levels resulting in mismatched forecasts that lead to conflicting decisions and a lack of coordination in the care service.
-To address this issue, we explore the inherently hierarchical and grouped structure of Emergency Medcine Services.
-We apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent, for the first time in the healthcare literature.
+  Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet.
+Such demand forecasts are required at national, regional and sub-regional levels, and must take account of the nature of incidents and their priorities.
+These forecasts are often generated independently by different teams within the organization.
+As a result, forecasts at different levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service.
+To address this issue, we exploit the hierarchical and grouped structure of the demand time series, and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation.
+The methods are applied to daily incident data from the Welsh Ambulance Service NHS Trust, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area.
+We use an ensemble of forecasting models, and show that the resulting forecasts are better than any individual forecasting model.
+We validate the forecasting approach using time-series cross-validation.
 keywords:
@@
   - ambulance demand
-  - atteneded incidents
+  - attended incidents
 

### 2022-11-16 manuscript/03-experiment.qmd
Cleaner code

@@
 
-library(ggraph)
-library(igraph)
-library(tidyverse)
-
 data <- data.frame(
-  level1="Total",
-  level2=c("Central & West", "Central & West","Central & West","North", "South & East","South & East","South & East"),
-  level3=c("HD", "SB", "PO", "BC","CV", "CT", "AB")
+  level1 = "Total",
+  level2 = c("Central & West", "Central & West", "Central & West", 
+             "North", "South & East", "South & East", "South & East"),
+  level3 = c("HD", "SB", "PO", "BC", "CV", "CT", "AB")
 )
 # transform it to a edge list!
-edges_level1_2 = data %>% select(level1, level2) %>% 
-  unique %>% rename(from=level1, to=level2)
-edges_level2_3 = data %>% select(level2, level3) %>% 
-  unique %>% rename(from=level2, to=level3)
-edge_list=rbind(edges_level1_2, edges_level2_3)
+edges_level1_2 <- data %>%
+  select(level1, level2) %>%
+  unique() %>%
+  rename(from = level1, to = level2)
+edges_level2_3 <- data %>%
+  select(level2, level3) %>%
+  unique() %>%
+  rename(from = level2, to = level3)
+edge_list <- rbind(edges_level1_2, edges_level2_3)
 
-mygraph <- graph_from_data_frame( edge_list )
-ggraph(mygraph, layout = 'dendrogram', circular = FALSE) + 
-  geom_edge_diagonal() +
-  geom_node_point(colour="#dddddd",size=10) +
-  geom_node_text(
-    aes(label=c("All country", "Central & West", "North","South & East", "HD", "SB", "PO", "BC","CV", "CT", "AB") )
-  )+
+mygraph <- igraph::graph_from_data_frame(edge_list)
+ggraph::ggraph(mygraph, layout = "dendrogram", circular = FALSE) +
+  ggraph::geom_edge_diagonal() +
+  ggraph::geom_node_point(color = "#dddddd", size = 10) +
+  ggraph::geom_node_text(
+    aes(label = c("All country", 
+                  "Central & West", "North", "South & East", 
+                  "HD", "SB", "PO", "BC", "CV", "CT", "AB"))
+  ) +
   theme_void()
@@
 )
-knitr::kable(agg_level, booktabs = T, position = "left") %>% kable_classic(full_width = F)
+knitr::kable(agg_level, booktabs = TRUE, position = "left") %>%
+  kable_classic(full_width = FALSE)
 ```
@@
 #| fig-cap: "Time series features of attended incidents across all levels (1530 series)"
-library(feasts)
-incident_gthf <- read_rds(paste0(storage_folder, "incidents_gt.rds"))
-incident_gthf <- read_rds(here::here("data/incident_gt.rds"))
+incident_gthf <- readr::read_rds(here::here("data/incidents_gt.rds"))
 incident_gthf %>%
-  features(incident,feat_stl) %>%
+  features(incident, feat_stl) %>%
   ggplot(aes(x = trend_strength, y = seasonal_strength_week)) +
-  geom_point(pch=1)+
-  ggthemes::theme_few()
+  geom_point(alpha = 0.5) +
+  ggthemes::theme_few() +
+  labs(x = "Strength of trend", y = "Strength of weekly seasonality")
 ```
@@
 #| out.width: "70%"
-#| fig-cap: "Time series of attended incidents at various levels"
+#| fig-cap: "Time series of attended incidents at various levels."
 #| fig-subcap:
-#|   - "all country"
-#|   - "Control area"
-#|   - "Health board and control area"
-#|   - "Healthboard and priority"
-#|   - "Nature of incident and health board"
+#|   - "Whole country"
+#|   - "By control area"
+#|   - "By health board and control area"
+#|   - "By health board and priority"
+#|   - "By nature of incident and health board"
 #| layout: [[44,-2, 44], [44,-2, 44], [100]]
@@
   filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-    autoplot()+
-    labs(x="Date", y="Incidents")+
+  autoplot() +
+  labs(x = "Date", y = "Incidents") +
   ggthemes::theme_few()
@@
   filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>% select(-nature,-category) %>% 
-  group_by(date,region) %>% summarise(incidents=sum(incidents), .groups="drop") %>% 
-  ggplot(aes(x=date, y= incidents, colour = factor(region))) +
-  geom_line()+
-  labs(x="Date", y="Incidents", colour = "Control")+
-  ggthemes::scale_color_colorblind()+
-  ggthemes::theme_few()+
+  as_tibble() %>%
+  select(-nature, -category) %>%
+  group_by(date, region) %>%
+  summarise(incident = sum(incident), .groups = "drop") %>%
+  ggplot(aes(x = date, y = incident, color = factor(region))) +
+  geom_line() +
+  labs(x = "Date", y = "Incidents", color = "Control") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few() +
   theme(legend.position = "bottom")
@@
   filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>% select(-nature,-category) %>% 
-  ggplot(aes(x=date, y= incidents, colour = factor(lhb))) +
-  geom_line()+
-  facet_wrap(vars(factor(region)), scales = "free_y")+
-  labs(x="Date", y="Incidents", colour = "Health board")+
-  ggthemes::scale_color_colorblind()+
-  ggthemes::theme_few()+
+  as_tibble() %>%
+  select(-nature, -category) %>%
+  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
+  geom_line() +
+  facet_wrap(vars(factor(region)), scales = "free_y") +
+  labs(x = "Date", y = "Incidents", color = "Health board") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few() +
   theme(legend.position = "bottom")
@@
   filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>% select(-nature,-region) %>% 
-  ggplot(aes(x=date, y= incidents, colour = factor(lhb))) +
-  geom_line()+
-  facet_wrap(vars(factor(category)), scales = "free_y")+
-  labs(x="Date", y="Incidents", colour = "Health board")+
-  ggthemes::scale_color_colorblind()+
-  ggthemes::theme_few()+
+  as_tibble() %>%
+  select(-nature, -region) %>%
+  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
+  geom_line() +
+  facet_wrap(vars(factor(category)), scales = "free_y") +
+  labs(x = "Date", y = "Incidents", color = "Health board") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few() +
   theme(legend.position = "bottom")
@@
   filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & !is_aggregated(nature)) %>%
-as_tibble() %>% mutate(nature=as.character(nature)) %>% 
-  filter(nature %in% selected_nature) %>% group_by(date,nature,lhb) %>% summarise(incidents=sum(incidents), .groups="drop") %>% 
-  ggplot(aes(x=date, y= incidents, colour = factor(lhb))) +
-  geom_line()+
-  facet_wrap(vars(factor(nature)), scales = "free_y")+
-  labs(x="Date", y="Incidents", colour = "Health board")+
-  ggthemes::scale_color_colorblind()+
-  ggthemes::theme_few()+
+  as_tibble() %>%
+  mutate(nature = as.character(nature)) %>%
+  filter(nature %in% selected_nature) %>%
+  group_by(date, nature, lhb) %>%
+  summarise(incident = sum(incident), .groups = "drop") %>%
+  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
+  geom_line() +
+  facet_wrap(vars(factor(nature)), scales = "free_y") +
+  labs(x = "Date", y = "Incidents", color = "Health board") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few() +
   theme(legend.position = "none")

### 2022-11-16 manuscript/main.qmd
Cleaner code

@@
 #| include: false
+library(fpp3)
 library(knitr)
-library(tidyverse)
-library(lubridate)
+library(kableExtra)
+library(ggthemes)
 library(patchwork)
-library(tsibble)
-library(kableExtra)
-library(fable)
-library(ggthemes)
-library(readxl)
-library(here)
-library(tsibble)
-library(fable)
+#library(ggraph)
+#library(igraph)
 ```

### 2022-11-16 rscript/result/analysing_results.Rmd
Cleaner code

@@
 
-ggplot(rmsse_control_h, aes(x=h,y=rmsse, colour=factor(approach)))+
+ggplot(rmsse_control_h, aes(x=h,y=rmsse, color=factor(approach)))+
   geom_point()+
@@
 
-ggplot(rmsse_hb_h, aes(x=h,y=rmsse, colour=factor(approach)))+
+ggplot(rmsse_hb_h, aes(x=h,y=rmsse, color=factor(approach)))+
   geom_point()+
@@
 
-ggplot(rmsse_priority_h, aes(x=h,y=rmsse, colour=factor(approach)))+
+ggplot(rmsse_priority_h, aes(x=h,y=rmsse, color=factor(approach)))+
   geom_point()+

### 2022-11-16 manuscript/graphs.R

@@ -1,9 +1,9 @@
 library(tidyverse)
 
 options(
-  ggplot2.continuous.colour = "viridis",
+  ggplot2.continuous.color = "viridis",
   ggplot2.continuous.fill = "viridis",
-  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9", "#F0E442"),
+  ggplot2.discrete.color = c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9", "#F0E442"),
   ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9", "#F0E442")
 )
 

### 2022-11-16 manuscript/main.qmd
Removed html output

@@
       cite-style: authoryear
-  elsevier-html:
-    toc: true
 ---

### 2022-11-16 manuscript/03-experiment.qmd
modify the method section

--- 2022-11-14 manuscript/03-experiment.qmd
@@
 
-We are interested in generating forecasts to inform the planning horizon of $ph= 42$ days, interested by planners in the ambulance services trust.
+We are interested in generating forecasts to inform the planning horizon of $ph= 42$ days, required by planners in the ambulance services trust.
 The forecast horizon in this study is $fh = 2 \times ph$ days ahead (2*42 days planning horizon).
@@
 
-The dataset used in this study is from a major ambulance service trust in the United Kingdom.
+The dataset used in this study is from a major ambulance service trust in Great Britain.
 It contains information relating to the daily number of attended incidents from 10 October 2015 to 31 July 2019 by nature of incidents, priority, the health board managing the service and the control area (or region).
@@
 Therefore, we first compute features of all 1530 time series and display the strength of trend and weekly seasonality strength in Figure @fig-feature.
-Each point represents one time series with strength of trend in  x-axis and strength of seasonality in y-axis.
+Each point represents one time series with the strength of trend in x-axis and the strength of seasonality in y-axis.
 It is clear that there are some series showing strong trends and/or seasonality, corresponding to series at the higher levels of the hierarchy.
-The majority show low trend and seasonality with .
-These are time series belonging to the bottom of the structure, series related to the nature of incidents for each given control, health board and priority.
+The majority of series show low trend and seasonality.
+These are time series belonging to the bottom series, series related to the nature of incidents for a given control, health board and priority level.
+Bottom series are dominated by noise with little or no systematic patterns.
 
@@
 
-In addition to displaying the trend and seasonality features, we also highlight few time series at various levels of the aggregation.
+In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation.
 Figure @fig-dataviz reveals different information such as trend, seasonality and noise.
@@
 
+We consider several forecasting models that account for the diverse patterns of the time series across the entire hierarchy.
+In developing the forecasting models, the time series of holidays are also used in addition to the attended incidents.
+We use public holidays, school holidays and Christmas Day and New Year's Day as predictors of incident attended.
+These type of holidays will  affect peoples' activities and may increase or decrease the number of attended incidents.
+
 ## Forecasting methods
@@
 
-**Naive:** We start with one of the simplest forecasting approaches used in practice - assuming that in the next few days, everything will be the same as in the similar days in the past.
+**Naive:** We start with one of the simplest forecasting approaches used in practice - assuming that in the next few days, everything will be the same as similar days in the past.
 This is called "Naïve".
@@
 
-<!-- **Auto-Regressive Integrated Moving Average (ARIMA):** ARIMA [@hyndman2021forecasting] is another effective family of forecasting models that capture autocorrelation in time series.
-They can model various range of time series considering the stationarity of series, the Auto-Regressive (AR) and moving average (MA) orders.
-In this study, an automatic ARIMA algorithm implemented using `ARIMA()` function of fable package in R is used to forecast the attended incidents.
-To select the best model for a given time series, the algorithm uses unit root test, Maximum Likelihood Estimation (MLE) and the AICc.
--->
+Despite the popularity and the relevance of automatic `ETS` in this study, however it may produce forecasts that are non-integer and negative.
+However, the number of attended incidents is an integer and non-negative.
+When using `ETS` , a time series transformation approach could be used to generate strictly positive forecasts, however forecasts are still not integer.
+An alternative is to use  forecasting models that produce integer, non-negative forecasts.
+in the following section we present Generalized Linear Models (GLMs) and Poisson Regression to produce count base forecasts.
 
-**Poisson Regression: ** Despite the popularity and the relevant of automatic `ETS` and `ARIMA` in this study, but they produce non-integer attended incidents and forecasts might also be negative.
-However, the number of attended incidents is an integer and non-negative.
-When using `ETS` and `ARIMA`, a time series transformation approach such as `sqrt()` could be used to generate strictly positive forecasts, however forecasts are still not integer.
-Another alternative is to use  forecasting models that produce integer, non-negative forecasts.
-One of the models that is frequently used in practice is the Poisson regression [see, for example, @mccarthy2008challenge], which can be summarized as
-$$
-  \begin{aligned}
-    & {y}_t \sim \mathrm{Poisson} \left( \lambda_t \right) \\
-    & \log \lambda_t = \mathbf{x}_t' \boldsymbol{\beta}
-  \end{aligned} \quad
-$$ {#eq-poisson-regression}
+**Generalized Linear Model (GLM): ** GLMs are a family of models developed to extend the concept of linear regression models.
+They perform a regression by modeling the response variable as coming from a particular member of the exponential family, and then transforming the mean of the response so that the transformed mean is a linear function of the predictors.
+One of the models that is frequently used in practice to generate count forecasts is the Poisson regression.
+We will consider forecasting attended incidents using the covariates spline trend,  day of the week dummy variables (from Monday to Sunday), Fourier terms to capture yearly seasonality, public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+We fit a Poisson regression model using the function `glm()` from the package in R, with the argument `family = poisson` to specify that we wish to fit a Poisson regression model.
 
-The logarithm in @eq-poisson-regression is needed to ensure that the parameter of Poisson distribution is always positive.
+**Poisson Regression using tscount: ** We consider another Poisson regression model that takes into account serial dependence in addition to covariates used in the GLM model.
+To that end, we use `tsglm()` function in _tscount_ package in R to model the attended incidents.
+The logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.
 This model can be estimated via maximization of the likelihood function based on Poisson mass function.
-For taking into account serial dependence we include a regression on the previous observation.
-We try to capture the short range serial dependence by a first order autoregressive term and the weekly seasonality by a 7th order autoregressive term.
-We use the `tscount()` function in the fable package in R, which is a wrapper function written based on `tscount` package.
-We fit a model to each time series using the function tsglm, with the identity link function, defined by the argument link.
+The regression model captures the short range serial dependence by including the three order autoregressive terms.
+
+**Ensemble method:** we also use an ensemble method that combines forecasts generated from Naive, ETS, GLM and Poisson regression together to make more accurate forecasts than any individual
+model.
+
+<!-- add a sentence to describe how the point forecast and forecast distribution is created for the ensemble -->
 
@@
 
+The data is spitted into training (up until 2018-07-31) and test (from  2018-08-01 to the end) sets, with all model development and hyper-parameter tuning performed using training data only.
+The time series cross validation is performed with a forecast horizon of $2*42$ days and advances in 42-day steps.
 Forecasting performance is evaluated using both point and probabilistic error measures.
+Forecast error is calculated by considering only the last 42 days and not the entire $2*42$ days.
+This corresponds on how forecasts are generated for planning in practice.
 
-The point forecast accuracy is measured via Root Mean Squared Scaled Error (RMSSE):
+The point forecast accuracy is measured via Root Mean Squared Scaled Error (RMSSE) and Mean Absolute Scaled Error (MASE) that are describes below:
 
 $$
-\text{RMSSE} = \sqrt{\text{mean}(q_{j}^2)},
+RMSSE = \sqrt{mean(q_{j}^2)},
 $$ {#eq-RMSSE}
@@
 $$
-q^2_{j} = \frac{\displaystyle e^2_{j}}
-    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
+q^2_{j} = \frac{ e^2_{j}}
+    {\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
 $$
-
 $e_{j}$ is the point forecast error $j$ and $m = 1$ for non-seasonal series and $m = 7$ for daily seasonal series, $y_t$ is the observation for period $t$ and $T$ is the sample size (observations used for training the forecasting model).
@@
 Note that the measure is scale-independent, thus allowing us to average the results across series.
+
+Mean Absolute Scaled Error (MASE) is calculated as:
+
+$$
+MASE = mean(|q_{j}|),
+$$
+
+where,
+$$
+q_{j} = \frac{ e_{j}}
+    {\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|}.
+$$
+
+The denominator is the mean absolute error of the naive method in the fitting sample of $n$ observations and is used to scale the error.
 

### 2022-11-16 manuscript/main.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Annals of Emergency Medicine}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Hierarchical time series forecasting in Emergency Medical Services},
  pdfauthor={Bahman Rostami-Tabar; Rob J. Hyndman},
  pdfkeywords={forecasting, healthcare, emergency services, forecast
reconciliation, hierarchical time series, ambulance demand, attended
incidents},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Hierarchical time series forecasting in Emergency Medical
Services}
\author[1]{Bahman Rostami-Tabar%
\corref{cor1}%
}
 \ead{rostami-tabarb@cardiff.ac.uk} 
\author[2]{Rob J. Hyndman%
%
}
 \ead{Rob.Hyndman@monash.edu}

\affiliation[1]{organization={Cardiff University, Cardiff Business
School},country={United Kingdom},countrysep={,},postcode={CF10
3EU},postcodesep={}}
\affiliation[2]{organization={Monash University, Department of
Econometrics and Business
Statistics},country={Australia},countrysep={,},postcode={VIC
3800},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
Accurate forecasts of ambulance demand are crucial inputs when planning
and deploying staff and fleet. Such demand forecasts are required at
national, regional and sub-regional levels, and must take account of the
nature of incidents and their priorities. These forecasts are often
generated independently by different teams within the organization. As a
result, forecasts at different levels may be inconsistent, resulting in
conflicting decisions and a lack of coherent coordination in the
service. To address this issue, we exploit the hierarchical and grouped
structure of the demand time series, and apply forecast reconciliation
methods to generate both point and probabilistic forecasts that are
coherent and use all the available data at all levels of disaggregation.
The methods are applied to daily incident data from the Welsh Ambulance
Service NHS Trust, from October 2015 to July 2019, disaggregated by
nature of incident, priority, managing health board, and control area.
We use an ensemble of forecasting models, and show that the resulting
forecasts are better than any individual forecasting model. We validate
the forecasting approach using time-series cross-validation.
\end{abstract}

\begin{keyword}
    forecasting \sep healthcare \sep emergency services \sep forecast
reconciliation \sep hierarchical time series \sep ambulance demand \sep 
    attended incidents
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{sec-intro}{%
\section{Introduction}\label{sec-intro}}

Inability to match the resources with the demand in Emergency Medical
Services (EMS) results in an overcrowding care system. This is a serious
problem causing challenging situations on patient flow with serious
consequences on patients, staff and the entire care system
\citep{ekstrom2015forecasting, ROSTAMITABAR20221197}. Demand forecasting
in EMS is a vital element that helps to depict various courses of action
to avoid the mismatch, which can result in massive savings in terms of
patient safety and lives. An accurate daily demand forecasting enables
planners and decision makers to manage resources to meet anticipated
patients, reconfigure units, redeploy staff and vehicles, where
necessary.

Demand forecasts at EMS are typically required at multiple levels of
cross-sectional granularities to inform various planning and
decision-making processes \citep{hulshof2012taxonomic}. There are some
planning process at the national level (more strategic or long-term)
such as workforce resource planning and budgeting; sub-national,
regional or healthcare level (tactical or medium-term) such as temporary
capacity expansions, resource sharing and staff-shift scheduling; and
hospital or station level (operational or short-term) such as planning
rosters for staff and ambulance deployment. Demand forecasts might also
be required at different level for a specific group of interest such as
nature of demand or priority. Moreover, the time series data in EMS has
an inherent hierarchical and grouped structure to support such
forecasting requirements. Demand for emergency medical services at the
country level can be disaggregated in a geographical hierarchy into
sub-national, regions, health boards, stations/hospitals or divided into
groups such as nature of incidents or demand priority. Therefore, using
forecasting methodologies that account for hierarchical and/or grouped
structures of time series in EMS seems to be a natural fit.

However, despite a large number of studies dedicated to forecasting for
EMS
\citep{mingliterature2022, gul2020exhaustive, ibrahim2016modeling, wargon2009systematic},
this area is neglected and the main focus has been on producing
forecasts at a single level, independently. Generating independent
forecasts not only ignore the inherent hierarchical and/or grouped
structure relationships of the time series demand but also results in a
lack of consistency and coordination. Consistent forecasts of the EMS
demand across all hierarchical and grouped levels are paramount for an
effective planning and decision making. The hierarchical forecasting
approaches can not only create consistent forecasts but also achieve
more accurate forecasts than the independent (base) forecasts
\citep{hyndman2011optimal}. Obtaining consistent forecasts at different
levels is important as it helps to avoid making conflicting decisions.
With hierarchical forecasting, plans at any level are based on coherent
forecasts and therefore can be aligned. Implementing and sustaining
improvements in EMS require alignments and coordination between
different stakeholders, without which teams operate in isolation leading
to conflicts, duplication work, rework, or work that runs counter to the
overall goal to improve the quality of delivery service. Hierarchical
forecasting framework can be used as a tool to improve coordination
between teams across the care services at the national, sub-national,
regional and local levels. To our knowledge, there is not only no
research in the EMS forecasting to account for the hierarchical and
grouped structure of the system but also in the entire field of
forecasting for healthcare management.

In this paper, we address this gap by investigating the application of
hierarchical forecasting approaches in the EMS using daily time series
of verified incidents from 2015 to 2020 in a major ambulance service in
Great Britain. The data has hierarchical and grouped structures with
hierarchies at the national, control (i.e.~sub-national), health board
(i.e.~regional) and groups by priority and nature of incidents. We
produce not only the point forecast but also the forecast distribution
across all levels, which is critical for an effective planning and
associated risk management. We compare the point and probabilistic
forecast accuracy of the independent forecasts, bottom-up and optimal
reconciliation approaches. We first generate independent/based forecasts
using Exponential Smoothing State Space (ETS), Generalized Linear Model
(GLM), Poisson regression, a simple empirical distribution and an
ensemble method followed by applying bottom-up and optimal
reconciliation approaches. Forecast performance is assessed by the Root
Mean Squared Scaled Error (RMSSE) for point forecasts and the Continuous
Ranked Probability Score (CRPS) for the probabilistic forecasts. This
paper complies with the principles of the reproducibility
\citep{stodden2013best, boylan2015reproducibility}. Therefore, the study
could equally be applied to any healthcare service (e.g.~emergency
department, primary or social care) subject to the time series having a
hierarchical and/or grouped structure, which is generally the case in
the healthcare sector.

The remainder of this article is structured as follows: In section 2, we
provide a brief review of the literature and discuss its limitation to
position our work; in section 3, we present the experiment design
describing the data set, forecasting methods and forecast evaluation
metrics. In Section 4, we discuss the hierarchical time series
forecasting approaches to generate both point and probabilistic
forecasts. In section 5, we present and discuss our results; in section
7, we summarize our findings and present ideas for future research.

\hypertarget{lit}{%
\section{Research background}\label{lit}}

Emergency medical services (EMS) are a critical component in the
delivery of urgent medical care to communities. An effective service
delivery requires accurate resource planning that are generally replying
on demand forecasts at operational, tactical and strategic levels.

There is a substantial number of studies on the application of time
series forecasting in the Emergency Medical Services. Various areas have
been the focus of the literature. Forecasting call volume arrivals is
one of the major research topics. \citet{ibrahim2016modeling} provide an
extensive review of the forecasting models in this context. Another
important area is related to forecasting ambulance demand. Although the
definition of demand might not be always clearly stated, however, this
is typically referring to a situation where a physical resource has been
deployed to respond to an incident. This might be also called
\emph{attended incidents}. Another demand related variable is verified
incidents. These are all incidents that require an action: either send a
physical vehicle, deal with via the Clinical Support Desk (e.g.~calls),
get an external (private) provider to respond to it, or send it through
to other channels such as police, firefighters or general practitioners.
Our study is aligned with this stream of the literature. Another similar
area that is largely studied in the literature, is Emergency Department
forecasting. We refer interested readers to \citet{mingliterature2022},
\citet{gul2020exhaustive} and \citet{wargon2009systematic} for extensive
reviews of the literature on Emergency Department forecasting. Although
crucial to EMS performance, \citet{aringhieri2017emergency} state that
demand forecasting has received limited research attention in the EMS
context. In this section, we provide a brief review of studies on
forecasting ambulance demand in EMS.

There are generally two main streams of research related to forecasting
ambulance demand in EMS: i) the first stream focuses on the application
of time series methods and regression approaches on forecasting
aggregate ambulance demand \citep{vile2012predicting, sasaki2010using};
ii) the second stream considers forecasting EMS demand in a more finer
temporal and geographical granularities by employing temporal-spatial
prediction methods \citep{zhou2016predicting, zhou2016predictinglit}.
The focus of our study is related to the first stream of research.

\citet{sasaki2010using} develop a multivariable regression model to
estimate future EMS demands. In addition to the historical demand, the
population census for different age groups and counts of the number of
companies employing more than five people are included in the
regression. The census variables describe groups who may be more likely
to need an ambulance. A stepwise ordinary least squares regression
analysis with SPSS is used for estimating the parameter and generating
forecast. The only performance measure reported in this study is
\(R^2\). The research design of this study is not rigorous and the study
is not reproducible. \citet{vile2012predicting} explore using a Singular
Spectrum Analysis (SSA) method to generate forecasts of the EMS demand
at the national level for 7-day, 14-day, 21-day and 28-day forecast
horizons using data provided by the Welsh Ambulance Service Trust
(WAST). The performance of this approach is compared with
Auto-Regressive Moving Average (ARIMA) and Holt-winter time series
methods using Root Mean Squared Error (RMSE). They concluded that point
forecasts generated by SSA are more accurate for longer-term, but ARIMA
and Holt-winter performance is superior for shorter-term horizons.
\citet{vile2016time} further develop a decision support system to
integrate forecasts generated by SSA. However, the study does not
compare and contrast the performance of forecasting methods based on
utility measures such as cost, resource utilization or response time.
The tool contains options that allow generating forecasts at various
levels of granularity, however, it ignore the hierarchical and grouped
relationships structure, preventing aligned decision making and
coordination.

\citet{9659837} investigate forecasting EMS demand in a high
spatio-temporal resolution of 1×1km spatial regions and 1-hr time
intervals using total incidents in Oslo, Norway from January 1st, 2015
up to and including February 11th, 2019. They used multi-layer
perceptron (MLP) and long short-term memory (LSTM) models to forecast
the EMS demand, and compare them to simple aggregation methods and
baselines. The point forecast accuracy is evaluated using Mean Absolute
Error (MAE) and Mean Squared Error (MSE) and the forecast distribution
is measures by Categorical Cross-Entropy. They shows that while Neural
Network models perform better in producing point forecast, a
distribution baseline method based on spatial distribution of the
incidents across all time steps provide more accurate forecast
distribution. \citet{zhou2016predictinglit} propose three methods based
on Gaussian mixture models, kernel density estimation, and kernel
warping to predict 4 weeks into future for a 1-km2 spatial region over
an hour. Two years of incidents attended from Toronto, Canada (years
2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011
and 2012 with 696,975 events) are used to build the model and examine
the performance on test data using mean negative log likelihood. They
show that forecasts generated by the proposed methods are significantly
more accurate than the current industry practice, a simple averaging
formula. \citet{grekousis2019will} investigate the combination of
spatial analysis methods with data mining techniques based on an
improved Hungarian algorithm and MLP neural network to identify the most
likely locations of future emergency events. The proposed approach is
tested using data of 2851 events attended by the EMS in Athens, Greece
over 24 weeks. They show that \(23.24\%\) of real emergency events lie
within 50 meter of the predicted ones and nearly \(70\%\) of the real
emergency events lie no further away than 150 meter, which is rather
accurate given the granularity of the problem at the city level.

We observed a number of limitations in the literature of EMS
forecasting, that encourage us to undertake this research. These
limitations are summarized as following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Current studies ignore the inherent hierarchical and/or grouped
  structure of the time series data and the relationship between series
  at different levels of hierarchy. This may result in incoherent
  forecasts leading to misaligned planning and decision making. While
  the hierarchical forecasting methodology has been developed and
  applied in various domains over the past 10 years
  \citep{panagiotelis2022probabilistic}, it has never been explored in
  this area.
\item
  Current research is mainly concerned with generating point forecast at
  a single level of hierarchy. There is a lack of studies presenting the
  entire forecast distribution of daily ambulance demand for the entire
  hierarchy to inform the whole decision-making process and to better
  represent the uncertainty of future demand, providing a risk
  management tool for planners.
\item
  Reproducibility is still a major challenge in EMS forecasting, as it
  is unlikely to reproduce the results without the help of the authors
  of those papers.
\item
  Another limitation is related to the generated forecasts that are not
  integer counts. Since actual ambulance counts cannot be negative or
  fractions, ambulance demand forecasts should be the same. While this
  might not be an issue when producing forecasts at a single level,
  producing non-negative count forecasts in a hierarchical/grouped
  structure is challenging and requires further investigation in the
  future.
\end{enumerate}

This paper concerns the problem of hierarchical forecasting in EMS and
generates and evaluates both point and probabilistic forecast across
different levels of the hierarchy, hence addressing some important gaps
identified in the literature.

\hypertarget{sec-experiment}{%
\section{Experiment setup}\label{sec-experiment}}

We are interested in generating forecasts to inform the planning horizon
of \(ph= 42\) days, interested by planners in the ambulance services
trust. The forecast horizon in this study is \(fh = 2 \times ph\) days
ahead (2*42 days planning horizon). This is because the planning is
generally frozen for \(ph\) days and considering a forecast horizon of
\(ph\) days might not be helpful for planning. While forecasts are
generated for \(2 \times ph\) days ahead, performance evaluation is only
assessed based on the last \(ph\) days and not the \(2 \times ph\) days.
The forecasts are produced for the holdout of 365 days using time series
cross-validation \citep{hyndman2021forecasting}. In the following
section, we discuss the dataset, describe the forecasting methods used
to generate base forecasts and present the point and probabilistic
accuracy measures.

\hypertarget{sec-data}{%
\subsection{Data}\label{sec-data}}

The dataset used in this study is from a major ambulance service trust
in the United Kingdom. It contains information relating to the daily
number of attended incidents from 10 October 2015 to 31 July 2019 by
nature of incidents, priority, the health board managing the service and
the control area (or region). Figure Figure~\ref{fig-hierarchy} depicts
both the hierarchical and grouped structure of the data. Figure
Figure~\ref{fig-hierarchy-1} illustrates the nested hierarchical
structure based on control area and health board and Figure
Figure~\ref{fig-hierarchy-2} shows the grouped structure by priority and
the nature of incident.

\begin{figure}

\begin{minipage}[t]{0.56\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{main_files/figure-pdf/fig-hierarchy-1.pdf}

}

}

\subcaption{\label{fig-hierarchy-1}Hierarchical structure: Attended
incidents in all country is disaggregated into 3 control areas /regions
and then into 7 different healthboard at the bottom}
\end{minipage}%
%
\begin{minipage}[t]{0.02\linewidth}

{\centering

~

}

\end{minipage}%
%
\begin{minipage}[t]{0.42\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{../img/group.png}

}

}

\subcaption{\label{fig-hierarchy-2}Grouped structure: Incidents could be
grouped into priority (i.e.~Red, Amber \& Green) and the nature of
attended incident (i.e.~there are 35 different nature of incidents
including chest pain , breathing problems, heart attack, stroke, and so
on). The symbol * refers to the crossed attributes between hierarchical
and grouped levels.}
\end{minipage}%

\caption{\label{fig-hierarchy}The hierarchical and grouped structure of
attended incidents (ambulance demand)}

\end{figure}

Table Table~\ref{tbl-hierarchy} also displays the structure of data with
the total number of series at each level. At the top level, we have the
total attended incidents for the country. We can split these total
attended incidents by control area, by health board, by priority or by
nature of incident. There are 3 control areas breakdown by 7 local
health boards. Attended incident data are categorized into 3 priority
classes of red, amber and green. There are also 35 different nature of
incidents such as chest pain, stroke, breathing problem, etc. In total,
across all levels of disaggregation, there are 1530 time series.

\hypertarget{tbl-hierarchy}{}
\begin{table}
\caption{\label{tbl-hierarchy}Number of time series in each level for the hierarchical \& grouped
structure of attended incidents }\tabularnewline

\centering
\begin{tabular}{lr}
\toprule
Level & Number of series\\
\midrule
All country & 1\\
Control & 3\\
Health board & 7\\
Priority & 3\\
Priority * Control & 9\\
\addlinespace
Priority * Health board & 21\\
Nature of incident & 35\\
Nature of incident * Control & 105\\
Nature of incident * Health board & 245\\
Priority * Nature of incident & 104\\
\addlinespace
Control * Priority * Nature of incident & 306\\
Control * Health board * Priority * Nature of incident (Bottom level) & 691\\
Total & 1530\\
\bottomrule
\end{tabular}
\end{table}

Given the total number of time series, direct visual analysis is
infeasible. Therefore, we first compute features of all 1530 time series
and display the strength of trend and weekly seasonality strength in
Figure Figure~\ref{fig-feature}. Each point represents one time series
with strength of trend in x-axis and strength of seasonality in y-axis.
It is clear that there are some series showing strong trends and/or
seasonality, corresponding to series at the higher levels of the
hierarchy. The majority show low trend and seasonality with . These are
time series belonging to the bottom of the structure, series related to
the nature of incidents for each given control, health board and
priority.

\begin{figure}

{\centering \includegraphics[width=0.7\textwidth,height=\textheight]{main_files/figure-pdf/fig-feature-1.pdf}

}

\caption{\label{fig-feature}Time series features of attended incidents
across all levels (1530 series)}

\end{figure}

In addition to displaying the trend and seasonality features, we also
highlight few time series at various levels of the aggregation. Figure
Figure~\ref{fig-dataviz} reveals different information such as trend,
seasonality and noise. For example, some series depict seasonality and
trend, whereas some other series report low volume of attended incidents
and entropy, making them more volatile and difficult to forecast. At the
level on nature of incidents combined with categories of other levels,
there are many series that contain zeros with low counts. As such, the
data set represents a diverse set of daily time series patterns.

\begin{figure}

\begin{minipage}[t]{0.49\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{main_files/figure-pdf/fig-dataviz-1.pdf}

}

}

\subcaption{\label{fig-dataviz-1}Whole country}
\end{minipage}%
%
\begin{minipage}[t]{0.02\linewidth}

{\centering

~

}

\end{minipage}%
%
\begin{minipage}[t]{0.49\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{main_files/figure-pdf/fig-dataviz-2.pdf}

}

}

\subcaption{\label{fig-dataviz-2}By control area}
\end{minipage}%
\newline
\begin{minipage}[t]{0.49\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{main_files/figure-pdf/fig-dataviz-3.pdf}

}

}

\subcaption{\label{fig-dataviz-3}By health board and control area}
\end{minipage}%
%
\begin{minipage}[t]{0.02\linewidth}

{\centering

~

}

\end{minipage}%
%
\begin{minipage}[t]{0.49\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{main_files/figure-pdf/fig-dataviz-4.pdf}

}

}

\subcaption{\label{fig-dataviz-4}By health board and priority}
\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{main_files/figure-pdf/fig-dataviz-5.pdf}

}

}

\subcaption{\label{fig-dataviz-5}By nature of incident and health board}
\end{minipage}%

\caption{\label{fig-dataviz}Time series of attended incidents at various
levels.}

\end{figure}

\hypertarget{forecasting-methods}{%
\subsection{Forecasting methods}\label{forecasting-methods}}

Given the presence of various significant patterns in the past attended
incidents, we consider three different forecasting models to generate
the base forecasts. Once the base forecasts are produced, hierarchical
and grouped time series methods are used to reconcile them across the
all levels. We briefly discussed forecasting models in the following
sections, and the hierarchical forecasting methods are discussed in
Section~\ref{sec-htc}.

\textbf{Naive:} We start with one of the simplest forecasting approaches
used in practice - assuming that in the next few days, everything will
be the same as in the similar days in the past. This is called
``Naïve''. In our case, given that we need a distribution of values, we
will use a modified approach, where the empirical distribution of the
daily attended incidents time series is used to forecast the future
attended incidents distribution. We consider the empirical distribution
of the most recent year of historic data on a rolling basis to capture
potential changes in behavior over time.

\textbf{Exponential Smoothing State Space model (ETS):} ETS models
\citep{hyndman2021forecasting} can combine trend, seasonality and error
components in a time series through various forms such as additive,
multiplicative or mixed. The trend component can be none (``N''),
Additive (``A''), damped (``Ad'') or multiplicative (``M''). The
seasonality can be none (``N''), Additive (``A''), or multiplicative
(``M''). The error term can also be additive (``A'') or multiplicative
(``M''). To forecast the attended incidents at each level, we use the
\texttt{ETS()} function in the \texttt{fable} package \citep{fable2022}
in R. To identify the best model for a given time series, the ETS
function uses the corrected Akaike's Information Criterion (AICc).

\textbf{Poisson Regression: } Despite the popularity and the relevant of
automatic \texttt{ETS} and \texttt{ARIMA} in this study, but they
produce non-integer attended incidents and forecasts might also be
negative. However, the number of attended incidents is an integer and
non-negative. When using \texttt{ETS} and \texttt{ARIMA}, a time series
transformation approach such as \texttt{sqrt()} could be used to
generate strictly positive forecasts, however forecasts are still not
integer. Another alternative is to use forecasting models that produce
integer, non-negative forecasts. One of the models that is frequently
used in practice is the Poisson regression \citep[see, for
example,][]{mccarthy2008challenge}, which can be summarized as
\begin{equation}\protect\hypertarget{eq-poisson-regression}{}{
  \begin{aligned}
    & {y}_t \sim \mathrm{Poisson} \left( \lambda_t \right) \\
    & \log \lambda_t = \mathbf{x}_t' \boldsymbol{\beta}
  \end{aligned} \quad
}\label{eq-poisson-regression}\end{equation}

The logarithm in Equation~\ref{eq-poisson-regression} is needed to
ensure that the parameter of Poisson distribution is always positive.
This model can be estimated via maximization of the likelihood function
based on Poisson mass function. For taking into account serial
dependence we include a regression on the previous observation. We try
to capture the short range serial dependence by a first order
autoregressive term and the weekly seasonality by a 7th order
autoregressive term. We use the \texttt{tscount()} function in the fable
package in R, which is a wrapper function written based on
\texttt{tscount} package. We fit a model to each time series using the
function tsglm, with the identity link function, defined by the argument
link.

\hypertarget{performance-evaluation}{%
\subsection{Performance evaluation}\label{performance-evaluation}}

Forecasting performance is evaluated using both point and probabilistic
error measures.

The point forecast accuracy is measured via Root Mean Squared Scaled
Error (RMSSE):

\begin{equation}\protect\hypertarget{eq-RMSSE}{}{
\text{RMSSE} = \sqrt{\text{mean}(q_{j}^2)},
}\label{eq-RMSSE}\end{equation}

where,

\[
q^2_{j} = \frac{\displaystyle e^2_{j}}
    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
\]

\(e_{j}\) is the point forecast error \(j\) and \(m = 1\) for
non-seasonal series and \(m = 7\) for daily seasonal series, \(y_t\) is
the observation for period \(t\) and \(T\) is the sample size
(observations used for training the forecasting model). Smaller RMSSE
values suggest more accurate forecasts. Note that the measure is
scale-independent, thus allowing us to average the results across
series.

To measure the forecast distribution performance, we calculate the
Continuous Rank Probability Score \citep{gneiting2014probabilistic}. It
rewards sharpness and penalizes miscalibration, so it measures overall
performance.

\begin{equation}\protect\hypertarget{eq-CRPS}{}{
CRPS=\frac{1}{h} \sum_{j=1}^{h} \int_{-\infty}^{\infty} \left(F^f_j(x) - F^0_j(x)\right)^2dx
}\label{eq-CRPS}\end{equation}

where \(F^f_j(x)\) is the forecasted Cumulative Density Function (CDF)
of period \(j\) and 𝐹\(F^0_j(x)\) is the true CDF of period \(j\).

\hypertarget{sec-htc}{%
\section{Hierarchical and grouped time series forecasting
techniques}\label{sec-htc}}

There are many applications in the healthcare and in particular in EMS
where a collection of time series is available. These series are
generally hierarchically organized based on multiple levels such as
area/region, health board and/or are aggregated at different levels in
groups based on nature of demand, priority of demand or some other
attributes. While series could be strictly hierarchical or only grouped
bases on some attributes, in many situation a more complex structures
arise when attributes of interest are both nested and crossed, having
hierarchical and grouped structure. This is also the case as discussed
in Section~\ref{sec-data}

\hypertarget{independent-base-forecast}{%
\subsection{Independent (base
forecast)}\label{independent-base-forecast}}

A common practice in healthcare (and EMS) to predict hierarchical and
grouped series relies on producing independent forecast, also refereed
to as base forecast, typically by different teams as the need for such
forecasts arise. This is also known as base forecast. We observe \(n\)
time series at time \(t\), across the entire hierarchical and grouped
structure, written as \(y_t\). The base forecasts of \(y_{T+h}\) given
data \(y_1,\dots,y_T\) are denoted by \(\hat{y}_h\) for h steps-ahead
for all \(n\) series (\(n=1530\) in this study). Forecasts generated in
this way are nor coherent.

\hypertarget{reconciliation-methos}{%
\subsection{Reconciliation methos}\label{reconciliation-methos}}

Traditionally, alternative approaches to produce coherent forecasts for
hierarchical and grouped time series involves using bottom-up and
top-down methods by generating forecasts at a single level and then
aggregate or disaggregate. Top-down requires having a unique structure
to disaggregated forecasts generated at top level by proportions.
However, given that we have multiple grouped attributes combined with
the hierarchical structure, there is no unique way to disaggregate top
forecast. Hence the top-down can not be used in this case, so either we
can do some kind of reconciliation or must define our own top-down
method for each hierarchy. The recommended approach is to use
reconciliation. In the following sections, we first discuss some
notations and then present bottom-up and forecast reconciliation
approach used in this study to generate coherent forecasts.

\hypertarget{notations}{%
\subsubsection{Notations}\label{notations}}

Let \(b_t\) be a vector of \(n_b\) ``bottom-level'' time series at time
\(t\), and let \(a_t\) be a corresponding vector of \(n_a = n-n_b\)
aggregated time series, where

\[
 a_t = Ab_t
\]

and \(A\) is the \(n_a\times n_b\) ``aggregation'' matrix specifying how
the bottom-level series \(b_t\) are to be aggregated to form \(a_t\).
The full vector of time series is given by \[
 y_t = \begin{bmatrix}a_t \\b_t\end{bmatrix}.
\] This leads to the \(n\times n_b\) ``summing'' or ``structural''
matrix given by \[
  S = \begin{bmatrix}A \\ I_{n_b}\end{bmatrix}
\] such that \(y_t = Sb_t\).

\hypertarget{bottom-up-bu-and-linear-reconciliation-methods}{%
\subsubsection{Bottom-up (BU) and linear reconciliation
methods}\label{bottom-up-bu-and-linear-reconciliation-methods}}

Bottom-Up is a simple approach to generate coherent forecasts. It first
involves creating the base forecasts for the bottom level series. These
forecasts are then aggregated to the upper levels which results in
generating coherent forecasts. BU approach can capture the dynamics of
the series at the bottom level, but they may be noisy and difficult to
forecast as well. BU approach is limited on using only forecasts at the
bottom level and does not utilize all the information available across
the hierarchical and grouped structure. Forecast reconciliation
approaches fill this gap and combine and reconcile all the base
forecasts in order to produce coherent forecasts.

Given the summing matrix and base forecasts, bottom-up and linear
reconciliation methods can be written as \(\tilde{y}_h = SG\hat{y}_h\)
for different matrices \(G\).

Optimal reconciled forecasts are obtained with
\(G=(S'W^{-1}S)^{-1}W^{-1}\), or \(\tilde{y}_h = M\hat{y}_h,\) where the
\(n\times n\) ``mapping'' matrix is given by
\(M = S(S'W^{-1}S)^{-1}W^{-1},\) where \(\hat{y}_h\) are the \(h\)-step
forecasts of \(y_{T+h}\) given data to time \(T\), and \(W\) is an
\(n \times n\) positive definite matrix. Different choices for \(W\)
lead to different solutions such as Ordinary Least Square (OLS) ,
Weighted Least Square (WLS) and Minimum Trace (MinT)
\citep{WicEtAl2019}. We use the implementation of these methods in the
fable package in R in the experiment.

\hypertarget{sec-result}{%
\section{Results and discussion}\label{sec-result}}

\hypertarget{point-forecast}{%
\subsection{Point forecast}\label{point-forecast}}

\hypertarget{probabilistic-forecast}{%
\subsection{Probabilistic forecast}\label{probabilistic-forecast}}

\hypertarget{sec-conclusion}{%
\section{Conclusion}\label{sec-conclusion}}

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}

### 2022-11-16 manuscript/03-experiment.qmd
render to pdf

--- 2022-11-16 manuscript/03-experiment.qmd
@@
 
-We are interested in generating forecasts to inform the planning horizon of $ph= 42$ days, interested by planners in the ambulance services trust.
+We are interested in generating forecasts to inform the planning horizon of $ph= 42$ days, required by planners in the ambulance services trust.
 The forecast horizon in this study is $fh = 2 \times ph$ days ahead (2*42 days planning horizon).
-This is because the planning is generally frozen for $ph$ days and considering a forecast horizon of $ph$ days might not be helpful for planning.
+This is because the planning is generally freezed for $ph$ days and considering a forecast horizon of $ph$ days might not be helpful for planning.
 While forecasts are generated for $2 \times ph$ days ahead, performance evaluation is only assessed based on the last $ph$ days and not the $2 \times ph$ days.
 The forecasts are produced for the holdout of 365 days using time series cross-validation [@hyndman2021forecasting].
+
 In the following section, we discuss the dataset, describe the forecasting methods used to generate base forecasts and present the point and probabilistic accuracy measures.
@@
 #| fig-cap: "Time series features of attended incidents across all levels (1530 series)"
-incident_gthf <- readr::read_rds(here::here("data/incidents_gt.rds"))
+incident_gthf <- readr::read_rds(paste0(storage_folder, "incidents_gt.rds"))
 incident_gthf %>%
@@
 
+The data is spitted into training (up until 2018-07-31) and test (from  2018-08-01 to the end) sets, with all model development and hyper-parameter tuning performed using training data only.
+The time series cross validation is performed with a forecast horizon of $2*42$ days and advances in 42-day steps.
 Forecasting performance is evaluated using both point and probabilistic error measures.
-
-The point forecast accuracy is measured via Root Mean Squared Scaled Error (RMSSE):
+Forecast error is calculated by considering only the last 42 days and not the entire $2*42$ days.
+This corresponds on how forecasts are generated for planning in practice.
+Forecasting performance is evaluated using both point and probabilistic error measures.
+The point forecast accuracy is measured via Root Mean Squared Scaled Error (RMSSE) and Mean Absolute Scaled Error (MASE) that are describes below:
 
 $$
-\text{RMSSE} = \sqrt{\text{mean}(q_{j}^2)},
+RMSSE = \sqrt{mean(q_{j}^2)},
 $$ {#eq-RMSSE}

### 2022-11-16 manuscript/main.tex
render to pdf

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \makeatletter
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, breakable, enhanced, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 We are interested in generating forecasts to inform the planning horizon
-of \(ph= 42\) days, interested by planners in the ambulance services
+of \(ph= 42\) days, required by planners in the ambulance services
 trust. The forecast horizon in this study is \(fh = 2 \times ph\) days
 ahead (2*42 days planning horizon). This is because the planning is
-generally frozen for \(ph\) days and considering a forecast horizon of
+generally freezed for \(ph\) days and considering a forecast horizon of
 \(ph\) days might not be helpful for planning. While forecasts are
@@
 The forecasts are produced for the holdout of 365 days using time series
-cross-validation \citep{hyndman2021forecasting}. In the following
-section, we discuss the dataset, describe the forecasting methods used
-to generate base forecasts and present the point and probabilistic
-accuracy measures.
+cross-validation \citep{hyndman2021forecasting}.
+
+In the following section, we discuss the dataset, describe the
+forecasting methods used to generate base forecasts and present the
+point and probabilistic accuracy measures.
 
@@
 The dataset used in this study is from a major ambulance service trust
-in the United Kingdom. It contains information relating to the daily
-number of attended incidents from 10 October 2015 to 31 July 2019 by
-nature of incidents, priority, the health board managing the service and
-the control area (or region). Figure Figure~\ref{fig-hierarchy} depicts
-both the hierarchical and grouped structure of the data. Figure
+in Great Britain. It contains information relating to the daily number
+of attended incidents from 10 October 2015 to 31 July 2019 by nature of
+incidents, priority, the health board managing the service and the
+control area (or region). Figure Figure~\ref{fig-hierarchy} depicts both
+the hierarchical and grouped structure of the data. Figure
 Figure~\ref{fig-hierarchy-1} illustrates the nested hierarchical
@@
 
-\includegraphics{../img/group.png}
+\includegraphics{/Users/bahmanrostami-tabar/Documents/1-research/1-research_papers/Work with Rob Hyndman/forecasting-emergency-medcine/img/group.png}
 
@@
 Figure Figure~\ref{fig-feature}. Each point represents one time series
-with strength of trend in x-axis and strength of seasonality in y-axis.
-It is clear that there are some series showing strong trends and/or
-seasonality, corresponding to series at the higher levels of the
-hierarchy. The majority show low trend and seasonality with . These are
-time series belonging to the bottom of the structure, series related to
-the nature of incidents for each given control, health board and
-priority.
+with the strength of trend in x-axis and the strength of seasonality in
+y-axis. It is clear that there are some series showing strong trends
+and/or seasonality, corresponding to series at the higher levels of the
+hierarchy. The majority of series show low trend and seasonality. These
+are time series belonging to the bottom series, series related to the
+nature of incidents for a given control, health board and priority
+level. Bottom series are dominated by noise with little or no systematic
+patterns.
 
@@
 In addition to displaying the trend and seasonality features, we also
-highlight few time series at various levels of the aggregation. Figure
+visualize few time series at various levels of the aggregation. Figure
 Figure~\ref{fig-dataviz} reveals different information such as trend,
@@
 
+We consider several forecasting models that account for the diverse
+patterns of the time series across the entire hierarchy. In developing
+the forecasting models, the time series of holidays are also used in
+addition to the attended incidents. We use public holidays, school
+holidays and Christmas Day and New Year's Day as predictors of incident
+attended. These type of holidays will affect peoples' activities and may
+increase or decrease the number of attended incidents.
+
 \hypertarget{forecasting-methods}{%
@@
 used in practice - assuming that in the next few days, everything will
-be the same as in the similar days in the past. This is called
-``Naïve''. In our case, given that we need a distribution of values, we
-will use a modified approach, where the empirical distribution of the
-daily attended incidents time series is used to forecast the future
-attended incidents distribution. We consider the empirical distribution
-of the most recent year of historic data on a rolling basis to capture
+be the same as similar days in the past. This is called ``Naïve''. In
+our case, given that we need a distribution of values, we will use a
+modified approach, where the empirical distribution of the daily
+attended incidents time series is used to forecast the future attended
+incidents distribution. We consider the empirical distribution of the
+most recent year of historic data on a rolling basis to capture
 potential changes in behavior over time.
@@
 
-\textbf{Poisson Regression: } Despite the popularity and the relevant of
-automatic \texttt{ETS} and \texttt{ARIMA} in this study, but they
-produce non-integer attended incidents and forecasts might also be
+Despite the popularity and the relevance of automatic \texttt{ETS} in
+this study, however it may produce forecasts that are non-integer and
 negative. However, the number of attended incidents is an integer and
-non-negative. When using \texttt{ETS} and \texttt{ARIMA}, a time series
-transformation approach such as \texttt{sqrt()} could be used to
-generate strictly positive forecasts, however forecasts are still not
-integer. Another alternative is to use forecasting models that produce
-integer, non-negative forecasts. One of the models that is frequently
-used in practice is the Poisson regression \citep[see, for
-example,][]{mccarthy2008challenge}, which can be summarized as
-\begin{equation}\protect\hypertarget{eq-poisson-regression}{}{
-  \begin{aligned}
-    & {y}_t \sim \mathrm{Poisson} \left( \lambda_t \right) \\
-    & \log \lambda_t = \mathbf{x}_t' \boldsymbol{\beta}
-  \end{aligned} \quad
-}\label{eq-poisson-regression}\end{equation}
+non-negative. When using \texttt{ETS} , a time series transformation
+approach could be used to generate strictly positive forecasts, however
+forecasts are still not integer. An alternative is to use forecasting
+models that produce integer, non-negative forecasts. in the following
+section we present Generalized Linear Models (GLMs) and Poisson
+Regression to produce count base forecasts.
 
-The logarithm in Equation~\ref{eq-poisson-regression} is needed to
-ensure that the parameter of Poisson distribution is always positive.
-This model can be estimated via maximization of the likelihood function
-based on Poisson mass function. For taking into account serial
-dependence we include a regression on the previous observation. We try
-to capture the short range serial dependence by a first order
-autoregressive term and the weekly seasonality by a 7th order
-autoregressive term. We use the \texttt{tscount()} function in the fable
-package in R, which is a wrapper function written based on
-\texttt{tscount} package. We fit a model to each time series using the
-function tsglm, with the identity link function, defined by the argument
-link.
+\textbf{Generalized Linear Model (GLM): } GLMs are a family of models
+developed to extend the concept of linear regression models. They
+perform a regression by modeling the response variable as coming from a
+particular member of the exponential family, and then transforming the
+mean of the response so that the transformed mean is a linear function
+of the predictors. One of the models that is frequently used in practice
+to generate count forecasts is the Poisson regression. We will consider
+forecasting attended incidents using the covariates spline trend, day of
+the week dummy variables (from Monday to Sunday), Fourier terms to
+capture yearly seasonality, public holidays (1 when is public holiday, 0
+otherwise), school holidays (1 when is school holiday, 0 otherwise) and
+Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day
+(1 when is New Year's Day, 0 otherwise). We fit a Poisson regression
+model using the function \texttt{glm()} from the package in R, with the
+argument \texttt{family\ =\ poisson} to specify that we wish to fit a
+Poisson regression model.
+
+\textbf{Poisson Regression using tscount: } We consider another Poisson
+regression model that takes into account serial dependence in addition
+to covariates used in the GLM model. To that end, we use
+\texttt{tsglm()} function in \emph{tscount} package in R to model the
+attended incidents. The logarithmic link function is used to ensure that
+the parameter of Poisson distribution is always positive. This model can
+be estimated via maximization of the likelihood function based on
+Poisson mass function. The regression model captures the short range
+serial dependence by including the three order autoregressive terms.
+
+\textbf{Ensemble method:} we also use an ensemble method that combines
+forecasts generated from Naive, ETS, GLM and Poisson regression together
+to make more accurate forecasts than any individual model.
 
@@
 
-Forecasting performance is evaluated using both point and probabilistic
-error measures.
-
-The point forecast accuracy is measured via Root Mean Squared Scaled
-Error (RMSSE):
+The data is spitted into training (up until 2018-07-31) and test (from
+2018-08-01 to the end) sets, with all model development and
+hyper-parameter tuning performed using training data only. The time
+series cross validation is performed with a forecast horizon of \(2*42\)
+days and advances in 42-day steps. Forecasting performance is evaluated
+using both point and probabilistic error measures. Forecast error is
+calculated by considering only the last 42 days and not the entire
+\(2*42\) days. This corresponds on how forecasts are generated for
+planning in practice. Forecasting performance is evaluated using both
+point and probabilistic error measures. The point forecast accuracy is
+measured via Root Mean Squared Scaled Error (RMSSE) and Mean Absolute
+Scaled Error (MASE) that are describes below:
 
 \begin{equation}\protect\hypertarget{eq-RMSSE}{}{
-\text{RMSSE} = \sqrt{\text{mean}(q_{j}^2)},
+RMSSE = \sqrt{mean(q_{j}^2)},
 }\label{eq-RMSSE}\end{equation}
@@
 \[
-q^2_{j} = \frac{\displaystyle e^2_{j}}
-    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
-\]
-
-\(e_{j}\) is the point forecast error \(j\) and \(m = 1\) for
+q^2_{j} = \frac{ e^2_{j}}
+    {\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
+\] \(e_{j}\) is the point forecast error \(j\) and \(m = 1\) for
 non-seasonal series and \(m = 7\) for daily seasonal series, \(y_t\) is
@@
 series.
+
+Mean Absolute Scaled Error (MASE) is calculated as:
+
+\[
+MASE = mean(|q_{j}|),
+\]
+
+where, \[
+q_{j} = \frac{ e_{j}}
+    {\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|}.
+\]
+
+The denominator is the mean absolute error of the naive method in the
+fitting sample of \(n\) observations and is used to scale the error.
 

### 2022-11-16 manuscript/main.qmd
Added colorblind palette and theme to setup

@@
   warning: false
-  cache: true
+pdf-engine: latexmk
 format:
@@
 library(patchwork)
-#library(ggraph)
-#library(igraph)
+# Color blind color palettes
+options(
+  ggplot2.discrete.colour = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
+  ggplot2.discrete.fill = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
+)
+# Few theme
+ggplot2::theme_set(ggthemes::theme_few())
 ```

### 2022-11-16 manuscript/03-experiment.qmd
Added alternative dataviz graph

@@
 #| out.width: "60%"
-#| fig-cap: "The hierarchical and grouped structure of attended incidents (ambulance demand)"
+#| fig-cap: "The hierarchical and grouped structure of attended incidents (ambulance demand)."
 #| fig-subcap:
@@
 #| fig-cap: "Time series features of attended incidents across all levels (1530 series)"
-incident_gthf <- readr::read_rds(paste0(storage_folder, "incidents_gt.rds"))
+incident_gthf <- readr::read_rds(here::here("data/incidents_gt.rds"))
 incident_gthf %>%
@@
   ggplot(aes(x = trend_strength, y = seasonal_strength_week)) +
-  geom_point(alpha = 0.5) +
-  ggthemes::theme_few() +
+  geom_point(alpha = 0.25) +
   labs(x = "Strength of trend", y = "Strength of weekly seasonality")
@@
 In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation.
-Figure @fig-dataviz reveals different information such as trend, seasonality and noise.
+Figure @fig-dataviz2 reveals different information such as trend, seasonality and noise.
 For example, some series depict seasonality and trend, whereas some other series report low volume of attended incidents and entropy, making them more volatile and difficult to forecast.
@@
 #| label: fig-dataviz
-#| cache: true
+#| eval: false
+#| include: false
 #| dependson: "fig-feature"
@@
 p_nature_healthboard
+```
+
+```{r}
+#| label: fig-dataviz2
+#| dependson: "fig-feature"
+#| out.width: "100%"
+#| fig-width: 8
+#| fig-height: 10
+#| fig-cap: "Time series of attended incidents at various levels."
+no_x_axis <- theme(axis.title.x=element_blank(),
+                   axis.text.x=element_blank(),
+                   axis.ticks.x=element_blank())
+p_total <- incident_gthf %>%
+  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
+  autoplot(incident) +
+  labs(x = "", y = "Incidents") +
+  no_x_axis
+
+p_control <- incident_gthf %>%
+  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
+  as_tibble() %>%
+  select(-nature, -category) %>%
+  group_by(date, region) %>%
+  summarise(incident = sum(incident), .groups = "drop") %>%
+  ggplot(aes(x = date, y = incident, color = factor(region))) +
+  geom_line() +
+  labs(y = "Incidents", color = "Control")  +
+  no_x_axis
+
+p_board <- incident_gthf %>%
+  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
+  as_tibble() %>%
+  select(-nature, -category) %>%
+  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
+  geom_line() +
+  #facet_wrap(vars(factor(region)), scales = "free_y") +
+  labs(y = "Incidents", color = "Health board") +
+  no_x_axis
+
+p_priority <- incident_gthf %>%
+  filter(is_aggregated(region) & is_aggregated(lhb) & !is_aggregated(category) & is_aggregated(nature)) %>%
+  mutate(
+    category = recode(category, RED = "Red", AMB = "Amber", GRE = "Green"),
+    category = factor(category, levels=c("Red","Amber","Green"))
+  ) |>
+  as_tibble() %>%
+  select(-nature, -region) %>%
+  ggplot(aes(x = date, y = incident, color = factor(category))) +
+  geom_line() +
+  scale_color_manual(values = c(Red="#ff3300", Amber = "#E69f00", Green="#009e73")) +
+  labs(y = "Incidents", color = "Priority") +
+  no_x_axis
+
+selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
+p_nature <- incident_gthf %>%
+  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & !is_aggregated(nature)) %>%
+  as_tibble() %>%
+  mutate(nature = as.character(nature)) %>%
+  filter(nature %in% selected_nature) %>%
+  group_by(date, nature, lhb) %>%
+  summarise(incident = sum(incident), .groups = "drop") %>%
+  ggplot(aes(x = date, y = incident, color = nature)) +
+  geom_line() +
+  labs(x = "Date", y = "Incidents", color = "Nature of incident")
+
+p_total /
+p_control /
+p_board /
+p_priority /
+p_nature
 ```

### 2022-11-16 manuscript/05-results.qmd
Added accuracy graph

@@
 ## Probabilistic forecast
+
+```{r}
+#| label: fig-accuracy
+#| fig-width: 8
+#| fig-height: 5
+#| out.width: "100%"
+#| fig-cap: "Average accuracy by week for 12 weeks. CRPS is relative to a naive ECDF. MASE and MSSE are relative to the corresponding values for the training set."
+rmsse <- readr::read_rds(here::here("results/rmsse.rds"))
+mase <- readr::read_rds(here::here("results/mase.rds"))
+crps <- readr::read_rds(here::here("results/crps.rds")) |>
+  pivot_wider(names_from = model, values_from = crps) |>
+  group_by(method, h, series) |>
+  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
+  ungroup() |>
+  pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")
+
+accuracy <- bind_rows(
+  rmsse |> mutate(measure = "msse", accuracy = rmsse^2),
+  mase |> mutate(measure = "mase", accuracy = mase),
+  crps |> mutate(measure = "crps", accuracy = crps),
+) |>
+  select(-rmsse, -mase, -crps)
+
+# Plot of average accuracy vs week for each method for Total
+acc_summary <- accuracy |>
+  filter(
+    series %in% c("Total", "Control areas", "Health boards"),
+    method == "mint", model != "qcomb"
+  ) |>
+  mutate(
+    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
+    model = factor(model, levels = c("naiveecdf", "ets", "tscount", "iglm", "ensemble")),
+    week = factor(trunc((h - 1) / 7) + 1)
+  ) |>
+  group_by(week, model, measure, series) |>
+  summarise(accuracy = mean(accuracy), .groups = "drop")
+
+acc_summary |>
+  ggplot(aes(x = week, y = accuracy, group = model)) +
+  geom_line(aes(col = model)) +
+  facet_grid(measure ~ series, scales = "free_y") +
+  labs(y = "Average accuracy", x = "Week ahead")
+```

### 2022-11-16 manuscript/main.qmd
Back to pdflatex

@@
   warning: false
-pdf-engine: latexmk
+pdf-engine: pdflatex
 format:

### 2022-11-16 manuscript/main.tex
Now compiling again

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, breakable, enhanced, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, boxrule=0pt, frame hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 ahead (2*42 days planning horizon). This is because the planning is
-generally freezed for \(ph\) days and considering a forecast horizon of
+generally frozen for \(ph\) days and considering a forecast horizon of
 \(ph\) days might not be helpful for planning. While forecasts are
@@
 
-\includegraphics{/Users/bahmanrostami-tabar/Documents/1-research/1-research_papers/Work with Rob Hyndman/forecasting-emergency-medcine/img/group.png}
+\includegraphics{../img/group.png}
 
@@
 \caption{\label{fig-hierarchy}The hierarchical and grouped structure of
-attended incidents (ambulance demand)}
+attended incidents (ambulance demand).}
 
@@
 visualize few time series at various levels of the aggregation. Figure
-Figure~\ref{fig-dataviz} reveals different information such as trend,
+Figure~\ref{fig-dataviz2} reveals different information such as trend,
 seasonality and noise. For example, some series depict seasonality and
@@
 
-\begin{minipage}[t]{0.49\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{main_files/figure-pdf/fig-dataviz-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-dataviz2-1.pdf}
 
@@
 
-}
-
-\subcaption{\label{fig-dataviz-1}Whole country}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.02\linewidth}
-
-{\centering
-
-~
-
-}
-
-\end{minipage}%
-%
-\begin{minipage}[t]{0.49\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{main_files/figure-pdf/fig-dataviz-2.pdf}
-
-}
-
-}
-
-\subcaption{\label{fig-dataviz-2}By control area}
-\end{minipage}%
-\newline
-\begin{minipage}[t]{0.49\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{main_files/figure-pdf/fig-dataviz-3.pdf}
-
-}
-
-}
-
-\subcaption{\label{fig-dataviz-3}By health board and control area}
-\end{minipage}%
-%
-\begin{minipage}[t]{0.02\linewidth}
-
-{\centering
-
-~
-
-}
-
-\end{minipage}%
-%
-\begin{minipage}[t]{0.49\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{main_files/figure-pdf/fig-dataviz-4.pdf}
-
-}
-
-}
-
-\subcaption{\label{fig-dataviz-4}By health board and priority}
-\end{minipage}%
-\newline
-\begin{minipage}[t]{\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{main_files/figure-pdf/fig-dataviz-5.pdf}
-
-}
-
-}
-
-\subcaption{\label{fig-dataviz-5}By nature of incident and health board}
-\end{minipage}%
-
-\caption{\label{fig-dataviz}Time series of attended incidents at various
-levels.}
+\caption{\label{fig-dataviz2}Time series of attended incidents at
+various levels.}
 
@@
 where \(F^f_j(x)\) is the forecasted Cumulative Density Function (CDF)
-of period \(j\) and 𝐹\(F^0_j(x)\) is the true CDF of period \(j\).
+of period \(j\) and \(F^0_j(x)\) is the true CDF of period \(j\).
 
@@
 
+\begin{figure}
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-accuracy-1.pdf}
+
+}
+
+\caption{\label{fig-accuracy}Average accuracy by week for 12 weeks. CRPS
+is relative to a naive ECDF. MASE and MSSE are relative to the
+corresponding values for the training set.}
+
+\end{figure}
+
 \hypertarget{sec-conclusion}{%

### 2022-11-16 manuscript/03-experiment.qmd
Removed stray character
Updating sec 3

@@
 The dataset used in this study is from a major ambulance service trust in Great Britain.
-It contains information relating to the daily number of attended incidents from 10 October 2015 to 31 July 2019 by nature of incidents, priority, the health board managing the service and the control area (or region).
-Figure @fig-hierarchy depicts both the hierarchical and grouped structure of the data.
-Figure @fig-hierarchy-1 illustrates the nested hierarchical structure based on control area and health board and Figure @fig-hierarchy-2 shows the grouped structure by priority and the nature of incident.
+It contains information relating to the daily number of attended incidents from 1 October 2015 to 31 July 2019, disaggregated by nature of incidents, priority, the health board managing the service and the control area (or region).
+@fig-hierarchy depicts both the hierarchical and grouped structure of the data.
+@fig-hierarchy-1 illustrates the nested hierarchical structure based on control area and health board and @fig-hierarchy-2 shows the grouped structure by priority and the nature of incident.
 
@@
 
-Table @tbl-hierarchy also displays the structure of data with the total number of series at each level.
+@tbl-hierarchy also displays the structure of data with the total number of series at each level.
 At the top level, we have the total attended incidents for the country.
@@
 Given the total number of time series, direct visual analysis is infeasible.
-Therefore, we first compute features of all 1530 time series and display the strength of trend and weekly seasonality strength in Figure @fig-feature.
+Therefore, we first compute features of all 1530 time series [@m3pca] and display the strength of trend and weekly seasonality strength in @fig-feature.
 Each point represents one time series with the strength of trend in x-axis and the strength of seasonality in y-axis.
@@
 In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation.
-Figure @fig-dataviz2 reveals different information such as trend, seasonality and noise.
+@fig-dataviz2 reveals different information such as trend, seasonality and noise.
 For example, some series depict seasonality and trend, whereas some other series report low volume of attended incidents and entropy, making them more volatile and difficult to forecast.
@@
 
-**Naive:** We start with one of the simplest forecasting approaches used in practice - assuming that in the next few days, everything will be the same as similar days in the past.
-This is called "Naïve".
-In our case, given that we need a distribution of values, we will use a modified approach, where the empirical distribution of the daily attended incidents time series is used to forecast the future attended incidents distribution.
-We consider the empirical distribution of the most recent year of historic data on a rolling basis to capture potential changes in behavior over time.
+**Naive:** We start with a simple forecasting approach, assuming that the future days will be similar to past days.
+We use the empirical distribution of the past daily attended incidents to create the forecast distribution of future attended incidents.
+We consider the empirical distribution of the most recent year of historic data on a rolling basis to capture potential changes in behavior over time.\todo{Actually we use all the data.}
 
@@
 
-Despite the popularity and the relevance of automatic `ETS` in this study, however it may produce forecasts that are non-integer and negative.
-However, the number of attended incidents is an integer and non-negative.
-When using `ETS` , a time series transformation approach could be used to generate strictly positive forecasts, however forecasts are still not integer.
+Despite the popularity and the relevance of automatic `ETS` in this study, however it may produce forecast distributions that are non-integer and include negative values, despite the number of attended incidents being integer and non-negative.
+When using `ETS`, a time series transformation approach could be used to generate strictly positive forecasts, although forecast distributions will still be non-integer.
 An alternative is to use  forecasting models that produce integer, non-negative forecasts.
@@
 
-**Generalized Linear Model (GLM): ** GLMs are a family of models developed to extend the concept of linear regression models.
+**Generalized Linear Model (GLM):** GLMs are a family of models developed to extend the concept of linear regression models.
 They perform a regression by modeling the response variable as coming from a particular member of the exponential family, and then transforming the mean of the response so that the transformed mean is a linear function of the predictors.
@@
 
-**Poisson Regression using tscount: ** We consider another Poisson regression model that takes into account serial dependence in addition to covariates used in the GLM model.
+**Poisson Regression using tscount:** We consider another Poisson regression model that takes into account serial dependence in addition to covariates used in the GLM model.
 To that end, we use `tsglm()` function in _tscount_ package in R to model the attended incidents.
@@
 
-**Ensemble method:** we also use an ensemble method that combines forecasts generated from Naive, ETS, GLM and Poisson regression together to make more accurate forecasts than any individual
-model.
+**Ensemble method:** we also use an ensemble method that combines forecasts generated from Naive, ETS, GLM and Poisson regression to form a mixture distribution [@combinations].
 
@@
 
-The data is spitted into training (up until 2018-07-31) and test (from  2018-08-01 to the end) sets, with all model development and hyper-parameter tuning performed using training data only.
-The time series cross validation is performed with a forecast horizon of $2*42$ days and advances in 42-day steps.
+To evaluate the performance of the various forecasting approaches, we split the data into a series of ten training and test sets.
+We use a time series cross-validation approach [@hyndman2021forecasting], with a forecast horizon of 84 days, and each training set expanding in 42-day steps.
+The first training set uses all data up to 2018-04-25, and the first test set uses the 84 days beginning 2018-04-26.
+The second training set uses all data up to 2018-06-06, with the second test set using the following 84 days.
+The largest training set ends on 2019-05-09, with the test set ending on 2019-07-31.
+Model development and hyper-parameter tuning is performed using the training data only.
 Forecasting performance is evaluated using both point and probabilistic error measures.
-Forecast error is calculated by considering only the last 42 days and not the entire $2*42$ days.
-This corresponds on how forecasts are generated for planning in practice.
+While we compute forecast errors for the entire 12 weeks, we are most interested in the last 42 days of each test set, because that corresponds to  how forecasts are generated for planning in practice.
 Forecasting performance is evaluated using both point and probabilistic error measures.
-The point forecast accuracy is measured via Root Mean Squared Scaled Error (RMSSE) and Mean Absolute Scaled Error (MASE) that are describes below:
 
+Point forecast accuracy is measured via the Mean Squared Scaled Error (MSSE) and the Mean Absolute Scaled Error (MASE). The Mean Absolute Scaled Error (MASE) [@HK06] is calculated as:
 $$
-RMSSE = \sqrt{mean(q_{j}^2)},
+  \text{MASE} = \text{mean}(|q_{j}|),
+$$
+where
+$$
+  q_{j} = \frac{ e_{j}}
+    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|},
+$$
+and $e_{j}$ is the point forecast error for forecast horizon $j$, $m = 7$ (as we have daily seasonal series), $y_t$ is the observation for period $t$, and $T$ is the sample size (the number of observations used for training the forecasting model).
+The denominator is the mean absolute error of the seasonal naive method in the fitting sample of $T$ observations and is used to scale the error.
+Smaller MASE values suggest more accurate forecasts.
+Note that the measure is scale-independent, thus allowing us to average the results across series.
+
+A related measure is MSSE [@hyndman2021forecasting], which uses squared errors rather than absolute errors:
+$$
+  \text{MSSE} = \text{mean}(q_{j}^2),
 $$ {#eq-RMSSE}
-
 where,
-
 $$
   q^2_{j} = \frac{ e^2_{j}}
-    {\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
+    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
 $$
-$e_{j}$ is the point forecast error $j$ and $m = 1$ for non-seasonal series and $m = 7$ for daily seasonal series, $y_t$ is the observation for period $t$ and $T$ is the sample size (observations used for training the forecasting model).
-Smaller RMSSE values suggest more accurate forecasts.
-Note that the measure is scale-independent, thus allowing us to average the results across series.
+Again, this is scale-independent, and smaller MSSE values suggest more accurate forecasts.
 
-Mean Absolute Scaled Error (MASE) is calculated as:
-
+To measure the forecast distribution accuracy, we calculate the Continuous Rank Probability Score [@gneiting2014probabilistic].
+It rewards sharpness and penalizes miscalibration, so it measures overall performance of the forecast distribution.
 $$
-MASE = mean(|q_{j}|),
+  \text{CRPS} = \text{mean}(p_j),
+$$ {#eq-CRPS}
+where
 $$
-
-where,
+  p_j = \int_{-\infty}^{\infty} \left(G_j(x) - F_j(x)\right)^2dx,
 $$
-q_{j} = \frac{ e_{j}}
-    {\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|}.
-$$
-
-The denominator is the mean absolute error of the naive method in the fitting sample of $n$ observations and is used to scale the error.
-
-To measure the forecast distribution performance, we calculate  the Continuous Rank Probability Score [@gneiting2014probabilistic].
-It rewards sharpness and penalizes miscalibration, so it measures overall performance.
-
-$$
-CRPS=\frac{1}{h} \sum_{j=1}^{h} \int_{-\infty}^{\infty} \left(F^f_j(x) - F^0_j(x)\right)^2dx
-$$ {#eq-CRPS}
-
-where $F^f_j(x)$ is the forecasted Cumulative Density Function (CDF) of period $j$ and 𝐹$F^0_j(x)$ is the true CDF of period $j$.
+where $G_j(x)$ is the forecasted probability distribution function for forecast horizon $j$, and $F_j(x)$ is the true probability distribution function for the same period.

### 2022-11-16 manuscript/main.qmd
Updating sec 3

--- 2022-11-16 manuscript/main.qmd
@@
     keep-tex: true
+    include-in-header: preamble.tex
     journal:

### 2022-11-16 manuscript/main.tex
Updating sec 3

@@
 \makeatother
+
+\usepackage{todonotes,mathtools}
+\mathtoolsset{showonlyrefs}
 
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, boxrule=0pt, frame hidden, breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 in Great Britain. It contains information relating to the daily number
-of attended incidents from 10 October 2015 to 31 July 2019 by nature of
-incidents, priority, the health board managing the service and the
-control area (or region). Figure Figure~\ref{fig-hierarchy} depicts both
-the hierarchical and grouped structure of the data. Figure
+of attended incidents from 1 October 2015 to 31 July 2019, disaggregated
+by nature of incidents, priority, the health board managing the service
+and the control area (or region). Figure~\ref{fig-hierarchy} depicts
+both the hierarchical and grouped structure of the data.
 Figure~\ref{fig-hierarchy-1} illustrates the nested hierarchical
-structure based on control area and health board and Figure
+structure based on control area and health board and
 Figure~\ref{fig-hierarchy-2} shows the grouped structure by priority and
@@
 
-Table Table~\ref{tbl-hierarchy} also displays the structure of data with
-the total number of series at each level. At the top level, we have the
+Table~\ref{tbl-hierarchy} also displays the structure of data with the
+total number of series at each level. At the top level, we have the
 total attended incidents for the country. We can split these total
@@
 infeasible. Therefore, we first compute features of all 1530 time series
-and display the strength of trend and weekly seasonality strength in
-Figure Figure~\ref{fig-feature}. Each point represents one time series
-with the strength of trend in x-axis and the strength of seasonality in
-y-axis. It is clear that there are some series showing strong trends
-and/or seasonality, corresponding to series at the higher levels of the
-hierarchy. The majority of series show low trend and seasonality. These
-are time series belonging to the bottom series, series related to the
-nature of incidents for a given control, health board and priority
-level. Bottom series are dominated by noise with little or no systematic
-patterns.
+\citep{m3pca} and display the strength of trend and weekly seasonality
+strength in Figure~\ref{fig-feature}. Each point represents one time
+series with the strength of trend in x-axis and the strength of
+seasonality in y-axis. It is clear that there are some series showing
+strong trends and/or seasonality, corresponding to series at the higher
+levels of the hierarchy. The majority of series show low trend and
+seasonality. These are time series belonging to the bottom series,
+series related to the nature of incidents for a given control, health
+board and priority level. Bottom series are dominated by noise with
+little or no systematic patterns.
 
@@
 In addition to displaying the trend and seasonality features, we also
-visualize few time series at various levels of the aggregation. Figure
+visualize few time series at various levels of the aggregation.
 Figure~\ref{fig-dataviz2} reveals different information such as trend,
@@
 
-\textbf{Naive:} We start with one of the simplest forecasting approaches
-used in practice - assuming that in the next few days, everything will
-be the same as similar days in the past. This is called ``Naïve''. In
-our case, given that we need a distribution of values, we will use a
-modified approach, where the empirical distribution of the daily
-attended incidents time series is used to forecast the future attended
-incidents distribution. We consider the empirical distribution of the
-most recent year of historic data on a rolling basis to capture
-potential changes in behavior over time.
+\textbf{Naive:} We start with a simple forecasting approach, assuming
+that the future days will be similar to past days. We use the empirical
+distribution of the past daily attended incidents to create the forecast
+distribution of future attended incidents. We consider the empirical
+distribution of the most recent year of historic data on a rolling basis
+to capture potential changes in behavior over
+time.\todo{Actually we use all the data.}
 
@@
 Despite the popularity and the relevance of automatic \texttt{ETS} in
-this study, however it may produce forecasts that are non-integer and
-negative. However, the number of attended incidents is an integer and
-non-negative. When using \texttt{ETS} , a time series transformation
-approach could be used to generate strictly positive forecasts, however
-forecasts are still not integer. An alternative is to use forecasting
-models that produce integer, non-negative forecasts. in the following
-section we present Generalized Linear Models (GLMs) and Poisson
-Regression to produce count base forecasts.
+this study, however it may produce forecast distributions that are
+non-integer and include negative values, despite the number of attended
+incidents being integer and non-negative. When using \texttt{ETS}, a
+time series transformation approach could be used to generate strictly
+positive forecasts, although forecast distributions will still be
+non-integer. An alternative is to use forecasting models that produce
+integer, non-negative forecasts. in the following section we present
+Generalized Linear Models (GLMs) and Poisson Regression to produce count
+base forecasts.
 
-\textbf{Generalized Linear Model (GLM): } GLMs are a family of models
+\textbf{Generalized Linear Model (GLM):} GLMs are a family of models
 developed to extend the concept of linear regression models. They
@@
 
-\textbf{Poisson Regression using tscount: } We consider another Poisson
+\textbf{Poisson Regression using tscount:} We consider another Poisson
 regression model that takes into account serial dependence in addition
@@
 \textbf{Ensemble method:} we also use an ensemble method that combines
-forecasts generated from Naive, ETS, GLM and Poisson regression together
-to make more accurate forecasts than any individual model.
+forecasts generated from Naive, ETS, GLM and Poisson regression to form
+a mixture distribution \citep{combinations}.
 
@@
 
-The data is spitted into training (up until 2018-07-31) and test (from
-2018-08-01 to the end) sets, with all model development and
-hyper-parameter tuning performed using training data only. The time
-series cross validation is performed with a forecast horizon of \(2*42\)
-days and advances in 42-day steps. Forecasting performance is evaluated
-using both point and probabilistic error measures. Forecast error is
-calculated by considering only the last 42 days and not the entire
-\(2*42\) days. This corresponds on how forecasts are generated for
-planning in practice. Forecasting performance is evaluated using both
-point and probabilistic error measures. The point forecast accuracy is
-measured via Root Mean Squared Scaled Error (RMSSE) and Mean Absolute
-Scaled Error (MASE) that are describes below:
+To evaluate the performance of the various forecasting approaches, we
+split the data into a series of ten training and test sets. We use a
+time series cross-validation approach \citep{hyndman2021forecasting},
+with a forecast horizon of 84 days, and each training set expanding in
+42-day steps. The first training set uses all data up to 2018-04-25, and
+the first test set uses the 84 days beginning 2018-04-26. The second
+training set uses all data up to 2018-06-06, with the second test set
+using the following 84 days. The largest training set ends on
+2019-05-09, with the test set ending on 2019-07-31. Model development
+and hyper-parameter tuning is performed using the training data only.
+Forecasting performance is evaluated using both point and probabilistic
+error measures. While we compute forecast errors for the entire 12
+weeks, we are most interested in the last 42 days of each test set,
+because that corresponds to how forecasts are generated for planning in
+practice. Forecasting performance is evaluated using both point and
+probabilistic error measures.
 
-\begin{equation}\protect\hypertarget{eq-RMSSE}{}{
-RMSSE = \sqrt{mean(q_{j}^2)},
-}\label{eq-RMSSE}\end{equation}
-
-where,
-
-\[
-q^2_{j} = \frac{ e^2_{j}}
-    {\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
-\] \(e_{j}\) is the point forecast error \(j\) and \(m = 1\) for
-non-seasonal series and \(m = 7\) for daily seasonal series, \(y_t\) is
-the observation for period \(t\) and \(T\) is the sample size
-(observations used for training the forecasting model). Smaller RMSSE
-values suggest more accurate forecasts. Note that the measure is
+Point forecast accuracy is measured via the Mean Squared Scaled Error
+(MSSE) and the Mean Absolute Scaled Error (MASE). The Mean Absolute
+Scaled Error (MASE) \citep{HK06} is calculated as: \[
+  \text{MASE} = \text{mean}(|q_{j}|),
+\] where \[
+  q_{j} = \frac{ e_{j}}
+    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|},
+\] and \(e_{j}\) is the point forecast error for forecast horizon \(j\),
+\(m = 7\) (as we have daily seasonal series), \(y_t\) is the observation
+for period \(t\), and \(T\) is the sample size (the number of
+observations used for training the forecasting model). The denominator
+is the mean absolute error of the seasonal naive method in the fitting
+sample of \(T\) observations and is used to scale the error. Smaller
+MASE values suggest more accurate forecasts. Note that the measure is
 scale-independent, thus allowing us to average the results across
@@
 
-Mean Absolute Scaled Error (MASE) is calculated as:
+A related measure is MSSE \citep{hyndman2021forecasting}, which uses
+squared errors rather than absolute errors:
+\begin{equation}\protect\hypertarget{eq-RMSSE}{}{
+  \text{MSSE} = \text{mean}(q_{j}^2),
+}\label{eq-RMSSE}\end{equation} where, \[
+  q^2_{j} = \frac{ e^2_{j}}
+    {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2},
+\] Again, this is scale-independent, and smaller MSSE values suggest
+more accurate forecasts.
 
-\[
-MASE = mean(|q_{j}|),
-\]
-
-where, \[
-q_{j} = \frac{ e_{j}}
-    {\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|}.
-\]
-
-The denominator is the mean absolute error of the naive method in the
-fitting sample of \(n\) observations and is used to scale the error.
-
-To measure the forecast distribution performance, we calculate the
+To measure the forecast distribution accuracy, we calculate the
 Continuous Rank Probability Score \citep{gneiting2014probabilistic}. It
 rewards sharpness and penalizes miscalibration, so it measures overall
-performance.
-
+performance of the forecast distribution.
 \begin{equation}\protect\hypertarget{eq-CRPS}{}{
-CRPS=\frac{1}{h} \sum_{j=1}^{h} \int_{-\infty}^{\infty} \left(F^f_j(x) - F^0_j(x)\right)^2dx
-}\label{eq-CRPS}\end{equation}
-
-where \(F^f_j(x)\) is the forecasted Cumulative Density Function (CDF)
-of period \(j\) and \(F^0_j(x)\) is the true CDF of period \(j\).
+  \text{CRPS} = \text{mean}(p_j),
+}\label{eq-CRPS}\end{equation} where \[
+  p_j = \int_{-\infty}^{\infty} \left(G_j(x) - F_j(x)\right)^2dx,
+\] where \(G_j(x)\) is the forecasted probability distribution function
+for forecast horizon \(j\), and \(F_j(x)\) is the true probability
+distribution function for the same period.
 

### 2022-11-16 manuscript/03-experiment.qmd
add new references and typos
typo

@@
 
-**Poisson Regression using tscount:** We consider another Poisson regression model that takes into account serial dependence in addition to covariates used in the GLM model.
-To that end, we use `tsglm()` function in _tscount_ package in R to model the attended incidents.
+**Poisson Regression using tscount: ** We consider another Poisson regression model that takes into account serial dependence in addition to covariates used in the GLM model.
+To that end, we use `tsglm()` function in _tscount_ package in R [@JSSv082i05] to model the attended incidents.
 The logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.

### 2022-11-17 manuscript/01-introduction.qmd
Cleaner code
Updated intro

@@
 
-Inability to match the resources with the demand in Emergency Medical Services (EMS) results in an overcrowding care system.
-This is a serious problem causing challenging situations on patient flow with serious consequences on patients, staff and the entire care system [@ekstrom2015forecasting; @ROSTAMITABAR20221197].
-Demand forecasting in EMS is a vital element that helps to depict various courses of action to avoid the mismatch, which can result in massive savings in terms of patient safety and lives.
-An accurate daily demand forecasting enables planners and decision makers to manage resources to meet anticipated patients, reconfigure units, redeploy staff and vehicles, where necessary.
+A failure to match available resources to demand in Emergency Medical Services (EMS) results in patient flow problems, with serious consequences for patients, staff and the entire care system [@ekstrom2015forecasting; @ROSTAMITABAR20221197].
+Demand forecasting in EMS helps service planners to avoid the mismatch, potentially providing massive savings in costs and lives, and leading to better patient outcomes.
+Accurate daily demand forecasting enables planners and decision makers to manage resources to meet anticipated patients, reconfigure units, and redeploy staff and vehicles as necessary.
 
-Demand forecasts at EMS are typically required at multiple levels of cross-sectional granularities to inform various planning and decision-making processes [@hulshof2012taxonomic].
-There are some planning process at the national level (more strategic or long-term) such as workforce resource planning and budgeting; sub-national, regional or healthcare level (tactical or medium-term) such as temporary capacity expansions, resource sharing and staff-shift scheduling; and hospital or station level (operational or short-term) such as planning rosters for staff and ambulance deployment.
-Demand forecasts might also be required at different level for a specific group of interest such as nature of demand or priority.
+Demand forecasts at EMS are typically required at multiple levels of an organization to inform various planning and decision-making processes [@hulshof2012taxonomic].
+There are some planning process at the national level (strategic and long-term) such as workforce resource planning and budgeting; sub-national, regional or healthcare level (tactical and medium-term) such as temporary capacity expansions, resource sharing and staff-shift scheduling; and hospital or station level (operational and short-term) such as planning rosters for staff and ambulance deployment.
+Demand forecasts might also be required at different levels for a specific area of interest such as the nature of demand or the priority level.
 Moreover, the time series data in EMS has an inherent hierarchical and grouped structure to support such forecasting requirements.
-Demand for emergency medical services at the country level can be disaggregated in a geographical hierarchy into sub-national, regions, health boards, stations/hospitals or divided into groups such as nature of incidents or demand priority.
-Therefore, using forecasting methodologies that account for hierarchical and/or grouped structures of time series in EMS seems to be a natural fit.
+Demand for emergency medical services at the country level can be disaggregated in a geographical hierarchy into sub-national, regions, health boards, and stations/hospitals, or divided into groups such as the nature of incidents or demand priority.
+Therefore, using forecasting methodologies that account for hierarchical and/or grouped structures of time series in EMS is a natural fit.
 
-However, despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], this area is neglected and the main focus has been on producing forecasts at a single level, independently.
-Generating independent forecasts not only ignore the inherent hierarchical and/or grouped structure relationships of the time series demand but also results in a lack of consistency and coordination.
-Consistent forecasts of the EMS demand across all hierarchical and grouped levels are paramount for an effective planning and decision making.
-The hierarchical forecasting approaches can not only create consistent forecasts but also achieve more accurate forecasts than the independent (base) forecasts [@hyndman2011optimal].
-Obtaining consistent forecasts at different levels is important as it helps to avoid making conflicting decisions.
+However, despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the hierarchical data structure has been largely ignored, and the main focus has been on producing independent forecasts at a single level.
+Generating independent forecasts can result in a lack of consistency and coordination, and therefore leads to less effective planning and decision making.
 With hierarchical forecasting, plans at any level are based on coherent forecasts and therefore can be aligned.
@@
 Hierarchical forecasting framework can be used as a tool to improve coordination between teams across the care services at the national, sub-national, regional and local levels.
-To our knowledge, there is not only no research in the EMS forecasting to account for the hierarchical and grouped structure of the system but also in the entire field of forecasting for healthcare management.
+The hierarchical forecasting approaches not only create consistent forecasts, but are usually also more accurate than the independent (base) forecasts [@hyndman2011optimal].
+To our knowledge, there has been no previous research involving hierarchical and grouped forecasting in the entire field of forecasting for healthcare management.
 
 In this paper, we address this gap by investigating the application of hierarchical forecasting approaches in the EMS using daily time series of verified incidents from 2015 to 2020 in a major ambulance service in Great Britain.
-The data has hierarchical and grouped structures with hierarchies at the national, control (i.e.
+The data has hierarchical and grouped structures, with hierarchies at the national, control (i.e.
 sub-national), health board (i.e.
-regional) and groups by priority and nature of incidents.
-We produce not only the point forecast but also the forecast distribution across all levels, which is critical for an effective planning and associated risk management.
+regional) levels, as well as groups by priority and nature of incidents.
+We produce consistent point forecasts and forecast distributions for all levels, which is critical for an effective planning and associated risk management.
 We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal reconciliation approaches.
-We first generate independent/based forecasts using Exponential Smoothing State Space (ETS),  Generalized Linear Model (GLM), Poisson regression, a simple empirical distribution and an ensemble method followed by applying bottom-up and optimal reconciliation approaches.
-Forecast performance is assessed by the Root Mean Squared Scaled Error (RMSSE) for point forecasts and the Continuous Ranked Probability Score (CRPS) for the probabilistic forecasts.
-This paper complies with the principles of the reproducibility  [@stodden2013best; @boylan2015reproducibility].
-Therefore, the study could equally be applied to any healthcare service (e.g.
-emergency department, primary or social care) subject to the time series having a hierarchical and/or grouped structure, which is generally the case in the healthcare sector.
+We first generate independent/base forecasts using Exponential Smoothing State Space (ETS), Generalized Linear Model (GLM), Poisson regression, a simple empirical distribution and an ensemble method, followed by applying bottom-up and optimal reconciliation approaches.
+Forecast performance is assessed by the Mean Squared Scaled Error (MSSE) for point forecasts and Continuous Ranked Probability Scores (CRPS) for the probabilistic forecasts.
+This paper complies with reproducibility principles [@stodden2013best; @boylan2015reproducibility], and can be applied to any healthcare service (e.g., emergency department, primary or social care) subject to the time series having a hierarchical and/or grouped structure.
 
-The remainder of this article is structured as follows: In section 2, we provide a brief review of the literature and discuss its limitation to position our work; in section 3, we present the experiment design describing the data set, forecasting methods and forecast evaluation metrics.
-In Section 4, we discuss the hierarchical time series forecasting approaches to generate both point and probabilistic forecasts.
-In section 5, we present and discuss our results; in section 7, we summarize our findings and present ideas for future research.
+The remainder of this article is structured as follows: In @sec-lit, we provide a brief review of the literature and discuss its limitation to position our work; in @sec-experiment, we present the experiment design describing the data set, forecasting methods and forecast evaluation metrics.
+In @sec-htc, we discuss the hierarchical time series forecasting approaches to generate both point and probabilistic forecasts.
+In @sec-results, we present and discuss our results; in @sec-conclusion, we summarize our findings and present ideas for future research.

### 2022-11-17 manuscript/main.qmd
Updated intro

@@
 ---
-title: "Hierarchical time series forecasting in Emergency Medical Services"
+title: "Hierarchical Time Series Forecasting in Emergency Medical Services"
 author:

### 2022-11-17 manuscript/main.qmd
Formatting and caching

@@
 keywords:
-  - forecasting
   - healthcare
@@
   - forecast reconciliation
-  - hierarchical time series
   - ambulance demand
   - attended incidents
-
 bibliography: bibliography.bib
@@
   warning: false
+  cache: true
 pdf-engine: pdflatex
@@
 #| label: setup
+#| context: setup
 #| include: false

### 2022-11-17 manuscript/03-experiment.qmd
Working on sec 3

@@
 
-We are interested in generating forecasts to inform the planning horizon of $ph= 42$ days, required by planners in the ambulance services trust.
-The forecast horizon in this study is $fh = 2 \times ph$ days ahead (2*42 days planning horizon).
-This is because the planning is generally frozen for $ph$ days and considering a forecast horizon of $ph$ days might not be helpful for planning.
-While forecasts are generated for $2 \times ph$ days ahead, performance evaluation is only assessed based on the last $ph$ days and not the $2 \times ph$ days.
-The forecasts are produced for the holdout of 365 days using time series cross-validation [@hyndman2021forecasting].
+We are interested in generating forecasts to inform the planning horizon of 42 days, required by planners in the ambulance service.
+The forecast horizon in this study is $2 \times 42 = 84$ days ahead, because the planning is generally frozen for 42 days and so forecasts of the next 42 days is not particularly helpful for planning.
+While forecasts are generated for 84 days ahead, performance evaluation is only assessed based on the last 42 days and not the whole forecast period.
+The forecasts are produced for various training and test sets using time series cross-validation [@hyndman2021forecasting].
 
-In the following section, we discuss the dataset, describe the forecasting methods used to generate base forecasts and present the point and probabilistic accuracy measures.
+In the following section, we discuss the dataset, describe the forecasting methods used to generate base forecasts, and present the point and probabilistic accuracy measures.
 
@@
 
-The dataset used in this study is from a major ambulance service trust in Great Britain.
+The dataset used in this study is from a major ambulance service in Great Britain.
 It contains information relating to the daily number of attended incidents from 1 October 2015 to 31 July 2019, disaggregated by nature of incidents, priority, the health board managing the service and the control area (or region).
@@
 #| label: fig-hierarchy
+#| cache: true
 #| out.width: "60%"
@@
   level1 = "Total",
-  level2 = c("Central & West", "Central & West", "Central & West",
-             "North", "South & East", "South & East", "South & East"),
+  level2 = c(
+    "Central & West", "Central & West", "Central & West",
+    "North", "South & East", "South & East", "South & East"
+  ),
   level3 = c("HD", "SB", "PO", "BC", "CV", "CT", "AB")
@@
   ggraph::geom_node_text(
-    aes(label = c("All country",
+    aes(label = c(
+      "All country",
       "Central & West", "North", "South & East",
-                  "HD", "SB", "PO", "BC", "CV", "CT", "AB"))
+      "HD", "SB", "PO", "BC", "CV", "CT", "AB"
+    ))
   ) +
@@
 #| label: tbl-hierarchy
+#| cache: true
 #| tbl-cap: "Number of time series in each level for the hierarchical & grouped structure of attended incidents"
@@
 Each point represents one time series with the strength of trend in x-axis and the strength of seasonality in y-axis.
+Both measures are on a scale of [0,1].
 It is clear that there are some series showing strong trends and/or seasonality, corresponding to series at the higher levels of the hierarchy.
@@
 
-In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation.
+In addition to displaying the trend and seasonality features, we also visualize a few time series at various levels of the aggregation.
 @fig-dataviz2 reveals different information such as trend, seasonality and noise.
@@
 ```{r}
-#| label: fig-dataviz
-#| eval: false
-#| include: false
-#| dependson: "fig-feature"
-#| out.width: "70%"
-#| fig-cap: "Time series of attended incidents at various levels."
-#| fig-subcap:
-#|   - "Whole country"
-#|   - "By control area"
-#|   - "By health board and control area"
-#|   - "By health board and priority"
-#|   - "By nature of incident and health board"
-#| layout: [[44,-2, 44], [44,-2, 44], [100]]
-p_total <- incident_gthf %>%
-  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  autoplot() +
-  labs(x = "Date", y = "Incidents") +
-  ggthemes::theme_few()
-
-p_total_control <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>%
-  select(-nature, -category) %>%
-  group_by(date, region) %>%
-  summarise(incident = sum(incident), .groups = "drop") %>%
-  ggplot(aes(x = date, y = incident, color = factor(region))) +
-  geom_line() +
-  labs(x = "Date", y = "Incidents", color = "Control") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "bottom")
-
-p_control <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>%
-  select(-nature, -category) %>%
-  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
-  geom_line() +
-  facet_wrap(vars(factor(region)), scales = "free_y") +
-  labs(x = "Date", y = "Incidents", color = "Health board") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "bottom")
-
-p_priority <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>%
-  select(-nature, -region) %>%
-  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
-  geom_line() +
-  facet_wrap(vars(factor(category)), scales = "free_y") +
-  labs(x = "Date", y = "Incidents", color = "Health board") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "bottom")
-
-selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
-
-p_nature_healthboard <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & !is_aggregated(nature)) %>%
-  as_tibble() %>%
-  mutate(nature = as.character(nature)) %>%
-  filter(nature %in% selected_nature) %>%
-  group_by(date, nature, lhb) %>%
-  summarise(incident = sum(incident), .groups = "drop") %>%
-  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
-  geom_line() +
-  facet_wrap(vars(factor(nature)), scales = "free_y") +
-  labs(x = "Date", y = "Incidents", color = "Health board") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "none")
-
-p_total
-p_total_control
-p_control
-p_priority
-p_nature_healthboard
-```
-
-```{r}
 #| label: fig-dataviz2
+#| cache: true
 #| dependson: "fig-feature"
@@
 #| fig-height: 10
-#| fig-cap: "Time series of attended incidents at various levels."
-no_x_axis <- theme(axis.title.x=element_blank(),
-                   axis.text.x=element_blank(),
-                   axis.ticks.x=element_blank())
+#| fig-cap: "Time series of attended incidents at various levels.
+The panels show data from the whole country, by control area, by health board, by priority level, and by nature of incident.
+Only four of the 35 nature of incident categories are shown to avoid too much overplotting."
+no_x_axis <- theme(
+  axis.title.x = element_blank(),
+  axis.text.x = element_blank(),
+  axis.ticks.x = element_blank()
+)
 p_total <- incident_gthf %>%
@@
   geom_line() +
-  #facet_wrap(vars(factor(region)), scales = "free_y") +
+  # facet_wrap(vars(factor(region)), scales = "free_y") +
   labs(y = "Incidents", color = "Health board") +
@@
     category = recode(category, RED = "Red", AMB = "Amber", GRE = "Green"),
-    category = factor(category, levels=c("Red","Amber","Green"))
+    category = factor(category, levels = c("Red", "Amber", "Green"))
   ) |>
@@
   geom_line() +
-  scale_color_manual(values = c(Red="#ff3300", Amber = "#E69f00", Green="#009e73")) +
+  scale_color_manual(values = c(Red = "#ff3300", Amber = "#E69f00", Green = "#009e73")) +
   labs(y = "Incidents", color = "Priority") +
@@
 Given the presence of various significant patterns in the past attended incidents, we consider three different forecasting models to generate the base forecasts.
-Once the base forecasts are produced, hierarchical and grouped time series methods are used to reconcile them across the all levels.
-We briefly discussed forecasting models in the following sections, and the hierarchical forecasting methods are discussed in @sec-htc.
+Once the base forecasts are produced, hierarchical and grouped time series methods are used to reconcile them across all levels.
+We briefly discuss forecasting models in the following sections, and the hierarchical forecasting methods are discussed in @sec-htc.
 
@@
 
-**Exponential Smoothing State Space model (ETS):** ETS models [@hyndman2021forecasting] can combine trend, seasonality and error components in a time series through various forms such as additive, multiplicative or mixed.
-The trend component can be none ("N"), Additive ("A"), damped ("Ad") or multiplicative ("M").
-The seasonality can be none ("N"), Additive ("A"), or multiplicative ("M").
-The error term can also be additive ("A") or multiplicative ("M").
+**Exponential Smoothing State Space model (ETS):** ETS models [@hyndman2021forecasting] can combine trend, seasonality and error components in a time series through various forms that can be additive, multiplicative or mixed.
+The trend component can be none ("N"), Additive ("A") or damped ("Ad"); the seasonality can be none ("N"), Additive ("A"), or multiplicative ("M"); and the error term can be additive ("A") or multiplicative ("M").
 To forecast the attended incidents at each level, we use the `ETS()` function in the `fable` package [@fable2022] in R.
-To identify the best model for a given time series, the ETS function uses the corrected Akaike’s Information Criterion (AICc).
+To identify the best model for a given time series, the `ETS` function uses the corrected Akaike’s Information Criterion (AICc).
 
-Despite the popularity and the relevance of automatic `ETS` in this study, however it may produce forecast distributions that are non-integer and include negative values, despite the number of attended incidents being integer and non-negative.
+Despite the popularity and the relevance of automatic `ETS` in this study, it may produce forecast distributions that are non-integer and include negative values, despite the number of attended incidents being integer and non-negative.
 When using `ETS`, a time series transformation approach could be used to generate strictly positive forecasts, although forecast distributions will still be non-integer.
 An alternative is to use forecasting models that produce integer, non-negative forecasts.
-in the following section we present Generalized Linear Models (GLMs) and Poisson Regression to produce count base forecasts.
+In the following section we present Generalized Linear Models (GLMs) and Poisson time series regression to produce count base forecasts.
 

### 2022-11-17 manuscript/03-experiment.qmd
Finished work on section 3 for now.

@@
 We use the empirical distribution of the past daily attended incidents to create the forecast distribution of future attended incidents.
-We consider the empirical distribution of the most recent year of historic data on a rolling basis to capture potential changes in behavior over time.\todo{Actually we use all the data.}
 
@@
 
-Despite the popularity and the relevance of automatic `ETS` in this study, it may produce forecast distributions that are non-integer and include negative values, despite the number of attended incidents being integer and non-negative.
+Despite the popularity and the relevance of automatic `ETS` in this study, it may produce forecast distributions that are non-integer and include negative values, although the number of attended incidents is always integer and non-negative.
 When using `ETS`, a time series transformation approach could be used to generate strictly positive forecasts, although forecast distributions will still be non-integer.
@@
 
-**Generalized Linear Model (GLM):** GLMs are a family of models developed to extend the concept of linear regression models.
-They perform a regression by modeling the response variable as coming from a particular member of the exponential family, and then transforming the mean of the response so that the transformed mean is a linear function of the predictors.
-One of the models that is frequently used in practice to generate count forecasts is the Poisson regression.
+**Generalized Linear Model (GLM):** GLMs are a family of models developed to extend the concept of linear regression models to non-Gaussian distributions [@Faraway2016].
+They model the response variable as a particular member of the exponential family, with the mean being a transformation of a linear function of the predictors.
+One of the models that is frequently used in practice to generate count forecasts is Poisson regression.
 We will consider forecasting attended incidents using the covariates spline trend, day of the week dummy variables (from Monday to Sunday), Fourier terms to capture yearly seasonality, public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
-We fit a Poisson regression model using the function `glm()` from the package in R, with the argument `family = poisson` to specify that we wish to fit a Poisson regression model.
+We fit a Poisson regression model using the function `glm()` from the *stats* package in R, with the argument `family = poisson` to specify that we wish to fit a Poisson regression model with a log link function.
 
-**Poisson Regression using tscount: ** We consider another Poisson regression model that takes into account serial dependence in addition to covariates used in the GLM model.
-To that end, we use `tsglm()` function in _tscount_ package in R [@JSSv082i05] to model the attended incidents.
-The logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.
-This model can be estimated via maximization of the likelihood function based on Poisson mass function.
-The regression model captures the short range serial dependence by including the three order autoregressive terms.
+**Poisson Regression using tscount:** We also consider another Poisson regression model that takes into account serial dependence, using the `tsglm()` function in the _tscount_ package in R [@JSSv082i05] to model the attended incidents.
+Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.
+This model captures the short range serial dependence by including three autoregressive terms, in addition to the same covariates that were used in the GLM model.
+To distinguish this from the previous GLM model, we will refer to this model as "tscount".
 
-**Ensemble method:** we also use an ensemble method that combines forecasts generated from Naive, ETS, GLM and Poisson regression to form a mixture distribution [@combinations].
+**Ensemble method:** Finally, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and tscount models to form a mixture distribution [@combinations].
 

### 2022-11-17 manuscript/03-experiment.qmd
first draft of results and conclusion

--- 2022-11-16 manuscript/03-experiment.qmd
@@
 ```{r}
-#| label: fig-dataviz
-#| eval: false
-#| include: false
-#| dependson: "fig-feature"
-#| out.width: "70%"
-#| fig-cap: "Time series of attended incidents at various levels."
-#| fig-subcap:
-#|   - "Whole country"
-#|   - "By control area"
-#|   - "By health board and control area"
-#|   - "By health board and priority"
-#|   - "By nature of incident and health board"
-#| layout: [[44,-2, 44], [44,-2, 44], [100]]
-p_total <- incident_gthf %>%
-  filter(is_aggregated(region) & is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  autoplot() +
-  labs(x = "Date", y = "Incidents") +
-  ggthemes::theme_few()
-
-p_total_control <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>%
-  select(-nature, -category) %>%
-  group_by(date, region) %>%
-  summarise(incident = sum(incident), .groups = "drop") %>%
-  ggplot(aes(x = date, y = incident, color = factor(region))) +
-  geom_line() +
-  labs(x = "Date", y = "Incidents", color = "Control") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "bottom")
-
-p_control <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>%
-  select(-nature, -category) %>%
-  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
-  geom_line() +
-  facet_wrap(vars(factor(region)), scales = "free_y") +
-  labs(x = "Date", y = "Incidents", color = "Health board") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "bottom")
-
-p_priority <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & is_aggregated(nature)) %>%
-  as_tibble() %>%
-  select(-nature, -region) %>%
-  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
-  geom_line() +
-  facet_wrap(vars(factor(category)), scales = "free_y") +
-  labs(x = "Date", y = "Incidents", color = "Health board") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "bottom")
-
-selected_nature <- c("CHESTPAIN", "STROKECVA", "BREATHING", "ABDOMINAL")
-
-p_nature_healthboard <- incident_gthf %>%
-  filter(!is_aggregated(region) & !is_aggregated(lhb) & !is_aggregated(category) & !is_aggregated(nature)) %>%
-  as_tibble() %>%
-  mutate(nature = as.character(nature)) %>%
-  filter(nature %in% selected_nature) %>%
-  group_by(date, nature, lhb) %>%
-  summarise(incident = sum(incident), .groups = "drop") %>%
-  ggplot(aes(x = date, y = incident, color = factor(lhb))) +
-  geom_line() +
-  facet_wrap(vars(factor(nature)), scales = "free_y") +
-  labs(x = "Date", y = "Incidents", color = "Health board") +
-  ggthemes::scale_color_colorblind() +
-  ggthemes::theme_few() +
-  theme(legend.position = "none")
-
-p_total
-p_total_control
-p_control
-p_priority
-p_nature_healthboard
-```
-
-```{r}
 #| label: fig-dataviz2
@@
   labs(x = "", y = "Incidents") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few() +
   no_x_axis
@@
   labs(y = "Incidents", color = "Control")  +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few() +
   no_x_axis
@@
   labs(y = "Incidents", color = "Health board") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few() +
   no_x_axis
@@
   labs(y = "Incidents", color = "Priority") +
+  ggthemes::theme_few() +
   no_x_axis
@@
   geom_line() +
-  labs(x = "Date", y = "Incidents", color = "Nature of incident")
+  labs(x = "Date", y = "Incidents", color = "Nature of incident")+
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few()
 

### 2022-11-17 manuscript/05-results.qmd
Updated intro
first draft of results and conclusion

@@
 
-## Point forecast
+```{r}
+#| include: false
+library(kableExtra)
+rmsse <- readr::read_rds(here::here("results/rmsse.rds"))
+mase <- readr::read_rds(here::here("results/mase.rds"))
+crps <- readr::read_rds(here::here("results/crps.rds")) |>
+  pivot_wider(names_from = model, values_from = crps) |>
+  group_by(method, h, series) |>
+  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
+  ungroup()
+```
 
-## Probabilistic forecast
+In this section, we compare the forecasting performance of Naive, ETS, GLM, Tscount and the ensemble model using base forecast and Minimum Trace (Mint) methods.
+We have also computed the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches.
+However, they are not reported here because their accuracy is outperformed by Mint.
+We should also note that forecasts, and consequently their corresponding errors, are generated for the entire hierarchy and they could be reported at any level, if required.
+
+The overall forecasting performance is reported in @tbl-result, in which the average forecast accuracy per model, method and the hierarchical level is presented.
+Reported forecast accuracy are averaged across all forecast horizons, rolling origins and series at each level.
+@tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom level series.
+Point forecast performance are reported using MASE and RMSSE in @tbl-result-1  and @tbl-result-2, respectively.
+Probabilistic forecast accuracy is reported using CRPS in @tbl-result-3.
+The bold entries in each table identify a combination of method and model that performs best for the corresponding level (i.e.
+each column), based on the smallest values of accuracy measures.
+
+@tbl-result-1 shows that forecast reconciliation (i.e.
+Mint) improves forecast accuracy at the higher levels of the hierarchy including total, control area and health board.
+However, it is does not result in accuracy improvement at the bottom level series, fro which base forecasts are more accurate.
+This might be due to the noisy structure of time series at the bottom level and the fact that base forecasts produced for the bottom of hierarchy are from well-specified forecasting models and therefore does not gain benefit from forecast reconciliation.
+Generally, if the bottom-level series are sufficiently smooth, base forecast is expected to outperform hierarchical reconciliation due to the better forecasts available at the level.
+It is also clear from @tbl-result-1 that the ensemble method improves forecast accuracy at total, control area and health board.
+However, this does not remain valid for bottom series where the Poisson regression by GLM outperforms other approaches.
+
+@tbl-result-2 also reports point forecast accuracy but using RMSSE.
+The outperformance of forecast reconciliation using Mint at higher levels of hierarchy is also confirmed by RMSSE.
+However, for the bottom series, we observe that the base forecasts generated by ETS are the most accurate, followed by its reconciliations forecasts.
+
+```{r}
+#| label: tbl-result
+#| tbl-cap: "Average forecast perfomance calculated on out-of-sample with time series cross validation applied to attended incident data. The best approach is highlighted in bold."
+#| tbl-subcap: 
+#|   - "Point forecast accuracy using MASE"
+#|   - "Point forecast accuracy using RMSSE"
+#|   - "Probabilistic forecast accuracy using CRPS"
+#| layout-nrow: 3
+rmsse1 <- rmsse |>
+  filter(
+    method %in% c("mint","base"), model != "qcomb"
+  ) |> group_by(method,model,series) |> summarise(rmsse = mean(rmsse), .groups = "drop") |> 
+  pivot_wider(names_from = series, values_from = rmsse) |> 
+  mutate_if(is.numeric, format, digits=4,nsmall = 0) |> 
+  select(method,model,Total,`Control areas`,`Health boards`,Bottom)
+mase1 <- mase |>
+  filter(
+    method %in% c("mint","base"), model != "qcomb"
+  ) |> group_by(method,model,series) |> summarise(mase = mean(mase), .groups = "drop") |> 
+  pivot_wider(names_from = series, values_from = mase) |> 
+  mutate_if(is.numeric, format, digits=4,nsmall = 0) |> 
+  select(method,model,Total,`Control areas`,`Health boards`,Bottom)
+crps1 <- readr::read_rds(here::here("results/crps.rds")) |>
+  filter(method %in% c("mint","base"), model != "qcomb") |> 
+  group_by(method,model,series) |> summarise(crps = mean(crps), .groups = "drop") |> 
+  pivot_wider(names_from = series, values_from = crps)|> 
+  mutate_if(is.numeric, format, digits=4,nsmall = 0) |> 
+  select(method,model,Total,`Control areas`,`Health boards`,Bottom)
+
+mase1[6, 3] <- cell_spec(mase1[6, 3], "latex", bold = T)
+mase1[6, 4] <- cell_spec(mase1[6, 4], "latex", bold = T)
+mase1[6, 5] <- cell_spec(mase1[6, 5], "latex", bold = T)
+mase1[3, 6] <- cell_spec(mase1[3, 6], "latex", bold = T)
+
+rmsse1[6, 3] <- cell_spec(rmsse1[6, 3], "latex", bold = T)
+rmsse1[6, 4] <- cell_spec(rmsse1[6, 4], "latex", bold = T)
+rmsse1[6, 5] <- cell_spec(rmsse1[6, 5], "latex", bold = T)
+rmsse1[2, 6] <- cell_spec(rmsse1[2, 6], "latex", bold = T)
+
+crps1[6, 3] <- cell_spec(crps1[6, 3], "latex", bold = T)
+crps1[6, 4] <- cell_spec(crps1[6, 4], "latex", bold = T)
+crps1[6, 5] <- cell_spec(crps1[6, 5], "latex", bold = T)
+crps1[7, 6] <- cell_spec(crps1[7, 6], "latex", bold = T)
+
+kbl(mase1, booktabs = T, escape = F) |> 
+  add_header_above(c(" " = 2, "MASE" = 4)) |> 
+   kable_styling(latex_options = c("hold_position"))
+
+kbl(rmsse1,booktabs = T, escape = F) %>%
+  add_header_above(c(" " = 2, "RMSSE" = 4)) |> 
+  kable_styling(latex_options = c("hold_position"))
+
+kbl(crps1,booktabs = T, escape = F) %>%
+  add_header_above(c(" " = 2, "CRPS" = 4)) |> 
+ kable_styling(latex_options = c("hold_position"))
+```
+
+@tbl-result-2 presents the accuracy of the forecast distributions measures by CRPS, which considers both forecasting reliability and interval sharpness.
+The smaller the value of CRPS, the better the comprehensive performance.
+We observe that forecast reconciliation results in forecast improvement, regardless of the hierarchical level.
+Ensemble method is also more accurate fore higher levels, but ETS performs better at the bottom level.
+While results show that forecast reconciliation does not improve point forecast at the bottom level, however @tbl-result-2 indicates that it generates more accurate forecasts than base method.
+The outperformance of probabilistic forecast reconciliation at the bottom level may highlight an important fact.
+The reconciliation method may improve forecast accuracy at certain number of quantiles at the tails of forecast distribution, which are critical for managing risks.
+
+Overall, our results indicate that forecast reconciliation using MinT method provides reliable forecasts, and all methods that reconcile using information at all levels of the forecast improve upon base (unreconciled) forecasts, except the bottom level series.
+For bottom series with less obvious patterns, base forecasts near center of forecast distribution might be more accurate, while forecast reconciliation using Mint could improve accuracy in the tails of of the distribution.
+
+In addition to the overall forecast accuracy presented in @tbl-result, we also report the point and probabilistic forecast accuracy measures for each forecast horizon in @fig-accuracy.
+The figure focuses on the hierarchical levels important for decision making including total, control area and health board, however this could be calculated for any level.
+We only illustrate the results of the Mint method, given the its outperformance described in @tbl-result.
+For the illustration purpose, we report the average weekly forecast accuracy instead of the daily forecast horizon (i.e.
+84 days including freezed planning (7 weeks = 42 days) plus future planing period (7 weeks = 42 days).
+Therefore, x-axis shows horizons from week 1 (h= 1,...,7) to week 12 (h= 78,...,84).
+The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision making.
+For the point forecast, we can see that ensemble model performs much better than others regardless of forecast horizons.
+For the probabilistic forecasts, the distinction between forecasting models is less obvious.
+It is important to highlight that, all forecasting models outperform the Naive empirical distribution that is used as a Benchmark for both point and probabilsitic forecasts.
 
@@
 #| fig-cap: "Average accuracy by week for 12 weeks. CRPS is relative to a naive ECDF. MASE and MSSE are relative to the corresponding values for the training set."
-rmsse <- readr::read_rds(here::here("results/rmsse.rds"))
-mase <- readr::read_rds(here::here("results/mase.rds"))
-crps <- readr::read_rds(here::here("results/crps.rds")) |>
-  pivot_wider(names_from = model, values_from = crps) |>
-  group_by(method, h, series) |>
-  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
-  ungroup() |>
+#| fig-pos: "H"
+crps <- crps |>
   pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")
@@
 acc_summary |>
-  ggplot(aes(x = week, y = accuracy, group = model)) +
-  geom_line(aes(col = model)) +
+  ggplot(aes(x = week, y = accuracy, group = model, col = model)) +
+  geom_line() +
+  geom_point(size=.5)+
   facet_grid(measure ~ series, scales = "free_y") +
-  labs(y = "Average accuracy", x = "Week ahead")
+  labs(y = "Average accuracy", x = "Week ahead")+
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few()
 ```

### 2022-11-17 manuscript/06-conclusion.qmd
first draft of results and conclusion

@@
-# Conclusion {#conclusion}
+# Conclusion {#sec-conclusion}
 
-Moreover, both models either underestimated or overestimated the number of calls.
-Authors suggested that the inability to capture a positive trend resulted in underestimation, and the opposite was due to a negative trend.
-Therefore, authors recommended that further research should add certain temporal variables able to capture the trend.
+Forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures.
+For example, looking at time series of arrival calls in a clinical desk service, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest.
+Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area.
+Alternative structure may arise when attributes of interest are crossed rather than nested.
+For example, the total demand could be disaggrgated by priority (e.g Red, Amber, Green) or nature of incidents.
+It is also natural to have a mixed structures, for example the total demand could be disaggregated by priority and also by health board.
+
+Despite the inherent hierarchical structure of the forecasting problem in EMS, the common practice is to produce point forecast for each time series independently.
+This practice may lead to the lack of coordination and possibly undesirable and conflicting outcomes.
+Furthermore, due to the asymmetric impact of resource allocation in this area, quantifying forecast uncertainty through probabilistic forecasts is also of value as it enables planners to manage associated risks.
+In this paper, we investigate the application of hierarchical forecasting methods for producing a probabilistic forecast of daily incident attended up to 84 days ahead, using different forecasting methods.
+
+Our results indicates that forecast reconciliation can not only contribute to a more coordinated efforts through producing coherent forecasts when it comes to using forecasting to inform decisions in EMS, but also it can result in forecast accuracy improvement.
+Our proposed forecasting models combined with reconciliation approaches outperform the empirical distribution benchmark.
+We show that a substantial forecast improvement can be achieved at higher levels of hierarchical and grouped structure time series by applying forecast reconciliation methods.
+When a point forecast is of interest at the bottom level of series, we observe that reconciliation may not improve the forecast accuracy, if the bottom series are noisy and lack systematic patterns.
+However, forecast reconciliation may result in more forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e.
+point forecast).
+Producing consistent forecast are also crucial for informing planning activities, we also demonstrate that proposed models produce consistent forecasts across all forecast horizons.
+Therefore, we recommend that forecast recociliation approaches to be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted.
+Moreover, we found that using an ensemble forecasting model, combining al models developed in this paper, instead of using each individualy, works remarkably well for our mixed hierarchical & grouped stucture.
+
+Further research could investigate the practical benefits of probabilistic hierarchical forecasting in EMS.
+Linking forecasts with its utilities (response time, resource utilisation, etc) can offer an opportunity to maximise benefits through more holistic planning approach.
+While, we generated count attend incident for the base forecast usign Poisson regression models, however the reconciles forecast is snot yet count.
+This could be also an avenue for further research.

### 2022-11-17 manuscript/main.tex
first draft of results and conclusion

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-\textbf{Poisson Regression using tscount:} We consider another Poisson
+\textbf{Poisson Regression using tscount: } We consider another Poisson
 regression model that takes into account serial dependence in addition
 to covariates used in the GLM model. To that end, we use
-\texttt{tsglm()} function in \emph{tscount} package in R to model the
-attended incidents. The logarithmic link function is used to ensure that
-the parameter of Poisson distribution is always positive. This model can
-be estimated via maximization of the likelihood function based on
-Poisson mass function. The regression model captures the short range
-serial dependence by including the three order autoregressive terms.
+\texttt{tsglm()} function in \emph{tscount} package in R
+\citep{JSSv082i05} to model the attended incidents. The logarithmic link
+function is used to ensure that the parameter of Poisson distribution is
+always positive. This model can be estimated via maximization of the
+likelihood function based on Poisson mass function. The regression model
+captures the short range serial dependence by including the three order
+autoregressive terms.
 
@@
 hierarchical and grouped structure. This is also the case as discussed
-in Section~\ref{sec-data}
+in Section~\ref{sec-data}.
 
@@
 
-\hypertarget{point-forecast}{%
-\subsection{Point forecast}\label{point-forecast}}
-
-\hypertarget{probabilistic-forecast}{%
-\subsection{Probabilistic forecast}\label{probabilistic-forecast}}
-
-\begin{figure}
+In this section, we compare the forecasting performance of Naive, ETS,
+GLM, Tscount and the ensemble model using base forecast and Minimum
+Trace (Mint) methods. We have also computed the forecast accuracy for
+Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches.
+However, they are not reported here because their accuracy is
+outperformed by Mint. We should also note that forecasts, and
+consequently their corresponding errors, are generated for the entire
+hierarchy and they could be reported at any level, if required.
+
+The overall forecasting performance is reported in
+Table~\ref{tbl-result}, in which the average forecast accuracy per
+model, method and the hierarchical level is presented. Reported forecast
+accuracy are averaged across all forecast horizons, rolling origins and
+series at each level. Table~\ref{tbl-result} presents both point and
+probabilistic forecast accuracy at total, control area, health board and
+bottom level series. Point forecast performance are reported using MASE
+and RMSSE in Table~\ref{tbl-result-1} and Table~\ref{tbl-result-2},
+respectively. Probabilistic forecast accuracy is reported using CRPS in
+Table~\ref{tbl-result-3}. The bold entries in each table identify a
+combination of method and model that performs best for the corresponding
+level (i.e.~each column), based on the smallest values of accuracy
+measures.
+
+Table~\ref{tbl-result-1} shows that forecast reconciliation (i.e.~Mint)
+improves forecast accuracy at the higher levels of the hierarchy
+including total, control area and health board. However, it is does not
+result in accuracy improvement at the bottom level series, fro which
+base forecasts are more accurate. This might be due to the noisy
+structure of time series at the bottom level and the fact that base
+forecasts produced for the bottom of hierarchy are from well-specified
+forecasting models and therefore does not gain benefit from forecast
+reconciliation. Generally, if the bottom-level series are sufficiently
+smooth, base forecast is expected to outperform hierarchical
+reconciliation due to the better forecasts available at the level. It is
+also clear from Table~\ref{tbl-result-1} that the ensemble method
+improves forecast accuracy at total, control area and health board.
+However, this does not remain valid for bottom series where the Poisson
+regression by GLM outperforms other approaches.
+
+Table~\ref{tbl-result-2} also reports point forecast accuracy but using
+RMSSE. The outperformance of forecast reconciliation using Mint at
+higher levels of hierarchy is also confirmed by RMSSE. However, for the
+bottom series, we observe that the base forecasts generated by ETS are
+the most accurate, followed by its reconciliations forecasts.
+
+\begin{table}
+
+\caption{\label{tbl-result}Average forecast perfomance calculated on
+out-of-sample with time series cross validation applied to attended
+incident data. The best approach is highlighted in
+bold.}\begin{minipage}[t]{\linewidth}
+\subcaption{\label{tbl-result-1}Point forecast accuracy using MASE }
+
+{\centering
+
+\tabularnewline
+
+\centering
+\begin{tabular}[t]{llllll}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{4}{c}{MASE} \\
+\cmidrule(l{3pt}r{3pt}){3-6}
+method & model & Total & Control areas & Health boards & Bottom\\
+\midrule
+base & ensemble & 0.7543 & 0.8432 & 0.8648 & 1.004\\
+base & ets & 0.9360 & 0.9133 & 0.8884 & 1.031\\
+base & iglm & 0.8609 & 0.9083 & 0.9039 & \textbf{0.999}\\
+base & naiveecdf & 1.1396 & 1.0633 & 1.0389 & 1.017\\
+base & tscount & 0.8704 & 0.9109 & 0.9041 & 1.000\\
+\addlinespace
+mint & ensemble & \textbf{0.7332} & \textbf{0.8346} & \textbf{0.8621} & 2.251\\
+mint & ets & 0.8593 & 0.9001 & 0.8997 & 1.258\\
+mint & iglm & 0.8176 & 0.8805 & 0.8883 & 2.490\\
+mint & naiveecdf & 1.1395 & 1.0631 & 1.0389 & 2.648\\
+mint & tscount & 0.8218 & 0.8846 & 0.8913 & 2.516\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{minipage}%
+\newline
+\begin{minipage}[t]{\linewidth}
+\subcaption{\label{tbl-result-2}Point forecast accuracy using RMSSE }
+
+{\centering
+
+\tabularnewline
+
+\centering
+\begin{tabular}[t]{llllll}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{4}{c}{RMSSE} \\
+\cmidrule(l{3pt}r{3pt}){3-6}
+method & model & Total & Control areas & Health boards & Bottom\\
+\midrule
+base & ensemble & 0.7338 & 0.8335 & 0.8652 & 0.9885\\
+base & ets & 0.9465 & 0.9126 & 0.8915 & \textbf{0.9795}\\
+base & iglm & 0.8383 & 0.9073 & 0.9125 & 0.9958\\
+base & naiveecdf & 1.0559 & 1.0213 & 1.0188 & 1.0081\\
+base & tscount & 0.8446 & 0.9090 & 0.9114 & 1.0069\\
+\addlinespace
+mint & ensemble & \textbf{0.7153} & \textbf{0.8244} & \textbf{0.8621} & 1.2405\\
+mint & ets & 0.8510 & 0.9002 & 0.9027 & 0.9879\\
+mint & iglm & 0.8015 & 0.8814 & 0.8972 & 1.3379\\
+mint & naiveecdf & 1.0558 & 1.0211 & 1.0188 & 1.4429\\
+mint & tscount & 0.8020 & 0.8849 & 0.9003 & 1.3515\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{minipage}%
+\newline
+\begin{minipage}[t]{\linewidth}
+\subcaption{\label{tbl-result-3}Probabilistic forecast accuracy using CRPS }
+
+{\centering
+
+\tabularnewline
+
+\centering
+\begin{tabular}[t]{llllll}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{4}{c}{CRPS} \\
+\cmidrule(l{3pt}r{3pt}){3-6}
+method & model & Total & Control areas & Health boards & Bottom\\
+\midrule
+base & ensemble & 12.44 & 5.681 & 3.385 & 0.2421\\
+base & ets & 13.96 & 5.974 & 3.430 & 0.2422\\
+base & iglm & 14.49 & 6.061 & 3.488 & 0.2420\\
+base & naiveecdf & 30.35 & 10.887 & 5.445 & 0.3000\\
+base & tscount & 14.98 & 6.124 & 3.500 & 0.2429\\
+\addlinespace
+mint & ensemble & \textbf{12.09} & \textbf{ 5.664} & \textbf{3.383} & 0.2455\\
+mint & ets & 13.28 & 5.884 & 3.472 & \textbf{0.2412}\\
+mint & iglm & 13.35 & 5.812 & 3.394 & 0.2443\\
+mint & naiveecdf & 30.36 & 10.904 & 5.446 & 0.3103\\
+mint & tscount & 13.55 & 5.868 & 3.423 & 0.2483\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{minipage}%
+
+\end{table}
+
+Table~\ref{tbl-result-2} presents the accuracy of the forecast
+distributions measures by CRPS, which considers both forecasting
+reliability and interval sharpness. The smaller the value of CRPS, the
+better the comprehensive performance. We observe that forecast
+reconciliation results in forecast improvement, regardless of the
+hierarchical level. Ensemble method is also more accurate fore higher
+levels, but ETS performs better at the bottom level. While results show
+that forecast reconciliation does not improve point forecast at the
+bottom level, however Table~\ref{tbl-result-2} indicates that it
+generates more accurate forecasts than base method. The outperformance
+of probabilistic forecast reconciliation at the bottom level may
+highlight an important fact. The reconciliation method may improve
+forecast accuracy at certain number of quantiles at the tails of
+forecast distribution, which are critical for managing risks.
+
+Overall, our results indicate that forecast reconciliation using MinT
+method provides reliable forecasts, and all methods that reconcile using
+information at all levels of the forecast improve upon base
+(unreconciled) forecasts, except the bottom level series. For bottom
+series with less obvious patterns, base forecasts near center of
+forecast distribution might be more accurate, while forecast
+reconciliation using Mint could improve accuracy in the tails of of the
+distribution.
+
+In addition to the overall forecast accuracy presented in
+Table~\ref{tbl-result}, we also report the point and probabilistic
+forecast accuracy measures for each forecast horizon in
+Figure~\ref{fig-accuracy}. The figure focuses on the hierarchical levels
+important for decision making including total, control area and health
+board, however this could be calculated for any level. We only
+illustrate the results of the Mint method, given the its outperformance
+described in Table~\ref{tbl-result}. For the illustration purpose, we
+report the average weekly forecast accuracy instead of the daily
+forecast horizon (i.e.~84 days including freezed planning (7 weeks = 42
+days) plus future planing period (7 weeks = 42 days). Therefore, x-axis
+shows horizons from week 1 (h= 1,\ldots,7) to week 12 (h= 78,\ldots,84).
+The forecast horizon from week 7 to week 12 corresponds to the upcoming
+planning horizon, which is used by planners and decision making. For the
+point forecast, we can see that ensemble model performs much better than
+others regardless of forecast horizons. For the probabilistic forecasts,
+the distinction between forecasting models is less obvious. It is
+important to highlight that, all forecasting models outperform the Naive
+empirical distribution that is used as a Benchmark for both point and
+probabilsitic forecasts.
+
+\begin{figure}[H]
 
@@
 
+Forecasting problems at Emergency Medical Services have often inherent
+hierarchical and grouped structures. For example, looking at time series
+of arrival calls in a clinical desk service, verified incidents, or
+attended incidents in a country, they could be disaggregated by various
+attributes of interest. Total demand in the country could be
+disaggregated by region, then within each region by health board, within
+each health board, by station/hospital, and so on down to the post code
+area. Alternative structure may arise when attributes of interest are
+crossed rather than nested. For example, the total demand could be
+disaggrgated by priority (e.g Red, Amber, Green) or nature of incidents.
+It is also natural to have a mixed structures, for example the total
+demand could be disaggregated by priority and also by health board.
+
+Despite the inherent hierarchical structure of the forecasting problem
+in EMS, the common practice is to produce point forecast for each time
+series independently. This practice may lead to the lack of coordination
+and possibly undesirable and conflicting outcomes. Furthermore, due to
+the asymmetric impact of resource allocation in this area, quantifying
+forecast uncertainty through probabilistic forecasts is also of value as
+it enables planners to manage associated risks. In this paper, we
+investigate the application of hierarchical forecasting methods for
+producing a probabilistic forecast of daily incident attended up to 84
+days ahead, using different forecasting methods.
+
+Our results indicates that forecast reconciliation can not only
+contribute to a more coordinated efforts through producing coherent
+forecasts when it comes to using forecasting to inform decisions in EMS,
+but also it can result in forecast accuracy improvement. Our proposed
+forecasting models combined with reconciliation approaches outperform
+the empirical distribution benchmark. We show that a substantial
+forecast improvement can be achieved at higher levels of hierarchical
+and grouped structure time series by applying forecast reconciliation
+methods. When a point forecast is of interest at the bottom level of
+series, we observe that reconciliation may not improve the forecast
+accuracy, if the bottom series are noisy and lack systematic patterns.
+However, forecast reconciliation may result in more forecast results for
+bottom series, if we are interested in the tails of forecast
+distribution rather than just center measures like mean (i.e.~point
+forecast). Producing consistent forecast are also crucial for informing
+planning activities, we also demonstrate that proposed models produce
+consistent forecasts across all forecast horizons. Therefore, we
+recommend that forecast recociliation approaches to be adopted for
+routine use in EMS, whenever hierarchical and/or grouped time series
+data need to be forecasted. Moreover, we found that using an ensemble
+forecasting model, combining al models developed in this paper, instead
+of using each individualy, works remarkably well for our mixed
+hierarchical \& grouped stucture.
+
+Further research could investigate the practical benefits of
+probabilistic hierarchical forecasting in EMS. Linking forecasts with
+its utilities (response time, resource utilisation, etc) can offer an
+opportunity to maximise benefits through more holistic planning
+approach. While, we generated count attend incident for the base
+forecast usign Poisson regression models, however the reconciles
+forecast is snot yet count. This could be also an avenue for further
+research.
+
 \renewcommand\refname{References}

### 2022-11-17 manuscript/02-literature.qmd

@@
-# Research background {#lit}
+# Research background {#sec-lit}
+
 Emergency medical services (EMS) are a critical component in the delivery of urgent medical care to communities.
-An effective service delivery requires accurate resource planning that are generally replying on demand forecasts at operational, tactical and strategic levels.
+An effective service delivery requires accurate resource planning that generally relies on demand forecasts at operational, tactical and strategic levels.
 
 There is a substantial number of studies on the application of time series forecasting in the Emergency Medical Services.
-Various areas have been the focus of the literature.
-Forecasting call volume arrivals is one of the major research topics.
-@ibrahim2016modeling provide an extensive review of the forecasting models in this context.
+For example, @ibrahim2016modeling provide an extensive review of the models used in forecasting call volume arrivals.
 Another important area is related to forecasting ambulance demand.
-Although the definition of demand might not be always clearly stated, however, this is typically referring to a situation where a physical resource has been deployed to respond to an incident.
+Although the definition of demand might not be always clearly stated, this is typically referring to a situation where a physical resource has been deployed to respond to an incident.
 This might be also called _attended incidents_.
-Another demand related variable is verified incidents.
-These are all incidents that require an action: either send a physical vehicle, deal with via the Clinical Support Desk (e.g.
-calls), get an external (private) provider to respond to it, or send it through to other channels such as police, firefighters or general practitioners.
+Another demand related variable is verified incidents; these are all incidents that require an action: either by sending a physical vehicle, responding via the Clinical Support Desk, requesting an external provider to respond to it, or forwarding it to other channels such as police, firefighters or general practitioners.
 Our study is aligned with this stream of the literature.
-Another similar area that is largely studied in the literature, is Emergency Department forecasting.
-We refer interested readers to @mingliterature2022, @gul2020exhaustive and @wargon2009systematic for extensive reviews of the literature on Emergency Department forecasting.
+Another similar area that has received considerable attention is Emergency Department forecasting; we refer interested readers to @mingliterature2022, @gul2020exhaustive and @wargon2009systematic for extensive reviews of the relevant literature.
 Although crucial to EMS performance, @aringhieri2017emergency state that demand forecasting has received limited research attention in the EMS context.
@@
 
-There are generally two main streams of research related to forecasting ambulance demand in EMS: i) the first stream focuses on the application of time series methods and regression approaches on forecasting aggregate ambulance demand [@vile2012predicting; @sasaki2010using]; ii) the second stream considers forecasting EMS demand in a more finer temporal and geographical granularities by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit].
+There are generally two main streams of research related to forecasting ambulance demand in EMS: (i) the first stream focuses on the application of time series methods and regression approaches to forecasting aggregate ambulance demand [@vile2012predicting; @sasaki2010using]; and (ii) the second stream considers forecasting EMS demand in finer temporal and geographical granularities by employing temporal-spatial prediction methods [@zhou2016predicting;@zhou2016predictinglit].
 The focus of our study is related to the first stream of research.
@@
 In addition to the historical demand, the population census for different age groups and counts of the number of companies employing more than five people are included in the regression.
-The census variables describe groups who may be more likely to need an ambulance.
-A stepwise ordinary least squares regression analysis with SPSS is used for estimating the parameter and generating forecast.
-The only performance measure reported in this study is $R^2$.
+The census variables describe groups who are more likely to need an ambulance.
+A stepwise ordinary least squares regression analysis is used for estimating the parameter and generating forecast.
+The only performance measure reported in this study is $R^2$, which is not an effective measure of forecast accuracy [@armstrong01, p457].
 The research design of this study is not rigorous and the study is not reproducible.
 @vile2012predicting explore using a Singular Spectrum Analysis (SSA) method to generate forecasts of the EMS demand at the national level for 7-day, 14-day, 21-day and 28-day forecast horizons using data provided by the Welsh Ambulance Service Trust (WAST).
-The performance of this approach is compared with Auto-Regressive Moving Average (ARIMA) and Holt-winter time series methods using Root Mean Squared Error (RMSE).
-They concluded that point forecasts generated by SSA are more accurate for longer-term, but ARIMA and Holt-winter performance is superior for shorter-term horizons.
+The performance of this approach is compared to Auto-Regressive Integrated Moving Average (ARIMA) and Holt-Winters time series methods using Root Mean Squared Error (RMSE).
+They concluded that point forecasts generated by SSA are more accurate for longer-term, but that ARIMA and Holt-Winters performance is superior for shorter-term horizons.
 @vile2016time further develop a decision support system to integrate forecasts generated by SSA.
 However, the study does not compare and contrast the performance of forecasting methods based on utility measures such as cost, resource utilization or response time.
-The tool contains options that allow generating forecasts at various levels of granularity, however, it ignore the hierarchical and grouped relationships structure, preventing  aligned decision making and coordination.
+The tool contains options that allow generating forecasts at various levels of granularity, however, it ignores the hierarchical and grouped relationships structure, preventing aligned decision making and coordination.
 
-@9659837 investigate forecasting EMS demand in a high spatio-temporal resolution of 1×1km spatial regions and 1-hr time intervals using total incidents in Oslo, Norway from January 1st, 2015 up to and including February 11th, 2019.
-They used multi-layer perceptron (MLP) and long short-term memory (LSTM) models to forecast the EMS demand, and compare them to simple aggregation methods and baselines.
-The point forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean Squared Error (MSE) and the forecast distribution is measures by Categorical Cross-Entropy.
-They shows that while Neural Network models perform better in producing point forecast, a distribution baseline method based on spatial distribution of the incidents across all time steps provide more accurate forecast distribution.
-@zhou2016predictinglit propose three methods based on Gaussian mixture models, kernel density estimation, and kernel warping to predict 4 weeks into future for a 1-km2 spatial region over an hour.
-Two years of incidents attended from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011 and 2012 with  696,975 events) are used to build the model and examine the  performance on test data using mean negative log likelihood.
-They show that forecasts generated by the proposed methods are significantly more accurate than the current industry practice,  a simple averaging formula.
-@grekousis2019will investigate the combination of spatial analysis methods with data mining techniques based on  an improved Hungarian algorithm and MLP neural network to identify the most likely locations of future emergency events.
-The proposed approach is tested using data of 2851 events attended by the EMS in Athens, Greece over 24 weeks.
-They show that $23.24\%$ of real emergency events lie within 50 meter of the predicted ones and nearly $70\%$ of the real emergency events lie no further away than 150 meter, which is rather accurate given the granularity of the problem at the city level.
+@9659837 investigate forecasting EMS demand in a high spatio-temporal resolution of 1km$^2$ spatial regions and 1-hr time intervals using total incidents in Oslo, Norway, from 1 January 2015 to 11 February 2019.
+They used multi-layer perceptron (MLP) and long short-term memory (LSTM) models to forecast the EMS demand, and compare the results to simple aggregation methods and baselines.
+The point forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean Squared Error (MSE), and the forecast distribution is measured by Categorical Cross-Entropy.
+They found that Neural Network models performed better in producing point forecasts, while a distribution baseline method based on spatial distribution of the incidents across all time steps provided more accurate forecast distributions.
+@zhou2016predictinglit proposed three methods based on Gaussian mixture models, kernel density estimation, and kernel warping to predict hourly data 4 weeks ahead for a 1km$^2$ spatial region.
+Two years of incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011 and 2012 with 696,975 events) were used to build the model and examine the performance on test data using mean negative log likelihood.
+They show that forecasts generated by the proposed methods were significantly more accurate than the current industry practice (a simple averaging formula).
+@grekousis2019will investigated the combination of spatial analysis methods with data mining techniques based on an improved Hungarian algorithm and a MLP neural network to identify the most likely locations of future emergency events.
+The proposed approach was tested using data of 2851 events attended by the EMS in Athens, Greece, over 24 weeks.
+They showed that 23% of real emergency events lie within 50 meters of the predicted ones and nearly 70% of the real emergency events lie no further than 150 meters away, which is rather accurate given the granularity of the problem at the city level.
 
-We observed a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research. These limitations are summarized as following:
+We note a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research. These limitations are summarized as following:
 
 1.
-Current studies ignore the inherent hierarchical and/or grouped structure of the time series data and the relationship between series at different levels of hierarchy.
+Current studies ignore the inherent hierarchical and/or grouped structure of the time series data, and the relationship between series at different levels of hierarchy.
 This may result in incoherent forecasts leading to misaligned planning and decision making.
@@
 2.
-Current research is mainly concerned with generating point forecast at a single level of hierarchy.
-There is a lack of studies presenting the entire forecast distribution of daily ambulance demand for the entire hierarchy to inform the whole decision-making process and to better represent the uncertainty of future demand, providing a risk management tool for planners.
+Current research is mainly concerned with generating point forecasts at a single level of hierarchy.
+There is a lack of studies considering the entire forecast distribution of daily ambulance demand for the whole hierarchy to inform the decision-making process and to better represent the uncertainty of future demand, providing a risk management tool for planners.
 
-3. Reproducibility is still a major challenge in EMS forecasting, as it is unlikely to reproduce the results without the help of the authors of those papers.
+3. Reproducibility is still a major challenge in EMS forecasting, as it is unlikely that any reader can  reproduce prior studies without the help of the authors of those papers.
 
 4.
-Another limitation is related to the generated forecasts that are not integer counts.
-Since actual ambulance counts cannot be negative or fractions, ambulance demand forecasts should be the same.
+Another limitation is related to the generated forecasts not being on the sample space of non-negative counts.
+Since actual ambulance counts cannot be negative or non-integer, ambulance demand forecast distributions should reflect the data.
+Of course, point forecasts represent means, so they should be non-negative, but may be non-integer.
 While this might not be an issue when producing forecasts at a single level, producing non-negative count forecasts in a hierarchical/grouped structure is challenging and requires further investigation in the future.

### 2022-11-17 manuscript/03-experiment.qmd
typo and removing pdf

--- 2022-11-17 manuscript/03-experiment.qmd
@@
 
-In addition to displaying the trend and seasonality features, we also visualize a few time series at various levels of the aggregation.
+In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation.
 @fig-dataviz2 reveals different information such as trend, seasonality and noise.
@@
 Model development and hyper-parameter tuning is performed using the training data only.
-Forecasting performance is evaluated using both point and probabilistic error measures.
 While we compute forecast errors for the entire 12 weeks, we are most interested in the last 42 days of each test set, because that corresponds to how forecasts are generated for planning in practice.

### 2022-11-17 manuscript/05-results.qmd
typo and removing pdf

@@
-# Results and discussion {#sec-result}
+# Results and discussion {#sec-results}
 
@@
 
-@tbl-result-2 also reports point forecast accuracy but using RMSSE.
-The outperformance of forecast reconciliation using Mint at higher levels of hierarchy is also confirmed by RMSSE.
-However, for the bottom series, we observe that the base forecasts generated by ETS are the most accurate, followed by its reconciliations forecasts.
+@tbl-result-2 also reports point forecast accuracy using RMSSE.
+The outperformance of forecast reconciliation using Mint at higher levels of hierarchy is also confirmed by this measure.
+However, for the bottom series, we observe that the base forecasts generated by ETS are the most accurate one, followed by its reconciliations forecasts.
 

### 2022-11-17 manuscript/06-conclusion.qmd
typo

@@
 Producing consistent forecast are also crucial for informing planning activities, we also demonstrate that proposed models produce consistent forecasts across all forecast horizons.
-Therefore, we recommend that forecast recociliation approaches to be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted.
-Moreover, we found that using an ensemble forecasting model, combining al models developed in this paper, instead of using each individualy, works remarkably well for our mixed hierarchical & grouped stucture.
+Therefore, we recommend that forecast reconciliation approaches to be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted.
+Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individualy, works remarkably well for our mixed hierarchical & grouped structure.
 
 Further research could investigate the practical benefits of probabilistic hierarchical forecasting in EMS.
-Linking forecasts with its utilities (response time, resource utilisation, etc) can offer an opportunity to maximise benefits through more holistic planning approach.
-While, we generated count attend incident for the base forecast usign Poisson regression models, however the reconciles forecast is snot yet count.
+Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through more holistic planning approach.
+While, we generated count attend incident for the base forecast using Poisson regression models, however the reconciles forecast is snot yet count.
 This could be also an avenue for further research.

### 2022-11-18 letter/cover_letter.qmd
add cover letetr and title page

---
author:
  - name: Bahman Rostami Tabar & Rob J. Hyndman 
    affiliation:
      - name: Bahman Rostami-Tabar
        department: Cardiff Business School
        address: 3, Colum Drive
        city: Cardiff
        country: United Kingdom
        postal-code: CF10 3EU
address:
  - Prof. Donald M. Yealy 
  - Editor-in-Chief
  - Annals of Emergency Medicine
subject: Paper submission for peer review
opening: "Dear Prof. Yealy ,"
closing: "Yours Faithfully,"
format: letter-pdf
---

We would be grateful if you could consider for publication our manuscript entitled "Hierarchical Time Series Forecasting in Emergency Medical Services" as an Original Research article in Annals of Emergency Medicine.

Time series forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures. For example the ambulance demand (e.g. arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area. Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents, .

In this paper, we aim at filling various gaps identified in the literature, and our contributions to the literature are summarised as follows

Thank you very much for your consideration of this manuscript. We are looking forward to hearing from you in due course.

### 2022-11-18 manuscript/title_page.qmd
add cover letetr and title page

--- 2022-11-17 manuscript/main.qmd
@@
   - ambulance demand
-  - attended incidents
-bibliography: bibliography.bib
+  - regression
 execute:
@@
 
-```{r}
-#| label: setup
-#| context: setup
-#| include: false
-library(fpp3)
-library(knitr)
-library(kableExtra)
-library(ggthemes)
-library(patchwork)
-# Color blind color palettes
-options(
-  ggplot2.discrete.colour = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
-  ggplot2.discrete.fill = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
-)
-# Few theme
-ggplot2::theme_set(ggthemes::theme_few())
-```
+- **Funding information:** A funding awarded by Cardiff Business School allowed BRT to travel to Melbourne to visit RJH to discuss the research design, run the experiment and write the article.
 
-```{r intro, child="01-introduction.qmd"}
-
-```
-
-```{r literature, child="02-literature.qmd"}
-
-```
-
-```{r experiment, child="03-experiment.qmd"}
-
-```
-
-```{r htc, child="04-htsgts.qmd"}
-
-```
-
-```{r results, child="05-results.qmd"}
-
-```
-
-```{r conclusion, child="06-conclusion.qmd"}
-
-```
-
-# References {#references .unnumbered}
+- **Author contributions:** BRT and RJH conceptualize the the study and designed the experiment.
+BRT obtained a funding to meet RHH at Monash University.
+BRT and RJH provided the analysis of the data.
+RJH developed the models; ML chaired the data oversight committee.
+BRT drafted the manuscript, and RJH contributed substantially to its revision.
+BRT takes responsibility for the paper as a whole.

### 2022-11-18 letter/cover_letter.qmd
Updated letter

@@
 author:
-  - name: Bahman Rostami Tabar & Rob J. Hyndman 
+  - name: Bahman Rostami Tabar & Rob&nbsp;J.&nbsp;Hyndman
     affiliation:
@@
 subject: Paper submission for peer review
-opening: "Dear Prof. Yealy ,"
-closing: "Yours Faithfully,"
+opening: "Dear Prof. Yealy,"
+closing: "Yours sincerely"
 format: letter-pdf
@@
 arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area.
-Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents, .
+Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents.
 

### 2022-11-18 manuscript/main.qmd
Removed Welsh

@@
 To address this issue, we exploit the hierarchical and grouped structure of the demand time series, and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation.
-The methods are applied to daily incident data from the Welsh Ambulance Service NHS Trust, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area.
+The methods are applied to daily incident data from an ambulance service in Great Britain, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area.
 We use an ensemble of forecasting models, and show that the resulting forecasts are better than any individual forecasting model.
@@
   - ambulance demand
-  - attended incidents
+  - regression
 bibliography: bibliography.bib

### 2022-11-18 letter/cover_letter.qmd
cover letter and title page draft

--- 2022-11-18 letter/cover_letter.qmd
@@
 
-We would be grateful if you could consider for publication our manuscript entitled "Hierarchical Time Series Forecasting in Emergency Medical Services" as an Original Research article in Annals of Emergency Medicine.
+We would be grateful if you could consider for publication our manuscript entitled _Hierarchical Time Series Forecasting in Emergency Medical Services_ as an Original Research article in Annals of Emergency Medicine.
 
@@
 arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area.
-Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents, .
+Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents.
+Despite the popularity of such data structure in EMS, the common practice is to produce point forecast for each time series independently and ignore the structure.
+This practice may lead to the lack of coordination and possibly undesirable and conflicting outcomes.
+Practices also rely on point forecasts, which does not  acknowledge forecast uncertainty.
 
-In this paper, we aim at filling various gaps identified in the literature, and our contributions to the literature are summarised as follows
+In this paper, we aim at filling these gaps identified in the academic literature and practice.
+We produce probabilistic forecasts of daily incident attended up to 84 days ahead using hierarchical forecasting methods for using state of the art forecasting methods.
+We apply hierarchical forecasting methods to produce a probabilistic and coherent forecast of attended incidents at multiple levels of aggregation.
+We use Exponential Smoothing State Space models, Poisson regression based on General Linear Models and tscount models, a Naive empirical distribution, and the ensemble model to generate base forecasts.
+These forecasting models collectively account for i) the spline trend, ii) multiple seasonalities, iii) holiday effects, and  iv) serial dependencies in the time series.
+Our study is fully reproducible and we provide data and code through a Github repository enabling reproduction and refinement of the proposed approach and benchmarks.
+The proposed approach could also be generalised to forecast requirements in other healthcare services, such as admissions in Emergency Departement, or call volumes in clinical desk services.
 

### 2022-11-18 manuscript/title_page.qmd
cover letter and title page draft

--- 2022-11-18 manuscript/title_page.qmd
@@
 - **Author contributions:** BRT and RJH conceptualize the the study and designed the experiment.
-BRT obtained a funding to meet RHH at Monash University.
-BRT and RJH provided the analysis of the data.
-RJH developed the models; ML chaired the data oversight committee.
+BRT obtained a funding to meet RHH at Monash University in Melbourne.
+BRT and RJH provided the analysis of the time series data.
+RJH developed forecasting models and hierarchical methods; BRT and RJH use R programming to conduct analysis.
 BRT drafted the manuscript, and RJH contributed substantially to its revision.

### 2022-11-18 manuscript/02-literature.qmd

@@
 The research design of this study is not rigorous and the study is not reproducible.
-@vile2012predicting explore using a Singular Spectrum Analysis (SSA) method to generate forecasts of the EMS demand at the national level for 7-day, 14-day, 21-day and 28-day forecast horizons using data provided by the Welsh Ambulance Service Trust (WAST).
+@vile2012predicting explore using a Singular Spectrum Analysis (SSA) method to generate forecasts of the EMS demand at the national level for 7-day, 14-day, 21-day and 28-day forecast horizons using data provided by an ambulance service in Great Britain.
 The performance of this approach is compared to Auto-Regressive Integrated Moving Average (ARIMA) and Holt-Winters time series methods using Root Mean Squared Error (RMSE).

### 2022-11-18 manuscript/04-htsgts.qmd
Spelling
typo

--- 2022-10-23 manuscript/04-htsgts.qmd
@@
-# Hierachcail and grouped time series forecasting techniques {#htc}
+# Hierarchical and grouped time series forecasting techniques {#sec-htc}
 
-Moreover, both models either underestimated or overestimated the number of calls.
-Authors suggested that the inability to capture a positive trend resulted in underestimation, and the opposite was due to a negative trend.
-Therefore, authors recommended that further research should add certain temporal variables able to capture the trend.
+There are many applications in healthcare, and in particular in EMS, where a collection of time series is available.
+These series are generally hierarchically organized based on multiple levels such as area/region, health board and/or are aggregated at different levels in groups based on nature of demand, priority of demand, or some other attributes.
+While series could be strictly hierarchical or only grouped bases on some attributes, in many situation a more complex structures arise when attributes of interest are both nested and crossed, having hierarchical and grouped structure.
+This is also the case for our application as discussed in @sec-data.
 
-## Notation
+## Independent (base forecast)
 
-## Base forecast
+A common practice in healthcare (and EMS) to predict hierarchical and grouped series relies on producing independent forecasts, also refereed to as base forecasts, typically by different teams as the need for such forecasts arise.
+We observe $n$ time series at time $t$, across the entire hierarchical and grouped structure, written as $y_t$.
+The base forecasts of $y_{T+h}$ given data $y_1,\dots,y_T$ are denoted by $\hat{y}_h$ for $h$ steps-ahead for all $n$ series ($n=1530$ in this study).
+Forecasts generated in this way are not coherent.
 
-## Bottom-up
+## Reconciliation methos
 
-## Optimal reconciliation
+Traditionally, approaches to produce coherent forecasts for hierarchical and grouped time series involve using bottom-up and top-down methods by generating forecasts at a single level and then aggregating or disaggregating.
+Top-down methods require having a unique hierarchical structure to disaggregate forecasts generated at the top level by proportions.
+However, given that we have multiple grouped attributes combined with the hierarchical structure, there is no unique way to disaggregate top forecasts.
+Hence the top-down cannot be used in our application.
+The recommended approach is to use forecast reconciliation [@hyndman2011optimal].
+In the following sections, we first discuss some notation, and then present bottom-up and forecast reconciliation approaches used in this study to generate coherent forecasts.
+
+### Notations
+
+Let $\bm{b}_t$ be a vector of $n_b$ "bottom-level" time series at time $t$, and let $\bm{a}_t$ be a corresponding vector of $n_a = n-n_b$ aggregated time series, where
+$$
+  \bm{a}_t = \bm{A}\bm{b}_t,
+$$
+and $\bm{A}$ is the $n_a\times n_b$ "aggregation" matrix specifying how the bottom-level series $\bm{b}_t$ are to be aggregated to form $\bm{a}_t$. The full vector of time series is given by
+$$
+ \bm{y}_t = \begin{bmatrix}\bm{a}_t \\\bm{b}_t\end{bmatrix}.
+$$
+This leads to the $n\times n_b$ "summing" or "structural" matrix given by
+$$
+  \bm{S} = \begin{bmatrix}\bm{A} \\ \bm{I}_{n_b}\end{bmatrix}
+$$
+such that $\bm{y}_t = \bm{S}\bm{b}_t$.
+
+### Bottom-up (BU) and linear reconciliation methods
+
+Bottom-Up is a simple approach to generate coherent forecasts.
+It involves first creating the base forecasts for the bottom level series (i.e., the most disaggregated series).
+These forecasts are then aggregated to the upper levels which naturally results in coherent forecasts.
+The BU approach can capture the dynamics of the series at the bottom level, but these series may be noisy and difficult to forecast.
+The approach using only the data at the most disaggregated level, and so does not utilize all the information available across the hierarchical and grouped structure.
+
+Forecast reconciliation approaches fill this gap by combining and reconciling all the base forecasts in order to produce coherent forecasts.
+Linear reconciliation methods can be written [@WicEtAl2019] as
+$$
+  \tilde{\bm{y}}_h = \bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{W}^{-1}\hat{\bm{y}}_h,
+$$
+where $\bm{W}$ is an $n \times n$ positive definite matrix, and $\hat{\bm{y}}_h$ contains the $h$-step forecasts of $\bm{y}_{T+h}$ given data to time $T$.
+Different choices for $\bm{W}$ lead to different solutions such as Ordinary Least Squares (OLS), Weighted Least Squares (WLS) and Minimum Trace (MinT).
+We use the implementation of these methods in the fable package in R in the experiment.

### 2022-11-18 manuscript/03-experiment.qmd
sec 5 and 6

@@
 
-**Poisson Regression using tscount:** We also consider another Poisson regression model that takes into account serial dependence, using the `tsglm()` function in the _tscount_ package in R [@JSSv082i05] to model the attended incidents.
+**Poisson Regression using tscount (TSGLM):** We also consider another Poisson regression model that takes into account serial dependence, using the `tsglm()` function in the _tscount_ package in R [@JSSv082i05] to model the attended incidents.
 Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.
 This model captures the short range serial dependence by including three autoregressive terms, in addition to the same covariates that were used in the GLM model.
-To distinguish this from the previous GLM model, we will refer to this model as "tscount".
+To distinguish this from the previous GLM model, we will refer to this model as "TSGLM".
 
-**Ensemble method:** Finally, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and tscount models to form a mixture distribution [@combinations].
+**Ensemble method:** Finally, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models to form a mixture distribution [@combinations].
 

### 2022-11-18 manuscript/05-results.qmd
sec 5 and 6

--- 2022-11-16 manuscript/05-results.qmd
@@
-# Results and discussion {#sec-result}
+# Results and discussion {#sec-results}
 
-## Point forecast
+```{r}
+#| include: false
+rmsse <- readr::read_rds(here::here("results/rmsse.rds")) |> 
+  mutate(msse = rmsse^2)
+mase <- readr::read_rds(here::here("results/mase.rds"))
+crps <- readr::read_rds(here::here("results/crps.rds")) |>
+  pivot_wider(names_from = model, values_from = crps) |>
+  group_by(method, h, series) |>
+  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
+  ungroup()
+```
 
-## Probabilistic forecast
+In this section, we compare the forecasting performance of the Naive, ETS, GLM and TSGLM models along with the ensemble, using base forecast and Minimum Trace (MinT) reconciliation methods.
+We have also computed the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches, along with bottom up forecasting.
+However, they are not reported here because their accuracy is outperformed by MinT.
+We should also note that forecasts, and consequently their corresponding errors, are generated for the entire hierarchy and they could be reported at any level, if required.
+But to save space, we have reported only the top level (Total), the bottom level, and the levels corresponding to Control areas and Health boards.
+The latter are chosen because this is where decision-making takes place, so these forecasts are the most important.
+
+The overall forecasting performance is reported in @tbl-result, in which the average forecast accuracy over horizons 43--84 days (corresponding to the planning horizon) is presented per model, method and the hierarchical level.
+Reported forecast accuracy are averaged across all forecast horizons, rolling origins and series at each level.
+@tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom level series.
+Point forecast performance are reported using MASE and MSSE in @tbl-result-1  and @tbl-result-2, respectively.
+Probabilistic forecast accuracy is reported using CRPS in @tbl-result-3.
+The bold entries in each table identify a combination of method and model that performs best for the corresponding level (i.e.
+each column), based on the smallest values of accuracy measures.
+
+@tbl-result-1 and 2b show that forecast reconciliation (i.e.
+MinT) improves forecast accuracy at the higher levels of the hierarchy including total, control area and health board.
+However, it is does not result in accuracy improvement at the bottom level series, for which base forecasts are more accurate.
+This might be due to the noisy structure of time series at the bottom level, and possibly due to very different patterns in the aggregated series.
+It is also clear from @tbl-result-1 that the ensemble method improves forecast accuracy at total, control area and health board.
+However, this does not remain valid for bottom series where different individual methods perform best, depending on the accuracy measure.
+
+```{r}
+#| label: tbl-result
+#| tbl-cap: "Average forecast perfomance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data.
+The best approach is highlighted in bold."
+#| tbl-subcap:
+#|   - "Point forecast accuracy using MASE"
+#|   - "Point forecast accuracy using MSSE"
+#|   - "Probabilistic forecast accuracy using CRPS"
+#| layout-nrow: 3
+#Set minimum in column to bold
+set_to_bold <- function(table) {
+  for(i in 3:6) {
+    best <- which.min(table[[i]])
+    table[[i]] <- sprintf(table[[i]], fmt="%1.3f")
+    table[[i]][best] <- cell_spec(table[[i]][best], "latex", bold=TRUE)
+  }
+  return(table)
+}
+
+msse1 <- rmsse |>
+  filter(
+    method %in% c("mint", "base"),
+    model != "qcomb",
+    h > 42
+  ) |>
+  mutate(model = factor(model,
+    levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
+    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
+    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT")),
+  ) |>
+  group_by(method, model, series) |>
+  summarise(msse = mean(msse), .groups = "drop") |>
+  pivot_wider(names_from = series, values_from = msse) |>
+  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
+  set_to_bold()
+mase1 <- mase |>
+  filter(
+    method %in% c("mint", "base"),
+    model != "qcomb",
+    h > 42
+  ) |>
+  mutate(
+    model = factor(model,
+      levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
+      labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
+    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
+  ) |>
+  group_by(method, model, series) |>
+  summarise(mase = mean(mase), .groups = "drop") |>
+  pivot_wider(names_from = series, values_from = mase) |>
+  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
+  set_to_bold()
+crps1 <- readr::read_rds(here::here("results/crps.rds")) |>
+  filter(method %in% c("mint", "base"),
+         model != "qcomb",
+         h > 42
+  ) |>
+  mutate(model = factor(model,
+    levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
+    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
+    method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
+  ) |>
+  group_by(method, model, series) |>
+  summarise(crps = mean(crps), .groups = "drop") |>
+  pivot_wider(names_from = series, values_from = crps) |>
+  select(Method = method, Model = model, Total, `Control areas`, `Health boards`, Bottom) |>
+  set_to_bold()
+
+kbl(mase1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
+  add_header_above(c(" " = 2, "MASE" = 4)) |>
+  kable_styling(latex_options = c("hold_position"))
+
+kbl(msse1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
+  add_header_above(c(" " = 2, "MSSE" = 4)) |>
+  kable_styling(latex_options = c("hold_position"))
+
+kbl(crps1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
+  add_header_above(c(" " = 2, "CRPS" = 4)) |>
+  kable_styling(latex_options = c("hold_position"))
+```
+
+@tbl-result-3 presents the accuracy of the forecast distributions measures by CRPS, which considers both forecasting reliability and interval sharpness.
+The smaller the value of CRPS, the better the comprehensive performance.
+We observe that forecast reconciliation results in forecast improvement, regardless of the hierarchical level.
+The ensemble method is also more accurate for higher levels, but ETS performs slightly better at the bottom level.
+While these results show that forecast reconciliation does not improve *point* forecasts at the bottom level, @tbl-result-3 indicates that it generates more accurate *distributional* forecasts than the base method.
+This is probably due to the reconciliation method giving improved forecast accuracy in the tails of the forecast distribution, which are critical for managing risks.
+
+Overall, our results indicate that forecast reconciliation using the MinT method provides reliable forecasts, improves upon the base (unreconciled) forecasts all levels except the bottom level series.
+But even there, forecast reconciliation using MinT improves accuracy in the tails of the distribution.
+
+In addition to the overall forecast accuracy presented in @tbl-result, we also report the point and probabilistic forecast accuracy measures for each forecast horizon in @fig-accuracy.
+The figure focuses on the hierarchical levels important for decision-making including total, control area and health board; however the accuracy could be calculated for any level.
+We only illustrate the results of the MinT method, given its strong performance described in @tbl-result.
+For illustration purposes, we report the average weekly forecast accuracy instead of the daily forecast horizon, as this reduces the visual noise in the figure.
+Thus the x-axis shows horizons from week 1 ($h= 1,\dots,7$) to week 12 ($h= 78,\dots,84$).
+The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision making.
+For both the point forecast and distributional accuracy  we can see that the ensemble approach performs best across almost all horizons, with the biggest differences at the highest levels of aggregation.
+It is important to highlight that, all forecasting models outperform the Naive empirical distribution that is used as a benchmark for both point and probabilistic forecasts.
 
@@
 #| out.width: "100%"
-#| fig-cap: "Average accuracy by week for 12 weeks. CRPS is relative to a naive ECDF. MASE and MSSE are relative to the corresponding values for the training set."
-rmsse <- readr::read_rds(here::here("results/rmsse.rds"))
-mase <- readr::read_rds(here::here("results/mase.rds"))
-crps <- readr::read_rds(here::here("results/crps.rds")) |>
-  pivot_wider(names_from = model, values_from = crps) |>
-  group_by(method, h, series) |>
-  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
-  ungroup() |>
+#| fig-cap: "Average accuracy by week for 12 weeks using MinT reconciliation. CRPS is relative to a naive ECDF. MASE and MSSE are relative to the corresponding values for the training set."
+#| fig-pos: "H"
+crps <- crps |>
   pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")
@@
   mutate(
+    measure = factor(measure, levels = c("mase","msse","crps"), labels = c("MASE","MSSE","CRPS")),
     series = factor(series, levels = c("Total", "Control areas", "Health boards")),
@@
 acc_summary |>
-  ggplot(aes(x = week, y = accuracy, group = model)) +
-  geom_line(aes(col = model)) +
+  ggplot(aes(x = week, y = accuracy, group = model, col = model)) +
+  geom_line() +
+  geom_point(size = .5) +
   facet_grid(measure ~ series, scales = "free_y") +
-  labs(y = "Average accuracy", x = "Week ahead")
+  labs(y = "Average accuracy", x = "Week ahead") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few()
 ```

### 2022-11-18 manuscript/06-conclusion.qmd
sec 5 and 6

@@
 
-Forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures.
+Forecasting problems at Emergency Medical Services often have inherent hierarchical and grouped structures.
 For example, looking at time series of arrival calls in a clinical desk service, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest.
 Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area.
-Alternative structure may arise when attributes of interest are crossed rather than nested.
-For example, the total demand could be disaggrgated by priority (e.g Red, Amber, Green) or nature of incidents.
+Alternative structures may arise when attributes of interest are crossed rather than nested.
+For example, the total demand could be disaggregated by priority (e.g., Red, Amber, Green) or by nature of incidents.
 It is also natural to have a mixed structures, for example the total demand could be disaggregated by priority and also by health board.
 
-Despite the inherent hierarchical structure of the forecasting problem in EMS, the common practice is to produce point forecast for each time series independently.
-This practice may lead to the lack of coordination and possibly undesirable and conflicting outcomes.
+Despite the inherent hierarchical structure of the forecasting problem in EMS, the common practice is to produce point forecasts for each time series independently.
+This practice may lead to a lack of coordination and possibly undesirable and conflicting outcomes.
 Furthermore, due to the asymmetric impact of resource allocation in this area, quantifying forecast uncertainty through probabilistic forecasts is also of value as it enables planners to manage associated risks.
-In this paper, we investigate the application of hierarchical forecasting methods for producing a probabilistic forecast of daily incident attended up to 84 days ahead, using different forecasting methods.
+In this paper, we investigate the application of hierarchical forecasting methods for producing probabilistic forecasts of daily incidents attended up to 84 days ahead, using different forecasting methods.
 
-Our results indicates that forecast reconciliation can not only contribute to a more coordinated efforts through producing coherent forecasts when it comes to using forecasting to inform decisions in EMS, but also it can result in forecast accuracy improvement.
-Our proposed forecasting models combined with reconciliation approaches outperform the empirical distribution benchmark.
-We show that a substantial forecast improvement can be achieved at higher levels of hierarchical and grouped structure time series by applying forecast reconciliation methods.
+Our results indicate that forecast reconciliation in EMS can not only contribute to a more coordinated approach to decision making through producing coherent forecasts, but also it can result in forecast accuracy improvement.
+Our proposed forecasting models, combined with reconciliation approaches, outperform the empirical distribution benchmark.
+We show that a substantial forecast improvement can be achieved at higher levels of aggregation by applying forecast reconciliation methods.
 When a point forecast is of interest at the bottom level of series, we observe that reconciliation may not improve the forecast accuracy, if the bottom series are noisy and lack systematic patterns.
@@
 Therefore, we recommend that forecast reconciliation approaches to be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted.
-Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individualy, works remarkably well for our mixed hierarchical & grouped structure.
+Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individually, works remarkably well for our mixed hierarchical & grouped structure.
 

### 2022-11-18 manuscript/05-results.qmd
typo

--- 2022-11-17 manuscript/05-results.qmd
@@
 
-In this section, we compare the forecasting performance of Naive, ETS, GLM, Tscount and the ensemble model using base forecast and Minimum Trace (Mint) methods.
-We have also computed the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches.
+In this section, we compare the forecasting performance of Naive, ETS, Poisson regression using GLM, Poisson regression with serial correlations using tscount and the ensemble model using base forecast and Minimum Trace (Mint) methods.
+We have also computed the forecast accuracy for Bottom-up (BU), Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches.
 However, they are not reported here because their accuracy is outperformed by Mint.

### 2022-11-18 manuscript/main.tex
draft of cover letetr and title page

--- 2022-11-16 manuscript/main.tex
@@
 
-\usepackage{todonotes,mathtools}
+\usepackage{todonotes,mathtools,bm,amsmath}
 \mathtoolsset{showonlyrefs}
-
+\usepackage{mathpazo}
+\setlength{\parindent}{0cm}
 \usepackage{booktabs}
@@
 \hypersetup{
-  pdftitle={Hierarchical time series forecasting in Emergency Medical Services},
+  pdftitle={Hierarchical Time Series Forecasting in Emergency Medical Services},
   pdfauthor={Bahman Rostami-Tabar; Rob J. Hyndman},
-  pdfkeywords={forecasting, healthcare, emergency services, forecast
-reconciliation, hierarchical time series, ambulance demand, attended
-incidents},
+  pdfkeywords={healthcare, emergency services, forecast
+reconciliation, ambulance demand, regression},
   colorlinks=true,
@@
 
-\setlength{\parindent}{6pt}
 \begin{document}
@@
 \begin{frontmatter}
-\title{Hierarchical time series forecasting in Emergency Medical
+\title{Hierarchical Time Series Forecasting in Emergency Medical
 Services}
@@
 coherent and use all the available data at all levels of disaggregation.
-The methods are applied to daily incident data from the Welsh Ambulance
-Service NHS Trust, from October 2015 to July 2019, disaggregated by
+The methods are applied to daily incident data from an ambulance service
+in Great Britain, from October 2015 to July 2019, disaggregated by
 nature of incident, priority, managing health board, and control area.
@@
 \begin{keyword}
-    forecasting \sep healthcare \sep emergency services \sep forecast
-reconciliation \sep hierarchical time series \sep ambulance demand \sep 
-    attended incidents
+    healthcare \sep emergency services \sep forecast
+reconciliation \sep ambulance demand \sep 
+    regression
 \end{keyword}
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, sharp corners, boxrule=0pt, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-Inability to match the resources with the demand in Emergency Medical
-Services (EMS) results in an overcrowding care system. This is a serious
-problem causing challenging situations on patient flow with serious
-consequences on patients, staff and the entire care system
+A failure to match available resources to demand in Emergency Medical
+Services (EMS) results in patient flow problems, with serious
+consequences for patients, staff and the entire care system
 \citep{ekstrom2015forecasting, ROSTAMITABAR20221197}. Demand forecasting
-in EMS is a vital element that helps to depict various courses of action
-to avoid the mismatch, which can result in massive savings in terms of
-patient safety and lives. An accurate daily demand forecasting enables
-planners and decision makers to manage resources to meet anticipated
-patients, reconfigure units, redeploy staff and vehicles, where
-necessary.
+in EMS helps service planners to avoid the mismatch, potentially
+providing massive savings in costs and lives, and leading to better
+patient outcomes. Accurate daily demand forecasting enables planners and
+decision makers to manage resources to meet anticipated patients,
+reconfigure units, and redeploy staff and vehicles as necessary.
 
-Demand forecasts at EMS are typically required at multiple levels of
-cross-sectional granularities to inform various planning and
-decision-making processes \citep{hulshof2012taxonomic}. There are some
-planning process at the national level (more strategic or long-term)
-such as workforce resource planning and budgeting; sub-national,
-regional or healthcare level (tactical or medium-term) such as temporary
-capacity expansions, resource sharing and staff-shift scheduling; and
-hospital or station level (operational or short-term) such as planning
-rosters for staff and ambulance deployment. Demand forecasts might also
-be required at different level for a specific group of interest such as
-nature of demand or priority. Moreover, the time series data in EMS has
+Demand forecasts at EMS are typically required at multiple levels of an
+organization to inform various planning and decision-making processes
+\citep{hulshof2012taxonomic}. There are some planning process at the
+national level (strategic and long-term) such as workforce resource
+planning and budgeting; sub-national, regional or healthcare level
+(tactical and medium-term) such as temporary capacity expansions,
+resource sharing and staff-shift scheduling; and hospital or station
+level (operational and short-term) such as planning rosters for staff
+and ambulance deployment. Demand forecasts might also be required at
+different levels for a specific area of interest such as the nature of
+demand or the priority level. Moreover, the time series data in EMS has
 an inherent hierarchical and grouped structure to support such
@@
 country level can be disaggregated in a geographical hierarchy into
-sub-national, regions, health boards, stations/hospitals or divided into
-groups such as nature of incidents or demand priority. Therefore, using
-forecasting methodologies that account for hierarchical and/or grouped
-structures of time series in EMS seems to be a natural fit.
+sub-national, regions, health boards, and stations/hospitals, or divided
+into groups such as the nature of incidents or demand priority.
+Therefore, using forecasting methodologies that account for hierarchical
+and/or grouped structures of time series in EMS is a natural fit.
 
@@
 \citep{mingliterature2022, gul2020exhaustive, ibrahim2016modeling, wargon2009systematic},
-this area is neglected and the main focus has been on producing
-forecasts at a single level, independently. Generating independent
-forecasts not only ignore the inherent hierarchical and/or grouped
-structure relationships of the time series demand but also results in a
-lack of consistency and coordination. Consistent forecasts of the EMS
-demand across all hierarchical and grouped levels are paramount for an
-effective planning and decision making. The hierarchical forecasting
-approaches can not only create consistent forecasts but also achieve
-more accurate forecasts than the independent (base) forecasts
-\citep{hyndman2011optimal}. Obtaining consistent forecasts at different
-levels is important as it helps to avoid making conflicting decisions.
-With hierarchical forecasting, plans at any level are based on coherent
-forecasts and therefore can be aligned. Implementing and sustaining
-improvements in EMS require alignments and coordination between
-different stakeholders, without which teams operate in isolation leading
-to conflicts, duplication work, rework, or work that runs counter to the
-overall goal to improve the quality of delivery service. Hierarchical
-forecasting framework can be used as a tool to improve coordination
-between teams across the care services at the national, sub-national,
-regional and local levels. To our knowledge, there is not only no
-research in the EMS forecasting to account for the hierarchical and
-grouped structure of the system but also in the entire field of
-forecasting for healthcare management.
+the hierarchical data structure has been largely ignored, and the main
+focus has been on producing independent forecasts at a single level.
+Generating independent forecasts can result in a lack of consistency and
+coordination, and therefore leads to less effective planning and
+decision making. With hierarchical forecasting, plans at any level are
+based on coherent forecasts and therefore can be aligned. Implementing
+and sustaining improvements in EMS require alignments and coordination
+between different stakeholders, without which teams operate in isolation
+leading to conflicts, duplication work, rework, or work that runs
+counter to the overall goal to improve the quality of delivery service.
+Hierarchical forecasting framework can be used as a tool to improve
+coordination between teams across the care services at the national,
+sub-national, regional and local levels. The hierarchical forecasting
+approaches not only create consistent forecasts, but are usually also
+more accurate than the independent (base) forecasts
+\citep{hyndman2011optimal}. To our knowledge, there has been no previous
+research involving hierarchical and grouped forecasting in the entire
+field of forecasting for healthcare management.
 
@@
 of verified incidents from 2015 to 2020 in a major ambulance service in
-Great Britain. The data has hierarchical and grouped structures with
+Great Britain. The data has hierarchical and grouped structures, with
 hierarchies at the national, control (i.e.~sub-national), health board
-(i.e.~regional) and groups by priority and nature of incidents. We
-produce not only the point forecast but also the forecast distribution
-across all levels, which is critical for an effective planning and
-associated risk management. We compare the point and probabilistic
-forecast accuracy of the independent forecasts, bottom-up and optimal
-reconciliation approaches. We first generate independent/based forecasts
-using Exponential Smoothing State Space (ETS), Generalized Linear Model
-(GLM), Poisson regression, a simple empirical distribution and an
-ensemble method followed by applying bottom-up and optimal
-reconciliation approaches. Forecast performance is assessed by the Root
-Mean Squared Scaled Error (RMSSE) for point forecasts and the Continuous
-Ranked Probability Score (CRPS) for the probabilistic forecasts. This
-paper complies with the principles of the reproducibility
-\citep{stodden2013best, boylan2015reproducibility}. Therefore, the study
-could equally be applied to any healthcare service (e.g.~emergency
-department, primary or social care) subject to the time series having a
-hierarchical and/or grouped structure, which is generally the case in
-the healthcare sector.
+(i.e.~regional) levels, as well as groups by priority and nature of
+incidents. We produce consistent point forecasts and forecast
+distributions for all levels, which is critical for an effective
+planning and associated risk management. We compare the point and
+probabilistic forecast accuracy of the independent forecasts, bottom-up
+and optimal reconciliation approaches. We first generate
+independent/base forecasts using Exponential Smoothing State Space
+(ETS), Generalized Linear Model (GLM), Poisson regression, a simple
+empirical distribution and an ensemble method, followed by applying
+bottom-up and optimal reconciliation approaches. Forecast performance is
+assessed by the Mean Squared Scaled Error (MSSE) for point forecasts and
+Continuous Ranked Probability Scores (CRPS) for the probabilistic
+forecasts. This paper complies with reproducibility principles
+\citep{stodden2013best, boylan2015reproducibility}, and can be applied
+to any healthcare service (e.g., emergency department, primary or social
+care) subject to the time series having a hierarchical and/or grouped
+structure.
 
-The remainder of this article is structured as follows: In section 2, we
-provide a brief review of the literature and discuss its limitation to
-position our work; in section 3, we present the experiment design
+The remainder of this article is structured as follows: In
+Section~\ref{sec-lit}, we provide a brief review of the literature and
+discuss its limitation to position our work; in
+Section~\ref{sec-experiment}, we present the experiment design
 describing the data set, forecasting methods and forecast evaluation
-metrics. In Section 4, we discuss the hierarchical time series
-forecasting approaches to generate both point and probabilistic
-forecasts. In section 5, we present and discuss our results; in section
-7, we summarize our findings and present ideas for future research.
+metrics. In Section~\ref{sec-htc}, we discuss the hierarchical time
+series forecasting approaches to generate both point and probabilistic
+forecasts. In Section~\ref{sec-results}, we present and discuss our
+results; in Section~\ref{sec-conclusion}, we summarize our findings and
+present ideas for future research.
 
-\hypertarget{lit}{%
-\section{Research background}\label{lit}}
+\hypertarget{sec-lit}{%
+\section{Research background}\label{sec-lit}}
 
@@
 delivery of urgent medical care to communities. An effective service
-delivery requires accurate resource planning that are generally replying
-on demand forecasts at operational, tactical and strategic levels.
+delivery requires accurate resource planning that generally relies on
+demand forecasts at operational, tactical and strategic levels.
 
 There is a substantial number of studies on the application of time
-series forecasting in the Emergency Medical Services. Various areas have
-been the focus of the literature. Forecasting call volume arrivals is
-one of the major research topics. \citet{ibrahim2016modeling} provide an
-extensive review of the forecasting models in this context. Another
-important area is related to forecasting ambulance demand. Although the
-definition of demand might not be always clearly stated, however, this
-is typically referring to a situation where a physical resource has been
-deployed to respond to an incident. This might be also called
-\emph{attended incidents}. Another demand related variable is verified
-incidents. These are all incidents that require an action: either send a
-physical vehicle, deal with via the Clinical Support Desk (e.g.~calls),
-get an external (private) provider to respond to it, or send it through
-to other channels such as police, firefighters or general practitioners.
-Our study is aligned with this stream of the literature. Another similar
-area that is largely studied in the literature, is Emergency Department
-forecasting. We refer interested readers to \citet{mingliterature2022},
+series forecasting in the Emergency Medical Services. For example,
+\citet{ibrahim2016modeling} provide an extensive review of the models
+used in forecasting call volume arrivals. Another important area is
+related to forecasting ambulance demand. Although the definition of
+demand might not be always clearly stated, this is typically referring
+to a situation where a physical resource has been deployed to respond to
+an incident. This might be also called \emph{attended incidents}.
+Another demand related variable is verified incidents; these are all
+incidents that require an action: either by sending a physical vehicle,
+responding via the Clinical Support Desk, requesting an external
+provider to respond to it, or forwarding it to other channels such as
+police, firefighters or general practitioners. Our study is aligned with
+this stream of the literature. Another similar area that has received
+considerable attention is Emergency Department forecasting; we refer
+interested readers to \citet{mingliterature2022},
 \citet{gul2020exhaustive} and \citet{wargon2009systematic} for extensive
-reviews of the literature on Emergency Department forecasting. Although
-crucial to EMS performance, \citet{aringhieri2017emergency} state that
-demand forecasting has received limited research attention in the EMS
-context. In this section, we provide a brief review of studies on
-forecasting ambulance demand in EMS.
+reviews of the relevant literature. Although crucial to EMS performance,
+\citet{aringhieri2017emergency} state that demand forecasting has
+received limited research attention in the EMS context. In this section,
+we provide a brief review of studies on forecasting ambulance demand in
+EMS.
 
 There are generally two main streams of research related to forecasting
-ambulance demand in EMS: i) the first stream focuses on the application
-of time series methods and regression approaches on forecasting
+ambulance demand in EMS: (i) the first stream focuses on the application
+of time series methods and regression approaches to forecasting
 aggregate ambulance demand \citep{vile2012predicting, sasaki2010using};
-ii) the second stream considers forecasting EMS demand in a more finer
+and (ii) the second stream considers forecasting EMS demand in finer
 temporal and geographical granularities by employing temporal-spatial
@@
 companies employing more than five people are included in the
-regression. The census variables describe groups who may be more likely
-to need an ambulance. A stepwise ordinary least squares regression
-analysis with SPSS is used for estimating the parameter and generating
-forecast. The only performance measure reported in this study is
-\(R^2\). The research design of this study is not rigorous and the study
-is not reproducible. \citet{vile2012predicting} explore using a Singular
+regression. The census variables describe groups who are more likely to
+need an ambulance. A stepwise ordinary least squares regression analysis
+is used for estimating the parameter and generating forecast. The only
+performance measure reported in this study is \(R^2\), which is not an
+effective measure of forecast accuracy \citep[p457]{armstrong01}. The
+research design of this study is not rigorous and the study is not
+reproducible. \citet{vile2012predicting} explore using a Singular
 Spectrum Analysis (SSA) method to generate forecasts of the EMS demand
 at the national level for 7-day, 14-day, 21-day and 28-day forecast
-horizons using data provided by the Welsh Ambulance Service Trust
-(WAST). The performance of this approach is compared with
-Auto-Regressive Moving Average (ARIMA) and Holt-winter time series
-methods using Root Mean Squared Error (RMSE). They concluded that point
-forecasts generated by SSA are more accurate for longer-term, but ARIMA
-and Holt-winter performance is superior for shorter-term horizons.
-\citet{vile2016time} further develop a decision support system to
-integrate forecasts generated by SSA. However, the study does not
+horizons using data provided by an ambulance service in Great Britain.
+The performance of this approach is compared to Auto-Regressive
+Integrated Moving Average (ARIMA) and Holt-Winters time series methods
+using Root Mean Squared Error (RMSE). They concluded that point
+forecasts generated by SSA are more accurate for longer-term, but that
+ARIMA and Holt-Winters performance is superior for shorter-term
+horizons. \citet{vile2016time} further develop a decision support system
+to integrate forecasts generated by SSA. However, the study does not
 compare and contrast the performance of forecasting methods based on
@@
 The tool contains options that allow generating forecasts at various
-levels of granularity, however, it ignore the hierarchical and grouped
+levels of granularity, however, it ignores the hierarchical and grouped
 relationships structure, preventing aligned decision making and
@@
 \citet{9659837} investigate forecasting EMS demand in a high
-spatio-temporal resolution of 1×1km spatial regions and 1-hr time
-intervals using total incidents in Oslo, Norway from January 1st, 2015
-up to and including February 11th, 2019. They used multi-layer
-perceptron (MLP) and long short-term memory (LSTM) models to forecast
-the EMS demand, and compare them to simple aggregation methods and
-baselines. The point forecast accuracy is evaluated using Mean Absolute
-Error (MAE) and Mean Squared Error (MSE) and the forecast distribution
-is measures by Categorical Cross-Entropy. They shows that while Neural
-Network models perform better in producing point forecast, a
-distribution baseline method based on spatial distribution of the
-incidents across all time steps provide more accurate forecast
-distribution. \citet{zhou2016predictinglit} propose three methods based
-on Gaussian mixture models, kernel density estimation, and kernel
-warping to predict 4 weeks into future for a 1-km2 spatial region over
-an hour. Two years of incidents attended from Toronto, Canada (years
-2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011
-and 2012 with 696,975 events) are used to build the model and examine
-the performance on test data using mean negative log likelihood. They
-show that forecasts generated by the proposed methods are significantly
-more accurate than the current industry practice, a simple averaging
-formula. \citet{grekousis2019will} investigate the combination of
-spatial analysis methods with data mining techniques based on an
-improved Hungarian algorithm and MLP neural network to identify the most
-likely locations of future emergency events. The proposed approach is
-tested using data of 2851 events attended by the EMS in Athens, Greece
-over 24 weeks. They show that \(23.24\%\) of real emergency events lie
-within 50 meter of the predicted ones and nearly \(70\%\) of the real
-emergency events lie no further away than 150 meter, which is rather
-accurate given the granularity of the problem at the city level.
+spatio-temporal resolution of 1km\(^2\) spatial regions and 1-hr time
+intervals using total incidents in Oslo, Norway, from 1 January 2015 to
+11 February 2019. They used multi-layer perceptron (MLP) and long
+short-term memory (LSTM) models to forecast the EMS demand, and compare
+the results to simple aggregation methods and baselines. The point
+forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean
+Squared Error (MSE), and the forecast distribution is measured by
+Categorical Cross-Entropy. They found that Neural Network models
+performed better in producing point forecasts, while a distribution
+baseline method based on spatial distribution of the incidents across
+all time steps provided more accurate forecast distributions.
+\citet{zhou2016predictinglit} proposed three methods based on Gaussian
+mixture models, kernel density estimation, and kernel warping to predict
+hourly data 4 weeks ahead for a 1km\(^2\) spatial region. Two years of
+incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events)
+and Melbourne, Australia (years 2011 and 2012 with 696,975 events) were
+used to build the model and examine the performance on test data using
+mean negative log likelihood. They show that forecasts generated by the
+proposed methods were significantly more accurate than the current
+industry practice (a simple averaging formula).
+\citet{grekousis2019will} investigated the combination of spatial
+analysis methods with data mining techniques based on an improved
+Hungarian algorithm and a MLP neural network to identify the most likely
+locations of future emergency events. The proposed approach was tested
+using data of 2851 events attended by the EMS in Athens, Greece, over 24
+weeks. They showed that 23\% of real emergency events lie within 50
+meters of the predicted ones and nearly 70\% of the real emergency
+events lie no further than 150 meters away, which is rather accurate
+given the granularity of the problem at the city level.
 
-We observed a number of limitations in the literature of EMS
-forecasting, that encourage us to undertake this research. These
-limitations are summarized as following:
+We note a number of limitations in the literature of EMS forecasting,
+that encourage us to undertake this research. These limitations are
+summarized as following:
 
@@
   Current studies ignore the inherent hierarchical and/or grouped
-  structure of the time series data and the relationship between series
+  structure of the time series data, and the relationship between series
   at different levels of hierarchy. This may result in incoherent
@@
 \item
-  Current research is mainly concerned with generating point forecast at
-  a single level of hierarchy. There is a lack of studies presenting the
-  entire forecast distribution of daily ambulance demand for the entire
-  hierarchy to inform the whole decision-making process and to better
+  Current research is mainly concerned with generating point forecasts
+  at a single level of hierarchy. There is a lack of studies considering
+  the entire forecast distribution of daily ambulance demand for the
+  whole hierarchy to inform the decision-making process and to better
   represent the uncertainty of future demand, providing a risk
@@
   Reproducibility is still a major challenge in EMS forecasting, as it
-  is unlikely to reproduce the results without the help of the authors
-  of those papers.
+  is unlikely that any reader can reproduce prior studies without the
+  help of the authors of those papers.
 \item
-  Another limitation is related to the generated forecasts that are not
-  integer counts. Since actual ambulance counts cannot be negative or
-  fractions, ambulance demand forecasts should be the same. While this
-  might not be an issue when producing forecasts at a single level,
-  producing non-negative count forecasts in a hierarchical/grouped
-  structure is challenging and requires further investigation in the
-  future.
+  Another limitation is related to the generated forecasts not being on
+  the sample space of non-negative counts. Since actual ambulance counts
+  cannot be negative or non-integer, ambulance demand forecast
+  distributions should reflect the data. Of course, point forecasts
+  represent means, so they should be non-negative, but may be
+  non-integer. While this might not be an issue when producing forecasts
+  at a single level, producing non-negative count forecasts in a
+  hierarchical/grouped structure is challenging and requires further
+  investigation in the future.
 \end{enumerate}
@@
 We are interested in generating forecasts to inform the planning horizon
-of \(ph= 42\) days, required by planners in the ambulance services
-trust. The forecast horizon in this study is \(fh = 2 \times ph\) days
-ahead (2*42 days planning horizon). This is because the planning is
-generally frozen for \(ph\) days and considering a forecast horizon of
-\(ph\) days might not be helpful for planning. While forecasts are
-generated for \(2 \times ph\) days ahead, performance evaluation is only
-assessed based on the last \(ph\) days and not the \(2 \times ph\) days.
-The forecasts are produced for the holdout of 365 days using time series
-cross-validation \citep{hyndman2021forecasting}.
+of 42 days, required by planners in the ambulance service. The forecast
+horizon in this study is \(2 \times 42 = 84\) days ahead, because the
+planning is generally frozen for 42 days and so forecasts of the next 42
+days is not particularly helpful for planning. While forecasts are
+generated for 84 days ahead, performance evaluation is only assessed
+based on the last 42 days and not the whole forecast period. The
+forecasts are produced for various training and test sets using time
+series cross-validation \citep{hyndman2021forecasting}.
 
 In the following section, we discuss the dataset, describe the
-forecasting methods used to generate base forecasts and present the
+forecasting methods used to generate base forecasts, and present the
 point and probabilistic accuracy measures.
@@
 
-The dataset used in this study is from a major ambulance service trust
-in Great Britain. It contains information relating to the daily number
-of attended incidents from 1 October 2015 to 31 July 2019, disaggregated
-by nature of incidents, priority, the health board managing the service
-and the control area (or region). Figure~\ref{fig-hierarchy} depicts
-both the hierarchical and grouped structure of the data.
+The dataset used in this study is from a major ambulance service in
+Great Britain. It contains information relating to the daily number of
+attended incidents from 1 October 2015 to 31 July 2019, disaggregated by
+nature of incidents, priority, the health board managing the service and
+the control area (or region). Figure~\ref{fig-hierarchy} depicts both
+the hierarchical and grouped structure of the data.
 Figure~\ref{fig-hierarchy-1} illustrates the nested hierarchical
@@
 series with the strength of trend in x-axis and the strength of
-seasonality in y-axis. It is clear that there are some series showing
-strong trends and/or seasonality, corresponding to series at the higher
-levels of the hierarchy. The majority of series show low trend and
-seasonality. These are time series belonging to the bottom series,
-series related to the nature of incidents for a given control, health
-board and priority level. Bottom series are dominated by noise with
-little or no systematic patterns.
+seasonality in y-axis. Both measures are on a scale of {[}0,1{]}. It is
+clear that there are some series showing strong trends and/or
+seasonality, corresponding to series at the higher levels of the
+hierarchy. The majority of series show low trend and seasonality. These
+are time series belonging to the bottom series, series related to the
+nature of incidents for a given control, health board and priority
+level. Bottom series are dominated by noise with little or no systematic
+patterns.
 
@@
 \caption{\label{fig-dataviz2}Time series of attended incidents at
-various levels.}
+various levels. The panels show data from the whole country, by control
+area, by health board, by priority level, and by nature of incident.
+Only four of the 35 nature of incident categories are shown to avoid too
+much overplotting.}
 
@@
 the base forecasts. Once the base forecasts are produced, hierarchical
-and grouped time series methods are used to reconcile them across the
-all levels. We briefly discussed forecasting models in the following
-sections, and the hierarchical forecasting methods are discussed in
+and grouped time series methods are used to reconcile them across all
+levels. We briefly discuss forecasting models in the following sections,
+and the hierarchical forecasting methods are discussed in
 Section~\ref{sec-htc}.
@@
 distribution of the past daily attended incidents to create the forecast
-distribution of future attended incidents. We consider the empirical
-distribution of the most recent year of historic data on a rolling basis
-to capture potential changes in behavior over
-time.\todo{Actually we use all the data.}
+distribution of future attended incidents.
 
@@
 \citep{hyndman2021forecasting} can combine trend, seasonality and error
-components in a time series through various forms such as additive,
+components in a time series through various forms that can be additive,
 multiplicative or mixed. The trend component can be none (``N''),
-Additive (``A''), damped (``Ad'') or multiplicative (``M''). The
-seasonality can be none (``N''), Additive (``A''), or multiplicative
-(``M''). The error term can also be additive (``A'') or multiplicative
-(``M''). To forecast the attended incidents at each level, we use the
-\texttt{ETS()} function in the \texttt{fable} package \citep{fable2022}
-in R. To identify the best model for a given time series, the ETS
-function uses the corrected Akaike's Information Criterion (AICc).
+Additive (``A'') or damped (``Ad''); the seasonality can be none
+(``N''), Additive (``A''), or multiplicative (``M''); and the error term
+can be additive (``A'') or multiplicative (``M''). To forecast the
+attended incidents at each level, we use the \texttt{ETS()} function in
+the \texttt{fable} package \citep{fable2022} in R. To identify the best
+model for a given time series, the \texttt{ETS} function uses the
+corrected Akaike's Information Criterion (AICc).
 
 Despite the popularity and the relevance of automatic \texttt{ETS} in
-this study, however it may produce forecast distributions that are
-non-integer and include negative values, despite the number of attended
-incidents being integer and non-negative. When using \texttt{ETS}, a
-time series transformation approach could be used to generate strictly
+this study, it may produce forecast distributions that are non-integer
+and include negative values, although the number of attended incidents
+is always integer and non-negative. When using \texttt{ETS}, a time
+series transformation approach could be used to generate strictly
 positive forecasts, although forecast distributions will still be
 non-integer. An alternative is to use forecasting models that produce
-integer, non-negative forecasts. in the following section we present
-Generalized Linear Models (GLMs) and Poisson Regression to produce count
-base forecasts.
+integer, non-negative forecasts. In the following section we present
+Generalized Linear Models (GLMs) and Poisson time series regression to
+produce count base forecasts.
 
 \textbf{Generalized Linear Model (GLM):} GLMs are a family of models
-developed to extend the concept of linear regression models. They
-perform a regression by modeling the response variable as coming from a
-particular member of the exponential family, and then transforming the
-mean of the response so that the transformed mean is a linear function
-of the predictors. One of the models that is frequently used in practice
-to generate count forecasts is the Poisson regression. We will consider
-forecasting attended incidents using the covariates spline trend, day of
-the week dummy variables (from Monday to Sunday), Fourier terms to
-capture yearly seasonality, public holidays (1 when is public holiday, 0
-otherwise), school holidays (1 when is school holiday, 0 otherwise) and
-Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day
-(1 when is New Year's Day, 0 otherwise). We fit a Poisson regression
-model using the function \texttt{glm()} from the package in R, with the
+developed to extend the concept of linear regression models to
+non-Gaussian distributions \citep{Faraway2016}. They model the response
+variable as a particular member of the exponential family, with the mean
+being a transformation of a linear function of the predictors. One of
+the models that is frequently used in practice to generate count
+forecasts is Poisson regression. We will consider forecasting attended
+incidents using the covariates spline trend, day of the week dummy
+variables (from Monday to Sunday), Fourier terms to capture yearly
+seasonality, public holidays (1 when is public holiday, 0 otherwise),
+school holidays (1 when is school holiday, 0 otherwise) and Christmas
+Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is
+New Year's Day, 0 otherwise). We fit a Poisson regression model using
+the function \texttt{glm()} from the \emph{stats} package in R, with the
 argument \texttt{family\ =\ poisson} to specify that we wish to fit a
-Poisson regression model.
+Poisson regression model with a log link function.
 
-\textbf{Poisson Regression using tscount:} We consider another Poisson
-regression model that takes into account serial dependence in addition
-to covariates used in the GLM model. To that end, we use
-\texttt{tsglm()} function in \emph{tscount} package in R to model the
-attended incidents. The logarithmic link function is used to ensure that
-the parameter of Poisson distribution is always positive. This model can
-be estimated via maximization of the likelihood function based on
-Poisson mass function. The regression model captures the short range
-serial dependence by including the three order autoregressive terms.
+\textbf{Poisson Regression using tscount (TSGLM):} We also consider
+another Poisson regression model that takes into account serial
+dependence, using the \texttt{tsglm()} function in the \emph{tscount}
+package in R \citep{JSSv082i05} to model the attended incidents. Again,
+the logarithmic link function is used to ensure that the parameter of
+Poisson distribution is always positive. This model captures the short
+range serial dependence by including three autoregressive terms, in
+addition to the same covariates that were used in the GLM model. To
+distinguish this from the previous GLM model, we will refer to this
+model as ``TSGLM''.
 
-\textbf{Ensemble method:} we also use an ensemble method that combines
-forecasts generated from Naive, ETS, GLM and Poisson regression to form
-a mixture distribution \citep{combinations}.
+\textbf{Ensemble method:} Finally, we use an ensemble method that
+combines the forecasts generated from the Naive, ETS, GLM and TSGLM
+models to form a mixture distribution \citep{combinations}.
 
@@
 and hyper-parameter tuning is performed using the training data only.
+While we compute forecast errors for the entire 12 weeks, we are most
+interested in the last 42 days of each test set, because that
+corresponds to how forecasts are generated for planning in practice.
 Forecasting performance is evaluated using both point and probabilistic
-error measures. While we compute forecast errors for the entire 12
-weeks, we are most interested in the last 42 days of each test set,
-because that corresponds to how forecasts are generated for planning in
-practice. Forecasting performance is evaluated using both point and
-probabilistic error measures.
+error measures.
 
@@
 
-There are many applications in the healthcare and in particular in EMS
+There are many applications in healthcare, and in particular in EMS,
 where a collection of time series is available. These series are
@@
 area/region, health board and/or are aggregated at different levels in
-groups based on nature of demand, priority of demand or some other
+groups based on nature of demand, priority of demand, or some other
 attributes. While series could be strictly hierarchical or only grouped
@@
 arise when attributes of interest are both nested and crossed, having
-hierarchical and grouped structure. This is also the case as discussed
-in Section~\ref{sec-data}
+hierarchical and grouped structure. This is also the case for our
+application as discussed in Section~\ref{sec-data}.
 
@@
 A common practice in healthcare (and EMS) to predict hierarchical and
-grouped series relies on producing independent forecast, also refereed
-to as base forecast, typically by different teams as the need for such
-forecasts arise. This is also known as base forecast. We observe \(n\)
-time series at time \(t\), across the entire hierarchical and grouped
-structure, written as \(y_t\). The base forecasts of \(y_{T+h}\) given
-data \(y_1,\dots,y_T\) are denoted by \(\hat{y}_h\) for h steps-ahead
-for all \(n\) series (\(n=1530\) in this study). Forecasts generated in
-this way are nor coherent.
+grouped series relies on producing independent forecasts, also refereed
+to as base forecasts, typically by different teams as the need for such
+forecasts arise. We observe \(n\) time series at time \(t\), across the
+entire hierarchical and grouped structure, written as \(y_t\). The base
+forecasts of \(y_{T+h}\) given data \(y_1,\dots,y_T\) are denoted by
+\(\hat{y}_h\) for \(h\) steps-ahead for all \(n\) series (\(n=1530\) in
+this study). Forecasts generated in this way are not coherent.
 
@@
 
-Traditionally, alternative approaches to produce coherent forecasts for
-hierarchical and grouped time series involves using bottom-up and
-top-down methods by generating forecasts at a single level and then
-aggregate or disaggregate. Top-down requires having a unique structure
-to disaggregated forecasts generated at top level by proportions.
-However, given that we have multiple grouped attributes combined with
-the hierarchical structure, there is no unique way to disaggregate top
-forecast. Hence the top-down can not be used in this case, so either we
-can do some kind of reconciliation or must define our own top-down
-method for each hierarchy. The recommended approach is to use
-reconciliation. In the following sections, we first discuss some
-notations and then present bottom-up and forecast reconciliation
-approach used in this study to generate coherent forecasts.
+Traditionally, approaches to produce coherent forecasts for hierarchical
+and grouped time series involve using bottom-up and top-down methods by
+generating forecasts at a single level and then aggregating or
+disaggregating. Top-down methods require having a unique hierarchical
+structure to disaggregate forecasts generated at the top level by
+proportions. However, given that we have multiple grouped attributes
+combined with the hierarchical structure, there is no unique way to
+disaggregate top forecasts. Hence the top-down cannot be used in our
+application. The recommended approach is to use forecast reconciliation
+\citep{hyndman2011optimal}. In the following sections, we first discuss
+some notation, and then present bottom-up and forecast reconciliation
+approaches used in this study to generate coherent forecasts.
 
@@
 
-Let \(b_t\) be a vector of \(n_b\) ``bottom-level'' time series at time
-\(t\), and let \(a_t\) be a corresponding vector of \(n_a = n-n_b\)
-aggregated time series, where
-
-\[
- a_t = Ab_t
-\]
-
-and \(A\) is the \(n_a\times n_b\) ``aggregation'' matrix specifying how
-the bottom-level series \(b_t\) are to be aggregated to form \(a_t\).
-The full vector of time series is given by \[
- y_t = \begin{bmatrix}a_t \\b_t\end{bmatrix}.
+Let \(\bm{b}_t\) be a vector of \(n_b\) ``bottom-level'' time series at
+time \(t\), and let \(\bm{a}_t\) be a corresponding vector of
+\(n_a = n-n_b\) aggregated time series, where \[
+  \bm{a}_t = \bm{A}\bm{b}_t,
+\] and \(\bm{A}\) is the \(n_a\times n_b\) ``aggregation'' matrix
+specifying how the bottom-level series \(\bm{b}_t\) are to be aggregated
+to form \(\bm{a}_t\). The full vector of time series is given by \[
+ \bm{y}_t = \begin{bmatrix}\bm{a}_t \\\bm{b}_t\end{bmatrix}.
 \] This leads to the \(n\times n_b\) ``summing'' or ``structural''
 matrix given by \[
-  S = \begin{bmatrix}A \\ I_{n_b}\end{bmatrix}
-\] such that \(y_t = Sb_t\).
+  \bm{S} = \begin{bmatrix}\bm{A} \\ \bm{I}_{n_b}\end{bmatrix}
+\] such that \(\bm{y}_t = \bm{S}\bm{b}_t\).
 
@@
 
-Bottom-Up is a simple approach to generate coherent forecasts. It first
-involves creating the base forecasts for the bottom level series. These
-forecasts are then aggregated to the upper levels which results in
-generating coherent forecasts. BU approach can capture the dynamics of
-the series at the bottom level, but they may be noisy and difficult to
-forecast as well. BU approach is limited on using only forecasts at the
-bottom level and does not utilize all the information available across
-the hierarchical and grouped structure. Forecast reconciliation
-approaches fill this gap and combine and reconcile all the base
-forecasts in order to produce coherent forecasts.
+Bottom-Up is a simple approach to generate coherent forecasts. It
+involves first creating the base forecasts for the bottom level series
+(i.e., the most disaggregated series). These forecasts are then
+aggregated to the upper levels which naturally results in coherent
+forecasts. The BU approach can capture the dynamics of the series at the
+bottom level, but these series may be noisy and difficult to forecast.
+The approach using only the data at the most disaggregated level, and so
+does not utilize all the information available across the hierarchical
+and grouped structure.
 
-Given the summing matrix and base forecasts, bottom-up and linear
-reconciliation methods can be written as \(\tilde{y}_h = SG\hat{y}_h\)
-for different matrices \(G\).
+Forecast reconciliation approaches fill this gap by combining and
+reconciling all the base forecasts in order to produce coherent
+forecasts. Linear reconciliation methods can be written
+\citep{WicEtAl2019} as \[
+  \tilde{\bm{y}}_h = \bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{W}^{-1}\hat{\bm{y}}_h,
+\] where \(\bm{W}\) is an \(n \times n\) positive definite matrix, and
+\(\hat{\bm{y}}_h\) contains the \(h\)-step forecasts of \(\bm{y}_{T+h}\)
+given data to time \(T\). Different choices for \(\bm{W}\) lead to
+different solutions such as Ordinary Least Squares (OLS), Weighted Least
+Squares (WLS) and Minimum Trace (MinT). We use the implementation of
+these methods in the fable package in R in the experiment.
 
-Optimal reconciled forecasts are obtained with
-\(G=(S'W^{-1}S)^{-1}W^{-1}\), or \(\tilde{y}_h = M\hat{y}_h,\) where the
-\(n\times n\) ``mapping'' matrix is given by
-\(M = S(S'W^{-1}S)^{-1}W^{-1},\) where \(\hat{y}_h\) are the \(h\)-step
-forecasts of \(y_{T+h}\) given data to time \(T\), and \(W\) is an
-\(n \times n\) positive definite matrix. Different choices for \(W\)
-lead to different solutions such as Ordinary Least Square (OLS) ,
-Weighted Least Square (WLS) and Minimum Trace (MinT)
-\citep{WicEtAl2019}. We use the implementation of these methods in the
-fable package in R in the experiment.
+\hypertarget{sec-results}{%
+\section{Results and discussion}\label{sec-results}}
 
-\hypertarget{sec-result}{%
-\section{Results and discussion}\label{sec-result}}
+In this section, we compare the forecasting performance of the Naive,
+ETS, GLM and TSGLM models along with the ensemble, using base forecast
+and Minimum Trace (MinT) reconciliation methods. We have also computed
+the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least
+Square (WLS) approaches, along with bottom up forecasting. However, they
+are not reported here because their accuracy is outperformed by MinT. We
+should also note that forecasts, and consequently their corresponding
+errors, are generated for the entire hierarchy and they could be
+reported at any level, if required. But to save space, we have reported
+only the top level (Total), the bottom level, and the levels
+corresponding to Control areas and Health boards. The latter are chosen
+because this is where decision-making takes place, so these forecasts
+are the most important.
 
-\hypertarget{point-forecast}{%
-\subsection{Point forecast}\label{point-forecast}}
+The overall forecasting performance is reported in
+Table~\ref{tbl-result}, in which the average forecast accuracy over
+horizons 43--84 days (corresponding to the planning horizon) is
+presented per model, method and the hierarchical level. Reported
+forecast accuracy are averaged across all forecast horizons, rolling
+origins and series at each level. Table~\ref{tbl-result} presents both
+point and probabilistic forecast accuracy at total, control area, health
+board and bottom level series. Point forecast performance are reported
+using MASE and MSSE in Table~\ref{tbl-result-1} and
+Table~\ref{tbl-result-2}, respectively. Probabilistic forecast accuracy
+is reported using CRPS in Table~\ref{tbl-result-3}. The bold entries in
+each table identify a combination of method and model that performs best
+for the corresponding level (i.e.~each column), based on the smallest
+values of accuracy measures.
 
-\hypertarget{probabilistic-forecast}{%
-\subsection{Probabilistic forecast}\label{probabilistic-forecast}}
+Table~\ref{tbl-result-1} and 2b show that forecast reconciliation
+(i.e.~MinT) improves forecast accuracy at the higher levels of the
+hierarchy including total, control area and health board. However, it is
+does not result in accuracy improvement at the bottom level series, for
+which base forecasts are more accurate. This might be due to the noisy
+structure of time series at the bottom level, and possibly due to very
+different patterns in the aggregated series. It is also clear from
+Table~\ref{tbl-result-1} that the ensemble method improves forecast
+accuracy at total, control area and health board. However, this does not
+remain valid for bottom series where different individual methods
+perform best, depending on the accuracy measure.
 
-\begin{figure}
+\begin{table}
+
+\caption{\label{tbl-result}Average forecast perfomance calculated on the
+test sets at forecast horizons \(h=43,\dots,84\) days, with time series
+cross validation applied to attended incident data. The best approach is
+highlighted in bold.}\begin{minipage}[t]{\linewidth}
+\subcaption{\label{tbl-result-1}Point forecast accuracy using MASE }
+
+{\centering
+
+\tabularnewline
+
+\centering
+\begin{tabular}[t]{llrrrr}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{4}{c}{MASE} \\
+\cmidrule(l{3pt}r{3pt}){3-6}
+Method & Model & Total & Control areas & Health boards & Bottom\\
+\midrule
+Base & Naive & 1.139 & 1.059 & 1.047 & 1.019\\
+Base & ETS & 0.963 & 0.930 & 0.899 & 1.038\\
+Base & GLM & 0.910 & 0.940 & 0.923 & \textbf{1.002}\\
+Base & TSGLM & 0.911 & 0.939 & 0.924 & 1.005\\
+Base & Ensemble & 0.782 & 0.856 & 0.876 & 1.008\\
+\addlinespace
+MinT & Naive & 1.138 & 1.059 & 1.047 & 2.651\\
+MinT & ETS & 0.877 & 0.916 & 0.915 & 1.289\\
+MinT & GLM & 0.848 & 0.901 & 0.902 & 2.493\\
+MinT & TSGLM & 0.852 & 0.903 & 0.903 & 2.513\\
+MinT & Ensemble & \textbf{0.753} & \textbf{0.844} & \textbf{0.872} & 2.260\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{minipage}%
+\newline
+\begin{minipage}[t]{\linewidth}
+\subcaption{\label{tbl-result-2}Point forecast accuracy using MSSE }
+
+{\centering
+
+\tabularnewline
+
+\centering
+\begin{tabular}[t]{llrrrr}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{4}{c}{MSSE} \\
+\cmidrule(l{3pt}r{3pt}){3-6}
+Method & Model & Total & Control areas & Health boards & Bottom\\
+\midrule
+Base & Naive & 1.169 & 1.056 & 1.062 & 1.031\\
+Base & ETS & 0.979 & 0.875 & 0.816 & \textbf{0.975}\\
+Base & GLM & 0.813 & 0.897 & 0.875 & 1.009\\
+Base & TSGLM & 0.822 & 0.901 & 0.875 & 1.050\\
+Base & Ensemble & 0.599 & 0.729 & 0.774 & 0.993\\
+\addlinespace
+MinT & Naive & 1.168 & 1.057 & 1.062 & 2.095\\
+MinT & ETS & 0.785 & 0.852 & 0.845 & 0.994\\
+MinT & GLM & 0.720 & 0.827 & 0.837 & 1.803\\
+MinT & TSGLM & 0.722 & 0.833 & 0.839 & 1.851\\
+MinT & Ensemble & \textbf{0.560} & \textbf{0.706} & \textbf{0.765} & 1.557\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{minipage}%
+\newline
+\begin{minipage}[t]{\linewidth}
+\subcaption{\label{tbl-result-3}Probabilistic forecast accuracy using CRPS }
+
+{\centering
+
+\tabularnewline
+
+\centering
+\begin{tabular}[t]{llrrrr}
+\toprule
+\multicolumn{2}{c}{ } & \multicolumn{4}{c}{CRPS} \\
+\cmidrule(l{3pt}r{3pt}){3-6}
+Method & Model & Total & Control areas & Health boards & Bottom\\
+\midrule
+Base & Naive & 30.387 & 10.882 & 5.500 & 0.302\\
+Base & ETS & 14.309 & 6.074 & 3.476 & 0.244\\
+Base & GLM & 15.396 & 6.253 & 3.576 & 0.244\\
+Base & TSGLM & 15.316 & 6.227 & 3.575 & 0.245\\
+Base & Ensemble & 12.978 & \textbf{5.727} & 3.430 & 0.243\\
+\addlinespace
+MinT & Naive & 30.368 & 10.902 & 5.498 & 0.313\\
+MinT & ETS & 13.515 & 5.967 & 3.547 & \textbf{0.243}\\
+MinT & GLM & 13.839 & 5.917 & 3.453 & 0.246\\
+MinT & TSGLM & 14.000 & 5.947 & 3.455 & 0.248\\
+MinT & Ensemble & \textbf{12.585} & 5.728 & \textbf{3.426} & 0.247\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{minipage}%
+
+\end{table}
+
+Table~\ref{tbl-result-3} presents the accuracy of the forecast
+distributions measures by CRPS, which considers both forecasting
+reliability and interval sharpness. The smaller the value of CRPS, the
+better the comprehensive performance. We observe that forecast
+reconciliation results in forecast improvement, regardless of the
+hierarchical level. The ensemble method is also more accurate for higher
+levels, but ETS performs slightly better at the bottom level. While
+these results show that forecast reconciliation does not improve
+\emph{point} forecasts at the bottom level, Table~\ref{tbl-result-3}
+indicates that it generates more accurate \emph{distributional}
+forecasts than the base method. This is probably due to the
+reconciliation method giving improved forecast accuracy in the tails of
+the forecast distribution, which are critical for managing risks.
+
+Overall, our results indicate that forecast reconciliation using the
+MinT method provides reliable forecasts, improves upon the base
+(unreconciled) forecasts all levels except the bottom level series. But
+even there, forecast reconciliation using MinT improves accuracy in the
+tails of the distribution.
+
+In addition to the overall forecast accuracy presented in
+Table~\ref{tbl-result}, we also report the point and probabilistic
+forecast accuracy measures for each forecast horizon in
+Figure~\ref{fig-accuracy}. The figure focuses on the hierarchical levels
+important for decision-making including total, control area and health
+board; however the accuracy could be calculated for any level. We only
+illustrate the results of the MinT method, given its strong performance
+described in Table~\ref{tbl-result}. For illustration purposes, we
+report the average weekly forecast accuracy instead of the daily
+forecast horizon, as this reduces the visual noise in the figure. Thus
+the x-axis shows horizons from week 1 (\(h= 1,\dots,7\)) to week 12
+(\(h= 78,\dots,84\)). The forecast horizon from week 7 to week 12
+corresponds to the upcoming planning horizon, which is used by planners
+and decision making. For both the point forecast and distributional
+accuracy we can see that the ensemble approach performs best across
+almost all horizons, with the biggest differences at the highest levels
+of aggregation. It is important to highlight that, all forecasting
+models outperform the Naive empirical distribution that is used as a
+benchmark for both point and probabilistic forecasts.
+
+\begin{figure}[H]
 
@@
 
-\caption{\label{fig-accuracy}Average accuracy by week for 12 weeks. CRPS
-is relative to a naive ECDF. MASE and MSSE are relative to the
-corresponding values for the training set.}
+\caption{\label{fig-accuracy}Average accuracy by week for 12 weeks using
+MinT reconciliation. CRPS is relative to a naive ECDF. MASE and MSSE are
+relative to the corresponding values for the training set.}
 
@@
 
+Forecasting problems at Emergency Medical Services often have inherent
+hierarchical and grouped structures. For example, looking at time series
+of arrival calls in a clinical desk service, verified incidents, or
+attended incidents in a country, they could be disaggregated by various
+attributes of interest. Total demand in the country could be
+disaggregated by region, then within each region by health board, within
+each health board, by station/hospital, and so on down to the post code
+area. Alternative structures may arise when attributes of interest are
+crossed rather than nested. For example, the total demand could be
+disaggregated by priority (e.g., Red, Amber, Green) or by nature of
+incidents. It is also natural to have a mixed structures, for example
+the total demand could be disaggregated by priority and also by health
+board.
+
+Despite the inherent hierarchical structure of the forecasting problem
+in EMS, the common practice is to produce point forecasts for each time
+series independently. This practice may lead to a lack of coordination
+and possibly undesirable and conflicting outcomes. Furthermore, due to
+the asymmetric impact of resource allocation in this area, quantifying
+forecast uncertainty through probabilistic forecasts is also of value as
+it enables planners to manage associated risks. In this paper, we
+investigate the application of hierarchical forecasting methods for
+producing probabilistic forecasts of daily incidents attended up to 84
+days ahead, using different forecasting methods.
+
+Our results indicate that forecast reconciliation in EMS can not only
+contribute to a more coordinated approach to decision making through
+producing coherent forecasts, but also it can result in forecast
+accuracy improvement. Our proposed forecasting models, combined with
+reconciliation approaches, outperform the empirical distribution
+benchmark. We show that a substantial forecast improvement can be
+achieved at higher levels of aggregation by applying forecast
+reconciliation methods. When a point forecast is of interest at the
+bottom level of series, we observe that reconciliation may not improve
+the forecast accuracy, if the bottom series are noisy and lack
+systematic patterns. However, forecast reconciliation may result in more
+forecast results for bottom series, if we are interested in the tails of
+forecast distribution rather than just center measures like mean
+(i.e.~point forecast). Producing consistent forecast are also crucial
+for informing planning activities, we also demonstrate that proposed
+models produce consistent forecasts across all forecast horizons.
+Therefore, we recommend that forecast reconciliation approaches to be
+adopted for routine use in EMS, whenever hierarchical and/or grouped
+time series data need to be forecasted. Moreover, we found that using an
+ensemble forecasting model, combining all models developed in this
+paper, instead of using each individually, works remarkably well for our
+mixed hierarchical \& grouped structure.
+
+Further research could investigate the practical benefits of
+probabilistic hierarchical forecasting in EMS. Linking forecasts with
+its utilities (response time, resource utilization, etc) can offer an
+opportunity to maximize benefits through more holistic planning
+approach. While, we generated count attend incident for the base
+forecast using Poisson regression models, however the reconciles
+forecast is snot yet count. This could be also an avenue for further
+research.
+
 \renewcommand\refname{References}

### 2022-11-18 manuscript/title_page.qmd
draft of cover letetr and title page

--- 2022-11-18 manuscript/title_page.qmd
@@
 - **Author contributions:** BRT and RJH conceptualize the the study and designed the experiment.
-BRT obtained a funding to meet RHH at Monash University.
-BRT and RJH provided the analysis of the data.
-RJH developed the models; ML chaired the data oversight committee.
+BRT obtained a funding to meet RJH at Monash University in Melbourne.
+BRT and RJH provided the analysis of the time series data.
+RJH developed forecasting models and hierarchical methods; BRT and RJH use R programming to conduct analysis.
 BRT drafted the manuscript, and RJH contributed substantially to its revision.

### 2022-11-18 manuscript/graphs.R
remove pdf

@@ -1,44 +1,0 @@
-library(tidyverse)
-
-options(
-  ggplot2.continuous.color = "viridis",
-  ggplot2.continuous.fill = "viridis",
-  ggplot2.discrete.color = c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9", "#F0E442"),
-  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#CC79A7", "#E69F00", "#009E73", "#56B4E9", "#F0E442")
-)
-
-rmsse <- readr::read_rds(here::here("results/rmsse.rds"))
-mase <- readr::read_rds(here::here("results/mase.rds"))
-crps <- readr::read_rds(here::here("results/crps.rds")) |>
-  pivot_wider(names_from = model, values_from = crps) |>
-  group_by(method, h, series) |>
-  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
-  ungroup() |>
-  pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")
-
-accuracy <- bind_rows(
-  rmsse |> mutate(measure = "msse", accuracy = rmsse^2),
-  mase |> mutate(measure = "mase", accuracy = mase),
-  crps |> mutate(measure = "crps", accuracy = crps),
-) |>
-  select(-rmsse, -mase, -crps)
-
-# Plot of average accuracy vs week for each method for Total
-acc_summary <- accuracy |>
-  filter(
-    series %in% c("Total", "Control areas", "Health boards"),
-    method == "mint", model != "qcomb"
-  ) |>
-  mutate(
-    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
-    model = factor(model, levels = c("naiveecdf", "ets", "tscount", "iglm", "ensemble")),
-    week = factor(trunc((h - 1) / 7) + 1)
-  ) |>
-  group_by(week, model, measure, series) |>
-  summarise(accuracy = mean(accuracy), .groups = "drop")
-
-acc_summary |>
-  ggplot(aes(x = week, y = accuracy, group = model)) +
-  geom_line(aes(col = model), size = 1) +
-  facet_grid(measure ~ series, scales = "free_y") +
-  labs(y = "Average accuracy", x = "Week ahead")

### 2022-11-18 letter/cover_letter.qmd
typo

@@
 Our study is fully reproducible and we provide data and code through a Github repository enabling reproduction and refinement of the proposed approach and benchmarks.
-The proposed approach could also be generalised to forecast requirements in other healthcare services, such as admissions in Emergency Departement, or call volumes in clinical desk services.
+The proposed approach could also be generalized to forecast requirements in other healthcare services, such as admissions in Emergency Department, or call volumes in clinical desk services.
 

### 2022-11-20 manuscript/01-introduction.qmd
final draft from bahman

@@
 
-However, despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the hierarchical data structure has been largely ignored, and the main focus has been on producing independent forecasts at a single level.
+However, despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the hierarchical data structure has been largely ignored, and the main focus has been on producing independent (base) forecasts at a single level.
 Generating independent forecasts can result in a lack of consistency and coordination, and therefore leads to less effective planning and decision making.
@@
 
-In this paper, we address this gap by investigating the application of hierarchical forecasting approaches in the EMS using daily time series of verified incidents from 2015 to 2020 in a major ambulance service in Great Britain.
+In this paper, we address this gap by investigating the application of hierarchical forecasting approaches in the EMS using daily time series of attended incidents from 2015 to 2020 in a major ambulance service in Great Britain.
 The data has hierarchical and grouped structures, with hierarchies at the national, control (i.e.
@@
 We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal reconciliation approaches.
-We first generate independent/base forecasts using Exponential Smoothing State Space (ETS), Generalized Linear Model (GLM), Poisson regression, a simple empirical distribution and an ensemble method, followed by applying bottom-up and optimal reconciliation approaches.
-Forecast performance is assessed by the Mean Squared Scaled Error (MSSE) for point forecasts and Continuous Ranked Probability Scores (CRPS) for the probabilistic forecasts.
-This paper complies with reproducibility principles [@stodden2013best; @boylan2015reproducibility], and can be applied to any healthcare service (e.g., emergency department, primary or social care) subject to the time series having a hierarchical and/or grouped structure.
+We first generate independent/base forecasts using Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method, followed by applying bottom-up and optimal reconciliation approaches.
+Forecast performance is assessed by the Mean Absolute Scaled Error (MASE) and  Mean Squared Scaled Error (MSSE) for point forecasts and Continuous Ranked Probability Scores (CRPS) for the probabilistic forecasts.
+This paper complies with reproducibility principles [@stodden2013best; @boylan2015reproducibility].
+We provide the R codes for the proposed models and benchmarks.
+Therefore, they can be applied to any healthcare service (e.g., emergency department, primary or social care) subject to the time series having a hierarchical and/or grouped structure.
 

### 2022-11-20 manuscript/03-experiment.qmd
final draft from bahman

@@
 
-Given the presence of various significant patterns in the past attended incidents, we consider three different forecasting models to generate the base forecasts.
+Given the presence of various patterns in the past attended incidents, we consider three different forecasting models to generate the base forecasts.
 Once the base forecasts are produced, hierarchical and grouped time series methods are used to reconcile them across all levels.
@@
 **Ensemble method:** Finally, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models to form a mixture distribution [@combinations].
-
-<!-- add a sentence to describe how the point forecast and forecast distribution is created for the ensemble -->
 

### 2022-11-20 manuscript/04-htsgts.qmd
final draft from bahman

@@
 
-Let $\bm{b}_t$ be a vector of $n_b$ "bottom-level" time series at time $t$, and let $\bm{a}_t$ be a corresponding vector of $n_a = n-n_b$ aggregated time series, where
+Let $\bm{b}_t$ be a vector of $n_b$ _bottom-level_ time series at time $t$, and let $\bm{a}_t$ be a corresponding vector of $n_a = n-n_b$ aggregated time series, where
 $$

### 2022-11-20 manuscript/05-results.qmd
final draft from bahman

--- 2022-11-18 manuscript/05-results.qmd
@@
 @tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom level series.
-Point forecast performance are reported using MASE and MSSE in @tbl-result-1  and @tbl-result-2, respectively.
+Point forecast performance are reported using MASE and MSSE in @tbl-result-1  and [-@tbl-result-2], respectively.
 Probabilistic forecast accuracy is reported using CRPS in @tbl-result-3.
@@
 
-@tbl-result-1 and 2b show that forecast reconciliation (i.e.
+@tbl-result-1 and [-@tbl-result-2] show that forecast reconciliation (i.e.
 MinT) improves forecast accuracy at the higher levels of the hierarchy including total, control area and health board.
@@
   ) |>
+  mutate(model = case_when(
+    model == "naiveecdf" ~ "Naive",
+    model == "ets" ~  "ETS",
+    model == "tscount" ~ "TSGLM",
+    model == "iglm" ~ "GLM",
+    model == "ensemble" ~ "Ensemble"
+  )) |> 
   mutate(
@@
     series = factor(series, levels = c("Total", "Control areas", "Health boards")),
-    model = factor(model, levels = c("naiveecdf", "ets", "tscount", "iglm", "ensemble")),
+    model = factor(model, levels = c("Naive", "ETS", "TSGLM", "GLM", "Ensemble")),
     week = factor(trunc((h - 1) / 7) + 1)

### 2022-11-20 manuscript/06-conclusion.qmd
final draft from bahman

@@
 Forecasting problems at Emergency Medical Services often have inherent hierarchical and grouped structures.
-For example, looking at time series of arrival calls in a clinical desk service, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest.
+For example, looking at time series of arrival calls in a clinical desk service, Emergency Department admissions, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest.
 Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area.
@@
 Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through more holistic planning approach.
-While, we generated count attend incident for the base forecast using Poisson regression models, however the reconciles forecast is snot yet count.
+While, we generated count attend incident for the base forecast using Poisson regression models, however the reconciles forecast is not yet count.
 This could be also an avenue for further research.

### 2022-11-20 manuscript/main.qmd
final draft from bahman

@@
     keep-tex: true
-    include-in-header: preamble.tex
+    include-in-header: 
+      - text: \usepackage{bm}
+      - file: preamble.tex
     journal:

### 2022-11-21 manuscript/05-results.qmd
Now producing docx as well as pdf

@@
 #| label: tbl-result
-#| tbl-cap: "Average forecast perfomance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data.
+#| tbl-cap: "Average forecast performance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data.
 The best approach is highlighted in bold."
@@
     table[[i]] <- sprintf(table[[i]], fmt="%1.3f")
-    table[[i]][best] <- cell_spec(table[[i]][best], "latex", bold=TRUE)
+    table[[i]][best] <- cell_spec(table[[i]][best], bold=TRUE)
   }

### 2022-11-21 manuscript/main.qmd
Now producing docx as well as pdf

--- 2022-11-18 manuscript/main.qmd
@@
 pdf-engine: pdflatex
+prefer-html: true
 format:
+  docx:
+    number-sections: true
   elsevier-pdf:
     keep-tex: true
-    include-in-header: preamble.tex
+    include-in-header:
+      text: |
+        \usepackage{todonotes,mathtools,bm,amsmath,mathpazo}
+        \mathtoolsset{showonlyrefs}
+        \setlength{\parindent}{0cm}
     journal:

### 2022-11-21 letter/cover_letter.qmd
typo

@@
 
-We would be grateful if you could consider for publication our manuscript entitled _Hierarchical Time Series Forecasting in Emergency Medical Services_ as an Original Research article in Annals of Emergency Medicine.
+We would be grateful if you could consider for publication our manuscript entitled _Hierarchical Time Series Forecasting in Emergency Medical Services_ as an Original Research article in the Annals of Emergency Medicine.
 
 Time series forecasting problems at Emergency Medical Services have often inherent hierarchical and grouped structures.
-For example the ambulance demand (e.g.
-arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area.
-Demand could also be disaggrgated by other attributes that such as priority (e.g Red, Amber, Green) or nature of incidents.
-Despite the popularity of such data structure in EMS, the common practice is to produce point forecast for each time series independently and ignore the structure.
-This practice may lead to the lack of coordination and possibly undesirable and conflicting outcomes.
-Practices also rely on point forecasts, which does not  acknowledge forecast uncertainty.
+For example, the ambulance demand (e.g.
+arrival calls, verified incidents, attended incidents) in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the postcode area.
+Demand could also be disaggrgated by other attributes such as priority (e.g Red, Amber, Green) or the nature of incidents.
+Despite the popularity of such data structure in EMS, the common practice is to produce point forecasts for each time series independently and ignore the structure.
+This practice may lead to a lack of coordination and possibly undesirable and conflicting outcomes.
+Practices also rely on point forecasts, which do not acknowledge forecast uncertainty.
 
 In this paper, we aim at filling these gaps identified in the academic literature and practice.
-We produce probabilistic forecasts of daily incident attended up to 84 days ahead using hierarchical forecasting methods for using state of the art forecasting methods.
+We produce probabilistic forecasts of daily incidents attended up to 84 days ahead using hierarchical forecasting methods for using state of the art forecasting methods.
 We apply hierarchical forecasting methods to produce a probabilistic and coherent forecast of attended incidents at multiple levels of aggregation.
-We use Exponential Smoothing State Space models, Poisson regression based on General Linear Models and tscount models, a Naive empirical distribution, and the ensemble model to generate base forecasts.
+We use Exponential Smoothing State Space models, Poisson regression models based on General Linear Models and serial dependencies, a Naive empirical distribution, and an ensemble model to generate base forecasts.
 These forecasting models collectively account for i) the spline trend, ii) multiple seasonalities, iii) holiday effects, and  iv) serial dependencies in the time series.
-Our study is fully reproducible and we provide data and code through a Github repository enabling reproduction and refinement of the proposed approach and benchmarks.
+Our study is fully reproducible and we provide the R code through a Github repository enabling reproduction and refinement of the proposed approach and benchmarks.
 The proposed approach could also be generalized to forecast requirements in other healthcare services, such as admissions in Emergency Department, or call volumes in clinical desk services.

### 2022-11-21 manuscript/01-introduction.qmd
typo

@@
 
-A failure to match available resources to demand in Emergency Medical Services (EMS) results in patient flow problems, with serious consequences for patients, staff and the entire care system [@ekstrom2015forecasting; @ROSTAMITABAR20221197].
+A failure to match available resources to demand in Emergency Medical Services (EMS) results in patient flow problems, with serious consequences for patients, staff, and the entire care system [@ekstrom2015forecasting; @ROSTAMITABAR20221197].
 Demand forecasting in EMS helps service planners to avoid the mismatch, potentially providing massive savings in costs and lives, and leading to better patient outcomes.
-Accurate daily demand forecasting enables planners and decision makers to manage resources to meet anticipated patients, reconfigure units, and redeploy staff and vehicles as necessary.
+Accurate daily demand forecasting enables planners and decision-makers to manage resources to meet anticipated patients, reconfigure units, and redeploy staff and vehicles as necessary.
 
 Demand forecasts at EMS are typically required at multiple levels of an organization to inform various planning and decision-making processes [@hulshof2012taxonomic].
-There are some planning process at the national level (strategic and long-term) such as workforce resource planning and budgeting; sub-national, regional or healthcare level (tactical and medium-term) such as temporary capacity expansions, resource sharing and staff-shift scheduling; and hospital or station level (operational and short-term) such as planning rosters for staff and ambulance deployment.
+There are some planning processes at the national level (strategic and long-term) such as workforce resource planning and budgeting; sub-national, regional, or healthcare level (tactical and medium-term) such as temporary capacity expansions, resource sharing, and staff-shift scheduling; and hospital or station level (operational and short-term) such as planning rosters for staff and ambulance deployment.
 Demand forecasts might also be required at different levels for a specific area of interest such as the nature of demand or the priority level.
@@
 The data has hierarchical and grouped structures, with hierarchies at the national, control (i.e.
-sub-national), health board (i.e.
+sub-national), and health board (i.e.
 regional) levels, as well as groups by priority and nature of incidents.

### 2022-11-21 manuscript/02-literature.qmd
final draft from bahman
typo

@@
 Another demand related variable is verified incidents; these are all incidents that require an action: either by sending a physical vehicle, responding via the Clinical Support Desk, requesting an external provider to respond to it, or forwarding it to other channels such as police, firefighters or general practitioners.
-Our study is aligned with this stream of the literature.
-Another similar area that has received considerable attention is Emergency Department forecasting; we refer interested readers to @mingliterature2022, @gul2020exhaustive and @wargon2009systematic for extensive reviews of the relevant literature.
-Although crucial to EMS performance, @aringhieri2017emergency state that demand forecasting has received limited research attention in the EMS context.
+Our study is aligned with this stream of literature.
+Another similar area that has received considerable attention is Emergency Department forecasting; we refer interested readers to @mingliterature2022, @gul2020exhaustive, and @wargon2009systematic for extensive reviews of the relevant literature.
 In this section, we provide a brief review of studies on forecasting ambulance demand in EMS.
@@
 The census variables describe groups who are more likely to need an ambulance.
-A stepwise ordinary least squares regression analysis is used for estimating the parameter and generating forecast.
+A stepwise ordinary least squares regression analysis is used for estimating the parameter and generating forecasts.
 The only performance measure reported in this study is $R^2$, which is not an effective measure of forecast accuracy [@armstrong01, p457].
 The research design of this study is not rigorous and the study is not reproducible.
-@vile2012predicting explore using a Singular Spectrum Analysis (SSA) method to generate forecasts of the EMS demand at the national level for 7-day, 14-day, 21-day and 28-day forecast horizons using data provided by an ambulance service in Great Britain.
+@vile2012predicting explore using a Singular Spectrum Analysis (SSA) method to generate forecasts of the EMS demand at the national level for 7-day, 14-day, 21-day, and 28-day forecast horizons using data provided by an ambulance service in Great Britain.
 The performance of this approach is compared to Auto-Regressive Integrated Moving Average (ARIMA) and Holt-Winters time series methods using Root Mean Squared Error (RMSE).
@@
 @vile2016time further develop a decision support system to integrate forecasts generated by SSA.
-However, the study does not compare and contrast the performance of forecasting methods based on utility measures such as cost, resource utilization or response time.
-The tool contains options that allow generating forecasts at various levels of granularity, however, it ignores the hierarchical and grouped relationships structure, preventing aligned decision making and coordination.
+However, the study does not compare and contrast the performance of forecasting methods based on utility measures such as cost, resource utilization and response time.
+The tool contains options that allow generating forecasts at various levels of granularity, however, it ignores the structure of the hierarchical and grouped relationships, preventing aligned decision making and coordination.
 
-@9659837 investigate forecasting EMS demand in a high spatio-temporal resolution of 1km$^2$ spatial regions and 1-hr time intervals using total incidents in Oslo, Norway, from 1 January 2015 to 11 February 2019.
+@9659837 investigate forecasting EMS demand in a high Spatio-temporal resolution of 1km$^2$ spatial regions and 1-hr time intervals using total incidents in Oslo, Norway, from 1 January 2015 to 11 February 2019.
 They used multi-layer perceptron (MLP) and long short-term memory (LSTM) models to forecast the EMS demand, and compare the results to simple aggregation methods and baselines.
 The point forecast accuracy is evaluated using Mean Absolute Error (MAE) and Mean Squared Error (MSE), and the forecast distribution is measured by Categorical Cross-Entropy.
-They found that Neural Network models performed better in producing point forecasts, while a distribution baseline method based on spatial distribution of the incidents across all time steps provided more accurate forecast distributions.
+They found that Neural Network models performed better in producing point forecasts, while a distribution baseline method based on the spatial distribution of the incidents across all time steps provided more accurate forecast distributions.
 @zhou2016predictinglit proposed three methods based on Gaussian mixture models, kernel density estimation, and kernel warping to predict hourly data 4 weeks ahead for a 1km$^2$ spatial region.
-Two years of incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011 and 2012 with 696,975 events) were used to build the model and examine the performance on test data using mean negative log likelihood.
+Two years of incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011 and 2012 with 696,975 events) were used to build the model and examine the performance on test data using mean negative log-likelihood.
 They show that forecasts generated by the proposed methods were significantly more accurate than the current industry practice (a simple averaging formula).
 @grekousis2019will investigated the combination of spatial analysis methods with data mining techniques based on an improved Hungarian algorithm and a MLP neural network to identify the most likely locations of future emergency events.
-The proposed approach was tested using data of 2851 events attended by the EMS in Athens, Greece, over 24 weeks.
+The proposed approach was tested using data from 2851 events attended by the EMS in Athens, Greece, over 24 weeks.
 They showed that 23% of real emergency events lie within 50 meters of the predicted ones and nearly 70% of the real emergency events lie no further than 150 meters away, which is rather accurate given the granularity of the problem at the city level.

### 2022-11-21 manuscript/03-experiment.qmd
typo

@@
 #| fig-subcap:
-#| - "Hierarchical structure: Attended incidents in all country is disaggregated into 3 control areas /regions and then into 7 different healthboard at the bottom"
+#| - "Hierarchical structure: Attended incidents in all country is disaggregated into 3 control areas and then into 7 different healthboard"
 #| - "Grouped structure: Incidents could be grouped into priority (i.e.

### 2022-11-21 manuscript/05-results.qmd
typo

@@
 
-In this section, we compare the forecasting performance of the Naive, ETS, GLM and TSGLM models along with the ensemble, using base forecast and Minimum Trace (MinT) reconciliation methods.
-We have also computed the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches, along with bottom up forecasting.
+In this section, we compare the forecasting performance of the Naive, ETS, GLM, and TSGLM models along with the ensemble, using base forecast and Minimum Trace (MinT) reconciliation methods.
+We have also computed the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches, along with bottom-up forecasting.
 However, they are not reported here because their accuracy is outperformed by MinT.
@@
 But to save space, we have reported only the top level (Total), the bottom level, and the levels corresponding to Control areas and Health boards.
-The latter are chosen because this is where decision-making takes place, so these forecasts are the most important.
+The latters are chosen because this is where decision-making takes place, so these forecasts are the most important.
 
-The overall forecasting performance is reported in @tbl-result, in which the average forecast accuracy over horizons 43--84 days (corresponding to the planning horizon) is presented per model, method and the hierarchical level.
-Reported forecast accuracy are averaged across all forecast horizons, rolling origins and series at each level.
+The overall forecasting performance is reported in @tbl-result, in which the average forecast accuracy over horizons 43--84 days (corresponding to the planning horizon) is presented per model, method, and the hierarchical level.
+Reported forecast accuracy is averaged across all forecast horizons, rolling origins, and series at each level.
 @tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom level series.
-Point forecast performance are reported using MASE and MSSE in @tbl-result-1  and [-@tbl-result-2], respectively.
+Point forecast performance is reported using MASE and MSSE in @tbl-result-1  and [-@tbl-result-2], respectively.
 Probabilistic forecast accuracy is reported using CRPS in @tbl-result-3.
@@
 The smaller the value of CRPS, the better the comprehensive performance.
-We observe that forecast reconciliation results in forecast improvement, regardless of the hierarchical level.
+We observe that forecast reconciliation results in forecast improvement for the total and health board level.
+CRPS is almost identical at the control area and health board levels.
+Base forecasts are slightly better at the control area level, while reconciliation is marginally accurate than base at the bottom level.
 The ensemble method is also more accurate for higher levels, but ETS performs slightly better at the bottom level.
-While these results show that forecast reconciliation does not improve *point* forecasts at the bottom level, @tbl-result-3 indicates that it generates more accurate *distributional* forecasts than the base method.
-This is probably due to the reconciliation method giving improved forecast accuracy in the tails of the forecast distribution, which are critical for managing risks.
+While these results show that forecast reconciliation does not improve *point* forecasts at the bottom level, @tbl-result-3 indicates that it generates accurate *distributional* forecasts.
+The marginal improvement in the average probabilistic forecast accuracy at the bottom level might be due to the reconciliation method giving improved forecast accuracy in the tails of the forecast distribution, which are critical for managing risks.
 
-Overall, our results indicate that forecast reconciliation using the MinT method provides reliable forecasts, improves upon the base (unreconciled) forecasts all levels except the bottom level series.
+Overall, our results indicate that forecast reconciliation using the MinT method provides reliable forecasts, and improves upon the base (unreconciled) forecasts at all levels except the bottom-level series.
 But even there, forecast reconciliation using MinT improves accuracy in the tails of the distribution.
@@
 In addition to the overall forecast accuracy presented in @tbl-result, we also report the point and probabilistic forecast accuracy measures for each forecast horizon in @fig-accuracy.
-The figure focuses on the hierarchical levels important for decision-making including total, control area and health board; however the accuracy could be calculated for any level.
+The figure focuses on the hierarchical levels important for decision-making including total, control area, and health board; however, the accuracy could be calculated for any level.
 We only illustrate the results of the MinT method, given its strong performance described in @tbl-result.
@@
 Thus the x-axis shows horizons from week 1 ($h= 1,\dots,7$) to week 12 ($h= 78,\dots,84$).
-The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision making.
+The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision-makers.
 For both the point forecast and distributional accuracy  we can see that the ensemble approach performs best across almost all horizons, with the biggest differences at the highest levels of aggregation.
@@
 #| out.width: "100%"
-#| fig-cap: "Average accuracy by week for 12 weeks using MinT reconciliation. CRPS is relative to a naive ECDF. MASE and MSSE are relative to the corresponding values for the training set."
+#| fig-cap: "Average accuracy by week for 12 weeks using MinT reconciliation.
+CRPS is relative to a naive Empirical Cumulative Distribution Function (ECDF).
+MASE and MSSE are relative to the corresponding values for the training set."
 #| fig-pos: "H"

### 2022-11-21 manuscript/06-conclusion.qmd
typo

@@
 For example, looking at time series of arrival calls in a clinical desk service, Emergency Department admissions, verified incidents, or attended incidents in a country, they could be disaggregated by various attributes of interest.
-Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the post code area.
+Total demand in the country could be disaggregated by region, then within each region by health board, within each health board, by station/hospital, and so on down to the postcode area.
 Alternative structures may arise when attributes of interest are crossed rather than nested.
-For example, the total demand could be disaggregated by priority (e.g., Red, Amber, Green) or by nature of incidents.
-It is also natural to have a mixed structures, for example the total demand could be disaggregated by priority and also by health board.
+For example, the total demand could be disaggregated by priority (e.g., Red, Amber, Green) or by the nature of incidents.
+It is also natural to have a mixed structure, for example, the total demand could be disaggregated by priority and also by health board.
 
@@
 
-Our results indicate that forecast reconciliation in EMS can not only contribute to a more coordinated approach to decision making through producing coherent forecasts, but also it can result in forecast accuracy improvement.
+Our results indicate that forecast reconciliation in EMS can not only contribute to a more coordinated approach to the planning and decision-making through producing coherent forecasts, but also it can result in forecast accuracy improvements.
 Our proposed forecasting models, combined with reconciliation approaches, outperform the empirical distribution benchmark.
 We show that a substantial forecast improvement can be achieved at higher levels of aggregation by applying forecast reconciliation methods.
-When a point forecast is of interest at the bottom level of series, we observe that reconciliation may not improve the forecast accuracy, if the bottom series are noisy and lack systematic patterns.
-However, forecast reconciliation may result in more forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e.
+When a point forecast is of interest at the bottom level of the series, we observe that reconciliation may not improve the forecast accuracy, if the bottom series are noisy and lack systematic patterns.
+However, forecast reconciliation may result in more accurate forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e.
 point forecast).
-Producing consistent forecast are also crucial for informing planning activities, we also demonstrate that proposed models produce consistent forecasts across all forecast horizons.
-Therefore, we recommend that forecast reconciliation approaches to be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted.
+Producing consistent forecasts are also crucial for informing planning activities, we also demonstrate that proposed models produce consistent forecasts across all forecast horizons.
+Therefore, we recommend that forecast reconciliation approaches be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted.
 Moreover, we found that using an ensemble forecasting model, combining all models developed in this paper, instead of using each individually, works remarkably well for our mixed hierarchical & grouped structure.
@@
 Further research could investigate the practical benefits of probabilistic hierarchical forecasting in EMS.
-Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through more holistic planning approach.
-While, we generated count attend incident for the base forecast using Poisson regression models, however the reconciles forecast is not yet count.
+Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.
+While we generated count attend incidents for the base forecast using Poisson regression models, however, the reconciled forecasts are not yet count.
 This could be also an avenue for further research.

### 2022-11-21 manuscript/main.qmd
typo

@@
   Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet.
-Such demand forecasts are required at national, regional and sub-regional levels, and must take account of the nature of incidents and their priorities.
+Such demand forecasts are required at national, regional, and sub-regional levels, and must take account of the nature of incidents and their priorities.
 These forecasts are often generated independently by different teams within the organization.
 As a result, forecasts at different levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service.
-To address this issue, we exploit the hierarchical and grouped structure of the demand time series, and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation.
+To address this issue, we exploit the hierarchical and grouped structure of the demand time series and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation.
 The methods are applied to daily incident data from an ambulance service in Great Britain, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area.
-We use an ensemble of forecasting models, and show that the resulting forecasts are better than any individual forecasting model.
+We use an ensemble of forecasting models and show that the resulting forecasts are better than any individual forecasting model.
 We validate the forecasting approach using time-series cross-validation.
@@
 format:
+  html:
+    number-sections: true
   docx:

### 2022-11-21 manuscript/03-experiment.qmd
Some final changes

@@
 
-We are interested in generating forecasts to inform the planning horizon of 42 days, required by planners in the ambulance service.
-The forecast horizon in this study is $2 \times 42 = 84$ days ahead, because the planning is generally frozen for 42 days and so forecasts of the next 42 days is not particularly helpful for planning.
-While forecasts are generated for 84 days ahead, performance evaluation is only assessed based on the last 42 days and not the whole forecast period.
+Planners in the ambulance service work with a planning horizon of 6 weeks.
+That is, planning is generally frozen for the next 42 days, so any forecasts will only affect plans for the time period beyond the next 42 days.
+Consequently, the forecast horizon in this study is $2 \times 42 = 84$ days ahead, with performance evaluation assessed based on the last 42 days and not the whole forecast period.
 The forecasts are produced for various training and test sets using time series cross-validation [@hyndman2021forecasting].
@@
 #| fig-subcap:
-#| - "Hierarchical structure: Attended incidents in all country is disaggregated into 3 control areas and then into 7 different healthboard"
+#| - "Hierarchical structure: Attended incidents in the whole country are disaggregated into 3 control areas and then into 7 different healthboards."
 #| - "Grouped structure: Incidents could be grouped into priority (i.e.
 Red, Amber & Green) and the nature of attended incident (i.e.
-there are 35 different nature of incidents including chest pain , breathing problems, heart attack, stroke, and so on).
+there are 35 different nature of incidents including chest pain, breathing problems, heart attack, stroke, and so on).
 The symbol * refers to the crossed attributes between hierarchical and grouped levels."
@@
 #| fig-cap: "Time series of attended incidents at various levels.
-The panels show data from the whole country, by control area, by health board, by priority level, and by nature of incident.
+The panels show data from the whole country (top panel), by control area, by health board, by priority level, and by nature of incident.
 Only four of the 35 nature of incident categories are shown to avoid too much overplotting."
@@
 The trend component can be none ("N"), Additive ("A") or damped ("Ad"); the seasonality can be none ("N"), Additive ("A"), or multiplicative ("M"); and the error term can be additive ("A") or multiplicative ("M").
-To forecast the attended incidents at each level, we use the `ETS()` function in the `fable` package [@fable2022] in R.
-To identify the best model for a given time series, the `ETS` function uses the corrected Akaike’s Information Criterion (AICc).
+To forecast the attended incidents at each level, we use the `ets()` function in the `forecast` package [@Rforecast;@HK08] in R.
+To identify the best model for a given time series, the `ets` function uses the corrected Akaike’s Information Criterion (AICc).
 

### 2022-11-21 manuscript/04-htsgts.qmd
Some final changes

@@
 However, given that we have multiple grouped attributes combined with the hierarchical structure, there is no unique way to disaggregate top forecasts.
-Hence the top-down cannot be used in our application.
+Hence the top-down approach cannot be used in our application.
 The recommended approach is to use forecast reconciliation [@hyndman2011optimal].
@@
 The BU approach can capture the dynamics of the series at the bottom level, but these series may be noisy and difficult to forecast.
-The approach using only the data at the most disaggregated level, and so does not utilize all the information available across the hierarchical and grouped structure.
+The approach uses only the data at the most disaggregated level, and so does not utilize all the information available across the hierarchical and grouped structure.
 

### 2022-11-21 manuscript/06-conclusion.qmd
Some final changes

@@
 point forecast).
-Producing consistent forecasts are also crucial for informing planning activities, we also demonstrate that proposed models produce consistent forecasts across all forecast horizons.
+Coherent forecasts are also crucial for informing planning activities, and we demonstrate that the proposed models produce coherent forecasts across all forecast horizons.
 Therefore, we recommend that forecast reconciliation approaches be adopted for routine use in EMS, whenever hierarchical and/or grouped time series data need to be forecasted.
@@
 Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.
-While we generated count attend incidents for the base forecast using Poisson regression models, however, the reconciled forecasts are not yet count.
+While we generated count distributions of attended incidents for the base forecasts using Poisson regression models, the reconciled forecast distributions are not counts.
 This could be also an avenue for further research.

### 2022-11-21 manuscript/title_page.qmd
Some final changes

@@
 We use an ensemble of forecasting models, and show that the resulting forecasts are better than any individual forecasting model.
-We validate the forecasting approach using time-series cross-validation.
+We validate the forecasting approach using time series cross validation.
 keywords:
@@
     keep-tex: true
-    include-in-header: preamble.tex
+    include-in-header:
+      text: |
+        \usepackage{mathpazo}
+        \setlength{\parindent}{0cm}
     journal:
@@
 - **Funding information:** A funding awarded by Cardiff Business School allowed BRT to travel to Melbourne to visit RJH to discuss the research design, run the experiment and write the article.
+RJH received funding through the Australian Research Council Industrial Transformation Training Centre in Optimisation Technologies, Integrated Methodologies, and Applications (OPTIMA), Project ID IC200100009.
 

### 2022-11-22 manuscript/01-introduction.qmd
MS word document generated for submission, plus typos corrected

@@
 Hierarchical forecasting framework can be used as a tool to improve coordination between teams across the care services at the national, sub-national, regional and local levels.
-The hierarchical forecasting approaches not only create consistent forecasts, but are usually also more accurate than the independent (base) forecasts [@hyndman2011optimal].
+The hierarchical forecasting approaches not only create consistent forecasts but are usually also more accurate than the independent (base) forecasts [@hyndman2011optimal].
 To our knowledge, there has been no previous research involving hierarchical and grouped forecasting in the entire field of forecasting for healthcare management.

### 2022-11-22 manuscript/02-literature.qmd
MS word document generated for submission, plus typos corrected

@@
 Emergency medical services (EMS) are a critical component in the delivery of urgent medical care to communities.
-An effective service delivery requires accurate resource planning that generally relies on demand forecasts at operational, tactical and strategic levels.
+An effective service delivery requires accurate resource planning that generally relies on demand forecasts at operational, tactical, and strategic levels.
 
@@
 This might be also called _attended incidents_.
-Another demand related variable is verified incidents; these are all incidents that require an action: either by sending a physical vehicle, responding via the Clinical Support Desk, requesting an external provider to respond to it, or forwarding it to other channels such as police, firefighters or general practitioners.
+Another demand related variable is verified incidents; these are all incidents that require an action: either by sending a physical vehicle, responding via the Clinical Support Desk, requesting an external provider to respond to it, or forwarding it to other channels such as police, firefighters, or general practitioners.
 Our study is aligned with this stream of literature.
@@
 However, the study does not compare and contrast the performance of forecasting methods based on utility measures such as cost, resource utilization and response time.
-The tool contains options that allow generating forecasts at various levels of granularity, however, it ignores the structure of the hierarchical and grouped relationships, preventing aligned decision making and coordination.
+The tool contains options that allow generating forecasts at various levels of granularity; however, it ignores the structure of the hierarchical and grouped relationships, preventing aligned decision making and coordination.
 

### 2022-11-22 manuscript/03-experiment.qmd
MS word document generated for submission, plus typos corrected

@@
 There are 3 control areas breakdown by 7 local health boards.
-Attended incident data are categorized into 3 priority classes of red, amber and green.
+Attended incident data are categorized into 3 priority classes of red, amber, and green.
 There are also 35 different nature of incidents such as chest pain, stroke, breathing problem, etc.
@@
 In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation.
-@fig-dataviz2 reveals different information such as trend, seasonality and noise.
+@fig-dataviz2 reveals different information such as trend, seasonality, and noise.
 For example, some series depict seasonality and trend, whereas some other series report low volume of attended incidents and entropy, making them more volatile and difficult to forecast.
@@
 We use public holidays, school holidays and Christmas Day and New Year's Day as predictors of incident attended.
-These type of holidays will affect peoples' activities and may increase or decrease the number of attended incidents.
+These types of holidays will affect peoples' activities and may increase or decrease the number of attended incidents.
 
@@
 
-**Exponential Smoothing State Space model (ETS):** ETS models [@hyndman2021forecasting] can combine trend, seasonality and error components in a time series through various forms that can be additive, multiplicative or mixed.
+**Exponential Smoothing State Space model (ETS):** ETS models [@hyndman2021forecasting] can combine trend, seasonality, and error components in a time series through various forms that can be additive, multiplicative or mixed.
 The trend component can be none ("N"), Additive ("A") or damped ("Ad"); the seasonality can be none ("N"), Additive ("A"), or multiplicative ("M"); and the error term can be additive ("A") or multiplicative ("M").

### 2022-11-22 manuscript/04-htsgts.qmd
MS word document generated for submission, plus typos corrected

@@
 These series are generally hierarchically organized based on multiple levels such as area/region, health board and/or are aggregated at different levels in groups based on nature of demand, priority of demand, or some other attributes.
-While series could be strictly hierarchical or only grouped bases on some attributes, in many situation a more complex structures arise when attributes of interest are both nested and crossed, having hierarchical and grouped structure.
+While series could be strictly hierarchical or only grouped bases on some attributes, in many situations a more complex structures arise when attributes of interest are both nested and crossed, having hierarchical and grouped structure.
 This is also the case for our application as discussed in @sec-data.

### 2022-11-22 manuscript/05-results.qmd
MS word document generated for submission, plus typos corrected

@@
 Base forecasts are slightly better at the control area level, while reconciliation is marginally accurate than base at the bottom level.
-The ensemble method is also more accurate for higher levels, but ETS performs slightly better at the bottom level.
-While these results show that forecast reconciliation does not improve *point* forecasts at the bottom level, @tbl-result-3 indicates that it generates accurate *distributional* forecasts.
+The ensemble method is also more accurate for higher levels, but ETS performs well at the bottom level.
+@tbl-result-3 indicates that it generates accurate *distributional* forecasts.
 The marginal improvement in the average probabilistic forecast accuracy at the bottom level might be due to the reconciliation method giving improved forecast accuracy in the tails of the forecast distribution, which are critical for managing risks.
 
-Overall, our results indicate that forecast reconciliation using the MinT method provides reliable forecasts, and improves upon the base (unreconciled) forecasts at all levels except the bottom-level series.
+Overall, our results indicate that forecast reconciliation using the MinT method provides reliable forecasts and improves upon the base (unreconciled) forecasts at all levels except the bottom-level series.
 But even there, forecast reconciliation using MinT improves accuracy in the tails of the distribution.
@@
 For illustration purposes, we report the average weekly forecast accuracy instead of the daily forecast horizon, as this reduces the visual noise in the figure.
-Thus the x-axis shows horizons from week 1 ($h= 1,\dots,7$) to week 12 ($h= 78,\dots,84$).
+Thus, the x-axis shows horizons from week 1 ($h= 1,\dots,7$) to week 12 ($h= 78,\dots,84$).
 The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision-makers.

### 2022-11-22 manuscript/06-conclusion.qmd
MS word document generated for submission, plus typos corrected

@@
 For example, the total demand could be disaggregated by priority (e.g., Red, Amber, Green) or by the nature of incidents.
-It is also natural to have a mixed structure, for example, the total demand could be disaggregated by priority and also by health board.
+It is also natural to have a mixed structure, for example, the total demand could be disaggregated by priority and by health board.
 
@@
 We show that a substantial forecast improvement can be achieved at higher levels of aggregation by applying forecast reconciliation methods.
-When a point forecast is of interest at the bottom level of the series, we observe that reconciliation may not improve the forecast accuracy, if the bottom series are noisy and lack systematic patterns.
+When a point forecast is of interest at the bottom level of the series, we observe that reconciliation may not improve the forecast accuracy if the bottom series are noisy and lack systematic patterns.
 However, forecast reconciliation may result in more accurate forecast results for bottom series, if we are interested in the tails of forecast distribution rather than just center measures like mean (i.e.

### 2022-11-22 manuscript/title_page.qmd
MS word document generated for submission, plus typos corrected

@@
   Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet.
-Such demand forecasts are required at national, regional and sub-regional levels, and must take account of the nature of incidents and their priorities.
+Such demand forecasts are required at national, regional, and sub-regional levels, and must take account of the nature of incidents and their priorities.
 These forecasts are often generated independently by different teams within the organization.
 As a result, forecasts at different levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service.
-To address this issue, we exploit the hierarchical and grouped structure of the demand time series, and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation.
+To address this issue, we exploit the hierarchical and grouped structure of the demand time series and apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all the available data at all levels of disaggregation.
 The methods are applied to daily incident data from an ambulance service in Great Britain, from October 2015 to July 2019, disaggregated by nature of incident, priority, managing health board, and control area.
-We use an ensemble of forecasting models, and show that the resulting forecasts are better than any individual forecasting model.
+We use an ensemble of forecasting models and show that the resulting forecasts are better than any individual forecasting model.
 We validate the forecasting approach using time series cross validation.

### 2022-11-22 manuscript/05-results.qmd
typos

@@
 We observe that forecast reconciliation results in forecast improvement for the total and health board level.
-CRPS is almost identical at the control area and health board levels.
+CRPS is almost identical at the control area and botom levels.
 Base forecasts are slightly better at the control area level, while reconciliation is marginally accurate than base at the bottom level.
 The ensemble method is also more accurate for higher levels, but ETS performs well at the bottom level.
-@tbl-result-3 indicates that it generates accurate *distributional* forecasts.
+@tbl-result-3 indicates that reconciliation using Mint  generates accurate *distributional* forecasts.
 The marginal improvement in the average probabilistic forecast accuracy at the bottom level might be due to the reconciliation method giving improved forecast accuracy in the tails of the forecast distribution, which are critical for managing risks.

### 2023-05-10 journal_correspondence/response_letter/response_letter_JSR.qmd
response letter - reposne to AE and reviewer 1

---
title: "Response letter"
format: pdf
---

Dear Sir or Madam,

Please find enclosed a revised version of our paper submitted for peer review for the Journal of Service Research. Your comments have been very much appreciated and we have tried to suitably modify/amend our paper for those comments to be reflected in the revised version.

We quote below the detailed referees’ comments that are followed by our response. In all cases, we point out, where necessary, the corresponding amendments in the new version of the paper and highlighted them in blue.

# Editor comment

Manuscript ID JSR-22-468 entitled "Hierarchical Time Series Forecasting in Emergency Medical Services" which you submitted to the Journal of Service Research, has been reviewed. I read the manuscript and assigned it to an Associate Editor (AE). The AE selected the reviewers. Three expert reviewers responded with detailed and constructive comments. The AE wrote a summary letter with additional comments. I read all the reports and reread the manuscript. The reviewer and AE reports are located below.

With the exception of R3, the review team finds the research topic to be very important and sees the potential for your manuscript to have substantial impact. I also like the topic and see similar potential. The AE and I think the issues raised by R3 can be addressed in a revision.

R1, R2 and the AE raised many clarifying questions and had concrete suggestions for improve the work. I ask that you give serious consideration to these comments. The current MS is 33 pages, and we allow up to about 50 pages, and so you should have enough space to add the clarifications, possibly as an appendix.

I agree with many of R3's comments. JSR is not a methods journal, and the readers will generally be interested in substantive implications of the work. The evaluation currently focuses on several accuracy metrics such as MACE and MSSE, but what does improved forecasting accuracy mean to business metrics? Perhaps you could do some calculations to demonstrate the importance of your research in terms of business outcomes? Another way to address this is to give more detail about the natural of the decisions that are being made. How specifically can a planner “reconfigure units” and what sorts of options are available in “redeploying staff”? Help the reader understand the managerial decisions being made.

R3 also raises an issue about the generalizability. The AE has some suggestions for broadening the introduction slightly, and there could be a bit more discussion about other possible contexts in which your model might be used. R3 gives two cites that seem to address the problem; I am interested in whether these methods already solve your problem, or whether they would require substantial modification.

I liked your point about reproducibility very much. It is great that you will released you code on Github. Would it be possible to release your data as a benchmark data set for the research community? I understand that there may be NDAs involved, but publicly available data sets help the field advance.

This is minor, but there seem to be issues with your bibliography software. Sometimes it inserts initials in the text citations, other times not. Sometimes commas are missing, e.g., "Haugsbø Hermansen and Mengshoel (2021)". There are other writing errors, e.g., sometimes "neural network" is capitalized and other times not. It is not a proper noun.

# Associate Editor

_General comments to author: This manuscript looks at how to best forecast ambulance demand. These forecasts are crucial for planning and deploying staff and fleet. They are required at different levels (e.g., national, regional, and sub-regional levels) and are often made independently of each. So, the resulting forecasts at the different levels could be inconsistent. To avoid this problem, the authors exploit the hierarchical and grouped structure of the demand time series and apply several forecasting methods. In their empirical study with daily incident data from an ambulance service in Great Britain from October 2015 to July 2019, an ensemble of forecasting models yields the best forecasts that are also coherent at all levels._

_We received three very excellent reviews that came up with mixed recommendations. Still, they share your excitement about the topic. They are mostly consistent in recognizing the manuscript’s strengths and weaknesses. All three reviewers read the paper as a “forecasting exercise”. Instead, you need to better highlight the substantive insights that your “forecasting exercise” provides. In addition to the recommendations of the reviewers, my recommendation below might help you achieve this aim._

_All reviewers have questions and recommendations regarding the details of your time series. They seem to be justified and not in conflict with each other. So, consider implementing them._

_All reviewers have questions regarding your look “beyond the point estimate”. Consider them. It might help to also come up with an example that outlines the advantage of going beyond point estimates in greater detail._

_R1 has a long list of precise and helpful suggestions for improving the manuscript. It looks like they are all justified, and it should be manageable to implement them. The same holds for R2 and R3. R3’s remark regarding Figure 4 and the performance of the naïve model that seems to be not much worse is particularly crucial._

Response: Thank you for your feedback and for summarizing the main points from the reviewers. We appreciate your suggestion to highlight the insights provided by the forecasting exercise, and we have worked on the manuscript to make this more explicit in the revised manuscript. We agree that the reviewers have raised important questions and recommendations regarding the details of our study, and we have addressed these in the revised manuscript. Regarding the recommendation to look beyond the point estimate, we have taken the reviewers' suggestions into account and provide an example that illustrates the advantage of doing so. We have implemented R1's suggestions in the revised manuscript. We have also responded to R3's remark regarding the performance of the Naïve model and Figure 4.

## AE comments

In addition to the reviewers, I recommend also addressing my following suggestions:

_Comment 1. Consider broadening the scope of the manuscript by not focusing instantly on ambulance demand. Probably, other services also share the problem of having different levels. Outline them and explain why they all face the core problem that you address. R3 makes a comparable statement (“generalizability”). Such broadening might also enable you to better link your work to previous research in the Journal of Service Research._

  **Response:** Thank you for the suggestion! We agree that broadening the scope of the study to include other services that face similar hierarchical and grouped time series forecasting issues would make it more relevant and interesting to the Journal of Service Research audience. We have added a new subsection to the manuscript in the conclusion section, where we discuss other services that share the same problem and how our approach could be applied to those services. We also linked our work to previous research in the Journal of Service Research to provide a more comprehensive understanding of our contribution.

_Comment 2 (Discuss with Rob). It would also help to have a rather detailed example, maybe even a numerical example, to better describe why separate forecasts at different levels can cause problems. Right now, it sounds plausible that such problems occur, but you are still rather vague concerning the details of these problems._

  **Response:** Thank you for your comment. We agree that providing a detailed example can better illustrate the problems that arise when separate forecasts at different levels are inconsistent. We have added a simple numerical example to the manuscript, which outlines the issue of incoherent in hierarchical forecasts.

_Comment 3: You sometimes mentioned that the data runs from 2015-2019 (e.g., in the abstract) and sometimes from 2015-2020 (page 3). Please harmonize._

  **Response:** Thank you for bringing this to our attention. We have gone through the manuscript and made sure that we consistently report the duration of the dataset as 2015 to 2019, which is the correct period

_Comment 4: Consider summarizing previous research in a table. Such a table should also enable you to better highlight your contribution (by also including your work in the table)._

  **Response:**  Thank you. We agree that a table could provide a clearer overview of the related literature and our contribution. Therefore, we have now included a table in the literature review section summarizing previous research, including our study's contribution.

_Comment 5: Make sure that all tables and figures are self-contained._

  **Response:** Thank you for your comment. We have reviewed each table and figure in the manuscript and ensure that they are complete and understandable without relying on the main text and all necessary information such as axis labels, legends, etc are provided.

_Comment 6: Table 1: Make it easier for the reader to understand all values. For example, instead of “105”, write “105 = 35 x3”, etc._

  **Response:** Thank you for your comment.We have updated Table 1 to make it easier for readers to understand all values by providing a breakdown of the calculation.

_Comment 7 (Check with Rob): If possible, display the number of observations in all tables and figures with statistical results._

  **Response:** Thank you for your comment. We agree that displaying the number of observations is important when reporting statistical results. However, in the context of forecasting, what matters more is testing the accuracy of the forecast using new data (i.e., test set) by employing time series cross-validation. Statistical significance matters more when we are interested in fitting a model to the data. Therefore, we believe that it is not necessary to display the number of observations in all tables and figures with statistical results in this manuscript.

# Reviewer 1

_**General comment:** In this study, the authors assess the efficacy of various hierarchical time series forecasting techniques within the domain of emergency medical services. Specifically, they consider the hierarchical and group structure of ambulance demand. Utilizing daily incident data from an ambulance service in Great Britain spanning from October 2015 to July 2019, the authors conduct forecasting and performance evaluations. They employ an ensemble of forecasting models, in which the predictions of multiple models are merged to produce superior results compared to any single forecasting model. The paper is commendably organized, clearly written, and easy to comprehend. However, it lacks sufficient detail on the methodology, as noted in my comments below. I encourage the authors to include more in-depth information to enhance the overall quality of the paper._

**Response:** Thank you for taking the time to review our paper. We appreciate your positive comments and understand your concerns regarding the lack of detail on the methodology. We have addressed your comments in the following responses.

_**Major Comments:** The description of the approaches in section 3 and section 4 are missing important details that would enhance the readers’ understanding of the paper._

**Response:** Thank you for pointing out that our descriptions in sections 3 and 4 could benefit from more detail. We have revised those sections to include more specific information on the forecasting models and hierarchical and grouped time series techniques utilized in our study.

_**Comment 1**: The data points used in this time series forecasting pertain to "attended incidents”, which refer to situations where a physical resource has been deployed to respond to an incident. However, this definition only accounts for actual responses, and does not consider potential failed responses. I suspect that the response rate will not always be 100% due to various reasons, such as lack of resources, and it is important to take these failed responses into account in forecasting future emergency medical services (EMS) needs. Is data on failed responses available? If not, it would be beneficial to at least acknowledge and discuss this aspect of the forecast in the analysis._

**Response:** Thank you for your thoughtful comment regarding the potential impact of failed responses on our time series forecasting. We agree that failed responses could have an impact on forecasts, as the actual demand for ambulance services could be higher than what we have observed in the attended incidents data. Unfortunately, as you noted, the dataset we used only includes information on attended incidents, and we do not have access to information on failed responses. Therefore, we were not able to directly account for the potential impact of failed responses in our analysis. However, we have acknowledged this limitation in our manuscript and discussed its potential impact on our forecasting results. We agree that it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available. In the meantime, we hope that our study can provide some insights into the hierarchical time series forecasting techniques that can be used with attended incident data, and how these techniques can potentially improve the accuracy of EMS demand forecasting. It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses. However, the forecasting results may differ depending on the specific characteristics of the data.

_Comment 2. It is important to note the value of this research in the context of emergency medical services (EMS). Both high-level and low-level forecasts are necessary for effective decision-making in EMS. For example, control area EMS forecasts are high-level forecasts that inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals. These types of decisions are typically made periodically, not on a daily or weekly basis. In this context, the research presented in this paper has clear value. However, it is not clear to me what specific value the research brings at a lower level. It would be beneficial if this aspect could be further explained in the paper._

**Response:** Thank you for your insightful comment. We appreciate your point about the importance of both high-level and low-level forecasts for effective decision-making in EMS. In our study, we use a methodology that provides forecasts for all levels of a given hierarchy including high-level and low-level of the hierarchy. We believe that wherever decisions are made about the future in the hierarchy, forecasts would be used as an input. While you pointed out values on high-level forecasts, but forecasts are also valuable in at lower levels for shorter-term decision-making and planning. For example, hospitals or ambulance stations could use such forecasts to plan for staffing and resource allocation, ambulance dispatching, staff-to-shift assignment, staff rescheduling based on the anticipated volume and type of incidents. Additionally, generating forecasts at lower levels could potentially improve the accuracy of the high-level forecasts, by providing more detailed information on the nature and priority of incidents. This could help to identify patterns and trends in demand that may not be apparent at the higher level.

_Comment 3. Second paragraph of page 9 mentions “the strength of trend” and “the strength of seasonality”. Can you explain what those are or how they are calculated?_

**Response:** Thank you for the comment. In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, which is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components. The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component. The strength of trend is defined as:

$$F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$

For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component. Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small. But for data with little or no trend, the two variances should be approximately the same.

The strength of seasonality is defined similarly:

$$F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).$$

 series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).

_Comment 4. One of my main concerns is with regard to section 3.2. The current discussions of the forecasting approaches are inadequate in terms of important details. To improve the understanding of the paper, please provide a more detailed description of each of these approaches._

**Response:** Thank you for your feedback. We appreciate your concern regarding the level of detail provided in Section 3.2 on the forecasting approaches used in this study. While we agree that it is important to provide adequate information in the paper itself, we also understand that some readers may be interested in more technical details and code implementation. For this reason, we have provided a Github repository with all technical details and R codes for the forecasting approaches used in this study. However, we acknowledge that more information can be provided in the paper itself to enhance readers' understanding. To address your concern, we have revised Section 3.2 to provide a more detailed description of each of the forecasting approaches used, including the mathematical framework, assumptions, and key parameters. We hope that this revised section, along with the Github repository, will provide readers with the necessary information to understand and replicate the methods used in this study.

_Comment 4.1: In this research, the ”naive forecasting approach” is used as a benchmark. Is this the forecasting approach currently used in practice? If not, why was it selected as the benchmark? I expect that more rigorous forecasting approaches will outperform the naive approach. It would be more compelling to demonstrate the real-world value of your approach if it were compared against the approach currently used in practice._

**Response:** Thank you for the suggestion. We appreciate your feedback. We would like to clarify that our intention in this paper is not to provide an exhaustive list of forecasting methods. Instead, we aim to provide a methodology that can incorporate forecasts produced by any method. Our focus is on developing a hierarchical time series forecasting approach that takes into account the hierarchical and group structure of ambulance demand. We have selected the naive method as a benchmark because it is widely used in forecasting literature and practice and is a simple and easily understandable method for users.

Forecasts are often used as input for other decision-making systems that often involve simulation, and it is quite common to use the empirical distribution of the demand as a forecast in simulations. Also, the naive method can be surprisingly accurate. Therefore, if any forecasting method should be able to provide more accurate results than naive, it would justify its use in practice, if not there is no need to use more complicated methods.

_Comment 4.2: In your discussion of ETS, you mention that the trend, seasonality, and error terms can take various forms, such as additive, multiplicative, or mixed. Can you specify which form you used for each of these components in your research and explain the reasoning behind your choice?_

**Response:** Thank you for the question. In our research, we used an automatic algorithm to determine the appropriate form for the trend, seasonality, and error terms in each time series. Specifically, we used the ETS() function in the fable package of R, which selects the most appropriate model for each time series based on statistical criteria such as Akaike information criterion (AIC). Since we had a large number of time series (1530), it was not practical to manually select the appropriate form for each component for each time series. Instead, the automatic algorithm selects the most appropriate components based on the characteristics of each individual time series. As a result, a mix of different component forms (additive or multiplicative) were used, depending on the specific characteristics of each time series.

_Comment 4.3 (Rob): In your research, the Generalized Linear Model (GLM) approach is used for forecasting. Can you provide the equation used for this approach in a concise form? Additionally, in your description of the Fourier terms, it is mentioned that they are used to capture yearly seasonality. Does your research also take into account monthly seasonality and if so, how is it captured? Furthermore, for yearly seasonality, what is the specific form of the Fourier term used?_

**Response:**

_Comment 4.4: Can you explain what “serial dependence” means in the TSGLM model in the context of EMS forecasting?_

**Response:** Thank you for the comment. In the context of EMS forecasting, serial dependence refers to the situations where ambulance demand at one point in time are correlated with the demand at previous points in time. In other words, the past values of the time series influence the current values, and this influence can be modeled using a variety of techniques. The TSGLM model, for example, explicitly accounts for serial dependence by including lagged values of the ambulance demand in the model. This is important in EMS forecasting because it allows the model to capture patterns and trends in the data that are dependent on the history of the time series.

_Comment 4.5 (Rob): It is clear that the naive method performed poorly in the results. Can you explain the reasoning behind its inclusion in the ensemble method? Have you also evaluated the performance of an ensemble method that combines forecasts from only ETS, GLM and TSGLM, without the naive method? I would be interested in seeing the results from this alternative ensemble method._

**Response:**

_Comment 5 (Rob to check the response): In the first paragraph of section 3.3, you stated that the model development and hyper-parameter tuning process is conducted using the training data alone. Is this a commonly employed approach within the literature? From my understanding, when hyper-parameter tuning is involved, a dataset is typically divided into three sets: training, validation, and test sets. The training set is used to develop the model, the validation set is employed for tuning the parameters, and the test set is utilized to report the error measurement metric. Can you provide further clarification on this methodology and how it compares to the standard approach in the literature?_

**Response:** Thank you for your comment and question. We agree that traditionally, datasets are divided into three sets: training, validation, and test sets, and the validation set is used to tune the hyperparameters of the model. However, please note that we employ time series cross-validation to assess the forecast accuracy, and thus the use of a separate "validation dataset" becomes unnecessary when estimating the model parameters through cross-validation with the training data.Time series cross-validation involves dividing the data into multiple training and testing sets based on the time series structure. Each testing set consists of a number of observations equal to the forecast horizon, and the corresponding training set only includes observations that occurred before the test set. This ensures that future observations are not used in constructing the forecast, making it more robust to evaluate forecast accuracy on new data. Therefore, we believe that our methodology is appropriate for time series forecasting and provides a more robust approach for evaluating forecast accuracy on new data.

_Comment 6: In Section 3.3, it is not clear what the forecasting horizon j represents. The unit of time, whether it is a day, a week, or multiple weeks, is not specified and this makes it difficult to understand the logic behind the point forecast error e_j . Additionally, it would be helpful to explain why the scale-independent measures such as MASE and MSSE were chosen and how it relates to other commonly used accuracy measures._

**Response:** Thank you for your comment. To clarify, the forecasting horizon j represents the number of time periods ahead that we are forecasting, which is j= 1,2, ..., 84 days in our study. We will make sure to include this information in the revised version of the paper.

In terms of the choice of accuracy measures, we chose MASE and MSSE because they are scale-independent. Using scale-independent measures like MASE and MSSE allows for more meaningful comparisons between time series at different levels and scales, as they are not affected by the magnitude of the data. This is particularly important in our study, as we have time series at various levels of the EMS hierarchy, which may have different scales and therefore different magnitudes of error. By using scale-independent measures, we can compare the accuracy of forecasts across the entire hierarchy in a more meaningful way.

MASE stands for Mean Absolute Scaled Error, and is a normalized measure of the mean absolute error (MAE) of the forecast relative to the naïve forecast, which is a simple forecasting method that uses the last observed value as the forecast for the next time period. Similarly, MSSE stands for Mean Squared Scaled Error and is a normalized measure of the mean squared error (MSE) of the forecast relative to the MSE of the naïve forecast.

_Comment 7 (Rob to Check the response): In the last paragraph of page 13, the Continuous Ranked Probability Score (CRPS) is mentioned as a metric that rewards sharpness and penalizes miscalibration. Can you provide an explanation of what these terms mean in the context of your problem or dataset? Additionally, in the definition of p_j, the forecasted probability distribution and true probability distribution are referenced. Can you explain how these are derived from the dataset? If a specific distribution is assumed, it would be helpful to be specific about it._

**Response:** Thank you for your question. We have provided responses to each of the two separate questions below:

Sharpness refers to the degree to which the forecast distribution concentrates around the actual observations. In other words, a sharp forecast distribution means that the probability mass is concentrated around the actual outcome, which indicates high confidence in the forecast. On the other hand, a flat forecast distribution implies low confidence in the forecast. The CRPS rewards sharpness by assigning lower scores to forecasts with sharper distributions. Miscalibration, on the other hand, refers to the degree to which the forecast distribution is shifted from the actual distribution of the observations. In other words, a miscalibrated forecast means that the forecast distribution is not centered around the actual outcome, which indicates that the model is not capturing the underlying patterns in the data. The CRPS penalizes miscalibration by assigning higher scores to forecasts that are further away from the actual outcome. Therefore, the CRPS is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts. We chose the CRPS as one of our evaluation metrics because it provides a comprehensive assessment of the forecast quality and allows us to compare the performance of different models.

To generate forecast probability distribution, we did not assume a specific distribution, instead we used bootstrapping. This involves repeatedly resampling the training data with replacement and fitting a model to each sample to generate a distribution of forecasts. For each time point in the test set, we generated a set of 1000 bootstrapped samples from the corresponding training set, and fit models to each sample to generate a set of 1000 forecasts for the next time point. We then used these 1000 forecasts to construct the forecast distribution at each time point in the test set. It is important to note that the distribution of an actual outcome is always equal to 1 because it has already happened. The comparison between the forecast distribution and the true probability distribution is then used to calculate the CRPS.

_Comment 8: Section 4.2 does not sufficiently relate to the current problem at hand. The definitions presented are given in a standard form without providing context or explaining how they relate to the problem or dataset being analyzed. For example, the term ”bottom-level series” is used in both the text and Table 2, but it is not clear how this term applies to the specific dataset being studied. It would be beneficial to provide a concrete example from the dataset to help readers understand the concept of ”bottom-level series” in the context of this problem._

**Response:** Thank you for your comment. We agree that further explanation and contextualization of the definitions presented in Section 4.2 would be beneficial for readers. In particular, we will provide a concrete example from our dataset to illustrate the concept of "bottom-level series" and how it applies to our problem. We will also revise the text to provide more context and explanation for each definition, to better relate them to the problem at hand.

_Comment 9: I am experiencing confusion while reading section 4.2.2. The section discusses a ”bottom-up approach” and it sounds like it is not a favorable approach. However, in the following paragraph, you mention ”forecast reconciliation approaches.” Are these two different approaches or is the ”forecast reconciliation approach” a type of ”bottom-up approach”? Additionally, you also mention the use of ”linear reconciliation method” in this section, which leads me to assume it is a type of ”forecast reconciliation approach.” Can you provide further clarification on these different approaches and how they relate to each other in this context?_

**Response:** Thank you for your comment on section 4.2.2. I apologize for any confusion that may have arisen from the text. Please allow us to provide some clarification.

The bottom-up approach is indeed a specific type of forecast reconciliation approach, where forecasts are generated at the bottom level (most disaggregate level) and then aggregated up to higher levels in the hierarchy. However, this approach has limitations because it only uses base forecasts from a single level of aggregation at the bottom level. On the other hand, the linear reconciliation method is a more advanced type of forecast reconciliation approach that takes all base forecasts generated within a hierarchical structure to regenerate coherent forecasts at all level of the hierarchy. This means that it does not only use base forecasts from a single level of aggregation, but rather uses all available information at all levels to generate forecasts that minimises the total forecast variance of the set of coherent forecasts. I hope this clarifies the difference between the two approaches and how they relate to each other in the context of the paper.

_Comment 10: In section 5, you present the results of your research. From my understanding, the ”forecast reconciliation approach” aggregates lower-level forecasts. If this understanding is correct, I am confused as to why you are comparing base forecasts to the aggregation of those base-level forecasts as they appear to be fundamentally different things. Also, in the second paragraph of page 20, you mention that the reconciliation may not improve the forecast accuracy if the bottom series are ”noisy” and lack systematic patterns. My question is: in such cases, what is the purpose or benefit of performing reconciliation at the bottom level?_

**Response:**

The forecast reconciliation approach combines base forecasts (not Lower level forecasts). It is important to note that one of the main purposes of comparing base forecasts to the reconciliation of those forecasts is to assess the improvement in forecast accuracy that can be achieved by using the forecast reconciliation approach. The base forecasts are generated independently at each level of the hierarchy, ignoring the hierarchical/grouped structure of time series or any aggregation constraints. As a result, the base forecasts are not guaranteed to be coherent, meaning that the sum of the lower-level forecasts will not necessarily add up to the higher-level forecast. On the other hand, the forecast reconciliation approach takes the base forecasts and regenerates them in a way that ensures coherence across all levels of the hierarchy, while also considering all available information. The comparison between the base forecasts and the reconciled forecasts allows us to see the improvement in forecast accuracy that can be achieved by using the forecast reconciliation approach.

While the approach is designed to improve forecast accuracy, it may not always be successful in doing so, particularly if the bottom-level series are too noisy and lack systematic patterns. In such cases, the benefit of performing reconciliation at the bottom level may still lie in creating coherent forecasts that can help align planning across different teams in an organization, improve coordination, and avoid conflicting decisions. Additionally, even if the bottom-level series are noisy and lack systematic patterns, reconciliation can still lead to more accurate forecasts at higher levels of the hierarchy by utilizing the information available in the upper-level series. Therefore, even if the bottom-level forecasts are not very accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and possibly more accurate forecasts at other levels.

_Comment 11 (Rob): In the second to last sentence of the paper, you state that ”the reconciled forecast distributions are not counts.” In your evaluation of forecast performance, did you round any of these non-integer values to whole numbers or were they kept as non-integers? For practical implementation, it is necessary to round these non-integers to whole numbers. Have you considered how your forecasting approaches perform if the non-integer forecast results are rounded up or rounded down?_

**Response:** Thank you for your comment. In our study, we kept the non-integer forecast values after reconciliation to evaluate the accuracy. However, we recognize the importance of considering practical implementation. We agree that this is an important consideration and will incorporate it into our future research as it requires further investigations. Additionally, we acknowledge that forecast reconciliation of count forecast is an active area for research and and it would be interesting to explore how our approach could be adapted to generate count reconciled probabilistic forecasts.

Rounding forecasts up or down and its effect on the forecast accuracy depends on the level of hierarchy and scale of data. For some situations, with high volume demand, forecast accuracy calculations can ignore integer effects as rounding becomes negligible. However, low volume demand settings, such as forecast at bottom level of hierarchy,  may be more susceptible to integer (rounding) effects.

_Comment 12: In Table 1, the number of parameters for “Priority * Nature of incident” is 104. Why is it not 105? Similar questions for the next two rows._

**Response:** Thank you for bringing this up. The reason for the discrepancy in the number of series is that for some combinations of priority and nature of incidents, there is no representation in the dataset. Therefore, these combinations are not included in the table For example, in the "Priority * Nature of incident" combination, one of the cells in the cross-tabulation have no incidents recorded for a combination of Red Green priority and Falls, so it is not included in the dataset. The same reasoning applies to the other two combinations in Table 1.

_Comment 13: Please fix the headings in Table 2 for “Control areas” and “Health boards”. Currently they are misaligned. I hope the authors will find those comments helpful in improving the paper._

**Response:** Thank you for bringing this to our attention. We have now fixed the headings in Table 2. 
We appreciate your feedback and we hope that our revised version will meet your expectations.

# Reviewer 2

_General comment: This paper proposes hierarchical algorithms to forecast the Emergency Medical Services (EMS) demands with both point estimates and distributional estimates. The EMS count data are naturally structured with hierarchies (with spatial levels corresponding to nation, sub-nation and region) and groups (including the priority and incident of the demand).  The author starts with independent/base forecast with some classical benchmark time series algorithms such as moving average, GLM and Poisson regression. The hierarchical forecasts are then introduced with the bottom-up structure and optimal reconciliation approaches. The outperformance of the algorithm is demonstrated by comparing the accuracy between combinations of methods and models under different spatial levels and groups. I appreciate the detailed background introduction and literature review for the application topic. The novel point of view that introduces the hierarchical and grouped forecasts to the EMS management community is promising. I provide some main suggestions, questions and minor comments listed below for reference._

**Response:** Thank you for your detailed review and positive feedback on our paper. We appreciate your valuable suggestions, questions, and comments, and we will address them accordingly.

_Comment 1 (Check with rOB): Section 3.2 of Forecasting methods covers the independent/base forecasting methods, and Section 3.3 of Performance evaluation is designed for both base and grouped forecasts. I would suggest to look further into the logic structure of Sections 3 and 4. One advice may be to push back the overview of base forecast algorithms (Section 3.2) to a new section before the main contribution of grouped forecasts in Section 4._

**Response:** Thank you for your comment. We appreciate your suggestion to improve the structure of the paper. We will consider moving the overview of base forecast algorithms to a separate section before Section 4 to provide a clearer delineation between the independent and hierarchical/grouped forecasts. This will help readers better understand the logic structure of the paper and facilitate easier navigation of the content.

_Comment 2: In Section 3.2, the author builds the base forecasting models._
_Comment 2-1: There should be more detail added to ETS and Ensemble method. What exactly do they do? Ensemble method refers to Wang et al., but no specific reference. Adding explanatory details to the used methods and a short discussion on other methods should be sufficient._

**Response:** Thank you for your helpful comment. We agree that providing more detail on the ETS and Ensemble methods, as well as other methods, would enhance the understanding of our proposed approach. In the revised version, we have added more detailed explanations on these methods, including their mathematical formulations, assumptions, and limitations. We will also provide additional references to the Ensemble method and further explanations on how it is applied in our study.

_Comment 2-2: In ETS, there are trend, seasonality and a property about the error terms. Have all combination options been considered? If not, how has it been evaluated?_

**Response:** Thank you for your comment. Since we had a large number of time series (1530), we did not manually evaluate all possible combinations of trend, seasonality, and error terms in ETS for each time series, instead we used an automatic algorithm to select the most appropriate model for each time series based on statistical criteria such as Akaike information criterion (AIC). The automatic algorithm considered a mix of different component forms, depending on the specific characteristics of each time series. We have added additional details on the automatic algorithm used in the revised version of the paper.

_Comment 2-3: GLM and TSGLM are regression approaches, sometimes with exogenous variables like holiday indicators. I would suggest to include explicit model formulas (probably R-formatted like y ~ a + b + c) to list all the input predictors._

**Response:**

_Comment 3: Section 3.3 stakes different measures for performance evaluation. I suggest to give more detailed references opposed to just the paper, i.e., equation numbers._

**Response:**

_Comment 4: The author extracts the trend and weekly seasonality of the data, and demonstrates their strengths in Figure 2. The follow-up analysis and forecast depend on the Poisson assumption most of the time due to the nature of non-negative count time series data._

**Response:**

_Comment 5: Is there any verification conducted to confirm the Poisson patterns in the data? Are over-dispersion and/or under-dispersion going to be potential issues? It may suffice to include statements from relative references._

**Response:**

_Comment 6: ETS seems to have the disadvantage of not producing integer valued forecasts. There are certainly other ways to do forecasting for integer valued time series. What about INAR or INGARCH? On the other hand, tsglm() seems to allow for those options but it is not clear what has been used._

**Response:**

_Comment 7: How to understand the stationarity of the time series? Is it going to impact the analysis if the errors are assumed additive and the heteroscedasticity is present?_

**Response:**

_Comment 8: And if the errors are assumed multiplicative in a Poisson sense, what are the possible adjustments to verify or handle the stationarity before forecasting?_

**Response:**

_Comment 9: The author introduces the algorithm for estimating the distributional forecasts of the data, and measures the accuracy using CRPS._

**Response:**

_– With the distributional forecasts, how do the confidence intervals behave?_

**Response:**

_– As mentioned in P20 L14, the algorithm gives more accurate forecast results concerning the tails of the distribution. How does such outperformance visualized regarding this statement, and what to be read from the forecast distributions for risk management?_

**Response:**

_Comment 10: The author briefly explains the model setup and estimation routines. The conciseness of the paper is greatly appreciated. In the meantime, an appendix chapter or a separate file of supplementary material might be expected for brief elaboration on some ideas and technical derivations. For example:_

**Response:**

_-how are the strengths of trend and weekly seasonality estimated and scaled to [0, 1]?_

**Response:**

_– for the aggregated time series a_t = A b_t, how is the ‘aggregation’ matrix A determined?_

**Response:**

_– for the linear reconciliation method  $\tilde{y}_h = S(S′W^{−1}S)^{−1}W^{−1} \hat{y}_h,$ how is it derived?_

**Response:**

_– for the positive definite matrix W, what are the intuitions and strengths for the different solutions (Ordinary Least Squares (OLS), Weighted Least Squares (WLS) and Minimum Trace (MinT))? And how is W defined under these options respectively?_

**Response:**

_Comment 11: Minor comments_
_Note that line count includes equations, and excludes (sub)section titles. Minus sign indicates counting from bottom._

**Response:**

_- P13, L1 & L3, it might be good to not indent after the equation if the paragraph does not end. The same concern applies to the other equations in the paper._

**Response:**

_- P14, L3, for the Continuous Rank Probability Score, how is the true probability distribution function F_j (x) determined?_

**Response:**

_- P17, L-6, “it is does” → “it does”._

**Response:**

_- P18, L8, “Mint” → “MinT”._

**Response:**

_In References, try to maintain a consistent format in terms of the inclusion (e.g., P22, L10 & L-4 & L-1) or exclusion of the url links._

**Response:**

_P26, Figure 1(a) overlaps with Figure 1(b) and turns to be cropped at the right side._

**Response:**

_In Table 2, the top row (of column names) is unclear with the line change (“areas” and “boards”)._

**Response:**

# Reviewer 3

_General comments: This paper looks at improving forecasts in emergency medical services by exploiting the hierarchical and grouped nature of the different timeseries that related to different levels of decision making. The paper is generally well written and the dataset is interesting. The abstract and introduction also have several elements that triggered my interest as a reader. Unfortunately, the rest of the paper did not live up to some of my expectations, which I will detail below._

**Response:**

_Comment 1: Method_
_Comment 1-a. The middle and backend of the paper feel more like a forecasting exercise. In many ways this paper reads more as a paper for a forecasting journal then a paper for JSR. Also, in this forecasting exercise many relevant choices, such as how the trend and seasonality enter the model happen in a black box for the reader, whereas the impact of these choices can be substantial, I think. For example, for a bottom level time series an additive seasonality component might be beneficial, but in its role for reconciling the higher level series it might be that a multiplicative term would have been better._

**Response:**

_Comment 1-b. Related to point 1.a I was also wondering if an integrated estimation method would not be beneficial. This might go at the expense of reproducibility, as those models might rely less on built-in functions of R. I want to be careful in making concrete suggestions, but my first intuition would be to consider a Dynamic Hierarchical Linear Model (e.g., Gamerman and Migon 1993; Neelamegham and Chintagunta 2004). The transfer functions in these models could be based on the “summing” or “structural” matrices the authors mention in section 4.2.1. The authors could still keep the two-step approach as alternative to show how it differs from an integrated approach. If differences are small, the latter could still serve as a reproducibility/managerial tool._

**Response:**

_Comment 2: Implications_
_Comment 2-a: The implications of the paper are not clear to me. The authors mention that the performance is better for the ensemble method, but they do not show what the impact of this is. Looking at Figure 4 it does not seem like the ensemble method is performing a lot better than the other non-naïve methods. The paper would benefit from some economic interpretation of the better performances. For example, how much more efficient can allocation of medical services be? This comes back to my point 1.a that currently it feels more like a forecasting exercise, but misses the actual impact on the medical services._

**Response:**

_Comment 2-b: Throughout the paper the authors mention that one of the strengths of their model is that they look beyond point estimates. What I understand is that they use the CRPS statistic for this. However, this is again very mechanic. It would be much more valuable to show when and how actual observations are within the prediction interval, as I can imagine that having both too wide or too narrow confidence intervals could lead to wrong allocation decisions._

**Response:**

_Comment 3: Generalizability_
_Comment 3-a: Another concern I have is the generalizability, especially for the service industry. The grouped and hierarchical nature seems very specific for the type of service of medical emergencies. And although the service is highly relevant and important, I am wondering whether the authors can’t broaden the scope somewhat in the introduction and/or implications. If not I would specifically address this in the limitations._

**Response:**

I wish the authors all the best with their research further

Gamerman, D., & Migon, H. S. (1993). Dynamic hierarchical models. Journal of the Royal Statistical Society: Series B (Methodological), 55(3), 629-642.

Neelamegham, R., & Chintagunta, P. K. (2004). Modeling and forecasting the sales of technology products. Quantitative Marketing and Economics, 2(3), 195.

### 2023-05-10 journal_correspondence/cover_letter.qmd
response letter - reposne to AE and reviewer 1

--- 2022-11-21 letter/cover_letter.qmd
@@
 address:
-  - Prof. Donald M. Yealy
+  - Prof. Ming-Hui Huang
   - Editor-in-Chief
-  - Annals of Emergency Medicine
+  -  Journal of Service Research
 subject: Paper submission for peer review
-opening: "Dear Prof. Yealy,"
+opening: "Dear Prof. Huang,"
 closing: "Yours sincerely"
@@
 
-We would be grateful if you could consider for publication our manuscript entitled _Hierarchical Time Series Forecasting in Emergency Medical Services_ as an Original Research article in the Annals of Emergency Medicine.
+We would be grateful if you could consider for publication our manuscript entitled _Hierarchical Time Series Forecasting in Emergency Medical Services_ as an Original Research article in the Journal of Service Research.
 

### 2023-05-10 manuscript/main.qmd
Some final changes
MS word document generated for submission, plus typos corrected
update the Github link
response letter - reposne to AE and reviewer 1

@@
 We use an ensemble of forecasting models and show that the resulting forecasts are better than any individual forecasting model.
-We validate the forecasting approach using time-series cross-validation.
+We validate the forecasting approach using time series cross validation.
 keywords:
@@
     journal:
-      name: Annals of Emergency Medicine
+      name: Journal of Service Research
       formatting: preprint
@@
 
+# Reproducibility {.unnumbered}
+
+R code to produce all results in this paper is available at
+
 # References {#references .unnumbered}

### 2023-05-15 journal_correspondence/response_letter/response_letter_JSR.qmd
Updated responses

@@
 title: "Response letter"
-format: pdf
+format:
+  pdf:
+    pdf-engine: pdflatex
+    papersize: a4
+    geometry: "margin = 2.5cm"
+    include-in-header: header.tex
 ---
 
-Dear Sir or Madam,
+Dear Professor Malthouse
 
-Please find enclosed a revised version of our paper submitted for peer review for the Journal of Service Research.
+Please find enclosed a revised version of our paper submitted for peer review for the *Journal of Service Research*.
 Your comments have been very much appreciated and we have tried to suitably modify/amend our paper for those comments to be reflected in the revised version.
@@
 
-Manuscript ID JSR-22-468 entitled "Hierarchical Time Series Forecasting in Emergency Medical Services" which you submitted to the Journal of Service Research, has been reviewed.
+*Manuscript ID JSR-22-468 entitled "Hierarchical Time Series Forecasting in Emergency Medical Services" which you submitted to the Journal of Service Research, has been reviewed.
 I read the manuscript and assigned it to an Associate Editor (AE).
@@
 I read all the reports and reread the manuscript.
-The reviewer and AE reports are located below.
+The reviewer and AE reports are located below.*
 
-With the exception of R3, the review team finds the research topic to be very important and sees the potential for your manuscript to have substantial impact.
+*With the exception of R3, the review team finds the research topic to be very important and sees the potential for your manuscript to have substantial impact.
 I also like the topic and see similar potential.
-The AE and I think the issues raised by R3 can be addressed in a revision.
+The AE and I think the issues raised by R3 can be addressed in a revision.*
 
-R1, R2 and the AE raised many clarifying questions and had concrete suggestions for improve the work.
+*R1, R2 and the AE raised many clarifying questions and had concrete suggestions for improve the work.
 I ask that you give serious consideration to these comments.
-The current MS is 33 pages, and we allow up to about 50 pages, and so you should have enough space to add the clarifications, possibly as an appendix.
+The current MS is 33 pages, and we allow up to about 50 pages, and so you should have enough space to add the clarifications, possibly as an appendix.*
 
-I agree with many of R3's comments.
+*I agree with many of R3's comments.
 JSR is not a methods journal, and the readers will generally be interested in substantive implications of the work.
@@
 How specifically can a planner “reconfigure units” and what sorts of options are available in “redeploying staff”?
-Help the reader understand the managerial decisions being made.
+Help the reader understand the managerial decisions being made.*
 
-R3 also raises an issue about the generalizability.
+*R3 also raises an issue about the generalizability.
 The AE has some suggestions for broadening the introduction slightly, and there could be a bit more discussion about other possible contexts in which your model might be used.
-R3 gives two cites that seem to address the problem; I am interested in whether these methods already solve your problem, or whether they would require substantial modification.
+R3 gives two cites that seem to address the problem; I am interested in whether these methods already solve your problem, or whether they would require substantial modification.*
 
-I liked your point about reproducibility very much.
+*I liked your point about reproducibility very much.
 It is great that you will released you code on Github.
 Would it be possible to release your data as a benchmark data set for the research community?
-I understand that there may be NDAs involved, but publicly available data sets help the field advance.
+I understand that there may be NDAs involved, but publicly available data sets help the field advance.*
 
-This is minor, but there seem to be issues with your bibliography software.
+*This is minor, but there seem to be issues with your bibliography software.
 Sometimes it inserts initials in the text citations, other times not.
@@
 There are other writing errors, e.g., sometimes "neural network" is capitalized and other times not.
-It is not a proper noun.
+It is not a proper noun.*
 
@@
 
-Response: Thank you for your feedback and for summarizing the main points from the reviewers.
+**Response:** Thank you for your feedback and for summarizing the main points from the reviewers.
 We appreciate your suggestion to highlight the insights provided by the forecasting exercise, and we have worked on the manuscript to make this more explicit in the revised manuscript.
@@
 
-_Comment 2 (Discuss with Rob).
+_Comment 2.
 It would also help to have a rather detailed example, maybe even a numerical example, to better describe why separate forecasts at different levels can cause problems.
@@
 
-  **Response:** Thank you for your comment.
-We agree that providing a detailed example can better illustrate the problems that arise when separate forecasts at different levels are inconsistent.
-We have added a simple numerical example to the manuscript, which outlines the issue of incoherent in hierarchical forecasts.
+  **Response:** We agree that providing a detailed example can better illustrate the problems that arise when separate forecasts at different levels are inconsistent.
+We have added the following example to the manuscript, which outlines the issue of incoherent in hierarchical forecasts.
+
+  > For example, the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
+If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area.
+Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others.
+Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly.
+Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).
 
@@
 
-  **Response:** Thank you for bringing this to our attention.
-We have gone through the manuscript and made sure that we consistently report the duration of the dataset as 2015 to 2019, which is the correct period
+  **Response:** Thank you for bringing this to our attention. We have gone through the manuscript and made sure that we consistently and correctly report the duration of the dataset as 2015 to 2019.
 
@@
 
-  **Response:** Thank you for your comment.
-We have reviewed each table and figure in the manuscript and ensure that they are complete and understandable without relying on the main text and all necessary information such as axis labels, legends, etc are provided.
+  **Response:** We have reviewed each table and figure in the manuscript and ensure that they are complete and understandable without relying on the main text and all necessary information such as axis labels, legends, etc are provided.
 
@@
 
-  **Response:** Thank you for your comment.We have updated Table 1 to make it easier for readers to understand all values by providing a breakdown of the calculation.
+  **Response:** We have updated Table 1 to make it easier for readers to understand all values by providing a breakdown of the calculation.
 
-_Comment 7 (Check with Rob): If possible, display the number of observations in all tables and figures with statistical results._
+_Comment 7: If possible, display the number of observations in all tables and figures with statistical results._
 
-  **Response:** Thank you for your comment.
-We agree that displaying the number of observations is important when reporting statistical results.
-However, in the context of forecasting, what matters more is testing the accuracy of the forecast using new data (i.e., test set) by employing time series cross-validation.
-Statistical significance matters more when we are interested in fitting a model to the data.
-Therefore, we believe that it is not necessary to display the number of observations in all tables and figures with statistical results in this manuscript.
+  **Response:** We agree that displaying the number of observations is important when reporting statistical results. We have added information about the size of the data set where relevant.
 
@@
 The strength of trend is defined as:
-
 $$F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$
-
 For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
@@
 The strength of seasonality is defined similarly:
-
 $$F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).$$
-
  series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).
@@
 In our research, we used an automatic algorithm to determine the appropriate form for the trend, seasonality, and error terms in each time series.
-Specifically, we used the ETS() function in the fable package of R, which selects the most appropriate model for each time series based on statistical criteria such as Akaike information criterion (AIC).
+Specifically, we used the `ETS()` function in the fable package of R, which selects the most appropriate model for each time series based on statistical criteria such as Akaike's Information Criterion (AIC).
 Since we had a large number of time series (1530), it was not practical to manually select the appropriate form for each component for each time series.
@@
 
-_Comment 4.3 (Rob): In your research, the Generalized Linear Model (GLM) approach is used for forecasting.
+_Comment 4.3: In your research, the Generalized Linear Model (GLM) approach is used for forecasting.
 Can you provide the equation used for this approach in a concise form?
@@
 
-**Response:**
+**Response:** Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
+\end{align*}
+and $\bm{x}_t$ is a vector of covariates.
+In our model, these include cubic splines for the time trend, day-of-week dummy variables, Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays, school holidays, Christmas Day and New Year's Day.
+The Fourier terms are as defined in Hyndman & Athanasopoulos (2021, Section 7.4).
+
+This model takes account of weekly seasonality and annual seasonality.
+Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
+There is no reason, for example, for incidents to occur more at some times of the month than others.
 
@@
 
-**Response:** Thank you for the comment.
-In the context of EMS forecasting, serial dependence refers to the situations where ambulance demand at one point in time are correlated with the demand at previous points in time.
+**Response:** In the context of EMS forecasting, serial dependence refers to the situations where ambulance demand at one point in time are correlated with the demand at previous points in time.
 In other words, the past values of the time series influence the current values, and this influence can be modeled using a variety of techniques.
 The TSGLM model, for example, explicitly accounts for serial dependence by including lagged values of the ambulance demand in the model.
-This is important in EMS forecasting because it allows the model to capture patterns and trends in the data that are dependent on the history of the time series.
+This is important in EMS forecasting because it allows the model to capture patterns and trends in the data that are dependent on the history of the time series, and which are not captured via the predictor variables.
 
-_Comment 4.5 (Rob): It is clear that the naive method performed poorly in the results.
+_Comment 4.5: It is clear that the naive method performed poorly in the results.
 Can you explain the reasoning behind its inclusion in the ensemble method?
@@
 
-**Response:**
+**Response:** We re-ran the results of our experience with the naive method excluded and .....
 
-_Comment 5 (Rob to check the response): In the first paragraph of section 3.3, you stated that the model development and hyper-parameter tuning process is conducted using the training data alone.
+_Comment 5: In the first paragraph of section 3.3, you stated that the model development and hyper-parameter tuning process is conducted using the training data alone.
 Is this a commonly employed approach within the literature?
@@
 
-**Response:** Thank you for your comment and question.
-We agree that traditionally, datasets are divided into three sets: training, validation, and test sets, and the validation set is used to tune the hyperparameters of the model.
-However, please note that we employ time series cross-validation to assess the forecast accuracy, and thus the use of a separate "validation dataset" becomes unnecessary when estimating the model parameters through cross-validation with the training data.Time series cross-validation involves dividing the data into multiple training and testing sets based on the time series structure.
+**Response:** In many machine learning applications, datasets are divided into three sets: training, validation, and test sets, and the validation set is used to tune the hyperparameters of the model.
+But we use a more efficient approach than this, whereby we employ time series cross-validation (Hyndman & Athanasopoulos, 2021, Section 5.10) to assess the forecast accuracy, and thus the use of a separate "validation dataset" becomes unnecessary.
+Time series cross-validation involves dividing the data into multiple training and testing sets based on the time series structure.
 Each testing set consists of a number of observations equal to the forecast horizon, and the corresponding training set only includes observations that occurred before the test set.
@@
 
-**Response:** Thank you for your comment.
-To clarify, the forecasting horizon j represents the number of time periods ahead that we are forecasting, which is j= 1,2, ..., 84 days in our study.
+**Response:** To clarify, the forecasting horizon j represents the number of time periods ahead that we are forecasting, which is $j= 1,2,\dots, 84$ days in our study.
 We will make sure to include this information in the revised version of the paper.
@@
 
-_Comment 7 (Rob to Check the response): In the last paragraph of page 13, the Continuous Ranked Probability Score (CRPS) is mentioned as a metric that rewards sharpness and penalizes miscalibration.
+_Comment 7: In the last paragraph of page 13, the Continuous Ranked Probability Score (CRPS) is mentioned as a metric that rewards sharpness and penalizes miscalibration.
 Can you provide an explanation of what these terms mean in the context of your problem or dataset?
@@
 
-Sharpness refers to the degree to which the forecast distribution concentrates around the actual observations.
-In other words, a sharp forecast distribution means that the probability mass is concentrated around the actual outcome, which indicates high confidence in the forecast.
-On the other hand, a flat forecast distribution implies low confidence in the forecast.
-The CRPS rewards sharpness by assigning lower scores to forecasts with sharper distributions.
-Miscalibration, on the other hand, refers to the degree to which the forecast distribution is shifted from the actual distribution of the observations.
-In other words, a miscalibrated forecast means that the forecast distribution is not centered around the actual outcome, which indicates that the model is not capturing the underlying patterns in the data.
-The CRPS penalizes miscalibration by assigning higher scores to forecasts that are further away from the actual outcome.
-Therefore, the CRPS is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.
+Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+It measures how well the predicted probabilities match the actual probabilities.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
+A model is well-calibrated if the predicted probabilities match the actual probabilities, and it is sharp if it is confident in its predictions.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.
 We chose the CRPS as one of our evaluation metrics because it provides a comprehensive assessment of the forecast quality and allows us to compare the performance of different models.
 
-To generate forecast probability distribution, we did not assume a specific distribution, instead we used bootstrapping.
-This involves repeatedly resampling the training data with replacement and fitting a model to each sample to generate a distribution of forecasts.
-For each time point in the test set, we generated a set of 1000 bootstrapped samples from the corresponding training set, and fit models to each sample to generate a set of 1000 forecasts for the next time point.
-We then used these 1000 forecasts to construct the forecast distribution at each time point in the test set.
-It is important to note that the distribution of an actual outcome is always equal to 1 because it has already happened.
-The comparison between the forecast distribution and the true probability distribution is then used to calculate the CRPS.
+To generate forecast probability distributions, we did not assume a specific distribution; instead we used a form of bootstrapping, described in Panagiotelis et al (EJOR, 2023).
+This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+In this way, we can generate an empirical distribution of forecasts for each model.
+The ensemble forecast distribution is a simple mixture of these empirical distributions.
+
+The comparison between the forecast distribution and the observations is then used to calculate the CRPS.
+In this context, the "true probability distribution" has all its weight on the observed value, and zero elsewhere.
+
+We have now updated the explanation of CRPS along the lines described above.
 
@@
 
-**Response:** Thank you for your comment.
-We agree that further explanation and contextualization of the definitions presented in Section 4.2 would be beneficial for readers.
+**Response:** We agree that further explanation and contextualization of the definitions presented in Section 4.2 would be beneficial for readers.
 In particular, we will provide a concrete example from our dataset to illustrate the concept of "bottom-level series" and how it applies to our problem.
@@
 
-**Response:** Thank you for your comment on section 4.2.2. I apologize for any confusion that may have arisen from the text. Please allow us to provide some clarification.
+**Response:** Thank you for your comment on section 4.2.2. We apologize for any confusion that may have arisen from the text. Please allow us to provide some clarification.
 
-The bottom-up approach is indeed a specific type of forecast reconciliation approach, where forecasts are generated at the bottom level (most disaggregate level) and then aggregated up to higher levels in the hierarchy.
-However, this approach has limitations because it only uses base forecasts from a single level of aggregation at the bottom level.
-On the other hand, the linear reconciliation method is a more advanced type of forecast reconciliation approach that takes all base forecasts generated within a hierarchical structure to regenerate coherent forecasts at all level of the hierarchy.
+While the bottom-up approach does lead to coherent forecasts, it is not a forecast reconcilation approach because no reconciling is done.
+In the bottom-up approach, forecasts are generated at the bottom (most disaggregate) level and then aggregated up to higher levels in the hierarchy.
+This approach has limitations because it only uses base forecasts from a single level of aggregation at the bottom level.
+On the other hand, the linear reconciliation method uses all base forecasts generated within a hierarchical structure to regenerate coherent forecasts at all level of the hierarchy.
 This means that it does not only use base forecasts from a single level of aggregation, but rather uses all available information at all levels to generate forecasts that minimises the total forecast variance of the set of coherent forecasts.
-I hope this clarifies the difference between the two approaches and how they relate to each other in the context of the paper.
+We hope this clarifies the difference between the two approaches and how they relate to each other in the context of the paper.
 
@@
 
-The forecast reconciliation approach combines base forecasts (not Lower level forecasts).
+The forecast reconciliation approach combines base forecasts (not only lower level forecasts).
 It is important to note that one of the main purposes of comparing base forecasts to the reconciliation of those forecasts is to assess the improvement in forecast accuracy that can be achieved by using the forecast reconciliation approach.
@@
 
-_Comment 11 (Rob): In the second to last sentence of the paper, you state that ”the reconciled forecast distributions are not counts.” In your evaluation of forecast performance, did you round any of these non-integer values to whole numbers or were they kept as non-integers?
+_Comment 11: In the second to last sentence of the paper, you state that ”the reconciled forecast distributions are not counts.” In your evaluation of forecast performance, did you round any of these non-integer values to whole numbers or were they kept as non-integers?
 For practical implementation, it is necessary to round these non-integers to whole numbers.
@@
 
-**Response:** Thank you for your comment.
-In our study, we kept the non-integer forecast values after reconciliation to evaluate the accuracy.
-However, we recognize the importance of considering practical implementation.
-We agree that this is an important consideration and will incorporate it into our future research as it requires further investigations.
-Additionally, we acknowledge that forecast reconciliation of count forecast is an active area for research and and it would be interesting to explore how our approach could be adapted to generate count reconciled probabilistic forecasts.
-
-Rounding forecasts up or down and its effect on the forecast accuracy depends on the level of hierarchy and scale of data.
+**Response:** Rounding forecasts up or down and its effect on the forecast accuracy depends on the level of hierarchy and scale of data.
 For some situations, with high volume demand, forecast accuracy calculations can ignore integer effects as rounding becomes negligible.
 However, low volume demand settings, such as forecast at bottom level of hierarchy,  may be more susceptible to integer (rounding) effects.
+In our study, we kept the non-integer forecast values after reconciliation when evaluating the accuracy.
+The difference in forecast accuracy results is negiglible.
+[Perhaps we should check this assertion?]  In practice, we may need to use integer forecasts, especially when the forecasts are relatively small.
+Count forecast reconciliation is an active area of research (e.g., https://arxiv.org/abs/2207.09322), and it would be interesting to explore in future research how our approach could be adapted to generate count reconciled probabilistic forecasts.
 
@@
 
-**Response:** Thank you for bringing this up.
-The reason for the discrepancy in the number of series is that for some combinations of priority and nature of incidents, there is no representation in the dataset.
+**Response:** The reason for the discrepancy in the number of series is that for some combinations of priority and nature of incidents, there is no representation in the dataset.
 Therefore, these combinations are not included in the table For example, in the "Priority * Nature of incident" combination, one of the cells in the cross-tabulation have no incidents recorded for a combination of Red Green priority and Falls, so it is not included in the dataset.
@@
 
-**Response:** Thank you for bringing this to our attention. We have now fixed the headings in Table 2. 
+**Response:** We have now fixed the headings in Table 2.
 We appreciate your feedback and we hope that our revised version will meet your expectations.
@@
 
-_Comment 1 (Check with rOB): Section 3.2 of Forecasting methods covers the independent/base forecasting methods, and Section 3.3 of Performance evaluation is designed for both base and grouped forecasts.
+_Comment 1: Section 3.2 of Forecasting methods covers the independent/base forecasting methods, and Section 3.3 of Performance evaluation is designed for both base and grouped forecasts.
 I would suggest to look further into the logic structure of Sections 3 and 4.
@@
 
-**Response:** Thank you for your comment.
-We appreciate your suggestion to improve the structure of the paper.
+**Response:** We appreciate your suggestion to improve the structure of the paper.
 We will consider moving the overview of base forecast algorithms to a separate section before Section 4 to provide a clearer delineation between the independent and hierarchical/grouped forecasts.
@@
 
-**Response:** Thank you for your helpful comment.
-We agree that providing more detail on the ETS and Ensemble methods, as well as other methods, would enhance the understanding of our proposed approach.
+**Response:**  We agree that providing more detail on the ETS and Ensemble methods, as well as other methods, would enhance the understanding of our proposed approach.
 In the revised version, we have added more detailed explanations on these methods, including their mathematical formulations, assumptions, and limitations.
@@
 
-**Response:** Thank you for your comment.
-Since we had a large number of time series (1530), we did not manually evaluate all possible combinations of trend, seasonality, and error terms in ETS for each time series, instead we used an automatic algorithm to select the most appropriate model for each time series based on statistical criteria such as Akaike information criterion (AIC).
+**Response:** Since we had a large number of time series (1530), we did not manually evaluate all possible combinations of trend, seasonality, and error terms in ETS for each time series.
+Instead we used an automatic algorithm to select the most appropriate model for each time series based on statistical criteria such as Akaike's Information Criterion (AIC).
 The automatic algorithm considered a mix of different component forms, depending on the specific characteristics of each time series.
@@
 
-_– for the linear reconciliation method  $\tilde{y}_h = S(S′W^{−1}S)^{−1}W^{−1} \hat{y}_h,$ how is it derived?_
+_– for the linear reconciliation method $\tilde{y}_h = S(S'W^{-1}S)^{-1}W^{-1}\hat{y}_h,$ how is it derived?_
 

### 2023-05-16 journal_correspondence/response_letter/response_letter_JSR.qmd
answer reviewer 2 and 3

@@
 **Response:** We appreciate your suggestion to improve the structure of the paper.
-We will consider moving the overview of base forecast algorithms to a separate section before Section 4 to provide a clearer delineation between the independent and hierarchical/grouped forecasts.
-This will help readers better understand the logic structure of the paper and facilitate easier navigation of the content.
+We have moved the overview of base forecast to a separate subsection before Section 4 to provide a clearer delineation between the independent and hierarchical/grouped forecasts.
+This can help readers better understand the logic structure of the paper and facilitate easier navigation of the content.
 
@@
 
-**Response:**
+**Response:** Thank you for your comment.
+We agree that it would be helpful to readers to have more information about the specific input predictors used in the GLM and TSGLM models.
+To address this, we have revised the manuscript to include the model formulas in R-format, as you suggested.
 
@@
 
-**Response:**
+**Response:** Thank you, we have revised the manuscript to include a more detailed reference for each equation used in the performance evaluation.
+Specifically, we have included the paper citation along with the equation number, so that readers can easily locate the relevant equation and read more about its derivation and interpretation.
 
 _Comment 4: The author extracts the trend and weekly seasonality of the data, and demonstrates their strengths in Figure 2.
-The follow-up analysis and forecast depend on the Poisson assumption most of the time due to the nature of non-negative count time series data._
-
-**Response:**
-
-_Comment 5: Is there any verification conducted to confirm the Poisson patterns in the data?
+The follow-up analysis and forecast depend on the Poisson assumption most of the time due to the nature of non-negative count time series data.
+Is there any verification conducted to confirm the Poisson patterns in the data?
 Are over-dispersion and/or under-dispersion going to be potential issues?
@@
 
-**Response:**
+**Response:** Thank you for your follow-up comment regarding the Poisson assumption used in our analysis.
+Regarding the potential issues of over-dispersion and under-dispersion, we acknowledge in the manuscript that these can be concerns when using Poisson regression models and included references to relevant literature that discuss these issues in more detail.
+We use the Poisson distribution to model the count data in our analysis.
+This is because the Poisson distribution is appropriate for count data that are non-negative and discrete, and it produces strictly positive forecasts.
+This is particularly important for some of the time series in our dataset, especially the bottom-level series, which have low volume and are more susceptible to producing negative forecasts if other distributions were used.
+Additionally, we have developed a flexible framework that can incorporate any forecasting model, but it is important to note that the intention of this paper is not to provide a comprehensive list of models.
+Instead, our focus is on demonstrating the usefulness of our framework for incorporating base forecasts generated by any model and producing hierarchical coherent forecasts.
 
-_Comment 6: ETS seems to have the disadvantage of not producing integer valued forecasts.
+_Comment 5: ETS seems to have the disadvantage of not producing integer valued forecasts.
 There are certainly other ways to do forecasting for integer valued time series.
@@
 
-**Response:**
+**Response:** We used ETS models because they are relatively simple, easy to understand, widely used, and well-established in the time series forecasting literature.
+They are particularly useful for modeling time series with multiple components, such as trend, seasonality, and/or cycles, and they are flexible and can be adapted to time series data with different patterns.
+This makes them appropriate for this study given the various patterns in the 1530 time series we analyzed.
 
-_Comment 7: How to understand the stationarity of the time series? Is it going to impact the analysis if the errors are assumed additive and the heteroscedasticity is present?_
+We acknowledge that ETS models do not necessarily produce integer-valued forecasts, which may be a disadvantage for certain applications.
+However, we note that rounding the forecasts can be a practical solution to address this issue.
+The impact of rounding on forecast accuracy depends on the level of hierarchy and the scale of the data.
+For high volume demand situations, rounding effects may be negligible, and forecast accuracy calculations can ignore integer effects.
+However, low volume demand settings, such as forecasting at the bottom level of the hierarchy, may be more susceptible to rounding effects.
 
-**Response:**
+Furthermore, we did consider other models specifically designed for integer-valued time series, such as TSGLM and GLM.
+We acknowledge that there are other methods available, such as INAR or INGARCH models.
+However, the intention of this study was not to provide a comprehensive list of forecasting models, but rather to develop a framework that can incorporate any forecasting model.
 
-_Comment 8: And if the errors are assumed multiplicative in a Poisson sense, what are the possible adjustments to verify or handle the stationarity before forecasting?_
+_Comment 6 (Check with Rob): How to understand the stationarity of the time series? Is it going to impact the analysis if the errors are assumed additive and the heteroscedasticity is present?_
 
-**Response:**
+**Response:** Thank you for your question.
 
-_Comment 9: The author introduces the algorithm for estimating the distributional forecasts of the data, and measures the accuracy using CRPS._
+Thank you for your comment.
+Stationarity refers to the property of a time series where its statistical properties (such as mean, variance) remain constant over time.
+In practical terms, a stationary time series has consistent behavior over time, regardless of when it is observed.
+To assess the stationarity of a time series we could extrcat features based on a test such as the Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test.
+These tests check whether a time series has a unit root, which is a stochastic trend that causes non-stationarity.
+Some models, such as ETS models, are designed to work with non-stationary time series.
+GLM and TSGLM can handle non-stationary time series by allowing for time-varying parameters that can capture changes in the underlying chnages of the data over time.
+They also allow for the inclusion of external regressors in the model, which can further help to capture different patterns in the data and be adapted to a wide range of time series , including non-stationary data.
 
-**Response:**
+Regarding your question about the impact of additive errors and heteroscedasticity on the analysis, in general, assuming additive errors does not require stationarity of the time series.
+However, if the time series is not stationary, the estimation of the error terms may be biased.
+Heteroscedasticity, which refers to the changing variance of the error terms over time, may also affect the estimation of the model parameters and the forecast accuracy that can be addressed by using a Box-Cox transformation.
 
-_– With the distributional forecasts, how do the confidence intervals behave?_
+_Comment 7: And if the errors are assumed multiplicative in a Poisson sense, what are the possible adjustments to verify or handle the stationarity before forecasting?_
 
-**Response:**
+**Response:**  Thank you for the question A possible adjustment could be applying the Box-Cox transformation before the forecasting.
+Also, alternative models such as the Negative Binomial Regression may be more appropriate.
+Both GLM and TSGLM allow for the use of  Negative Binomial, if required.
 
-_– As mentioned in P20 L14, the algorithm gives more accurate forecast results concerning the tails of the distribution.
+_Comment 8: The author introduces the algorithm for estimating the distributional forecasts of the data, and measures the accuracy using CRPS.
+With the distributional forecasts, how do the confidence intervals behave?_
+
+**Response:** Thank you.
+The performance of the distributional forecasts can be evaluated using the CRPS, which is a scoring rule that account for the shape and width of the predictive distribution.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.
+
+Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+It measures how well the predicted probabilities match the actual probabilities.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
+A model is well-calibrated if the predicted probabilities match the actual probabilities, and it is sharp if it is confident in its predictions.
+
+_Comment 9: As mentioned in P20 L14, the algorithm gives more accurate forecast results concerning the tails of the distribution.
 How does such outperformance visualized regarding this statement, and what to be read from the forecast distributions for risk management?_
 
-**Response:**
+**Response:** Thank you for your comment.
+In response to your question, we have included an additional visualization to demonstrate the tails of the forecast distribution for a given time series.
+While the center of the distribution may represent the most likely scenario, it may not necessarily capture the full range of potential outcomes that can affect decision-making at EMS.
+Therefore, a better understanding of the tails of the distribution can provide more valuable insights for managers to better prepare for and mitigate risks.
+Therefore, the accuracy of the tail forecasts can be critical for emergency planning and response.
 
@@
 
-**Response:**
+**Response:** Thank you for your comment and feedback.
+We appreciate your suggestion of adding supplementary material to provide more technical details and elaboration on some ideas.
+To respond to your queries, we have included the GitHub repository which includes all materials including data and R codes used in the paper.
+Additionally, we have included further explanation in the revisions of the paper.
 
@@
 
-**Response:**
+**Response:** Thank you for the comment.
+In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, which is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
+The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
+The strength of trend is defined as:
+$$F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$
+For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
+Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small.
+But for data with little or no trend, the two variances should be approximately the same.
+
+The strength of seasonality is defined similarly:
+$$F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).$$
+ series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).
 
@@
 
-**Response:**
+**Response:** Thank you.
+The aggregation matrix A is determined by the structure of the hierarchy.
+It maps the bottom-level time series to the corresponding higher-level time series, and recursively aggregates the time series at each level up to the top-level time series.
+The matrix A is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
+(Maybe we can show an example)
 
-_– for the linear reconciliation method $\tilde{y}_h = S(S'W^{-1}S)^{-1}W^{-1}\hat{y}_h,$ how is it derived?_
+_– (check with Rob) for the linear reconciliation method $\tilde{y}_h = S(S'W^{-1}S)^{-1}W^{-1}\hat{y}_h,$ how is it derived?_
 
@@
 
-**Response:**
+**Response:** Thank you or your comment.
+
+Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
+In this approach, the estimation of W is based on the assumption that all the errors have equal variance.
+Hence, W is simply defined as the identity matrix multiplied by a constant factor.
+The intuitions behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable, assuming the errors are normally distributed.
+
+Weighted Least Squares (WLS): WLS is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each observation.
+In this approach, W is defined as a diagonal matrix with the reciprocal of the variance of the errors on the diagonal.
+The intuitions behind WLS is that it assigns higher weight to observations with smaller variance of the errors, resulting in a better fit for the entire data set.
+
+Minimum Trace (MinT): MinT is a method that aims to minimize the trace of W, subject to the constraint that W is a positive definite matrix.
+It is often used in cases where the number of variables is larger than the number of observations, resulting in an underdetermined system of equations.
+In this approach, W is defined as the inverse of the covariance matrix of the errors.
+The intuitions behind MinT is that it shrinks the estimates towards zero and produces more stable results, particularly when the data has multicollinearity.
 
@@
 
-**Response:**
+**Response:** Thank you for providing minor comment. We have corrected the text accordingly.
 
@@
 
-**Response:**
+**Response:** Thank you for your comment. We have removed the indentation after equations in paragraphs that do not end.
 
@@
 
-**Response:**
+**Response:** Thank you.
+CRPS measures the difference between the forecasted cumulative distribution function (CDF) and the actual CDF of the forecast variable.
+The CRPS is computed as the integral of the squared difference between the forecasted CDF and the actual CDF, over the range of possible values.
+To provide a more specific example, suppose we have a time series of ambulance demand, and we want to forecast the demand for the next day The true probability in each day is the probability of observing the actual demand, which is always 1.
+The CRPS measures the difference between the forecasted probability distribution of the demand and the actual probability distribution of the demand.
+The yellow area in the plot of the forecasted CDF and the actual CDF highlights this difference.
+
+```{r, out.width= "60%"}
+knitr::include_graphics("figure/crps.jpg")
+```
 
@@
 
-**Response:**
+**Response:** This is now corrected.
 
@@
 
-**Response:**
+**Response:** The typo has been fixed now.
 
@@
 
-**Response:**
+**Response:** Thank you for your suggestion.Throughout, we ensured that a consistent format was maintained in the References section regarding the inclusion or exclusion of URL links.
 
@@
 
-**Response:**
+**Response:** The issue has been fixed now.
 
@@
 
-**Response:**
+**Response:** Thank you. We have reviewed Table 2 revised the table to make the column names more clear and consistent with Table 1.
 
@@
 
-**Response:**
+**Response:** We are glad to hear that the abstract and introduction piqued your interest as a reader and that you found the dataset engaging.
+However, we are sorry to learn that other aspects of the paper did not meet your expectations.
+Your feedback is valuable to us, and we have carefully considered it in our revision to address your concerns.
+Thank you for bringing these points to our attention.
 
@@
 
-**Response:**
+**Response:** Thank you for your feedback.
+We completely understand your concerns and acknowledge that transparency in our modeling process is crucial, and we would like to provide further information on how different forms of the trend and seasonality enter the model.
+For the exponential smoothing model (ETS), using an automatic algorithm, specifically the `ETS()` function in the fable package of R, to select the most appropriate model for each time series based on statistical criteria such as Akaike's Information Criterion (AIC).
+This approach allowed us to select the most appropriate components based on the individual characteristics of each time series, which was necessary given the large number of time series (1530) in our dataset.
+As a result, a mix of different component forms (additive or multiplicative) were used, depending on the specific characteristics of each time series at different level of the hierarchy.
+We hope that this information provides a better understanding of our approach to modeling.
 
-_Comment 1-b.
+_Comment 1-b (Rob to check).
 Related to point 1.a I was also wondering if an integrated estimation method would not be beneficial.
@@
 
-**Response:**
+**Response:** Thank you for your valuable feedback.
+We appreciate your comment and agree that providing an economic interpretation of the improved performances would greatly enhance the significance of our findings.
+We acknowledge that a forecast serves a purpose beyond its own existence and ideally it should enable the best utility such as the efficiency of allocating medical services, response time, and cost, informed by the forecast.
+
+While we recognize the importance of evaluating the quality of a forecast by considering its impact on decision-making processes, it is crucial to address the data requirements and methodology involved in measuring this impact.
+To comprehensively assess the implications of forecast, it would be necessary to have access to further data beyond ambulance demand across the hierarchy such as different decision types, capacity data, constraint in the decision system, etc.
+This additional data would provide insights into the specific decisions that rely on the forecasts, allowing for a more accurate evaluation of the impact on medical services.
+Unfortunately, obtaining such data is not feasible and accessible within the scope of our current study.
+Moreover, measuring the impact of forecasts would require a different approach that extends beyond the forecasting itself.
+It would involve developing and implementing simulation models that can simulate the decision-making processes based on the forecast inputs.
+These models would evaluate the quality of the final decision, considering the utilities that are of particular importance to the EMS context.
+
+While we were unable to measure the specific economic impacts and utilities of forecasts largely due to data limitations in the current study, we believe that our research lays a strong foundation for future research investigations, as well as practical implementation in EMS.
+The proposed forecasting framework and forecast reconciliation methods offer a robust framework that can be extended and integrated with operational information to evaluate decision quality and forecast utility in EMS.
+EMS can use this advanced forecasting framework that exploit the hierarchical and grouped structure of demand time series, and apply forecast reconciliation methods to generate coherent point and probabilistic forecasts that use all available data at all levels of disaggregation.
+
+Moreover, future research can build upon this foundation by incorporating additional data and implementing simulation modeling techniques.
+This would enable a more comprehensive evaluation of the economic impact, efficiency gains, and utility of the forecasted results in the context of EMS.
+In future investigations, we aim to explore this avenue by incorporating operational information, simulating decision processes, and assessing the impact on utilities that are significant to the EMS.
+To that end, we have started presenting this study to different ambulance services and we hope to establish a collaboration that allow us to access data and decisions made across the hierarchy to evaluate its impact.
+
+We appreciate your thoughtful comment, and we will take it into account for future research.
 
@@
 
-**Response:**
+**Response:** Thank you for your insightful comment.
+In our study, we chose the Continuous Ranked Probability Score (CRPS) as one of our evaluation metrics because it provides a comprehensive assessment of forecast quality and allows for the comparison of different models.
+The CRPS strikes a balance between calibration and sharpness of forecast distributions.
+Calibration refers to the statistical consistency between the forecasted distributions and the actual observations.
+It measures how well the predicted probabilities align with the observed probabilities.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating a high level of confidence in the forecast.
+A well-calibrated model matches the predicted probabilities to the actual probabilities, while a sharp model is confident in its predictions.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.
+
+In the context of emergency medical services, it is crucial to consider the implications of forecasts that are not well-calibrated or not sharp.
+If a forecast is not well-calibrated, it means that the predicted probabilities do not accurately reflect the actual probabilities of events, leading to potential biases in decision-making processes.
+For example, if the forecast consistently underestimates the demand for ambulance services, it could result in inadequate resource allocation and delayed responses to emergencies.
+Conversely, if the forecast consistently overestimates the demand, it may lead to unnecessary allocation of resources and increased costs.
+On the other hand, the sharpness of a forecast is an indicator of the level of confidence in the predictions.
+When a forecast is not sharp, it suggests a lack of confidence in the estimated outcomes.
+This can have implications for decision-making processes within emergency medical services.
+If the intervals are excessively wide, decision-makers may adopt overly conservative strategies, leading to underutilization of resources and potentially delayed responses.
+Conversely, if the intervals are excessively narrow, decision-makers may take overly aggressive actions, risking resource shortages and potential errors in resource allocation.
+Therefore, it is crucial to assess both calibration and sharpness of forecast to ensure effective decision-making within emergency medical services.
 
@@
 
-**Response:**
+**Response:** Thank you for raising the concern about the generalizability of our proposed framework in the context of the service industry.
+We appreciate your feedback and have addressed it in the revised paper.
+
+While our study primarily focuses on the domain of emergency medical services, we would like to emphasize that the framework we propose is applicable to a wide range of industries, including the service industry.
+Wherever there are collections of related time series structured hierarchically and/or in a grouped manner, our framework can be utilized to provide accurate forecasts.
+The grouped and hierarchical structure of time series data is a common characteristic across many sectors, and it arises when data can be naturally organized into various levels of aggregation or when there are dependencies and relationships among different entities within the system.
+
+For example, in the supply chain industry, accurately forecasting demand at different levels of the distribution network, such as regional warehouses or retail stores, is essential for efficient inventory management and minimizing stockouts.
+Our framework enables the reconciliation of forecasts to ensure consistency and alignment across the supply chain, leading to improved decision-making and operational efficiency.
+
+In the financial industry, where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
+By considering the hierarchical structure of the data, financial institutions can make informed investment decisions, manage risk effectively, and meet the specific demands of different customer segments.
+
+In the transportation sector, our framework can support forecasting transportation demand at various levels, such as across different states or regions, to optimize route planning and resource allocation.
+Similarly, in the hospitality and tourism industry, our framework can be utilized to forecast demand rates at state, regional, and department levels, facilitating strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses.
+
+Furthermore, in call centers, accurately forecasting call volumes at different levels of the call center hierarchy is crucial for workforce management and resource allocation.
+By applying our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, enabling efficient staffing and optimal service levels to meet customer demands.
+
+In summary, the strength of our proposed framework lies in its ability to handle and reconcile time series data structured hierarchically and/or in a grouped manner.
+This characteristic can be found across a multitude of service industries, making our framework applicable in various contexts.
+By considering the relationships and dependencies within the data, our approach can contribute to more accurate forecasts, more coordinated activities due to consistent forecasts and potentially improved decision-making.
 

### 2023-05-16 journal_correspondence/response_letter/response_letter_JSR.qmd
typos

@@
 **Response:** Thank you for your question.
-
-Thank you for your comment.
 Stationarity refers to the property of a time series where its statistical properties (such as mean, variance) remain constant over time.
@@
 Some models, such as ETS models, are designed to work with non-stationary time series.
-GLM and TSGLM can handle non-stationary time series by allowing for time-varying parameters that can capture changes in the underlying chnages of the data over time.
-They also allow for the inclusion of external regressors in the model, which can further help to capture different patterns in the data and be adapted to a wide range of time series , including non-stationary data.
+GLM and TSGLM can handle non-stationary time series by allowing for time-varying parameters that can capture changes in the underlying changes of the data over time.
+They also allow for the inclusion of external predictors in the model, which can further help to capture different patterns in the data and be adapted to a wide range of time series , including non-stationary data.
 

### 2023-05-16 journal_correspondence/response_letter/response_letter_JSR.qmd
typos

@@
 
-**Response:** Thank you or your comment.
+**Response: (Check with Rob)** Thank you or your comment.
 
@@
 
-Weighted Least Squares (WLS): WLS is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each observation.
+Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each observation.
 In this approach, W is defined as a diagonal matrix with the reciprocal of the variance of the errors on the diagonal.
@@
 
-Minimum Trace (MinT): MinT is a method that aims to minimize the trace of W, subject to the constraint that W is a positive definite matrix.
+Minimum Trace (MinT) is a method that aims to minimize the trace of W, subject to the constraint that W is a positive definite matrix.
 It is often used in cases where the number of variables is larger than the number of observations, resulting in an underdetermined system of equations.

### 2023-05-19 journal_correspondence/response_letter/response_letter_JSR.qmd
typos
typo
Updated some responses

--- 2023-05-16 journal_correspondence/response_letter/response_letter_JSR.qmd
@@
 
-  > For example, the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
+  > For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
 If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area.
@@
 
-  **Response:** Thank you for bringing this to our attention. We have gone through the manuscript and made sure that we consistently and correctly report the duration of the dataset as 2015 to 2019.
+  **Response:**  We have gone through the manuscript and made sure that we consistently and correctly report the duration of the dataset as 2015 to 2019.
 
@@
 
-  **Response:**  Thank you.
-We agree that a table could provide a clearer overview of the related literature and our contribution.
+  **Response:**  We agree that a table could provide a clearer overview of the related literature and our contribution.
 Therefore, we have now included a table in the literature review section summarizing previous research, including our study's contribution.
@@
 
-**Response:** Thank you for pointing out that our descriptions in sections 3 and 4 could benefit from more detail.
-We have revised those sections to include more specific information on the forecasting models and hierarchical and grouped time series techniques utilized in our study.
+**Response:** We have now revised sections 3 and 4 to include more specific information on the forecasting models and hierarchical and grouped time series techniques utilized in our study.
 
@@
 
-**Response:** Thank you for your thoughtful comment regarding the potential impact of failed responses on our time series forecasting.
-We agree that failed responses could have an impact on forecasts, as the actual demand for ambulance services could be higher than what we have observed in the attended incidents data.
+**Response:** We agree that failed responses could have an impact on forecasts, as the actual demand for ambulance services could be higher than what we have observed in the attended incidents data.
 Unfortunately, as you noted, the dataset we used only includes information on attended incidents, and we do not have access to information on failed responses.
@@
 
-**Response:** Thank you for your insightful comment.
-We appreciate your point about the importance of both high-level and low-level forecasts for effective decision-making in EMS.
+**Response:** We appreciate your point about the importance of both high-level and low-level forecasts for effective decision-making in EMS.
 In our study, we use a methodology that provides forecasts for all levels of a given hierarchy including high-level and low-level of the hierarchy.
@@
 
-**Response:** Thank you for the comment.
-In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, which is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
+**Response:**  In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by Kang et al (2017).
+STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
 The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
@@
 
-**Response:** Thank you for your feedback.
-We appreciate your concern regarding the level of detail provided in Section 3.2 on the forecasting approaches used in this study.
+**Response:**  To address your concern, we have revised Section 3.2 to provide a more detailed description of each of the forecasting approaches used, including the mathematical framework, assumptions, and key parameters.
 While we agree that it is important to provide adequate information in the paper itself, we also understand that some readers may be interested in more technical details and code implementation.
 For this reason, we have provided a Github repository with all technical details and R codes for the forecasting approaches used in this study.
-However, we acknowledge that more information can be provided in the paper itself to enhance readers' understanding.
-To address your concern, we have revised Section 3.2 to provide a more detailed description of each of the forecasting approaches used, including the mathematical framework, assumptions, and key parameters.
 We hope that this revised section, along with the Github repository, will provide readers with the necessary information to understand and replicate the methods used in this study.
@@
 
-**Response:** Thank you for the suggestion.
-We appreciate your feedback.
-We would like to clarify that our intention in this paper is not to provide an exhaustive list of forecasting methods.
+**Response:**  We would like to clarify that our intention in this paper is not to provide an exhaustive list of forecasting methods.
 Instead, we aim to provide a methodology that can incorporate forecasts produced by any method.
 Our focus is on developing a hierarchical time series forecasting approach that takes into account the hierarchical and group structure of ambulance demand.
-We have selected the naive method as a benchmark because it is widely used in forecasting literature and practice and is a simple and easily understandable method for users.
+We have selected the naive method as a benchmark because it is widely used in the forecasting literature and practice and is a simple and easily understandable method for users.
 
@@
 
-**Response:** Thank you for the question.
-In our research, we used an automatic algorithm to determine the appropriate form for the trend, seasonality, and error terms in each time series.
+**Response:** In our research, we used an automatic algorithm to determine the appropriate form for the trend, seasonality, and error terms in each time series.
 Specifically, we used the `ETS()` function in the fable package of R, which selects the most appropriate model for each time series based on statistical criteria such as Akaike's Information Criterion (AIC).
@@
 
-**Response:** Thank you for your question. We have provided responses to each of the two separate questions below:
+**Response:**  We have provided responses to each of the two separate questions below:
 
@@
 
-**Response:** Thank you for your comment on section 4.2.2. We apologize for any confusion that may have arisen from the text. Please allow us to provide some clarification.
+**Response:** We apologize for any confusion that may have arisen from the text. Please allow us to provide some clarification.
 
@@
 For some situations, with high volume demand, forecast accuracy calculations can ignore integer effects as rounding becomes negligible.
-However, low volume demand settings, such as forecast at bottom level of hierarchy,  may be more susceptible to integer (rounding) effects.
+However, low volume demand settings, such as forecasts at the bottom level of the hierarchy, may be more susceptible to integer (rounding) effects.
 In our study, we kept the non-integer forecast values after reconciliation when evaluating the accuracy.
@@
 **Response:** The reason for the discrepancy in the number of series is that for some combinations of priority and nature of incidents, there is no representation in the dataset.
-Therefore, these combinations are not included in the table For example, in the "Priority * Nature of incident" combination, one of the cells in the cross-tabulation have no incidents recorded for a combination of Red Green priority and Falls, so it is not included in the dataset.
+Therefore, these combinations are not included in the table.
+For example, in the "Priority * Nature of incident" combination, one of the cells in the cross-tabulation has no incidents recorded for the combination of Red Green priority and Falls, so it is not included in the dataset.
 The same reasoning applies to the other two combinations in Table 1.
@@
 **Response:** We appreciate your suggestion to improve the structure of the paper.
-We have moved the overview of base forecast to a separate subsection before Section 4 to provide a clearer delineation between the independent and hierarchical/grouped forecasts.
+We have moved the overview of base forecasts to a separate subsection before Section 4 to provide a clearer delineation between the independent and hierarchical/grouped forecasts.
 This can help readers better understand the logic structure of the paper and facilitate easier navigation of the content.
@@
 
-**Response:**  We agree that providing more detail on the ETS and Ensemble methods, as well as other methods, would enhance the understanding of our proposed approach.
-In the revised version, we have added more detailed explanations on these methods, including their mathematical formulations, assumptions, and limitations.
-We will also provide additional references to the Ensemble method and further explanations on how it is applied in our study.
+**Response:**  In the revised version, we have added more detailed explanations on these methods, including their mathematical formulations, assumptions, and limitations.
+We also provide additional references to the Ensemble method and further explanations on how it is applied in our study.
 
@@
 
-**Response:** Thank you for your comment.
-We agree that it would be helpful to readers to have more information about the specific input predictors used in the GLM and TSGLM models.
-To address this, we have revised the manuscript to include the model formulas in R-format, as you suggested.
+**Response:** We have revised the manuscript to include the model formulas, as you suggested.
 
@@
 
-**Response:** Thank you, we have revised the manuscript to include a more detailed reference for each equation used in the performance evaluation.
+**Response:** We have revised the manuscript to include a more detailed reference for each equation used in the performance evaluation.
 Specifically, we have included the paper citation along with the equation number, so that readers can easily locate the relevant equation and read more about its derivation and interpretation.
@@
 
-**Response:** Thank you for your follow-up comment regarding the Poisson assumption used in our analysis.
-Regarding the potential issues of over-dispersion and under-dispersion, we acknowledge in the manuscript that these can be concerns when using Poisson regression models and included references to relevant literature that discuss these issues in more detail.
-We use the Poisson distribution to model the count data in our analysis.
-This is because the Poisson distribution is appropriate for count data that are non-negative and discrete, and it produces strictly positive forecasts.
-This is particularly important for some of the time series in our dataset, especially the bottom-level series, which have low volume and are more susceptible to producing negative forecasts if other distributions were used.
-Additionally, we have developed a flexible framework that can incorporate any forecasting model, but it is important to note that the intention of this paper is not to provide a comprehensive list of models.
+**Response:** Provided accidents occur independently, they will follow a Poisson distribution by definition (Feller, 3rd ed, 1968, p156-158).
+So it is reasonable to assume a Poisson distribution here.
+We allow for the mean of the Poisson distribution to change over time by including trend and seasonality covariates, and public holiday effects.
+If there are other factors that affect the mean of the Poisson distribution, but which aren't in our model, we may observe over-dispersion or under-dispersion.
+We have tested for this ??????
+
+We have developed a flexible framework that can incorporate any forecasting model, and it is important to note that the intention of this paper is not to provide a comprehensive list of models, or to argue for a particular model class.
 Instead, our focus is on demonstrating the usefulness of our framework for incorporating base forecasts generated by any model and producing hierarchical coherent forecasts.
@@
 We acknowledge that there are other methods available, such as INAR or INGARCH models.
-However, the intention of this study was not to provide a comprehensive list of forecasting models, but rather to develop a framework that can incorporate any forecasting model.
+However, neither INAR or INGARCH are able to handle the non-stationary trends and seasonality, or public holiday effects, that are evident in the data.
+Further, we note that the intention of this study was not to provide a comprehensive list of forecasting models, but rather to develop a framework that can incorporate any forecasting model.
 
-_Comment 6 (Check with Rob): How to understand the stationarity of the time series? Is it going to impact the analysis if the errors are assumed additive and the heteroscedasticity is present?_
+_Comment 6: How to understand the stationarity of the time series? Is it going to impact the analysis if the errors are assumed additive and the heteroscedasticity is present?_
 
-**Response:** Thank you for your question.
-Stationarity refers to the property of a time series where its statistical properties (such as mean, variance) remain constant over time.
-In practical terms, a stationary time series has consistent behavior over time, regardless of when it is observed.
-To assess the stationarity of a time series we could extrcat features based on a test such as the Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test.
-These tests check whether a time series has a unit root, which is a stochastic trend that causes non-stationarity.
-Some models, such as ETS models, are designed to work with non-stationary time series.
-GLM and TSGLM can handle non-stationary time series by allowing for time-varying parameters that can capture changes in the underlying changes of the data over time.
-They also allow for the inclusion of external predictors in the model, which can further help to capture different patterns in the data and be adapted to a wide range of time series , including non-stationary data.
-
-Regarding your question about the impact of additive errors and heteroscedasticity on the analysis, in general, assuming additive errors does not require stationarity of the time series.
-However, if the time series is not stationary, the estimation of the error terms may be biased.
-Heteroscedasticity, which refers to the changing variance of the error terms over time, may also affect the estimation of the model parameters and the forecast accuracy that can be addressed by using a Box-Cox transformation.
+**Response:** All of the time series in our data set are non-stationary, due to the changing rate of accidents with the time-of-year, day-of-week, and other factors.
+All our models, other than Naive, allow for these non-stationary features of the data.
+None of our models assume additive errors or homoscedasticity.
+The ETS models allow for multiplicative errors and heteroscedasticity where appropriate.
+The GLM and TSGLM models also do not have additive errors and have implicity heteroscedasticity due to the Poisson distribution.
 
@@
 
-**Response:**  Thank you for the question A possible adjustment could be applying the Box-Cox transformation before the forecasting.
-Also, alternative models such as the Negative Binomial Regression may be more appropriate.
-Both GLM and TSGLM allow for the use of  Negative Binomial, if required.
+**Response:** We do not understand this question. The data are not stationary.
 
@@
 
-**Response:** Thank you.
-The performance of the distributional forecasts can be evaluated using the CRPS, which is a scoring rule that account for the shape and width of the predictive distribution.
+**Response:** The performance of the distributional forecasts can be evaluated using the CRPS, which is a scoring rule that accounts for the shape and width of the predictive distribution.
 The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
@@
 
-Calibration refers to the statistical consistency between the distributional forecasts and the observations.
-It measures how well the predicted probabilities match the actual probabilities.
-On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
-A model is well-calibrated if the predicted probabilities match the actual probabilities, and it is sharp if it is confident in its predictions.
+A specific prediction interval could be evaluated using a Winkler score (Winkler, JASA, 1992).
+CRPS can be considered an average of all possible Winkler scores [@fpp3, Section 5.9], and thus provides an evaluation of all possible prediction intervals.
 
@@
 
-**Response:** Thank you for your comment.
-In response to your question, we have included an additional visualization to demonstrate the tails of the forecast distribution for a given time series.
+**Response:** In response to your question, we have included an additional visualization to demonstrate the tails of the forecast distribution for a given time series.
 While the center of the distribution may represent the most likely scenario, it may not necessarily capture the full range of potential outcomes that can affect decision-making at EMS.
@@
 
-**Response:** Thank you for your comment and feedback.
-We appreciate your suggestion of adding supplementary material to provide more technical details and elaboration on some ideas.
+**Response:** We appreciate your suggestion of adding supplementary material to provide more technical details and elaboration on some ideas.
 To respond to your queries, we have included the GitHub repository which includes all materials including data and R codes used in the paper.
@@
 
-**Response:** Thank you.
-The aggregation matrix A is determined by the structure of the hierarchy.
-It maps the bottom-level time series to the corresponding higher-level time series, and recursively aggregates the time series at each level up to the top-level time series.
-The matrix A is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
-(Maybe we can show an example)
+**Response:**  The aggregation matrix $A$ is determined by the structure of the hierarchy.
+It maps the bottom-level time series to the corresponding higher-level time series.
+The matrix $A$ is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
+For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $A = \begin{bmatrix} 1 1 \end{bmatrix}$.
 
-_– (check with Rob) for the linear reconciliation method $\tilde{y}_h = S(S'W^{-1}S)^{-1}W^{-1}\hat{y}_h,$ how is it derived?_
+_– for the linear reconciliation method $\tilde{y}_h = S(S'W^{-1}S)^{-1}W^{-1}\hat{y}_h,$ how is it derived?_
 
-**Response:**
+**Response:** Linear reconciliation involves projecting the base forecasts onto the coherent space.
+It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased (Wickramasuriya et al, JASA 2019).
 
@@
 
-**Response:** Thank you or your comment.
+**Response:**
 
@@
 Hence, W is simply defined as the identity matrix multiplied by a constant factor.
-The intuitions behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable, assuming the errors are normally distributed.
+The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
+The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
+A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.
 
-Weighted Least Squares (WLS): WLS is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each observation.
-In this approach, W is defined as a diagonal matrix with the reciprocal of the variance of the errors on the diagonal.
-The intuitions behind WLS is that it assigns higher weight to observations with smaller variance of the errors, resulting in a better fit for the entire data set.
+Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
+But it assumes that the errors of each series are uncorrelated with each other.
+In this approach, W is defined as a diagonal matrix with the variance of the errors on the diagonal.
+The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
+The main weakness of this approach is that it ignores the relationships between series.
+A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.
 
-Minimum Trace (MinT): MinT is a method that aims to minimize the trace of W, subject to the constraint that W is a positive definite matrix.
-It is often used in cases where the number of variables is larger than the number of observations, resulting in an underdetermined system of equations.
-In this approach, W is defined as the inverse of the covariance matrix of the errors.
-The intuitions behind MinT is that it shrinks the estimates towards zero and produces more stable results, particularly when the data has multicollinearity.
+Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.
+Wickramasuriya et al (JASA 2019) showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
+The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
+In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.
 
@@
 
-**Response:** Thank you for providing minor comment. We have corrected the text accordingly.
+**Response:** We have corrected the text accordingly.
 
@@
 
-**Response:** Thank you for your comment. We have removed the indentation after equations in paragraphs that do not end.
+**Response:**  We have removed the indentation after equations in paragraphs that do not end.
 
@@
 
-**Response:** Thank you.
-CRPS measures the difference between the forecasted cumulative distribution function (CDF) and the actual CDF of the forecast variable.
-The CRPS is computed as the integral of the squared difference between the forecasted CDF and the actual CDF, over the range of possible values.
-To provide a more specific example, suppose we have a time series of ambulance demand, and we want to forecast the demand for the next day The true probability in each day is the probability of observing the actual demand, which is always 1.
+**Response:**  CRPS measures the integrated squared difference between the forecasted cumulative distribution function (CDF), $G_j(x)$$, and the true CDF of the forecast variable, $F_j(x)$.
+We estimate the latter using the empirical distribution function.
+Thus, once observed, $\hat{F}_j(x)$ is the empirical distribution function taking value 0 before $x=y_t$ and value 1 after $x=y_t$, as shown in the figure below.
 The CRPS measures the difference between the forecasted probability distribution of the demand and the actual probability distribution of the demand.
@@
 
-**Response:** Thank you for your suggestion.Throughout, we ensured that a consistent format was maintained in the References section regarding the inclusion or exclusion of URL links.
+**Response:** Throughout, we ensured that a consistent format was maintained in the References section regarding the inclusion or exclusion of URL links.
 
@@
 
-**Response:** Thank you. We have reviewed Table 2 revised the table to make the column names more clear and consistent with Table 1.
+**Response:**  We have reviewed Table 2 revised the table to make the column names more clear and consistent with Table 1.
 
@@
 
-**Response:** Thank you for your feedback.
-We completely understand your concerns and acknowledge that transparency in our modeling process is crucial, and we would like to provide further information on how different forms of the trend and seasonality enter the model.
+**Response:**  We completely understand your concerns and acknowledge that transparency in our modeling process is crucial, and we would like to provide further information on how different forms of the trend and seasonality enter the model.
 For the exponential smoothing model (ETS), using an automatic algorithm, specifically the `ETS()` function in the fable package of R, to select the most appropriate model for each time series based on statistical criteria such as Akaike's Information Criterion (AIC).
@@
 
-_Comment 1-b (Rob to check).
+_Comment 1-b.
 Related to point 1.a I was also wondering if an integrated estimation method would not be beneficial.
@@
 
-**Response:**
+**Response:** Indeed, it is possible to define a state space model that ensures coherent forecasts.
+This was proposed by Pennings & Dalen (2017), and Villegas & Pedregal (2018).
+However, in these models, the covariance matrices are extremely difficult to estimate with anything other than very small hierarchies.
+The strength of our approach is that it allows for any type of forecasting method to be used for the base forecasts, including different methods for different series.
+It decouples the time series models from the reconciliation step, thus allowing much more flexibility in the types of models that can be considered.
 
@@
 
-**Response:** Thank you for your valuable feedback.
-We appreciate your comment and agree that providing an economic interpretation of the improved performances would greatly enhance the significance of our findings.
+**Response:**  We appreciate your comment and agree that providing an economic interpretation of the improved performances would greatly enhance the significance of our findings.
 We acknowledge that a forecast serves a purpose beyond its own existence and ideally it should enable the best utility such as the efficiency of allocating medical services, response time, and cost, informed by the forecast.
@@
 While we recognize the importance of evaluating the quality of a forecast by considering its impact on decision-making processes, it is crucial to address the data requirements and methodology involved in measuring this impact.
-To comprehensively assess the implications of forecast, it would be necessary to have access to further data beyond ambulance demand across the hierarchy such as different decision types, capacity data, constraint in the decision system, etc.
+To comprehensively assess the implications of the forecasts, it would be necessary to have access to further data beyond ambulance demand across the hierarchy, such as different decision types, capacity data, constraint in the decision system, etc.
 This additional data would provide insights into the specific decisions that rely on the forecasts, allowing for a more accurate evaluation of the impact on medical services.
@@
 
-**Response:** Thank you for your insightful comment.
-In our study, we chose the Continuous Ranked Probability Score (CRPS) as one of our evaluation metrics because it provides a comprehensive assessment of forecast quality and allows for the comparison of different models.
+**Response:**  In our study, we chose the Continuous Ranked Probability Score (CRPS) as one of our evaluation metrics because it provides a comprehensive assessment of forecast quality and allows for the comparison of different models.
 The CRPS strikes a balance between calibration and sharpness of forecast distributions.

### 2023-06-06 manuscript/05-results.qmd
Ensemble without naive added as Ensemble2

@@
   mutate(model = factor(model,
-    levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
-    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
+    levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
+    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT")),
@@
     model = factor(model,
-      levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
-      labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
+      levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
+      labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
@@
   mutate(model = factor(model,
-    levels = c("naiveecdf","ets","iglm","tscount","ensemble"),
-    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble")),
+    levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
+    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
@@
     model == "iglm" ~ "GLM",
-    model == "ensemble" ~ "Ensemble"
+    model == "ensemble" ~ "Ensemble",
+    model == "ensemble2" ~ "Ensemble2"
   )) |> 
@@
     series = factor(series, levels = c("Total", "Control areas", "Health boards")),
-    model = factor(model, levels = c("Naive", "ETS", "TSGLM", "GLM", "Ensemble")),
+    model = factor(model, levels = c("Naive", "ETS", "TSGLM", "GLM", "Ensemble", "Ensemble2")),
     week = factor(trunc((h - 1) / 7) + 1)

### 2023-06-06 rscript/hts/accuracy.R

@@ -13,7 +13,7 @@ compute_accuracy <- function(train, measure = "mse") {
   files <- fs::dir_ls(storage_folder, glob = paste0("*_sim_*.rds"))
   # Find models
   models <- str_remove(files, storage_folder) |>
-    str_extract("[a-zA-Z]*_") |>
+    str_extract("[a-zA-Z0-9]*_") |>
     str_remove("_") |>
     unique()
   # Find methods

### 2023-07-18 journal_correspondence/response_letter/response_letter_JSR.qmd
response letter and paper ready for submission

@@
 ---
-title: "Response letter"
+title: "Response letter - Manuscript ID JSR-22-468"
+bibliography: bibliography.bib
 format:
@@
 We quote below the detailed referees’ comments that are followed by our response.
-In all cases, we point out, where necessary, the corresponding amendments in the new version of the paper and highlighted them in blue.
+In all cases, we point out, where necessary, the corresponding amendments in the new version of the paper and highlighted them in teal (greenish-blue color).
 
@@
 
+Response:  Thank you very much for considering our paper for publication.
+We are sincerely grateful to the reviewers for providing such valuable feedback, and we highly value the comments raised by all reviewers and the Associate Editor.
+We have devoted significant effort to carefully reflect on the concerns raised and have made every effort to address them in the revised version of the manuscript.
+
+We agree that the reviewers have raised important questions and provided valuable recommendations regarding the finer details of our study, and we have taken all of these into account in the revised manuscript.
+In response to the suggestion to look beyond the point estimate, we have incorporated an example that highlights the advantages of doing so.
+Furthermore, we have thoroughly addressed the specific suggestions provided by Reviewers 1 and 2 in the revised manuscript.
+Additionally, we have considered and responded to Reviewer 3's remarks concerning the performance of the Naïve model and Figure 4.
+To emphasize the significance of our study, we have included additional paragraphs in the paper that highlight its importance and potential extensions.
+Specifically, in response to R3's remark, we have discussed how the framework can be extended to incorporate the decision-making process and assess the implications of forecasts on crucial aspects like cost allocation and resource management, which are vital to planners.
+Finally, we elaborate on how our current study has broader implications and can be generalized to other areas within the service industries.
+This highlights the wide applicability and potential impact of our research in various contexts.
+
+We quote below the detailed referees’ comments that are followed by our response.
+In all cases, we point out, where necessary, the corresponding amendments in the new version of the paper and highlighted them in teal (greenish-blue color).
+
 # Associate Editor
@@
 Regarding the recommendation to look beyond the point estimate, we have taken the reviewers' suggestions into account and provide an example that illustrates the advantage of doing so.
-We have implemented R1's suggestions in the revised manuscript.
+We have responded and implemented suggestion provided by R1 and R2 in the revised manuscript.
 We have also responded to R3's remark regarding the performance of the Naïve model and Figure 4.
+We have also included additional paragraphs in the paper to highlight the importance of the current study and how it can be extended to incorporate the decision making process and evaluate the implication of forecasts on decision impacts and utilities important to planners such as cost, allocating resources, etc.
+Finally, we added a paragraph to highlight how the current study can be generalized to other areas in service industries.
 
@@
 We agree that broadening the scope of the study to include other services that face similar hierarchical and grouped time series forecasting issues would make it more relevant and interesting to the Journal of Service Research audience.
-We have added a new subsection to the manuscript in the conclusion section, where we discuss other services that share the same problem and how our approach could be applied to those services.
-We also linked our work to previous research in the Journal of Service Research to provide a more comprehensive understanding of our contribution.
+We have added a new paragraph to the manuscript in the conclusion section, where we discuss other services that share the same problem and how our approach could be applied to those services.
+We also linked our work to previous research in the Journal of Service Research.
 
@@
   **Response:** We agree that providing a detailed example can better illustrate the problems that arise when separate forecasts at different levels are inconsistent.
-We have added the following example to the manuscript, which outlines the issue of incoherent in hierarchical forecasts.
+We have added the following example to the introduction section in the manuscript, which outlines the issue of incoherent in hierarchical forecasts.
 
-  > For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
+\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
 If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area.
@@
 Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly.
-Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).
+Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}
 
@@
 Therefore, we have now included a table in the literature review section summarizing previous research, including our study's contribution.
+We also included two additional references in the table and added the relevant texts in the literature section.
+
+```{r summarylit, echo=FALSE}
+library(kableExtra)
+li <- readxl::read_xlsx("../../literature_table/table.xlsx")
+knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies in forecasting FOR AMBULANCE SERVICES") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
+  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |>  kableExtra::landscape()
+```
+
+@al2021empirical \textcolor{teal}{utilises real data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand accurately.
+The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
+Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE) are calculated.
+The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
+However, when it comes to long-term demand, the SSA method proves to be the most effective.} @ibrahim2016modeling \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
+Real-life call centre data was utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
+The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
+The findings indicate that the ME consistently produces the most accurate point forecasts.
+On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
+For a two-week leading period, MU shows better coverage probability.}
 
@@
   **Response:** We have reviewed each table and figure in the manuscript and ensure that they are complete and understandable without relying on the main text and all necessary information such as axis labels, legends, etc are provided.
+Please refer to notes or captions in each figure and table in the revised manuscript.
 
@@
 
-  **Response:** We have updated Table 1 to make it easier for readers to understand all values by providing a breakdown of the calculation.
+  **Response:** We have updated Table 1 to make it easier for readers to understand all values by providing a breakdown of the calculation, where feasible. Please see the footnote added to Table 1:
+
+\textcolor{teal}{Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset.
+As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.}
 
@@
 
-  **Response:** We agree that displaying the number of observations is important when reporting statistical results. We have added information about the size of the data set where relevant.
+  **Response:** We agree that displaying the number of observations could be important when reporting results.
+We have added information about the size of the data set where relevant.
+Please see captions od figures and tables for reference.
 
@@
 **Response:** We have now revised sections 3 and 4 to include more specific information on the forecasting models and hierarchical and grouped time series techniques utilized in our study.
+Please refer to section 3 and 4 in the revised manuscript where we highlighted changes in teal.
 
@@
 However, the forecasting results may differ depending on the specific characteristics of the data.
+We have added the following text in the conclusion section of the revised manuscript:
+
+\textcolor{teal}{The dataset used in this study only includes information on attended incidents.
+It would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
+It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.}
 
@@
 **Response:** We appreciate your point about the importance of both high-level and low-level forecasts for effective decision-making in EMS.
-In our study, we use a methodology that provides forecasts for all levels of a given hierarchy including high-level and low-level of the hierarchy.
+In our study, we use a methodology that provides forecasts for all levels of a given hierarchy including high-level and low-level of the hierarchy for a given daily temporal period.
+The methodology could be used for any temporal interval such as sub-daily, weekly, monthly or quarterly.
 We believe that wherever decisions are made about the future in the hierarchy, forecasts would be used as an input.
@@
 
+We have included the following text in the introduction section of the revised manuscript:
+
+\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchy are necessary for effective decision-making in EMS.
+For example, control area EMS forecasts can inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals.
+At the lower levels, hospitals or ambulance stations could use such forecasts to plan for staffing and resource allocation, ambulance dispatching, staff-to-shift assignment, staff rescheduling based on the anticipated volume and priority and nature of incidents.
+Additionally, generating forecasts at lower levels could potentially improve the accuracy of the high-level forecasts, by providing more detailed information on the nature and priority of incidents.
+This could help to identify patterns in demand that may not be apparent at the higher level.
+Therefore, employing forecasting techniques that consider the hierarchical and/or grouped patterns of time series in EMS aligns naturally, offering the possibility to enhance forecast accuracy and facilitate coordination.}
+
 _Comment 3. Second paragraph of page 9 mentions “the strength of trend” and “the strength of seasonality”. Can you explain what those are or how they are calculated?_
 
-**Response:**  In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by Kang et al (2017).
-STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
+**Response:**  Thank you. We have added the following text to the revised manuscript to address this comment:
+
+\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by} @kang2017visualising.
+\textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
 The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
-The strength of trend is defined as:
+The strength of trend is defined as:}
 $$F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$
-For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
+\textcolor{teal}{For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
 Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small.
-But for data with little or no trend, the two variances should be approximately the same.
+But for data with little or no trend, the two variances should be approximately the same.}
 
-The strength of seasonality is defined similarly:
+\textcolor{teal}{The strength of seasonality is defined similarly:}
 $$F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).$$
- series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).
+\textcolor{teal}{series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).}
 
@@
 
-**Response:**  To address your concern, we have revised Section 3.2 to provide a more detailed description of each of the forecasting approaches used, including the mathematical framework, assumptions, and key parameters.
+**Response:**  To address your concern, we have revised Section 3.2 to provide a more detailed description of each of the forecasting approaches used.
 While we agree that it is important to provide adequate information in the paper itself, we also understand that some readers may be interested in more technical details and code implementation.
-For this reason, we have provided a Github repository with all technical details and R codes for the forecasting approaches used in this study.
+For this reason, we have provided a Github repository containing the entire paper written in Quarto with all technical details and R codes for the forecasting approaches used in this study.
 We hope that this revised section, along with the Github repository, will provide readers with the necessary information to understand and replicate the methods used in this study.
+
+Please see our response to comments 4.1, 4.2 and 4.3 below, where we included further description of the methods and added the related texts to the revised manuscript.
 
@@
 Instead, we aim to provide a methodology that can incorporate forecasts produced by any method.
-Our focus is on developing a hierarchical time series forecasting approach that takes into account the hierarchical and group structure of ambulance demand.
+Our focus is on developing a hierarchical time series forecasting approach that takes into account the hierarchical and group structure of data.
 We have selected the naive method as a benchmark because it is widely used in the forecasting literature and practice and is a simple and easily understandable method for users.
@@
 
+\textcolor{teal}{We have chosen the naive method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
+Additionally, the naive method has shown surprisingly high accuracy.
+Hence, any forecasting approach that can offer superior results compared to the naive method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
+
 _Comment 4.2: In your discussion of ETS, you mention that the trend, seasonality, and error terms can take various forms, such as additive, multiplicative, or mixed.
@@
 
-**Response:** In our research, we used an automatic algorithm to determine the appropriate form for the trend, seasonality, and error terms in each time series.
-Specifically, we used the `ETS()` function in the fable package of R, which selects the most appropriate model for each time series based on statistical criteria such as Akaike's Information Criterion (AIC).
-Since we had a large number of time series (1530), it was not practical to manually select the appropriate form for each component for each time series.
-Instead, the automatic algorithm selects the most appropriate components based on the characteristics of each individual time series.
-As a result, a mix of different component forms (additive or multiplicative) were used, depending on the specific characteristics of each time series.
+**Response:** In response to your comment, we have included the following text in the revised version of the manuscript:
+
+\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable form for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
+Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
 
@@
 
-**Response:** Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as
+**Response:** Thank you for your comment. We have included the following text in revised manuscript to address your comment:
+
+\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
+
 \begin{align*}
@@
 \end{align*}
-and $\bm{x}_t$ is a vector of covariates.
-In our model, these include cubic splines for the time trend, day-of-week dummy variables, Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays, school holidays, Christmas Day and New Year's Day.
-The Fourier terms are as defined in Hyndman & Athanasopoulos (2021, Section 7.4).
 
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+The Fourier terms are as defined in} @hyndman2021forecasting \textcolor{teal}{(Section 7.4).
 This model takes account of weekly seasonality and annual seasonality.
 Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
-There is no reason, for example, for incidents to occur more at some times of the month than others.
+There is no reason, for example, for incidents to occur more at some times of the month than others.}
 
@@
 
+We have now included the following text in revised manuscript:
+
+\textcolor{teal}{The Poisson TSGLM is similar to GLM with an aditional component accounting for serial dependence (i.e.
+autoregressive terms).
+The term serial dependence refers to situations where the ambulance demand at a given time exhibits correlation with the demand at previous time intervals.
+Put simply, past values of the time series influence the current values.
+TSGLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= (\bm{y}_{t-k}'+\bm{x}_t')\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{${y}_{t-k}$ is a vector of k lagged values.
+The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
+past values) of the ambulance demand in the model.
+This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}
+
+\textcolor{teal}{We use the `tsglm()` function in the tscount package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
+Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
+
 _Comment 4.5: It is clear that the naive method performed poorly in the results.
@@
 
-**Response:** We re-ran the results of our experience with the naive method excluded and .....
+**Response:** We re-ran the results of our experiment with the naive method excluded and name it "Ensemble 2".
+We illustrated the results in the following plot.
+As the plot shows, excluding the naive from the ensemble deteriorates the performance overall, so we keep it in the ensemble and show the original plot in the revised manuscript.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-accuracy
+#| fig-width: 8
+#| fig-height: 5
+#| out.width: "100%"
+library(knitr)
+library(kableExtra)
+library(ggthemes)
+library(patchwork)
+library(tidyverse)
+rmsse <- readr::read_rds(here::here("results/rmsse.rds")) |> 
+  mutate(msse = rmsse^2)
+mase <- readr::read_rds(here::here("results/mase.rds"))
+crps <- readr::read_rds(here::here("results/crps.rds")) |>
+  pivot_wider(names_from = model, values_from = crps) |>
+  group_by(method, h, series) |>
+  mutate(across(where(is.numeric), ~ .x / naiveecdf)) |>
+  ungroup()
+crps <- crps |>
+  pivot_longer(ensemble:tscount, names_to = "model", values_to = "crps")
+
+accuracy <- bind_rows(
+  rmsse |> mutate(measure = "msse", accuracy = rmsse^2),
+  mase |> mutate(measure = "mase", accuracy = mase),
+  crps |> mutate(measure = "crps", accuracy = crps),
+) |>
+  select(-rmsse, -mase, -crps)
+
+# Plot of average accuracy vs week for each method for Total
+acc_summary <- accuracy |>
+  filter(
+    series %in% c("Total", "Control areas", "Health boards"),
+    method == "mint", model != "qcomb"
+  ) |>
+  mutate(model = case_when(
+    model == "naiveecdf" ~ "Naive",
+    model == "ets" ~  "ETS",
+    model == "tscount" ~ "TSGLM",
+    model == "iglm" ~ "GLM",
+    model == "ensemble" ~ "Ensemble",
+    model == "ensemble2" ~ "Ensemble2"
+  )) |> 
+  mutate(
+    measure = factor(measure, levels = c("mase","msse","crps"), labels = c("MASE","MSSE","CRPS")),
+    series = factor(series, levels = c("Total", "Control areas", "Health boards")),
+    model = factor(model, levels = c("Naive", "ETS", "TSGLM", "GLM", "Ensemble", "Ensemble2")),
+    week = factor(trunc((h - 1) / 7) + 1)
+  ) |>
+  group_by(week, model, measure, series) |>
+  summarise(accuracy = mean(accuracy), .groups = "drop")
+
+acc_summary |>
+  ggplot(aes(x = week, y = accuracy, group = model, col = model)) +
+  geom_line() +
+  geom_point(size = .5) +
+  facet_grid(measure ~ series, scales = "free_y") +
+  labs(y = "Average accuracy", x = "Week ahead") +
+  ggthemes::scale_color_colorblind() +
+  ggthemes::theme_few()
+```
 
@@
 
-**Response:** In many machine learning applications, datasets are divided into three sets: training, validation, and test sets, and the validation set is used to tune the hyperparameters of the model.
-But we use a more efficient approach than this, whereby we employ time series cross-validation (Hyndman & Athanasopoulos, 2021, Section 5.10) to assess the forecast accuracy, and thus the use of a separate "validation dataset" becomes unnecessary.
+**Response:** In machine learning applications, datasets might be divided into three sets: training, validation, and test sets, and the validation set is used to tune the hyperparameters of the model.
+But we use a more efficient approach than this, whereby we employ time series cross-validation [@hyndman2021forecasting,Section 5.10] to assess the forecast accuracy, and thus the use of a separate "validation dataset" becomes unnecessary.
 Time series cross-validation involves dividing the data into multiple training and testing sets based on the time series structure.
@@
 
+\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by j, representing the number of time periods ahead we are predicting.
+In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
+
+\textcolor{teal}{Using of scale-independent measures, such as MASE and MSSE, enables more significant comparisons between time series at different levels and scales, as these measures are not influenced by the data's magnitude.
+This aspect holds particular importance in our study, as we work with time series at various levels of hierarchy.
+These time series show varying scales, resulting in different magnitudes of error.
+By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
+
 _Comment 7: In the last paragraph of page 13, the Continuous Ranked Probability Score (CRPS) is mentioned as a metric that rewards sharpness and penalizes miscalibration.
@@
 
-To generate forecast probability distributions, we did not assume a specific distribution; instead we used a form of bootstrapping, described in Panagiotelis et al (EJOR, 2023).
+To generate forecast probability distributions, we did not assume a specific distribution; instead we used a form of bootstrapping, described in @panagiotelis2023probabilistic.
 This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
@@
 
+We have added the following text to section 3.2 in the revised version of the paper:
+
+\textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in} @panagiotelis2023probabilistic.
+\textcolor{teal}{This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+In this way, we can generate an empirical distribution of forecasts for each model.
+The ensemble forecast distribution is a simple mixture of these empirical distributions.}
+
+\textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+It measures how well the predicted probabilities match the actual.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
+A model is well-calibrated if the predicted probabilities match the actual, and it is sharp if it is confident in its predictions.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.}
+
 _Comment 8: Section 4.2 does not sufficiently relate to the current problem at hand.
@@
 
-**Response:** We agree that further explanation and contextualization of the definitions presented in Section 4.2 would be beneficial for readers.
-In particular, we will provide a concrete example from our dataset to illustrate the concept of "bottom-level series" and how it applies to our problem.
-We will also revise the text to provide more context and explanation for each definition, to better relate them to the problem at hand.
+**Response:** We agree that further explanation of the definitions presented in Section 4.2 would be beneficial for readers.
+In particular, we will provide an example from our dataset to illustrate the concept of "bottom-level series" and how it applies to our problem.
+We have ioncluded the following text in the revised version of the manuscript to address this comment:.
+
+\textcolor{teal}{The term "level bottom series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
+For instance, in Table 2, each distinct combination of values in Control area (e.g.
+South \& East), Health board (e.g.
+CV), Priority (e.g.
+Green), and Nature of incident (e.g.
+Chest pain) corresponds to one individual time series.
+In the dataset at hand, there are 691 unique combinations, resulting in 691 bottom level time series.
+The "aggregate time series" describes how these bottom-level series are combined to create higher-level series.
+For instance, to obtain the incidents at the national level (i.e.
+all country level), the time series are aggregated across all Control areas, Health boards, Priorities, and Natures of incidents.
+Any desired aggregation level can be achieved based on the data structure, utilizing the bottom level series available.}
 
@@
 We hope this clarifies the difference between the two approaches and how they relate to each other in the context of the paper.
+There are also different approaches in the linear reconciliation method, we have now updated section 4.4 and included the following paragraphs to address your cocnerns:
+
+\textcolor{teal}{The bottom-up (BU) approach is constrained by its reliance solely on base forecasts from a single level of aggregation at the bottom level.
+While it does result in consistent forecasts, the BU approach lacks forecast reconciliation since no reconciliation is performed.}
+
+\textcolor{teal}{Forecast reconciliation approaches bridge this gap by combining and reconciling all base forecasts to generate coherent forecasts.
+This technique utilizes all the base forecasts produced within a hierarchical structure to create consistent forecasts at every level of the hierarchy.
+As a result, it goes beyond relying solely on base forecasts from a single level of aggregation, and instead leverages all available information at each level to generate forecasts that minimize the total forecast variance of the set of coherent forecasts.}
+
+\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
+In this approach, the estimation of W is based on the assumption that all the errors have equal variance.
+Hence, W is simply defined as the identity matrix multiplied by a constant factor.
+The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
+The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
+A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}
+
+\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
+But it assumes that the errors of each series are uncorrelated with each other.
+In this approach, W is defined as a diagonal matrix with the variance of the errors on the diagonal.
+The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
+The main weakness of this approach is that it ignores the relationships between series.
+A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.}
+
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.} @WicEtAl2019 \textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
+The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
+In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}
 
@@
 
-**Response:**
+**Response:** Thank you for the comment.
 
-The forecast reconciliation approach combines base forecasts (not only lower level forecasts).
-It is important to note that one of the main purposes of comparing base forecasts to the reconciliation of those forecasts is to assess the improvement in forecast accuracy that can be achieved by using the forecast reconciliation approach.
+The forecast reconciliation approach does not aggregate lower level forecasts, instead it combines all base forecasts.
 The base forecasts are generated independently at each level of the hierarchy, ignoring the hierarchical/grouped structure of time series or any aggregation constraints.
@@
 In such cases, the benefit of performing reconciliation at the bottom level may still lie in creating coherent forecasts that can help align planning across different teams in an organization, improve coordination, and avoid conflicting decisions.
-Additionally, even if the bottom-level series are noisy and lack systematic patterns, reconciliation can still lead to more accurate forecasts at higher levels of the hierarchy by utilizing the information available in the upper-level series.
-Therefore, even if the bottom-level forecasts are not very accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and possibly more accurate forecasts at other levels.
+Additionally, even if the bottom-level series are noisy and lack systematic patterns, reconciliation can still lead to more accurate forecasts at higher levels of the hierarchy by utilizing the information available across the hierarchy.
+Therefore, even if the bottom-level forecasts might not be very accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and possibly more accurate forecasts at other levels.
+
+We have added the following paragraph in the revised version of the manuscript to address this comment:
+
+\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness may not always be guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
+Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy.
+Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}
 
@@
 
-**Response:** Rounding forecasts up or down and its effect on the forecast accuracy depends on the level of hierarchy and scale of data.
+**Response:** Thank you for highlighting this point, we have now included a paragraph at the end of the result section to acknowledge this point.
+
+Rounding forecasts up or down and its effect on the forecast accuracy depends on the level of hierarchy and scale of data.
 For some situations, with high volume demand, forecast accuracy calculations can ignore integer effects as rounding becomes negligible.
 However, low volume demand settings, such as forecasts at the bottom level of the hierarchy, may be more susceptible to integer (rounding) effects.
-In our study, we kept the non-integer forecast values after reconciliation when evaluating the accuracy.
-The difference in forecast accuracy results is negiglible.
-[Perhaps we should check this assertion?]  In practice, we may need to use integer forecasts, especially when the forecasts are relatively small.
-Count forecast reconciliation is an active area of research (e.g., https://arxiv.org/abs/2207.09322), and it would be interesting to explore in future research how our approach could be adapted to generate count reconciled probabilistic forecasts.
+In practice, we may need to use integer forecasts, especially when the forecasts are relatively small.
+Count forecast reconciliation is an active area of research, and it would be interesting to explore in future research how our approach could be adapted to generate count reconciled probabilistic forecasts.
+
+\textcolor{teal}{Despite using Poisson regression models to create count distributions of attended incidents for the base forecasts, it is important to note that the reconciled forecast distributions do not maintain a count format.
+In practical scenarios, there might be a need to use integer forecasts.
+Count forecast reconciliation is an active area of research, and it would be interesting to explore how our approach could be adapted to generate count-reconciled probabilistic forecasts in future studies.
+One possible approach to address this is by rounding the forecasts.
+However, the impact of rounding on forecast accuracy varies depending on the level of hierarchy and the scale of the data.
+In situations with high-volume demand, the effects of rounding may become negligible, and forecast accuracy calculations can overlook integer effects.
+On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.}
 
@@
 The same reasoning applies to the other two combinations in Table 1.
+we have added a note to Table 1 to explain this further.
+Please refer to Table 1 in the revised paper.
 
@@
 
-**Response:** We appreciate your suggestion to improve the structure of the paper.
-We have moved the overview of base forecasts to a separate subsection before Section 4 to provide a clearer delineation between the independent and hierarchical/grouped forecasts.
-This can help readers better understand the logic structure of the paper and facilitate easier navigation of the content.
+**Response:** Thank you for your valuable suggestion to enhance the paper's structure.
+We have thoroughly considered your proposal of moving the base forecast to a new section.
+However, after careful consideration, we believe that the current structure maintains a smooth and easily understandable flow.
+As it stands, Section 3 provides a detailed discussion of the forecasting models, while Section 4 focuses on using the models described in Section 3 to generate forecasts for the hierarchical structure, including both base and reconciled forecasts.
+We hope this arrangement aligns with your expectations and requirements.
 
@@
 
-**Response:**  In the revised version, we have added more detailed explanations on these methods, including their mathematical formulations, assumptions, and limitations.
-We also provide additional references to the Ensemble method and further explanations on how it is applied in our study.
+**Response:**  In the revised version, we have added more detailed explanations on ETS, Ensemble as well as Naive, GLM and TSGLM. Please refer to section 3.2 to see all details:
+
+We added the following text in the revised version of the manuscript:
+
+\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable form for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
+Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
+
+\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by averaging the forecasts, rather than relying on separate forecasts generated by each individual method} [@clemen1989combining].
+\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models using a simple average to form a mixture distribution} [@combinations].
+
+For the rest of the methods, we added:
+
+\textcolor{teal}{We have chosen the naive method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
+Additionally, the naive method has shown surprisingly high accuracy.
+Hence, any forecasting approach that can offer superior results compared to the naive method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
+
+\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+The Fourier terms are as defined in} @hyndman2021forecasting \textcolor{teal}{(Section 7.4).
+This model takes account of weekly seasonality and annual seasonality.
+Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
+There is no reason, for example, for incidents to occur more at some times of the month than others.}
+
+\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional component accounting for serial correlation (i.e.
+autoregressive).
+The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points.
+Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
+TSGLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= (\bm{y}_{t-k}'+\bm{x}_t')\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{${y}_{t-k}$ is a vector of k lagged values.
+The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
+past values) of the ambulance demand in the model.
+This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}
+
+\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
+Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
 
@@
 
+We added the following text in the revised version of the manuscript to address this comment:
+
+\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable form for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
+Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
+
 _Comment 2-3: GLM and TSGLM are regression approaches, sometimes with exogenous variables like holiday indicators.
@@
 
-**Response:** We have revised the manuscript to include the model formulas, as you suggested.
+**Response:** We have revised the manuscript to include the model formulas, as you suggested for both GLM and TSGLM.
+We have also included the entire publication in Quarto, including R codes, in a GitHub repository for further details:
+
+\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+The Fourier terms are as defined in} @hyndman2021forecasting \textcolor{teal}{(Section 7.4).
+This model takes account of weekly seasonality and annual seasonality.
+Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
+There is no reason, for example, for incidents to occur more at some times of the month than others.}
+
+\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional component accounting for serial correlation (i.e.
+autoregressive).
+The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points.
+Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
+TSGLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= (\bm{y}_{t-k}'+\bm{x}_t')\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{${y}_{t-k}$ is a vector of k lagged values.
+The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
+past values) of the ambulance demand in the model.
+This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}
+
+\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
+Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
 
@@
 Specifically, we have included the paper citation along with the equation number, so that readers can easily locate the relevant equation and read more about its derivation and interpretation.
+Please refer to section 3.3 to see changes, also highlighted below:
+
+\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by j, representing the number of time periods ahead we are predicting.
+In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
+
+\textcolor{teal}{Using of scale-independent measures, such as MASE and MSSE, enables more significant comparisons between time series at different levels and scales, as these measures are not influenced by the data's magnitude.
+This aspect holds particular importance in our study, as we work with time series at various levels of hierarchy.
+These time series show varying scales, resulting in different magnitudes of error.
+By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
+
+\textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+It measures how well the predicted probabilities match the actual.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
+A model is well-calibrated if the predicted probabilities match the actual, and it is sharp if it is confident in its predictions.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.}
 
@@
 
-**Response:** Provided accidents occur independently, they will follow a Poisson distribution by definition (Feller, 3rd ed, 1968, p156-158).
+**Response:** Thank you for your comment.
+Provided accidents occur independently, they will follow a Poisson distribution by definition (@feller1991introduction, p156-158).
 So it is reasonable to assume a Poisson distribution here.
@@
 If there are other factors that affect the mean of the Poisson distribution, but which aren't in our model, we may observe over-dispersion or under-dispersion.
-We have tested for this ??????
-
 We have developed a flexible framework that can incorporate any forecasting model, and it is important to note that the intention of this paper is not to provide a comprehensive list of models, or to argue for a particular model class.
 Instead, our focus is on demonstrating the usefulness of our framework for incorporating base forecasts generated by any model and producing hierarchical coherent forecasts.
+
+To address this comment, we have added the following to section 3 before ensemble method:
+
+\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution} [@feller1991introduction, p156-158].
+\textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
+To account for changes over time, we incorporate trend and seasonality covariates, as well as public holiday effects, allowing the mean of the Poisson distribution to vary.
+However, it is important to note that if there are additional factors influencing the mean of the Poisson distribution, which are not accounted for in our model, we might observe over-dispersion or under-dispersion in the data.}
+
+We have also added the following paragraph to the end of section 3:
+
+\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models or promote a particular model class.
+Instead, we have developed a flexible framework that can accommodate any forecasting models.
+Our primary objective is to demonstrate its practicality and effectiveness in integrating base forecasts from any model and generating coherent forecasts within a hierarchical structure.}
 
@@
 
-**Response:** We used ETS models because they are relatively simple, easy to understand, widely used, and well-established in the time series forecasting literature.
+**Response:** Thank you for the comment.
+We used ETS models because they are relatively simple, easy to understand, widely used, and well-established in the time series forecasting literature.
 They are particularly useful for modeling time series with multiple components, such as trend, seasonality, and/or cycles, and they are flexible and can be adapted to time series data with different patterns.
@@
 Furthermore, we did consider other models specifically designed for integer-valued time series, such as TSGLM and GLM.
+TSGLM is also capable of accounting for autoregressive lags (i.e.
+previous observations) and the number of lags to beconsidered can be determined by user.
+
 We acknowledge that there are other methods available, such as INAR or INGARCH models.
@@
 
-**Response:** All of the time series in our data set are non-stationary, due to the changing rate of accidents with the time-of-year, day-of-week, and other factors.
+**Response:** Thank you for your comment.
+All of the time series in our data set are non-stationary, due to the changing rate of accidents with the time-of-year, day-of-week, and other factors.
 All our models, other than Naive, allow for these non-stationary features of the data.
@@
 
-**Response:** We do not understand this question. The data are not stationary.
+**Response:** Thank you for your comment.
+First of all, we apologize if our understanding of the question is not entirely clear.
+Despite thoroughly discussing it, we acknowledge that we may not have grasped it completely.
+Building on the response to the previous comment, we want to emphasize that the data used in this study exhibit non-stationarity.
+Nevertheless, the models employed possess the capability to handle non-stationarity through various methods, as explained in our response to the previous concern.
 
@@
 
-**Response:** The performance of the distributional forecasts can be evaluated using the CRPS, which is a scoring rule that accounts for the shape and width of the predictive distribution.
+**Response:** Thank you for the comment.
+
+The performance of the distributional forecasts can be evaluated using the CRPS, which is a scoring rule that accounts for the shape and width of the predictive distribution.
 The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
 Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.
+A specific prediction interval could be evaluated using a Winkler score [@winkler1972decision].
+CRPS can be considered an average of all possible Winkler scores [@hyndman2021forecasting, Section 5.9], and thus provides an evaluation of all possible prediction intervals.
 
-A specific prediction interval could be evaluated using a Winkler score (Winkler, JASA, 1992).
-CRPS can be considered an average of all possible Winkler scores [@fpp3, Section 5.9], and thus provides an evaluation of all possible prediction intervals.
+We have clarified this further in the revised manuscript:
+
+\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores} [@winkler1972decision;@hyndman2021forecasting, Section 5.9] \textcolor{teal}{or percentile scores} [@hyndman2021forecasting, section 5.9], \textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.} \textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
+Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.
+95\%) quantiles.
+In such cases, a percentile score  becomes useful in meeting this specific requirement.}
 
@@
 
-**Response:** In response to your question, we have included an additional visualization to demonstrate the tails of the forecast distribution for a given time series.
-While the center of the distribution may represent the most likely scenario, it may not necessarily capture the full range of potential outcomes that can affect decision-making at EMS.
-Therefore, a better understanding of the tails of the distribution can provide more valuable insights for managers to better prepare for and mitigate risks.
-Therefore, the accuracy of the tail forecasts can be critical for emergency planning and response.
+**Response:** Thank you for this valuable question.
+While the central part of this distribution may indicate the most probable scenario, it might not include the entire spectrum of potential outcomes that could significantly impact EMS decision-making.
+Thus, gaining a deeper understanding of the distribution's tails can offer valuable insights to managers, enabling them to better prepare for and manage risks effectively.
+Ensuring the accuracy of these tail forecasts becomes crucial in emergency planning and response, as it plays a pivotal role in anticipating and addressing critical situations.
+In our research, we use CRPS (Continuous Ranked Probability Score) as a metric to assess the accuracy of probabilistic forecasts.
+CRPS provides an average measure across all quantiles, including both low and high quantiles.
+However, in some cases, there might be a need to calculate the accuracy for a specific quantile only like 5\% or 95\%.
+In such situations, a percentile score can be used to fulfill this requirement.
+
+As a response to your comment, we have incorporated an extra subsection (section 5.1) that includes a visualization representing the forecast distribution for a specific time series of incidents.
+We have added the following new section to the revised manuscript:
+
+```{r}
+#| label: fig-forecast-density
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8
+#| fig-height: 5
+#| out.width: "100%"
+#| fig-cap: "An example of forecast distribution of ambulance demand (i.e.
+total incidence attended) in PO health board for 1 day ahead.
+x-axis shows all possible outcomes that may occur, with thier likelihhod in y-aixs"
+library(tidyverse)
+library(scales)
+library(fpp3)
+library(ggridges)
+incident_gthf <- readr::read_rds(here::here("data/incidents_gt.rds"))
+
+incident_cv <- incident_gthf |> filter(lhb == "PO",is_aggregated(category),is_aggregated(nature))
+fit <- incident_cv %>%
+  model(
+    ets = ETS(sqrt(incident))
+  )
+set.seed(1)
+fcst <- fit %>% generate(h=1, times = 5000, bootstrap = TRUE)
+p1 <- ggplot(data=fcst,aes(x=.sim, y=..density..)) +
+  geom_histogram(binwidth = 1, fill="#d3d3d3", color="black")+
+  geom_vline(data=fcst,aes(xintercept = (mean(.sim))), color="black", linetype="dashed", linewidth=1)+
+  scale_x_continuous(breaks = seq(25,70,5))+
+  scale_y_continuous(labels=percent)+
+  ggthemes::theme_few()+
+    labs(y="Density", x = "Demand", title = "Probabilistic forecast of ambulance demand")
+p1
+```
+
+@fig-forecast-density \textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, regarding the total attended incidents in the PO health board.
+The black dashed line in the graph represents the point forecast, which presents a single numerical estimate for the future number of attended incidents.
+Due to the complexity of including such plots for the entire hierarchy and 84 days ahead, only one example is presented here for 1 day ahead.
+However, it is feasible to generate these plots for the entire hierarchy and for any forecast horizon if necessary.}
+
+\textcolor{teal}{In practice, point forecasts are commonly used, but they have limitations as they ignore the uncertainty associated with the forecast, leading to wrong decisions.
+In contrast, probabilistic forecasts offer an alternative approach to anticipate future demand.
+Rather than providing a single value, they assign likelihoods to all possible demand outcomes, acknowledging that different numbers of attended incidents are possible, but with varying likelihoods.}
+
+\textcolor{teal}{The purpose of probabilistic forecasting, as demonstrated in} @fig-forecast-density, \textcolor{teal}{is to quantify uncertainty.
+Decisions based on these forecasts could focus on the tails of the distribution: unexpected high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
+Such forecasts are valuable tools for decision-makers and planners, especially when dealing with low-probability, high-cost situations.
+Different EMS managements may have varying risk attitudes depending on resource availability, making it crucial to consider the entire distribution when making decisions.
+For instance, these forecasts enable management to calculate the probability of demand exceeding a certain threshold of available resources (e.g., 90\%), which can serve as an informative early warning measure for crowdedness.}
+
+\textcolor{teal}{It is important to note that while point forecasts and prediction intervals can be obtained from the probabilistic forecasts, the reverse is not possible.
+A single number cannot be used to directly derive a probabilistic forecast.
+Prediction intervals, although helpful in indicating possible ranges, do not provide information on the probabilities of low or high demand.}
+
+\textcolor{teal}{In EMS planning, future demand is just one aspect to consider.
+Other inputs, such as capacity, should also be treated as probability distributions to adopt a probabilistic approach to planning.
+To extract valuable insights and make informed decisions from probabilistic forecasts, specialized numerical tools are required, as the forecasts themselves are typically represented as explicit probability density functions or Monte Carlo generators.}
+
+We have also included the following paragraph in section 3.3:
+
+\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores} [@winkler1972decision;@hyndman2021forecasting, Section 5.9] \textcolor{teal}{or percentile scores} [@hyndman2021forecasting, section 5.9], \textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.} \textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
+Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.
+95\%) quantiles.
+In such cases, a percentile score  becomes useful in meeting this specific requirement.}
 
@@
 **Response:** We appreciate your suggestion of adding supplementary material to provide more technical details and elaboration on some ideas.
-To respond to your queries, we have included the GitHub repository which includes all materials including data and R codes used in the paper.
-Additionally, we have included further explanation in the revisions of the paper.
+To respond to your queries, we have included the GitHub repository which includes all files use to produce the paper including Quarto files, data and R codes used to generate forecasts, evaluate accuracy and produce plots and tables.
+Additionally, we have included further explanation in the revisions of the paper highlighted them in teal.
 
@@
 
-**Response:**  In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by Kang et al (2017).
+**Response:** We have added the following text to the revised manuscript to discuss how the strengths of trend and weekly seasonality are estimated :
+
+\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by Kang et al (2017).
 STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
 The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
-The strength of trend is defined as:
+The strength of trend is defined as:}
 $$F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$
-For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
+\textcolor{teal}{For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
 Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small.
-But for data with little or no trend, the two variances should be approximately the same.
+But for data with little or no trend, the two variances should be approximately the same.}
 
-The strength of seasonality is defined similarly:
+\textcolor{teal}{The strength of seasonality is defined similarly:}
 $$F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).$$
- series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).
+\textcolor{teal}{series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).}
 
@@
 
-**Response:**  The aggregation matrix $A$ is determined by the structure of the hierarchy.
+**Response:**  We have included the following paragraph in section 4.3 to address this comment:
+
+\textcolor{teal}{The aggregation matrix $A$ is determined by the structure of the hierarchy.
 It maps the bottom-level time series to the corresponding higher-level time series.
@@
 For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $A = \begin{bmatrix} 1 1 \end{bmatrix}$.
+}
 
@@
 
-**Response:** Linear reconciliation involves projecting the base forecasts onto the coherent space.
-It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased (Wickramasuriya et al, JASA 2019).
+**Response:** We have added the following text to section 4.4 to further clarify this:
+
+\textcolor{teal}{Linear reconciliation involves projecting the base forecasts onto the coherent space.
+It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased} [@WicEtAl2019].
 
@@
 
-**Response:**
+**Response:** Thank you for asking.
+We have clarified different solutions (Ordinary Least Squares (OLS), Weighted Least Squares (WLS) and Minimum Trace (MinT)) and how W is defined.
+We have added the following paragraphs to section 4.4 in the revised manuscript:
 
-Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
+\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
 In this approach, the estimation of W is based on the assumption that all the errors have equal variance.
@@
 The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
-A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.
+A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}
 
-Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
+\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
 But it assumes that the errors of each series are uncorrelated with each other.
@@
 The main weakness of this approach is that it ignores the relationships between series.
-A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.
+A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.}
 
-Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
-So it takes account of both the scale of each series, and the relationships between the series.
-Wickramasuriya et al (JASA 2019) showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.} @WicEtAl2019 \textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
 The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
-In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.
+In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}
 
@@
 
-**Response:** We have corrected the text accordingly.
+**Response:** Thank you. We believe that line counts are automatically inserted by the system following the paper's submission.
 
@@
 
-```{r, out.width= "60%"}
+```{r, out.width= "60%", echo=FALSE, fig.align='center'}
 knitr::include_graphics("figure/crps.jpg")
@@
 
-**Response:** Throughout, we ensured that a consistent format was maintained in the References section regarding the inclusion or exclusion of URL links.
+**Response:** Thank you.
+We have removed URL links from the references.
+Throughout, we ensured that a consistent format was maintained in the References section.
+Also, we note that the entire citation and reference management is automatic and handled by Quarto and the format of the journal.
 
@@
 
-**Response:**  We have reviewed Table 2 revised the table to make the column names more clear and consistent with Table 1.
+**Response:**  We have reviewed Table 2, revised the table to make the column names more clear and consistent with Table 1.
 
@@
 
+We have included the following paragraph to further clarify this in the revised manuscript:
+
+\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable form for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
+Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
+
 _Comment 1-b.
@@
 
-**Response:** Indeed, it is possible to define a state space model that ensures coherent forecasts.
-This was proposed by Pennings & Dalen (2017), and Villegas & Pedregal (2018).
+**Response:** Thank you for your comment.
+Indeed, it is possible to define a state space model that ensures coherent forecasts.
+This was proposed by @pennings2017integrated, and @villegas2018supply.
 However, in these models, the covariance matrices are extremely difficult to estimate with anything other than very small hierarchies.
+The hierarchy used in our study is extremely big with 1530 time series.
 The strength of our approach is that it allows for any type of forecasting method to be used for the base forecasts, including different methods for different series.
 It decouples the time series models from the reconciliation step, thus allowing much more flexibility in the types of models that can be considered.
+
+We have added the following paragraph in the revised version of the paper for further clarification:
+
+\textcolor{teal}{Certainly, other approaches can be applied to hierarchical forecasting problems.} @pennings2017integrated and @villegas2018supply \textcolor{teal}{proposed the idea of using a state space model to ensure consistent forecasts.
+However, when dealing with larger hierarchies, these models encounter difficulties in estimating covariance matrices.
+In contrast, our approach provides a clear advantage by allowing the incorporation of different forecasting methods for the base forecasts, and even accommodating distinct methods for individual series.
+The decoupling of time series models from the reconciliation step adds significant flexibility in exploring a wide range of models.}
 
@@
 
-We appreciate your thoughtful comment, and we will take it into account for future research.
+We have added the following paragraphs to the conclusion section to further discuss this crucial point and address the comment:
+
+\textcolor{teal}{Our research establishes a strong basis for future investigations and practical implementation in EMS.
+Leveraging the hierarchical and grouped structure of demand time series, EMS can use this advanced forecasting framework to generate coherent point and probabilistic forecasts, making the most of all available data at every level of the hierarchy.
+We acknowledge that a forecast serves a greater purpose beyond its mere existence, ideally enabling the best utility in terms of efficient allocation of medical services, response time, and cost, all informed by the forecast.
+While we fully appreciate the importance of evaluating forecast quality based on its impact on decision-making processes, it is essential to address the data requirements and methodology involved in measuring this impact.
+For a comprehensive assessment of the forecasts' implications, access to additional data beyond ambulance demand, covering various decision types, capacity information, constraints in the decision system, and more, becomes necessary.
+This additional data would offer valuable insights into the specific decisions relying on the forecasts, resulting in a more accurate evaluation of their impact on medical services.
+Furthermore, measuring the actual impact of forecasts would necessitate an approach that goes beyond forecasting itself.
+This would involve developing and implementing simulation models capable of replicating decision-making processes based on the forecast inputs.
+These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.}
+
+\textcolor{teal}{Future research can build upon this study by incorporating additional data and implementing simulation modeling techniques.
+In future investigations, we aim to explore this avenue by incorporating operational information, simulating decision processes, and assessing the impact on utilities that are significant to the EMS.
+Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.}
 
@@
 Therefore, it is crucial to assess both calibration and sharpness of forecast to ensure effective decision-making within emergency medical services.
+
+We have added the following paragraph to section 3.3 to furthe clarify this:
+
+\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores} [@winkler1972decision;@hyndman2021forecasting, Section 5.9] \textcolor{teal}{or percentile scores} [@hyndman2021forecasting, section 5.9], \textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.} \textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
+Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.
+95\%) quantiles.
+In such cases, a percentile score  becomes useful in meeting this specific requirement.}
 
@@
 Neelamegham, R., & Chintagunta, P. K. (2004). Modeling and forecasting the sales of technology products. Quantitative Marketing and Economics, 2(3), 195.
+
+We have added the following paragraph to the end of the conclusion section to discuss how our approach is generalisable across different service industries:
+
+\textcolor{teal}{Although our study primarily focuses on Emergency Medical Services, it is essential to emphasize that the framework we propose has broad applicability across various service industries} [@ostrom2010moving].
+\textcolor{teal}{Our approach is particularly valuable in situations where time series data is structured hierarchically and/or grouped, a common characteristic found in many sectors.
+This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
+For instance, in the supply chains} [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
+Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency.
+In the financial industry } [@kimes1998strategic] \textcolor{teal}{, where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
+Similarly, in transportation, the framework supports forecasting transportation demand at various levels, optimizing route planning and resource allocation.
+Likewise, in the hospitality and tourism industry} [@dekimpe2016impact], \textcolor{teal}{it facilitates forecasting demand rates at state, regional, and department levels, enabling strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses.
+Additionally, in call centers, accurate call volume forecasting at different levels of the call center hierarchy or grouped structure is crucial for workforce management and resource allocation.
+Implementing our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, ensuring efficient staffing and optimal service levels to meet customer demands.}
+
+# References {#references .unnumbered}

### 2023-07-18 manuscript/01-introduction.qmd
response letter and paper ready for submission

@@
 Demand forecasts at EMS are typically required at multiple levels of an organization to inform various planning and decision-making processes [@hulshof2012taxonomic].
-There are some planning processes at the national level (strategic and long-term) such as workforce resource planning and budgeting; sub-national, regional, or healthcare level (tactical and medium-term) such as temporary capacity expansions, resource sharing, and staff-shift scheduling; and hospital or station level (operational and short-term) such as planning rosters for staff and ambulance deployment.
+There are some planning processes at the national level (strategic and long-term) such as workforce resource planning and budgeting; sub-national, regional, or healthcare level (tactical and medium-term) such as temporary capacity expansions, resource sharing; and hospital or station level (operational and short-term) such as planning rosters for staff and ambulance deployment.
 Demand forecasts might also be required at different levels for a specific area of interest such as the nature of demand or the priority level.
 Moreover, the time series data in EMS has an inherent hierarchical and grouped structure to support such forecasting requirements.
-Demand for emergency medical services at the country level can be disaggregated in a geographical hierarchy into sub-national, regions, health boards, and stations/hospitals, or divided into groups such as the nature of incidents or demand priority.
-Therefore, using forecasting methodologies that account for hierarchical and/or grouped structures of time series in EMS is a natural fit.
+Demand for emergency medical services at the national level can be disaggregated in a geographical hierarchy into sub-national, regions, health boards, and stations/hospitals, or divided into groups such as the nature of incidents or demand priority.
+\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchies are necessary for effective decision-making in EMS.
+For example, control area EMS forecasts can inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals.
+At the lower levels, hospitals or ambulance stations could use such forecasts to plan for staffing and resource allocation, ambulance dispatching, staff-to-shift assignment, staff rescheduling based on the anticipated volume and priority and nature of incidents.
+Additionally, generating forecasts at lower levels could potentially improve the accuracy of the high-level forecasts, by providing more detailed information on the nature and priority of incidents.
+This could help to identify patterns in demand that may not be apparent at the higher level.
+Therefore, employing forecasting techniques that consider the hierarchical and/or grouped patterns of time series in EMS aligns naturally, offering the possibility to enhance forecast accuracy and facilitate coordination.}
 
@@
 Generating independent forecasts can result in a lack of consistency and coordination, and therefore leads to less effective planning and decision making.
+\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
+If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area.
+Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others.
+Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly.
+Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}
+
 With hierarchical forecasting, plans at any level are based on coherent forecasts and therefore can be aligned.
@@
 
-In this paper, we address this gap by investigating the application of hierarchical forecasting approaches in the EMS using daily time series of attended incidents from 2015 to 2020 in a major ambulance service in Great Britain.
+In this paper, we address this gap by investigating the application of hierarchical forecasting approaches in the EMS using daily time series of attended incidents from 2015 to 2019 in a major ambulance service in Great Britain.
 The data has hierarchical and grouped structures, with hierarchies at the national, control (i.e.

### 2023-07-18 manuscript/02-literature.qmd
response letter and paper ready for submission

@@
 An effective service delivery requires accurate resource planning that generally relies on demand forecasts at operational, tactical, and strategic levels.
-
 There is a substantial number of studies on the application of time series forecasting in the Emergency Medical Services.
@@
 The tool contains options that allow generating forecasts at various levels of granularity; however, it ignores the structure of the hierarchical and grouped relationships, preventing aligned decision making and coordination.
+@al2021empirical \textcolor{teal}{utilises real data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand accurately.
+The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
+Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE) are calculated.
+The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
+However, when it comes to long-term demand, the SSA method proves to be the most effective.} @ibrahim2016modeling \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
+Real-life call centre data was utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
+The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
+The findings indicate that the ME consistently produces the most accurate point forecasts.
+On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
+For a two-week leading period, MU shows better coverage probability.}
 
@@
 @zhou2016predictinglit proposed three methods based on Gaussian mixture models, kernel density estimation, and kernel warping to predict hourly data 4 weeks ahead for a 1km$^2$ spatial region.
-Two years of incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia (years 2011 and 2012 with 696,975 events) were used to build the model and examine the performance on test data using mean negative log-likelihood.
+Two years of incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events) and Melbourne, Australia were used to build the model and examine the performance on test data using mean negative log-likelihood.
 They show that forecasts generated by the proposed methods were significantly more accurate than the current industry practice (a simple averaging formula).
@@
 Current studies ignore the inherent hierarchical and/or grouped structure of the time series data, and the relationship between series at different levels of hierarchy.
-This may result in incoherent forecasts leading to misaligned planning and decision making.
 While the hierarchical forecasting methodology has been developed and applied in various domains over the past 10 years [@panagiotelis2022probabilistic], it has never been explored in this area.
@@
 Current research is mainly concerned with generating point forecasts at a single level of hierarchy.
-There is a lack of studies considering the entire forecast distribution of daily ambulance demand for the whole hierarchy to inform the decision-making process and to better represent the uncertainty of future demand, providing a risk management tool for planners.
+There is a lack of studies considering the entire forecast distribution of daily ambulance demand for the whole hierarchy to better represent the uncertainty of future demand, providing a risk management tool for planners.
 
@@
 
-This paper concerns the problem of hierarchical forecasting in EMS and generates and evaluates both point and probabilistic forecast across different levels of the hierarchy, hence addressing some important gaps identified in the literature.
+```{r summarylit, echo=FALSE, cache=FALSE}
+library(kableExtra)
+li <- readxl::read_xlsx("../literature_table/table.xlsx")
+knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies in forecasting FOR AMBULANCE SERVICES") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
+  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |>  kableExtra::landscape()
+```

### 2023-07-18 manuscript/03-experiment.qmd
response letter and paper ready for submission

@@
   "Priority", 3,
-  "Priority * Control", 9,
-  "Priority * Health board", 21,
+  "Priority * Control", 3*3,
+  "Priority * Health board", 3*7,
   "Nature of incident", 35,
-  "Nature of incident * Control", 105,
-  "Nature of incident * Health board", 245,
-  "Priority * Nature of incident", 104,
+  "Nature of incident * Control", 3*35,
+  "Nature of incident * Health board", 7*35,
+  "Priority * Nature of incident", 3*35-1,
   "Control * Priority * Nature of incident", 306,
@@
 knitr::kable(agg_level, booktabs = TRUE, position = "left") %>%
-  kable_classic(full_width = FALSE)
+  kable_classic(full_width = FALSE) |> footnote(general = "Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset.
+As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.",threeparttable = TRUE)
 ```
@@
 Both measures are on a scale of [0,1].
+
+\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by} @kang2017visualising.
+\textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
+The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
+The strength of trend is defined as:}
+$$F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$
+\textcolor{teal}{For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
+Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small.
+But for data with little or no trend, the two variances should be approximately the same.}
+
+\textcolor{teal}{The strength of seasonality is defined similarly:}
+$$F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).$$
+\textcolor{teal}{series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).}
+
 It is clear that there are some series showing strong trends and/or seasonality, corresponding to series at the higher levels of the hierarchy.
@@
 #| fig.align: center
-#| fig-cap: "Time series features of attended incidents across all levels (1530 series)"
+#| fig-cap: "The strength of the trend and weekly seasonality in the time series of attended incidents.
+The scatter plot shows a total of 1530 data points, with each point corresponding to a specific time series."
 incident_gthf <- readr::read_rds(here::here("data/incidents_gt.rds"))
@@
 #| fig-height: 10
-#| fig-cap: "Time series of attended incidents at various levels.
+#| fig-cap: "Daily time plot of attended incidents at various levels.
+X-axis shows the date of incidents, consisting of 1400 data points (days) and y-axis shows the number of attended incidents.
 The panels show data from the whole country (top panel), by control area, by health board, by priority level, and by nature of incident.
-Only four of the 35 nature of incident categories are shown to avoid too much overplotting."
+Only four of the 35 nature of incident categories are shown to avoid too much overplotting.
+Each time plot .
+"
 no_x_axis <- theme(
@@
 We use the empirical distribution of the past daily attended incidents to create the forecast distribution of future attended incidents.
+\textcolor{teal}{We have chosen the naive method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
+Additionally, the naive method has shown surprisingly high accuracy.
+Hence, any forecasting approach that can offer superior results compared to the naive method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
 
@@
 To identify the best model for a given time series, the `ets` function uses the corrected Akaike’s Information Criterion (AICc).
+
+\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
+Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
 
@@
 One of the models that is frequently used in practice to generate count forecasts is Poisson regression.
-We will consider forecasting attended incidents using the covariates spline trend, day of the week dummy variables (from Monday to Sunday), Fourier terms to capture yearly seasonality, public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+
+\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+The Fourier terms are as defined in} @hyndman2021forecasting \textcolor{teal}{(Section 7.4).
+This model takes account of weekly seasonality and annual seasonality.
+Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
+There is no reason, for example, for incidents to occur more at some times of the month than others.}
+
 We fit a Poisson regression model using the function `glm()` from the *stats* package in R, with the argument `family = poisson` to specify that we wish to fit a Poisson regression model with a log link function.
 
-**Poisson Regression using tscount (TSGLM):** We also consider another Poisson regression model that takes into account serial dependence, using the `tsglm()` function in the _tscount_ package in R [@JSSv082i05] to model the attended incidents.
-Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.
-This model captures the short range serial dependence by including three autoregressive terms, in addition to the same covariates that were used in the GLM model.
-To distinguish this from the previous GLM model, we will refer to this model as "TSGLM".
+**Poisson Regression using tscount (TSGLM):** We also consider another Poisson regression model that takes into account serial dependence.
+This model captures the short range serial dependence by including autoregressive terms, in addition to the same covariates that were used in the GLM model.
+To distinguish this from the previous GLM model, we will refer to this model as `TSGLM`.
 
-**Ensemble method:** Finally, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models to form a mixture distribution [@combinations].
+\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional component accounting for serial correlation (i.e.
+autoregressive).
+The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points.
+Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
+TSGLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= (\bm{y}_{t-k}'+\bm{x}_t')\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{${y}_{t-k}$ is a vector of k lagged values.
+The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
+past values) of the ambulance demand in the model.
+This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}
+
+\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
+Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
+
+\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution} [@feller1991introduction, p156-158].
+\textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
+To account for changes over time, we incorporate trend and seasonality covariates, as well as public holiday effects, allowing the mean of the Poisson distribution to vary.
+However, it is important to note that if there are additional factors influencing the mean of the Poisson distribution, which are not accounted for in our model, we might observe over-dispersion or under-dispersion in the data.}
+
+**Ensemble method:** \textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by averaging the forecasts, rather than relying on separate forecasts generated by each individual method} [@clemen1989combining].
+\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models using a simple average to form a mixture distribution} [@combinations].
+
+\textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in} @panagiotelis2023probabilistic.
+\textcolor{teal}{This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+In this way, we can generate an empirical distribution of forecasts for each model.
+The ensemble forecast distribution is a simple mixture of these empirical distributions.}
+
+\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models or promote a particular model class.
+Instead, we have developed a flexible framework that can accommodate any forecasting models.
+Our primary objective is to demonstrate its practicality and effectiveness in integrating base forecasts from any model and generating coherent forecasts within a hierarchical structure.}
 
@@
 The largest training set ends on 2019-05-09, with the test set ending on 2019-07-31.
-Model development and hyper-parameter tuning is performed using the training data only.
+Model development and hyper-parameter tuning is performed using the training data and the errors are assess using the corresponding test set.
 While we compute forecast errors for the entire 12 weeks, we are most interested in the last 42 days of each test set, because that corresponds to how forecasts are generated for planning in practice.
@@
 
-Point forecast accuracy is measured via the Mean Squared Scaled Error (MSSE) and the Mean Absolute Scaled Error (MASE). The Mean Absolute Scaled Error (MASE) [@HK06] is calculated as:
+\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by j, representing the number of time periods ahead we are predicting.
+In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
+
+Point forecast accuracy is measured via the Mean Squared Scaled Error (MSSE) and the Mean Absolute Scaled Error (MASE).
+The Mean Absolute Scaled Error (MASE) [@HK06; @hyndman2021forecasting] is calculated as:
 $$
@@
 
-A related measure is MSSE [@hyndman2021forecasting], which uses squared errors rather than absolute errors:
+A related measure is MSSE [@hyndman2021forecasting;@makridakis2022m5], which uses squared errors rather than absolute errors:
 $$
@@
 
-To measure the forecast distribution accuracy, we calculate the Continuous Rank Probability Score [@gneiting2014probabilistic].
+\textcolor{teal}{Using of scale-independent measures, such as MASE and MSSE, enables more significant comparisons between time series at different levels and scales, as these measures are not influenced by the data's magnitude.
+This aspect holds particular importance in our study, as we work with time series at various levels of hierarchy.
+These time series show varying scales, resulting in different magnitudes of error.
+By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
+
+To measure the forecast distribution accuracy, we calculate the Continuous Rank Probability Score [@gneiting2014probabilistic;@hyndman2021forecasting].
 It rewards sharpness and penalizes miscalibration, so it measures overall performance of the forecast distribution.
@@
 where $G_j(x)$ is the forecasted probability distribution function for forecast horizon $j$, and $F_j(x)$ is the true probability distribution function for the same period.
+
+\textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+It measures how well the predicted probabilities match the actual.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
+A model is well-calibrated if the predicted probabilities match the actual, and it is sharp if it is confident in its predictions.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.}
+
+\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores} [@winkler1972decision;@hyndman2021forecasting, Section 5.9] \textcolor{teal}{or percentile scores} [@hyndman2021forecasting, section 5.9], \textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.} \textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
+Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.
+95\%) quantiles.
+In such cases, a percentile score  becomes useful in meeting this specific requirement.}

### 2023-07-18 manuscript/04-htsgts.qmd
typos
response letter and paper ready for submission

@@
 
-## Reconciliation methos
+## Reconciliation methods
 
@@
 $$
-and $\bm{A}$ is the $n_a\times n_b$ "aggregation" matrix specifying how the bottom-level series $\bm{b}_t$ are to be aggregated to form $\bm{a}_t$. The full vector of time series is given by
+and $\bm{A}$ is the $n_a\times n_b$ "aggregation" matrix specifying how the bottom-level series $\bm{b}_t$ are to be aggregated to form $\bm{a}_t$.
+\textcolor{teal}{The aggregation matrix $A$ is determined by the structure of the hierarchy.
+It maps the bottom-level time series to the corresponding higher-level time series.
+The matrix $A$ is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
+For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $A = \begin{bmatrix} 1 1 \end{bmatrix}$.
+}
+
+The full vector of time series is given by
 $$
@@
 
+\textcolor{teal}{The term "level bottom series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
+For instance, in Table 2, each distinct combination of values in Control area (e.g.
+South \& East), Health board (e.g.
+CV), Priority (e.g.
+Green), and Nature of incident (e.g.
+Chest pain) corresponds to one individual time series.
+In the dataset at hand, there are 691 unique combinations, resulting in 691 bottom level time series.
+The "aggregate time series" describes how these bottom-level series are combined to create higher-level series.
+For instance, to obtain the incidents at the national level (i.e.
+all country level), the time series are aggregated across all Control areas, Health boards, Priorities, and Natures of incidents.
+Any desired aggregation level can be achieved based on the data structure, utilizing the bottom level series available.}
+
 ### Bottom-up (BU) and linear reconciliation methods
@@
 
-Forecast reconciliation approaches fill this gap by combining and reconciling all the base forecasts in order to produce coherent forecasts.
+The bottom-up (BU) approach is constrained by its reliance solely on base forecasts from a single level of aggregation at the bottom level.
+While it does result in consistent forecasts, the BU approach lacks forecast reconciliation since no reconciliation is performed.
+
+\textcolor{teal}{Forecast reconciliation approaches bridge this gap by combining and reconciling all base forecasts to generate coherent forecasts.
+This technique utilizes all the base forecasts produced within a hierarchical structure to create consistent forecasts at every level of the hierarchy.
+As a result, it goes beyond relying solely on base forecasts from a single level of aggregation, and instead leverages all available information at each level to generate forecasts that minimize the total forecast variance of the set of coherent forecasts.} \textcolor{teal}{Linear reconciliation involves projecting the base forecasts onto the coherent space.
+It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased} [@WicEtAl2019].
+
 Linear reconciliation methods can be written [@WicEtAl2019] as
@@
 Different choices for $\bm{W}$ lead to different solutions such as Ordinary Least Squares (OLS), Weighted Least Squares (WLS) and Minimum Trace (MinT).
+
+\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
+In this approach, the estimation of W is based on the assumption that all the errors have equal variance.
+Hence, W is simply defined as the identity matrix multiplied by a constant factor.
+The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
+The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
+A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}
+
+\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
+But it assumes that the errors of each series are uncorrelated with each other.
+In this approach, W is defined as a diagonal matrix with the variance of the errors on the diagonal.
+The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
+The main weakness of this approach is that it ignores the relationships between series.
+A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.}
+
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.} @WicEtAl2019 \textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
+The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
+In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}
+
 We use the implementation of these methods in the fable package in R in the experiment.
+
+\textcolor{teal}{Certainly, other approaches can be applied to hierarchical forecasting problems} @pennings2017integrated and @villegas2018supply \textcolor{teal}{proposed the idea of using a state space model to ensure consistent forecasts.
+However, when dealing with larger hierarchies, these models encounter difficulties in estimating covariance matrices.
+In contrast, our approach provides a clear advantage by allowing the incorporation of different forecasting methods for the base forecasts, and even accommodating distinct methods for individual series.
+The decoupling of time series models from the reconciliation step adds significant flexibility in exploring a wide range of models.}

### 2023-07-18 manuscript/05-results.qmd
response letter and paper ready for submission

@@
 MinT) improves forecast accuracy at the higher levels of the hierarchy including total, control area and health board.
-However, it is does not result in accuracy improvement at the bottom level series, for which base forecasts are more accurate.
+However, it does not result in accuracy improvement at the bottom level series, for which base forecasts are more accurate.
 This might be due to the noisy structure of time series at the bottom level, and possibly due to very different patterns in the aggregated series.
@@
 However, this does not remain valid for bottom series where different individual methods perform best, depending on the accuracy measure.
+\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness may not always be guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
+Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy.
+Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}
 
@@
 #| tbl-cap: "Average forecast performance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data.
+The testset consists of 462 days.
 The best approach is highlighted in bold."
@@
 
-kbl(mase1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
+kbl(mase1, booktabs = TRUE, escape = FALSE, align="llcccc") |>
   add_header_above(c(" " = 2, "MASE" = 4)) |>
@@
 
-kbl(msse1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
+kbl(msse1, booktabs = TRUE, escape = FALSE, align="llcccc") |>
   add_header_above(c(" " = 2, "MSSE" = 4)) |>
@@
 
-kbl(crps1, booktabs = TRUE, escape = FALSE, align="llrrrr") |>
+kbl(crps1, booktabs = TRUE, escape = FALSE, align="llcccc") |>
   add_header_above(c(" " = 2, "CRPS" = 4)) |>
@@
 
-In addition to the overall forecast accuracy presented in @tbl-result, we also report the point and probabilistic forecast accuracy measures for each forecast horizon in @fig-accuracy.
-The figure focuses on the hierarchical levels important for decision-making including total, control area, and health board; however, the accuracy could be calculated for any level.
-We only illustrate the results of the MinT method, given its strong performance described in @tbl-result.
-For illustration purposes, we report the average weekly forecast accuracy instead of the daily forecast horizon, as this reduces the visual noise in the figure.
-Thus, the x-axis shows horizons from week 1 ($h= 1,\dots,7$) to week 12 ($h= 78,\dots,84$).
-The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision-makers.
-For both the point forecast and distributional accuracy  we can see that the ensemble approach performs best across almost all horizons, with the biggest differences at the highest levels of aggregation.
-It is important to highlight that, all forecasting models outperform the Naive empirical distribution that is used as a benchmark for both point and probabilistic forecasts.
-
 ```{r}
@@
 #| fig-cap: "Average accuracy by week for 12 weeks using MinT reconciliation.
+The total number of days used to calculate the accuracy in the test set is 462.
+Forecasts are generateled every 42 days, therefore we use 11 samples to calculate the average accuracy.
 CRPS is relative to a naive Empirical Cumulative Distribution Function (ECDF).
@@
 ```
+
+In addition to the overall forecast accuracy presented in @tbl-result, we also report the point and probabilistic forecast accuracy measures for each forecast horizon in @fig-accuracy.
+The figure focuses on the hierarchical levels important for decision-making including total, control area, and health board; however, the accuracy could be calculated for any level.
+We only illustrate the results of the MinT method, given its strong performance described in @tbl-result.
+For illustration purposes, we report the average weekly forecast accuracy instead of the daily forecast horizon, as this reduces the visual noise in the figure.
+Thus, the x-axis shows horizons from week 1 ($h= 1,\dots,7$) to week 12 ($h= 78,\dots,84$).
+The forecast horizon from week 7 to week 12 corresponds to the upcoming planning horizon, which is used by planners and decision-makers.
+For both the point forecast and distributional accuracy  we can see that the ensemble approach performs best across almost all horizons, with the biggest differences at the highest levels of aggregation.
+It is important to highlight that, all forecasting models outperform the Naive empirical distribution that is used as a benchmark for both point and probabilistic forecasts.
+
+\textcolor{teal}{Despite using Poisson regression models to create count distributions of attended incidents for the base forecasts, it is important to note that the reconciled forecast distributions do not maintain a count format.
+In practical scenarios, there might be a need to use integer forecasts.
+Count forecast reconciliation is an active area of research, and it would be interesting to explore how our approach could be adapted to generate count-reconciled probabilistic forecasts in future studies.
+One possible approach to address this is by rounding the forecasts.
+However, the impact of rounding on forecast accuracy varies depending on the level of hierarchy and the scale of the data.
+In situations with high-volume demand, the effects of rounding may become negligible, and forecast accuracy calculations can overlook integer effects.
+On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.}
+
+## An illustration of probabilistic forecast for EMS demand
+
+@fig-forecast-density \textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, regarding the total attended incidents in the PO health board.
+The black dashed line in the graph represents the point forecast, which presents a single numerical estimate for the future number of attended incidents.
+Due to the complexity of including such plots for the entire hierarchy and 84 days ahead, only one example is presented here for 1 day ahead.
+However, it is feasible to generate these plots for the entire hierarchy and for any forecast horizon if necessary.}
+
+\textcolor{teal}{In practice, point forecasts are commonly used, but they have limitations as they ignore the uncertainty associated with the forecast, leading to wrong decisions.
+In contrast, probabilistic forecasts offer an alternative approach to anticipate future demand.
+Rather than providing a single value, they assign likelihoods to all possible demand outcomes, acknowledging that different numbers of attended incidents are possible, but with varying likelihoods.}
+
+\textcolor{teal}{The purpose of probabilistic forecasting, as demonstrated in} @fig-forecast-density, \textcolor{teal}{is to quantify uncertainty.
+Decisions based on these forecasts could focus on the tails of the distribution: unexpected high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
+Such forecasts are valuable tools for decision-makers and planners, especially when dealing with low-probability, high-cost situations.
+Different EMS managements may have varying risk attitudes depending on resource availability, making it crucial to consider the entire distribution when making decisions.
+For instance, these forecasts enable management to calculate the probability of demand exceeding a certain threshold of available resources (e.g., 90\%), which can serve as an informative early warning measure for crowdedness.}
+
+\textcolor{teal}{It is important to note that while point forecasts and prediction intervals can be obtained from the probabilistic forecasts, the reverse is not possible.
+A single number cannot be used to directly derive a probabilistic forecast.
+Prediction intervals, although helpful in indicating possible ranges, do not provide information on the probabilities of low or high demand.}
+
+\textcolor{teal}{In EMS planning, future demand is just one aspect to consider.
+Other inputs, such as capacity, should also be treated as probability distributions to adopt a probabilistic approach to planning.
+To extract valuable insights and make informed decisions from probabilistic forecasts, specialized numerical tools are required, as the forecasts themselves are typically represented as explicit probability density functions or Monte Carlo generators.}
+
+```{r}
+#| label: fig-forecast-density
+#| fig-width: 8
+#| fig-height: 5
+#| out.width: "100%"
+#| fig-cap: "An illustrative example of forecast distribution of ambulance demand (i.e.
+total incidence attended) in PO health board for 1 day ahead.
+x-axis shows all possible outcomes that may occur, with thier likelihhod in y-aixs"
+#| fig-pos: "H"
+incident_gthf <- readr::read_rds(here::here("data/incidents_gt.rds"))
+
+incident_cv <- incident_gthf |> filter(lhb == "PO",is_aggregated(category),is_aggregated(nature))
+fit <- incident_cv %>%
+  model(
+    ets = ETS(sqrt(incident))
+  )
+set.seed(1)
+fcst <- fit %>% generate(h=1, times = 5000, bootstrap = TRUE)
+p1 <- ggplot(data=fcst,aes(x=.sim, y=..density..)) +
+  geom_histogram(binwidth = 1, fill="#d3d3d3", color="black")+
+  geom_vline(data=fcst,aes(xintercept = (mean(.sim))), color="black", linetype="dashed", linewidth=1)+
+  scale_x_continuous(breaks = seq(25,70,5))+
+  scale_y_continuous(labels=percent)+
+  ggthemes::theme_few()+
+    labs(y="Density", x = "Demand", title = "Probabilistic forecast of ambulance demand for 1 day ahead")
+p1
+```

### 2023-07-18 manuscript/06-conclusion.qmd
response letter and paper ready for submission

@@
 
-Further research could investigate the practical benefits of probabilistic hierarchical forecasting in EMS.
-Linking forecasts with its utilities (response time, resource utilization, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.
-While we generated count distributions of attended incidents for the base forecasts using Poisson regression models, the reconciled forecast distributions are not counts.
-This could be also an avenue for further research.
+\textcolor{teal}{Our research establishes a strong basis for future investigations and practical implementation in EMS.
+Leveraging the hierarchical and grouped structure of demand time series, EMS can use this advanced forecasting framework to generate coherent point and probabilistic forecasts, making the most of all available data at every level of the hierarchy.
+We acknowledge that a forecast serves a greater purpose beyond its mere existence, ideally enabling the best utility in terms of efficient allocation of medical services, response time, and cost, all informed by the forecast.
+While we fully appreciate the importance of evaluating forecast quality based on its impact on decision-making processes, it is essential to address the data requirements and methodology involved in measuring this impact.
+For a comprehensive assessment of the forecasts' implications, access to additional data beyond ambulance demand, covering various decision types, capacity information, constraints in the decision system, and more, becomes necessary.
+This additional data would offer valuable insights into the specific decisions relying on the forecasts, resulting in a more accurate evaluation of their impact on medical services.
+Furthermore, measuring the actual impact of forecasts would necessitate an approach that goes beyond forecasting itself.
+This would involve developing and implementing simulation models capable of replicating decision-making processes based on the forecast inputs.
+These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.}
+
+\textcolor{teal}{Future research can build upon this study in several ways.
+In future investigations, we aim to explore this avenue by incorporating operational information, simulating decision processes, and assessing the decision impact of this framework on utilities that are significant to the EMS.
+Linking forecasts with its utilities (e.g.
+response time, allocation of medical services, resource utilization, cost, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.} Aditionally, in our study, we employed Poisson regression models to generate count distributions of attended incidents for the base forecasts.
+However, it is essential to note that the reconciled forecast distributions are not counts.
+This observation presents an interesting avenue for future research.
+\textcolor{teal}{Also, the dataset used in this study only includes information on attended incidents, it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
+It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.}
+
+\textcolor{teal}{Although our study primarily focuses on Emergency Medical Services, it is essential to emphasize that the framework we propose has broad applicability across various service industries} [@ostrom2010moving].
+\textcolor{teal}{Our approach is particularly valuable in situations where time series data is structured hierarchically and/or grouped, a common characteristic found in many sectors.
+This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
+For instance, in the supply chains } [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
+Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency.
+In the financial industry} [@kimes1998strategic], \textcolor{teal}{where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
+Similarly, in transportation, the framework supports forecasting transportation demand at various levels, optimizing route planning and resource allocation.
+Likewise, in the hospitality and tourism industry} [@dekimpe2016impact], \textcolor{teal}{it facilitates forecasting demand rates at state, regional, and department levels, enabling strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses.
+Additionally, in call centers, accurate call volume forecasting at different levels of the call center hierarchy or grouped structure is crucial for workforce management and resource allocation.
+Implementing our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, ensuring efficient staffing and optimal service levels to meet customer demands.}

### 2023-07-18 manuscript/main.qmd
response letter and paper ready for submission

@@
 library(patchwork)
+library(scales)
 # Color blind color palettes

### 2023-07-18 manuscript/05-results.qmd
ensemble 2 removed from table 3

@@
     model != "qcomb",
-    h > 42
+    h > 42,
+    model != "ensemble2"
   ) |>
@@
          model != "qcomb",
-         h > 42
+         h > 42,
+         model != "ensemble2"
   ) |>

### 2023-07-18 manuscript/main.tex
ensemble 2 removed from table 3

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
 \fi
@@
 
-\usepackage{todonotes,mathtools,bm,amsmath}
-\mathtoolsset{showonlyrefs}
-\usepackage{mathpazo}
-\setlength{\parindent}{0cm}
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
+\usepackage{todonotes,mathtools,bm,amsmath,mathpazo}
+\mathtoolsset{showonlyrefs}
+\setlength{\parindent}{0cm}
 \makeatletter
@@
 \makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
 \makeatother
@@
 \makeatother
-\journal{Annals of Emergency Medicine}
+\makeatletter
+\makeatother
+\journal{Journal of Service Research}
 \ifLuaTeX
@@
 and deploying staff and fleet. Such demand forecasts are required at
-national, regional and sub-regional levels, and must take account of the
-nature of incidents and their priorities. These forecasts are often
+national, regional, and sub-regional levels, and must take account of
+the nature of incidents and their priorities. These forecasts are often
 generated independently by different teams within the organization. As a
@@
 service. To address this issue, we exploit the hierarchical and grouped
-structure of the demand time series, and apply forecast reconciliation
+structure of the demand time series and apply forecast reconciliation
 methods to generate both point and probabilistic forecasts that are
@@
 nature of incident, priority, managing health board, and control area.
-We use an ensemble of forecasting models, and show that the resulting
+We use an ensemble of forecasting models and show that the resulting
 forecasts are better than any individual forecasting model. We validate
-the forecasting approach using time-series cross-validation.
+the forecasting approach using time series cross validation.
 \end{abstract}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 Services (EMS) results in patient flow problems, with serious
-consequences for patients, staff and the entire care system
+consequences for patients, staff, and the entire care system
 \citep{ekstrom2015forecasting, ROSTAMITABAR20221197}. Demand forecasting
@@
 patient outcomes. Accurate daily demand forecasting enables planners and
-decision makers to manage resources to meet anticipated patients,
+decision-makers to manage resources to meet anticipated patients,
 reconfigure units, and redeploy staff and vehicles as necessary.
@@
 organization to inform various planning and decision-making processes
-\citep{hulshof2012taxonomic}. There are some planning process at the
+\citep{hulshof2012taxonomic}. There are some planning processes at the
 national level (strategic and long-term) such as workforce resource
-planning and budgeting; sub-national, regional or healthcare level
+planning and budgeting; sub-national, regional, or healthcare level
 (tactical and medium-term) such as temporary capacity expansions,
-resource sharing and staff-shift scheduling; and hospital or station
-level (operational and short-term) such as planning rosters for staff
-and ambulance deployment. Demand forecasts might also be required at
-different levels for a specific area of interest such as the nature of
-demand or the priority level. Moreover, the time series data in EMS has
-an inherent hierarchical and grouped structure to support such
-forecasting requirements. Demand for emergency medical services at the
-country level can be disaggregated in a geographical hierarchy into
+resource sharing; and hospital or station level (operational and
+short-term) such as planning rosters for staff and ambulance deployment.
+Demand forecasts might also be required at different levels for a
+specific area of interest such as the nature of demand or the priority
+level. Moreover, the time series data in EMS has an inherent
+hierarchical and grouped structure to support such forecasting
+requirements. Demand for emergency medical services at the national
+level can be disaggregated in a geographical hierarchy into
 sub-national, regions, health boards, and stations/hospitals, or divided
 into groups such as the nature of incidents or demand priority.
-Therefore, using forecasting methodologies that account for hierarchical
-and/or grouped structures of time series in EMS is a natural fit.
+\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchies are necessary for effective decision-making in EMS.
+For example, control area EMS forecasts can inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals.
+At the lower levels, hospitals or ambulance stations could use such forecasts to plan for staffing and resource allocation, ambulance dispatching, staff-to-shift assignment, staff rescheduling based on the anticipated volume and priority and nature of incidents.
+Additionally, generating forecasts at lower levels could potentially improve the accuracy of the high-level forecasts, by providing more detailed information on the nature and priority of incidents.
+This could help to identify patterns in demand that may not be apparent at the higher level.
+Therefore, employing forecasting techniques that consider the hierarchical and/or grouped patterns of time series in EMS aligns naturally, offering the possibility to enhance forecast accuracy and facilitate coordination.}
 
@@
 the hierarchical data structure has been largely ignored, and the main
-focus has been on producing independent forecasts at a single level.
-Generating independent forecasts can result in a lack of consistency and
-coordination, and therefore leads to less effective planning and
-decision making. With hierarchical forecasting, plans at any level are
-based on coherent forecasts and therefore can be aligned. Implementing
-and sustaining improvements in EMS require alignments and coordination
-between different stakeholders, without which teams operate in isolation
-leading to conflicts, duplication work, rework, or work that runs
-counter to the overall goal to improve the quality of delivery service.
-Hierarchical forecasting framework can be used as a tool to improve
-coordination between teams across the care services at the national,
-sub-national, regional and local levels. The hierarchical forecasting
-approaches not only create consistent forecasts, but are usually also
-more accurate than the independent (base) forecasts
-\citep{hyndman2011optimal}. To our knowledge, there has been no previous
-research involving hierarchical and grouped forecasting in the entire
-field of forecasting for healthcare management.
+focus has been on producing independent (base) forecasts at a single
+level. Generating independent forecasts can result in a lack of
+consistency and coordination, and therefore leads to less effective
+planning and decision making.
+\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
+If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area.
+Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others.
+Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly.
+Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}
+
+With hierarchical forecasting, plans at any level are based on coherent
+forecasts and therefore can be aligned. Implementing and sustaining
+improvements in EMS require alignments and coordination between
+different stakeholders, without which teams operate in isolation leading
+to conflicts, duplication work, rework, or work that runs counter to the
+overall goal to improve the quality of delivery service. Hierarchical
+forecasting framework can be used as a tool to improve coordination
+between teams across the care services at the national, sub-national,
+regional and local levels. The hierarchical forecasting approaches not
+only create consistent forecasts but are usually also more accurate than
+the independent (base) forecasts \citep{hyndman2011optimal}. To our
+knowledge, there has been no previous research involving hierarchical
+and grouped forecasting in the entire field of forecasting for
+healthcare management.
 
@@
 hierarchical forecasting approaches in the EMS using daily time series
-of verified incidents from 2015 to 2020 in a major ambulance service in
+of attended incidents from 2015 to 2019 in a major ambulance service in
 Great Britain. The data has hierarchical and grouped structures, with
-hierarchies at the national, control (i.e.~sub-national), health board
-(i.e.~regional) levels, as well as groups by priority and nature of
-incidents. We produce consistent point forecasts and forecast
+hierarchies at the national, control (i.e.~sub-national), and health
+board (i.e.~regional) levels, as well as groups by priority and nature
+of incidents. We produce consistent point forecasts and forecast
 distributions for all levels, which is critical for an effective
@@
 independent/base forecasts using Exponential Smoothing State Space
-(ETS), Generalized Linear Model (GLM), Poisson regression, a simple
-empirical distribution and an ensemble method, followed by applying
-bottom-up and optimal reconciliation approaches. Forecast performance is
-assessed by the Mean Squared Scaled Error (MSSE) for point forecasts and
+(ETS), Poisson regression using Generalized Linear Model (GLM) and
+tscount (TSGLM), a simple empirical distribution and an ensemble method,
+followed by applying bottom-up and optimal reconciliation approaches.
+Forecast performance is assessed by the Mean Absolute Scaled Error
+(MASE) and Mean Squared Scaled Error (MSSE) for point forecasts and
 Continuous Ranked Probability Scores (CRPS) for the probabilistic
 forecasts. This paper complies with reproducibility principles
-\citep{stodden2013best, boylan2015reproducibility}, and can be applied
-to any healthcare service (e.g., emergency department, primary or social
-care) subject to the time series having a hierarchical and/or grouped
-structure.
+\citep{stodden2013best, boylan2015reproducibility}. We provide the R
+codes for the proposed models and benchmarks. Therefore, they can be
+applied to any healthcare service (e.g., emergency department, primary
+or social care) subject to the time series having a hierarchical and/or
+grouped structure.
 
@@
 delivery requires accurate resource planning that generally relies on
-demand forecasts at operational, tactical and strategic levels.
-
-There is a substantial number of studies on the application of time
-series forecasting in the Emergency Medical Services. For example,
+demand forecasts at operational, tactical, and strategic levels. There
+is a substantial number of studies on the application of time series
+forecasting in the Emergency Medical Services. For example,
 \citet{ibrahim2016modeling} provide an extensive review of the models
@@
 provider to respond to it, or forwarding it to other channels such as
-police, firefighters or general practitioners. Our study is aligned with
-this stream of the literature. Another similar area that has received
+police, firefighters, or general practitioners. Our study is aligned
+with this stream of literature. Another similar area that has received
 considerable attention is Emergency Department forecasting; we refer
 interested readers to \citet{mingliterature2022},
-\citet{gul2020exhaustive} and \citet{wargon2009systematic} for extensive
-reviews of the relevant literature. Although crucial to EMS performance,
-\citet{aringhieri2017emergency} state that demand forecasting has
-received limited research attention in the EMS context. In this section,
-we provide a brief review of studies on forecasting ambulance demand in
+\citet{gul2020exhaustive}, and \citet{wargon2009systematic} for
+extensive reviews of the relevant literature. In this section, we
+provide a brief review of studies on forecasting ambulance demand in
 EMS.
@@
 need an ambulance. A stepwise ordinary least squares regression analysis
-is used for estimating the parameter and generating forecast. The only
+is used for estimating the parameter and generating forecasts. The only
 performance measure reported in this study is \(R^2\), which is not an
@@
 Spectrum Analysis (SSA) method to generate forecasts of the EMS demand
-at the national level for 7-day, 14-day, 21-day and 28-day forecast
+at the national level for 7-day, 14-day, 21-day, and 28-day forecast
 horizons using data provided by an ambulance service in Great Britain.
@@
 compare and contrast the performance of forecasting methods based on
-utility measures such as cost, resource utilization or response time.
+utility measures such as cost, resource utilization and response time.
 The tool contains options that allow generating forecasts at various
-levels of granularity, however, it ignores the hierarchical and grouped
-relationships structure, preventing aligned decision making and
-coordination.
+levels of granularity; however, it ignores the structure of the
+hierarchical and grouped relationships, preventing aligned decision
+making and coordination. \citet{al2021empirical}
+\textcolor{teal}{utilises real data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand accurately.
+The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
+Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE) are calculated.
+The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
+However, when it comes to long-term demand, the SSA method proves to be the most effective.}
+\citet{ibrahim2016modeling}
+\textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
+Real-life call centre data was utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
+The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
+The findings indicate that the ME consistently produces the most accurate point forecasts.
+On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
+For a two-week leading period, MU shows better coverage probability.}
 
 \citet{9659837} investigate forecasting EMS demand in a high
-spatio-temporal resolution of 1km\(^2\) spatial regions and 1-hr time
+Spatio-temporal resolution of 1km\(^2\) spatial regions and 1-hr time
 intervals using total incidents in Oslo, Norway, from 1 January 2015 to
@@
 performed better in producing point forecasts, while a distribution
-baseline method based on spatial distribution of the incidents across
-all time steps provided more accurate forecast distributions.
+baseline method based on the spatial distribution of the incidents
+across all time steps provided more accurate forecast distributions.
 \citet{zhou2016predictinglit} proposed three methods based on Gaussian
@@
 incidents from Toronto, Canada (years 2007 and 2008 with 391,296 events)
-and Melbourne, Australia (years 2011 and 2012 with 696,975 events) were
-used to build the model and examine the performance on test data using
-mean negative log likelihood. They show that forecasts generated by the
-proposed methods were significantly more accurate than the current
-industry practice (a simple averaging formula).
-\citet{grekousis2019will} investigated the combination of spatial
-analysis methods with data mining techniques based on an improved
-Hungarian algorithm and a MLP neural network to identify the most likely
-locations of future emergency events. The proposed approach was tested
-using data of 2851 events attended by the EMS in Athens, Greece, over 24
-weeks. They showed that 23\% of real emergency events lie within 50
-meters of the predicted ones and nearly 70\% of the real emergency
-events lie no further than 150 meters away, which is rather accurate
-given the granularity of the problem at the city level.
+and Melbourne, Australia were used to build the model and examine the
+performance on test data using mean negative log-likelihood. They show
+that forecasts generated by the proposed methods were significantly more
+accurate than the current industry practice (a simple averaging
+formula). \citet{grekousis2019will} investigated the combination of
+spatial analysis methods with data mining techniques based on an
+improved Hungarian algorithm and a MLP neural network to identify the
+most likely locations of future emergency events. The proposed approach
+was tested using data from 2851 events attended by the EMS in Athens,
+Greece, over 24 weeks. They showed that 23\% of real emergency events
+lie within 50 meters of the predicted ones and nearly 70\% of the real
+emergency events lie no further than 150 meters away, which is rather
+accurate given the granularity of the problem at the city level.
 
@@
   structure of the time series data, and the relationship between series
-  at different levels of hierarchy. This may result in incoherent
-  forecasts leading to misaligned planning and decision making. While
-  the hierarchical forecasting methodology has been developed and
-  applied in various domains over the past 10 years
-  \citep{panagiotelis2022probabilistic}, it has never been explored in
-  this area.
+  at different levels of hierarchy. While the hierarchical forecasting
+  methodology has been developed and applied in various domains over the
+  past 10 years \citep{panagiotelis2022probabilistic}, it has never been
+  explored in this area.
 \item
@@
   the entire forecast distribution of daily ambulance demand for the
-  whole hierarchy to inform the decision-making process and to better
-  represent the uncertainty of future demand, providing a risk
-  management tool for planners.
+  whole hierarchy to better represent the uncertainty of future demand,
+  providing a risk management tool for planners.
 \item
@@
 
-This paper concerns the problem of hierarchical forecasting in EMS and
-generates and evaluates both point and probabilistic forecast across
-different levels of the hierarchy, hence addressing some important gaps
-identified in the literature.
+\begin{landscape}\begin{table}[!h]
+
+\caption{Summary of some studies in forecasting FOR AMBULANCE SERVICES}
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{11}{13}\selectfont
+\begin{tabular}[t]{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}ll}
+\toprule
+Reference & Year & Variable & Horizon & Method & Metric & Probabilistic & Reconciliation\\
+\midrule
+Current study & 2023 & Ambulance demand & 84 days & Naïve;  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES\\
+Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO\\
+Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & Yes & NO\\
+Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO\\
+Ibrahim et al.
+& 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO\\
+Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & No & NO\\
+Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO\\
+\bottomrule
+\end{tabular}}
+\end{table}
+\end{landscape}
 
@@
 
-We are interested in generating forecasts to inform the planning horizon
-of 42 days, required by planners in the ambulance service. The forecast
-horizon in this study is \(2 \times 42 = 84\) days ahead, because the
-planning is generally frozen for 42 days and so forecasts of the next 42
-days is not particularly helpful for planning. While forecasts are
-generated for 84 days ahead, performance evaluation is only assessed
+Planners in the ambulance service work with a planning horizon of 6
+weeks. That is, planning is generally frozen for the next 42 days, so
+any forecasts will only affect plans for the time period beyond the next
+42 days. Consequently, the forecast horizon in this study is
+\(2 \times 42 = 84\) days ahead, with performance evaluation assessed
 based on the last 42 days and not the whole forecast period. The
@@
 \subcaption{\label{fig-hierarchy-1}Hierarchical structure: Attended
-incidents in all country is disaggregated into 3 control areas /regions
-and then into 7 different healthboard at the bottom}
+incidents in the whole country are disaggregated into 3 control areas
+and then into 7 different healthboards.}
 \end{minipage}%
@@
 attended incident (i.e.~there are 35 different nature of incidents
-including chest pain , breathing problems, heart attack, stroke, and so
+including chest pain, breathing problems, heart attack, stroke, and so
 on). The symbol * refers to the crossed attributes between hierarchical
@@
 health boards. Attended incident data are categorized into 3 priority
-classes of red, amber and green. There are also 35 different nature of
+classes of red, amber, and green. There are also 35 different nature of
 incidents such as chest pain, stroke, breathing problem, etc. In total,
@@
 \centering
+\begin{threeparttable}
 \begin{tabular}{lr}
@@
 \end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset.
+As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.
+\end{tablenotes}
+\end{threeparttable}
 \end{table}
@@
 series with the strength of trend in x-axis and the strength of
-seasonality in y-axis. Both measures are on a scale of {[}0,1{]}. It is
-clear that there are some series showing strong trends and/or
+seasonality in y-axis. Both measures are on a scale of {[}0,1{]}.
+
+\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by}
+\citet{kang2017visualising}.
+\textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
+The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
+The strength of trend is defined as:}
+\[F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)\]
+\textcolor{teal}{For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
+Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small.
+But for data with little or no trend, the two variances should be approximately the same.}
+
+\textcolor{teal}{The strength of seasonality is defined similarly:}
+\[F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).\]
+\textcolor{teal}{series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).}
+
+It is clear that there are some series showing strong trends and/or
 seasonality, corresponding to series at the higher levels of the
@@
 
-\caption{\label{fig-feature}Time series features of attended incidents
-across all levels (1530 series)}
+\caption{\label{fig-feature}The strength of the trend and weekly
+seasonality in the time series of attended incidents. The scatter plot
+shows a total of 1530 data points, with each point corresponding to a
+specific time series.}
 
@@
 Figure~\ref{fig-dataviz2} reveals different information such as trend,
-seasonality and noise. For example, some series depict seasonality and
+seasonality, and noise. For example, some series depict seasonality and
 trend, whereas some other series report low volume of attended incidents
@@
 
-\caption{\label{fig-dataviz2}Time series of attended incidents at
-various levels. The panels show data from the whole country, by control
+\caption{\label{fig-dataviz2}Daily time plot of attended incidents at
+various levels. X-axis shows the date of incidents, consisting of 1400
+data points (days) and y-axis shows the number of attended incidents.
+The panels show data from the whole country (top panel), by control
 area, by health board, by priority level, and by nature of incident.
 Only four of the 35 nature of incident categories are shown to avoid too
-much overplotting.}
+much overplotting. Each time plot .}
 
@@
 holidays and Christmas Day and New Year's Day as predictors of incident
-attended. These type of holidays will affect peoples' activities and may
-increase or decrease the number of attended incidents.
+attended. These types of holidays will affect peoples' activities and
+may increase or decrease the number of attended incidents.
 
@@
 
-Given the presence of various significant patterns in the past attended
-incidents, we consider three different forecasting models to generate
-the base forecasts. Once the base forecasts are produced, hierarchical
-and grouped time series methods are used to reconcile them across all
+Given the presence of various patterns in the past attended incidents,
+we consider three different forecasting models to generate the base
+forecasts. Once the base forecasts are produced, hierarchical and
+grouped time series methods are used to reconcile them across all
 levels. We briefly discuss forecasting models in the following sections,
@@
 distribution of future attended incidents.
+\textcolor{teal}{We have chosen the naive method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
+Additionally, the naive method has shown surprisingly high accuracy.
+Hence, any forecasting approach that can offer superior results compared to the naive method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
 
 \textbf{Exponential Smoothing State Space model (ETS):} ETS models
-\citep{hyndman2021forecasting} can combine trend, seasonality and error
+\citep{hyndman2021forecasting} can combine trend, seasonality, and error
 components in a time series through various forms that can be additive,
@@
 can be additive (``A'') or multiplicative (``M''). To forecast the
-attended incidents at each level, we use the \texttt{ETS()} function in
-the \texttt{fable} package \citep{fable2022} in R. To identify the best
-model for a given time series, the \texttt{ETS} function uses the
-corrected Akaike's Information Criterion (AICc).
+attended incidents at each level, we use the \texttt{ets()} function in
+the \texttt{forecast} package \citep{Rforecast, HK08} in R. To identify
+the best model for a given time series, the \texttt{ets} function uses
+the corrected Akaike's Information Criterion (AICc).
+
+\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
+Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
 
@@
 the models that is frequently used in practice to generate count
-forecasts is Poisson regression. We will consider forecasting attended
-incidents using the covariates spline trend, day of the week dummy
-variables (from Monday to Sunday), Fourier terms to capture yearly
-seasonality, public holidays (1 when is public holiday, 0 otherwise),
-school holidays (1 when is school holiday, 0 otherwise) and Christmas
-Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is
-New Year's Day, 0 otherwise). We fit a Poisson regression model using
-the function \texttt{glm()} from the \emph{stats} package in R, with the
-argument \texttt{family\ =\ poisson} to specify that we wish to fit a
-Poisson regression model with a log link function.
+forecasts is Poisson regression.
+
+\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+The Fourier terms are as defined in}
+\citet{hyndman2021forecasting}
+\textcolor{teal}{(Section 7.4).
+This model takes account of weekly seasonality and annual seasonality.
+Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
+There is no reason, for example, for incidents to occur more at some times of the month than others.}
+
+We fit a Poisson regression model using the function \texttt{glm()} from
+the \emph{stats} package in R, with the argument
+\texttt{family\ =\ poisson} to specify that we wish to fit a Poisson
+regression model with a log link function.
 
@@
 another Poisson regression model that takes into account serial
-dependence, using the \texttt{tsglm()} function in the \emph{tscount}
-package in R \citep{JSSv082i05} to model the attended incidents. Again,
-the logarithmic link function is used to ensure that the parameter of
-Poisson distribution is always positive. This model captures the short
-range serial dependence by including three autoregressive terms, in
-addition to the same covariates that were used in the GLM model. To
-distinguish this from the previous GLM model, we will refer to this
-model as ``TSGLM''.
-
-\textbf{Ensemble method:} Finally, we use an ensemble method that
-combines the forecasts generated from the Naive, ETS, GLM and TSGLM
-models to form a mixture distribution \citep{combinations}.
+dependence. This model captures the short range serial dependence by
+including autoregressive terms, in addition to the same covariates that
+were used in the GLM model. To distinguish this from the previous GLM
+model, we will refer to this model as \texttt{TSGLM}.
+
+\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional component accounting for serial correlation (i.e.
+autoregressive).
+The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points.
+Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
+TSGLM can be written as}
+
+\begin{align*}
+  y_t &\sim \text{Poisson}(\lambda_t) \\
+  \text{where}\qquad
+  \log(\lambda_t) &= (\bm{y}_{t-k}'+\bm{x}_t')\bm{\beta},
+\end{align*}
+
+\textcolor{teal}{${y}_{t-k}$ is a vector of k lagged values.
+The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
+past values) of the ambulance demand in the model.
+This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}
+
+\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R}
+\citep{JSSv082i05}
+\textcolor{teal}{to model the attended incidents. Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
+
+\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution}
+\citep[p156-158]{feller1991introduction}.
+\textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
+To account for changes over time, we incorporate trend and seasonality covariates, as well as public holiday effects, allowing the mean of the Poisson distribution to vary.
+However, it is important to note that if there are additional factors influencing the mean of the Poisson distribution, which are not accounted for in our model, we might observe over-dispersion or under-dispersion in the data.}
+
+\textbf{Ensemble method:}
+\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by averaging the forecasts, rather than relying on separate forecasts generated by each individual method}
+\citep{clemen1989combining}.
+\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models using a simple average to form a mixture distribution}
+\citep{combinations}.
+
+\textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in}
+\citet{panagiotelis2023probabilistic}.
+\textcolor{teal}{This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+In this way, we can generate an empirical distribution of forecasts for each model.
+The ensemble forecast distribution is a simple mixture of these empirical distributions.}
+
+\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models or promote a particular model class.
+Instead, we have developed a flexible framework that can accommodate any forecasting models.
+Our primary objective is to demonstrate its practicality and effectiveness in integrating base forecasts from any model and generating coherent forecasts within a hierarchical structure.}
 
@@
 2019-05-09, with the test set ending on 2019-07-31. Model development
-and hyper-parameter tuning is performed using the training data only.
-While we compute forecast errors for the entire 12 weeks, we are most
-interested in the last 42 days of each test set, because that
-corresponds to how forecasts are generated for planning in practice.
-Forecasting performance is evaluated using both point and probabilistic
-error measures.
+and hyper-parameter tuning is performed using the training data and the
+errors are assess using the corresponding test set. While we compute
+forecast errors for the entire 12 weeks, we are most interested in the
+last 42 days of each test set, because that corresponds to how forecasts
+are generated for planning in practice. Forecasting performance is
+evaluated using both point and probabilistic error measures.
+
+\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by j, representing the number of time periods ahead we are predicting.
+In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
 
@@
 (MSSE) and the Mean Absolute Scaled Error (MASE). The Mean Absolute
-Scaled Error (MASE) \citep{HK06} is calculated as: \[
+Scaled Error (MASE) \citep{HK06, hyndman2021forecasting} is calculated
+as: \[
   \text{MASE} = \text{mean}(|q_{j}|),
@@
 
-A related measure is MSSE \citep{hyndman2021forecasting}, which uses
-squared errors rather than absolute errors:
+A related measure is MSSE
+\citep{hyndman2021forecasting, makridakis2022m5}, which uses squared
+errors rather than absolute errors:
 \begin{equation}\protect\hypertarget{eq-RMSSE}{}{
@@
 
+\textcolor{teal}{Using of scale-independent measures, such as MASE and MSSE, enables more significant comparisons between time series at different levels and scales, as these measures are not influenced by the data's magnitude.
+This aspect holds particular importance in our study, as we work with time series at various levels of hierarchy.
+These time series show varying scales, resulting in different magnitudes of error.
+By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
+
 To measure the forecast distribution accuracy, we calculate the
-Continuous Rank Probability Score \citep{gneiting2014probabilistic}. It
-rewards sharpness and penalizes miscalibration, so it measures overall
+Continuous Rank Probability Score
+\citep{gneiting2014probabilistic, hyndman2021forecasting}. It rewards
+sharpness and penalizes miscalibration, so it measures overall
 performance of the forecast distribution.
@@
 
+\textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+It measures how well the predicted probabilities match the actual.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
+A model is well-calibrated if the predicted probabilities match the actual, and it is sharp if it is confident in its predictions.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.}
+
+\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores}
+\citep[Section 5.9]{winkler1972decision, hyndman2021forecasting}
+\textcolor{teal}{or percentile scores} \citep[section
+5.9]{hyndman2021forecasting},
+\textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.}
+\textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
+Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.
+95\%) quantiles.
+In such cases, a percentile score  becomes useful in meeting this specific requirement.}
+
 \hypertarget{sec-htc}{%
@@
 attributes. While series could be strictly hierarchical or only grouped
-bases on some attributes, in many situation a more complex structures
+bases on some attributes, in many situations a more complex structures
 arise when attributes of interest are both nested and crossed, having
@@
 
-\hypertarget{reconciliation-methos}{%
-\subsection{Reconciliation methos}\label{reconciliation-methos}}
+\hypertarget{reconciliation-methods}{%
+\subsection{Reconciliation methods}\label{reconciliation-methods}}
 
@@
 combined with the hierarchical structure, there is no unique way to
-disaggregate top forecasts. Hence the top-down cannot be used in our
-application. The recommended approach is to use forecast reconciliation
-\citep{hyndman2011optimal}. In the following sections, we first discuss
-some notation, and then present bottom-up and forecast reconciliation
-approaches used in this study to generate coherent forecasts.
+disaggregate top forecasts. Hence the top-down approach cannot be used
+in our application. The recommended approach is to use forecast
+reconciliation \citep{hyndman2011optimal}. In the following sections, we
+first discuss some notation, and then present bottom-up and forecast
+reconciliation approaches used in this study to generate coherent
+forecasts.
 
@@
 
-Let \(\bm{b}_t\) be a vector of \(n_b\) ``bottom-level'' time series at
-time \(t\), and let \(\bm{a}_t\) be a corresponding vector of
+Let \(\bm{b}_t\) be a vector of \(n_b\) \emph{bottom-level} time series
+at time \(t\), and let \(\bm{a}_t\) be a corresponding vector of
 \(n_a = n-n_b\) aggregated time series, where \[
@@
 specifying how the bottom-level series \(\bm{b}_t\) are to be aggregated
-to form \(\bm{a}_t\). The full vector of time series is given by \[
+to form \(\bm{a}_t\).
+\textcolor{teal}{The aggregation matrix $A$ is determined by the structure of the hierarchy.
+It maps the bottom-level time series to the corresponding higher-level time series.
+The matrix $A$ is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
+For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $A = \begin{bmatrix} 1 1 \end{bmatrix}$.
+}
+
+The full vector of time series is given by \[
  \bm{y}_t = \begin{bmatrix}\bm{a}_t \\\bm{b}_t\end{bmatrix}.
@@
 \] such that \(\bm{y}_t = \bm{S}\bm{b}_t\).
+
+\textcolor{teal}{The term "level bottom series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
+For instance, in Table 2, each distinct combination of values in Control area (e.g.
+South \& East), Health board (e.g.
+CV), Priority (e.g.
+Green), and Nature of incident (e.g.
+Chest pain) corresponds to one individual time series.
+In the dataset at hand, there are 691 unique combinations, resulting in 691 bottom level time series.
+The "aggregate time series" describes how these bottom-level series are combined to create higher-level series.
+For instance, to obtain the incidents at the national level (i.e.
+all country level), the time series are aggregated across all Control areas, Health boards, Priorities, and Natures of incidents.
+Any desired aggregation level can be achieved based on the data structure, utilizing the bottom level series available.}
 
@@
 bottom level, but these series may be noisy and difficult to forecast.
-The approach using only the data at the most disaggregated level, and so
+The approach uses only the data at the most disaggregated level, and so
 does not utilize all the information available across the hierarchical
@@
 
-Forecast reconciliation approaches fill this gap by combining and
-reconciling all the base forecasts in order to produce coherent
-forecasts. Linear reconciliation methods can be written
-\citep{WicEtAl2019} as \[
+The bottom-up (BU) approach is constrained by its reliance solely on
+base forecasts from a single level of aggregation at the bottom level.
+While it does result in consistent forecasts, the BU approach lacks
+forecast reconciliation since no reconciliation is performed.
+
+\textcolor{teal}{Forecast reconciliation approaches bridge this gap by combining and reconciling all base forecasts to generate coherent forecasts.
+This technique utilizes all the base forecasts produced within a hierarchical structure to create consistent forecasts at every level of the hierarchy.
+As a result, it goes beyond relying solely on base forecasts from a single level of aggregation, and instead leverages all available information at each level to generate forecasts that minimize the total forecast variance of the set of coherent forecasts.}
+\textcolor{teal}{Linear reconciliation involves projecting the base forecasts onto the coherent space.
+It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased}
+\citep{WicEtAl2019}.
+
+Linear reconciliation methods can be written \citep{WicEtAl2019} as \[
   \tilde{\bm{y}}_h = \bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{W}^{-1}\hat{\bm{y}}_h,
@@
 different solutions such as Ordinary Least Squares (OLS), Weighted Least
-Squares (WLS) and Minimum Trace (MinT). We use the implementation of
-these methods in the fable package in R in the experiment.
+Squares (WLS) and Minimum Trace (MinT).
+
+\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
+In this approach, the estimation of W is based on the assumption that all the errors have equal variance.
+Hence, W is simply defined as the identity matrix multiplied by a constant factor.
+The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
+The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
+A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}
+
+\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
+But it assumes that the errors of each series are uncorrelated with each other.
+In this approach, W is defined as a diagonal matrix with the variance of the errors on the diagonal.
+The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
+The main weakness of this approach is that it ignores the relationships between series.
+A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.}
+
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.}
+\citet{WicEtAl2019}
+\textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
+The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
+In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}
+
+We use the implementation of these methods in the fable package in R in
+the experiment.
+
+\textcolor{teal}{Certainly, other approaches can be applied to hierarchical forecasting problems}
+\citet{pennings2017integrated} and \citet{villegas2018supply}
+\textcolor{teal}{proposed the idea of using a state space model to ensure consistent forecasts.
+However, when dealing with larger hierarchies, these models encounter difficulties in estimating covariance matrices.
+In contrast, our approach provides a clear advantage by allowing the incorporation of different forecasting methods for the base forecasts, and even accommodating distinct methods for individual series.
+The decoupling of time series models from the reconciliation step adds significant flexibility in exploring a wide range of models.}
 
@@
 In this section, we compare the forecasting performance of the Naive,
-ETS, GLM and TSGLM models along with the ensemble, using base forecast
+ETS, GLM, and TSGLM models along with the ensemble, using base forecast
 and Minimum Trace (MinT) reconciliation methods. We have also computed
 the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least
-Square (WLS) approaches, along with bottom up forecasting. However, they
+Square (WLS) approaches, along with bottom-up forecasting. However, they
 are not reported here because their accuracy is outperformed by MinT. We
@@
 only the top level (Total), the bottom level, and the levels
-corresponding to Control areas and Health boards. The latter are chosen
+corresponding to Control areas and Health boards. The latters are chosen
 because this is where decision-making takes place, so these forecasts
@@
 horizons 43--84 days (corresponding to the planning horizon) is
-presented per model, method and the hierarchical level. Reported
-forecast accuracy are averaged across all forecast horizons, rolling
-origins and series at each level. Table~\ref{tbl-result} presents both
+presented per model, method, and the hierarchical level. Reported
+forecast accuracy is averaged across all forecast horizons, rolling
+origins, and series at each level. Table~\ref{tbl-result} presents both
 point and probabilistic forecast accuracy at total, control area, health
-board and bottom level series. Point forecast performance are reported
-using MASE and MSSE in Table~\ref{tbl-result-1} and
-Table~\ref{tbl-result-2}, respectively. Probabilistic forecast accuracy
-is reported using CRPS in Table~\ref{tbl-result-3}. The bold entries in
-each table identify a combination of method and model that performs best
-for the corresponding level (i.e.~each column), based on the smallest
-values of accuracy measures.
-
-Table~\ref{tbl-result-1} and 2b show that forecast reconciliation
-(i.e.~MinT) improves forecast accuracy at the higher levels of the
-hierarchy including total, control area and health board. However, it is
-does not result in accuracy improvement at the bottom level series, for
-which base forecasts are more accurate. This might be due to the noisy
-structure of time series at the bottom level, and possibly due to very
-different patterns in the aggregated series. It is also clear from
-Table~\ref{tbl-result-1} that the ensemble method improves forecast
+board and bottom level series. Point forecast performance is reported
+using MASE and MSSE in Table~\ref{tbl-result-1} and \ref{tbl-result-2},
+respectively. Probabilistic forecast accuracy is reported using CRPS in
+Table~\ref{tbl-result-3}. The bold entries in each table identify a
+combination of method and model that performs best for the corresponding
+level (i.e.~each column), based on the smallest values of accuracy
+measures.
+
+Table~\ref{tbl-result-1} and \ref{tbl-result-2} show that forecast
+reconciliation (i.e.~MinT) improves forecast accuracy at the higher
+levels of the hierarchy including total, control area and health board.
+However, it does not result in accuracy improvement at the bottom level
+series, for which base forecasts are more accurate. This might be due to
+the noisy structure of time series at the bottom level, and possibly due
+to very different patterns in the aggregated series. It is also clear
+from Table~\ref{tbl-result-1} that the ensemble method improves forecast
 accuracy at total, control area and health board. However, this does not
@@
 perform best, depending on the accuracy measure.
+\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness may not always be guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
+Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy.
+Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}
 
@@
 
-\caption{\label{tbl-result}Average forecast perfomance calculated on the
-test sets at forecast horizons \(h=43,\dots,84\) days, with time series
-cross validation applied to attended incident data. The best approach is
-highlighted in bold.}\begin{minipage}[t]{\linewidth}
+\caption{\label{tbl-result}Average forecast performance calculated on
+the test sets at forecast horizons \(h=43,\dots,84\) days, with time
+series cross validation applied to attended incident data. The testset
+consists of 462 days. The best approach is highlighted in
+bold.}\begin{minipage}[t]{\linewidth}
 \subcaption{\label{tbl-result-1}Point forecast accuracy using MASE }
@@
 \centering
-\begin{tabular}[t]{llrrrr}
+\begin{tabular}[t]{llcccc}
 \toprule
@@
 better the comprehensive performance. We observe that forecast
-reconciliation results in forecast improvement, regardless of the
-hierarchical level. The ensemble method is also more accurate for higher
-levels, but ETS performs slightly better at the bottom level. While
-these results show that forecast reconciliation does not improve
-\emph{point} forecasts at the bottom level, Table~\ref{tbl-result-3}
-indicates that it generates more accurate \emph{distributional}
-forecasts than the base method. This is probably due to the
+reconciliation results in forecast improvement for the total and health
+board level. CRPS is almost identical at the control area and botom
+levels. Base forecasts are slightly better at the control area level,
+while reconciliation is marginally accurate than base at the bottom
+level. The ensemble method is also more accurate for higher levels, but
+ETS performs well at the bottom level. Table~\ref{tbl-result-3}
+indicates that reconciliation using Mint generates accurate
+\emph{distributional} forecasts. The marginal improvement in the average
+probabilistic forecast accuracy at the bottom level might be due to the
 reconciliation method giving improved forecast accuracy in the tails of
@@
 Overall, our results indicate that forecast reconciliation using the
-MinT method provides reliable forecasts, improves upon the base
-(unreconciled) forecasts all levels except the bottom level series. But
-even there, forecast reconciliation using MinT improves accuracy in the
-tails of the distribution.
+MinT method provides reliable forecasts and improves upon the base
+(unreconciled) forecasts at all levels except the bottom-level series.
+But even there, forecast reconciliation using MinT improves accuracy in
+the tails of the distribution.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-accuracy-1.pdf}
+
+}
+
+\caption{\label{fig-accuracy}Average accuracy by week for 12 weeks using
+MinT reconciliation. The total number of days used to calculate the
+accuracy in the test set is 462. Forecasts are generateled every 42
+days, therefore we use 11 samples to calculate the average accuracy.
+CRPS is relative to a naive Empirical Cumulative Distribution Function
+(ECDF). MASE and MSSE are relative to the corresponding values for the
+training set.}
+
+\end{figure}
 
@@
 Figure~\ref{fig-accuracy}. The figure focuses on the hierarchical levels
-important for decision-making including total, control area and health
-board; however the accuracy could be calculated for any level. We only
+important for decision-making including total, control area, and health
+board; however, the accuracy could be calculated for any level. We only
 illustrate the results of the MinT method, given its strong performance
@@
 report the average weekly forecast accuracy instead of the daily
-forecast horizon, as this reduces the visual noise in the figure. Thus
+forecast horizon, as this reduces the visual noise in the figure. Thus,
 the x-axis shows horizons from week 1 (\(h= 1,\dots,7\)) to week 12
@@
 corresponds to the upcoming planning horizon, which is used by planners
-and decision making. For both the point forecast and distributional
+and decision-makers. For both the point forecast and distributional
 accuracy we can see that the ensemble approach performs best across
@@
 
+\textcolor{teal}{Despite using Poisson regression models to create count distributions of attended incidents for the base forecasts, it is important to note that the reconciled forecast distributions do not maintain a count format.
+In practical scenarios, there might be a need to use integer forecasts.
+Count forecast reconciliation is an active area of research, and it would be interesting to explore how our approach could be adapted to generate count-reconciled probabilistic forecasts in future studies.
+One possible approach to address this is by rounding the forecasts.
+However, the impact of rounding on forecast accuracy varies depending on the level of hierarchy and the scale of the data.
+In situations with high-volume demand, the effects of rounding may become negligible, and forecast accuracy calculations can overlook integer effects.
+On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.}
+
+\hypertarget{an-illustration-of-probabilistic-forecast-for-ems-demand}{%
+\subsection{An illustration of probabilistic forecast for EMS
+demand}\label{an-illustration-of-probabilistic-forecast-for-ems-demand}}
+
+Figure~\ref{fig-forecast-density}
+\textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, regarding the total attended incidents in the PO health board.
+The black dashed line in the graph represents the point forecast, which presents a single numerical estimate for the future number of attended incidents.
+Due to the complexity of including such plots for the entire hierarchy and 84 days ahead, only one example is presented here for 1 day ahead.
+However, it is feasible to generate these plots for the entire hierarchy and for any forecast horizon if necessary.}
+
+\textcolor{teal}{In practice, point forecasts are commonly used, but they have limitations as they ignore the uncertainty associated with the forecast, leading to wrong decisions.
+In contrast, probabilistic forecasts offer an alternative approach to anticipate future demand.
+Rather than providing a single value, they assign likelihoods to all possible demand outcomes, acknowledging that different numbers of attended incidents are possible, but with varying likelihoods.}
+
+\textcolor{teal}{The purpose of probabilistic forecasting, as demonstrated in}
+Figure~\ref{fig-forecast-density},
+\textcolor{teal}{is to quantify uncertainty.
+Decisions based on these forecasts could focus on the tails of the distribution: unexpected high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
+Such forecasts are valuable tools for decision-makers and planners, especially when dealing with low-probability, high-cost situations.
+Different EMS managements may have varying risk attitudes depending on resource availability, making it crucial to consider the entire distribution when making decisions.
+For instance, these forecasts enable management to calculate the probability of demand exceeding a certain threshold of available resources (e.g., 90\%), which can serve as an informative early warning measure for crowdedness.}
+
+\textcolor{teal}{It is important to note that while point forecasts and prediction intervals can be obtained from the probabilistic forecasts, the reverse is not possible.
+A single number cannot be used to directly derive a probabilistic forecast.
+Prediction intervals, although helpful in indicating possible ranges, do not provide information on the probabilities of low or high demand.}
+
+\textcolor{teal}{In EMS planning, future demand is just one aspect to consider.
+Other inputs, such as capacity, should also be treated as probability distributions to adopt a probabilistic approach to planning.
+To extract valuable insights and make informed decisions from probabilistic forecasts, specialized numerical tools are required, as the forecasts themselves are typically represented as explicit probability density functions or Monte Carlo generators.}
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-accuracy-1.pdf}
-
-}
-
-\caption{\label{fig-accuracy}Average accuracy by week for 12 weeks using
-MinT reconciliation. CRPS is relative to a naive ECDF. MASE and MSSE are
-relative to the corresponding values for the training set.}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}
+
+}
+
+\caption{\label{fig-forecast-density}An illustrative example of forecast
+distribution of ambulance demand (i.e.~total incidence attended) in PO
+health board for 1 day ahead. x-axis shows all possible outcomes that
+may occur, with thier likelihhod in y-aixs}
 
@@
 hierarchical and grouped structures. For example, looking at time series
-of arrival calls in a clinical desk service, verified incidents, or
-attended incidents in a country, they could be disaggregated by various
-attributes of interest. Total demand in the country could be
-disaggregated by region, then within each region by health board, within
-each health board, by station/hospital, and so on down to the post code
-area. Alternative structures may arise when attributes of interest are
-crossed rather than nested. For example, the total demand could be
-disaggregated by priority (e.g., Red, Amber, Green) or by nature of
-incidents. It is also natural to have a mixed structures, for example
-the total demand could be disaggregated by priority and also by health
-board.
+of arrival calls in a clinical desk service, Emergency Department
+admissions, verified incidents, or attended incidents in a country, they
+could be disaggregated by various attributes of interest. Total demand
+in the country could be disaggregated by region, then within each region
+by health board, within each health board, by station/hospital, and so
+on down to the postcode area. Alternative structures may arise when
+attributes of interest are crossed rather than nested. For example, the
+total demand could be disaggregated by priority (e.g., Red, Amber,
+Green) or by the nature of incidents. It is also natural to have a mixed
+structure, for example, the total demand could be disaggregated by
+priority and by health board.
 
@@
 Our results indicate that forecast reconciliation in EMS can not only
-contribute to a more coordinated approach to decision making through
-producing coherent forecasts, but also it can result in forecast
-accuracy improvement. Our proposed forecasting models, combined with
-reconciliation approaches, outperform the empirical distribution
-benchmark. We show that a substantial forecast improvement can be
-achieved at higher levels of aggregation by applying forecast
-reconciliation methods. When a point forecast is of interest at the
-bottom level of series, we observe that reconciliation may not improve
-the forecast accuracy, if the bottom series are noisy and lack
+contribute to a more coordinated approach to the planning and
+decision-making through producing coherent forecasts, but also it can
+result in forecast accuracy improvements. Our proposed forecasting
+models, combined with reconciliation approaches, outperform the
+empirical distribution benchmark. We show that a substantial forecast
+improvement can be achieved at higher levels of aggregation by applying
+forecast reconciliation methods. When a point forecast is of interest at
+the bottom level of the series, we observe that reconciliation may not
+improve the forecast accuracy if the bottom series are noisy and lack
 systematic patterns. However, forecast reconciliation may result in more
-forecast results for bottom series, if we are interested in the tails of
-forecast distribution rather than just center measures like mean
-(i.e.~point forecast). Producing consistent forecast are also crucial
-for informing planning activities, we also demonstrate that proposed
-models produce consistent forecasts across all forecast horizons.
-Therefore, we recommend that forecast reconciliation approaches to be
+accurate forecast results for bottom series, if we are interested in the
+tails of forecast distribution rather than just center measures like
+mean (i.e.~point forecast). Coherent forecasts are also crucial for
+informing planning activities, and we demonstrate that the proposed
+models produce coherent forecasts across all forecast horizons.
+Therefore, we recommend that forecast reconciliation approaches be
 adopted for routine use in EMS, whenever hierarchical and/or grouped
@@
 
-Further research could investigate the practical benefits of
-probabilistic hierarchical forecasting in EMS. Linking forecasts with
-its utilities (response time, resource utilization, etc) can offer an
-opportunity to maximize benefits through more holistic planning
-approach. While, we generated count attend incident for the base
-forecast using Poisson regression models, however the reconciles
-forecast is snot yet count. This could be also an avenue for further
-research.
+\textcolor{teal}{Our research establishes a strong basis for future investigations and practical implementation in EMS.
+Leveraging the hierarchical and grouped structure of demand time series, EMS can use this advanced forecasting framework to generate coherent point and probabilistic forecasts, making the most of all available data at every level of the hierarchy.
+We acknowledge that a forecast serves a greater purpose beyond its mere existence, ideally enabling the best utility in terms of efficient allocation of medical services, response time, and cost, all informed by the forecast.
+While we fully appreciate the importance of evaluating forecast quality based on its impact on decision-making processes, it is essential to address the data requirements and methodology involved in measuring this impact.
+For a comprehensive assessment of the forecasts' implications, access to additional data beyond ambulance demand, covering various decision types, capacity information, constraints in the decision system, and more, becomes necessary.
+This additional data would offer valuable insights into the specific decisions relying on the forecasts, resulting in a more accurate evaluation of their impact on medical services.
+Furthermore, measuring the actual impact of forecasts would necessitate an approach that goes beyond forecasting itself.
+This would involve developing and implementing simulation models capable of replicating decision-making processes based on the forecast inputs.
+These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.}
+
+\textcolor{teal}{Future research can build upon this study in several ways.
+In future investigations, we aim to explore this avenue by incorporating operational information, simulating decision processes, and assessing the decision impact of this framework on utilities that are significant to the EMS.
+Linking forecasts with its utilities (e.g.
+response time, allocation of medical services, resource utilization, cost, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.}
+Aditionally, in our study, we employed Poisson regression models to
+generate count distributions of attended incidents for the base
+forecasts. However, it is essential to note that the reconciled forecast
+distributions are not counts. This observation presents an interesting
+avenue for future research.
+\textcolor{teal}{Also, the dataset used in this study only includes information on attended incidents, it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
+It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.}
+
+\textcolor{teal}{Although our study primarily focuses on Emergency Medical Services, it is essential to emphasize that the framework we propose has broad applicability across various service industries}
+\citep{ostrom2010moving}.
+\textcolor{teal}{Our approach is particularly valuable in situations where time series data is structured hierarchically and/or grouped, a common characteristic found in many sectors.
+This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
+For instance, in the supply chains }
+\citep{shugan2000advance},
+\textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
+Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency.
+In the financial industry}
+\citep{kimes1998strategic},
+\textcolor{teal}{where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
+Similarly, in transportation, the framework supports forecasting transportation demand at various levels, optimizing route planning and resource allocation.
+Likewise, in the hospitality and tourism industry}
+\citep{dekimpe2016impact},
+\textcolor{teal}{it facilitates forecasting demand rates at state, regional, and department levels, enabling strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses.
+Additionally, in call centers, accurate call volume forecasting at different levels of the call center hierarchy or grouped structure is crucial for workforce management and resource allocation.
+Implementing our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, ensuring efficient staffing and optimal service levels to meet customer demands.}
+
+\hypertarget{reproducibility}{%
+\section*{Reproducibility}\label{reproducibility}}
+\addcontentsline{toc}{section}{Reproducibility}
+
+R code to produce all results in this paper is available at
 

### 2023-07-19 manuscript/02-literature.qmd
Minor rewording, typos and fixed some refs

@@
 The tool contains options that allow generating forecasts at various levels of granularity; however, it ignores the structure of the hierarchical and grouped relationships, preventing aligned decision making and coordination.
-@al2021empirical \textcolor{teal}{utilises real data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand accurately.
+@al2021empirical \textcolor{teal}{utilises data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand.
 The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
-Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE) are calculated.
+Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE).
 The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
 However, when it comes to long-term demand, the SSA method proves to be the most effective.} @ibrahim2016modeling \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
-Real-life call centre data was utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
+Call centre data were utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
 The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.

### 2023-07-19 manuscript/03-experiment.qmd
Minor rewording, typos and fixed some refs

@@
 
-\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by} @kang2017visualising.
+\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by} @mstl.
 \textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
@@
 
-In addition to displaying the trend and seasonality features, we also visualize few time series at various levels of the aggregation.
+In addition to displaying the trend and seasonality features, we also visualize a few time series at various levels of aggregation.
 @fig-dataviz2 reveals different information such as trend, seasonality, and noise.
@@
 
-**Naive:** We start with a simple forecasting approach, assuming that the future days will be similar to past days.
+**Stationary:** We start with a simple forecasting approach, assuming that the future days will be similar to past days.
 We use the empirical distribution of the past daily attended incidents to create the forecast distribution of future attended incidents.
-\textcolor{teal}{We have chosen the naive method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+\textcolor{teal}{We have chosen this "stationary" method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
 Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
-Additionally, the naive method has shown surprisingly high accuracy.
-Hence, any forecasting approach that can offer superior results compared to the naive method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
+Additionally, the stationary method has shown surprisingly high accuracy.
+Hence, any forecasting approach that can offer superior results compared to the stationary method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
 
@@
 \textcolor{teal}{In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series.
-Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
 Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
-Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+Consequently, the automated algorithm selects the best model based on the unique characteristics of each individual time series.
 As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
@@
 \textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
-
 \begin{align*}
@@
 \end{align*}
-
-\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the mean of the Poisson distribution.
 In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
-The Fourier terms are as defined in} @hyndman2021forecasting \textcolor{teal}{(Section 7.4).
-This model takes account of weekly seasonality and annual seasonality.
+The Fourier terms are as defined in} @hyndman2021forecasting [Section 7.4] \textcolor{teal}{This model takes account of weekly seasonality and annual seasonality.
 Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
@@
 
-\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional component accounting for serial correlation (i.e.
-autoregressive).
-The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points.
+\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional autoregressive component accounting for serial correlation.
+The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points, after accounting for the exogenous covariates.
 Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
 TSGLM can be written as}
-
 \begin{align*}
@@
   \text{where}\qquad
-  \log(\lambda_t) &= (\bm{y}_{t-k}'+\bm{x}_t')\bm{\beta},
+  \log(\lambda_t) &= (\bm{y}_{t-k}' , \bm{x}_t')\bm{\beta},
 \end{align*}
-
-\textcolor{teal}{${y}_{t-k}$ is a vector of k lagged values.
+\textcolor{teal}{and ${y}_{t-k}$ is a vector of $k$ lagged values.
 The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
@@
 \textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
-Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
+Again, the logarithmic link function is used to ensure that the mean of the Poisson distribution is always positive.}
 
-\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution} [@feller1991introduction, p156-158].
+\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution} [@feller1991introduction, p156--158].
 \textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
@@
 
-**Ensemble method:** \textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by averaging the forecasts, rather than relying on separate forecasts generated by each individual method} [@clemen1989combining].
-\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models using a simple average to form a mixture distribution} [@combinations].
+**Ensemble method:** \textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by combining the forecasts, rather than relying on separate forecasts generated by each individual method} [@clemen1989combining].
+\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Stationary, ETS, GLM and TSGLM models using a simple average to form a mixture distribution} [@combinations].
 
 \textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in} @panagiotelis2023probabilistic.
-\textcolor{teal}{This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+\textcolor{teal}{This involves simulating 1000 future sample paths from each of the models, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
 In this way, we can generate an empirical distribution of forecasts for each model.
@@
 
-\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models or promote a particular model class.
+\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models, or to promote a particular model class.
 Instead, we have developed a flexible framework that can accommodate any forecasting models.
@@
 
-\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by j, representing the number of time periods ahead we are predicting.
+\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by $j$, representing the number of time periods ahead we are predicting.
 In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
@@
 
-\textcolor{teal}{Using of scale-independent measures, such as MASE and MSSE, enables more significant comparisons between time series at different levels and scales, as these measures are not influenced by the data's magnitude.
-This aspect holds particular importance in our study, as we work with time series at various levels of hierarchy.
-These time series show varying scales, resulting in different magnitudes of error.
+\textcolor{teal}{Using scale-independent measures, such as MASE and MSSE, enables more appropriate comparisons between time series at different levels and scales, as these measures are not influenced by the magnitude of the data.
+This is of particular importance in our study, as we work with time series at various levels of hierarchy, with varying scales, resulting in different magnitudes of error.
 By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
@@
 \textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
-It measures how well the predicted probabilities match the actual.
+It measures how well the predicted probabilities match the observations.
 On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
-A model is well-calibrated if the predicted probabilities match the actual, and it is sharp if it is confident in its predictions.
+A model is well-calibrated if the predicted probabilities match the distribution of the observations, and it is sharp if it is confident in its predictions.
 The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
@@
 
-\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores} [@winkler1972decision;@hyndman2021forecasting, Section 5.9] \textcolor{teal}{or percentile scores} [@hyndman2021forecasting, section 5.9], \textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.} \textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
+\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores} [@winkler1972decision;@hyndman2021forecasting, Section 5.9] \textcolor{teal}{or percentile scores} [@hyndman2021forecasting, Section 5.9], \textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.} \textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
 Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.

### 2023-07-19 manuscript/04-htsgts.qmd
Minor rewording, typos and fixed some refs

@@
 and $\bm{A}$ is the $n_a\times n_b$ "aggregation" matrix specifying how the bottom-level series $\bm{b}_t$ are to be aggregated to form $\bm{a}_t$.
-\textcolor{teal}{The aggregation matrix $A$ is determined by the structure of the hierarchy.
+\textcolor{teal}{The aggregation matrix $\bm{A}$ is determined by the structure of the hierarchy.
 It maps the bottom-level time series to the corresponding higher-level time series.
-The matrix $A$ is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
-For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $A = \begin{bmatrix} 1 1 \end{bmatrix}$.
+For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $\bm{A} = \begin{bmatrix} 1 ~~ 1 \end{bmatrix}$.
 }
-
 The full vector of time series is given by
@@
 
-\textcolor{teal}{The term "level bottom series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
+\textcolor{teal}{The term "bottom-level series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
 For instance, in Table 2, each distinct combination of values in Control area (e.g.
@@
 Green), and Nature of incident (e.g.
-Chest pain) corresponds to one individual time series.
+Chest pain), corresponds to one individual time series.
 In the dataset at hand, there are 691 unique combinations, resulting in 691 bottom level time series.
@@
 all country level), the time series are aggregated across all Control areas, Health boards, Priorities, and Natures of incidents.
-Any desired aggregation level can be achieved based on the data structure, utilizing the bottom level series available.}
+Any desired aggregation level can be achieved based on the data structure, utilizing the bottom-level series available.}
 
@@
 Bottom-Up is a simple approach to generate coherent forecasts.
-It involves first creating the base forecasts for the bottom level series (i.e., the most disaggregated series).
+It involves first creating the base forecasts for the bottom-level series (i.e., the most disaggregated series).
 These forecasts are then aggregated to the upper levels which naturally results in coherent forecasts.
@@
 \textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
-In this approach, the estimation of W is based on the assumption that all the errors have equal variance.
-Hence, W is simply defined as the identity matrix multiplied by a constant factor.
+In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors have equal variance.
+Hence, $\bm{W}$ is simply defined as the identity matrix multiplied by a constant factor.
 The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
@@
 But it assumes that the errors of each series are uncorrelated with each other.
-In this approach, W is defined as a diagonal matrix with the variance of the errors on the diagonal.
+In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal.
 The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
 The main weakness of this approach is that it ignores the relationships between series.
-A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.}
+A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}
 
-\textcolor{teal}{Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the base forecast errors.
 So it takes account of both the scale of each series, and the relationships between the series.} @WicEtAl2019 \textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.

### 2023-07-19 manuscript/05-results.qmd
Minor rewording, typos and fixed some refs

--- 2023-07-18 manuscript/05-results.qmd
@@
 
-In this section, we compare the forecasting performance of the Naive, ETS, GLM, and TSGLM models along with the ensemble, using base forecast and Minimum Trace (MinT) reconciliation methods.
+In this section, we compare the forecasting performance of the Stationary, ETS, GLM, and TSGLM models along with the ensemble, using base forecast and Minimum Trace (MinT) reconciliation methods.
 We have also computed the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least Square (WLS) approaches, along with bottom-up forecasting.
@@
 Reported forecast accuracy is averaged across all forecast horizons, rolling origins, and series at each level.
-@tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom level series.
+@tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom-level series.
 Point forecast performance is reported using MASE and MSSE in @tbl-result-1  and [-@tbl-result-2], respectively.
@@
 MinT) improves forecast accuracy at the higher levels of the hierarchy including total, control area and health board.
-However, it does not result in accuracy improvement at the bottom level series, for which base forecasts are more accurate.
+However, it does not result in accuracy improvement at the bottom-level series, for which base forecasts are more accurate.
 This might be due to the noisy structure of time series at the bottom level, and possibly due to very different patterns in the aggregated series.
@@
 However, this does not remain valid for bottom series where different individual methods perform best, depending on the accuracy measure.
-\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness may not always be guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
 Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
@@
     levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
-    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
+    labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT")),
@@
       levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
-      labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
+      labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
@@
     levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
-    labels = c("Naive", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
+    labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
@@
 Forecasts are generateled every 42 days, therefore we use 11 samples to calculate the average accuracy.
-CRPS is relative to a naive Empirical Cumulative Distribution Function (ECDF).
+CRPS is relative to a stationary Empirical Cumulative Distribution Function (ECDF).
 MASE and MSSE are relative to the corresponding values for the training set."
@@
   mutate(model = case_when(
-    model == "naiveecdf" ~ "Naive",
+    model == "naiveecdf" ~ "Stationary",
     model == "ets" ~  "ETS",
@@
     series = factor(series, levels = c("Total", "Control areas", "Health boards")),
-    model = factor(model, levels = c("Naive", "ETS", "TSGLM", "GLM", "Ensemble", "Ensemble2")),
+    model = factor(model, levels = c("Stationary", "ETS", "TSGLM", "GLM", "Ensemble", "Ensemble2")),
     week = factor(trunc((h - 1) / 7) + 1)
@@
 For both the point forecast and distributional accuracy  we can see that the ensemble approach performs best across almost all horizons, with the biggest differences at the highest levels of aggregation.
-It is important to highlight that, all forecasting models outperform the Naive empirical distribution that is used as a benchmark for both point and probabilistic forecasts.
+It is important to highlight that, all forecasting models outperform the stationary empirical distribution that is used as a benchmark for both point and probabilistic forecasts.
 
@@
 However, the impact of rounding on forecast accuracy varies depending on the level of hierarchy and the scale of the data.
-In situations with high-volume demand, the effects of rounding may become negligible, and forecast accuracy calculations can overlook integer effects.
+In situations with high-volume demand, the effects of rounding may be negligible, and forecast accuracy calculations can overlook integer effects.
 On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.}
@@
 
-@fig-forecast-density \textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, regarding the total attended incidents in the PO health board.
-The black dashed line in the graph represents the point forecast, which presents a single numerical estimate for the future number of attended incidents.
+@fig-forecast-density \textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, of the total attended incidents in the PO health board.
+The black dashed line in the graph represents the point forecast (the mean of the distribution), which presents a single numerical estimate for the future number of attended incidents.
 Due to the complexity of including such plots for the entire hierarchy and 84 days ahead, only one example is presented here for 1 day ahead.
@@
 \textcolor{teal}{The purpose of probabilistic forecasting, as demonstrated in} @fig-forecast-density, \textcolor{teal}{is to quantify uncertainty.
-Decisions based on these forecasts could focus on the tails of the distribution: unexpected high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
+Decisions based on these forecasts could focus on the tails of the distribution: unexpectedly high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
 Such forecasts are valuable tools for decision-makers and planners, especially when dealing with low-probability, high-cost situations.
@@
 #| out.width: "100%"
-#| fig-cap: "An illustrative example of forecast distribution of ambulance demand (i.e.
-total incidence attended) in PO health board for 1 day ahead.
-x-axis shows all possible outcomes that may occur, with thier likelihhod in y-aixs"
+#| fig-cap: "An illustrative example of the forecast distribution of ambulance demand (i.e.
+total incidence attended) for the PO health board for 1 day ahead.
+The horizontal axis shows all possible outcomes that may occur, with their likelihood shown on the vertical axis.
+The dashed line indicates the mean of the distribution."
 #| fig-pos: "H"

### 2023-07-19 manuscript/06-conclusion.qmd
Minor rewording, typos and fixed some refs

@@
 This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
-For instance, in the supply chains } [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
+For instance, in supply chains} [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
 Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency.

### 2023-07-19 manuscript/main.tex
Minor rewording, typos and fixed some refs

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 making and coordination. \citet{al2021empirical}
-\textcolor{teal}{utilises real data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand accurately.
+\textcolor{teal}{utilises data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand.
 The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
-Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE) are calculated.
+Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE).
 The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
@@
 \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
-Real-life call centre data was utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
+Call centre data were utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
 The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
@@
 Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO\\
-Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & Yes & NO\\
+Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO\\
 Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO\\
@@
 & 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO\\
-Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & No & NO\\
+Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO\\
 Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO\\
@@
 \textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by}
-\citet{kang2017visualising}.
+\citet{mstl}.
 \textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
@@
 In addition to displaying the trend and seasonality features, we also
-visualize few time series at various levels of the aggregation.
+visualize a few time series at various levels of aggregation.
 Figure~\ref{fig-dataviz2} reveals different information such as trend,
@@
 
-\textbf{Naive:} We start with a simple forecasting approach, assuming
-that the future days will be similar to past days. We use the empirical
-distribution of the past daily attended incidents to create the forecast
-distribution of future attended incidents.
-\textcolor{teal}{We have chosen the naive method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+\textbf{Stationary:} We start with a simple forecasting approach,
+assuming that the future days will be similar to past days. We use the
+empirical distribution of the past daily attended incidents to create
+the forecast distribution of future attended incidents.
+\textcolor{teal}{We have chosen this "stationary" method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
 Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
-Additionally, the naive method has shown surprisingly high accuracy.
-Hence, any forecasting approach that can offer superior results compared to the naive method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
+Additionally, the stationary method has shown surprisingly high accuracy.
+Hence, any forecasting approach that can offer superior results compared to the stationary method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
 
@@
 \end{align*}
-
-\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the mean of the Poisson distribution.
 In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
 The Fourier terms are as defined in}
-\citet{hyndman2021forecasting}
-\textcolor{teal}{(Section 7.4).
-This model takes account of weekly seasonality and annual seasonality.
+\citet[Section 7.4]{hyndman2021forecasting}
+\textcolor{teal}{This model takes account of weekly seasonality and annual seasonality.
 Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
@@
 \citep{JSSv082i05}
-\textcolor{teal}{to model the attended incidents. Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
+\textcolor{teal}{to model the attended incidents. Again, the logarithmic link function is used to ensure that the mean of the Poisson distribution is always positive.}
 
 \textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution}
-\citep[p156-158]{feller1991introduction}.
+\citep[p156--158]{feller1991introduction}.
 \textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
@@
 \textbf{Ensemble method:}
-\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by averaging the forecasts, rather than relying on separate forecasts generated by each individual method}
+\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by combining the forecasts, rather than relying on separate forecasts generated by each individual method}
 \citep{clemen1989combining}.
-\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models using a simple average to form a mixture distribution}
+\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Stationary, ETS, GLM and TSGLM models using a simple average to form a mixture distribution}
 \citep{combinations}.
@@
 \citet{panagiotelis2023probabilistic}.
-\textcolor{teal}{This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+\textcolor{teal}{This involves simulating 1000 future sample paths from each of the models, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
 In this way, we can generate an empirical distribution of forecasts for each model.
@@
 \citep[Section 5.9]{winkler1972decision, hyndman2021forecasting}
-\textcolor{teal}{or percentile scores} \citep[section
+\textcolor{teal}{or percentile scores} \citep[Section
 5.9]{hyndman2021forecasting},
@@
 to form \(\bm{a}_t\).
-\textcolor{teal}{The aggregation matrix $A$ is determined by the structure of the hierarchy.
+\textcolor{teal}{The aggregation matrix $\bm{A}$ is determined by the structure of the hierarchy.
 It maps the bottom-level time series to the corresponding higher-level time series.
-The matrix $A$ is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
-For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $A = \begin{bmatrix} 1 1 \end{bmatrix}$.
-}
-
-The full vector of time series is given by \[
+For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $\bm{A} = \begin{bmatrix} 1 ~~ 1 \end{bmatrix}$.
+} The full vector of time series is given by \[
  \bm{y}_t = \begin{bmatrix}\bm{a}_t \\\bm{b}_t\end{bmatrix}.
@@
 Bottom-Up is a simple approach to generate coherent forecasts. It
-involves first creating the base forecasts for the bottom level series
+involves first creating the base forecasts for the bottom-level series
 (i.e., the most disaggregated series). These forecasts are then
@@
 But it assumes that the errors of each series are uncorrelated with each other.
-In this approach, W is defined as a diagonal matrix with the variance of the errors on the diagonal.
+In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal.
 The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
 The main weakness of this approach is that it ignores the relationships between series.
-A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.}
-
-\textcolor{teal}{Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}
+
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the base forecast errors.
 So it takes account of both the scale of each series, and the relationships between the series.}
@@
 
-In this section, we compare the forecasting performance of the Naive,
-ETS, GLM, and TSGLM models along with the ensemble, using base forecast
-and Minimum Trace (MinT) reconciliation methods. We have also computed
-the forecast accuracy for Ordinary Least Square (OLS) and Weighted Least
-Square (WLS) approaches, along with bottom-up forecasting. However, they
-are not reported here because their accuracy is outperformed by MinT. We
-should also note that forecasts, and consequently their corresponding
-errors, are generated for the entire hierarchy and they could be
-reported at any level, if required. But to save space, we have reported
-only the top level (Total), the bottom level, and the levels
-corresponding to Control areas and Health boards. The latters are chosen
-because this is where decision-making takes place, so these forecasts
-are the most important.
+In this section, we compare the forecasting performance of the
+Stationary, ETS, GLM, and TSGLM models along with the ensemble, using
+base forecast and Minimum Trace (MinT) reconciliation methods. We have
+also computed the forecast accuracy for Ordinary Least Square (OLS) and
+Weighted Least Square (WLS) approaches, along with bottom-up
+forecasting. However, they are not reported here because their accuracy
+is outperformed by MinT. We should also note that forecasts, and
+consequently their corresponding errors, are generated for the entire
+hierarchy and they could be reported at any level, if required. But to
+save space, we have reported only the top level (Total), the bottom
+level, and the levels corresponding to Control areas and Health boards.
+The latters are chosen because this is where decision-making takes
+place, so these forecasts are the most important.
 
@@
 point and probabilistic forecast accuracy at total, control area, health
-board and bottom level series. Point forecast performance is reported
+board and bottom-level series. Point forecast performance is reported
 using MASE and MSSE in Table~\ref{tbl-result-1} and \ref{tbl-result-2},
@@
 levels of the hierarchy including total, control area and health board.
-However, it does not result in accuracy improvement at the bottom level
+However, it does not result in accuracy improvement at the bottom-level
 series, for which base forecasts are more accurate. This might be due to
@@
 perform best, depending on the accuracy measure.
-\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness may not always be guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
 Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
@@
 \midrule
-Base & Naive & 1.139 & 1.059 & 1.047 & 1.019\\
+Base & Stationary & 1.139 & 1.059 & 1.047 & 1.019\\
 Base & ETS & 0.963 & 0.930 & 0.899 & 1.038\\
@@
 \addlinespace
-MinT & Naive & 1.138 & 1.059 & 1.047 & 2.651\\
+MinT & Stationary & 1.138 & 1.059 & 1.047 & 2.651\\
 MinT & ETS & 0.877 & 0.916 & 0.915 & 1.289\\
@@
 \midrule
-Base & Naive & 1.169 & 1.056 & 1.062 & 1.031\\
+Base & Stationary & 1.169 & 1.056 & 1.062 & 1.031\\
 Base & ETS & 0.979 & 0.875 & 0.816 & \textbf{0.975}\\
@@
 \addlinespace
-MinT & Naive & 1.168 & 1.057 & 1.062 & 2.095\\
+MinT & Stationary & 1.168 & 1.057 & 1.062 & 2.095\\
 MinT & ETS & 0.785 & 0.852 & 0.845 & 0.994\\
@@
 \midrule
-Base & Naive & 30.387 & 10.882 & 5.500 & 0.302\\
+Base & Stationary & 30.387 & 10.882 & 5.500 & 0.302\\
 Base & ETS & 14.309 & 6.074 & 3.476 & 0.244\\
@@
 \addlinespace
-MinT & Naive & 30.368 & 10.902 & 5.498 & 0.313\\
+MinT & Stationary & 30.368 & 10.902 & 5.498 & 0.313\\
 MinT & ETS & 13.515 & 5.967 & 3.547 & \textbf{0.243}\\
@@
 days, therefore we use 11 samples to calculate the average accuracy.
-CRPS is relative to a naive Empirical Cumulative Distribution Function
-(ECDF). MASE and MSSE are relative to the corresponding values for the
-training set.}
+CRPS is relative to a stationary Empirical Cumulative Distribution
+Function (ECDF). MASE and MSSE are relative to the corresponding values
+for the training set.}
 
@@
 of aggregation. It is important to highlight that, all forecasting
-models outperform the Naive empirical distribution that is used as a
-benchmark for both point and probabilistic forecasts.
+models outperform the stationary empirical distribution that is used as
+a benchmark for both point and probabilistic forecasts.
 
@@
 Figure~\ref{fig-forecast-density}
-\textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, regarding the total attended incidents in the PO health board.
-The black dashed line in the graph represents the point forecast, which presents a single numerical estimate for the future number of attended incidents.
+\textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, of the total attended incidents in the PO health board.
+The black dashed line in the graph represents the point forecast (the mean of the distribution), which presents a single numerical estimate for the future number of attended incidents.
 Due to the complexity of including such plots for the entire hierarchy and 84 days ahead, only one example is presented here for 1 day ahead.
@@
 \textcolor{teal}{is to quantify uncertainty.
-Decisions based on these forecasts could focus on the tails of the distribution: unexpected high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
+Decisions based on these forecasts could focus on the tails of the distribution: unexpectedly high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
 Such forecasts are valuable tools for decision-makers and planners, especially when dealing with low-probability, high-cost situations.
@@
 
-\caption{\label{fig-forecast-density}An illustrative example of forecast
-distribution of ambulance demand (i.e.~total incidence attended) in PO
-health board for 1 day ahead. x-axis shows all possible outcomes that
-may occur, with thier likelihhod in y-aixs}
+\caption{\label{fig-forecast-density}An illustrative example of the
+forecast distribution of ambulance demand (i.e.~total incidence
+attended) for the PO health board for 1 day ahead. The horizontal axis
+shows all possible outcomes that may occur, with their likelihood shown
+on the vertical axis. The dashed line indicates the mean of the
+distribution.}
 
@@
 This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
-For instance, in the supply chains }
+For instance, in supply chains}
 \citep{shugan2000advance},

### 2023-07-19 journal_correspondence/response_letter/response_letter_JSR.qmd
edit response letter based on Rob's chnages

@@
 We also included two additional references in the table and added the relevant texts in the literature section.
+@al2021empirical \textcolor{teal}{utilises data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand.
+The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
+Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE).
+The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
+However, when it comes to long-term demand, the SSA method proves to be the most effective.} @ibrahim2016modeling \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
+Call centre data were utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
+The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
+The findings indicate that the ME consistently produces the most accurate point forecasts.
+On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
+For a two-week leading period, MU shows better coverage probability.}
 
@@
 ```
-
-@al2021empirical \textcolor{teal}{utilises real data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand accurately.
-The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
-Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE) are calculated.
-The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
-However, when it comes to long-term demand, the SSA method proves to be the most effective.} @ibrahim2016modeling \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
-Real-life call centre data was utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
-The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
-The findings indicate that the ME consistently produces the most accurate point forecasts.
-On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
-For a two-week leading period, MU shows better coverage probability.}
 
@@
 
-\textcolor{teal}{The dataset used in this study only includes information on attended incidents.
-It would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
+\textcolor{teal}{Also, the dataset used in this study only includes information on attended incidents, it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
 It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.}
@@
 
-\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchy are necessary for effective decision-making in EMS.
+\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchies are necessary for effective decision-making in EMS.
 For example, control area EMS forecasts can inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals.
@@
 Our focus is on developing a hierarchical time series forecasting approach that takes into account the hierarchical and group structure of data.
-We have selected the naive method as a benchmark because it is widely used in the forecasting literature and practice and is a simple and easily understandable method for users.
+We changed the name of the method to "Stationary", to avoid confusion with Naive method that is known in time series forecasting.
+We have selected the Stationary method as a benchmark because it is widely used in the forecasting literature and practice and is a simple and easily understandable method for users.
 
 Forecasts are often used as input for other decision-making systems that often involve simulation, and it is quite common to use the empirical distribution of the demand as a forecast in simulations.
-Also, the naive method can be surprisingly accurate.
+Also, the Stationary method can be surprisingly accurate.
 Therefore, if any forecasting method should be able to provide more accurate results than naive, it would justify its use in practice, if not there is no need to use more complicated methods.
 
-\textcolor{teal}{We have chosen the naive method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+**Stationary:** We start with a simple forecasting approach, assuming that the future days will be similar to past days.
+We use the empirical distribution of the past daily attended incidents to create the forecast distribution of future attended incidents.
+\textcolor{teal}{We have chosen this "stationary" method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
 Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
-Additionally, the naive method has shown surprisingly high accuracy.
-Hence, any forecasting approach that can offer superior results compared to the naive method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
+Additionally, the stationary method has shown surprisingly high accuracy.
+Hence, any forecasting approach that can offer superior results compared to the stationary method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
 
@@
 
-\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable form for the trend, seasonality, and error terms in each time series.
-Specifically, we utilize the `ETS()` function in the fable package of R, which employs statistical criteria like Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ETS()` function in the fable package of R, which employs Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
 Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
-Consequently, the automated algorithm selects the most fitting components based on the unique characteristics of each individual time series.
+Consequently, the automated algorithm selects the best model based on the unique characteristics of each individual time series.
 As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
@@
 
-\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the parameter of Poisson distribution.
+\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the mean of the Poisson distribution.
 In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
-The Fourier terms are as defined in} @hyndman2021forecasting \textcolor{teal}{(Section 7.4).
-This model takes account of weekly seasonality and annual seasonality.
+The Fourier terms are as defined in} @hyndman2021forecasting [Section 7.4] \textcolor{teal}{This model takes account of weekly seasonality and annual seasonality.
 Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
@@
 
-\textcolor{teal}{The Poisson TSGLM is similar to GLM with an aditional component accounting for serial dependence (i.e.
-autoregressive terms).
-The term serial dependence refers to situations where the ambulance demand at a given time exhibits correlation with the demand at previous time intervals.
-Put simply, past values of the time series influence the current values.
+\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional autoregressive component accounting for serial correlation.
+The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points, after accounting for the exogenous covariates.
+Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
 TSGLM can be written as}
-
 \begin{align*}
@@
 
-\textcolor{teal}{We use the `tsglm()` function in the tscount package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
-Again, the logarithmic link function is used to ensure that the parameter of Poisson distribution is always positive.}
+\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
+Again, the logarithmic link function is used to ensure that the mean of the Poisson distribution is always positive.}
 
@@
 
-\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by j, representing the number of time periods ahead we are predicting.
+\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by $j$, representing the number of time periods ahead we are predicting.
 In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
 
-\textcolor{teal}{Using of scale-independent measures, such as MASE and MSSE, enables more significant comparisons between time series at different levels and scales, as these measures are not influenced by the data's magnitude.
-This aspect holds particular importance in our study, as we work with time series at various levels of hierarchy.
-These time series show varying scales, resulting in different magnitudes of error.
+\textcolor{teal}{Using scale-independent measures, such as MASE and MSSE, enables more appropriate comparisons between time series at different levels and scales, as these measures are not influenced by the magnitude of the data.
+This is of particular importance in our study, as we work with time series at various levels of hierarchy, with varying scales, resulting in different magnitudes of error.
 By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
@@
 \textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in} @panagiotelis2023probabilistic.
-\textcolor{teal}{This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+\textcolor{teal}{This involves simulating 1000 future sample paths from each of the models, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
 In this way, we can generate an empirical distribution of forecasts for each model.
@@
 But it assumes that the errors of each series are uncorrelated with each other.
-In this approach, W is defined as a diagonal matrix with the variance of the errors on the diagonal.
+In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal.
 The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
 The main weakness of this approach is that it ignores the relationships between series.
-A strength of WLS is that it is relatively easy to compute W as it is based only on error variances which are readily estimated.}
-
-\textcolor{teal}{Minimum Trace (MinT) is a further generalization where W is defined as the covariance matrix of the base forecast errors.
+A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}
+
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the base forecast errors.
 So it takes account of both the scale of each series, and the relationships between the series.} @WicEtAl2019 \textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
@@
 
-\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness may not always be guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
 Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
@@
 
-\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by averaging the forecasts, rather than relying on separate forecasts generated by each individual method} [@clemen1989combining].
-\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Naive, ETS, GLM and TSGLM models using a simple average to form a mixture distribution} [@combinations].
+\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by combining the forecasts, rather than relying on separate forecasts generated by each individual method} [@clemen1989combining].
+\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Stationary, ETS, GLM and TSGLM models using a simple average to form a mixture distribution} [@combinations].
 
@@
 
-\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution} [@feller1991introduction, p156-158].
+\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution} [@feller1991introduction, p156--158].
 \textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
@@
 
-\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by Kang et al (2017).
-STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
+\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by} @mstl.
+\textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
 The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
@@
 
-\textcolor{teal}{The aggregation matrix $A$ is determined by the structure of the hierarchy.
+\textcolor{teal}{The aggregation matrix $\bm{A}$ is determined by the structure of the hierarchy.
 It maps the bottom-level time series to the corresponding higher-level time series.
-The matrix $A$ is typically predefined based on the structure of the hierarchy, and it remains fixed throughout the forecasting process.
-For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $A = \begin{bmatrix} 1 1 \end{bmatrix}$.
+For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $\bm{A} = \begin{bmatrix} 1 ~~ 1 \end{bmatrix}$.
 }
@@
 This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
-For instance, in the supply chains} [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
+For instance, in supply chains} [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
 Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency.
-In the financial industry } [@kimes1998strategic] \textcolor{teal}{, where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
+In the financial industry} [@kimes1998strategic], \textcolor{teal}{where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
 Similarly, in transportation, the framework supports forecasting transportation demand at various levels, optimizing route planning and resource allocation.

### 2023-07-19 manuscript/05-results.qmd
edit response letter based on Rob's chnages

@@
 
-acc_summary |>
+acc_summary |>   filter(
+    model != "Ensemble2"
+  ) |> 
   ggplot(aes(x = week, y = accuracy, group = model, col = model)) +

### 2023-07-19 manuscript/main.tex
edit response letter based on Rob's chnages

@@
   pdftitle={Hierarchical Time Series Forecasting in Emergency Medical Services},
-  pdfauthor={Bahman Rostami-Tabar; Rob J. Hyndman},
   pdfkeywords={healthcare, emergency services, forecast
@@
 Services}
-\author[1]{Bahman Rostami-Tabar%
-\corref{cor1}%
-}
- \ead{rostami-tabarb@cardiff.ac.uk} 
-\author[2]{Rob J. Hyndman%
-%
-}
- \ead{Rob.Hyndman@monash.edu}
-
-\affiliation[1]{organization={Cardiff University, Cardiff Business
-School},country={United Kingdom},countrysep={,},postcode={CF10
-3EU},postcodesep={}}
-\affiliation[2]{organization={Monash University, Department of
-Econometrics and Business
-Statistics},country={Australia},countrysep={,},postcode={VIC
-3800},postcodesep={}}
 
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, frame hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi
 

### 2023-07-19 manuscript/main.tex
update naive to Stationary in literature table

--- 2023-07-19 manuscript/main.tex
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, frame hidden, boxrule=0pt, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, breakable, sharp corners, boxrule=0pt, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 \midrule
-Current study & 2023 & Ambulance demand & 84 days & Naïve;  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES\\
+Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES\\
 Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO\\
@@
 R code to produce all results in this paper is available at
+https://github.com/bahmanrostamitabar/forecasting-emergency-medcine
 

### 2023-07-22 journal_correspondence/response_letter/response_letter_JSR.qmd
resubmit word document

@@
 Furthermore, we have thoroughly addressed the specific suggestions provided by Reviewers 1 and 2 in the revised manuscript.
-Additionally, we have considered and responded to Reviewer 3's remarks concerning the performance of the Naïve model and Figure 4.
+Additionally, we have conducted further analysses and responded to Reviewer 3's remarks concerning the performance of the Naïve model (name changed to Stationary in the revised version) and Figure 4.
 To emphasize the significance of our study, we have included additional paragraphs in the paper that highlight its importance and potential extensions.
@@
 This highlights the wide applicability and potential impact of our research in various contexts.
+
+Our manuscript is written using [Quarto](https://quarto.org/), an open-source scientific and technical publishing system designed for publishing reproducible and production-quality articles.
+To ensure transparency and accessibility, we have established a GitHub repository containing all the necessary files, analyses, R codes, and related materials used in creating the article.
+Currently, the repository link remains private to avoid revealing the authors' identities in the paper.
+However, upon request, we are prepared to share this link with the Editor, Associate Editor, and all reviewers.
+Once the paper is accepted for publication, we will include the repository link in the article to facilitate further exploration and validation of our work.
 
@@
 We have responded and implemented suggestion provided by R1 and R2 in the revised manuscript.
-We have also responded to R3's remark regarding the performance of the Naïve model and Figure 4.
+We have also conducted further analyses to respond to R3's remark regarding the performance of the Naïve model (name changed to Stationary in the revised version) and Figure 4.
 We have also included additional paragraphs in the paper to highlight the importance of the current study and how it can be extended to incorporate the decision making process and evaluate the implication of forecasts on decision impacts and utilities important to planners such as cost, allocating resources, etc.
-Finally, we added a paragraph to highlight how the current study can be generalized to other areas in service industries.
+Finally, we added a paragraph to the end of the conclusion section to highlight how the current study can be generalized to other areas in service industries.
+
+Our manuscript is written using [Quarto](https://quarto.org/), an open-source scientific and technical publishing system designed for publishing reproducible and production-quality articles.
+To ensure transparency and accessibility, we have established a GitHub repository containing all the necessary files, analyses, R codes, and related materials used in creating the article.
+Currently, the repository link remains private to avoid revealing the authors' identities in the paper.
+However, upon request, we are prepared to share this link with the Editor, Associate Editor, and all reviewers.
+Once the paper is accepted for publication, we will include the repository link in the article to facilitate further exploration and validation of our work.
 

### 2023-07-22 manuscript/02-literature.qmd
resubmit word document

@@
 
-We note a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research. These limitations are summarized as following:
+@tbl-literature provides a summary of some studies in the literature on forecasting in Emergency Medical Services.
+We note a number of limitations in the literature of EMS forecasting, that encourage us to undertake this research.
+These limitations are summarized as following:
 
@@
 
+This paper concerns the problem of hierarchical forecasting in EMS and generates and evaluates both point and probabilistic forecast across different levels of the hierarchy, hence addressing some important gaps identified in the literature.
+
 ```{r summarylit, echo=FALSE, cache=FALSE}
+#| label: tbl-literature
 library(kableExtra)
 li <- readxl::read_xlsx("../literature_table/table.xlsx")
-knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies in forecasting FOR AMBULANCE SERVICES") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
+knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies on forecasting in Emergency Medical Services") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
   column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |>  kableExtra::landscape()

### 2023-07-22 manuscript/main.qmd
add the Github repo
resubmit word document

@@
 title: "Hierarchical Time Series Forecasting in Emergency Medical Services"
-author:
-  - name: Bahman Rostami-Tabar
-    email: rostami-tabarb@cardiff.ac.uk
-    affiliations:
-        - id: carbs
-          name: Cardiff University
-          department: Cardiff Business School
-          country: United Kingdom
-          postal-code: CF10 3EU
-    attributes:
-        corresponding: true
-  - name: Rob J. Hyndman
-    email: Rob.Hyndman@monash.edu
-    affiliations:
-        - id: monash
-          name: Monash University
-          department: Department of Econometrics and Business Statistics
-          country: Australia
-          postal-code: VIC 3800
+# author:
+#   - name: Bahman Rostami-Tabar
+#     email: rostami-tabarb@cardiff.ac.uk
+#     affiliations:
+#         - id: carbs
+#           name: Cardiff University
+#           department: Cardiff Business School
+#           country: United Kingdom
+#           postal-code: CF10 3EU
+#     attributes:
+#         corresponding: true
+#   - name: Rob J. Hyndman
+#     email: Rob.Hyndman@monash.edu
+#     affiliations:
+#         - id: monash
+#           name: Monash University
+#           department: Department of Econometrics and Business Statistics
+#           country: Australia
+#           postal-code: VIC 3800
 abstract: |
@@
 
-R code to produce all results in this paper is available at
+R code to produce all results in this paper is available at https://github.com/bahmanrostamitabar/forecasting-emergency-medcine
 

### 2023-11-06 journal_correspondence/response_letter/R2_response_letter_JSR.qmd
response letetr R2

---
title: "Response letter - Manuscript ID JSR-22-468.R1"
bibliography: bibliography.bib
format:
  pdf:
    pdf-engine: pdflatex
    papersize: a4
    geometry: "margin = 2.5cm"
    include-in-header: header.tex
---

Dear Professor Malthouse

Please find enclosed a revised version of our paper submitted for peer review for the *Journal of Service Research*. Your comments have been very much appreciated and we have tried to suitably modify/amend our paper for those comments to be reflected in the revised version.

We quote below the detailed referees’ comments that are followed by our response. In all cases, we point out, where necessary, the corresponding amendments in the new version of the paper and highlighted them in teal (greenish-blue color).

# Editor comment

*My decision on this manuscript was not easy. On the one hand, the reviewers rate the importance of the work as "important" and "very important." They also see "major" potential impact. They acknowledge that you made major changes to the first version of the MS. I personally appreciate you highlighting the changes you made to the manuscript in a blue font and see your substantial improvements to R1. On the other hand the review team feels like many comments from the first round were not addressed. R1 and the AE are asking for model details, clarifications and elaborations. They continue to find typos (R2 notes issues with subscripts and both reviewers mention formatting issues). Simple suggestions like mentioning extended application areas in the introduction, which will help in the positioning of the article for JSR readers, were ignored. They appreciate the example you have added, but suggest adding details. I see these comments as constructive in that their goal is to improve the readability of your manuscript for the JSR and other readers, which will ultimately mean that readers will understand your work so that it realizes its full potential impact. While many of our readers have strong methodological training, JSR is also read by more managerial and policy-oriented scholars; it is important that articles are written in a way to reach both audiences. This expands the reach and impact of your work.*

*In order to gain final acceptance, I ask for you to go back and address the clarifications, details, etc. Please write a point-by-point response and highlight the changes in the manuscript by using a different colored font. I do not plan to send this out for review again. I reserve the right to reject the work if there is not sufficient progress made in addressing the issues identified by the review team in the next submission. (Bahman)*

Response:

# Associate Editor

We received two very good reviews that came up with justified concern and mixed recommendations. Let me add additional comments.

## AE comments

_Comment 1. The authors need to score higher on the economic insights. Their example did not impress me, and I was hoping to receive more intuition as to why their approach works better. (Rob?)_

  **Response:**

_Comment 2. The authors miss opportunities. I appreciate that the authors summarize previous research in Table 1. Yet, there is nothing in Table 1 that helps the reader understand why the current research has the limitations that the authors state on page 8 (Bahman)._

  **Response:**

_Comment 3. I highly appreciate that the authors intend to provide data and code via their GitHub repository. No doubt, the resulting transparency is a major strength of this manuscript. Yet, the authors could still do a much better job in outlining to the readers that the authors have done a good job. (Rob?)_

  **Response:**

_Comment 4.I felt that the authors did not address the following concern that I outlined in the last round of review: “It would also help to have a rather detailed example, maybe even a numerical example, to better describe why separate forecasts at different levels can cause problems. Right now, it sounds plausible that such problems occur, but you are still rather vague concerning the details of these problems. (Rob?)”_

  **Response:**

_Comment 5. I also had the following comments in the previous round of review: “Make sure that all tables and figures are self-contained” and “If possible, display the number of observations in all tables and figures with statistical results”. Unfortunately, Figure 1 is not self-contained (e.g., abbreviations are not explained). Figure 2 could also contain the correlation and the number of observations. (Bahman)_

  **Response:**

_Comment 6. I was also hoping for an example that not only describes the problem that the authors solve, but that allows us to better understand the basic idea behind the authors’ solution. (Rob?)_

  **Response:**

# Reviewer 1

_**Overall Review:** _

_The authors have made revisions to address some of the concerns raised in the initial review, which has improved certain aspects of the manuscript. Additions have been made to explain the trend and seasonality strength measures, provide the GLM equation, and resolve minor formatting issues. These changes are beneficial. (Bahman)_

_However, critical concerns around model specifics, validation methodology, results analysis, and terminology remain unclear or insufficiently addressed. Major issues persist with detailing the ETS configurations, clarifying serial dependence in TSGLM, justifying the naive method’s inclusion, explaining training/testing data splits for tuning, defining the forecast horizon, elaborating the CRPS metric, providing dataset-specific examples for hierarchical forecasting concepts, distinguishing between hierarchical forecasting approaches, validating the comparison technique, and discussing practical implications of non-integer forecasts._

_The lack of clarity and details in these important methodological areas continues to undermine the credibility and reproducibility of the study. Furthermore, the authors do not sufficiently leverage the results to provide managerial insights for EMS planning. More in-depth analysis of the forecast accuracy patterns across timescales, hierarchy levels, and model configurations would strengthen the practical implications._

_In summary, while I acknowledge the efforts to improve the manuscript in certain aspects, there remains a critical need for enhancing model specifics, methodology descriptions, results analysis, and terminology usage._

  **Response:**

_Comment 1. The authors added brief explanations of the trend and seasonality strength measures in response to Comment #3. While this provides more context, a citation or reference for these specific definitions would further strengthen this section (Bahman)._

  **Response:**

_Comment 2. The revisions partially address the lack of model detail raised in Comment #4, but some components are still unclear (Bahman):_

  **Response:**

_Comment 2.1. For ETS, it is now stated that an automated algorithm selects the best configuration, but no information is given on what options are available for trend, seasonality, etc. Listing the choices would help the reader. (Bahman)._

  **Response:**

_Comment 2.2. The equation for the GLM model is now provided, but the description still does not mention monthly seasonality. This should be explicitly addressed. (Bahman)_

  **Response:**

_Comment 2.3. The concept of ”serial dependence” in TSGLM needs more clarification and specificity to the EMS context. (Bahman)_

  **Response:**

_Comment 2.4. The rationale for including the naive method in the ensemble is still not provided. Its apparent poor performance suggests it should be excluded (Rob, Bahman)._

  **Response:**

_Comment 3. The explanation of the training/validation/testing data split in Comment #5 remains unclear. The authors should explicitly state if tuning is done on the training set only or if a separate validation set is used. (Bahman)_

  **Response:**

_Comment 4. The forecast horizon explanation in Comment #6 is still ambiguous. The time unit (days, weeks, etc.) should be clearly defined. (Bahman)_

  **Response:**

_Comment 5. Comment #7 regarding the CRPS metric lacks sufficient clarification. Explicitly defining concepts like sharpness and miscalibration would strengthen this section. Additionally, more specifics are needed on the forecast distribution derivation. (Rob)_

  **Response:**

_Comment 6. Section 4.2 still lacks concrete examples tying the hierarchical forecasting concepts to the EMS dataset, as noted in Comment #8. This example-based explanation needs enhancement. (Bahman)_

  **Response:**

_Comment 7. Comment #9 about clarifying the differences between various hierarchical forecasting approaches does not seem to be addressed. This terminology is still confusing in Section 4.2.2 and requires revision. (Rob)_

  **Response:**

_Comment 8. Comment #10 questioning the validation approach of comparing base and reconciled forecasts is not addressed. The authors should explain why this is an appropriate evaluation technique. (Rob)_

  **Response:**

_Comment 9. My concerns in Comment #11 about non-integer forecasts and rounding do not appear to be addressed. This practical issue needs to be discussed. (Bahman)_

  **Response:**

# Reviewer 2

_General comment: The author revised the manuscript upon the review comments and I appreciate the detailed introductions added to the body of the paper.I provide some follow-up comments below for consideration and reference. (Bahman)_

  **Response:**

_Comment 1. Might be good to at least briefly mention the extended applicability of the idea early in the introduction section. (Bahman)_

  **Response:**

_Comment 2. It is confusing to have the abbreviations BME1 and BME2 listed with the same full name while no explanations or references are given. This is at Section 2, Page 7. (Bahman)_

  **Response:**

_Comment 3. The author added great details to the methods mentioned in Section 3.2, such as the well-explained model specifications of GLM and TSGLM. Make sure typos are corrected, especially notation-wise such as subscripts. (Bahman)_

  **Response:**

_Comment 4. It may be good to consider an alternatively way of providing details to the linear reconciliation method. For instance, the main equation \tilde{y}_h is still missing important reasoning about how it is derived. With some better descriptions of how it is derived, such as an objective function, the follow-up paragraphs for the three methods can be more concisely summarised with better readability. (Rob)_

  **Response:**

_Comment 5. The new Section 5.1 provided details about how to obtain and understand the distributional forecasts of the method. Indeed, this can be considered a major selling point of the work, and hence I would always appreciate further addressing and discussions about the probabilistic outperformance of the idea. For example, it might be good to include figures with fitting coverage of the training data, as well as some additional appendix figures for other example series probably at different hierarchy levels. It is not necessary to refer them in the main paper, but some self-explanatory captions should suffice. Besides, comparison with competing methods regarding the same aspect would be a plus. (Bahman)_

  **Response:**

_Comment 6. There are still formatting issues in the included tables. (Bahman)_

  **Response:**

# References {#references .unnumbered}

### 2023-11-06 manuscript/02-literature.qmd
response letetr R2

@@
 
-```{r summarylit, echo=FALSE, cache=FALSE}
+```{r}
 #| label: tbl-literature
+#| echo: false
+#| cache: false
+#| tbl-cap: "Summary of some studies on forecasting in Emergency Medical Services"
 library(kableExtra)
 li <- readxl::read_xlsx("../literature_table/table.xlsx")
-knitr::kable(li, booktabs = T, linesep = "",caption = "Summary of some studies on forecasting in Emergency Medical Services") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
-  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |>  kableExtra::landscape()
+knitr::kable(li, booktabs = T, linesep = "") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
+  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |> landscape()
 ```

### 2023-11-06 manuscript/main.tex
response letetr R2

--- 2023-07-19 manuscript/main.tex
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, frame hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, sharp corners, breakable, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-We note a number of limitations in the literature of EMS forecasting,
-that encourage us to undertake this research. These limitations are
+Table~\ref{tbl-literature} provides a summary of some studies in the
+literature on forecasting in Emergency Medical Services. We note a
+number of limitations in the literature of EMS forecasting, that
+encourage us to undertake this research. These limitations are
 summarized as following:
@@
 
+This paper concerns the problem of hierarchical forecasting in EMS and
+generates and evaluates both point and probabilistic forecast across
+different levels of the hierarchy, hence addressing some important gaps
+identified in the literature.
+
+\hypertarget{tbl-literature}{}
 \begin{landscape}\begin{table}[!h]
-
-\caption{Summary of some studies in forecasting FOR AMBULANCE SERVICES}
-\centering
+\caption{\label{tbl-literature}Summary of some studies on forecasting in Emergency Medical Services }\tabularnewline
+
+\centering\begingroup\fontsize{11}{13}\selectfont
+
 \resizebox{\linewidth}{!}{
-\fontsize{11}{13}\selectfont
-\begin{tabular}[t]{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}ll}
+\begin{tabular}{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}ll}
 \toprule
@@
 \end{tabular}}
+\endgroup{}
 \end{table}

### 2023-11-10 journal_correspondence/response_letter/R2_response_letter_JSR.qmd
Some revisions based on comments from AE and reviewer 2

@@
 
+We have now extended the reproducibility section to say the following:
+
+> R code to produce all results in this paper is available at [github.com/bahmanrostamitabar/forecasting-emergency-medicine](https://github.com/bahmanrostamitabar/forecasting-emergency-medicine).
+The repository contains the raw data, all R scripts used in experiments as well as the results used in this paper, along with the raw files for producing this paper.
+Full instructions are provided in the repository.
+
+[Bahman: can you please fix the spelling of medicine in the name of the repo]
+
   **Response:**
@@
 (Rob?)”_
+
+We have now added a more detailed but simple example in the introduction.
 
@@
 _Comment 6. I was also hoping for an example that not only describes the problem that the authors solve, but that allows us to better understand the basic idea behind the authors’ solution. (Rob?)_
+
+The example in the introduction should serve this purpose.
 
@@
 
+We have now reworded this section, and added some more details about the objective function which is optimized.
+
   **Response:**

### 2023-11-10 manuscript/01-introduction.qmd
Some revisions based on comments from AE and reviewer 2

@@
 
-However, despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the hierarchical data structure has been largely ignored, and the main focus has been on producing independent (base) forecasts at a single level.
+To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
+There is a total national budget to be split between the regions in proportion to the forecast number of incidents in each region.
+The two regions have very different incident patterns, and so must be forecast using different models.
+However, the data are noisy at regional level, so the national forecasts are best obtained by summing the demand from the two regions.
+The resulting national forecasts are not equal to the sum of the regional forecasts, and so are not coherent.
+In fact, the national forecasts show a decreasing trend in demand, and so the national governing body decides to cut the budget for the next year.
+But neither of the regional forecasts shows a trend, and so the regions argue that the budget cut is unfair.
+In addition, Region A has much more variable demand than Region B, and so to cope with periods of peak demand, Region A needs to hold more resources in reserve.
+So the budget distribution needs to be made in a way that ensures the probability of each region being unable to meet demand is equal.
+Our solution to this problem is to use a hierarchical forecasting approach that ensures the forecasts are probabilistically coherent.
+Then any trends or other forecast characteristics at national level will also be reflected in the regional forecasts, and the probabilistic forecasts allow for the different levels of uncertainty in the two regions.
+Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.
+
+Despite a large number of studies dedicated to forecasting for EMS [@mingliterature2022; @gul2020exhaustive; @ibrahim2016modeling; @wargon2009systematic], the hierarchical data structure has been largely ignored, and the main focus has been on producing independent (base) forecasts at a single level.
 Generating independent forecasts can result in a lack of consistency and coordination, and therefore leads to less effective planning and decision making.

### 2023-11-10 manuscript/04-htsgts.qmd
Some revisions based on comments from AE and reviewer 2

@@
 
-Linear reconciliation methods can be written [@WicEtAl2019] as
+Linear forecast reconciliation methods can be written [@WicEtAl2019] as
 $$
@@
 where $\bm{W}$ is an $n \times n$ positive definite matrix, and $\hat{\bm{y}}_h$ contains the $h$-step forecasts of $\bm{y}_{T+h}$ given data to time $T$.
-Different choices for $\bm{W}$ lead to different solutions such as Ordinary Least Squares (OLS), Weighted Least Squares (WLS) and Minimum Trace (MinT).
+When $\bm{W}_h$ is the covariance matrix of $\hat{\bm{y}}_h$, the resulting forecasts are optimal in the sense that the sum of the variances of the reconciled forecasts is minimized, provided the base forecasts $\hat{\bm{y}}_h$ are unbiased.
+However, $\bm{W}_h$ is difficult to estimate, and so there have been various suggested approximations to $\bm{W}_h$, leading to different types of reconciliation such as Ordinary Least Squares (OLS), Weighted Least Squares (WLS) and Minimum Trace (MinT).
 
-\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
-In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors have equal variance.
-Hence, $\bm{W}$ is simply defined as the identity matrix multiplied by a constant factor.
-The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
+\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method.
+In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors are uncorrelated and have equal variance, and that multi-step forecast variances are proportion to one-step forecast variances.
+Then, $\bm{W}$ is simply the identity matrix multiplied by a constant factor.
 The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
@@
 \textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
-But it assumes that the errors of each series are uncorrelated with each other.
+But it assumes that the errors of each series are uncorrelated with each other, and that multi-step forecast variances are proportion to one-step forecast variances.
 In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal.
@@
 
-\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the base forecast errors.
-So it takes account of both the scale of each series, and the relationships between the series.} @WicEtAl2019 \textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
-The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the one-step base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.} But it still assumes
+that multi-step forecast variances are proportion to one-step forecast variances.
+The main weakness of this approach is that it is difficult to estimate the full covariance matrix, even of the one-step errors.
 In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}

### 2023-11-10 manuscript/main.qmd
Some revisions based on comments from AE and reviewer 2

@@
 
-R code to produce all results in this paper is available at https://github.com/bahmanrostamitabar/forecasting-emergency-medcine
+R code to produce all results in this paper is available at [github.com/bahmanrostamitabar/forecasting-emergency-medicine](https://github.com/bahmanrostamitabar/forecasting-emergency-medicine).
+The repository contains the raw data, all R scripts used in experiments as well as the results used in this paper, along with the raw files for producing this paper.
+Full instructions are provided in the repository.
 

### 2023-11-10 manuscript/02-literature.qmd
Fixed some bib refs and tables

@@
 Current studies ignore the inherent hierarchical and/or grouped structure of the time series data, and the relationship between series at different levels of hierarchy.
-While the hierarchical forecasting methodology has been developed and applied in various domains over the past 10 years [@panagiotelis2022probabilistic], it has never been explored in this area.
+While the hierarchical forecasting methodology has been developed and applied in various domains over the past 10 years [@panagiotelis2023probabilistic], it has never been explored in this area.
 
@@
 library(kableExtra)
-li <- readxl::read_xlsx("../literature_table/table.xlsx")
-knitr::kable(li, booktabs = T, linesep = "") |>  kableExtra::kable_styling(latex_options = c("scale_down","hold_position"), font_size =11) |> column_spec(3,"6em") |>
-  column_spec(4,"4em") |> column_spec(5,"10em")  |> column_spec(6,"8em") |> landscape()
+readxl::read_xlsx("../literature_table/table.xlsx") |>
+  knitr::kable(booktabs = T, linesep = "") |>
+  kable_styling(latex_options = "scale_down", font_size =11) |>
+  column_spec(3,"6em") |>
+  column_spec(4,"4em") |>
+  column_spec(5,"10em")  |>
+  column_spec(6,"8em") |>
+  landscape()
 ```

### 2023-11-10 manuscript/05-results.qmd
Fixed some bib refs and tables

@@
 @tbl-result presents both point and probabilistic forecast accuracy at total, control area, health board and bottom-level series.
-Point forecast performance is reported using MASE and MSSE in @tbl-result-1  and [-@tbl-result-2], respectively.
-Probabilistic forecast accuracy is reported using CRPS in @tbl-result-3.
+Point forecast performance is reported using MASE and MSSE, while probabilistic forecast accuracy is reported using CRPS.
 The bold entries in each table identify a combination of method and model that performs best for the corresponding level (i.e.
@@
 
-@tbl-result-1 and [-@tbl-result-2] show that forecast reconciliation (i.e.
-MinT) improves forecast accuracy at the higher levels of the hierarchy including total, control area and health board.
+@tbl-result shows that forecast reconciliation (i.e.
+MinT) improves point forecast accuracy at the higher levels of the hierarchy including total, control area and health board.
 However, it does not result in accuracy improvement at the bottom-level series, for which base forecasts are more accurate.
 This might be due to the noisy structure of time series at the bottom level, and possibly due to very different patterns in the aggregated series.
-It is also clear from @tbl-result-1 that the ensemble method improves forecast accuracy at total, control area and health board.
+It is also clear from @tbl-result that the ensemble method improves forecast accuracy at total, control area and health board.
 However, this does not remain valid for bottom series where different individual methods perform best, depending on the accuracy measure.
@@
 #| tbl-cap: "Average forecast performance calculated on the test sets at forecast horizons $h=43,\\dots,84$ days, with time series cross validation applied to attended incident data.
-The testset consists of 462 days.
-The best approach is highlighted in bold."
-#| tbl-subcap:
-#|   - "Point forecast accuracy using MASE"
-#|   - "Point forecast accuracy using MSSE"
-#|   - "Probabilistic forecast accuracy using CRPS"
-#| layout-nrow: 3
+The test set consists of 462 days.
+The best approach is highlighted in bold.
+Point forecast accuracy is measured using MASE and MSSE, while probabilistic forecast accuracy is measured using CRPS."
 #Set minimum in column to bold
@@
     table[[i]] <- sprintf(table[[i]], fmt="%1.3f")
-    table[[i]][best] <- cell_spec(table[[i]][best], bold=TRUE)
+    table[[i]][best] <- paste0("\\textbf{",table[[i]][best],"}")
   }
@@
   set_to_bold()
-
-kbl(mase1, booktabs = TRUE, escape = FALSE, align="llcccc") |>
-  add_header_above(c(" " = 2, "MASE" = 4)) |>
-  kable_styling(latex_options = c("hold_position"))
-
-kbl(msse1, booktabs = TRUE, escape = FALSE, align="llcccc") |>
-  add_header_above(c(" " = 2, "MSSE" = 4)) |>
-  kable_styling(latex_options = c("hold_position"))
-
-kbl(crps1, booktabs = TRUE, escape = FALSE, align="llcccc") |>
-  add_header_above(c(" " = 2, "CRPS" = 4)) |>
-  kable_styling(latex_options = c("hold_position"))
+bind_rows(mase1, msse1, crps1) |> 
+  kable(format = "latex", booktabs = TRUE, escape = FALSE, align = "llrrrr") |> 
+  pack_rows("MSSE", 1, 10) |> 
+  pack_rows("MASE", 11, 20, hline_before = TRUE) |>
+  pack_rows("CRPS", 21, 30, hline_before = TRUE) 
 ```
 
-@tbl-result-3 presents the accuracy of the forecast distributions measures by CRPS, which considers both forecasting reliability and interval sharpness.
+@tbl-result presents the accuracy of the forecast distributions measures by CRPS, which considers both forecasting reliability and interval sharpness.
 The smaller the value of CRPS, the better the comprehensive performance.
@@
 The ensemble method is also more accurate for higher levels, but ETS performs well at the bottom level.
-@tbl-result-3 indicates that reconciliation using Mint  generates accurate *distributional* forecasts.
+@tbl-result also indicates that reconciliation using Mint  generates accurate *distributional* forecasts.
 The marginal improvement in the average probabilistic forecast accuracy at the bottom level might be due to the reconciliation method giving improved forecast accuracy in the tails of the forecast distribution, which are critical for managing risks.

### 2023-11-10 manuscript/02-literature.qmd
Ran styler to obtain better code formatting

@@
   knitr::kable(booktabs = T, linesep = "") |>
-  kable_styling(latex_options = "scale_down", font_size =11) |>
-  column_spec(3,"6em") |>
-  column_spec(4,"4em") |>
-  column_spec(5,"10em")  |>
-  column_spec(6,"8em") |>
+  kable_styling(latex_options = "scale_down", font_size = 11) |>
+  column_spec(3, "6em") |>
+  column_spec(4, "4em") |>
+  column_spec(5, "10em") |>
+  column_spec(6, "8em") |>
   landscape()

### 2023-11-10 manuscript/03-experiment.qmd
Ran styler to obtain better code formatting

@@
   "Priority", 3,
-  "Priority * Control", 3*3,
-  "Priority * Health board", 3*7,
+  "Priority * Control", 3 * 3,
+  "Priority * Health board", 3 * 7,
   "Nature of incident", 35,
-  "Nature of incident * Control", 3*35,
-  "Nature of incident * Health board", 7*35,
-  "Priority * Nature of incident", 3*35-1,
+  "Nature of incident * Control", 3 * 35,
+  "Nature of incident * Health board", 7 * 35,
+  "Priority * Nature of incident", 3 * 35 - 1,
   "Control * Priority * Nature of incident", 306,
@@
 knitr::kable(agg_level, booktabs = TRUE, position = "left") %>%
-  kable_classic(full_width = FALSE) |> footnote(general = "Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset.
-As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.",threeparttable = TRUE)
+  kable_classic(full_width = FALSE) |>
+  footnote(general = "Due to certain combinations of the nature of incident with other variables, there is a lack of representation in the dataset.
+As a result, for example, instead of the calculation 3 * 35 = 105, it would be modified to 3 * 35-1 = 104.", threeparttable = TRUE)
 ```
@@
   geom_line() +
-  labs(x = "Date", y = "Incidents", color = "Nature of incident")+
+  labs(x = "Date", y = "Incidents", color = "Nature of incident") +
   ggthemes::scale_color_colorblind() +

### 2023-11-10 manuscript/05-results.qmd
Ran styler to obtain better code formatting

@@
 Point forecast accuracy is measured using MASE and MSSE, while probabilistic forecast accuracy is measured using CRPS."
-#Set minimum in column to bold
+# Set minimum in column to bold
 set_to_bold <- function(table) {
-  for(i in 3:6) {
+  for (i in 3:6) {
     best <- which.min(table[[i]])
-    table[[i]] <- sprintf(table[[i]], fmt="%1.3f")
-    table[[i]][best] <- paste0("\\textbf{",table[[i]][best],"}")
+    table[[i]] <- sprintf(table[[i]], fmt = "%1.3f")
+    table[[i]][best] <- paste0("\\textbf{", table[[i]][best], "}")
   }
@@
   ) |>
-  mutate(model = factor(model,
-    levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
-    labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
+  mutate(
+    model = factor(model,
+      levels = c("naiveecdf", "ets", "iglm", "tscount", "ensemble", "ensemble2"),
+      labels = c("Stationary", "ETS", "GLM", "TSGLM", "Ensemble", "Ensemble2")
+    ),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT")),
@@
     model = factor(model,
-      levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
-      labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
+      levels = c("naiveecdf", "ets", "iglm", "tscount", "ensemble", "ensemble2"),
+      labels = c("Stationary", "ETS", "GLM", "TSGLM", "Ensemble", "Ensemble2")
+    ),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
@@
 crps1 <- readr::read_rds(here::here("results/crps.rds")) |>
-  filter(method %in% c("mint", "base"),
+  filter(
+    method %in% c("mint", "base"),
     model != "qcomb",
@@
   ) |>
-  mutate(model = factor(model,
-    levels = c("naiveecdf","ets","iglm","tscount","ensemble","ensemble2"),
-    labels = c("Stationary", "ETS", "GLM", "TSGLM","Ensemble","Ensemble2")),
+  mutate(
+    model = factor(model,
+      levels = c("naiveecdf", "ets", "iglm", "tscount", "ensemble", "ensemble2"),
+      labels = c("Stationary", "ETS", "GLM", "TSGLM", "Ensemble", "Ensemble2")
+    ),
     method = factor(method, levels = c("base", "mint"), labels = c("Base", "MinT"))
@@
   mutate(
-    measure = factor(measure, levels = c("mase","msse","crps"), labels = c("MASE","MSSE","CRPS")),
+    measure = factor(measure, levels = c("mase", "msse", "crps"), labels = c("MASE", "MSSE", "CRPS")),
     series = factor(series, levels = c("Total", "Control areas", "Health boards")),
@@
 
-acc_summary |>   filter(
+acc_summary |>
+  filter(
     model != "Ensemble2"
@@
 
-incident_cv <- incident_gthf |> filter(lhb == "PO",is_aggregated(category),is_aggregated(nature))
+incident_cv <- incident_gthf |> filter(lhb == "PO", is_aggregated(category), is_aggregated(nature))
 fit <- incident_cv %>%
@@
 set.seed(1)
-fcst <- fit %>% generate(h=1, times = 5000, bootstrap = TRUE)
-p1 <- ggplot(data=fcst,aes(x=.sim, y=..density..)) +
-  geom_histogram(binwidth = 1, fill="#d3d3d3", color="black")+
-  geom_vline(data=fcst,aes(xintercept = (mean(.sim))), color="black", linetype="dashed", linewidth=1)+
-  scale_x_continuous(breaks = seq(25,70,5))+
-  scale_y_continuous(labels=percent)+
-  ggthemes::theme_few()+
-    labs(y="Density", x = "Demand", title = "Probabilistic forecast of ambulance demand for 1 day ahead")
+fcst <- fit %>% generate(h = 1, times = 5000, bootstrap = TRUE)
+p1 <- ggplot(data = fcst, aes(x = .sim, y = ..density..)) +
+  geom_histogram(binwidth = 1, fill = "#d3d3d3", color = "black") +
+  geom_vline(data = fcst, aes(xintercept = (mean(.sim))), color = "black", linetype = "dashed", linewidth = 1) +
+  scale_x_continuous(breaks = seq(25, 70, 5)) +
+  scale_y_continuous(labels = percent) +
+  ggthemes::theme_few() +
+  labs(y = "Density", x = "Demand", title = "Probabilistic forecast of ambulance demand for 1 day ahead")
 p1

### 2023-11-10 manuscript/main.qmd
Ran styler to obtain better code formatting

@@
 options(
-  ggplot2.discrete.colour = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
-  ggplot2.discrete.fill = c("#D55E00", "#0072B2","#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
+  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
+  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
 )
@@
 ```{r intro, child="01-introduction.qmd"}
-
 ```
@@
 ```{r literature, child="02-literature.qmd"}
-
 ```
@@
 ```{r experiment, child="03-experiment.qmd"}
-
 ```
@@
 ```{r htc, child="04-htsgts.qmd"}
-
 ```
@@
 ```{r results, child="05-results.qmd"}
-
 ```
@@
 ```{r conclusion, child="06-conclusion.qmd"}
-
 ```

### 2023-11-10 manuscript/main.tex
Ran styler to obtain better code formatting

@@
 \makeatletter
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
 \@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
-\makeatother
-\makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
-\makeatother
-\makeatletter
-\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
-\makeatother
-\makeatletter
 \makeatother
@@
 
+\setlength{\parindent}{6pt}
 \begin{document}
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, sharp corners, breakable, enhanced]}{\end{tcolorbox}}\fi
-
-\hypertarget{sec-intro}{%
-\section{Introduction}\label{sec-intro}}
+    \section{Introduction}\label{sec-intro}
 
@@
 
-However, despite a large number of studies dedicated to forecasting for
-EMS
+To illustrate the problem, let's consider a very simple example where we
+have an EMS provider with a national level of governance, and two
+regions (A and B), each with a health board and a station. There is a
+total national budget to be split between the regions in proportion to
+the forecast number of incidents in each region. The two regions have
+very different incident patterns, and so must be forecast using
+different models. However, the data are noisy at regional level, so the
+national forecasts are best obtained by summing the demand from the two
+regions. The resulting national forecasts are not equal to the sum of
+the regional forecasts, and so are not coherent. In fact, the national
+forecasts show a decreasing trend in demand, and so the national
+governing body decides to cut the budget for the next year. But neither
+of the regional forecasts shows a trend, and so the regions argue that
+the budget cut is unfair. In addition, Region A has much more variable
+demand than Region B, and so to cope with periods of peak demand, Region
+A needs to hold more resources in reserve. So the budget distribution
+needs to be made in a way that ensures the probability of each region
+being unable to meet demand is equal. Our solution to this problem is to
+use a hierarchical forecasting approach that ensures the forecasts are
+probabilistically coherent. Then any trends or other forecast
+characteristics at national level will also be reflected in the regional
+forecasts, and the probabilistic forecasts allow for the different
+levels of uncertainty in the two regions. Budget can be allocated by
+controlling the probability of demand exceeding available resources,
+rather than being simply in proportion to the expected demand.
+
+Despite a large number of studies dedicated to forecasting for EMS
 \citep{mingliterature2022, gul2020exhaustive, ibrahim2016modeling, wargon2009systematic},
@@
 
-\hypertarget{sec-lit}{%
-\section{Research background}\label{sec-lit}}
+\section{Research background}\label{sec-lit}
 
@@
   methodology has been developed and applied in various domains over the
-  past 10 years \citep{panagiotelis2022probabilistic}, it has never been
+  past 10 years \citep{panagiotelis2023probabilistic}, it has never been
   explored in this area.
@@
 
-\hypertarget{tbl-literature}{}
-\begin{landscape}\begin{table}[!h]
-\caption{\label{tbl-literature}Summary of some studies on forecasting in Emergency Medical Services }\tabularnewline
+\begin{table}
+
+\caption{\label{tbl-literature}Summary of some studies on forecasting in
+Emergency Medical Services}
+
+\centering{
 
@@
 \endgroup{}
-\end{table}
-\end{landscape}
-
-\hypertarget{sec-experiment}{%
-\section{Experiment setup}\label{sec-experiment}}
+
+}
+
+\end{table}%
+
+\section{Experiment setup}\label{sec-experiment}
 
@@
 
-\hypertarget{sec-data}{%
-\subsection{Data}\label{sec-data}}
+\subsection{Data}\label{sec-data}
 
@@
 
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{main_files/figure-pdf/fig-hierarchy-1.pdf}
-
-}
+\centering{
+
+\includegraphics[width=0.6\textwidth,height=\textheight]{main_files/figure-pdf/fig-hierarchy-1.pdf}
 
@@
 and then into 7 different healthboards.}
+
 \end{minipage}%
@@
 \begin{minipage}[t]{0.02\linewidth}
-
-{\centering
-
-~
-
-}
-
-\end{minipage}%
+~\end{minipage}%
 %
@@
 
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{../img/group.png}
-
-}
+\centering{
+
+\includegraphics[width=0.6\textwidth,height=\textheight]{../img/group.png}
 
@@
 and grouped levels.}
+
 \end{minipage}%
@@
 
-\end{figure}
+\end{figure}%
 
@@
 
-\hypertarget{tbl-hierarchy}{}
 \begin{table}
-\caption{\label{tbl-hierarchy}Number of time series in each level for the hierarchical \& grouped
-structure of attended incidents }\tabularnewline
+
+\caption{\label{tbl-hierarchy}Number of time series in each level for
+the hierarchical \& grouped structure of attended incidents}
+
+\centering{
 
@@
 \end{threeparttable}
-\end{table}
+
+}
+
+\end{table}%
 
@@
 
-{\centering \includegraphics[width=0.7\textwidth,height=\textheight]{main_files/figure-pdf/fig-feature-1.pdf}
+\centering{
+
+\includegraphics[width=0.7\textwidth,height=\textheight]{main_files/figure-pdf/fig-feature-1.pdf}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-dataviz2-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-dataviz2-1.pdf}
 
@@
 
-\hypertarget{forecasting-methods}{%
-\subsection{Forecasting methods}\label{forecasting-methods}}
+\subsection{Forecasting methods}\label{forecasting-methods}
 
@@
 
-\hypertarget{performance-evaluation}{%
-\subsection{Performance evaluation}\label{performance-evaluation}}
+\subsection{Performance evaluation}\label{performance-evaluation}
 
@@
 errors rather than absolute errors:
-\begin{equation}\protect\hypertarget{eq-RMSSE}{}{
+\begin{equation}\phantomsection\label{eq-RMSSE}{
   \text{MSSE} = \text{mean}(q_{j}^2),
-}\label{eq-RMSSE}\end{equation} where, \[
+}\end{equation} where, \[
   q^2_{j} = \frac{ e^2_{j}}
@@
 performance of the forecast distribution.
-\begin{equation}\protect\hypertarget{eq-CRPS}{}{
+\begin{equation}\phantomsection\label{eq-CRPS}{
   \text{CRPS} = \text{mean}(p_j),
-}\label{eq-CRPS}\end{equation} where \[
+}\end{equation} where \[
   p_j = \int_{-\infty}^{\infty} \left(G_j(x) - F_j(x)\right)^2dx,
@@
 
-\hypertarget{sec-htc}{%
 \section{Hierarchical and grouped time series forecasting
-techniques}\label{sec-htc}}
+techniques}\label{sec-htc}
 
@@
 
-\hypertarget{independent-base-forecast}{%
 \subsection{Independent (base
-forecast)}\label{independent-base-forecast}}
+forecast)}\label{independent-base-forecast}
 
@@
 
-\hypertarget{reconciliation-methods}{%
-\subsection{Reconciliation methods}\label{reconciliation-methods}}
+\subsection{Reconciliation methods}\label{reconciliation-methods}
 
@@
 
-\hypertarget{notations}{%
-\subsubsection{Notations}\label{notations}}
+\subsubsection{Notations}\label{notations}
 
@@
 
-\hypertarget{bottom-up-bu-and-linear-reconciliation-methods}{%
 \subsubsection{Bottom-up (BU) and linear reconciliation
-methods}\label{bottom-up-bu-and-linear-reconciliation-methods}}
+methods}\label{bottom-up-bu-and-linear-reconciliation-methods}
 
@@
 
-Linear reconciliation methods can be written \citep{WicEtAl2019} as \[
+Linear forecast reconciliation methods can be written
+\citep{WicEtAl2019} as \[
   \tilde{\bm{y}}_h = \bm{S}(\bm{S}'\bm{W}^{-1}\bm{S})^{-1}\bm{W}^{-1}\hat{\bm{y}}_h,
@@
 \(\hat{\bm{y}}_h\) contains the \(h\)-step forecasts of \(\bm{y}_{T+h}\)
-given data to time \(T\). Different choices for \(\bm{W}\) lead to
-different solutions such as Ordinary Least Squares (OLS), Weighted Least
+given data to time \(T\). When \(\bm{W}_h\) is the covariance matrix of
+\(\hat{\bm{y}}_h\), the resulting forecasts are optimal in the sense
+that the sum of the variances of the reconciled forecasts is minimized,
+provided the base forecasts \(\hat{\bm{y}}_h\) are unbiased. However,
+\(\bm{W}_h\) is difficult to estimate, and so there have been various
+suggested approximations to \(\bm{W}_h\), leading to different types of
+reconciliation such as Ordinary Least Squares (OLS), Weighted Least
 Squares (WLS) and Minimum Trace (MinT).
 
-\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
-In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors have equal variance.
-Hence, $\bm{W}$ is simply defined as the identity matrix multiplied by a constant factor.
-The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
+\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method.
+In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors are uncorrelated and have equal variance, and that multi-step forecast variances are proportion to one-step forecast variances.
+Then, $\bm{W}$ is simply the identity matrix multiplied by a constant factor.
 The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
@@
 
-\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the base forecast errors.
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the one-step base forecast errors.
 So it takes account of both the scale of each series, and the relationships between the series.}
-\citet{WicEtAl2019}
-\textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
-The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
-In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}
+But it still assumes that multi-step forecast variances are proportion
+to one-step forecast variances. The main weakness of this approach is
+that it is difficult to estimate the full covariance matrix, even of the
+one-step errors. In practice, we usually need to use a shrinkage
+estimate where the off-diagonal elements are shrunk towards zero.\}
 
@@
 
-\hypertarget{sec-results}{%
-\section{Results and discussion}\label{sec-results}}
+\section{Results and discussion}\label{sec-results}
 
@@
 board and bottom-level series. Point forecast performance is reported
-using MASE and MSSE in Table~\ref{tbl-result-1} and \ref{tbl-result-2},
-respectively. Probabilistic forecast accuracy is reported using CRPS in
-Table~\ref{tbl-result-3}. The bold entries in each table identify a
-combination of method and model that performs best for the corresponding
-level (i.e.~each column), based on the smallest values of accuracy
-measures.
-
-Table~\ref{tbl-result-1} and \ref{tbl-result-2} show that forecast
-reconciliation (i.e.~MinT) improves forecast accuracy at the higher
-levels of the hierarchy including total, control area and health board.
-However, it does not result in accuracy improvement at the bottom-level
-series, for which base forecasts are more accurate. This might be due to
-the noisy structure of time series at the bottom level, and possibly due
-to very different patterns in the aggregated series. It is also clear
-from Table~\ref{tbl-result-1} that the ensemble method improves forecast
+using MASE and MSSE, while probabilistic forecast accuracy is reported
+using CRPS. The bold entries in each table identify a combination of
+method and model that performs best for the corresponding level
+(i.e.~each column), based on the smallest values of accuracy measures.
+
+Table~\ref{tbl-result} shows that forecast reconciliation (i.e.~MinT)
+improves point forecast accuracy at the higher levels of the hierarchy
+including total, control area and health board. However, it does not
+result in accuracy improvement at the bottom-level series, for which
+base forecasts are more accurate. This might be due to the noisy
+structure of time series at the bottom level, and possibly due to very
+different patterns in the aggregated series. It is also clear from
+Table~\ref{tbl-result} that the ensemble method improves forecast
 accuracy at total, control area and health board. However, this does not
@@
 the test sets at forecast horizons \(h=43,\dots,84\) days, with time
-series cross validation applied to attended incident data. The testset
-consists of 462 days. The best approach is highlighted in
-bold.}\begin{minipage}[t]{\linewidth}
-\subcaption{\label{tbl-result-1}Point forecast accuracy using MASE }
-
-{\centering
-
-\tabularnewline
-
-\centering
-\begin{tabular}[t]{llcccc}
+series cross validation applied to attended incident data. The test set
+consists of 462 days. The best approach is highlighted in bold. Point
+forecast accuracy is measured using MASE and MSSE, while probabilistic
+forecast accuracy is measured using CRPS.}
+
+\centering{
+
+\begin{tabular}{llrrrr}
 \toprule
-\multicolumn{2}{c}{ } & \multicolumn{4}{c}{MASE} \\
-\cmidrule(l{3pt}r{3pt}){3-6}
 Method & Model & Total & Control areas & Health boards & Bottom\\
 \midrule
-Base & Stationary & 1.139 & 1.059 & 1.047 & 1.019\\
-Base & ETS & 0.963 & 0.930 & 0.899 & 1.038\\
-Base & GLM & 0.910 & 0.940 & 0.923 & \textbf{1.002}\\
-Base & TSGLM & 0.911 & 0.939 & 0.924 & 1.005\\
-Base & Ensemble & 0.782 & 0.856 & 0.876 & 1.008\\
-\addlinespace
-MinT & Stationary & 1.138 & 1.059 & 1.047 & 2.651\\
-MinT & ETS & 0.877 & 0.916 & 0.915 & 1.289\\
-MinT & GLM & 0.848 & 0.901 & 0.902 & 2.493\\
-MinT & TSGLM & 0.852 & 0.903 & 0.903 & 2.513\\
-MinT & Ensemble & \textbf{0.753} & \textbf{0.844} & \textbf{0.872} & 2.260\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{MSSE}}\\
+\hspace{1em}Base & Stationary & 1.139 & 1.059 & 1.047 & 1.019\\
+\hspace{1em}Base & ETS & 0.963 & 0.930 & 0.899 & 1.038\\
+\hspace{1em}Base & GLM & 0.910 & 0.940 & 0.923 & \textbf{1.002}\\
+\hspace{1em}Base & TSGLM & 0.911 & 0.939 & 0.924 & 1.005\\
+\hspace{1em}Base & Ensemble & 0.782 & 0.856 & 0.876 & 1.008\\
+\hspace{1em}MinT & Stationary & 1.138 & 1.059 & 1.047 & 2.651\\
+\hspace{1em}MinT & ETS & 0.877 & 0.916 & 0.915 & 1.289\\
+\hspace{1em}MinT & GLM & 0.848 & 0.901 & 0.902 & 2.493\\
+\hspace{1em}MinT & TSGLM & 0.852 & 0.903 & 0.903 & 2.513\\
+\hspace{1em}MinT & Ensemble & \textbf{0.753} & \textbf{0.844} & \textbf{0.872} & 2.260\\
+\addlinespace[0.3em]
+\hline
+\multicolumn{6}{l}{\textbf{MASE}}\\
+\hspace{1em}Base & Stationary & 1.169 & 1.056 & 1.062 & 1.031\\
+\hspace{1em}Base & ETS & 0.979 & 0.875 & 0.816 & \textbf{0.975}\\
+\hspace{1em}Base & GLM & 0.813 & 0.897 & 0.875 & 1.009\\
+\hspace{1em}Base & TSGLM & 0.822 & 0.901 & 0.875 & 1.050\\
+\hspace{1em}Base & Ensemble & 0.599 & 0.729 & 0.774 & 0.993\\
+\hspace{1em}MinT & Stationary & 1.168 & 1.057 & 1.062 & 2.095\\
+\hspace{1em}MinT & ETS & 0.785 & 0.852 & 0.845 & 0.994\\
+\hspace{1em}MinT & GLM & 0.720 & 0.827 & 0.837 & 1.803\\
+\hspace{1em}MinT & TSGLM & 0.722 & 0.833 & 0.839 & 1.851\\
+\hspace{1em}MinT & Ensemble & \textbf{0.560} & \textbf{0.706} & \textbf{0.765} & 1.557\\
+\addlinespace[0.3em]
+\hline
+\multicolumn{6}{l}{\textbf{CRPS}}\\
+\hspace{1em}Base & Stationary & 30.387 & 10.882 & 5.500 & 0.302\\
+\hspace{1em}Base & ETS & 14.309 & 6.074 & 3.476 & 0.244\\
+\hspace{1em}Base & GLM & 15.396 & 6.253 & 3.576 & 0.244\\
+\hspace{1em}Base & TSGLM & 15.316 & 6.227 & 3.575 & 0.245\\
+\hspace{1em}Base & Ensemble & 12.978 & \textbf{5.727} & 3.430 & 0.243\\
+\hspace{1em}MinT & Stationary & 30.368 & 10.902 & 5.498 & 0.313\\
+\hspace{1em}MinT & ETS & 13.515 & 5.967 & 3.547 & \textbf{0.243}\\
+\hspace{1em}MinT & GLM & 13.839 & 5.917 & 3.453 & 0.246\\
+\hspace{1em}MinT & TSGLM & 14.000 & 5.947 & 3.455 & 0.248\\
+\hspace{1em}MinT & Ensemble & \textbf{12.585} & 5.728 & \textbf{3.426} & 0.247\\
 \bottomrule
@@
 
-\end{minipage}%
-\newline
-\begin{minipage}[t]{\linewidth}
-\subcaption{\label{tbl-result-2}Point forecast accuracy using MSSE }
-
-{\centering
-
-\tabularnewline
-
-\centering
-\begin{tabular}[t]{llcccc}
-\toprule
-\multicolumn{2}{c}{ } & \multicolumn{4}{c}{MSSE} \\
-\cmidrule(l{3pt}r{3pt}){3-6}
-Method & Model & Total & Control areas & Health boards & Bottom\\
-\midrule
-Base & Stationary & 1.169 & 1.056 & 1.062 & 1.031\\
-Base & ETS & 0.979 & 0.875 & 0.816 & \textbf{0.975}\\
-Base & GLM & 0.813 & 0.897 & 0.875 & 1.009\\
-Base & TSGLM & 0.822 & 0.901 & 0.875 & 1.050\\
-Base & Ensemble & 0.599 & 0.729 & 0.774 & 0.993\\
-\addlinespace
-MinT & Stationary & 1.168 & 1.057 & 1.062 & 2.095\\
-MinT & ETS & 0.785 & 0.852 & 0.845 & 0.994\\
-MinT & GLM & 0.720 & 0.827 & 0.837 & 1.803\\
-MinT & TSGLM & 0.722 & 0.833 & 0.839 & 1.851\\
-MinT & Ensemble & \textbf{0.560} & \textbf{0.706} & \textbf{0.765} & 1.557\\
-\bottomrule
-\end{tabular}
-
-}
-
-\end{minipage}%
-\newline
-\begin{minipage}[t]{\linewidth}
-\subcaption{\label{tbl-result-3}Probabilistic forecast accuracy using CRPS }
-
-{\centering
-
-\tabularnewline
-
-\centering
-\begin{tabular}[t]{llcccc}
-\toprule
-\multicolumn{2}{c}{ } & \multicolumn{4}{c}{CRPS} \\
-\cmidrule(l{3pt}r{3pt}){3-6}
-Method & Model & Total & Control areas & Health boards & Bottom\\
-\midrule
-Base & Stationary & 30.387 & 10.882 & 5.500 & 0.302\\
-Base & ETS & 14.309 & 6.074 & 3.476 & 0.244\\
-Base & GLM & 15.396 & 6.253 & 3.576 & 0.244\\
-Base & TSGLM & 15.316 & 6.227 & 3.575 & 0.245\\
-Base & Ensemble & 12.978 & \textbf{5.727} & 3.430 & 0.243\\
-\addlinespace
-MinT & Stationary & 30.368 & 10.902 & 5.498 & 0.313\\
-MinT & ETS & 13.515 & 5.967 & 3.547 & \textbf{0.243}\\
-MinT & GLM & 13.839 & 5.917 & 3.453 & 0.246\\
-MinT & TSGLM & 14.000 & 5.947 & 3.455 & 0.248\\
-MinT & Ensemble & \textbf{12.585} & 5.728 & \textbf{3.426} & 0.247\\
-\bottomrule
-\end{tabular}
-
-}
-
-\end{minipage}%
-
-\end{table}
-
-Table~\ref{tbl-result-3} presents the accuracy of the forecast
+\end{table}%
+
+Table~\ref{tbl-result} presents the accuracy of the forecast
 distributions measures by CRPS, which considers both forecasting
@@
 level. The ensemble method is also more accurate for higher levels, but
-ETS performs well at the bottom level. Table~\ref{tbl-result-3}
+ETS performs well at the bottom level. Table~\ref{tbl-result} also
 indicates that reconciliation using Mint generates accurate
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-accuracy-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-accuracy-1.pdf}
 
@@
 
-\hypertarget{an-illustration-of-probabilistic-forecast-for-ems-demand}{%
 \subsection{An illustration of probabilistic forecast for EMS
-demand}\label{an-illustration-of-probabilistic-forecast-for-ems-demand}}
+demand}\label{an-illustration-of-probabilistic-forecast-for-ems-demand}
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}
+\centering{
+
+\includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}
 
@@
 
-\end{figure}
-
-\hypertarget{sec-conclusion}{%
-\section{Conclusion}\label{sec-conclusion}}
+\end{figure}%
+
+\section{Conclusion}\label{sec-conclusion}
 
@@
 
-\hypertarget{reproducibility}{%
-\section*{Reproducibility}\label{reproducibility}}
+\section*{Reproducibility}\label{reproducibility}
 \addcontentsline{toc}{section}{Reproducibility}
@@
 R code to produce all results in this paper is available at
-https://github.com/bahmanrostamitabar/forecasting-emergency-medcine
+\href{https://github.com/bahmanrostamitabar/forecasting-emergency-medicine}{github.com/bahmanrostamitabar/forecasting-emergency-medicine}.
+The repository contains the raw data, all R scripts used in experiments
+as well as the results used in this paper, along with the raw files for
+producing this paper. Full instructions are provided in the repository.
 

### 2023-11-10 rscript/result/analysing_results.Rmd
Ran styler to obtain better code formatting

@@
 ```{r read-result, options}
-
 hts_result_ets <- read_rds("result/ets_hts_rmsse_long_h.rds")
@@
 ```{r label, options}
+rmsse_approach_level_id <- rmse_for_analysis %>%
+  group_by(level, approach, .id) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-rmsse_approach_level_id <-  rmse_for_analysis %>% group_by(level,approach,.id) %>%
-  summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
-
-rmsse_approach_id <-  rmsse_approach_level_id %>% group_by(approach,level) %>%
-  summarise(rmsse=mean(rmsse), .groups = "drop")
+rmsse_approach_id <- rmsse_approach_level_id %>%
+  group_by(approach, level) %>%
+  summarise(rmsse = mean(rmsse), .groups = "drop")
 ```
@@
 ```{r label, options}
-r <- rmsse_approach_level_id %>% group_by(approach) %>%
-  summarise(mean_rmsse=mean(rmsse), median_rmsse=median(rmsse), .groups = "drop")
+r <- rmsse_approach_level_id %>%
+  group_by(approach) %>%
+  summarise(mean_rmsse = mean(rmsse), median_rmsse = median(rmsse), .groups = "drop")
 r
-
 ```
@@
 ```{r label, options}
-rmsse_total <- rmse_for_analysis %>% filter(level=="Total") %>% dplyr::select(-level) %>%
-  group_by(approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_total <- rmse_for_analysis %>%
+  filter(level == "Total") %>%
+  dplyr::select(-level) %>%
+  group_by(approach, .id) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_total, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
-  geom_boxplot()+
+ggplot(rmsse_total, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
+  geom_boxplot() +
   ggthemes::theme_few()
@@
 ```{r label, options}
-rmsse_control <- rmse_for_analysis %>% filter(level %in% c("G1/S","G1/N","G1/C")) %>%
-  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_control <- rmse_for_analysis %>%
+  filter(level %in% c("G1/S", "G1/N", "G1/C")) %>%
+  group_by(level, approach, .id) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_control, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
-  geom_boxplot()+
-  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
+ggplot(rmsse_control, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
+  geom_boxplot() +
+  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
   ggthemes::theme_few()
@@
 ```{r label, options}
-rmsse_control_h <- rmse_for_analysis %>% filter(level %in% c("G1/S","G1/N","G1/C")) %>%
-  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_control_h <- rmse_for_analysis %>%
+  filter(level %in% c("G1/S", "G1/N", "G1/C")) %>%
+  group_by(level, approach, h) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_control_h, aes(x=h,y=rmsse, color=factor(approach)))+
-  geom_point()+
-  geom_line()+
-  scale_x_continuous(breaks = seq(1,35,1))+
-  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
-  ggthemes::theme_few()+
+ggplot(rmsse_control_h, aes(x = h, y = rmsse, color = factor(approach))) +
+  geom_point() +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, 35, 1)) +
+  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
+  ggthemes::theme_few() +
   ggthemes::scale_color_colorblind()
@@
 ```{r label, options}
-rmsse_hb <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G2/")) %>%
-  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_hb <- rmse_for_analysis %>%
+  filter(stringr::str_detect(level, "G2/")) %>%
+  group_by(level, approach, .id) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_hb, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
-  geom_boxplot()+
-  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
+ggplot(rmsse_hb, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
+  geom_boxplot() +
+  facet_wrap(vars(level), ncol = 2, scales = "free_y") +
   ggthemes::theme_few()
@@
 ```{r label, options}
-rmsse_hb_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G2/")) %>%
-  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_hb_h <- rmse_for_analysis %>%
+  filter(stringr::str_detect(level, "G2/")) %>%
+  group_by(level, approach, h) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_hb_h, aes(x=h,y=rmsse, color=factor(approach)))+
-  geom_point()+
-  geom_line()+
-  scale_x_continuous(breaks = seq(1,35,1))+
-  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
-  ggthemes::theme_few()+
+ggplot(rmsse_hb_h, aes(x = h, y = rmsse, color = factor(approach))) +
+  geom_point() +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, 35, 1)) +
+  facet_wrap(vars(level), ncol = 2, scales = "free_y") +
+  ggthemes::theme_few() +
   ggthemes::scale_color_colorblind()
@@
 ```{r label, options}
-rmsse_priority <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G3/")) %>%
-  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_priority <- rmse_for_analysis %>%
+  filter(stringr::str_detect(level, "G3/")) %>%
+  group_by(level, approach, .id) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_priority, aes(x=fct_reorder(approach,rmsse),y=rmsse))+
-  geom_boxplot()+
-  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
+ggplot(rmsse_priority, aes(x = fct_reorder(approach, rmsse), y = rmsse)) +
+  geom_boxplot() +
+  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
   ggthemes::theme_few()
@@
 ```{r label, options}
-rmsse_priority_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G3/")) %>%
-  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_priority_h <- rmse_for_analysis %>%
+  filter(stringr::str_detect(level, "G3/")) %>%
+  group_by(level, approach, h) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_priority_h, aes(x=h,y=rmsse, color=factor(approach)))+
-  geom_point()+
-  geom_line()+
-  scale_x_continuous(breaks = seq(1,35,1))+
-  facet_wrap(vars(level),ncol = 1, scales = "free_y")+
-  ggthemes::theme_few()+
+ggplot(rmsse_priority_h, aes(x = h, y = rmsse, color = factor(approach))) +
+  geom_point() +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, 35, 1)) +
+  facet_wrap(vars(level), ncol = 1, scales = "free_y") +
+  ggthemes::theme_few() +
   ggthemes::scale_color_colorblind()
@@
 ```{r label, options}
-rmsse_nature <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G4/")) %>%
-  group_by(level,approach,.id) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_nature <- rmse_for_analysis %>%
+  filter(stringr::str_detect(level, "G4/")) %>%
+  group_by(level, approach, .id) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-rmsse_nature_all <- rmsse_nature %>% group_by(approach,level) %>% summarise(rmsse=mean(rmsse), .groups = "drop")
-rmsse_nature_all %>% pivot_wider(names_from = approach, values_from = rmsse)%>% arrange(arima_bu) %>% View()
+rmsse_nature_all <- rmsse_nature %>%
+  group_by(approach, level) %>%
+  summarise(rmsse = mean(rmsse), .groups = "drop")
+rmsse_nature_all %>%
+  pivot_wider(names_from = approach, values_from = rmsse) %>%
+  arrange(arima_bu) %>%
+  View()
 ```
@@
 ```{r label, options}
-rmsse_nature_h <- rmse_for_analysis %>% filter(stringr::str_detect(level, "G4/")) %>%
-  group_by(level,approach,h) %>% summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
+rmsse_nature_h <- rmse_for_analysis %>%
+  filter(stringr::str_detect(level, "G4/")) %>%
+  group_by(level, approach, h) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
 
-ggplot(rmsse_hb_h, aes(x=h,y=rmsse, color=factor(approach)))+
-  geom_point()+
-  geom_line()+
-  scale_x_continuous(breaks = seq(1,35,1))+
-  facet_wrap(vars(level),ncol = 2, scales = "free_y")+
-  ggthemes::theme_few()+
+ggplot(rmsse_hb_h, aes(x = h, y = rmsse, color = factor(approach))) +
+  geom_point() +
+  geom_line() +
+  scale_x_continuous(breaks = seq(1, 35, 1)) +
+  facet_wrap(vars(level), ncol = 2, scales = "free_y") +
+  ggthemes::theme_few() +
   ggthemes::scale_color_colorblind()
@@
 ```{r label, options}
-rr <- rmsse_approach_level_id %>% group_by(approach,level) %>%
-  summarise(mean_rmsse=mean(rmsse),.groups = "drop")
-ggplot(rmsse_approach_id)+
-  geom_boxplot(aes(y=rmsse, x=approach))+
-  coord_cartesian(ylim = c(0,2))
+rr <- rmsse_approach_level_id %>%
+  group_by(approach, level) %>%
+  summarise(mean_rmsse = mean(rmsse), .groups = "drop")
+ggplot(rmsse_approach_id) +
+  geom_boxplot(aes(y = rmsse, x = approach)) +
+  coord_cartesian(ylim = c(0, 2))
 ```
@@
 ```{r label, options}
-rmsse_hd <-  rmse_for_analysis %>% group_by(h,.id,approach) %>%
-  summarise(rmsse=sqrt(mean(error2_scale)), .groups = "drop")
-ggplot(rmsse_hd)+
-  geom_boxplot(aes(y=rmsse, x=factor(h)))+
+rmsse_hd <- rmse_for_analysis %>%
+  group_by(h, .id, approach) %>%
+  summarise(rmsse = sqrt(mean(error2_scale)), .groups = "drop")
+ggplot(rmsse_hd) +
+  geom_boxplot(aes(y = rmsse, x = factor(h))) +
   facet_wrap(vars(approach))

### 2023-11-10 rscript/data/holidays.R

@@ -32,7 +32,7 @@ holiday_dummy <- holiday |>
   as_tsibble(index = date)
 
 # Save as RDS for reading in by other scripts
-write_rds(holiday_dummy, paste0(storage_folder, "holiday_dummy.rds"), compress="bz2")
+write_rds(holiday_dummy, paste0(storage_folder, "holiday_dummy.rds"), compress = "bz2")
 
 
 # Other holiday tibbles created below, no longer used (?)

### 2023-11-10 rscript/data/incidents.R

@@ -43,7 +43,7 @@ incidents_tsbl <- incidents |>
   as_tsibble(index = date, key = c(lhb_code, category, nature_of_incident)) |>
   fill_gaps(incidents = 0, .full = TRUE)
 # Save as rds
-write_rds(incidents_tsbl, paste0(storage_folder, "incidents_tsbl.rds"), compress="bz2")
+write_rds(incidents_tsbl, paste0(storage_folder, "incidents_tsbl.rds"), compress = "bz2")
 
 # Small data set in gts format
 incidents_tsbl |>
@@ -52,12 +52,12 @@ incidents_tsbl |>
   select(-nature_of_incident, -lhb_code) |>
   mutate(category = recode(category, "GREEN" = "GRE", "AMBER" = "AMB")) |>
   pivot_wider(names_from = category, values_from = incidents) |>
-  #filter(date <= "2015-12-31") |>
+  # filter(date <= "2015-12-31") |>
   select(-date) |>
   mutate(BLU = 0) |>
-  ts(frequency=7) |>
+  ts(frequency = 7) |>
   hts() |>
-  write_rds(paste0(storage_folder, "incidents_test_gts.rds"), compress="bz2")
+  write_rds(paste0(storage_folder, "incidents_test_gts.rds"), compress = "bz2")
 
 # Prepare full data in gts format
 incident_modify <- incidents_tsbl |>
@@ -132,7 +132,7 @@ incident_ready <- incident_modify |>
 incident_gt <- incident_ready |>
   as_tsibble(index = date, key = c(region, lhb, category, nature)) |>
   aggregate_key((region / lhb) * category * nature, incident = sum(incident))
-write_rds(incident_gt, paste0(storage_folder, "incidents_gt.rds"), compress="bz2")
+write_rds(incident_gt, paste0(storage_folder, "incidents_gt.rds"), compress = "bz2")
 
 
 a1 <- incident_ready |>
@@ -145,14 +145,14 @@ nat <- unique(incident_ready$nature)
 region <- sort(unique(incident_ready$region))
 for (i in seq_along(hb)) {
   region <- case_when(
-      hb[i] == "CT" ~ "S",
-      hb[i] == "PO" ~ "C",
-      hb[i] == "AB" ~ "S",
-      hb[i] == "BC" ~ "N",
-      hb[i] == "CV" ~ "S",
[…]

### 2023-11-10 rscript/empirical_distribution.R

@@ -1,23 +1,22 @@
-
 # empirical distribution
 
 library(fpp3)
 library(distributional)
 set.seed(123)
 
-train <- tsibble(t = 1:100, y=rnorm(100), index=t)
-test <- tsibble(t = 101:120, y=rnorm(20), index=t)
+train <- tsibble(t = 1:100, y = rnorm(100), index = t)
+test <- tsibble(t = 101:120, y = rnorm(20), index = t)
 fc_ar1 <- train %>%
   model(
-    ar1 = ARIMA(y ~ pdq(1,0,0))
+    ar1 = ARIMA(y ~ pdq(1, 0, 0))
   ) %>%
-  fabletools::forecast(h=20)
-fc_benchmark <- fc_ar1 %>% as_tibble() %>% 
+  fabletools::forecast(h = 20)
+fc_benchmark <- fc_ar1 %>%
+  as_tibble() %>%
   mutate(
     y = distributional::dist_sample(list(train$y)),
-    .mean1=mean(train$y),
-    .model = 'benchmark'
+    .mean1 = mean(train$y),
+    .model = "benchmark"
   )
 bind_rows(fc_ar1, fc_benchmark) %>%
   accuracy(test, measures = distribution_accuracy_measures)
-

### 2023-11-10 rscript/fable/fit_model_fable.R

@@ -1,5 +1,5 @@
 # This script needs the "tolerant-inverse" branch of fabletools
-#remotes::install_github("tidyverts/fabletools", "tolerant-inverse")
+# remotes::install_github("tidyverts/fabletools", "tolerant-inverse")
 
 library(tidyverse)
 library(tsibble)
@@ -11,13 +11,13 @@ library(fable.tscount)
 # Read data previously processed in data folder and add aggregation levels
 incidents <- readRDS(here::here("data/incidents_tsbl.rds")) |>
   # Temporarily only consider a small part of data until everything works
-  #filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |> 
-  #select(-nature_of_incident, -lhb_code) |> 
-  #aggregate_key(category, incidents = sum(incidents))
+  # filter(nature_of_incident == "BREATHING PROBLEMS", lhb_code == "BC") |>
+  # select(-nature_of_incident, -lhb_code) |>
+  # aggregate_key(category, incidents = sum(incidents))
   # End of temporary section of code
   # Uncomment next 3 lines when ready to scale up
   aggregate_key(
-    nature_of_incident * category * lhb_code, 
+    nature_of_incident * category * lhb_code,
     incidents = sum(incidents)
   )
 
@@ -29,9 +29,9 @@ incidents <- incidents |>
   )
 
 # First pass using a simple training/test set keeping last 6 weeks for testing
-train <- incidents |> 
-  filter(date <= max(date) - 42) 
-test <- incidents |> 
+train <- incidents |>
+  filter(date <= max(date) - 42)
+test <- incidents |>
   filter(date > max(date) - 42)
 
 # Fit some models
@@ -39,52 +39,54 @@ test <- incidents |>
 # Issue raised at https://github.com/mitchelloharawild/fable.tscount/issues/2
 # library(future)
 # plan(multisession)
-if(fs::file_exists(here::here("rscript/fable/fit_incidents.rds"))) {
-    fit_incidents <- read_rds(here::here("rscript/fable/fit_incidents.rds"))
-  } else {
-    fit_incidents <- train |> 
-      model(
-        NAIVE = NAIVE(sqrt(incidents)),
-        ETS = ETS(sqrt(incidents)),
[…]

### 2023-11-10 rscript/figure/hierarchical_graph.R

@@ -1,22 +1,26 @@
-library(collapsibleTree) 
+library(collapsibleTree)
 
 # input data must be a nested data frame:
 head(warpbreaks)
 
 # Represent this tree:
-p <- collapsibleTree( warpbreaks, 
-                      c("wool", "tension", "breaks"))
+p <- collapsibleTree(
+  warpbreaks,
+  c("wool", "tension", "breaks")
+)
 
-Total <- incident_ready %>% 
-  select(region, lhb,category,nature)
+Total <- incident_ready %>%
+  select(region, lhb, category, nature)
 
-p <- collapsibleTree( Total, 
-                      c("region","lhb","category","nature"))
+p <- collapsibleTree(
+  Total,
+  c("region", "lhb", "category", "nature")
+)
 p
 
 
 
-widgetToPng <- function(widget, file = "widget.pdf",  ...) {
+widgetToPng <- function(widget, file = "widget.pdf", ...) {
   temp <- tempfile(fileext = ".html")
   file <- R.utils::getAbsolutePath(file)
   htmlwidgets::saveWidget(widget, temp)
@@ -30,7 +34,8 @@ widgetToPng <- function(widget, file = "widget.pdf",  ...) {
 }
 
 widgetToPng(
-  collapsibleTreeSummary(warpbreaks, c("wool", "tension", "breaks"), 
-                         maxPercent = 50, collapsed = FALSE),
+  collapsibleTreeSummary(warpbreaks, c("wool", "tension", "breaks"),
+    maxPercent = 50, collapsed = FALSE
+  ),
   "collapsibleTree.png"
 )

### 2023-11-10 rscript/hts/ensemble.R

@@ -35,7 +35,7 @@ create_ensembles <- function(models_to_use = "all") {
 
 create_specific_ensemble <- function(models, method, origin) {
   # Form file name for saving results
-  if(all(models == "all")) {
+  if (all(models == "all")) {
     filename <- paste0(storage_folder, "ensemble_", origin, "_sim_", method, ".rds")
   } else {
     filename <- paste0(storage_folder, "ensemble2_", origin, "_sim_", method, ".rds")
@@ -50,7 +50,7 @@ create_specific_ensemble <- function(models, method, origin) {
   files <- fs::dir_ls(storage_folder, glob = paste0("*_", origin, "_sim_", method, ".rds"))
   files <- files[!grepl("ensemble", files)]
   files <- files[!grepl("qcomb", files)]
-  if(all(models != "all")) {
+  if (all(models != "all")) {
     # Remove naive
     files <- files[!grepl("naive*", files)]
   }

### 2023-11-10 rscript/hts/forecast_hts.R

@@ -33,9 +33,9 @@ for (i in seq(origins)) {
 }
 
 create_ensembles() # Include naive
-create_ensembles(models_to_use = c("ets","iglm","tscount")) #Don't include naive
+create_ensembles(models_to_use = c("ets", "iglm", "tscount")) # Don't include naive
 create_qcomb() # Include naive
-create_qcomb(models_to_use = c("ets","iglm","tscount")) #Don't include naive
+create_qcomb(models_to_use = c("ets", "iglm", "tscount")) # Don't include naive
 
 # Accuracy
 

### 2023-11-10 rscript/hts/qcombinations.R

@@ -35,7 +35,7 @@ create_qcomb <- function(models_to_use = "all") {
 
 create_specific_qcomb <- function(models, method, origin) {
   # Form file name for saving results
-  if(all(models == "all")) {
+  if (all(models == "all")) {
     filename <- paste0(storage_folder, "qcomb_", origin, "_sim_", method, ".rds")
   } else {
     filename <- paste0(storage_folder, "qcomb2_", origin, "_sim_", method, ".rds")
@@ -50,29 +50,29 @@ create_specific_qcomb <- function(models, method, origin) {
   files <- fs::dir_ls(storage_folder, glob = paste0("*_", origin, "_sim_", method, ".rds"))
   files <- files[!grepl("ensemble", files)]
   files <- files[!grepl("qcomb", files)]
-  if(all(models != "all")) {
+  if (all(models != "all")) {
     # Remove naive
     files <- files[!grepl("naive*", files)]
   }
 
   # For each training set, load available simulation files and compute quantiles
   # with same average quantiles
-  prob <- seq(0.001, 0.999, by=0.001)
+  prob <- seq(0.001, 0.999, by = 0.001)
   q <- array(0, c(1224, 84, length(files), length(prob)))
-  for(k in seq_along(files)) {
+  for (k in seq_along(files)) {
     tmp <- readr::read_rds(files[k])
-    for(i in seq(1224)) {
-      for(j in seq(84)) {
-        q[i, j, k, ] <- quantile(tmp[i,j,], prob=prob)
+    for (i in seq(1224)) {
+      for (j in seq(84)) {
+        q[i, j, k, ] <- quantile(tmp[i, j, ], prob = prob)
       }
     }
   }
-  aveq <- apply(q, c(1,2,4), mean) 
+  aveq <- apply(q, c(1, 2, 4), mean)
   # Generate simulated data with same distribution
   sim <- array(0, c(1224, 84, 1000))
-  for(j in seq(1224)) {
-    for(k in seq(84)) {
-      sim[j,k,] <- sample(aveq[j,k,], size=1000, replace=TRUE)
+  for (j in seq(1224)) {
+    for (k in seq(84)) {
+      sim[j, k, ] <- sample(aveq[j, k, ], size = 1000, replace = TRUE)
     }
   }
   write_rds(sim, filename, compress = "bz2")

### 2023-11-10 rscript/minimal_examples/minimal_example.R

@@ -50,7 +50,7 @@ fit_incident <- train %>%
   model(
     NAIVE = NAIVE(sqrt(incidents)),
     ETS = ETS(sqrt(incidents)),
-    TSCOUNT = fable.tscount::TSCOUNT(incidents ~ trend() + season("week") + fourier("year", 10), link = "log", model=list(past_obs=1:4))
+    TSCOUNT = fable.tscount::TSCOUNT(incidents ~ trend() + season("week") + fourier("year", 10), link = "log", model = list(past_obs = 1:4))
   ) %>%
   reconcile(
     bu_ETS = bottom_up(ETS),
@@ -60,4 +60,3 @@ fit_incident <- train %>%
   )
 
 fcst_incident <- fit_incident %>% forecast(h = f_horizon)
-

### 2023-11-10 rscript/minimal_examples/negative_forecast.R

@@ -2,64 +2,77 @@ library(hts)
 
 
 # functions
-tscount_f <- function(x,h) {
-  xFourier <- forecast::fourier(msts(x,
-                                     seasonal.periods=c(7,365.25)),
-                                K=c(3,10))
-  xreg_holiday_fourier <- cbind(xFourier,
-                                holiday_dummy_matrix_train)
+tscount_f <- function(x, h) {
+  xFourier <- forecast::fourier(
+    msts(x,
+      seasonal.periods = c(7, 365.25)
+    ),
+    K = c(3, 10)
+  )
+  xreg_holiday_fourier <- cbind(
+    xFourier,
+    holiday_dummy_matrix_train
+  )
   tsglm_fit <- tscount::tsglm(as.ts(x),
-                              model = list(past_obs = 1), 
-                              link = "log", 
-                              distr = "poisson",
-                              xreg= xreg_holiday_fourier
+    model = list(past_obs = 1),
+    link = "log",
+    distr = "poisson",
+    xreg = xreg_holiday_fourier
   )
   fFourier <- forecast::fourier(msts(x,
-                                     seasonal.periods=c(7,365.25)),K=c(3,10),h=fh)
-  newxreg_holiday_fourier <- cbind(fFourier,holiday_dummy_matrix_test)
-  pf <- predict(tsglm_fit, 
-                n.ahead = fh, 
-                newxreg = newxreg_holiday_fourier,
-                level=0
+    seasonal.periods = c(7, 365.25)
+  ), K = c(3, 10), h = fh)
+  newxreg_holiday_fourier <- cbind(fFourier, holiday_dummy_matrix_test)
+  pf <- predict(tsglm_fit,
+    n.ahead = fh,
+    newxreg = newxreg_holiday_fourier,
+    level = 0
   )$p
   pf
 }
 set.seed(2021)
 fh <- 35
-holiday_dummy_matrix <- holiday_dummy %>% as_tibble() %>% select(-date) %>% as.matrix()
-holiday_dummy_matrix_train <- holiday_dummy_matrix[1:(1400-fh),]
-holiday_dummy_matrix_test <- holiday_dummy_matrix[((1400-fh+1):1400),]
+holiday_dummy_matrix <- holiday_dummy %>%
+  as_tibble() %>%
[…]

### 2023-11-10 rscript/minimal_examples/reconcile_negative.R

@@ -1,8 +1,8 @@
 library(hts)
 set.seed(1)
 # Bottom level series
-bts <- msts(matrix(rnorm(1500), ncol=2), seasonal.periods=c(7,365.25))
-bts[bts<0] <- 0
+bts <- msts(matrix(rnorm(1500), ncol = 2), seasonal.periods = c(7, 365.25))
+bts[bts < 0] <- 0
 bts <- round(bts)
 
 # Create hts object
@@ -13,9 +13,10 @@ ally <- aggts(y)
 # Produce forecasts of all series
 h <- 28
 allf <- matrix(NA, nrow = h, ncol = NCOL(ally))
-for(i in seq(NCOL(ally)))
-  allf[,i] <- stlf(ally[,i], h = h)$mean
-allf[allf<0] <- 0
+for (i in seq(NCOL(ally))) {
+  allf[, i] <- stlf(ally[, i], h = h)$mean
+}
+allf[allf < 0] <- 0
 # Reconcile results
 y.f <- combinef(allf, get_nodes(y))
 y.f_all <- aggts(y.f)
@@ -25,8 +26,8 @@ y.f_all <- aggts(y.f)
 library(hts)
 set.seed(1)
 # Bottom level series
-bts <- msts(matrix(rnorm(1500), ncol=2), seasonal.periods=c(7,365.25))
-bts[bts<0] <- 0
+bts <- msts(matrix(rnorm(1500), ncol = 2), seasonal.periods = c(7, 365.25))
+bts[bts < 0] <- 0
 bts <- round(bts)
 
 # Create hts object
@@ -34,18 +35,19 @@ y <- hts(bts)
 # Produce forecasts
 h <- 28
 comb_ets <- forecast(y,
-                     h=h,
-                     method = "comb",
-                     fmethod="ets",
-                     keep_interval = TRUE,
-                     lambda = .5,
-                     parallel = TRUE)
+  h = h,
+  method = "comb",
+  fmethod = "ets",
+  keep_interval = TRUE,
+  lambda = .5,
+  parallel = TRUE
+)
 
 bu_ets <- forecast(y,
-                   h=h,
-                   method="bu",
-                   fmethod="ets",
-                   keep_interval = TRUE,
-                   lambda = .5,
-                   parallel = TRUE
+  h = h,
+  method = "bu",
+  fmethod = "ets",
+  keep_interval = TRUE,
+  lambda = .5,
+  parallel = TRUE
 )

### 2023-11-10 rscript/minimal_examples/test_msts_hts.R

@@ -1,29 +1,33 @@
 library(hts)
 set.seed(2021)
 # Bottom level series
-bts <- msts(matrix(rnorm(1500), ncol=2), seasonal.periods=c(7,365.25))
-bts[bts<0] <- 0
+bts <- msts(matrix(rnorm(1500), ncol = 2), seasonal.periods = c(7, 365.25))
+bts[bts < 0] <- 0
 bts <- round(bts)
 
 # Create hts object
 y <- hts(bts)
-#plot(y)
+# plot(y)
 # Create aggregated series
 ally <- aggts(y)
 # Produce forecasts of all series
 h <- 28
 allf <- matrix(NA, nrow = h, ncol = NCOL(ally))
-for(i in seq(NCOL(ally)))
-  allf[,i] <- stlf(ally[,i], h = h)$mean
-allf[allf<0] <- 0
+for (i in seq(NCOL(ally))) {
+  allf[, i] <- stlf(ally[, i], h = h)$mean
+}
+allf[allf < 0] <- 0
 # Reconcile results
-y.f <- combinef(allf, 
-                nodes = get_nodes(y), 
-                nonnegative = TRUE)
+y.f <- combinef(allf,
+  nodes = get_nodes(y),
+  nonnegative = TRUE
+)
 y.f_all <- aggts(y.f)
-y.f_all[y.f_all<0]
+y.f_all[y.f_all < 0]
 
 # Add back multiple seasonality attributes
-y.f$bts <- msts(y.f$bts, start=tsp(y$bts)[2]+1/tsp(y$bts)[3],
-                seasonal.periods = attributes(y$bts)$msts)
+y.f$bts <- msts(y.f$bts,
+  start = tsp(y$bts)[2] + 1 / tsp(y$bts)[3],
+  seasonal.periods = attributes(y$bts)$msts
+)
 plot(y.f)

### 2023-11-10 rscript/oldhts/forecast_hts.R

@@ -3,58 +3,67 @@ library(hts)
 library(forecast)
 library(tscount)
 
-fh <- 5*7
+fh <- 5 * 7
 incident_all <- read_rds("data/incident_all.rds")
 holidays_p <- read_rds("data/holiday_prophet.rds")
 holiday_dummy <- read_rds("data/holiday_dummy.rds")
-holiday_dummy <- holiday_dummy %>% 
-  mutate_if(is.factor,as.character)%>% 
+holiday_dummy <- holiday_dummy %>%
+  mutate_if(is.factor, as.character) %>%
   mutate_if(is.character, as.integer)
 
-holiday_dummy_matrix <- holiday_dummy %>% as_tibble() %>% select(-date) %>% as.matrix()
-holiday_dummy_matrix_train <- holiday_dummy_matrix[1:(1400-fh),]
-holiday_dummy_matrix_test <- holiday_dummy_matrix[((1400-fh+1):1400),]
+holiday_dummy_matrix <- holiday_dummy %>%
+  as_tibble() %>%
+  select(-date) %>%
+  as.matrix()
+holiday_dummy_matrix_train <- holiday_dummy_matrix[1:(1400 - fh), ]
+holiday_dummy_matrix_test <- holiday_dummy_matrix[((1400 - fh + 1):1400), ]
 
-incident_all_train <- incident_all[1:(1400-fh),]
-incident_all_test <- incident_all[((1400-fh+1):1400),]
+incident_all_train <- incident_all[1:(1400 - fh), ]
+incident_all_test <- incident_all[((1400 - fh + 1):1400), ]
 #----
 
-#incident_msts_train <- msts(incident_all_train,seasonal.periods = c(7,365.25))
-#create hierarchical/grouped time series
+# incident_msts_train <- msts(incident_all_train,seasonal.periods = c(7,365.25))
+# create hierarchical/grouped time series
 incident_ts_train <- ts(incident_all_train, frequency = 7)
-gts_incident_train <- hts::gts(incident_ts_train, 
-                         characters = list(c(1,2),3,9))
-#tets set
-#incident_msts_test <- msts(incident_all_test, seasonal.periods = c(7,365.25))
+gts_incident_train <- hts::gts(incident_ts_train,
+  characters = list(c(1, 2), 3, 9)
+)
+# tets set
+# incident_msts_test <- msts(incident_all_test, seasonal.periods = c(7,365.25))
[…]

### 2023-11-10 rscript/oldhts/form JASA paper.R

@@ -1,38 +1,38 @@
-It is easy enough to do using the equations from the JASA paper. Here is an example using the infantgts data from the hts package.
+# It is easy enough to do using the equations from the JASA paper. Here is an example using the infantgts data from the hts package.
 
 library(hts)
 
 h <- 10
 ally <- aggts(infantgts)
 fmean <- fvar <- matrix(NA, nrow = h, ncol = NCOL(ally))
-for(i in seq(NCOL(ally))) {
-  fc <- forecast(auto.arima(ally[,i]), h = h, level=95)
-  fmean[,i] <- fc$mean
-  fvar[,i] <- ((fc$upper - fc$lower)/(2*qnorm(0.975)))^2
+for (i in seq(NCOL(ally))) {
+  fc <- forecast(auto.arima(ally[, i]), h = h, level = 95)
+  fmean[, i] <- fc$mean
+  fvar[, i] <- ((fc$upper - fc$lower) / (2 * qnorm(0.975)))^2
 }
 S <- smatrix(infantgts)
-Lambda <- diag(1/fvar[1,])
+Lambda <- diag(1 / fvar[1, ])
 G <- solve(t(S) %*% Lambda %*% S) %*% t(S) %*% Lambda
 SG <- S %*% G
 ytilde <- SG %*% t(fmean)
 vartilde <- ytilde * NA
-for(i in seq(h)) {
-  vartilde <- diag(SG %*% diag(fvar[i,]) %*% t(SG))
+for (i in seq(h)) {
+  vartilde <- diag(SG %*% diag(fvar[i, ]) %*% t(SG))
 }
 lower <- qnorm(0.025, mean = ytilde, sd = sqrt(vartilde))
 upper <- qnorm(0.975, mean = ytilde, sd = sqrt(vartilde))
 
-I have computed the mean directly (as ytilde). This should give the same results as 
-combinef(fmean, groups = get_groups(infantgts), weights=1/fvar[1,])
+# I have computed the mean directly (as ytilde). This should give the same results as
+# combinef(fmean, groups = get_groups(infantgts), weights=1/fvar[1,])
 
-But since we needed all the matrices for the variance calculation, I thought it would be more obvious what is happening if I computed ytilde as well.
+# But since we needed all the matrices for the variance calculation, I thought it would be more obvious what is happening if I computed ytilde as well.
 
-Because we are using a transformation of the data (sqrt) before fitting the models, you need to compute ytilde, lower and upper on the transformed data. Then square them all to give the results on the original scale.
[…]

### 2023-11-10 rscript/oldhts/generate_quantile.R

@@ -1,23 +1,24 @@
 library(forecast)
-#Normal distribution
+# Normal distribution
 fit <- auto.arima(WWWusage)
-fc <- forecast(fit, h=20, level=95)
-qf <- matrix(0, nrow=19, ncol=20)
+fc <- forecast(fit, h = 20, level = 95)
+qf <- matrix(0, nrow = 19, ncol = 20)
 m <- fc$mean
-s <- (fc$upper-fc$lower)/1.96/2
-for(h in 1:20)
-  qf[,h] <- qnorm(seq(0.05,.95,.05), m[h], s[h])
+s <- (fc$upper - fc$lower) / 1.96 / 2
+for (h in 1:20) {
+  qf[, h] <- qnorm(seq(0.05, .95, .05), m[h], s[h])
+}
 
 plot(fc)
-matlines(101:120, t(qf), col=rainbow(120), lty=1)
+matlines(101:120, t(qf), col = rainbow(120), lty = 1)
 
 # poisson
-qf <- matrix(0, nrow=19, ncol=20)
+qf <- matrix(0, nrow = 19, ncol = 20)
 m <- fc$mean # mean should be the point forecast generate from a forecasting model sy=uch as poinsson regression , tscount , etc
-for(h in 1:20)
-  qf[,h] <- qpois(seq(0.05,.95,.05), m[h])
+for (h in 1:20) {
+  qf[, h] <- qpois(seq(0.05, .95, .05), m[h])
+}
 
 # for poisson
 lower <- qpois(0.025, lambda = ytilde)
 upper <- qpois(0.975, lambda = ytilde)
-

### 2023-11-10 rscript/oldhts/interval and probability without transformation/hts_arima_interval.R

@@ -6,130 +6,143 @@ library(forecast)
 library(tscount)
 library(tsibble)
 fcst_inciddent <- function(train, test) {
-incident_all_train <- train %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-incident_all_test <- test %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-holiday_dummy_matrix_train <- train %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
-holiday_dummy_matrix_test <- test %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
+  incident_all_train <- train %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  incident_all_test <- test %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  holiday_dummy_matrix_train <- train %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
+  holiday_dummy_matrix_test <- test %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
 
-incident_ts_train <- ts((incident_all_train), frequency = 7)
-#create hierarchical/grouped time series
-gts_incident_train <- hts::gts(incident_ts_train , 
-                               characters = list(c(1,2),3,9))
-incident_ts_test <- ts(incident_all_test, frequency = 7)
-gts_incident_test <- hts::gts(incident_ts_test , 
-                              characters = list(c(1,2),3,9))
-comd_arima <- forecast(gts_incident_train, 
-                       h=fh,
-                       method = "comb",
-                       fmethod="arima", 
-                       parallel = TRUE,
-                       lambda = .5)
-bu_arima <- forecast(gts_incident_train, 
-                     h=fh,method="bu", 
-                     fmethod="arima", 
-                     parallel = TRUE,
-                     lambda = .5)
+  incident_ts_train <- ts((incident_all_train), frequency = 7)
[…]

### 2023-11-10 rscript/oldhts/interval and probability without transformation/hts_ets_interval.R

@@ -5,164 +5,176 @@ library(forecast)
 library(tscount)
 library(tsibble)
 fcst_inciddent <- function(train, test) {
-incident_all_train <- train %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-incident_all_test <- test %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-holiday_dummy_matrix_train <- train %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
-holiday_dummy_matrix_test <- test %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
+  incident_all_train <- train %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  incident_all_test <- test %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  holiday_dummy_matrix_train <- train %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
+  holiday_dummy_matrix_test <- test %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
 
-incident_ts_train <- ts((incident_all_train), frequency = 7)
-#create hierarchical/grouped time series
-gts_incident_train <- hts::gts(incident_ts_train , 
-                               characters = list(c(1,2),3,9))
-incident_ts_test <- ts(incident_all_test, frequency = 7)
-gts_incident_test <- hts::gts(incident_ts_test , 
-                              characters = list(c(1,2),3,9))
-comb_ets <- forecast(gts_incident_train, 
-                     h=fh,
-                     method = "comb",
-                     fmethod="ets", 
-                     lambda = .5,
-                     parallel = TRUE,
-                     )
+  incident_ts_train <- ts((incident_all_train), frequency = 7)
+  # create hierarchical/grouped time series
+  gts_incident_train <- hts::gts(incident_ts_train,
+    characters = list(c(1, 2), 3, 9)
+  )
[…]

### 2023-11-10 rscript/oldhts/interval_probability with transformation/htsplus_with_boxcox_transformation.R

@@ -10,78 +10,78 @@ library(hts)
 # htsplus = hts with BoxCox transformation
 htsplus <- function(object, h,
                     model_function = auto.arima,
-                    lambda=NULL, level=95,
-                    method = c("wls","bu"), nsim=2000) {
+                    lambda = NULL, level = 95,
+                    method = c("wls", "bu"), nsim = 2000) {
   method <- match.arg(method)
   # Generate all time series at all levels
   ally <- aggts(object)
   nseries <- NCOL(ally)
   # Do Box-Cox transformation if required
-  if(!is.null(lambda)) {
+  if (!is.null(lambda)) {
     ally <- BoxCox(ally, lambda)
   }
   # Set up matrices to store results
   fmean <- fvar <- flower <- fupper <- matrix(NA, nrow = h, ncol = nseries)
   models <- list()
   # Loop over all time series and compute forecast means and variances
-  for(i in seq(NCOL(ally))) {
-    models[[i]] <- model_function(ally[,i])
-    fc <- forecast(models[[i]], h=h, level=level)
-    fmean[,i] <- fc$mean
-    fvar[,i] <- ((fc$upper - fc$lower)/(2*qnorm(0.5+level/200)))^2
+  for (i in seq(NCOL(ally))) {
+    models[[i]] <- model_function(ally[, i])
+    fc <- forecast(models[[i]], h = h, level = level)
+    fmean[, i] <- fc$mean
+    fvar[, i] <- ((fc$upper - fc$lower) / (2 * qnorm(0.5 + level / 200)))^2
   }
   # Reverse Box-Cox transformation if required
   # using Taylor series approximation (delta method) to get mean and variance
   # on original scale
-  if(!is.null(lambda)) {
+  if (!is.null(lambda)) {
     fmean <- InvBoxCox(fmean, lambda)
     tmp <- lambda * fmean + 1
-    fp2 <- fmean^(2-2*lambda)
-    fmean <- fmean * (1 + 0.5*fvar*(1-lambda)/(fmean)^(2*lambda))
+    fp2 <- fmean^(2 - 2 * lambda)
+    fmean <- fmean * (1 + 0.5 * fvar * (1 - lambda) / (fmean)^(2 * lambda))
     fvar <- fvar * fp2
   }
   # S matrix
   S <- smatrix(object)
   # Find G matrix
[…]

### 2023-11-10 rscript/oldhts/point forecast/hts_arima.R

@@ -6,101 +6,113 @@ library(forecast)
 library(tscount)
 library(tsibble)
 fcst_inciddent <- function(train, test) {
-incident_all_train <- train %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-incident_all_test <- test %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-holiday_dummy_matrix_train <- train %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
-holiday_dummy_matrix_test <- test %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
+  incident_all_train <- train %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  incident_all_test <- test %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  holiday_dummy_matrix_train <- train %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
+  holiday_dummy_matrix_test <- test %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
 
-incident_ts_train <- ts((incident_all_train), frequency = 7)
-#create hierarchical/grouped time series
-gts_incident_train <- hts::gts(incident_ts_train , 
-                               characters = list(c(1,2),3,9))
-incident_ts_test <- ts(incident_all_test, frequency = 7)
-gts_incident_test <- hts::gts(incident_ts_test , 
-                              characters = list(c(1,2),3,9))
-comd_arima <- forecast(gts_incident_train, 
-                       h=fh,
-                       method = "comb",
-                       fmethod="arima", 
-                       parallel = TRUE,
-                       lambda = .5)
-bu_arima <- forecast(gts_incident_train, 
-                     h=fh,method="bu", 
-                     fmethod="arima", 
-                     parallel = TRUE,
-                     lambda = .5)
+  incident_ts_train <- ts((incident_all_train), frequency = 7)
[…]

### 2023-11-10 rscript/oldhts/point forecast/hts_ets.R

@@ -5,135 +5,146 @@ library(forecast)
 library(tscount)
 library(tsibble)
 fcst_inciddent <- function(train, test) {
-incident_all_train <- train %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-incident_all_test <- test %>% select(-public_holiday_d,-school_holiday_d,-xmas,-new_years_day)
-holiday_dummy_matrix_train <- train %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
-holiday_dummy_matrix_test <- test %>% select(public_holiday_d,school_holiday_d,xmas,new_years_day) %>% as.matrix()
+  incident_all_train <- train %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  incident_all_test <- test %>% select(-public_holiday_d, -school_holiday_d, -xmas, -new_years_day)
+  holiday_dummy_matrix_train <- train %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
+  holiday_dummy_matrix_test <- test %>%
+    select(public_holiday_d, school_holiday_d, xmas, new_years_day) %>%
+    as.matrix()
 
-incident_ts_train <- ts((incident_all_train), frequency = 7)
-#create hierarchical/grouped time series
-gts_incident_train <- hts::gts(incident_ts_train , 
-                               characters = list(c(1,2),3,9))
-incident_ts_test <- ts(incident_all_test, frequency = 7)
-gts_incident_test <- hts::gts(incident_ts_test , 
-                              characters = list(c(1,2),3,9))
-comb_ets <- forecast(gts_incident_train, 
-                     h=fh,
-                     method = "comb",
-                     fmethod="ets", 
-                     lambda = .5,
-                     parallel = TRUE,
-                     )
+  incident_ts_train <- ts((incident_all_train), frequency = 7)
+  # create hierarchical/grouped time series
+  gts_incident_train <- hts::gts(incident_ts_train,
+    characters = list(c(1, 2), 3, 9)
+  )
[…]

### 2023-11-10 rscript/oldhts/point forecast/hts_prophet.R

@@ -9,122 +9,134 @@ library(prophet)
 
 #---- prophet parameter setting -------
 prophetfit_control <- prophet(
-  yearly.seasonality = FALSE, 
-  weekly.seasonality = FALSE, 
+  yearly.seasonality = FALSE,
+  weekly.seasonality = FALSE,
   daily.seasonality = FALSE,
   holidays = holidays_p,
   interval.width = 0.95,
   mcmc.samples = 0, # invokes rstan::optimizing (fast MAP estimation)
-  )
-
-prophetfit_control <- add_seasonality(prophetfit_control,
-                                      name = "weekly",
-                                      period = 7,
-                                      fourier.order = 3
 )
 
 prophetfit_control <- add_seasonality(prophetfit_control,
-                                      name = "yearly",
-                                      period = 365.25,
-                                      fourier.order = 10
+  name = "weekly",
+  period = 7,
+  fourier.order = 3
+)
+
+prophetfit_control <- add_seasonality(prophetfit_control,
+  name = "yearly",
+  period = 365.25,
+  fourier.order = 10
 )
 
 #-------end of prophet parameter setting-------
 #--- Begin of function to generate prediction-------
-#x is the time series in the training set, h =forecast horizon, date_train=date corresponding to the training,date_test=date corresponding to the test set
-prophet_f <- function(x,h,date_train,date_test) {
-  df <-  data.frame(date_train, 
-                    y = x)
-  df$y = BoxCox(df$y, lambda=0.5)# use Box-Cox transformation to avoid negarive forecast
+# x is the time series in the training set, h =forecast horizon, date_train=date corresponding to the training,date_test=date corresponding to the test set
+prophet_f <- function(x, h, date_train, date_test) {
+  df <- data.frame(date_train,
+    y = x
+  )
+  df$y <- BoxCox(df$y, lambda = 0.5) # use Box-Cox transformation to avoid negarive forecast
   prophetfit <- fit.prophet(
     m = prophetfit_control,
     df = df
   )
-  
[…]

### 2023-11-10 rscript/oldhts/point forecast/hts_tscount.R

@@ -5,142 +5,161 @@ library(forecast)
 library(tscount)
 library(tsibble)
 #---------------- tscount_f----
-tscount_f <- function(x,h,holiday_dummy_matrix_train,holiday_dummy_matrix_test) {
-  if ((sum(x>0))>20) {
-  xFourier <- forecast::fourier(msts(x,
-                                     seasonal.periods=c(7,365.25)),
-                                K=c(3,10))
-  xreg_holiday_fourier <- cbind(xFourier,holiday_dummy_matrix_train)
-  #count non-zeros and depending on the length, fit a simple model like mean
-  tsglm_fit <- tsglm(as.ts(x),
-                     model = list(past_obs = 13), 
-                     link = "log", 
-                     distr = "poisson",
-                     xreg= xreg_holiday_fourier
-  )
-  fFourier <- forecast::fourier(msts(x,
-                                     seasonal.periods=c(7,365.25)),K=c(3,10),h=fh)
-  newxreg_holiday_fourier <- cbind(fFourier,holiday_dummy_matrix_test)
-  pf <- predict(tsglm_fit, 
-                n.ahead = fh, 
-                newxreg = newxreg_holiday_fourier,
-                level=0
-  )$p
-  } else{
-  pf <- meanf(x,fh)
+tscount_f <- function(x, h, holiday_dummy_matrix_train, holiday_dummy_matrix_test) {
+  if ((sum(x > 0)) > 20) {
+    xFourier <- forecast::fourier(
+      msts(x,
+        seasonal.periods = c(7, 365.25)
+      ),
+      K = c(3, 10)
+    )
+    xreg_holiday_fourier <- cbind(xFourier, holiday_dummy_matrix_train)
+    # count non-zeros and depending on the length, fit a simple model like mean
+    tsglm_fit <- tsglm(as.ts(x),
+      model = list(past_obs = 13),
+      link = "log",
+      distr = "poisson",
+      xreg = xreg_holiday_fourier
+    )
+    fFourier <- forecast::fourier(msts(x,
+      seasonal.periods = c(7, 365.25)
+    ), K = c(3, 10), h = fh)
+    newxreg_holiday_fourier <- cbind(fFourier, holiday_dummy_matrix_test)
+    pf <- predict(tsglm_fit,
+      n.ahead = fh,
[…]

### 2023-11-10 rscript/oldhts/point forecast/tscv_hts.R

@@ -10,162 +10,193 @@ library(forecast)
 library(tscount)
 
 incident_all <- read_rds("incident_all.rds")
-#holidays_p <- read_rds("data/holiday_prophet.rds")
+# holidays_p <- read_rds("data/holiday_prophet.rds")
 holiday_dummy <- read_rds("holiday_dummy.rds")
-holiday_dummy <- holiday_dummy %>% 
-  mutate_if(is.factor,as.character)%>% 
+holiday_dummy <- holiday_dummy %>%
+  mutate_if(is.factor, as.character) %>%
   mutate_if(is.character, as.integer)
-holiday_dummy_matrix <- holiday_dummy %>% as_tibble() %>% select(-date) %>% as.matrix()
-#Functions
-tscount_f <- function(x,h,holiday_dummy_matrix_train,holiday_dummy_matrix_test) {
-  xFourier <- forecast::fourier(msts(x,
-                                     seasonal.periods=c(7,365.25)),
-                                K=c(3,10))
-  xreg_holiday_fourier <- cbind(xFourier,holiday_dummy_matrix_train)
+holiday_dummy_matrix <- holiday_dummy %>%
+  as_tibble() %>%
+  select(-date) %>%
+  as.matrix()
+# Functions
+tscount_f <- function(x, h, holiday_dummy_matrix_train, holiday_dummy_matrix_test) {
+  xFourier <- forecast::fourier(
+    msts(x,
+      seasonal.periods = c(7, 365.25)
+    ),
+    K = c(3, 10)
+  )
+  xreg_holiday_fourier <- cbind(xFourier, holiday_dummy_matrix_train)
   tsglm_fit <- tsglm(as.ts(x),
-                     model = list(past_obs = 1), 
-                     link = "log", 
-                     distr = "poisson",
-                     xreg= xreg_holiday_fourier
+    model = list(past_obs = 1),
+    link = "log",
+    distr = "poisson",
+    xreg = xreg_holiday_fourier
   )
   fFourier <- forecast::fourier(msts(x,
-                                     seasonal.periods=c(7,365.25)),K=c(3,10),h=fh)
-  newxreg_holiday_fourier <- cbind(fFourier,holiday_dummy_matrix_test)
-  pf <- predict(tsglm_fit, 
-                n.ahead = fh, 
-                newxreg = newxreg_holiday_fourier,
[…]

### 2023-11-10 rscript/result/accuracy_hts.R

@@ -1,4 +1,4 @@
-# RMSSE <- function (.error, .train, na.rm = TRUE) 
+# RMSSE <- function (.error, .train, na.rm = TRUE)
 # {
 #   .train <- diff(.train, lag = 7)
 #   scale <- mean(.train^2, na.rm = na.rm)
@@ -7,41 +7,41 @@
 
 library(tidyverse)
 library(hts)
-#hts_result_ets <- read_rds("result/hts_result_ets.rds")
-#hts_result_ets <- read_rds("result/hts_result_arima.rds")
+# hts_result_ets <- read_rds("result/hts_result_ets.rds")
+# hts_result_ets <- read_rds("result/hts_result_arima.rds")
 hts_result_ets <- read_rds("result/hts_result_tscount.rds")
 
 
 #-----forecast------
 fct <- hts_result_ets$fct_all
-id <- tibble(.id=rep(1,70))
+id <- tibble(.id = rep(1, 70))
 a1 <- (fct[[1]]$forecast_all_level) %>% bind_cols(id)
-a1[,] <- NA
+a1[, ] <- NA
 for (i in (1:nrow(hts_result_ets))) {
   a <- (fct[[i]]$forecast_all_level)
-  id <- tibble(.id=rep(i,70))
+  id <- tibble(.id = rep(i, 70))
   a <- a %>% bind_cols(id)
   a1 <- a1 %>% bind_rows(a)
 }
 forecast_hts <- a1 %>% drop_na()
-head(forecast_hts[,c(1,1225:1226)])
-write_rds(forecast_hts,"result/forecast/forecast_tscount.rds")
+head(forecast_hts[, c(1, 1225:1226)])
+write_rds(forecast_hts, "result/forecast/forecast_tscount.rds")
 
 #----error---------
 
 fct <- hts_result_ets$fct_all
-id <- tibble(.id=rep(1,70))
+id <- tibble(.id = rep(1, 70))
 a1 <- (fct[[1]]$error_all_level) %>% bind_cols(id)
-a1[,] <- NA
+a1[, ] <- NA
 for (i in (1:nrow(hts_result_ets))) {
   a <- (fct[[i]]$error_all_level)
-  id <- tibble(.id=rep(i,70))
+  id <- tibble(.id = rep(i, 70))
   a <- a %>% bind_cols(id)
   a1 <- a1 %>% bind_rows(a)
 }
 error <- a1 %>% drop_na()
 
-write_rds(error,"result/error/error_tscount.rds")
[…]

### 2023-11-10 rscript/result/error_metric.R

@@ -1,12 +1,8 @@
-
-
 d <- .train %>% features(incident, unitroot_ndiffs)
 D <- .train %>% features(incident, unitroot_nsdiffs)
 
-RMSSE <- function (.error, .train, na.rm = TRUE) 
-{
-    .train <- diff(.train, lag = .period, differences = D)
+RMSSE <- function(.error, .train, na.rm = TRUE) {
+  .train <- diff(.train, lag = .period, differences = D)
   scale <- mean(.train^2, na.rm = na.rm)
-  sqrt(mean(.error^2/scale, na.rm = na.rm))
+  sqrt(mean(.error^2 / scale, na.rm = na.rm))
 }
-

### 2023-11-10 rscript/result/visualise_results.R

@@ -8,43 +8,66 @@ f_ets <- read_rds("result/forecast/forecast_ets.rds")
 f_arima <- read_rds("result/forecast/forecast_arima.rds")
 f_tscount <- read_rds("result/forecast/forecast_tscount.rds")
 
-f1 <- f_ets %>% filter(.id %in% c(1,120,332)) %>% select(Total, .id,approach)
-f2 <- f_arima %>% filter(.id %in% c(1,120,332)) %>% select(Total,.id,approach)
-f3 <- f_tscount %>% select(Total,.id,approach) %>% filter(Total > 4000)
+f1 <- f_ets %>%
+  filter(.id %in% c(1, 120, 332)) %>%
+  select(Total, .id, approach)
+f2 <- f_arima %>%
+  filter(.id %in% c(1, 120, 332)) %>%
+  select(Total, .id, approach)
+f3 <- f_tscount %>%
+  select(Total, .id, approach) %>%
+  filter(Total > 4000)
 View(f1)
 View(f2)
 View(f3)
 
 test <- read_rds("result/test/test.rds")
-t1 <- test %>% filter(.id %in% c(1,2,3)) %>% select(Total, .id, approach)
+t1 <- test %>%
+  filter(.id %in% c(1, 2, 3)) %>%
+  select(Total, .id, approach)
 View(t1)
-all_hts_result <- hts_result_ets %>% bind_rows(hts_result_arima) %>% bind_rows(hts_result_tscount)
-ets_hts_rmsse_long_h %>% filter(approach == "ets_comb", .id==1|.id==2,level=="Total") %>% View()
+all_hts_result <- hts_result_ets %>%
+  bind_rows(hts_result_arima) %>%
+  bind_rows(hts_result_tscount)
+ets_hts_rmsse_long_h %>%
+  filter(approach == "ets_comb", .id == 1 | .id == 2, level == "Total") %>%
+  View()
 
 
-write_rds(all_hts_result,"result/all_hts_result.rds")
-all_hts_result %>% filter(is.na(error2_scale)) %>% View()
+write_rds(all_hts_result, "result/all_hts_result.rds")
+all_hts_result %>%
+  filter(is.na(error2_scale)) %>%
+  View()
 
 
 
 rm <- all_hts_result %>% filter(!is.nan(error2_scale))
 rm_noninf <- rm %>% filter(!is.infinite(error2_scale))
 
-rmss_hts <- rm_noninf %>% 
[…]

### 2023-12-14 journal_correspondence/response_letter/R2_response_letter_JSR.qmd
finalised comments of AE

--- 2023-11-06 journal_correspondence/response_letter/R2_response_letter_JSR.qmd
@@
 
-_Comment 1. The authors need to score higher on the economic insights. Their example did not impress me, and I was hoping to receive more intuition as to why their approach works better. (Rob?)_
+_Comment 1. The authors need to score higher on the economic insights. Their example did not impress me, and I was hoping to receive more intuition as to why their approach works better._
 
-  **Response:**
+  **Response:** Thank you for your feedback.
+In this paper, we introduce a comprehensive framework for probabilistic forecast reconciliation in Emergency Medical Services.
+Our work not only outlines the rationale behind the proposed approach's suitability for this specific context but also rigorously evaluates its effectiveness through statistical forecast accuracy metrics, as presented in Table 3 and Figure 4.
+In response to the comment on insights into the economic implications of our approach, we have added the following paragraphs to the paper to further discuss this point:
+
+\textcolor{teal}{Our research establishes a strong basis for future investigations and practical implementation in EMS.
+Leveraging the hierarchical and grouped structure of demand time series, EMS can use this advanced forecasting framework to generate coherent point and probabilistic forecasts, making the most of all available data at every level of the hierarchy.
+We acknowledge that a forecast serves a greater purpose beyond its mere existence, ideally enabling the best utility in terms of efficient allocation of medical services, response time, and cost, all informed by the forecast.
+While we fully appreciate the importance of evaluating forecast quality based on its impact on decision-making processes, it is essential to address the data requirements and methodology involved in measuring this impact.
+For a comprehensive assessment of the forecasts' implications, access to additional data beyond ambulance demand, covering various decision types, capacity information, constraints in the decision system, and more, becomes necessary.
+This additional data would offer valuable insights into the specific decisions relying on the forecasts, resulting in a more accurate evaluation of their impact on medical services.
+Furthermore, measuring the actual impact of forecasts would necessitate an approach that goes beyond forecasting itself.
+This would involve developing and implementing simulation models capable of replicating decision-making processes based on the forecast inputs.
+These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.}
 
@@
 I appreciate that the authors summarize previous research in Table 1.
-Yet, there is nothing in Table 1 that helps the reader understand why the current research has the limitations that the authors state on page 8 (Bahman)._
+Yet, there is nothing in Table 1 that helps the reader understand why the current research has the limitations that the authors state on page 8._
 
-  **Response:**
+  **Response:** Thank you for your comment.
+In Table 1, we present three dimensions that highlights the limitations identified in prior research.
+These dimensions include reconciliation, probabilistic forecasting, and reproducibility.
+It is noteworthy that there is a notable gap in the existing literature regarding the consideration of probabilistic forecast reconciliation while adhering to the principles of reproducibility.
+The first column in the table relates to our research, indicating 'YES' in all these dimensions, whereas previous studies have been characterized by a 'NO' in this regard.
+Table 1 in the previous version had only Reconciliation, and Probabilistic.
+We now added the reproducibility column to the table, aligning with limitations discussed in the literature section.
 
@@
 No doubt, the resulting transparency is a major strength of this manuscript.
-Yet, the authors could still do a much better job in outlining to the readers that the authors have done a good job.
-(Rob?)_
+Yet, the authors could still do a much better job in outlining to the readers that the authors have done a good job._
 
-  **Response:**
+  **Response:** Thank you.
+We would like to emphasize that our commitment to transparency and reproducibility goes beyond merely sharing data and code.
+Our entire paper is authored in R using [Quarto](https://quarto.org/), setting it apart as a uniquely transparent and reproducible workflow and this could be used as an outstanding example of a transparent and reproducible research in academic communities.
+In this way, we provide not only the code and data but also the entire narrative of the research process.
+To further highlight that, we have expanded the the reproducibility section in the revised manuscript:
+
+\textcolor{teal}{To enhance transparency and reproducibility, we not only provide data and the code, but also the entire paper that is written in R using [Quarto](https://quarto.org/).
+All materials to reproduce this paper is available at [github.com/bahmanrostamitabar/forecasting-emergency-medicine](https://github.com/bahmanrostamitabar/forecasting-emergency-medicine).
+The repository contains the raw data, all R scripts used in experiments, the results used in the paper, as well as the quarto files for producing this paper.
+Full instructions are provided in the repository.}
 
 _Comment 4.I felt that the authors did not address the following concern that I outlined in the last round of review: “It would also help to have a rather detailed example, maybe even a numerical example, to better describe why separate forecasts at different levels can cause problems.
-Right now, it sounds plausible that such problems occur, but you are still rather vague concerning the details of these problems.
-(Rob?)”_
+Right now, it sounds plausible that such problems occur, but you are still rather vague concerning the details of these problems.”_
 
-  **Response:**
+  **Response:** Thank you. We have now added a more detailed but simple example in the introduction based on the ambulance data and added the following text to the revised version of the paper:
+
+\textcolor{teal}{To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
+There is a total national budget to be split between the regions in proportion to the forecast number of incidents in each region.
+The two regions have very different incident patterns, and so must be forecast using different models.
+However, the data are noisy at regional level, so the national forecasts are best obtained by summing the demand from the two regions.
+The resulting national forecasts are not equal to the sum of the regional forecasts, and so are not coherent.
+In fact, the national forecasts show a decreasing trend in demand, and so the national governing body decides to cut the budget for the next year.
+But neither of the regional forecasts shows a trend, and so the regions argue that the budget cut is unfair.
+In addition, Region A has much more variable demand than Region B, and so to cope with periods of peak demand, Region A needs to hold more resources in reserve.
+So the budget distribution needs to be made in a way that ensures the probability of each region being unable to meet demand is equal.
+Our solution to this problem is to use a hierarchical forecasting approach that ensures the forecasts are probabilistically coherent.
+Then any trends or other forecast characteristics at national level will also be reflected in the regional forecasts, and the probabilistic forecasts allow for the different levels of uncertainty in the two regions.
+Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.}
 
@@
 Unfortunately, Figure 1 is not self-contained (e.g., abbreviations are not explained).
-Figure 2 could also contain the correlation and the number of observations.
-(Bahman)_
+Figure 2 could also contain the correlation and the number of observations._
 
-  **Response:**
+  **Response:** Thank you for the comment.
+We have anonymized the name of the health board using two letters (e.g.) that hold no meaningful association, so these are not abbreviations.
+We have clarified this in the caption of Figure 1.
 
-_Comment 6. I was also hoping for an example that not only describes the problem that the authors solve, but that allows us to better understand the basic idea behind the authors’ solution. (Rob?)_
+  With regard to Figure 2, we appreciate the interest in examining correlations.
+However, we would like to clarify that the purpose of the scatter plot in Figure 2 is not to show the correlation.
+Each point on the plot corresponds to an individual time series, and the primary objective of this plot is to facilitate a clear understanding of time series features in our dataset such as the strength of trend and seasonality.
+Introducing correlation values in this specific plot may not add meaningful insights due to the nature of the data representation.
 
-  **Response:**
+_Comment 6. I was also hoping for an example that not only describes the problem that the authors solve, but that allows us to better understand the basic idea behind the authors’ solution._
+
+  **Response:** Thank you.
+We've addressed this comment by including a detailed example in the introduction.
+This addition is designed to clarify the problem and elaborate on the idea behind the forecast reconciliation.
+We added the following paragraph to the revised version of the paper:
+
+\textcolor{teal}{To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
+There is a total national budget to be split between the regions in proportion to the forecast number of incidents in each region.
+The two regions have very different incident patterns, and so must be forecast using different models.
+However, the data are noisy at regional level, so the national forecasts are best obtained by summing the demand from the two regions.
+The resulting national forecasts are not equal to the sum of the regional forecasts, and so are not coherent.
+In fact, the national forecasts show a decreasing trend in demand, and so the national governing body decides to cut the budget for the next year.
+But neither of the regional forecasts shows a trend, and so the regions argue that the budget cut is unfair.
+In addition, Region A has much more variable demand than Region B, and so to cope with periods of peak demand, Region A needs to hold more resources in reserve.
+So the budget distribution needs to be made in a way that ensures the probability of each region being unable to meet demand is equal.
+Our solution to this problem is to use a hierarchical forecasting approach that ensures the forecasts are probabilistically coherent.
+Then any trends or other forecast characteristics at national level will also be reflected in the regional forecasts, and the probabilistic forecasts allow for the different levels of uncertainty in the two regions.
+Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.}
 

### 2023-12-14 manuscript/01-introduction.qmd
finalised comments of AE

@@
 Demand for emergency medical services at the national level can be disaggregated in a geographical hierarchy into sub-national, regions, health boards, and stations/hospitals, or divided into groups such as the nature of incidents or demand priority.
-\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchies are necessary for effective decision-making in EMS.
+Forecasts produced at both higher and lower levels of hierarchies are necessary for effective decision-making in EMS.
 For example, control area EMS forecasts can inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals.
@@
 This could help to identify patterns in demand that may not be apparent at the higher level.
-Therefore, employing forecasting techniques that consider the hierarchical and/or grouped patterns of time series in EMS aligns naturally, offering the possibility to enhance forecast accuracy and facilitate coordination.}
+Therefore, employing forecasting techniques that consider the hierarchical and/or grouped patterns of time series in EMS aligns naturally, offering the possibility to enhance forecast accuracy and facilitate coordination.
 
-To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
+\textcolor{teal}{To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
 There is a total national budget to be split between the regions in proportion to the forecast number of incidents in each region.
@@
 Then any trends or other forecast characteristics at national level will also be reflected in the regional forecasts, and the probabilistic forecasts allow for the different levels of uncertainty in the two regions.
-Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.
+Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.}
 
@@
 Generating independent forecasts can result in a lack of consistency and coordination, and therefore leads to less effective planning and decision making.
-\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
-If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area.
-Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others.
-Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly.
-Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}
 

### 2023-12-14 manuscript/main.qmd
finalised comments of AE

@@
 
-R code to produce all results in this paper is available at [github.com/bahmanrostamitabar/forecasting-emergency-medicine](https://github.com/bahmanrostamitabar/forecasting-emergency-medicine).
-The repository contains the raw data, all R scripts used in experiments as well as the results used in this paper, along with the raw files for producing this paper.
+To enhance transparency and reproducibility, we not only provide data and the code, but also the entire paper that is written in R using [Quarto](https://quarto.org/).
+All materials to reproduce this paper is available at [github.com/bahmanrostamitabar/forecasting-emergency-medicine](https://github.com/bahmanrostamitabar/forecasting-emergency-medicine).
+The repository contains the raw data, all R scripts used in experiments, the results used in the paper, as well as the quarto files for producing this paper.
 Full instructions are provided in the repository.

### 2023-12-14 manuscript/main.tex
finalised comments of AE

@@
 \makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \@ifpackageloaded{caption}{}{\usepackage{caption}}
@@
 \makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
 \makeatother
 \makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \makeatother
@@
 \end{frontmatter}
-    \section{Introduction}\label{sec-intro}
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, enhanced, frame hidden]}{\end{tcolorbox}}\fi
+
+\hypertarget{sec-intro}{%
+\section{Introduction}\label{sec-intro}}
 
@@
 
-\section{Research background}\label{sec-lit}
+\hypertarget{sec-lit}{%
+\section{Research background}\label{sec-lit}}
 
@@
 
-\begin{table}
-
-\caption{\label{tbl-literature}Summary of some studies on forecasting in
-Emergency Medical Services}
-
-\centering{
+\hypertarget{tbl-literature}{}
+\begin{landscape}\begin{table}
+\caption{\label{tbl-literature}Summary of some studies on forecasting in Emergency Medical Services }\tabularnewline
 
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}ll}
+\begin{tabular}{lr>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{4em}>{\raggedright\arraybackslash}p{10em}>{\raggedright\arraybackslash}p{8em}lll}
 \toprule
-Reference & Year & Variable & Horizon & Method & Metric & Probabilistic & Reconciliation\\
+Reference & Year & Variable & Horizon & Method & Metric & Probabilistic & Reconciliation & Reproducibility\\
 \midrule
-Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES\\
-Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO\\
-Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO\\
-Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO\\
+Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES & YES\\
+Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO & No\\
+Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO & No\\
+Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO & No\\
 Ibrahim et al.
-& 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO\\
-Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO\\
-Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO\\
+& 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO & No\\
+Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO & No\\
+Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO & No\\
 \bottomrule
@@
 \endgroup{}
-
-}
-
-\end{table}%
-
-\section{Experiment setup}\label{sec-experiment}
+\end{table}
+\end{landscape}
+
+\hypertarget{sec-experiment}{%
+\section{Experiment setup}\label{sec-experiment}}
 
@@
 
-\subsection{Data}\label{sec-data}
+\hypertarget{sec-data}{%
+\subsection{Data}\label{sec-data}}
 
@@
 
-\centering{
-
-\includegraphics[width=0.6\textwidth,height=\textheight]{main_files/figure-pdf/fig-hierarchy-1.pdf}
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{main_files/figure-pdf/fig-hierarchy-1.pdf}
+
+}
 
@@
 and then into 7 different healthboards.}
-
 \end{minipage}%
@@
 \begin{minipage}[t]{0.02\linewidth}
-~\end{minipage}%
+
+{\centering
+
+~
+
+}
+
+\end{minipage}%
 %
@@
 
-\centering{
-
-\includegraphics[width=0.6\textwidth,height=\textheight]{../img/group.png}
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{../img/group.png}
+
+}
 
@@
 and grouped levels.}
-
 \end{minipage}%
@@
 
-\end{figure}%
+\end{figure}
 
@@
 
+\hypertarget{tbl-hierarchy}{}
 \begin{table}
-
-\caption{\label{tbl-hierarchy}Number of time series in each level for
-the hierarchical \& grouped structure of attended incidents}
-
-\centering{
+\caption{\label{tbl-hierarchy}Number of time series in each level for the hierarchical \& grouped
+structure of attended incidents }\tabularnewline
 
@@
 \end{threeparttable}
-
-}
-
-\end{table}%
+\end{table}
 
@@
 
-\centering{
-
-\includegraphics[width=0.7\textwidth,height=\textheight]{main_files/figure-pdf/fig-feature-1.pdf}
+{\centering \includegraphics[width=0.7\textwidth,height=\textheight]{main_files/figure-pdf/fig-feature-1.pdf}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-dataviz2-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-dataviz2-1.pdf}
 
@@
 
-\subsection{Forecasting methods}\label{forecasting-methods}
+\hypertarget{forecasting-methods}{%
+\subsection{Forecasting methods}\label{forecasting-methods}}
 
@@
 
-\subsection{Performance evaluation}\label{performance-evaluation}
+\hypertarget{performance-evaluation}{%
+\subsection{Performance evaluation}\label{performance-evaluation}}
 
@@
 errors rather than absolute errors:
-\begin{equation}\phantomsection\label{eq-RMSSE}{
+\begin{equation}\protect\hypertarget{eq-RMSSE}{}{
   \text{MSSE} = \text{mean}(q_{j}^2),
-}\end{equation} where, \[
+}\label{eq-RMSSE}\end{equation} where, \[
   q^2_{j} = \frac{ e^2_{j}}
@@
 performance of the forecast distribution.
-\begin{equation}\phantomsection\label{eq-CRPS}{
+\begin{equation}\protect\hypertarget{eq-CRPS}{}{
   \text{CRPS} = \text{mean}(p_j),
-}\end{equation} where \[
+}\label{eq-CRPS}\end{equation} where \[
   p_j = \int_{-\infty}^{\infty} \left(G_j(x) - F_j(x)\right)^2dx,
@@
 
+\hypertarget{sec-htc}{%
 \section{Hierarchical and grouped time series forecasting
-techniques}\label{sec-htc}
+techniques}\label{sec-htc}}
 
@@
 
+\hypertarget{independent-base-forecast}{%
 \subsection{Independent (base
-forecast)}\label{independent-base-forecast}
+forecast)}\label{independent-base-forecast}}
 
@@
 
-\subsection{Reconciliation methods}\label{reconciliation-methods}
+\hypertarget{reconciliation-methods}{%
+\subsection{Reconciliation methods}\label{reconciliation-methods}}
 
@@
 
-\subsubsection{Notations}\label{notations}
+\hypertarget{notations}{%
+\subsubsection{Notations}\label{notations}}
 
@@
 
+\hypertarget{bottom-up-bu-and-linear-reconciliation-methods}{%
 \subsubsection{Bottom-up (BU) and linear reconciliation
-methods}\label{bottom-up-bu-and-linear-reconciliation-methods}
+methods}\label{bottom-up-bu-and-linear-reconciliation-methods}}
 
@@
 
-\section{Results and discussion}\label{sec-results}
+\hypertarget{sec-results}{%
+\section{Results and discussion}\label{sec-results}}
 
@@
 
+\hypertarget{tbl-result}{}
 \begin{table}
-
-\caption{\label{tbl-result}Average forecast performance calculated on
-the test sets at forecast horizons \(h=43,\dots,84\) days, with time
-series cross validation applied to attended incident data. The test set
-consists of 462 days. The best approach is highlighted in bold. Point
-forecast accuracy is measured using MASE and MSSE, while probabilistic
-forecast accuracy is measured using CRPS.}
-
-\centering{
-
+\caption{\label{tbl-result}Average forecast performance calculated on the test sets at forecast
+horizons \(h=43,\dots,84\) days, with time series cross validation
+applied to attended incident data. The test set consists of 462 days.
+The best approach is highlighted in bold. Point forecast accuracy is
+measured using MASE and MSSE, while probabilistic forecast accuracy is
+measured using CRPS. }\tabularnewline
+
+\centering
 \begin{tabular}{llrrrr}
@@
 \end{tabular}
-
-}
-
-\end{table}%
+\end{table}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-accuracy-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-accuracy-1.pdf}
 
@@
 
+\hypertarget{an-illustration-of-probabilistic-forecast-for-ems-demand}{%
 \subsection{An illustration of probabilistic forecast for EMS
-demand}\label{an-illustration-of-probabilistic-forecast-for-ems-demand}
+demand}\label{an-illustration-of-probabilistic-forecast-for-ems-demand}}
 
@@
 
-\centering{
-
-\includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}
 
@@
 
-\end{figure}%
-
-\section{Conclusion}\label{sec-conclusion}
+\end{figure}
+
+\hypertarget{sec-conclusion}{%
+\section{Conclusion}\label{sec-conclusion}}
 
@@
 
-\section*{Reproducibility}\label{reproducibility}
+\hypertarget{reproducibility}{%
+\section*{Reproducibility}\label{reproducibility}}
 \addcontentsline{toc}{section}{Reproducibility}

### 2023-12-16 journal_correspondence/response_letter/R2_response_letter_JSR.qmd
second revision finalised

@@
     include-in-header: header.tex
+execute: 
+  echo: false
+  warning: false
 ---
+
+```{r}
+#| label: setup
+#| context: setup
+#| include: false
+#| echo: false
+library(fpp3)
+library(knitr)
+library(kableExtra)
+library(ggthemes)
+library(patchwork)
+library(scales)
+library(ggdist)
+# Color blind color palettes
+options(
+  ggplot2.discrete.colour = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442"),
+  ggplot2.discrete.fill = c("#D55E00", "#0072B2", "#009E73", "#CC79A7", "#E69F00", "#56B4E9", "#F0E442")
+)
+# Few theme
+ggplot2::theme_set(ggthemes::theme_few())
+```
 
@@
 I do not plan to send this out for review again.
-I reserve the right to reject the work if there is not sufficient progress made in addressing the issues identified by the review team in the next submission.
-(Bahman)*
+I reserve the right to reject the work if there is not sufficient progress made in addressing the issues identified by the review team in the next submission.*
 
-Response:
+  **Response:**  Thank you for your detailed feedback and thorough review of the manuscript.
+We appreciate the acknowledgement of the significance of our work as well as the improvements made in response to the initial reviews.
+We have supplied a point-by-point response to the individual points identified, and the revised version includes the appropriate clarifications, details, and corrections.
+To indicate the changes made in the revised manuscript, we use a different colored font.
+We have addressed specific concerns mentioned about model specifics, explanations, and elaborations, as well as identified typos, formatting issues, and ideas for improvement.
+We also recognize the significance of catering to readers who are methodologically trained as well as managerial or policy-oriented.
+We presented the material in a way that effectively communicates the importance of our work to a wide range of audiences.
 
@@
 Additions have been made to explain the trend and seasonality strength measures, provide the GLM equation, and resolve minor formatting issues.
-These changes are beneficial.
-(Bahman)_
+These changes are beneficial._
 
@@
 
-  **Response:**
+  **Response:** Thank you for the feedback.
+Wherever necessary, we responded to all comments raised in both the previous and current versions.
+We are confident that the current study is one of the first to publish a publication using a cutting-edge scientific report, such as Quarto, sticking to transparency and reproducibility principles.
+This study is completely reproducible and transparent, and the reviewer comments criticizing the research's credibility and reproducibility astound us.
 
@@
 The authors added brief explanations of the trend and seasonality strength measures in response to Comment #3.
-While this provides more context, a citation or reference for these specific definitions would further strengthen this section (Bahman)._
+While this provides more context, a citation or reference for these specific definitions would further strengthen this section._
 
-  **Response:**
+  **Response:** Thank you.
+We have included @mstl, which describes the approach to decomposing the time series used to compute the strength of trend and seasonality.
+Additionally, we have included @hyndman2021forecasting, which also describe the strength of trend and seasonality.
 
-_Comment 2. The revisions partially address the lack of model detail raised in Comment #4, but some components are still unclear (Bahman):_
+_Comment 2. The revisions partially address the lack of model detail raised in Comment #4, but some components are still unclear:_
 
-  **Response:**
+  **Response:** Thank you very much.
+Section 3.2 contains further information on all models.
+The section contains significantly more details than the first version of the manuscript.
+We have also made available the Github repository, which contains the whole manuscript written in Quarto, as well as the technical details and R codes for the forecasting methodologies used in this study.
+We think that the amended section, together with the Github repository, will provide readers with the information they need to understand and replicate the methodologies employed in this study.
 
@@
 For ETS, it is now stated that an automated algorithm selects the best configuration, but no information is given on what options are available for trend, seasonality, etc.
-Listing the choices would help the reader.
-(Bahman)._
+Listing the choices would help the reader._
 
-  **Response:**
+  **Response:** Thank you.
+This has been included in the previosu version of the paper.
+The papers highlights that ETS models [@hyndman2021forecasting] can combine trend, seasonality, and error components in a time series through various forms that can be additive, multiplicative or mixed.
+The trend component can be none ("N"), Additive ("A") or damped ("Ad"); the seasonality can be none ("N"), Additive ("A"), or multiplicative ("M"); and the error term can be additive ("A") or multiplicative ("M").
+Please refer to "Exponential Smoothing State Space model (ETS)" in section 3.2..
 
-_Comment 2.2. The equation for the GLM model is now provided, but the description still does not mention monthly seasonality. This should be explicitly addressed. (Bahman)_
+_Comment 2.2. The equation for the GLM model is now provided, but the description still does not mention monthly seasonality. This should be explicitly addressed._
 
-  **Response:**
+  **Response:** Thank you.
+The GLM model employed in this study takes account of weekly seasonality and annual seasonality.
+Monthly seasonality in time series data is extremely rare, and it does not exist in the ambulance demand used in this study.
+There is no reason for occurrences to occur more frequently at certain times of the month than others.
 
-_Comment 2.3. The concept of ”serial dependence” in TSGLM needs more clarification and specificity to the EMS context. (Bahman)_
+  We have highlighted this in the revised paper:
 
-  **Response:**
+  \textcolor{teal}{Monthly seasonality in time series data is extremely rare, and it does not exist in the ambulance demand used in this study.
+There is no reason for occurrences to occur more frequently at certain times of the month than others.}
 
-_Comment 2.4. The rationale for including the naive method in the ensemble is still not provided. Its apparent poor performance suggests it should be excluded (Rob, Bahman)._
+_Comment 2.3. The concept of ”serial dependence” in TSGLM needs more clarification and specificity to the EMS context._
 
-  **Response:**
+  **Response:** Thank you.
+We have clarified the concept of serial dependence and its specificity to the EMS context in the revised manuscript.
+Serial dependence in the context of EMS forecasting refers to scenarios in which ambulance demand at one point in time correlates with demand at previous points in time.
+In other words, the time series' previous values are associated with its current values, which can be described in a variety of ways.
+The TSGLM model, for example, clearly accounts for serial dependence by incorporating lagged ambulance demand numbers into the model.
+We have included the following paragraph in the revised manuscript to address this comment:
+
+  \textcolor{teal}{The term serial dependence refers to instances in which the number of incidents on a current day correlates with the number of incidents on previous days.}
+
+_Comment 2.4. The rationale for including the naive method in the ensemble is still not provided. Its apparent poor performance suggests it should be excluded._
+
+  **Response:** Thank you very much.
+In response to your prior comment, we re-ran the experiment, removing the naive method from the ensemble.
+We then produced a graph that shows the outcomes with and without the naive method.
+We showed that removing the naive from the ensemble worsens overall performance, thus we kept it in the ensemble and included the original plot in the revised publication.
 
@@
 The explanation of the training/validation/testing data split in Comment #5 remains unclear.
-The authors should explicitly state if tuning is done on the training set only or if a separate validation set is used.
-(Bahman)_
+The authors should explicitly state if tuning is done on the training set only or if a separate validation set is used._
 
-  **Response:**
+  **Response:** Thank you.
+We explicitly mentioned that we employed time series cross validation [@hyndman2021forecasting,Section 5.10] to assess the forecast accuracy.
+Time series cross-validation involves dividing the data into multiple training and testing sets based on the time series structure.
+Model development and hyper-parameter tuning is performed using the training data and the errors are assess using the corresponding test set.
+Each testing set consists of a number of observations equal to the forecast horizon, and the corresponding training set only includes observations that occurred before the test set.
+This ensures that future observations are not used in constructing the forecast, making it more robust to evaluate forecast accuracy on new data.
+Therefore, we believe that our methodology is appropriate for time series forecasting and provides a more robust approach for evaluating forecast accuracy on new data.
+This has been highlighted in the revised paper in section 3.3 :
 
-_Comment 4. The forecast horizon explanation in Comment #6 is still ambiguous. The time unit (days, weeks, etc.) should be clearly defined. (Bahman)_
+  \textcolor{teal}{Model development and hyper-parameter tuning is performed using the training data and the errors are assess using the corresponding test set.}
 
-  **Response:**
+_Comment 4. The forecast horizon explanation in Comment #6 is still ambiguous. The time unit (days, weeks, etc.) should be clearly defined._
+
+  **Response:** Thank you very much.
+We explicitly stated in several sections of the paper, including the introduction and Section 3, that the forecasting horizon ranges from 1 to 84 days.
+Please see the Introduction, the beginning of section 3, and section 3.4.
 
@@
 Explicitly defining concepts like sharpness and miscalibration would strengthen this section.
-Additionally, more specifics are needed on the forecast distribution derivation.
-(Rob)_
+Additionally, more specifics are needed on the forecast distribution derivation._
 
-  **Response:**
+  **Response:** We have provided detail explanation of CRPS and cleanly elaborated on the concepts of calibration and sharpness.
+We have also discussed how probabilsitic forecasts are computed.
+We have included the following text in the revised version to address this comment:
 
-_Comment 6.
-Section 4.2 still lacks concrete examples tying the hierarchical forecasting concepts to the EMS dataset, as noted in Comment #8.
-This example-based explanation needs enhancement.
-(Bahman)_
+  \textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+It measures how well the predicted probabilities match the actual probabilities.
+On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
+A model is well-calibrated if the predicted probabilities match the actual probabilities, and it is sharp if it is confident in its predictions.
+The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.
+We chose the CRPS as one of our evaluation metrics because it provides a comprehensive assessment of the forecast quality and allows us to compare the performance of different models.}
 
-  **Response:**
+\textcolor{teal}{To generate forecast probability distributions, we did not assume a specific distribution; instead we used a form of bootstrapping, described in} @panagiotelis2023probabilistic.
+\textcolor{teal}{This involves simulating 1000 future sample paths from each of the model, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+In this way, we can generate an empirical distribution of forecasts for each model.
+The ensemble forecast distribution is a simple mixture of these empirical distributions.}
+
+_Comment 6. Section 4.2 still lacks concrete examples tying the hierarchical forecasting concepts to the EMS dataset, as noted in Comment #8. This example-based explanation needs enhancement._
+
+  **Response:** Thank you. We have now added a more detailed but simple example in the introduction based on the ambulance data and added the following text to the revised version of the paper:
+
+\textcolor{teal}{To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
+There is a total national budget to be split between the regions in proportion to the forecast number of incidents in each region.
+The two regions have very different incident patterns, and so must be forecast using different models.
+However, the data are noisy at regional level, so the national forecasts are best obtained by summing the demand from the two regions.
+The resulting national forecasts are not equal to the sum of the regional forecasts, and so are not coherent.
+In fact, the national forecasts show a decreasing trend in demand, and so the national governing body decides to cut the budget for the next year.
+But neither of the regional forecasts shows a trend, and so the regions argue that the budget cut is unfair.
+In addition, Region A has much more variable demand than Region B, and so to cope with periods of peak demand, Region A needs to hold more resources in reserve.
+So the budget distribution needs to be made in a way that ensures the probability of each region being unable to meet demand is equal.
+Our solution to this problem is to use a hierarchical forecasting approach that ensures the forecasts are probabilistically coherent.
+Then any trends or other forecast characteristics at national level will also be reflected in the regional forecasts, and the probabilistic forecasts allow for the different levels of uncertainty in the two regions.
+Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.}
 
@@
 Comment #9 about clarifying the differences between various hierarchical forecasting approaches does not seem to be addressed.
-This terminology is still confusing in Section 4.2.2 and requires revision.
-(Rob)_
+This terminology is still confusing in Section 4.2.2 and requires revision._
 
-  **Response:**
+  **Response:** In the amended version, we have given a detailed overview of various hierarchical forecasting approaches, which are presented below:
+
+\textcolor{teal}{The bottom-up (BU) approach is constrained by its reliance solely on base forecasts from a single level of aggregation at the bottom level.
+While it does result in consistent forecasts, the BU approach lacks forecast reconciliation since no reconciliation is performed.}
+
+\textcolor{teal}{Forecast reconciliation approaches bridge this gap by combining and reconciling all base forecasts to generate coherent forecasts.
+This technique utilizes all the base forecasts produced within a hierarchical structure to create consistent forecasts at every level of the hierarchy.
+As a result, it goes beyond relying solely on base forecasts from a single level of aggregation, and instead leverages all available information at each level to generate forecasts that minimize the total forecast variance of the set of coherent forecasts.}
+
+\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method for estimating the parameters in linear regression models.
+In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors have equal variance.
+Hence, $\bm{W}$ is simply defined as the identity matrix multiplied by a constant factor.
+The intuition behind OLS is that it minimizes the sum of squared residuals between the observed and predicted values of the dependent variable.
+The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
+A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}
+
+\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
+But it assumes that the errors of each series are uncorrelated with each other.
+In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal.
+The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
+The main weakness of this approach is that it ignores the relationships between series.
+A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}
+
+\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.} @WicEtAl2019 \textcolor{teal}{showed that this approach gives the optimal reconciled forecasts in the sense that the sum of the forecast variances is minimized.
+The main weakness of this approach is that it is difficult to estimate the full covariance matrix.
+In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}
 
@@
 
-  **Response:**
+  **Response:** Thank you for the comment.
 
-_Comment 9. My concerns in Comment #11 about non-integer forecasts and rounding do not appear to be addressed. This practical issue needs to be discussed. (Bahman)_
+The forecast reconciliation approach does not aggregate lower level forecasts, instead it combines all base forecasts.
+The base forecasts are generated independently at each level of the hierarchy, ignoring the hierarchical/grouped structure of time series or any aggregation constraints.
+As a result, the base forecasts are not guaranteed to be coherent, meaning that the sum of the lower-level forecasts will not necessarily add up to the higher-level forecast.
+On the other hand, the forecast reconciliation approach takes the base forecasts and regenerates them in a way that ensures coherence across all levels of the hierarchy, while also considering all available information.
+The comparison between the base forecasts and the reconciled forecasts allows us to see the improvement in forecast accuracy that can be achieved by using the forecast reconciliation approach.
 
-  **Response:**
+While the approach is designed to improve forecast accuracy, it may not always be successful in doing so, particularly if the bottom-level series are too noisy and lack systematic patterns.
+In such cases, the benefit of performing reconciliation at the bottom level may still lie in creating coherent forecasts that can help align planning across different teams in an organization, improve coordination, and avoid conflicting decisions.
+Additionally, even if the bottom-level series are noisy and lack systematic patterns, reconciliation can still lead to more accurate forecasts at higher levels of the hierarchy by utilizing the information available across the hierarchy.
+Therefore, even if the bottom-level forecasts might not be very accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and possibly more accurate forecasts at other levels.
+
+We have added the following paragraph in the revised version of the manuscript to address this comment:
+
+\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
+Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy.
+Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}
+
+_Comment 9. My concerns in Comment #11 about non-integer forecasts and rounding do not appear to be addressed. This practical issue needs to be discussed._
+
+  **Response:** Thank you for highlighting this point, we have now included a paragraph at the end of the result section to acknowledge this point.
+
+Rounding forecasts up or down and its effect on the forecast accuracy depends on the level of hierarchy and scale of data.
+For some situations, with high volume demand, forecast accuracy calculations can ignore integer effects as rounding becomes negligible.
+However, low volume demand settings, such as forecasts at the bottom level of the hierarchy, may be more susceptible to integer (rounding) effects.
+In practice, we may need to use integer forecasts, especially when the forecasts are relatively small.
+Count forecast reconciliation is an active area of research, and it would be interesting to explore in future research how our approach could be adapted to generate count reconciled probabilistic forecasts.
+
+\textcolor{teal}{Despite using Poisson regression models to create count distributions of attended incidents for the base forecasts, it is important to note that the reconciled forecast distributions do not maintain a count format.
+In practical scenarios, there might be a need to use integer forecasts.
+Count forecast reconciliation is an active area of research, and it would be interesting to explore how our approach could be adapted to generate count-reconciled probabilistic forecasts in future studies.
+Rounding the forecasts is one possible solution to this problem.
+However, the impact of rounding on forecast accuracy varies depending on the level of hierarchy and the scale of the data.
+In situations with high volume demand, the effects of rounding may be negligible, and forecast accuracy calculations can overlook integer effects.
+On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.}
 
@@
 
-_General comment: The author revised the manuscript upon the review comments and I appreciate the detailed introductions added to the body of the paper.I provide some follow-up comments below for consideration and reference.
-(Bahman)_
+_General comment: The author revised the manuscript upon the review comments and I appreciate the detailed introductions added to the body of the paper.I provide some follow-up comments below for consideration and reference._
 
-  **Response:**
+  **Response:** Thank you for your positive feedback and appreciation of the detailed added to the paper.
+We have carefully reviewed your follow-up comments and take them into consideration as we continue to refine the manuscript.
 
-_Comment 1. Might be good to at least briefly mention the extended applicability of the idea early in the introduction section. (Bahman)_
+_Comment 1. Might be good to at least briefly mention the extended applicability of the idea early in the introduction section._
 
-  **Response:**
+  **Response:** Thank you for the suggestion. We have included the following paragraph in the introduction to address this comment.
 
-_Comment 2. It is confusing to have the abbreviations BME1 and BME2 listed with the same full name while no explanations or references are given. This is at Section 2, Page 7. (Bahman)_
+  \textcolor{teal}{While our research focuses on emergency medical services, it is important to emphasize the suggested framework's adaptability, which expands its relevance to a variety of service sectors such as supply chains, tourism, finance, and call centers.
+Our approach can be generalized in cases with hierarchically structured and/or grouped time series data, which is common in many sectors.}
 
-  **Response:**
+_Comment 2. It is confusing to have the abbreviations BME1 and BME2 listed with the same full name while no explanations or references are given. This is at Section 2, Page 7._
+
+  **Response:** Thank you for the feedback.
+We used the notation mentioned in the literature study.
+However, since it appears that this could be confusing, we have changed the wording to refer to these models as versions of bivariate mixed-effects models.
+We included a reference to the study that contains details regarding these variations.
 
@@
 The author added great details to the methods mentioned in Section 3.2, such as the well-explained model specifications of GLM and TSGLM.
-Make sure typos are corrected, especially notation-wise such as subscripts.
-(Bahman)_
+Make sure typos are corrected, especially notation-wise such as subscripts._
 
-  **Response:**
+  **Response:** Thank you for your suggestions.
+The document has been proofread thoroughly.
+We carefully corrected all typos, including punctuation and those relating to notation and subscripts, to ensure the manuscript's correctness and clarity.
 
@@
 It may be good to consider an alternatively way of providing details to the linear reconciliation method.
-For instance, the main equation \tilde{y}_h is still missing important reasoning about how it is derived.
-With some better descriptions of how it is derived, such as an objective function, the follow-up paragraphs for the three methods can be more concisely summarised with better readability.
-(Rob)_
+For instance, the main equation tilde{y}_h is still missing important reasoning about how it is derived.
+With some better descriptions of how it is derived, such as an objective function, the follow-up paragraphs for the three methods can be more concisely summarised with better readability._
 
+  **Response:** Thank you for the comment.
 We have now reworded this section, and added some more details about the objective function which is optimized.
+We have included the following text in the section 4.2.2 of the revised manuscript:
 
-  **Response:**
+  \textcolor{teal}{When $\bm{W}_h$ is the covariance matrix of $\hat{\bm{y}}_h$, the resulting forecasts are optimal in the sense that the sum of the variances of the reconciled forecasts is minimized, provided the base forecasts $\hat{\bm{y}}_h$ are unbiased.
+However, $\bm{W}_h$ is difficult to estimate, and so there have been various suggested approximations to $\bm{W}_h$, leading to different types of reconciliation such as Ordinary Least Squares (OLS), Weighted Least Squares (WLS) and Minimum Trace (MinT).}
 
@@
 It is not necessary to refer them in the main paper, but some self-explanatory captions should suffice.
-Besides, comparison with competing methods regarding the same aspect would be a plus.
-(Bahman)_
+Besides, comparison with competing methods regarding the same aspect would be a plus._
 
-  **Response:**
+  **Response:** Thank you for your feedback.
+Although the distribution is generated for the whole 84-day prediction and all 1530 time series across the hierarchy, visualizing each of these distributions and forecast horizons to fit into the manuscript is not feasible.
+As a result, we've given a sample example at the total country level that illustrates the probability distribution over a 7-day period.
+This gives readers a clear understanding that can be accommodated within the constraints of the manuscript.
+The code is available on our GitHub repository for those interested in investigating additional time series or forecast horizons, allowing readers to reproduce the plot for any chosen time series and forecast length.
 
-_Comment 6. There are still formatting issues in the included tables. (Bahman)_
+We have included the following plots in the revised version of the paper:
 
-  **Response:**
+@fig-forecast-density-hstep \textcolor{teal}{depicts the forecast distribution of total incidents in one health board over a 7-day period.
+It also gives the point forecast as well as the 80\% and 90\% prediction intervals.} @fig-forecast-density \textcolor{teal}{zooms in on the first day to show the histogram more clearly, illustrating the range of possible outcomes and their likelihood.}
+
+```{r}
+#| label: fig-forecast-density-hstep
+#| echo: false
+#| out.width: "85%"
+#| fig-cap: "A graphical illustration of the forecast distribution of ambulance demand (i.e.
+total incidence attended) for the SB health board for a horizon of seven days.
+For each day, we display the point forecast (black point), the histogram, and 80% (thick line) and 90% (thin line) prediction intervals.
+It also shows a portion of a historical time series as well as its fitted values."
+#| fig-pos: "H"
+incident_gthf <- readr::read_rds(here::here("data/incidents_gt.rds"))
+
+incident_cv <- incident_gthf |> filter(lhb == "SB", is_aggregated(category), is_aggregated(nature))
+fit <- incident_cv %>%
+  model(
+    ets = ETS(sqrt(incident))
+  )
+set.seed(2023)
+fcst <- fit %>% generate(h = 7, times = 5000, bootstrap = TRUE)
+
+fitted_ets <- fit |> augment()
+fcst_point <- fcst |> index_by(date) |> summarise(.mean=mean(.sim))
+p1 <- ggplot(data = fcst, mapping = aes(x = date, y = .sim))+
+  stat_histinterval(point_interval = "mean_qi",.width = c(0.80, 0.95),breaks=15)+
+  geom_line(data=fcst_point,aes(y=.mean, shape ="Point Forecast"))+
+  geom_line(aes(y = .fitted, shape ="Fitted"), data = filter_index(fitted_ets, "2019-07-17" ~ .))+
+  geom_point(aes(y = .fitted, shape ="Fitted"), data = filter_index(fitted_ets, "2019-07-17" ~ .))+
+  geom_line(aes(y = incident, shape ="Actual"),data = filter_index(incident_cv, "2019-07-17" ~ .))+
+  geom_point(aes(y = incident, shape ="Actual"),data = filter_index(incident_cv, "2019-07-17" ~ .))+
+  scale_shape_manual(name=NULL,
+                   breaks=c('Actual','Fitted','Point Forecast'),
+                   values=c('Actual'=17,'Fitted'=15, 'Point Forecast'=19))+
+  labs(x= "Date", y="Incidents")+
+  theme_few()+
+  theme(legend.position = "top")
+p1
+```
+
+```{r}
+#| label: fig-forecast-density
+#| out.width: 85%
+#| fig-cap: "An illustrative example of the forecast distribution of ambulance demand (i.e.
+total incidence attended) for the SB health board for one day ahead.
+This corresponds to the first forecast distribution in Figure 5.
+The horizontal axis shows all possible outcomes that may occur, with their likelihood shown on the vertical axis.
+The point in the middle shows the point forecast.
+Two lines at the bottom of the distribution highlights 80% (thick line) and 90% (thin line) prediction intervals."
+#| fig-pos: "H"
+
+fcst1 <- fcst %>% filter_index("2019-08-01")
+
+p2 <- ggplot(data = fcst1, mapping = aes(x = .sim))+
+  stat_histinterval(point_interval = "mean_qi",.width = c(0.80, 0.95),breaks=40)+
+  scale_x_continuous(breaks = seq(60, 140, 5))+
+  scale_y_continuous(labels = percent)+
+  labs(y = "Density", x = "Incidents")+
+  theme_few()
+p2
+```
+
+_Comment 6. There are still formatting issues in the included tables._
+
+  **Response:**  Thank you very much. We have now reviewed the format of all tables and resolved any issues that occurred. For more information, refer to Tables 1, 2, and 3.
 

### 2023-12-16 manuscript/01-introduction.qmd
second revision finalised

@@
 
-\textcolor{teal}{To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
+To illustrate the problem, let's consider a very simple example where we have an EMS provider with a national level of governance, and two regions (A and B), each with a health board and a station.
 There is a total national budget to be split between the regions in proportion to the forecast number of incidents in each region.
@@
 Then any trends or other forecast characteristics at national level will also be reflected in the regional forecasts, and the probabilistic forecasts allow for the different levels of uncertainty in the two regions.
-Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.}
+Budget can be allocated by controlling the probability of demand exceeding available resources, rather than being simply in proportion to the expected demand.
 
@@
 regional) levels, as well as groups by priority and nature of incidents.
-We produce consistent point forecasts and forecast distributions for all levels, which is critical for an effective planning and associated risk management.
+We produce consistent point forecasts and forecast distributions for all levels over 84 days horizon, which is critical for an effective planning and associated risk management.
 We compare the point and probabilistic forecast accuracy of the independent forecasts, bottom-up and optimal reconciliation approaches.
@@
 Therefore, they can be applied to any healthcare service (e.g., emergency department, primary or social care) subject to the time series having a hierarchical and/or grouped structure.
+While our research focuses on emergency medical services, it is important to emphasize the suggested framework's adaptability, which expands its relevance to a variety of service sectors such as supply chains, tourism, finance, and call centers.
+Our approach can be generalized in cases with hierarchically structured and/or grouped time series data, which is common in many service sectors.
 

### 2023-12-16 manuscript/02-literature.qmd
second revision finalised

@@
 The tool contains options that allow generating forecasts at various levels of granularity; however, it ignores the structure of the hierarchical and grouped relationships, preventing aligned decision making and coordination.
-@al2021empirical \textcolor{teal}{utilises data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand.
+@al2021empirical utilises data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand.
 The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
@@
 The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
-However, when it comes to long-term demand, the SSA method proves to be the most effective.} @ibrahim2016modeling \textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
-Call centre data were utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
+However, when it comes to long-term demand, the SSA method proves to be the most effective.
+@ibrahim2016modeling conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), and two variations of bivariate mixed-effects models (BME).
+Call centre data were utilized to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
 The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
 The findings indicate that the ME consistently produces the most accurate point forecasts.
-On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
-For a two-week leading period, MU shows better coverage probability.}
+On the other hand, BME models demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
+For a two-week leading period, MU shows better coverage probability.
 

### 2023-12-16 manuscript/03-experiment.qmd
second revision finalised

@@
 #| fig-subcap:
-#| - "Hierarchical structure: Attended incidents in the whole country are disaggregated into 3 control areas and then into 7 different healthboards."
+#| - "Hierarchical structure: Attended incidents in the whole country are disaggregated into 3 control areas and then into 7 different healthboards, anonymized using two letters (e.g. AB)"
 #| - "Grouped structure: Incidents could be grouped into priority (i.e.
@@
 
-\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by} @mstl.
-\textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
+In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by @mstl.
+STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
 The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
-The strength of trend is defined as:}
+The strength of trend is defined as:
 $$F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)$$
-\textcolor{teal}{For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
+For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
 Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small.
-But for data with little or no trend, the two variances should be approximately the same.}
+But for data with little or no trend, the two variances should be approximately the same.
 
-\textcolor{teal}{The strength of seasonality is defined similarly:}
+The strength of seasonality is defined similarly:
 $$F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).$$
-\textcolor{teal}{series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).}
+series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).
 
@@
 
-In addition to displaying the trend and seasonality features, we also visualize a few time series at various levels of aggregation.
+In addition to displaying the trend and seasonality strength [@hyndman2021forecasting], we also visualize a few time series at various levels of aggregation.
 @fig-dataviz2 reveals different information such as trend, seasonality, and noise.
@@
 We use the empirical distribution of the past daily attended incidents to create the forecast distribution of future attended incidents.
-\textcolor{teal}{We have chosen this "stationary" method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
+We have chosen this "stationary" method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
 Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
 Additionally, the stationary method has shown surprisingly high accuracy.
-Hence, any forecasting approach that can offer superior results compared to the stationary method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
+Hence, any forecasting approach that can offer superior results compared to the stationary method would validate its practical use, otherwise there is no necessity for employing more complex methods.
 
@@
 
-\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series.
-Specifically, we utilize the `ETS()` function in the fable package of R, which employs Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
+In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series.
+Specifically, we utilize the `ets()` function in the forecast package of R, which employs Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
 Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
 Consequently, the automated algorithm selects the best model based on the unique characteristics of each individual time series.
-As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
+As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.
 
@@
 
-\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
+Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as
 \begin{align*}
@@
 \end{align*}
-\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the mean of the Poisson distribution.
-In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
-The Fourier terms are as defined in} @hyndman2021forecasting [Section 7.4] \textcolor{teal}{This model takes account of weekly seasonality and annual seasonality.
-Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
-There is no reason, for example, for incidents to occur more at some times of the month than others.}
+and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the mean of the Poisson distribution.
+In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is a public holiday, 0 otherwise), school holidays (1 when is a school holiday, 0 otherwise), and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
+The Fourier terms are as defined in @hyndman2021forecasting [Section 7.4].
+This model takes account of weekly seasonality and annual seasonality.
+Monthly seasonality in time series data is extremely rare, and it does not exist in the ambulance demand used in this study.
+There is no reason for occurrences to occur more frequently at certain times of the month than others.
 
@@
 **Poisson Regression using tscount (TSGLM):** We also consider another Poisson regression model that takes into account serial dependence.
-This model captures the short range serial dependence by including autoregressive terms, in addition to the same covariates that were used in the GLM model.
+This model captures the short-range serial dependence by including autoregressive terms in addition to the same covariates that were used in the GLM model.
 To distinguish this from the previous GLM model, we will refer to this model as `TSGLM`.
 
-\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional autoregressive component accounting for serial correlation.
-The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points, after accounting for the exogenous covariates.
-Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
-TSGLM can be written as}
+The Poisson TSGLM is similar to the GLM, with an additional autoregressive component accounting for serial dependence.
+The term serial dependence refers to instances in which the number of incidents on a current day correlates with the number of incidents on previous days.
 \begin{align*}
@@
 \end{align*}
-\textcolor{teal}{and ${y}_{t-k}$ is a vector of $k$ lagged values.
-The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
-past values) of the ambulance demand in the model.
-This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}
+and ${y}_{t-k}$ is a vector of $k$ lagged values.
+The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e., past values) of the ambulance demand in the model.
+This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.
 
-\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R} [@JSSv082i05] \textcolor{teal}{to model the attended incidents.
-Again, the logarithmic link function is used to ensure that the mean of the Poisson distribution is always positive.}
+We use the `tsglm()` function in the `tscount` package in R [@JSSv082i05] to model the attended incidents.
+Again, the logarithmic link function is used to ensure that the mean of the Poisson distribution is always positive.
 
-\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution} [@feller1991introduction, p156--158].
-\textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
+Provided accidents occur independently, they will inherently follow a Poisson distribution [@feller1991introduction, p156--158].
+Hence, it is reasonable to assume a Poisson distribution in this context.
 To account for changes over time, we incorporate trend and seasonality covariates, as well as public holiday effects, allowing the mean of the Poisson distribution to vary.
-However, it is important to note that if there are additional factors influencing the mean of the Poisson distribution, which are not accounted for in our model, we might observe over-dispersion or under-dispersion in the data.}
+However, it is important to note that if there are additional factors influencing the mean of the Poisson distribution that are not accounted for in our model, we might observe over- or under-dispersion in the data.
 
-**Ensemble method:** \textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by combining the forecasts, rather than relying on separate forecasts generated by each individual method} [@clemen1989combining].
-\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Stationary, ETS, GLM and TSGLM models using a simple average to form a mixture distribution} [@combinations].
+**Ensemble method:** Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by combining the forecasts rather than relying on separate forecasts generated by each individual method [@clemen1989combining].
+In this paper, we use an ensemble method that combines the forecasts generated from the Stationary, ETS, GLM, and TSGLM models using a simple average to form a mixture distribution [@combinations].
 
-\textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in} @panagiotelis2023probabilistic.
-\textcolor{teal}{This involves simulating 1000 future sample paths from each of the models, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
+To generate forecast probability distributions using the above methods, we use a form of bootstrapping, described in @panagiotelis2023probabilistic.
+This involves simulating 1000 future sample paths from each of the models by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
 In this way, we can generate an empirical distribution of forecasts for each model.
-The ensemble forecast distribution is a simple mixture of these empirical distributions.}
+The ensemble forecast distribution is a simple mixture of these empirical distributions.
 
-\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models, or to promote a particular model class.
-Instead, we have developed a flexible framework that can accommodate any forecasting models.
-Our primary objective is to demonstrate its practicality and effectiveness in integrating base forecasts from any model and generating coherent forecasts within a hierarchical structure.}
+It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models or to promote a particular model class.
+Instead, we have developed a flexible framework that can accommodate any forecasting model.
+Our primary objective is to demonstrate its practicality and effectiveness in integrating base forecasts from any model and generating coherent forecasts within a hierarchical structure.
 
@@
 
-\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by $j$, representing the number of time periods ahead we are predicting.
-In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
+The error metrics provided below consider a forecasting horizon denoted by $j$, representing the number of time periods ahead we are predicting.
+In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.
 
@@
 
-\textcolor{teal}{Using scale-independent measures, such as MASE and MSSE, enables more appropriate comparisons between time series at different levels and scales, as these measures are not influenced by the magnitude of the data.
+Using scale-independent measures, such as MASE and MSSE, enables more appropriate comparisons between time series at different levels and scales, as these measures are not influenced by the magnitude of the data.
 This is of particular importance in our study, as we work with time series at various levels of hierarchy, with varying scales, resulting in different magnitudes of error.
-By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
+By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.
 
@@
 
-\textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
+Calibration refers to the statistical consistency between the distributional forecasts and the observations.
 It measures how well the predicted probabilities match the observations.
@@
 The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
-Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.}
+Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.
 
-\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores} [@winkler1972decision;@hyndman2021forecasting, Section 5.9] \textcolor{teal}{or percentile scores} [@hyndman2021forecasting, Section 5.9], \textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.} \textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
+CRPS can be considered an average of all possible Winkler scores [@winkler1972decision;@hyndman2021forecasting, Section 5.9] or percentile scores [@hyndman2021forecasting, Section 5.9], and thus provides an evaluation of all possible prediction intervals or quantiles.
+A specific prediction interval could be evaluated using a Winkler score.
 Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.
 95\%) quantiles.
-In such cases, a percentile score becomes useful in meeting this specific requirement.}
+In such cases, a percentile score becomes useful in meeting this specific requirement.

### 2023-12-16 manuscript/04-htsgts.qmd
second revision finalised

@@
 and $\bm{A}$ is the $n_a\times n_b$ "aggregation" matrix specifying how the bottom-level series $\bm{b}_t$ are to be aggregated to form $\bm{a}_t$.
-\textcolor{teal}{The aggregation matrix $\bm{A}$ is determined by the structure of the hierarchy.
+The aggregation matrix $\bm{A}$ is determined by the structure of the hierarchy.
 It maps the bottom-level time series to the corresponding higher-level time series.
 For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $\bm{A} = \begin{bmatrix} 1 ~~ 1 \end{bmatrix}$.
-}
+
 The full vector of time series is given by
@@
 
-\textcolor{teal}{The term "bottom-level series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
+The term "bottom-level series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
 For instance, in Table 2, each distinct combination of values in Control area (e.g.
@@
 all country level), the time series are aggregated across all Control areas, Health boards, Priorities, and Natures of incidents.
-Any desired aggregation level can be achieved based on the data structure, utilizing the bottom-level series available.}
+Any desired aggregation level can be achieved based on the data structure, utilizing the bottom-level series available.
 
@@
 
-\textcolor{teal}{Forecast reconciliation approaches bridge this gap by combining and reconciling all base forecasts to generate coherent forecasts.
+Forecast reconciliation approaches bridge this gap by combining and reconciling all base forecasts to generate coherent forecasts.
 This technique utilizes all the base forecasts produced within a hierarchical structure to create consistent forecasts at every level of the hierarchy.
-As a result, it goes beyond relying solely on base forecasts from a single level of aggregation, and instead leverages all available information at each level to generate forecasts that minimize the total forecast variance of the set of coherent forecasts.} \textcolor{teal}{Linear reconciliation involves projecting the base forecasts onto the coherent space.
-It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased} [@WicEtAl2019].
+As a result, it goes beyond relying solely on base forecasts from a single level of aggregation, and instead leverages all available information at each level to generate forecasts that minimize the total forecast variance of the set of coherent forecasts.
+Linear reconciliation involves projecting the base forecasts onto the coherent space.
+It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased [@WicEtAl2019].
 
@@
 
-\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method.
+Ordinary Least Squares (OLS) is the simplest and most commonly used method.
 In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors are uncorrelated and have equal variance, and that multi-step forecast variances are proportion to one-step forecast variances.
@@
 The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
-A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}
+A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.
 
-\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
+Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
 But it assumes that the errors of each series are uncorrelated with each other, and that multi-step forecast variances are proportion to one-step forecast variances.
@@
 The main weakness of this approach is that it ignores the relationships between series.
-A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}
+A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.
 
-\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the one-step base forecast errors.
-So it takes account of both the scale of each series, and the relationships between the series.} But it still assumes
-that multi-step forecast variances are proportion to one-step forecast variances.
+Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the one-step base forecast errors.
+So it takes account of both the scale of each series, and the relationships between the series.
+But it still assumes that multi-step forecast variances are proportion to one-step forecast variances.
 The main weakness of this approach is that it is difficult to estimate the full covariance matrix, even of the one-step errors.
-In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.}
+In practice, we usually need to use a shrinkage estimate where the off-diagonal elements are shrunk towards zero.
 
@@
 
-\textcolor{teal}{Certainly, other approaches can be applied to hierarchical forecasting problems} @pennings2017integrated and @villegas2018supply \textcolor{teal}{proposed the idea of using a state space model to ensure consistent forecasts.
+Certainly, other approaches can be applied to hierarchical forecasting problems @pennings2017integrated and @villegas2018supply proposed the idea of using a state space model to ensure consistent forecasts.
 However, when dealing with larger hierarchies, these models encounter difficulties in estimating covariance matrices.
 In contrast, our approach provides a clear advantage by allowing the incorporation of different forecasting methods for the base forecasts, and even accommodating distinct methods for individual series.
-The decoupling of time series models from the reconciliation step adds significant flexibility in exploring a wide range of models.}
+The decoupling of time series models from the reconciliation step adds significant flexibility in exploring a wide range of models.

### 2023-12-16 manuscript/05-results.qmd
second revision finalised

@@
 However, this does not remain valid for bottom series where different individual methods perform best, depending on the accuracy measure.
-\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
+While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
 Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
 Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy.
-Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}
+Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.
 
@@
 
-\textcolor{teal}{Despite using Poisson regression models to create count distributions of attended incidents for the base forecasts, it is important to note that the reconciled forecast distributions do not maintain a count format.
+Despite using Poisson regression models to create count distributions of attended incidents for the base forecasts, it is important to note that the reconciled forecast distributions do not maintain a count format.
 In practical scenarios, there might be a need to use integer forecasts.
 Count forecast reconciliation is an active area of research, and it would be interesting to explore how our approach could be adapted to generate count-reconciled probabilistic forecasts in future studies.
-One possible approach to address this is by rounding the forecasts.
+Rounding the forecasts is one possible solution to this problem.
 However, the impact of rounding on forecast accuracy varies depending on the level of hierarchy and the scale of the data.
-In situations with high-volume demand, the effects of rounding may be negligible, and forecast accuracy calculations can overlook integer effects.
-On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.}
+In situations with high volume demand, the effects of rounding may be negligible, and forecast accuracy calculations can overlook integer effects.
+On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.
 
@@
 
-@fig-forecast-density \textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, of the total attended incidents in the PO health board.
-The black dashed line in the graph represents the point forecast (the mean of the distribution), which presents a single numerical estimate for the future number of attended incidents.
-Due to the complexity of including such plots for the entire hierarchy and 84 days ahead, only one example is presented here for 1 day ahead.
+In this section, we provide an illustrative example of a probabilistic forecast for future demand, of the total attended incidents in the SB health board.
+Due to the complexity of including such plots in the manuscript for the entire hierarchy and 84 days ahead, only one example is presented here.
 However, it is feasible to generate these plots for the entire hierarchy and for any forecast horizon if necessary.}
 
-\textcolor{teal}{In practice, point forecasts are commonly used, but they have limitations as they ignore the uncertainty associated with the forecast, leading to wrong decisions.
-In contrast, probabilistic forecasts offer an alternative approach to anticipate future demand.
-Rather than providing a single value, they assign likelihoods to all possible demand outcomes, acknowledging that different numbers of attended incidents are possible, but with varying likelihoods.}
+In practice, point forecasts are commonly used, but they have limitations as they ignore the uncertainty associated with the forecast.
+The future is inherently characterized by an irreducible level of uncertainty.
+Being prepared entails considering alternate courses of action.
+Probabilistic forecasts offer an alternative approach to anticipate future demand.
+Rather than providing a single value, they assign likelihoods to all possible demand outcomes, acknowledging that different numbers of attended incidents are possible, but with varying likelihoods.
 
-\textcolor{teal}{The purpose of probabilistic forecasting, as demonstrated in} @fig-forecast-density, \textcolor{teal}{is to quantify uncertainty.
+The purpose of probabilistic forecasting, as demonstrated in @fig-forecast-density-hstep and @fig-forecast-density, is to quantify uncertainty.
+@fig-forecast-density-hstep depicts the forecast distribution of total incidents in one health board over a 7-day period.
+It also gives the point forecast as well as the 80\% and 90\% prediction intervals.
+@fig-forecast-density zooms in on the first day to show the histogram more clearly, illustrating the range of possible outcomes and their likelihood.
+
 Decisions based on these forecasts could focus on the tails of the distribution: unexpectedly high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
@@
 Different EMS managements may have varying risk attitudes depending on resource availability, making it crucial to consider the entire distribution when making decisions.
-For instance, these forecasts enable management to calculate the probability of demand exceeding a certain threshold of available resources (e.g., 90\%), which can serve as an informative early warning measure for crowdedness.}
+For instance, these forecasts enable management to calculate the probability of demand exceeding a certain threshold of available resources (e.g., 90\%), which can serve as an informative early warning measure for overcrowding.
 
-\textcolor{teal}{It is important to note that while point forecasts and prediction intervals can be obtained from the probabilistic forecasts, the reverse is not possible.
+It is important to note that while point forecasts and prediction intervals can be obtained from the probabilistic forecasts, the reverse is not possible.
 A single number cannot be used to directly derive a probabilistic forecast.
-Prediction intervals, although helpful in indicating possible ranges, do not provide information on the probabilities of low or high demand.}
+Prediction intervals, although helpful in indicating possible ranges, do not provide information on the probabilities of low or high demand.
 
-\textcolor{teal}{In EMS planning, future demand is just one aspect to consider.
+In EMS planning, future demand is just one aspect to consider.
 Other inputs, such as capacity, should also be treated as probability distributions to adopt a probabilistic approach to planning.
-To extract valuable insights and make informed decisions from probabilistic forecasts, specialized numerical tools are required, as the forecasts themselves are typically represented as explicit probability density functions or Monte Carlo generators.}
+To extract valuable insights and make informed decisions from probabilistic forecasts, specialized numerical tools are required, as the forecasts themselves are typically represented as explicit probability density functions or Monte Carlo generators.
 
 ```{r}
-#| label: fig-forecast-density
-#| fig-width: 8
-#| fig-height: 5
-#| out.width: "100%"
-#| fig-cap: "An illustrative example of the forecast distribution of ambulance demand (i.e.
-total incidence attended) for the PO health board for 1 day ahead.
-The horizontal axis shows all possible outcomes that may occur, with their likelihood shown on the vertical axis.
-The dashed line indicates the mean of the distribution."
+#| label: fig-forecast-density-hstep
+#| out.width: "80%"
+#| fig-cap: "A graphical illustration of the forecast distribution of ambulance demand (i.e.
+total incidence attended) for the SB health board for a horizon of seven days.
+For each day, we display the point forecast (black point), the histogram, and 80% (thick line) and 90% (thin line) prediction intervals.
+It also shows a portion of a historical time series as well as its fitted values."
 #| fig-pos: "H"
@@
 
-incident_cv <- incident_gthf |> filter(lhb == "PO", is_aggregated(category), is_aggregated(nature))
+incident_cv <- incident_gthf |> filter(lhb == "SB", is_aggregated(category), is_aggregated(nature))
 fit <- incident_cv %>%
@@
   )
-set.seed(1)
-fcst <- fit %>% generate(h = 1, times = 5000, bootstrap = TRUE)
-p1 <- ggplot(data = fcst, aes(x = .sim, y = ..density..)) +
-  geom_histogram(binwidth = 1, fill = "#d3d3d3", color = "black") +
-  geom_vline(data = fcst, aes(xintercept = (mean(.sim))), color = "black", linetype = "dashed", linewidth = 1) +
-  scale_x_continuous(breaks = seq(25, 70, 5)) +
-  scale_y_continuous(labels = percent) +
-  ggthemes::theme_few() +
-  labs(y = "Density", x = "Demand", title = "Probabilistic forecast of ambulance demand for 1 day ahead")
+set.seed(2023)
+fcst <- fit %>% generate(h = 7, times = 5000, bootstrap = TRUE)
+
+fitted_ets <- fit |> augment()
+fcst_point <- fcst |> index_by(date) |> summarise(.mean=mean(.sim))
+p1 <- ggplot(data = fcst, mapping = aes(x = date, y = .sim))+
+  stat_histinterval(point_interval = "mean_qi",.width = c(0.80, 0.95),breaks=15)+
+  geom_line(data=fcst_point,aes(y=.mean, shape ="Point Forecast"))+
+  geom_line(aes(y = .fitted, shape ="Fitted"), data = filter_index(fitted_ets, "2019-07-17" ~ .))+
+  geom_point(aes(y = .fitted, shape ="Fitted"), data = filter_index(fitted_ets, "2019-07-17" ~ .))+
+  geom_line(aes(y = incident, shape ="Actual"),data = filter_index(incident_cv, "2019-07-17" ~ .))+
+  geom_point(aes(y = incident, shape ="Actual"),data = filter_index(incident_cv, "2019-07-17" ~ .))+
+  scale_shape_manual(name=NULL,
+                   breaks=c('Actual','Fitted','Point Forecast'),
+                   values=c('Actual'=17,'Fitted'=15, 'Point Forecast'=19))+
+  labs(x= "Date", y="Incidents")+
+  theme_few()+
+  theme(legend.position = "top")
 p1
 ```
+
+```{r}
+#| label: fig-forecast-density
+#| out.width: 80%
+#| fig-cap: "An illustrative example of the forecast distribution of ambulance demand (i.e.
+total incidence attended) for the SB health board for one day ahead.
+This corresponds to the first forecast distribution in Figure 5.
+The horizontal axis shows all possible outcomes that may occur, with their likelihood shown on the vertical axis.
+The point in the middle shows the point forecast.
+Two lines at the bottom of the distribution highlights 80% (thick line) and 90% (thin line) prediction intervals."
+#| fig-pos: "H"
+
+fcst1 <- fcst %>% filter_index("2019-08-01")
+
+p2 <- ggplot(data = fcst1, mapping = aes(x = .sim))+
+  stat_histinterval(point_interval = "mean_qi",.width = c(0.80, 0.95),breaks=40)+
+  scale_x_continuous(breaks = seq(60, 140, 5))+
+  scale_y_continuous(labels = percent)+
+  labs(y = "Density", x = "Incidents")+
+  theme_few()
+p2
+```

### 2023-12-16 manuscript/06-conclusion.qmd
second revision finalised

@@
 
-\textcolor{teal}{Our research establishes a strong basis for future investigations and practical implementation in EMS.
+Our research establishes a strong basis for future investigations and practical implementation in EMS.
 Leveraging the hierarchical and grouped structure of demand time series, EMS can use this advanced forecasting framework to generate coherent point and probabilistic forecasts, making the most of all available data at every level of the hierarchy.
@@
 This would involve developing and implementing simulation models capable of replicating decision-making processes based on the forecast inputs.
-These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.}
+These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.
 
-\textcolor{teal}{Future research can build upon this study in several ways.
+Future research can build upon this study in several ways.
 In future investigations, we aim to explore this avenue by incorporating operational information, simulating decision processes, and assessing the decision impact of this framework on utilities that are significant to the EMS.
 Linking forecasts with its utilities (e.g.
-response time, allocation of medical services, resource utilization, cost, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.} Aditionally, in our study, we employed Poisson regression models to generate count distributions of attended incidents for the base forecasts.
+response time, allocation of medical services, resource utilization, cost, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.
+Aditionally, in our study, we employed Poisson regression models to generate count distributions of attended incidents for the base forecasts.
 However, it is essential to note that the reconciled forecast distributions are not counts.
 This observation presents an interesting avenue for future research.
-\textcolor{teal}{Also, the dataset used in this study only includes information on attended incidents, it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
-It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.}
+Also, the dataset used in this study only includes information on attended incidents, it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
+It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.
 
-\textcolor{teal}{Although our study primarily focuses on Emergency Medical Services, it is essential to emphasize that the framework we propose has broad applicability across various service industries} [@ostrom2010moving].
-\textcolor{teal}{Our approach is particularly valuable in situations where time series data is structured hierarchically and/or grouped, a common characteristic found in many sectors.
+Although our study primarily focuses on Emergency Medical Services, it is essential to emphasize that the framework we propose has broad applicability across various service industries [@ostrom2010moving].
+Our approach is particularly valuable in situations where time series data is structured hierarchically and/or grouped, a common characteristic found in many sectors.
 This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
-For instance, in supply chains} [@shugan2000advance], \textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
+For instance, in supply chains [@shugan2000advance], demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
 Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency.
-In the financial industry} [@kimes1998strategic], \textcolor{teal}{where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
+In the financial industry [@kimes1998strategic], where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
 Similarly, in transportation, the framework supports forecasting transportation demand at various levels, optimizing route planning and resource allocation.
-Likewise, in the hospitality and tourism industry} [@dekimpe2016impact], \textcolor{teal}{it facilitates forecasting demand rates at state, regional, and department levels, enabling strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses.
+Likewise, in the hospitality and tourism industry [@dekimpe2016impact], it facilitates forecasting demand rates at state, regional, and department levels, enabling strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses.
 Additionally, in call centers, accurate call volume forecasting at different levels of the call center hierarchy or grouped structure is crucial for workforce management and resource allocation.
-Implementing our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, ensuring efficient staffing and optimal service levels to meet customer demands.}
+Implementing our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, ensuring efficient staffing and optimal service levels to meet customer demands.

### 2023-12-16 manuscript/main.qmd
second revision finalised

@@
 library(scales)
+library(ggdist)
 # Color blind color palettes

### 2023-12-16 manuscript/main.tex
second revision finalised

@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, enhanced, frame hidden]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
 
@@
 into groups such as the nature of incidents or demand priority.
-\textcolor{teal}{Forecasts produced at both higher and lower levels of hierarchies are necessary for effective decision-making in EMS.
-For example, control area EMS forecasts can inform strategic decisions about how to allocate limited resources to lower levels, such as health boards and stations/hospitals.
-At the lower levels, hospitals or ambulance stations could use such forecasts to plan for staffing and resource allocation, ambulance dispatching, staff-to-shift assignment, staff rescheduling based on the anticipated volume and priority and nature of incidents.
-Additionally, generating forecasts at lower levels could potentially improve the accuracy of the high-level forecasts, by providing more detailed information on the nature and priority of incidents.
-This could help to identify patterns in demand that may not be apparent at the higher level.
-Therefore, employing forecasting techniques that consider the hierarchical and/or grouped patterns of time series in EMS aligns naturally, offering the possibility to enhance forecast accuracy and facilitate coordination.}
+Forecasts produced at both higher and lower levels of hierarchies are
+necessary for effective decision-making in EMS. For example, control
+area EMS forecasts can inform strategic decisions about how to allocate
+limited resources to lower levels, such as health boards and
+stations/hospitals. At the lower levels, hospitals or ambulance stations
+could use such forecasts to plan for staffing and resource allocation,
+ambulance dispatching, staff-to-shift assignment, staff rescheduling
+based on the anticipated volume and priority and nature of incidents.
+Additionally, generating forecasts at lower levels could potentially
+improve the accuracy of the high-level forecasts, by providing more
+detailed information on the nature and priority of incidents. This could
+help to identify patterns in demand that may not be apparent at the
+higher level. Therefore, employing forecasting techniques that consider
+the hierarchical and/or grouped patterns of time series in EMS aligns
+naturally, offering the possibility to enhance forecast accuracy and
+facilitate coordination.
 
@@
 planning and decision making.
-\textcolor{teal}{For example, suppose the annual budget for the whole organization is to be allocated to each area health board based on demand within the corresponding area.
-If the forecasts are incoherent, there is a mismatch between the total resources to be allocated, and the resources allocated to each area.
-Further, suppose the budget also needs to take account of the nature of incidents that occur within each area, with more money allocated for some types of incidents than others.
-Then we need forecasts of the demand disaggregated by nature of incidents and health board, but these data are often too noisy to forecast directly.
-Forecast reconciliation addresses these problems by ensuring the forecasts are coherent across all aggregated and disaggregated series (avoiding mismatches), and by using the signal in the aggregated data to allow forecasting of highly disaggregated data (allowing better targetting of the budget).}
 
@@
 of incidents. We produce consistent point forecasts and forecast
-distributions for all levels, which is critical for an effective
-planning and associated risk management. We compare the point and
-probabilistic forecast accuracy of the independent forecasts, bottom-up
-and optimal reconciliation approaches. We first generate
+distributions for all levels over 84 days horizon, which is critical for
+an effective planning and associated risk management. We compare the
+point and probabilistic forecast accuracy of the independent forecasts,
+bottom-up and optimal reconciliation approaches. We first generate
 independent/base forecasts using Exponential Smoothing State Space
@@
 or social care) subject to the time series having a hierarchical and/or
-grouped structure.
+grouped structure. While our research focuses on emergency medical
+services, it is important to emphasize the suggested framework's
+adaptability, which expands its relevance to a variety of service
+sectors such as supply chains, tourism, finance, and call centers. Our
+approach can be generalized in cases with hierarchically structured
+and/or grouped time series data, which is common in many service
+sectors.
 
@@
 hierarchical and grouped relationships, preventing aligned decision
-making and coordination. \citet{al2021empirical}
-\textcolor{teal}{utilises data from the Welsh Ambulance Service to explore the forecast accuracy of four forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis (SSA) in predicting call volume demand.
-The aim is to compare these approaches with the current method across various planning horizons (7 days, 30 days, and 90 days) for both total call volume and category-specific demand.
-Forecast accuracy performance is evaluated using root mean square error (RMSE) and mean absolute percentage error (MAPE).
-The findings indicate that ARIMA performs the best in predicting weekly and monthly demand.
-However, when it comes to long-term demand, the SSA method proves to be the most effective.}
-\citet{ibrahim2016modeling}
-\textcolor{teal}{conducted a case study to assess the effectiveness of multiple forecasting methods: the multiplicative univariate forecasting model (MU), univariate mixed-effects model (ME), bivariate mixed-effects model (BME1), and bivariate mixed-effects model (BME2).
-Call centre data were utilised to forecast for periods of 1, 7, and 14 days ahead, using only a limited dataset of 42 days.
-The performance of these forecasting methods was evaluated using two metrics: RMSE for point forecasts and coverage probability for the 95\% prediction interval.
-The findings indicate that the ME consistently produces the most accurate point forecasts.
-On the other hand, BME1 and BME2 demonstrate superior coverage probabilities when forecasting for one day or one week ahead.
-For a two-week leading period, MU shows better coverage probability.}
+making and coordination. \citet{al2021empirical} utilises data from the
+Welsh Ambulance Service to explore the forecast accuracy of four
+forecasting approaches: ARIMA, Holt Winters, Multiple Regression, and
+Singular Spectrum Analysis (SSA) in predicting call volume demand. The
+aim is to compare these approaches with the current method across
+various planning horizons (7 days, 30 days, and 90 days) for both total
+call volume and category-specific demand. Forecast accuracy performance
+is evaluated using root mean square error (RMSE) and mean absolute
+percentage error (MAPE). The findings indicate that ARIMA performs the
+best in predicting weekly and monthly demand. However, when it comes to
+long-term demand, the SSA method proves to be the most effective.
+\citet{ibrahim2016modeling} conducted a case study to assess the
+effectiveness of multiple forecasting methods: the multiplicative
+univariate forecasting model (MU), univariate mixed-effects model (ME),
+and two variations of bivariate mixed-effects models (BME). Call centre
+data were utilized to forecast for periods of 1, 7, and 14 days ahead,
+using only a limited dataset of 42 days. The performance of these
+forecasting methods was evaluated using two metrics: RMSE for point
+forecasts and coverage probability for the 95\% prediction interval. The
+findings indicate that the ME consistently produces the most accurate
+point forecasts. On the other hand, BME models demonstrate superior
+coverage probabilities when forecasting for one day or one week ahead.
+For a two-week leading period, MU shows better coverage probability.
 
@@
 Current study & 2023 & Ambulance demand & 84 days & Stationary,  Exponential Smoothing State Space (ETS), Poisson regression using Generalized Linear Model (GLM) and tscount (TSGLM), a simple empirical distribution and an ensemble method & MASE, MSSE, CRPS & YES & YES & YES\\
-Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO & No\\
-Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO & No\\
-Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO & No\\
+Al-Azzani et al. & 2021 & Call volume & 7 , 30 ,  90 days & ARIMA, Holt Winters, Multiple Regression, and Singular Spectrum Analysis & RMSE, MAPE & NO & NO & NO\\
+Haugsbø et al. & 2021 & Ambulance demand in Spatio-temporal & 1hour & MLP, LSTM & MAE, MSE, Cross-Entropy & YES & NO & NO\\
+Grekousis et al. & 2019 & Locations of incidents & 1 hour & MLP and Hungarian algorithm & RMSE & NO & NO & NO\\
 Ibrahim et al.
-& 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO & No\\
-Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO & No\\
-Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO & No\\
+& 2016 & Call volume & 1, 7 , 14 days & multiplicative univariate forecasting, univariate mixed-effects, bivariate mixed-effects model, and bivariate mixed-effects & RMSE, prediction interval coverage & Partial & NO & NO\\
+Vile et al. & 2012 & Ambulance demand & 7, 14, 21,  28 day & Singular Spectrum Analysis, ARIMA,  Holt-Winters & RMSE & NO & NO & NO\\
+Sasaki et al. & 2010 & Ambulance demand & 5 years & OLS regression & R\textasciicircum{}2 & NO & NO & NO\\
 \bottomrule
@@
 incidents in the whole country are disaggregated into 3 control areas
-and then into 7 different healthboards.}
+and then into 7 different healthboards, anonymized using two letters
+(e.g.~AB)}
 \end{minipage}%
@@
 
-\textcolor{teal}{In this paper, the strength of trend and seasonality were calculated using the "STL" (Seasonal and Trend decomposition using Loess) decomposition method, as described by}
-\citet{mstl}.
-\textcolor{teal}{STL is a widely used and flexible method for decomposing time series data into trend, seasonal, and remainder components.
-The decomposition of a time series  $y_t$ is written as $y_t = T_t + S_{t} + R_t$, where $T_t$ is the smoothed trend component, $S_t$ is the seasonal component and $R_t$ is a remainder component.
-The strength of trend is defined as:}
+In this paper, the strength of trend and seasonality were calculated
+using the ``STL'' (Seasonal and Trend decomposition using Loess)
+decomposition method, as described by \citet{mstl}. STL is a widely used
+and flexible method for decomposing time series data into trend,
+seasonal, and remainder components. The decomposition of a time series
+\(y_t\) is written as \(y_t = T_t + S_{t} + R_t\), where \(T_t\) is the
+smoothed trend component, \(S_t\) is the seasonal component and \(R_t\)
+is a remainder component. The strength of trend is defined as:
 \[F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)\]
-\textcolor{teal}{For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component.
-Therefore Var($R_t$)/Var($T_t+R_t$) should be relatively small.
-But for data with little or no trend, the two variances should be approximately the same.}
-
-\textcolor{teal}{The strength of seasonality is defined similarly:}
+For strongly trended data, the seasonally adjusted data should have much
+more variation than the remainder component. Therefore
+Var(\(R_t\))/Var(\(T_t+R_t\)) should be relatively small. But for data
+with little or no trend, the two variances should be approximately the
+same.
+
+The strength of seasonality is defined similarly:
 \[F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right).\]
-\textcolor{teal}{series with seasonal strength  $F_S$, close to 0 exhibits almost no seasonality, while a series with strong seasonality will have $F_S$ close to 1 because Var($R_t$) will be much smaller than Var($S_t+R_t$).}
+series with seasonal strength \(F_S\), close to 0 exhibits almost no
+seasonality, while a series with strong seasonality will have \(F_S\)
+close to 1 because Var(\(R_t\)) will be much smaller than
+Var(\(S_t+R_t\)).
 
@@
 
-In addition to displaying the trend and seasonality features, we also
-visualize a few time series at various levels of aggregation.
-Figure~\ref{fig-dataviz2} reveals different information such as trend,
-seasonality, and noise. For example, some series depict seasonality and
-trend, whereas some other series report low volume of attended incidents
-and entropy, making them more volatile and difficult to forecast. At the
-level on nature of incidents combined with categories of other levels,
-there are many series that contain zeros with low counts. As such, the
-data set represents a diverse set of daily time series patterns.
+In addition to displaying the trend and seasonality strength
+\citep{hyndman2021forecasting}, we also visualize a few time series at
+various levels of aggregation. Figure~\ref{fig-dataviz2} reveals
+different information such as trend, seasonality, and noise. For
+example, some series depict seasonality and trend, whereas some other
+series report low volume of attended incidents and entropy, making them
+more volatile and difficult to forecast. At the level on nature of
+incidents combined with categories of other levels, there are many
+series that contain zeros with low counts. As such, the data set
+represents a diverse set of daily time series patterns.
 
@@
 empirical distribution of the past daily attended incidents to create
-the forecast distribution of future attended incidents.
-\textcolor{teal}{We have chosen this "stationary" method as a benchmark due to its widespread usage and simplicity, making it easily understandable for users.
-Forecasts serve as inputs for various decision-making systems that frequently employ simulations, wherein it is common to utilize the empirical distribution of demand as a forecast.
-Additionally, the stationary method has shown surprisingly high accuracy.
-Hence, any forecasting approach that can offer superior results compared to the stationary method would validate its practical use, otherwise there is no necessity for employing more complex methods.}
+the forecast distribution of future attended incidents. We have chosen
+this ``stationary'' method as a benchmark due to its widespread usage
+and simplicity, making it easily understandable for users. Forecasts
+serve as inputs for various decision-making systems that frequently
+employ simulations, wherein it is common to utilize the empirical
+distribution of demand as a forecast. Additionally, the stationary
+method has shown surprisingly high accuracy. Hence, any forecasting
+approach that can offer superior results compared to the stationary
+method would validate its practical use, otherwise there is no necessity
+for employing more complex methods.
 
@@
 
-\textcolor{teal}{In our study, we use an automated algorithm to determine the suitable configuration for the trend, seasonality, and error terms in each time series.
-Specifically, we utilize the `ETS()` function in the fable package of R, which employs Akaike's Information Criterion (AIC) to identify the optimal model for each time series.
-Given the large number of time series we work with (1530), it is impractical to manually select the appropriate form for each component in every time series.
-Consequently, the automated algorithm selects the best model based on the unique characteristics of each individual time series.
-As a result, a combination of additive or multiplicative forms for the components are employed, depending on the specific attributes of each time series.}
+In our study, we use an automated algorithm to determine the suitable
+configuration for the trend, seasonality, and error terms in each time
+series. Specifically, we utilize the \texttt{ets()} function in the
+forecast package of R, which employs Akaike's Information Criterion
+(AIC) to identify the optimal model for each time series. Given the
+large number of time series we work with (1530), it is impractical to
+manually select the appropriate form for each component in every time
+series. Consequently, the automated algorithm selects the best model
+based on the unique characteristics of each individual time series. As a
+result, a combination of additive or multiplicative forms for the
+components are employed, depending on the specific attributes of each
+time series.
 
@@
 
-\textcolor{teal}{Suppose the time series is denoted by $y_1,\dots,y_T$, then the Poisson GLM can be written as}
-\begin{align*}
+Suppose the time series is denoted by \(y_1,\dots,y_T\), then the
+Poisson GLM can be written as \begin{align*}
   y_t &\sim \text{Poisson}(\lambda_t) \\
@@
   \log(\lambda_t) &= \bm{x}_t'\bm{\beta},
-\end{align*}
-\textcolor{teal}{and $\bm{x}_t$ is a vector of covariates, $\beta$ is a vector of coefficients, and $\lambda_t$ is the mean of the Poisson distribution.
-In our model, these include cubic splines for the time trend, day-of-week dummy variables (from Monday to Sunday), Fourier terms to capture the yearly seasonality, dummy variables indicating public holidays (1 when is public holiday, 0 otherwise), school holidays (1 when is school holiday, 0 otherwise) and Christmas Day (1 when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New Year's Day, 0 otherwise).
-The Fourier terms are as defined in}
-\citet[Section 7.4]{hyndman2021forecasting}
-\textcolor{teal}{This model takes account of weekly seasonality and annual seasonality.
-Monthly seasonality is exceedingly rare in time series data, and does not occur in ambulance demand.
-There is no reason, for example, for incidents to occur more at some times of the month than others.}
+\end{align*} and \(\bm{x}_t\) is a vector of covariates, \(\beta\) is a
+vector of coefficients, and \(\lambda_t\) is the mean of the Poisson
+distribution. In our model, these include cubic splines for the time
+trend, day-of-week dummy variables (from Monday to Sunday), Fourier
+terms to capture the yearly seasonality, dummy variables indicating
+public holidays (1 when is a public holiday, 0 otherwise), school
+holidays (1 when is a school holiday, 0 otherwise), and Christmas Day (1
+when is Christmas Day, 0 otherwise) and New Year's Day (1 when is New
+Year's Day, 0 otherwise). The Fourier terms are as defined in
+\citet[Section 7.4]{hyndman2021forecasting}. This model takes account of
+weekly seasonality and annual seasonality. Monthly seasonality in time
+series data is extremely rare, and it does not exist in the ambulance
+demand used in this study. There is no reason for occurrences to occur
+more frequently at certain times of the month than others.
 
@@
 another Poisson regression model that takes into account serial
-dependence. This model captures the short range serial dependence by
-including autoregressive terms, in addition to the same covariates that
+dependence. This model captures the short-range serial dependence by
+including autoregressive terms in addition to the same covariates that
 were used in the GLM model. To distinguish this from the previous GLM
@@
 
-\textcolor{teal}{The Poisson TSGLM is similar to GLM with an additional autoregressive component accounting for serial correlation.
-The term serial dependence refers to situations where the ambulance demand at a given time point exhibits correlation with the demand at previous time points, after accounting for the exogenous covariates.
-Put simply, past values of the time series influence the current values, and various techniques can be employed to model this influence.
-TSGLM can be written as}
-\begin{align*}
+The Poisson TSGLM is similar to the GLM, with an additional
+autoregressive component accounting for serial dependence. The term
+serial dependence refers to instances in which the number of incidents
+on a current day correlates with the number of incidents on previous
+days. \begin{align*}
   y_t &\sim \text{Poisson}(\lambda_t) \\
@@
   \log(\lambda_t) &= (\bm{y}_{t-k}' , \bm{x}_t')\bm{\beta},
-\end{align*}
-\textcolor{teal}{and ${y}_{t-k}$ is a vector of $k$ lagged values.
-The TSGLM model explicitly accounts for serial dependence by including lagged values (i.e.
-past values) of the ambulance demand in the model.
-This is important in EMS forecasting because it allows the model to capture patterns in the data that are dependent on the past values of the time series, which might not be captured via the predictor variables.}
-
-\textcolor{teal}{We use the `tsglm()` function in the `tscount` package in R}
-\citep{JSSv082i05}
-\textcolor{teal}{to model the attended incidents. Again, the logarithmic link function is used to ensure that the mean of the Poisson distribution is always positive.}
-
-\textcolor{teal}{Provided accidents occur independently, they will inherently follow a Poisson distribution}
-\citep[p156--158]{feller1991introduction}.
-\textcolor{teal}{Hence, it is reasonable to assume a Poisson distribution in this context.
-To account for changes over time, we incorporate trend and seasonality covariates, as well as public holiday effects, allowing the mean of the Poisson distribution to vary.
-However, it is important to note that if there are additional factors influencing the mean of the Poisson distribution, which are not accounted for in our model, we might observe over-dispersion or under-dispersion in the data.}
-
-\textbf{Ensemble method:}
-\textcolor{teal}{Finally, one effective strategy for improving forecast accuracy includes the simultaneous application of multiple forecasting methods on a given time series, followed by combining the forecasts, rather than relying on separate forecasts generated by each individual method}
-\citep{clemen1989combining}.
-\textcolor{teal}{In this paper, we use an ensemble method that combines the forecasts generated from the Stationary, ETS, GLM and TSGLM models using a simple average to form a mixture distribution}
-\citep{combinations}.
-
-\textcolor{teal}{To generate forecast probability distributions using above methods, we use a form of bootstrapping, described in}
-\citet{panagiotelis2023probabilistic}.
-\textcolor{teal}{This involves simulating 1000 future sample paths from each of the models, by bootstrapping the model residuals, taking into account the cross-sectional correlations between the different aggregated and disaggregated series.
-In this way, we can generate an empirical distribution of forecasts for each model.
-The ensemble forecast distribution is a simple mixture of these empirical distributions.}
-
-\textcolor{teal}{It is important to emphasize that the aim of this study is not to provide an exhaustive compilation of forecasting models, or to promote a particular model class.
-Instead, we have developed a flexible framework that can accommodate any forecasting models.
-Our primary objective is to demonstrate its practicality and effectiveness in integrating base forecasts from any model and generating coherent forecasts within a hierarchical structure.}
+\end{align*} and \({y}_{t-k}\) is a vector of \(k\) lagged values. The
+TSGLM model explicitly accounts for serial dependence by including
+lagged values (i.e., past values) of the ambulance demand in the model.
+This is important in EMS forecasting because it allows the model to
+capture patterns in the data that are dependent on the past values of
+the time series, which might not be captured via the predictor
+variables.
+
+We use the \texttt{tsglm()} function in the \texttt{tscount} package in
+R \citep{JSSv082i05} to model the attended incidents. Again, the
+logarithmic link function is used to ensure that the mean of the Poisson
+distribution is always positive.
+
+Provided accidents occur independently, they will inherently follow a
+Poisson distribution \citep[p156--158]{feller1991introduction}. Hence,
+it is reasonable to assume a Poisson distribution in this context. To
+account for changes over time, we incorporate trend and seasonality
+covariates, as well as public holiday effects, allowing the mean of the
+Poisson distribution to vary. However, it is important to note that if
+there are additional factors influencing the mean of the Poisson
+distribution that are not accounted for in our model, we might observe
+over- or under-dispersion in the data.
+
+\textbf{Ensemble method:} Finally, one effective strategy for improving
+forecast accuracy includes the simultaneous application of multiple
+forecasting methods on a given time series, followed by combining the
+forecasts rather than relying on separate forecasts generated by each
+individual method \citep{clemen1989combining}. In this paper, we use an
+ensemble method that combines the forecasts generated from the
+Stationary, ETS, GLM, and TSGLM models using a simple average to form a
+mixture distribution \citep{combinations}.
+
+To generate forecast probability distributions using the above methods,
+we use a form of bootstrapping, described in
+\citet{panagiotelis2023probabilistic}. This involves simulating 1000
+future sample paths from each of the models by bootstrapping the model
+residuals, taking into account the cross-sectional correlations between
+the different aggregated and disaggregated series. In this way, we can
+generate an empirical distribution of forecasts for each model. The
+ensemble forecast distribution is a simple mixture of these empirical
+distributions.
+
+It is important to emphasize that the aim of this study is not to
+provide an exhaustive compilation of forecasting models or to promote a
+particular model class. Instead, we have developed a flexible framework
+that can accommodate any forecasting model. Our primary objective is to
+demonstrate its practicality and effectiveness in integrating base
+forecasts from any model and generating coherent forecasts within a
+hierarchical structure.
 
@@
 
-\textcolor{teal}{The error metrics provided below consider a forecasting horizon denoted by $j$, representing the number of time periods ahead we are predicting.
-In our study, this forecasting horizon ranges from 1 to 84 days, $j= 1,2,\dots, 84$.}
+The error metrics provided below consider a forecasting horizon denoted
+by \(j\), representing the number of time periods ahead we are
+predicting. In our study, this forecasting horizon ranges from 1 to 84
+days, \(j= 1,2,\dots, 84\).
 
@@
 
-\textcolor{teal}{Using scale-independent measures, such as MASE and MSSE, enables more appropriate comparisons between time series at different levels and scales, as these measures are not influenced by the magnitude of the data.
-This is of particular importance in our study, as we work with time series at various levels of hierarchy, with varying scales, resulting in different magnitudes of error.
-By employing scale-independent measures, we can meaningfully assess the forecast accuracy across the entire hierarchy, ensuring a more robust comparison.}
+Using scale-independent measures, such as MASE and MSSE, enables more
+appropriate comparisons between time series at different levels and
+scales, as these measures are not influenced by the magnitude of the
+data. This is of particular importance in our study, as we work with
+time series at various levels of hierarchy, with varying scales,
+resulting in different magnitudes of error. By employing
+scale-independent measures, we can meaningfully assess the forecast
+accuracy across the entire hierarchy, ensuring a more robust comparison.
 
@@
 
-\textcolor{teal}{Calibration refers to the statistical consistency between the distributional forecasts and the observations.
-It measures how well the predicted probabilities match the observations.
-On the other hand, sharpness refers to the concentration of the forecast distributions --- a sharp forecast distribution results in narrow prediction intervals, indicating high confidence in the forecast.
-A model is well-calibrated if the predicted probabilities match the distribution of the observations, and it is sharp if it is confident in its predictions.
-The CRPS rewards sharpness and calibration by assigning lower scores to forecasts with sharper distributions, and to forecasts that are well-calibrated.
-Thus, it is a metric that combines both sharpness and miscalibration into a single score, making it a useful tool for evaluating the performance of probabilistic forecasts.}
-
-\textcolor{teal}{CRPS can be considered an average of all possible Winkler scores}
-\citep[Section 5.9]{winkler1972decision, hyndman2021forecasting}
-\textcolor{teal}{or percentile scores} \citep[Section
-5.9]{hyndman2021forecasting},
-\textcolor{teal}{and thus provides an evaluation of all possible prediction intervals or quantiles.}
-\textcolor{teal}{A specific prediction interval could be evaluated using a Winkler score.
-Certain situations may also require assessing accuracy for a particular quantile, such as lower (e.g 5\%) or higher (e.g.
-95\%) quantiles.
-In such cases, a percentile score becomes useful in meeting this specific requirement.}
+Calibration refers to the statistical consistency between the
+distributional forecasts and the observations. It measures how well the
+predicted probabilities match the observations. On the other hand,
+sharpness refers to the concentration of the forecast distributions ---
+a sharp forecast distribution results in narrow prediction intervals,
+indicating high confidence in the forecast. A model is well-calibrated
+if the predicted probabilities match the distribution of the
+observations, and it is sharp if it is confident in its predictions. The
+CRPS rewards sharpness and calibration by assigning lower scores to
+forecasts with sharper distributions, and to forecasts that are
+well-calibrated. Thus, it is a metric that combines both sharpness and
+miscalibration into a single score, making it a useful tool for
+evaluating the performance of probabilistic forecasts.
+
+CRPS can be considered an average of all possible Winkler scores
+\citep[Section 5.9]{winkler1972decision, hyndman2021forecasting} or
+percentile scores \citep[Section 5.9]{hyndman2021forecasting}, and thus
+provides an evaluation of all possible prediction intervals or
+quantiles. A specific prediction interval could be evaluated using a
+Winkler score. Certain situations may also require assessing accuracy
+for a particular quantile, such as lower (e.g 5\%) or higher (e.g.~95\%)
+quantiles. In such cases, a percentile score becomes useful in meeting
+this specific requirement.
 
@@
 specifying how the bottom-level series \(\bm{b}_t\) are to be aggregated
-to form \(\bm{a}_t\).
-\textcolor{teal}{The aggregation matrix $\bm{A}$ is determined by the structure of the hierarchy.
-It maps the bottom-level time series to the corresponding higher-level time series.
-For example, if there are two bottom-level series, and one aggregated series (equal to the sum of the two bottom-level series), then $\bm{A} = \begin{bmatrix} 1 ~~ 1 \end{bmatrix}$.
-} The full vector of time series is given by \[
+to form \(\bm{a}_t\). The aggregation matrix \(\bm{A}\) is determined by
+the structure of the hierarchy. It maps the bottom-level time series to
+the corresponding higher-level time series. For example, if there are
+two bottom-level series, and one aggregated series (equal to the sum of
+the two bottom-level series), then
+\(\bm{A} = \begin{bmatrix} 1 ~~ 1 \end{bmatrix}\).
+
+The full vector of time series is given by \[
  \bm{y}_t = \begin{bmatrix}\bm{a}_t \\\bm{b}_t\end{bmatrix}.
@@
 
-\textcolor{teal}{The term "bottom-level series" relates to the most disaggregated series within the hierarchical and grouped time series structure.
-For instance, in Table 2, each distinct combination of values in Control area (e.g.
-South \& East), Health board (e.g.
-CV), Priority (e.g.
-Green), and Nature of incident (e.g.
-Chest pain), corresponds to one individual time series.
-In the dataset at hand, there are 691 unique combinations, resulting in 691 bottom level time series.
-The "aggregate time series" describes how these bottom-level series are combined to create higher-level series.
-For instance, to obtain the incidents at the national level (i.e.
-all country level), the time series are aggregated across all Control areas, Health boards, Priorities, and Natures of incidents.
-Any desired aggregation level can be achieved based on the data structure, utilizing the bottom-level series available.}
+The term ``bottom-level series'' relates to the most disaggregated
+series within the hierarchical and grouped time series structure. For
+instance, in Table 2, each distinct combination of values in Control
+area (e.g.~South \& East), Health board (e.g.~CV), Priority
+(e.g.~Green), and Nature of incident (e.g.~Chest pain), corresponds to
+one individual time series. In the dataset at hand, there are 691 unique
+combinations, resulting in 691 bottom level time series. The ``aggregate
+time series'' describes how these bottom-level series are combined to
+create higher-level series. For instance, to obtain the incidents at the
+national level (i.e.~all country level), the time series are aggregated
+across all Control areas, Health boards, Priorities, and Natures of
+incidents. Any desired aggregation level can be achieved based on the
+data structure, utilizing the bottom-level series available.
 
@@
 
-\textcolor{teal}{Forecast reconciliation approaches bridge this gap by combining and reconciling all base forecasts to generate coherent forecasts.
-This technique utilizes all the base forecasts produced within a hierarchical structure to create consistent forecasts at every level of the hierarchy.
-As a result, it goes beyond relying solely on base forecasts from a single level of aggregation, and instead leverages all available information at each level to generate forecasts that minimize the total forecast variance of the set of coherent forecasts.}
-\textcolor{teal}{Linear reconciliation involves projecting the base forecasts onto the coherent space.
-It is derived by minimizing the sum of the variances of the reconciled forecasts subject to the resulting forecasts being coherent and unbiased}
-\citep{WicEtAl2019}.
+Forecast reconciliation approaches bridge this gap by combining and
+reconciling all base forecasts to generate coherent forecasts. This
+technique utilizes all the base forecasts produced within a hierarchical
+structure to create consistent forecasts at every level of the
+hierarchy. As a result, it goes beyond relying solely on base forecasts
+from a single level of aggregation, and instead leverages all available
+information at each level to generate forecasts that minimize the total
+forecast variance of the set of coherent forecasts. Linear
+reconciliation involves projecting the base forecasts onto the coherent
+space. It is derived by minimizing the sum of the variances of the
+reconciled forecasts subject to the resulting forecasts being coherent
+and unbiased \citep{WicEtAl2019}.
 
@@
 
-\textcolor{teal}{Ordinary Least Squares (OLS) is the simplest and most commonly used method.
-In this approach, the estimation of $\bm{W}$ is based on the assumption that all the errors are uncorrelated and have equal variance, and that multi-step forecast variances are proportion to one-step forecast variances.
-Then, $\bm{W}$ is simply the identity matrix multiplied by a constant factor.
-The main weakness of this approach is that it does not take account of the different scales of the base time series; the aggregated series will usually have higher variance than the disaggregated series, simply because the values are larger, but OLS treats all series the same.
-A strength of the approach is that it is simple, and does not involve estimating a covariance matrix.}
-
-\textcolor{teal}{Weighted Least Squares (WLS) is an extension of OLS where the variance of the errors is assumed to be heteroscedastic, i.e., different for each series.
-But it assumes that the errors of each series are uncorrelated with each other, and that multi-step forecast variances are proportion to one-step forecast variances.
-In this approach, $\bm{W}$ is defined as a diagonal matrix with the variance of the errors on the diagonal.
-The intuition behind WLS is that it assigns higher weight to series with smaller error variance, and thereby takes into account the different scales of the base time series.
-The main weakness of this approach is that it ignores the relationships between series.
-A strength of WLS is that it is relatively easy to compute $\bm{W}$ as it is based only on error variances which are readily estimated.}
-
-\textcolor{teal}{Minimum Trace (MinT) is a further generalization where $\bm{W}$ is defined as the covariance matrix of the one-step base forecast errors.
-So it takes account of both the scale of each series, and the relationships between the series.}
-But it still assumes that multi-step forecast variances are proportion
-to one-step forecast variances. The main weakness of this approach is
-that it is difficult to estimate the full covariance matrix, even of the
-one-step errors. In practice, we usually need to use a shrinkage
-estimate where the off-diagonal elements are shrunk towards zero.\}
+Ordinary Least Squares (OLS) is the simplest and most commonly used
+method. In this approach, the estimation of \(\bm{W}\) is based on the
+assumption that all the errors are uncorrelated and have equal variance,
+and that multi-step forecast variances are proportion to one-step
+forecast variances. Then, \(\bm{W}\) is simply the identity matrix
+multiplied by a constant factor. The main weakness of this approach is
+that it does not take account of the different scales of the base time
+series; the aggregated series will usually have higher variance than the
+disaggregated series, simply because the values are larger, but OLS
+treats all series the same. A strength of the approach is that it is
+simple, and does not involve estimating a covariance matrix.
+
+Weighted Least Squares (WLS) is an extension of OLS where the variance
+of the errors is assumed to be heteroscedastic, i.e., different for each
+series. But it assumes that the errors of each series are uncorrelated
+with each other, and that multi-step forecast variances are proportion
+to one-step forecast variances. In this approach, \(\bm{W}\) is defined
+as a diagonal matrix with the variance of the errors on the diagonal.
+The intuition behind WLS is that it assigns higher weight to series with
+smaller error variance, and thereby takes into account the different
+scales of the base time series. The main weakness of this approach is
+that it ignores the relationships between series. A strength of WLS is
+that it is relatively easy to compute \(\bm{W}\) as it is based only on
+error variances which are readily estimated.
+
+Minimum Trace (MinT) is a further generalization where \(\bm{W}\) is
+defined as the covariance matrix of the one-step base forecast errors.
+So it takes account of both the scale of each series, and the
+relationships between the series. But it still assumes that multi-step
+forecast variances are proportion to one-step forecast variances. The
+main weakness of this approach is that it is difficult to estimate the
+full covariance matrix, even of the one-step errors. In practice, we
+usually need to use a shrinkage estimate where the off-diagonal elements
+are shrunk towards zero.
 
@@
 
-\textcolor{teal}{Certainly, other approaches can be applied to hierarchical forecasting problems}
-\citet{pennings2017integrated} and \citet{villegas2018supply}
-\textcolor{teal}{proposed the idea of using a state space model to ensure consistent forecasts.
-However, when dealing with larger hierarchies, these models encounter difficulties in estimating covariance matrices.
-In contrast, our approach provides a clear advantage by allowing the incorporation of different forecasting methods for the base forecasts, and even accommodating distinct methods for individual series.
-The decoupling of time series models from the reconciliation step adds significant flexibility in exploring a wide range of models.}
+Certainly, other approaches can be applied to hierarchical forecasting
+problems \citet{pennings2017integrated} and \citet{villegas2018supply}
+proposed the idea of using a state space model to ensure consistent
+forecasts. However, when dealing with larger hierarchies, these models
+encounter difficulties in estimating covariance matrices. In contrast,
+our approach provides a clear advantage by allowing the incorporation of
+different forecasting methods for the base forecasts, and even
+accommodating distinct methods for individual series. The decoupling of
+time series models from the reconciliation step adds significant
+flexibility in exploring a wide range of models.
 
@@
 remain valid for bottom series where different individual methods
-perform best, depending on the accuracy measure.
-\textcolor{teal}{While the forecast reconciliation approach aims to enhance forecast accuracy, its effectiveness is not guaranteed, especially if the bottom-level series exhibit excessive noise and lack systematic patterns.
-Despite this, reconciling forecasts at the bottom level can offer advantages by generating coherent forecasts that facilitate alignment in planning across various teams within an organization, promote better coordination, and prevent conflicting decisions.
-Moreover, even when dealing with noisy and irregular bottom-level series, reconciliation can still improve forecast accuracy at higher levels of the hierarchy by leveraging the information available across the hierarchy.
-Therefore, although the bottom-level forecasts may not be highly accurate on their own, reconciling them with higher-level forecasts can still provide a more consistent view of future demand and potentially yield more accurate forecasts at other levels.}
+perform best, depending on the accuracy measure. While the forecast
+reconciliation approach aims to enhance forecast accuracy, its
+effectiveness is not guaranteed, especially if the bottom-level series
+exhibit excessive noise and lack systematic patterns. Despite this,
+reconciling forecasts at the bottom level can offer advantages by
+generating coherent forecasts that facilitate alignment in planning
+across various teams within an organization, promote better
+coordination, and prevent conflicting decisions. Moreover, even when
+dealing with noisy and irregular bottom-level series, reconciliation can
+still improve forecast accuracy at higher levels of the hierarchy by
+leveraging the information available across the hierarchy. Therefore,
+although the bottom-level forecasts may not be highly accurate on their
+own, reconciling them with higher-level forecasts can still provide a
+more consistent view of future demand and potentially yield more
+accurate forecasts at other levels.
 
@@
 
-\textcolor{teal}{Despite using Poisson regression models to create count distributions of attended incidents for the base forecasts, it is important to note that the reconciled forecast distributions do not maintain a count format.
-In practical scenarios, there might be a need to use integer forecasts.
-Count forecast reconciliation is an active area of research, and it would be interesting to explore how our approach could be adapted to generate count-reconciled probabilistic forecasts in future studies.
-One possible approach to address this is by rounding the forecasts.
-However, the impact of rounding on forecast accuracy varies depending on the level of hierarchy and the scale of the data.
-In situations with high-volume demand, the effects of rounding may be negligible, and forecast accuracy calculations can overlook integer effects.
-On the other hand, in low-volume demand settings, such as forecasts at the bottom level of the hierarchy, integer (rounding) effects may have a more noticeable influence on forecast accuracy.}
+Despite using Poisson regression models to create count distributions of
+attended incidents for the base forecasts, it is important to note that
+the reconciled forecast distributions do not maintain a count format. In
+practical scenarios, there might be a need to use integer forecasts.
+Count forecast reconciliation is an active area of research, and it
+would be interesting to explore how our approach could be adapted to
+generate count-reconciled probabilistic forecasts in future studies.
+Rounding the forecasts is one possible solution to this problem.
+However, the impact of rounding on forecast accuracy varies depending on
+the level of hierarchy and the scale of the data. In situations with
+high volume demand, the effects of rounding may be negligible, and
+forecast accuracy calculations can overlook integer effects. On the
+other hand, in low-volume demand settings, such as forecasts at the
+bottom level of the hierarchy, integer (rounding) effects may have a
+more noticeable influence on forecast accuracy.
 
@@
 
-Figure~\ref{fig-forecast-density}
-\textcolor{teal}{provides an illustrative example of a probabilistic forecast for future demand, one day ahead, of the total attended incidents in the PO health board.
-The black dashed line in the graph represents the point forecast (the mean of the distribution), which presents a single numerical estimate for the future number of attended incidents.
-Due to the complexity of including such plots for the entire hierarchy and 84 days ahead, only one example is presented here for 1 day ahead.
-However, it is feasible to generate these plots for the entire hierarchy and for any forecast horizon if necessary.}
-
-\textcolor{teal}{In practice, point forecasts are commonly used, but they have limitations as they ignore the uncertainty associated with the forecast, leading to wrong decisions.
-In contrast, probabilistic forecasts offer an alternative approach to anticipate future demand.
-Rather than providing a single value, they assign likelihoods to all possible demand outcomes, acknowledging that different numbers of attended incidents are possible, but with varying likelihoods.}
-
-\textcolor{teal}{The purpose of probabilistic forecasting, as demonstrated in}
-Figure~\ref{fig-forecast-density},
-\textcolor{teal}{is to quantify uncertainty.
-Decisions based on these forecasts could focus on the tails of the distribution: unexpectedly high demand leading to crowding and inefficiency, or unexpectedly low demand resulting in wasted resources.
-Such forecasts are valuable tools for decision-makers and planners, especially when dealing with low-probability, high-cost situations.
-Different EMS managements may have varying risk attitudes depending on resource availability, making it crucial to consider the entire distribution when making decisions.
-For instance, these forecasts enable management to calculate the probability of demand exceeding a certain threshold of available resources (e.g., 90\%), which can serve as an informative early warning measure for crowdedness.}
-
-\textcolor{teal}{It is important to note that while point forecasts and prediction intervals can be obtained from the probabilistic forecasts, the reverse is not possible.
-A single number cannot be used to directly derive a probabilistic forecast.
-Prediction intervals, although helpful in indicating possible ranges, do not provide information on the probabilities of low or high demand.}
-
-\textcolor{teal}{In EMS planning, future demand is just one aspect to consider.
-Other inputs, such as capacity, should also be treated as probability distributions to adopt a probabilistic approach to planning.
-To extract valuable insights and make informed decisions from probabilistic forecasts, specialized numerical tools are required, as the forecasts themselves are typically represented as explicit probability density functions or Monte Carlo generators.}
+In this section, we provide an illustrative example of a probabilistic
+forecast for future demand, of the total attended incidents in the SB
+health board. Due to the complexity of including such plots in the
+manuscript for the entire hierarchy and 84 days ahead, only one example
+is presented here. However, it is feasible to generate these plots for
+the entire hierarchy and for any forecast horizon if necessary.\}
+
+In practice, point forecasts are commonly used, but they have
+limitations as they ignore the uncertainty associated with the forecast.
+The future is inherently characterized by an irreducible level of
+uncertainty. Being prepared entails considering alternate courses of
+action. Probabilistic forecasts offer an alternative approach to
+anticipate future demand. Rather than providing a single value, they
+assign likelihoods to all possible demand outcomes, acknowledging that
+different numbers of attended incidents are possible, but with varying
+likelihoods.
+
+The purpose of probabilistic forecasting, as demonstrated in
+Figure~\ref{fig-forecast-density-hstep} and
+Figure~\ref{fig-forecast-density}, is to quantify uncertainty.
+Figure~\ref{fig-forecast-density-hstep} depicts the forecast
+distribution of total incidents in one health board over a 7-day period.
+It also gives the point forecast as well as the 80\% and 90\% prediction
+intervals. Figure~\ref{fig-forecast-density} zooms in on the first day
+to show the histogram more clearly, illustrating the range of possible
+outcomes and their likelihood.
+
+Decisions based on these forecasts could focus on the tails of the
+distribution: unexpectedly high demand leading to crowding and
+inefficiency, or unexpectedly low demand resulting in wasted resources.
+Such forecasts are valuable tools for decision-makers and planners,
+especially when dealing with low-probability, high-cost situations.
+Different EMS managements may have varying risk attitudes depending on
+resource availability, making it crucial to consider the entire
+distribution when making decisions. For instance, these forecasts enable
+management to calculate the probability of demand exceeding a certain
+threshold of available resources (e.g., 90\%), which can serve as an
+informative early warning measure for overcrowding.
+
+It is important to note that while point forecasts and prediction
+intervals can be obtained from the probabilistic forecasts, the reverse
+is not possible. A single number cannot be used to directly derive a
+probabilistic forecast. Prediction intervals, although helpful in
+indicating possible ranges, do not provide information on the
+probabilities of low or high demand.
+
+In EMS planning, future demand is just one aspect to consider. Other
+inputs, such as capacity, should also be treated as probability
+distributions to adopt a probabilistic approach to planning. To extract
+valuable insights and make informed decisions from probabilistic
+forecasts, specialized numerical tools are required, as the forecasts
+themselves are typically represented as explicit probability density
+functions or Monte Carlo generators.
 
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-hstep-1.pdf}
+
+}
+
+\caption{\label{fig-forecast-density-hstep}A graphical illustration of
+the forecast distribution of ambulance demand (i.e.~total incidence
+attended) for the SB health board for a horizon of seven days. For each
+day, we display the point forecast (black point), the histogram, and
+80\% (thick line) and 90\% (thin line) prediction intervals. It also
+shows a portion of a historical time series as well as its fitted
+values.}
+
+\end{figure}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{main_files/figure-pdf/fig-forecast-density-1.pdf}
 
@@
 forecast distribution of ambulance demand (i.e.~total incidence
-attended) for the PO health board for 1 day ahead. The horizontal axis
-shows all possible outcomes that may occur, with their likelihood shown
-on the vertical axis. The dashed line indicates the mean of the
-distribution.}
+attended) for the SB health board for one day ahead. This corresponds to
+the first forecast distribution in Figure 5. The horizontal axis shows
+all possible outcomes that may occur, with their likelihood shown on the
+vertical axis. The point in the middle shows the point forecast. Two
+lines at the bottom of the distribution highlights 80\% (thick line) and
+90\% (thin line) prediction intervals.}
 
@@
 
-\textcolor{teal}{Our research establishes a strong basis for future investigations and practical implementation in EMS.
-Leveraging the hierarchical and grouped structure of demand time series, EMS can use this advanced forecasting framework to generate coherent point and probabilistic forecasts, making the most of all available data at every level of the hierarchy.
-We acknowledge that a forecast serves a greater purpose beyond its mere existence, ideally enabling the best utility in terms of efficient allocation of medical services, response time, and cost, all informed by the forecast.
-While we fully appreciate the importance of evaluating forecast quality based on its impact on decision-making processes, it is essential to address the data requirements and methodology involved in measuring this impact.
-For a comprehensive assessment of the forecasts' implications, access to additional data beyond ambulance demand, covering various decision types, capacity information, constraints in the decision system, and more, becomes necessary.
-This additional data would offer valuable insights into the specific decisions relying on the forecasts, resulting in a more accurate evaluation of their impact on medical services.
-Furthermore, measuring the actual impact of forecasts would necessitate an approach that goes beyond forecasting itself.
-This would involve developing and implementing simulation models capable of replicating decision-making processes based on the forecast inputs.
-These simulation models would then evaluate the quality of the final decisions, taking into consideration the utilities that are particularly significant in the context of EMS.}
-
-\textcolor{teal}{Future research can build upon this study in several ways.
-In future investigations, we aim to explore this avenue by incorporating operational information, simulating decision processes, and assessing the decision impact of this framework on utilities that are significant to the EMS.
-Linking forecasts with its utilities (e.g.
-response time, allocation of medical services, resource utilization, cost, etc) can offer an opportunity to maximize benefits through a more holistic planning approach.}
-Aditionally, in our study, we employed Poisson regression models to
-generate count distributions of attended incidents for the base
-forecasts. However, it is essential to note that the reconciled forecast
-distributions are not counts. This observation presents an interesting
-avenue for future research.
-\textcolor{teal}{Also, the dataset used in this study only includes information on attended incidents, it would be valuable for future research to investigate the impact of failed responses on EMS forecasting, if data on these incidents becomes available.
-It is also important to note that our methodology for hierarchical time series forecasting can be applied to any time series data in EMS, including those that may include failed responses.}
-
-\textcolor{teal}{Although our study primarily focuses on Emergency Medical Services, it is essential to emphasize that the framework we propose has broad applicability across various service industries}
-\citep{ostrom2010moving}.
-\textcolor{teal}{Our approach is particularly valuable in situations where time series data is structured hierarchically and/or grouped, a common characteristic found in many sectors.
-This occurs when data can be naturally organized into different levels of hierarchies or when dependencies and relationships exist among entities within the system.
-For instance, in supply chains}
-\citep{shugan2000advance},
-\textcolor{teal}{demand forecasting at different levels of the distribution network, such as regional warehouses or retail stores, is vital for efficient inventory management and minimizing stockouts.
-Our framework allows the reconciliation of forecasts, ensuring consistency and alignment throughout the supply chain, leading to improved decision-making and operational efficiency.
-In the financial industry}
-\citep{kimes1998strategic},
-\textcolor{teal}{where investments span multiple asset classes, geographical regions, or customer segments, our framework can be applied to forecast portfolio performance, asset allocation, or customer demand.
-Similarly, in transportation, the framework supports forecasting transportation demand at various levels, optimizing route planning and resource allocation.
-Likewise, in the hospitality and tourism industry}
-\citep{dekimpe2016impact},
-\textcolor{teal}{it facilitates forecasting demand rates at state, regional, and department levels, enabling strategic pricing, capacity planning, and revenue management for hotels and other travel-related businesses.
-Additionally, in call centers, accurate call volume forecasting at different levels of the call center hierarchy or grouped structure is crucial for workforce management and resource allocation.
-Implementing our framework, call centers can generate accurate forecasts for different skill groups, shifts, and locations, ensuring efficient staffing and optimal service levels to meet customer demands.}
+Our research establishes a strong basis for future investigations and
+practical implementation in EMS. Leveraging the hierarchical and grouped
+structure of demand time series, EMS can use this advanced forecasting
+framework to generate coherent point and probabilistic forecasts, making
+the most of all available data at every level of the hierarchy. We
+acknowledge that a forecast serves a greater purpose beyond its mere
+existence, ideally enabling the best utility in terms of efficient
+allocation of medical services, response time, and cost, all informed by
+the forecast. While we fully appreciate the importance of evaluating
+forecast quality based on its impact on decision-making processes, it is
+essential to address the data requirements and methodology involved in
+measuring this impact. For a comprehensive assessment of the forecasts'
+implications, access to additional data beyond ambulance demand,
+covering various decision types, capacity information, constraints in
+the decision system, and more, becomes necessary. This additional data
+would offer valuable insights into the specific decisions relying on the
+forecasts, resulting in a more accurate evaluation of their impact on
+medical services. Furthermore, measuring the actual impact of forecasts
+would necessitate an approach that goes beyond forecasting itself. This
+would involve developing and implementing simulation models capable of
+replicating decision-making processes based on the forecast inputs.
+These simulation models would then evaluate the quality of the final
+decisions, taking into consideration the utilities that are particularly
+significant in the context of EMS.
+
+Future research can build upon this study in several ways. In future
+investigations, we aim to explore this avenue by incorporating
+operational information, simulating decision processes, and assessing
+the decision impact of this framework on utilities that are significant
+to the EMS. Linking forecasts with its utilities (e.g.~response time,
+allocation of medical services, resource utilization, cost, etc) can
+offer an opportunity to maximize benefits through a more holistic
+planning approach. Aditionally, in our study, we employed Poisson
+regression models to generate count distributions of attended incidents
+for the base forecasts. However, it is essential to note that the
+reconciled forecast distributions are not counts. This observation
+presents an interesting avenue for future research. Also, the dataset
+used in this study only includes information on attended incidents, it
+would be valuable for future research to investigate the impact of
+failed responses on EMS forecasting, if data on these incidents becomes
+available. It is also important to note that our methodology for
+hierarchical time series forecasting can be applied to any time series
+data in EMS, including those that may include failed responses.
+
+Although our study primarily focuses on Emergency Medical Services, it
+is essential to emphasize that the framework we propose has broad
+applicability across various service industries
+\citep{ostrom2010moving}. Our approach is particularly valuable in
+situations where time series data is structured hierarchically and/or
+grouped, a common characteristic found in many sectors. This occurs when
+data can be naturally organized into different levels of hierarchies or
+when dependencies and relationships exist among entities within the
+system. For instance, in supply chains \citep{shugan2000advance}, demand
+forecasting at different levels of the distribution network, such as
+regional warehouses or retail stores, is vital for efficient inventory
+management and minimizing stockouts. Our framework allows the
+reconciliation of forecasts, ensuring consistency and alignment
+throughout the supply chain, leading to improved decision-making and
+operational efficiency. In the financial industry
+\citep{kimes1998strategic}, where investments span multiple asset
+classes, geographical regions, or customer segments, our framework can
+be applied to forecast portfolio performance, asset allocation, or
+customer demand. Similarly, in transportation, the framework supports
+forecasting transportation demand at various levels, optimizing route
+planning and resource allocation. Likewise, in the hospitality and
+tourism industry \citep{dekimpe2016impact}, it facilitates forecasting
+demand rates at state, regional, and department levels, enabling
+strategic pricing, capacity planning, and revenue management for hotels
+and other travel-related businesses. Additionally, in call centers,
+accurate call volume forecasting at different levels of the call center
+hierarchy or grouped structure is crucial for workforce management and
+resource allocation. Implementing our framework, call centers can
+generate accurate forecasts for different skill groups, shifts, and
+locations, ensuring efficient staffing and optimal service levels to
+meet customer demands.
 
@@
 
-R code to produce all results in this paper is available at
+To enhance transparency and reproducibility, we not only provide data
+and the code, but also the entire paper that is written in R using
+\href{https://quarto.org/}{Quarto}. All materials to reproduce this
+paper is available at
 \href{https://github.com/bahmanrostamitabar/forecasting-emergency-medicine}{github.com/bahmanrostamitabar/forecasting-emergency-medicine}.
-The repository contains the raw data, all R scripts used in experiments
-as well as the results used in this paper, along with the raw files for
-producing this paper. Full instructions are provided in the repository.
+The repository contains the raw data, all R scripts used in experiments,
+the results used in the paper, as well as the quarto files for producing
+this paper. Full instructions are provided in the repository.
 

### 2023-12-28 journal_correspondence/executive_summary/executive_summary.qmd
add summary executive and powerpoint slide

---
title: "Executive summary"
format: docx
---

Accurate forecasts of ambulance demand are critical inputs when planning and deploying staff and fleet. Such forecasts are required at the national, regional, and sub-regional levels and must take into account the nature of incidents as well as their priorities. Different teams within the organization frequently produce these forecasts independently. As a result, forecasts at various levels may be inconsistent, resulting in conflicting decisions and a lack of coherent coordination in the service. To address this issue, we use forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all available data at all levels of disaggregation.  Implementing and sustaining improvements in EMS necessitate alignments and coordination among various stakeholders, without which teams operate in isolation, leading to conflicts, duplication of work, rework, or work that runs counter to the overall goal of improving service quality. A hierarchical forecasting framework can be used to improve coordination between teams across care services at the national, sub-national, regional, and local levels. Therefore, because hierarchical forecasting is based on coherent forecasts, plans at any level can be aligned. Hierarchical forecasting approaches not only produce consistent forecasts, but they are also more accurate than independent (base) forecasts. As a result, the proposed framework in this study can assist EMS managers and professionals in not only improving coordination across the organization but also improving forecast accuracy across different levels by leveraging data available throughout the hierarchy.

### 2023-12-28 manuscript/main.qmd
add summary executive and powerpoint slide

@@
 title: "Hierarchical Time Series Forecasting in Emergency Medical Services"
-# author:
-#   - name: Bahman Rostami-Tabar
-#     email: rostami-tabarb@cardiff.ac.uk
-#     affiliations:
-#         - id: carbs
-#           name: Cardiff University
-#           department: Cardiff Business School
-#           country: United Kingdom
-#           postal-code: CF10 3EU
-#     attributes:
-#         corresponding: true
-#   - name: Rob J. Hyndman
-#     email: Rob.Hyndman@monash.edu
-#     affiliations:
-#         - id: monash
-#           name: Monash University
-#           department: Department of Econometrics and Business Statistics
-#           country: Australia
+author:
+  - name: Bahman Rostami-Tabar
+    email: rostami-tabarb@cardiff.ac.uk
+    affiliations:
+        - id: carbs
+          name: Cardiff University
+          department: Cardiff Business School
+          country: United Kingdom
+          postal-code: CF10 3EU
+    attributes:
+        corresponding: true
+  - name: Rob J. Hyndman
+    email: Rob.Hyndman@monash.edu
+    affiliations:
+        - id: monash
+          name: Monash University
+          department: Department of Econometrics and Business Statistics
+          country: Australia
 #           postal-code: VIC 3800

### 2023-12-28 manuscript/main.tex
add summary executive and powerpoint slide

@@
   pdftitle={Hierarchical Time Series Forecasting in Emergency Medical Services},
+  pdfauthor={Bahman Rostami-Tabar; Rob J. Hyndman},
   pdfkeywords={healthcare, emergency services, forecast
@@
 Services}
+\author[1]{Bahman Rostami-Tabar%
+\corref{cor1}%
+}
+ \ead{rostami-tabarb@cardiff.ac.uk} 
+\author[2]{Rob J. Hyndman%
+%
+}
+ \ead{Rob.Hyndman@monash.edu}
+
+\affiliation[1]{organization={Cardiff University, Cardiff Business
+School},country={United Kingdom},countrysep={,},postcode={CF10
+3EU},postcodesep={}}
+\affiliation[2]{organization={Monash University, Department of
+Econometrics and Business
+Statistics},country={Australia},countrysep={,},postcodesep={}}
 
@@
 \end{frontmatter}
-    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, interior hidden, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi
+    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, enhanced, sharp corners, frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 

### 2023-12-28 journal_correspondence/title_page.qmd
response letter - reposne to AE and reviewer 1
update title page

--- 2022-11-22 manuscript/title_page.qmd
@@
   Accurate forecasts of ambulance demand are crucial inputs when planning and deploying staff and fleet.
-Such demand forecasts are required at national, regional, and sub-regional levels, and must take account of the nature of incidents and their priorities.
+Such demand forecasts are required at national, regional, and sub-regional levels and must take account of the nature of incidents and their priorities.
 These forecasts are often generated independently by different teams within the organization.

### 2023-12-29 journal_correspondence/executive_summary/executive_summary.qmd
Updated exec summary

@@
 To address this issue, we use forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent and use all available data at all levels of disaggregation.
+
 Implementing and sustaining improvements in EMS necessitate alignments and coordination among various stakeholders, without which teams operate in isolation, leading to conflicts, duplication of work, rework, or work that runs counter to the overall goal of improving service quality.
 A hierarchical forecasting framework can be used to improve coordination between teams across care services at the national, sub-national, regional, and local levels.
-Therefore, because hierarchical forecasting is based on coherent forecasts, plans at any level can be aligned.
-Hierarchical forecasting approaches not only produce consistent forecasts, but they are also more accurate than independent (base) forecasts.
-As a result, the proposed framework in this study can assist EMS managers and professionals in not only improving coordination across the organization but also improving forecast accuracy across different levels by leveraging data available throughout the hierarchy.
+Because hierarchical forecasting is based on coherent forecasts, plans at any level can be aligned.
+
+Hierarchical forecasting approaches not only produce consistent forecasts, but they are also more accurate than independent (base) forecasts because they leverage data available throughout the hierarchy.
+As a result, the proposed framework in this study can assist EMS managers and professionals in improving forecast accuracy across different levels, as well as improving coordination across the organization.
+
+In summary, the use of forecast reconciliation methods and a hierarchical forecasting framework is crucial for improving coordination and forecast accuracy in emergency medical service, supporting more effective planning and deployment of ambulance staff and resources.
